### 2023-10-03 paper/paper.Rmd
upload rp training script

---
title: |
  Automated reading of residual plot with computer vision models
type: ARTICLE TEMPLATE
author:
  - name: Weihao Li
    affil: a
    email: weihao.li@monash.edu
affiliation:
  - num: a
    address: |
      Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
# bibliography: ref.bib
abstract: |
  TBD.
keywords: |
  TBD
header-includes: |
  \usepackage{lscape}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
output: rticles::tf_article
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  echo = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = "100%",
  fig.align = "center")
```

```{r}
# OOP supports needed by `visage`
# remotes::install_github("TengMCing/bandicoot")
# 
# Visual inference models and p-value calculation
# remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.0.tar.gz")

library(tidyverse)
library(visage)

# To control the simulation in this file
set.seed(10086)
```

# Introduction

# Methods for training computer vision models to analyze residual plots

In the field of statistical modeling and data analysis, interpreting residual plots plays a pivotal role in assessing the validity of our models. These plots can reveal deviations from modeling assumptions, including non-linearity, heteroskedasticity, and non-normality. The automation of residual plot analysis through computer vision provides a valuable tool for model validation and informed decision-making.

This section discusses various methodologies for training computer vision models to effectively read and interpret residual plots. We explore three distinct approaches, each with its unique characteristics and strengths.

## Method 1: Single Plot Model

The first method involves training a computer vision model to analyze a single residual plot and determine the probability that it was generated from a model without any violations. This method aligns with the principles of traditional deep learning classification problem, making it an intuitive and straightforward approach.

To implement the Single Plot Model effectively, we utilize established neural network architectures like VGG16. The model is trained on a diverse dataset containing both "not null" plots (generated from model with violations) and "null" plots (generated from model without violations).

This approach provides the model with certain capabilities:

1. **Distinguishing null and not null plots**: Through training, the model is expected to be able to distinguish between null plots and not null plots. This distinction forms the foundation for identifying model violations in new residual plots.

2. **Capturing variance in null plots**: Null plots often exhibit varying shapes and outliers due to the randomness of the error term and the distribution of fitted values. The model learns to recognize and characterize this variance, aiding in the detection of violations.

3. **Recognizing patterns in not null plots**: The model would develop the ability to identify common patterns in not null plots generated through simulations.

4. **Analyzing Aesthetic Elements**: Beyond content, the model takes into account aesthetic elements within the residual plot, including point size, auxiliary lines, layers, background colors, and other visual attributes. This is an undesired capability but it is unavoidable if only a single design of plot is used as inputs.

training sample:

improvements:
1. enrich the factors in simulation
2. sigma same, shape same, same fitted value distribution, stack residual plots  
3. test other residual plots generated from factors not seen by the model
4. replace the output with the effect size

what the function tries to approximate?

## Lineup model

The second method involves training a computer vision model to analyze a lineup of residual plots, typically consisting of around 20 plots. It then determines whether the data plot within the lineup is a null plot or a not null plot. This method closely resembles a visual test, where the input is a lineup, and the output is a decision indicating whether the data plot exhibits visual features corresponding to model violations. This method still adheres to the standard classification scheme, allowing for the utilization of the same model architecture.

This approach enable the model to learn all four points detailed in subsection 1 and three additional key points:

1. **Locating the data plot**: It learns to identify and locate the data plot within a lineup of residual plots.

2. **Comparing data plot and null plots**: The model may potentially compare the data plot with null plots within the lineup. This provide additional information for decision making.

3. **Understanding variance within a lineup**: It may also understand the variance among null plots within a lineup.

training sample:

improvement:
1. similarly, other lineups generated from factors not seen by the model
2. multiple lineups from each data plot, different positions

## Position model

The third method shares similarities to the second method, but with a different output objective: determining the position of the data plot within the lineup. Essentially, this approach asks the model to locate the data plot within a set of residual plots. It resembles a visual testing procedure where an individual is asked to identify the most distinct plot from a lineup. Even though the data plot may not always be the most distinct, training the model exclusively on lineups with substantial effect sizes can yield effective results.

The Position Model can potentially learn all the aspects mentioned in method 2 while providing extra predictive information: the probability that the data plot resides in a position different from the true one. This information is valuable as it can be utilized to calculate the p-value of a visual test.

training sample:

## Effect size model

##

### 2023-10-03 paper/paper.tex
upload rp training script

--- 2023-10-03 paper/paper.Rmd
@@
----
-title: |
-  Automated reading of residual plot with computer vision models
-type: ARTICLE TEMPLATE
-author:
-  - name: Weihao Li
-    affil: a
-    email: weihao.li@monash.edu
-affiliation:
-  - num: a
-    address: |
-      Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
-# bibliography: ref.bib
-abstract: |
-  TBD.
-keywords: |
-  TBD
-header-includes: |
+% interactcadsample.tex
+% v1.03 - April 2017
+
+\documentclass[]{interact}
+
+\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
+\usepackage{subfigure}% Support for small, `sub' figures and tables
+%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
+
+\usepackage{natbib}% Citation support using natbib.sty
+\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
+\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty
+
+\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{proposition}[theorem]{Proposition}
+
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
+
+\theoremstyle{remark}
+\newtheorem{remark}{Remark}
+\newtheorem{notation}{Notation}
+
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
 \usepackage{lscape}
@@
 \usepackage{setspace}
-output: rticles::tf_article
-editor_options: 
-  chunk_output_type: console
----
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-  message = FALSE, 
-  warning = FALSE, 
-  echo = FALSE,
-  fig.width = 8,
-  fig.height = 6,
-  out.width = "100%",
-  fig.align = "center")
-```
+\begin{document}
 
-```{r}
-# OOP supports needed by `visage`
-# remotes::install_github("TengMCing/bandicoot")
-# 
-# Visual inference models and p-value calculation
-# remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.0.tar.gz")
+\articletype{ARTICLE TEMPLATE}
 
-library(tidyverse)
-library(visage)
+\title{Automated reading of residual plot with computer vision models}
 
-# To control the simulation in this file
-set.seed(10086)
-```
+\author{\name{Weihao Li$^{a}$}
+\affil{$^{a}$Department of Econometrics and Business Statistics, Monash
+University, Clayton, VIC, Australia}
+}
 
-# Introduction
+\thanks{CONTACT Weihao
+Li. Email: \href{mailto:weihao.li@monash.edu}{\nolinkurl{weihao.li@monash.edu}}}
 
-# Methods for training computer vision models to analyze residual plots
+\maketitle
 
-In the field of statistical modeling and data analysis, interpreting residual plots plays a pivotal role in assessing the validity of our models.
-These plots can reveal deviations from modeling assumptions, including non-linearity, heteroskedasticity, and non-normality.
-The automation of residual plot analysis through computer vision provides a valuable tool for model validation and informed decision-making.
+\begin{abstract}
+TBD.
+\end{abstract}
 
-This section discusses various methodologies for training computer vision models to effectively read and interpret residual plots.
-We explore three distinct approaches, each with its unique characteristics and strengths.
+\begin{keywords}
+TBD
+\end{keywords}
 
-## Method 1: Single Plot Model
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-The first method involves training a computer vision model to analyze a single residual plot and determine the probability that it was generated from a model without any violations.
-This method aligns with the principles of traditional deep learning classification problem, making it an intuitive and straightforward approach.
+\hypertarget{methods-for-training-computer-vision-models-to-analyze-residual-plots}{%
+\section{Methods for training computer vision models to analyze residual
+plots}\label{methods-for-training-computer-vision-models-to-analyze-residual-plots}}
 
-To implement the Single Plot Model effectively, we utilize established neural network architectures like VGG16.
-The model is trained on a diverse dataset containing both "not null" plots (generated from model with violations) and "null" plots (generated from model without violations).
+In the field of statistical modeling and data analysis, interpreting
+residual plots plays a pivotal role in assessing the validity of our
+models. These plots can reveal deviations from modeling assumptions,
+including non-linearity, heteroskedasticity, and non-normality. The
+automation of residual plot analysis through computer vision provides a
+valuable tool for model validation and informed decision-making.
+
+This section discusses various methodologies for training computer
+vision models to effectively read and interpret residual plots. We
+explore three distinct approaches, each with its unique characteristics
+and strengths.
+
+\hypertarget{method-1-single-plot-model}{%
+\subsection{Method 1: Single Plot
+Model}\label{method-1-single-plot-model}}
+
+The first method involves training a computer vision model to analyze a
+single residual plot and determine the probability that it was generated
+from a model without any violations. This method aligns with the
+principles of traditional deep learning classification problem, making
+it an intuitive and straightforward approach.
+
+To implement the Single Plot Model effectively, we utilize established
+neural network architectures like VGG16. The model is trained on a
+diverse dataset containing both ``not null'' plots (generated from model
+with violations) and ``null'' plots (generated from model without
+violations).
 
@@
 
-1.
-**Distinguishing null and not null plots**: Through training, the model is expected to be able to distinguish between null plots and not null plots.
-This distinction forms the foundation for identifying model violations in new residual plots.
-
-2.
-**Capturing variance in null plots**: Null plots often exhibit varying shapes and outliers due to the randomness of the error term and the distribution of fitted values.
-The model learns to recognize and characterize this variance, aiding in the detection of violations.
-
-3. **Recognizing patterns in not null plots**: The model would develop the ability to identify common patterns in not null plots generated through simulations.
-
-4.
-**Analyzing Aesthetic Elements**: Beyond content, the model takes into account aesthetic elements within the residual plot, including point size, auxiliary lines, layers, background colors, and other visual attributes.
-This is an undesired capability but it is unavoidable if only a single design of plot is used as inputs.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  \textbf{Distinguishing null and not null plots}: Through training, the
+  model is expected to be able to distinguish between null plots and not
+  null plots. This distinction forms the foundation for identifying
+  model violations in new residual plots.
+\item
+  \textbf{Capturing variance in null plots}: Null plots often exhibit
+  varying shapes and outliers due to the randomness of the error term
+  and the distribution of fitted values. The model learns to recognize
+  and characterize this variance, aiding in the detection of violations.
+\item
+  \textbf{Recognizing patterns in not null plots}: The model would
+  develop the ability to identify common patterns in not null plots
+  generated through simulations.
+\item
+  \textbf{Analyzing Aesthetic Elements}: Beyond content, the model takes
+  into account aesthetic elements within the residual plot, including
+  point size, auxiliary lines, layers, background colors, and other
+  visual attributes. This is an undesired capability but it is
+  unavoidable if only a single design of plot is used as inputs.
+\end{enumerate}
 
@@
 
-improvements:
-1. enrich the factors in simulation
-2. sigma same, shape same, same fitted value distribution, stack residual plots  
-3. test other residual plots generated from factors not seen by the model
-4. replace the output with the effect size
+improvements: 1. enrich the factors in simulation 2. sigma same, shape
+same, same fitted value distribution, stack residual plots\\
+3. test other residual plots generated from factors not seen by the
+model 4. replace the output with the effect size
 
@@
 
-## Lineup model
+\hypertarget{lineup-model}{%
+\subsection{Lineup model}\label{lineup-model}}
 
-The second method involves training a computer vision model to analyze a lineup of residual plots, typically consisting of around 20 plots.
-It then determines whether the data plot within the lineup is a null plot or a not null plot.
-This method closely resembles a visual test, where the input is a lineup, and the output is a decision indicating whether the data plot exhibits visual features corresponding to model violations.
-This method still adheres to the standard classification scheme, allowing for the utilization of the same model architecture.
+The second method involves training a computer vision model to analyze a
+lineup of residual plots, typically consisting of around 20 plots. It
+then determines whether the data plot within the lineup is a null plot
+or a not null plot. This method closely resembles a visual test, where
+the input is a lineup, and the output is a decision indicating whether
+the data plot exhibits visual features corresponding to model
+violations. This method still adheres to the standard classification
+scheme, allowing for the utilization of the same model architecture.
 
-This approach enable the model to learn all four points detailed in subsection 1 and three additional key points:
+This approach enable the model to learn all four points detailed in
+subsection 1 and three additional key points:
 
-1. **Locating the data plot**: It learns to identify and locate the data plot within a lineup of residual plots.
-
-2. **Comparing data plot and null plots**: The model may potentially compare the data plot with null plots within the lineup. This provide additional information for decision making.
-
-3. **Understanding variance within a lineup**: It may also understand the variance among null plots within a lineup.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  \textbf{Locating the data plot}: It learns to identify and locate the
+  data plot within a lineup of residual plots.
+\item
+  \textbf{Comparing data plot and null plots}: The model may potentially
+  compare the data plot with null plots within the lineup. This provide
+  additional information for decision making.
+\item
+  \textbf{Understanding variance within a lineup}: It may also
+  understand the variance among null plots within a lineup.
+\end{enumerate}
 
@@
 
-improvement:
-1. similarly, other lineups generated from factors not seen by the model
-2. multiple lineups from each data plot, different positions
+improvement: 1. similarly, other lineups generated from factors not seen
+by the model 2. multiple lineups from each data plot, different
+positions
 
-## Position model
+\hypertarget{position-model}{%
+\subsection{Position model}\label{position-model}}
 
-The third method shares similarities to the second method, but with a different output objective: determining the position of the data plot within the lineup.
-Essentially, this approach asks the model to locate the data plot within a set of residual plots.
-It resembles a visual testing procedure where an individual is asked to identify the most distinct plot from a lineup.
-Even though the data plot may not always be the most distinct, training the model exclusively on lineups with substantial effect sizes can yield effective results.
+The third method shares similarities to the second method, but with a
+different output objective: determining the position of the data plot
+within the lineup. Essentially, this approach asks the model to locate
+the data plot within a set of residual plots. It resembles a visual
+testing procedure where an individual is asked to identify the most
+distinct plot from a lineup. Even though the data plot may not always be
+the most distinct, training the model exclusively on lineups with
+substantial effect sizes can yield effective results.
 
-The Position Model can potentially learn all the aspects mentioned in method 2 while providing extra predictive information: the probability that the data plot resides in a position different from the true one.
-This information is valuable as it can be utilized to calculate the p-value of a visual test.
+The Position Model can potentially learn all the aspects mentioned in
+method 2 while providing extra predictive information: the probability
+that the data plot resides in a position different from the true one.
+This information is valuable as it can be utilized to calculate the
+p-value of a visual test.
 
@@
 
-## Effect size model
+\hypertarget{effect-size-model}{%
+\subsection{Effect size model}\label{effect-size-model}}
 
-##
+\end{document}

### 2023-10-03 scripts/evaluation/compare_d_c.R

@@ -1,0 +1,227 @@
+library(tensorflow)
+library(keras)
+library(tidyverse)
+library(yardstick)
+library(visage)
+
+mod <- load_model_tf("temp_model/mixed_224")
+d_test_set <- flow_images_from_directory("data/experiment_factor/residual_plots/native/mixed/test",
+                                         target_size = c(224L, 224L), 
+                                         shuffle = FALSE)
+c_test_set <- flow_images_from_directory("data/phn_extended/residual_plots/native/mixed/test",
+                                         target_size = c(224L, 224L), 
+                                         shuffle = FALSE)
+d_pred <- mod$predict(d_test_set)
+c_pred <- mod$predict(c_test_set)
+
+d_meta <- readRDS("data/experiment_factor/residual_plots/meta.rds")
+c_meta <- readRDS("data/phn_extended/residual_plots/meta.rds")
+
+d_pred <- d_pred %>%
+  as.data.frame() %>%
+  mutate(truth = d_test_set$labels) %>%
+  mutate(filename = d_test_set$filenames) %>%
+  mutate(plot_uid = as.integer(gsub(".*/(.*).png", "\\1", filename))) %>%
+  left_join(d_meta)
+
+c_pred <- c_pred %>%
+  as.data.frame() %>%
+  mutate(truth = c_test_set$labels) %>%
+  mutate(filename = c_test_set$filenames) %>%
+  mutate(plot_uid = as.integer(gsub(".*/(.*).png", "\\1", filename))) %>%
+  left_join(c_meta)
+
+simple_es <- function(this_mod, n) {
+  map_dbl(1:10, ~this_mod$gen(n) %>% this_mod$sample_effect_size()) %>%
+    mean()
+}
+
+# Ensure the support of the predictor is [-1, 1]
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+get_x_var <- function(dist_name) {
+  
+  # Define the x variable
+  rand_uniform_x <- rand_uniform(-1, 1)
+  rand_normal_raw_x <- rand_normal(sigma = 0.3)
+  rand_normal_x <- closed_form(~stand_dist(rand_normal_raw_x))
+  rand_lognormal_raw_x <- rand_lognormal(sigma = 0.6)
+  rand_lognormal_x <- closed_form(~stand_dist(rand_lognormal_raw_x/3 - 1))
[…]

### 2023-10-03 scripts/evaluation/experiment_factor_vector_32.R

@@ -1,0 +1,227 @@
+library(tensorflow)
+library(keras)
+library(tidyverse)
+library(yardstick)
+library(visage)
+
+mod <- load_model_tf("keras_tuner/best_models/experiment_factor/vector_32")
+test_set <- read_csv("data/experiment_factor/vector/32/test.csv")
+
+test_set$pred <- mod$predict(test_set[,1:32]) %>% as.numeric()
+test_set$pred_label <- ifelse(test_set$pred > 0.5, "not_null", "null")
+
+conf_mat(test_set %>%
+           mutate(pred_label = factor(pred_label),
+                  response = factor(response)), 
+         estimate = pred_label, 
+         truth = response)
+
+bal_accuracy(test_set %>%
+               mutate(pred_label = factor(pred_label),
+                      response = factor(response)), 
+             estimate = pred_label, 
+             truth = response)
+
+vi_lineup <- get_vi_lineup()
+
+std_var <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+residual_to_vector <- function(resids, fitted, target_len = TARGET_LEN) {
+  ordered_resids <- resids[order(fitted)]
+  ordered_fitted <- sort(fitted)
+  
+  fitted_gap <- (max(ordered_fitted) - min(ordered_fitted))/target_len
+  
+  fitted_position <- floor((ordered_fitted - min(ordered_fitted))/fitted_gap) + 1L
+  fitted_position <- ifelse(fitted_position > target_len, target_len, fitted_position)
+  
+  map_dbl(1:target_len, function(i) {
+    mean(ordered_resids[i == fitted_position])
+  }) %>%
+    {ifelse(is.nan(.), 0, .)} %>%
+    std_var()
+}
+
+lineup_dat <- map(vi_lineup, function(lineup) {
+  this_dat <- lineup$data %>%
+    filter(null == FALSE)
+  residual_to_vector(this_dat$.resid, this_dat$.fitted, 32)
+}) %>%
+  reduce(rbind) %>%
+  `rownames<-`(NULL) %>%
+  as.data.frame() %>%
+  mutate(unique_lineup_id = names(vi_lineup))
+
+lineup_dat$pred <- mod$predict(lineup_dat[,1:32]) %>% as.numeric()
+lineup_dat$pred_label <- ifelse(lineup_dat$pred > 0.5, "not_null", "null")
+
+lineup_dat <- lineup_dat %>%
+  left_join(vi_survey %>%
+              group_by(unique_lineup_id) %>%
[…]

### 2023-10-03 scripts/evaluation/experiment_factor_vector_64.R

@@ -1,0 +1,232 @@
+library(tensorflow)
+library(keras)
+library(tidyverse)
+library(yardstick)
+library(visage)
+
+mod <- load_model_tf("keras_tuner/best_models/experiment_factor/vector/64")
+test_set <- read_csv("data/experiment_factor/vector/64/test.csv")
+
+test_set$pred <- mod$predict(test_set[,1:64]) %>% as.numeric()
+test_set$pred_label <- ifelse(test_set$pred > 0.5, "not_null", "null")
+
+conf_mat(test_set %>%
+           mutate(pred_label = factor(pred_label),
+                  response = factor(response)), 
+         estimate = pred_label, 
+         truth = response)
+
+bal_accuracy(test_set %>%
+               mutate(pred_label = factor(pred_label),
+                      response = factor(response)), 
+             estimate = pred_label, 
+             truth = response)
+
+vi_lineup <- get_vi_lineup()
+
+std_var <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+residual_to_vector <- function(resids, fitted, target_len = TARGET_LEN) {
+  ordered_resids <- resids[order(fitted)]
+  ordered_fitted <- sort(fitted)
+  
+  fitted_gap <- (max(ordered_fitted) - min(ordered_fitted))/target_len
+  
+  fitted_position <- floor((ordered_fitted - min(ordered_fitted))/fitted_gap) + 1L
+  fitted_position <- ifelse(fitted_position > target_len, target_len, fitted_position)
+  
+  map_dbl(1:target_len, function(i) {
+    mean(ordered_resids[i == fitted_position])
+  }) %>%
+    {ifelse(is.nan(.), 0, .)} %>%
+    std_var()
+}
+
+lineup_dat <- map(vi_lineup, function(lineup) {
+  this_dat <- lineup$data %>%
+    filter(null == FALSE)
+  residual_to_vector(this_dat$.resid, this_dat$.fitted, 64)
+}) %>%
+  reduce(rbind) %>%
+  `rownames<-`(NULL) %>%
+  as.data.frame() %>%
+  mutate(unique_lineup_id = names(vi_lineup))
+
+lineup_dat$pred <- mod$predict(lineup_dat[,1:64]) %>% as.numeric()
+lineup_dat$pred_label <- ifelse(lineup_dat$pred > 0.5, "not_null", "null")
+
+lineup_dat <- lineup_dat %>%
+  left_join(vi_survey %>%
+              group_by(unique_lineup_id) %>%
[…]

### 2023-10-03 scripts/evaluation/stack_residual_plots.R

@@ -1,0 +1,87 @@
+library(visage)
+
+
+# get_x_var ---------------------------------------------------------------
+
+# Ensure the support of the predictor is [-1, 1]
+stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1
+
+get_x_var <- function(dist_name) {
+  
+  # Define the x variable
+  rand_uniform_x <- rand_uniform(-1, 1)
+  rand_normal_raw_x <- rand_normal(sigma = 0.3)
+  rand_normal_x <- closed_form(~stand_dist(rand_normal_raw_x))
+  rand_lognormal_raw_x <- rand_lognormal(sigma = 0.6)
+  rand_lognormal_x <- closed_form(~stand_dist(rand_lognormal_raw_x/3 - 1))
+  rand_discrete_x <- rand_uniform_d(-1, 1, k = 5, even = TRUE)
+  
+  switch(as.character(dist_name),
+         uniform = rand_uniform_x,
+         normal = rand_normal_x,
+         lognormal = rand_lognormal_x,
+         even_discrete = rand_discrete_x)
+} 
+
+mod_ <- poly_model(shape = 1, x = get_x_var("uniform"), sigma = 0.5)
+map(1:500, ~mod_$gen(300, test = FALSE) %>% 
+      mutate(.resid = stand_dist(.resid), 
+             .fitted = stand_dist(.fitted)) %>%
+      select(.resid, .fitted)) %>%
+  reduce(bind_rows) %>%
+  ggplot() +
+  geom_point(aes(.fitted, .resid), alpha = 0.1)
+
+mod_ <- poly_model(shape = 1, x = get_x_var("uniform"), sigma = 1)
+map(1:500, ~mod_$gen(300, test = FALSE) %>% 
+      mutate(.resid = stand_dist(.resid), 
+             .fitted = stand_dist(.fitted)) %>%
+      select(.resid, .fitted)) %>%
+  reduce(bind_rows) %>%
+  ggplot() +
+  geom_point(aes(.fitted, .resid), alpha = 0.1)
+
+mod_ <- poly_model(shape = 1, x = get_x_var("uniform"), sigma = 2)
+map(1:500, ~mod_$gen(300, test = FALSE) %>% 
+      mutate(.resid = stand_dist(.resid), 
+             .fitted = stand_dist(.fitted)) %>%
[…]

### 2023-10-03 scripts/training/experiment_factor_residual_plots_32.py

@@ -1,0 +1,206 @@
+import sys
+import subprocess
+import os
+import tensorflow as tf
+from tensorflow import keras
+import keras_tuner
+
+# Global setttings
+DATA_CLASS = 'experiment_factor'
+INPUT_TYPE = 'residual_plots'
+RES = 32
+COLOR_MODE = 'rgb'
+CHANNELS = 3
+CLASS_NAMES = ('not_null', "null")
+BATCH_SIZE = 32
+
+project_dir = subprocess.run(['Rscript', '-e', 'cat(here::here())'],
+                             check=True,
+                             capture_output=True,
+                             text=True).stdout
+
+train_dir = os.path.join(project_dir, 
+                         'data', 
+                         f'{DATA_CLASS}',
+                         f'{INPUT_TYPE}',
+                         f'{RES}',
+                         'train')
+                         
+                         
+data_gen = keras.preprocessing.image.ImageDataGenerator(validation_split=0.2)
+train_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="training")
+val_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="validation")
+                             
+def build_model(hp):
+    
+    # Preprocess the input image
+    model_input = keras.layers.Input(shape=(RES, RES, CHANNELS))
+    processed_input = keras.applications.vgg16.preprocess_input(model_input)
+    grey_scale_input = keras.layers.Lambda(
+      lambda image: tf.image.rgb_to_grayscale(image), name='grey_scale'
+    )(processed_input)
+    
[…]

### 2023-10-03 scripts/training/experiment_factor_residual_plots_32.py
correct the training script

@@ -186,11 +186,8 @@ callbacks.append(keras.callbacks.ReduceLROnPlateau(
                        
 tuner.search_space_summary()
 
-tuner.search(x=train_x, 
-             y=train_y,
+tuner.search(x=train_dat, 
              epochs=100, 
-             validation_split=0.2,
-             shuffle=False,
              callbacks=callbacks)
 
 tuner.results_summary()

### 2023-10-03 scripts/training/experiment_factor_residual_plots_32.py
update train script

@@ -24,6 +24,7 @@ train_dir = os.path.join(project_dir,
                          f'{DATA_CLASS}',
                          f'{INPUT_TYPE}',
                          f'{RES}',
+                         'mixed',
                          'train')
                          
                          

### 2023-10-03 scripts/training/experiment_factor_residual_plots_32.py
update

@@ -187,7 +187,8 @@ callbacks.append(keras.callbacks.ReduceLROnPlateau(
                        
 tuner.search_space_summary()
 
-tuner.search(x=train_dat, 
+tuner.search(x=train_dat,
+             validation_data=val_dat,
              epochs=100, 
              callbacks=callbacks)
 

### 2023-10-03 scripts/training/experiment_factor_residual_plots_32.py
update

@@ -151,7 +151,7 @@ def build_model(hp):
   
 
 tuner = keras_tuner.BayesianOptimization(hypermodel=build_model,
-                                         objective='val_binary_accuracy',
+                                         objective='val_categorical_accuracy',
                                          max_trials=30,
                                          executions_per_trial=1,
                                          overwrite=False,

### 2023-10-04 scripts/training/experiment_factor_residual_plots_64.py
residual plots 64

@@ -1,0 +1,205 @@
+import sys
+import subprocess
+import os
+import tensorflow as tf
+from tensorflow import keras
+import keras_tuner
+
+# Global setttings
+DATA_CLASS = 'experiment_factor'
+INPUT_TYPE = 'residual_plots'
+RES = 64
+COLOR_MODE = 'rgb'
+CHANNELS = 3
+CLASS_NAMES = ('not_null', "null")
+BATCH_SIZE = 32
+
+project_dir = subprocess.run(['Rscript', '-e', 'cat(here::here())'],
+                             check=True,
+                             capture_output=True,
+                             text=True).stdout
+
+train_dir = os.path.join(project_dir, 
+                         'data', 
+                         f'{DATA_CLASS}',
+                         f'{INPUT_TYPE}',
+                         f'{RES}',
+                         'mixed',
+                         'train')
+                         
+                         
+data_gen = keras.preprocessing.image.ImageDataGenerator(validation_split=0.2)
+train_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="training")
+val_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="validation")
+                             
+def build_model(hp):
+    
+    # Preprocess the input image
+    model_input = keras.layers.Input(shape=(RES, RES, CHANNELS))
+    processed_input = keras.applications.vgg16.preprocess_input(model_input)
+    grey_scale_input = keras.layers.Lambda(
[…]

### 2023-10-04 scripts/training/experiment_factor_residual_plots_128.py
residual plots 128

@@ -1,0 +1,205 @@
+import sys
+import subprocess
+import os
+import tensorflow as tf
+from tensorflow import keras
+import keras_tuner
+
+# Global setttings
+DATA_CLASS = 'experiment_factor'
+INPUT_TYPE = 'residual_plots'
+RES = 128
+COLOR_MODE = 'rgb'
+CHANNELS = 3
+CLASS_NAMES = ('not_null', "null")
+BATCH_SIZE = 32
+
+project_dir = subprocess.run(['Rscript', '-e', 'cat(here::here())'],
+                             check=True,
+                             capture_output=True,
+                             text=True).stdout
+
+train_dir = os.path.join(project_dir, 
+                         'data', 
+                         f'{DATA_CLASS}',
+                         f'{INPUT_TYPE}',
+                         f'{RES}',
+                         'mixed',
+                         'train')
+                         
+                         
+data_gen = keras.preprocessing.image.ImageDataGenerator(validation_split=0.2)
+train_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="training")
+val_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="validation")
+                             
+def build_model(hp):
+    
+    # Preprocess the input image
+    model_input = keras.layers.Input(shape=(RES, RES, CHANNELS))
+    processed_input = keras.applications.vgg16.preprocess_input(model_input)
+    grey_scale_input = keras.layers.Lambda(
[…]

### 2023-10-04 scripts/training/experiment_factor_residual_plots_256.py
residual plots 256

@@ -1,0 +1,205 @@
+import sys
+import subprocess
+import os
+import tensorflow as tf
+from tensorflow import keras
+import keras_tuner
+
+# Global setttings
+DATA_CLASS = 'experiment_factor'
+INPUT_TYPE = 'residual_plots'
+RES = 256
+COLOR_MODE = 'rgb'
+CHANNELS = 3
+CLASS_NAMES = ('not_null', "null")
+BATCH_SIZE = 32
+
+project_dir = subprocess.run(['Rscript', '-e', 'cat(here::here())'],
+                             check=True,
+                             capture_output=True,
+                             text=True).stdout
+
+train_dir = os.path.join(project_dir, 
+                         'data', 
+                         f'{DATA_CLASS}',
+                         f'{INPUT_TYPE}',
+                         f'{RES}',
+                         'mixed',
+                         'train')
+                         
+                         
+data_gen = keras.preprocessing.image.ImageDataGenerator(validation_split=0.2)
+train_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="training")
+val_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="validation")
+                             
+def build_model(hp):
+    
+    # Preprocess the input image
+    model_input = keras.layers.Input(shape=(RES, RES, CHANNELS))
+    processed_input = keras.applications.vgg16.preprocess_input(model_input)
+    grey_scale_input = keras.layers.Lambda(
[…]

### 2023-10-04 scripts/training/experiment_factor_residual_plots_128.py
correct the saved model path

@@ -201,5 +201,5 @@ model_dir = os.path.join(project_dir,
                          "best_models",
                          f"{DATA_CLASS}",
                          f"{INPUT_TYPE}",
-                         f'{VECTOR_LEN}')
+                         f'{RES}')
 best_model.save(model_dir)

### 2023-10-05 scripts/evaluation/tensorboard.R
move slurm to new places

@@ -1,0 +1,3 @@
+library(tensorflow)
+tensorboard("keras_tuner/logs/residual_plots/32", "start")
+tensorboard("keras_tuner/logs/residual_plots/32", "stop")

### 2023-10-06 scripts/evaluation/tensorboard.R
add tensorboard

@@ -1,3 +1,3 @@
 library(tensorflow)
-tensorboard("keras_tuner/logs/residual_plots/32", "start")
-tensorboard("keras_tuner/logs/residual_plots/32", "stop")
+tensorboard("keras_tuner/logs/residual_plots/64", "start")
+tensorboard("keras_tuner/logs/residual_plots/64", "stop")

### 2023-10-07 scripts/evaluation/experiment_factor.R
update experiment factor eval script

@@ -1,0 +1,221 @@
+library(tensorflow)
+library(keras)
+library(tidyverse)
+library(yardstick)
+library(visage)
+
+mod_32 <- load_model_tf("keras_tuner/best_models/experiment_factor/residual_plots/32")
+mod_64 <- load_model_tf("keras_tuner/best_models/experiment_factor/residual_plots/64")
+
+test_32 <- flow_images_from_directory("data/experiment_factor/residual_plots/32/mixed/test",
+                                      target_size = c(32L, 32L),
+                                      shuffle = FALSE)
+test_64 <- flow_images_from_directory("data/experiment_factor/residual_plots/64/mixed/test",
+                                      target_size = c(64L, 64L),
+                                      shuffle = FALSE)
+
+
+# get predictions ---------------------------------------------------------
+
+pred_32 <- mod_32$predict(test_32) %>%
+  as.data.frame() %>%
+  mutate(truth = test_32$classes + 1) %>%
+  mutate(truth = names(test_32$class_indices)[truth]) %>%
+  mutate(filenames = test_32$filenames) %>%
+  mutate(plot_uid = gsub(".*/(.*).png", "\\1", filenames)) %>%
+  mutate(plot_uid = as.integer(plot_uid))
+
+pred_32 <- pred_32 %>%
+  mutate(pred = V1 > 0.5) %>%
+  mutate(pred = ifelse(pred, "not_null", "null")) %>%
+  mutate(pred = factor(pred)) %>%
+  mutate(truth = factor(truth))
+
+pred_32 %>% bal_accuracy(estimate = pred, truth = truth)
+
+pred_64 <- mod_64$predict(test_64) %>%
+  as.data.frame() %>%
+  mutate(truth = test_64$classes + 1) %>%
+  mutate(truth = names(test_64$class_indices)[truth]) %>%
+  mutate(filenames = test_64$filenames) %>%
+  mutate(plot_uid = gsub(".*/(.*).png", "\\1", filenames)) %>%
+  mutate(plot_uid = as.integer(plot_uid))
+
+pred_64 <- pred_64 %>%
+  mutate(pred = V1 > 0.5) %>%
+  mutate(pred = ifelse(pred, "not_null", "null")) %>%
+  mutate(pred = factor(pred)) %>%
+  mutate(truth = factor(truth))
+
[…]

### 2023-10-08 scripts/evaluation/experiment_factor.Rmd
update eval script

---
title: "Analysis of residual plot model"
output: html_document
date: "2023-10-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load_packages}
library(tensorflow)
library(keras)
library(tidyverse)
library(yardstick)
library(visage)
library(glue)
library(here)
```

```{r set_global_variables}
ALL_RES <- c(32L, 64L)
```

```{r load_model_and_test_set}
mod <- list()
test <- list()
for (res in ALL_RES) {
  mod[[res]] <- load_model_tf(here(glue("keras_tuner/best_models/experiment_factor/residual_plots/{res}")))
  test[[res]] <- flow_images_from_directory(here(glue("data/experiment_factor/residual_plots/{res}/mixed/test")),
                                            target_size = c(res, res),
                                            shuffle = FALSE)
}
```

```{r get_predictions}
pred <- list()
for (res in ALL_RES) {
  pred[[res]] <- mod[[res]]$predict(test[[res]]) %>%
    as.data.frame() %>%
    mutate(truth = test[[res]]$classes + 1) %>%
    mutate(truth = names(test[[res]]$class_indices)[truth]) %>%
    mutate(filenames = test[[res]]$filenames) %>%
    mutate(plot_uid = gsub(".*/(.*).png", "\\1", filenames)) %>%
    mutate(plot_uid = as.integer(plot_uid)) %>%
    mutate(pred = V1 > 0.5) %>%
    mutate(pred = ifelse(pred, "not_null", "null")) %>%
    mutate(pred = factor(pred)) %>%
    mutate(truth = factor(truth)) %>%
    mutate(res = res)
}
```

```{r get_meta}
meta <- readRDS(here("data/experiment_factor/residual_plots/meta.rds"))

effect_size_ref <- vi_survey %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  select(shape, e_sigma, a, b, n, x_dist, effect_size) %>%
  group_by(shape, e_sigma, a, b, n, x_dist) %>%
  summarise(effect_size = first(effect_size))
```

```{r merge_meta}
for (res in ALL_RES) {
  pred[[res]] <- pred[[res]] %>%
    left_join(meta) %>%
    left_join(effect_size_ref) %>%
    mutate(type = ifelse(!is.na(shape), "polynomial", "heteroskedasticity")) %>%
    mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete", x_dist))
}
```

# Basic summary

```{r basic_summary}

basic_summary <- function(dat, res = 32L) {
  dat %>%
  ggplot() +
  geom_bar(aes(pred, fill = type), position = "dodge") +
  geom_text(data = dat %>%
              group_by(truth, type, pred) %>%
              summarise(prop = n()/2000) %>%
              filter(type == "heteroskedasticity"),
  aes(pred, prop * 2000, label = format(prop, digits = 3)),
  nudge_x = -0.25, nudge_y = 30) +
  geom_text(data = dat %>%
              group_by(truth, type, pred) %>%
              summarise(prop = n()/2000) %>%
              filter(type == "polynomial"),
  aes(pred, prop * 2000, label = format(prop, digits = 3)),
  nudge_x = 0.25, nudge_y = 30) +
  xlab("Prediction") +
  theme_light() +
  facet_wrap(~truth, labeller = label_both) +
  ylab("Count") +
  scale_fill_brewer(palette = "Set2") +
  ggtitle("Summary of prediction", subtitle = glue("Image size {res} * {res}"))
}

for (res in ALL_RES) {
  print(basic_summary(pred[[res]], res))
}

```

```{r balance_accuracy}
pred %>%
  reduce(bind_rows) %>%
  group_by(res) %>%
  bal_accuracy(truth, pred) %>%
  kableExtra::kable(col.names = c("Input size", "Metric", "Estimator", "Estimate")) %>%
  kableExtra::kable_material()
```

### 2023-10-08 scripts/evaluation/experiment_factor.R

@@ -3,6 +3,7 @@ library(keras)
 library(tidyverse)
 library(yardstick)
 library(visage)
+library(glue)
 
 mod_32 <- load_model_tf("keras_tuner/best_models/experiment_factor/residual_plots/32")
 mod_64 <- load_model_tf("keras_tuner/best_models/experiment_factor/residual_plots/64")
@@ -83,30 +84,54 @@ bind_rows(mutate(pred_32, res = 32),
   autoplot()
 
 
-# check 32 ----------------------------------------------------------------
+
+# check model -------------------------------------------------------------
+
+check_model <- list()
 
 # Distribution of predicted probability on the test set
-pred_32 %>%
-  ggplot() +
-  geom_histogram(aes(V1)) +
-  ggtitle(quote("Distribution of " ~ hat(P)(y == not_null) ~ "on the test set"), 
-          subtitle = "Input size 32 * 32") +
-  xlab(quote(hat(P)(y == not_null))) +
-  facet_grid(type ~ truth, labeller = label_both) +
-  theme_light()
+check_model$p_hat$distribution <- function(dat, res = 32) {
+  dat %>%
+    ggplot() +
+    geom_histogram(aes(V1)) +
+    ggtitle(quote("Distribution of " ~ hat(P)(y == not_null) ~ "on the test set"), 
+            subtitle = glue("Input size {res} * {res}")) +
+    xlab(quote(hat(P)(y == not_null))) +
+    facet_grid(type ~ truth, labeller = label_both) +
+    theme_light()
+}
 
 # Boxplot of predicted probability conditional on factors
-# x_dist | null
-pred_32 %>%
-  filter(truth == "null") %>%
-  ggplot() +
-  geom_boxplot(aes(factor(x_dist), V1)) +
-  ylab(quote(hat(P)(y == not_null))) +
-  xlab(quote(distribution ~ "of fitted values")) +
-  scale_x_discrete(labels = c("discrete uniform", "lognormal", "normal", "uniform")) +
-  ggtitle(quote("Distribution of " ~ hat(P)(y == not_null ~"|"~ truth == null) ~ "on the test set"), 
-          subtitle = "Input size 32 * 32") +
-  theme_light()
+check_model$p_hat$null$x_dist$distribution <- function(dat, res = 32) {
+  dat %>%
+    filter(truth == "null") %>%
+    ggplot() +
+    geom_boxplot(aes(factor(x_dist), V1)) +
+    ylab(quote(hat(P)(y == not_null))) +
[…]

### 2023-10-09 scripts/evaluation/experiment_factor.Rmd
update training script

@@
 ```{r set_global_variables}
-ALL_RES <- c(32L, 64L)
+ALL_RES <- c(32L, 64L, 128L)
 ```
@@
 
-```{r basic_summary}
-
-basic_summary <- function(dat, res = 32L) {
-  dat %>%
-  ggplot() +
-  geom_bar(aes(pred, fill = type), position = "dodge") +
-  geom_text(data = dat %>%
-              group_by(truth, type, pred) %>%
+```{r}
+pred %>%
+  reduce(bind_rows) %>%
+  group_by(res, truth, type, pred) %>%
   summarise(prop = n()/2000)  %>%
-              filter(type == "heteroskedasticity"),
-  aes(pred, prop * 2000, label = format(prop, digits = 3)),
-  nudge_x = -0.25, nudge_y = 30) +
-  geom_text(data = dat %>%
-              group_by(truth, type, pred) %>%
-              summarise(prop = n()/2000) %>%
-              filter(type == "polynomial"),
-  aes(pred, prop * 2000, label = format(prop, digits = 3)),
-  nudge_x = 0.25, nudge_y = 30) +
-  xlab("Prediction") +
-  theme_light() +
-  facet_wrap(~truth, labeller = label_both) +
-  ylab("Count") +
-  scale_fill_brewer(palette = "Set2") +
-  ggtitle("Summary of prediction", subtitle = glue("Image size {res} * {res}"))
-}
-
-for (res in ALL_RES) {
-  print(basic_summary(pred[[res]], res))
-}
-
+  filter(truth == pred) %>%
+  select(-pred) %>%
+  pivot_wider(names_from = truth, values_from = prop) %>%
+  arrange(type, res) %>%
+  kableExtra::kable(col.names = c("Input size", "Violation", "Accuracy: not_null", "Accuracy: null")) %>%
+  kableExtra::kable_material()
 ```
@@
 ```
+
+# ROC
+
+```{r roc}
+pred %>%
+  reduce(bind_rows) %>%
+  group_by(res) %>%
+  roc_curve(truth, V1) %>%
+  autoplot() +
+  labs(col = "Input size") +
+  ggtitle("ROC")
+```
+
+# Distribution of predicted probability
+
+```{r p_hat_distribution}
+p_hat_distribution <- function(dat, res = 32) {
+  dat %>%
+    ggplot() +
+    geom_histogram(aes(V1)) +
+    ggtitle(quote("Distribution of " ~ hat(P)(y == not_null) ~ "on the test set"), 
+            subtitle = glue("Input size {res} * {res}")) +
+    xlab(quote(hat(P)(y == not_null))) +
+    facet_grid(type ~ truth, labeller = label_both) +
+    theme_light()
+}
+
+for (res in ALL_RES) {
+  p_hat_distribution(pred[[res]], res) %>%
+    print()
+}
+```
+
+# Distribution of predicted probability on null set
+
+```{r p_hat_null_shape}
+p_hat_null_shape <- function(dat, res = 32) {
+  dat %>%
+    filter(truth == "null") %>%
+    ggplot() +
+    geom_boxplot(aes(factor(x_dist), V1)) +
+    ylab(quote(hat(P)(y == not_null))) +
+    xlab("Distribution of fitted values") +
+    scale_x_discrete(labels = c("discrete uniform", "lognormal", "normal", "uniform")) +
+    ggtitle(quote("Distribution of " ~ hat(P)(y == not_null ~"|"~ truth == null) ~ "on the test set"), 
+            subtitle = glue("Input size {res} * {res}")) +
+    theme_light()
+}
+
+for (res in ALL_RES) {
+  p_hat_null_shape(pred[[res]], res) %>%
+    print()
+}
+```

### 2023-10-09 scripts/training/phn_extended_residual_plots_32.py

@@ -1,0 +1,205 @@
+import sys
+import subprocess
+import os
+import tensorflow as tf
+from tensorflow import keras
+import keras_tuner
+
+# Global setttings
+DATA_CLASS = 'phn_extended'
+INPUT_TYPE = 'residual_plots'
+RES = 32
+COLOR_MODE = 'rgb'
+CHANNELS = 3
+CLASS_NAMES = ('not_null', "null")
+BATCH_SIZE = 32
+
+project_dir = subprocess.run(['Rscript', '-e', 'cat(here::here())'],
+                             check=True,
+                             capture_output=True,
+                             text=True).stdout
+
+train_dir = os.path.join(project_dir, 
+                         'data', 
+                         f'{DATA_CLASS}',
+                         f'{INPUT_TYPE}',
+                         f'{RES}',
+                         'mixed',
+                         'train')
+                         
+                         
+data_gen = keras.preprocessing.image.ImageDataGenerator(validation_split=0.2)
+train_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="training")
+val_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="validation")
+                             
+def build_model(hp):
+    
+    # Preprocess the input image
+    model_input = keras.layers.Input(shape=(RES, RES, CHANNELS))
+    processed_input = keras.applications.vgg16.preprocess_input(model_input)
+    grey_scale_input = keras.layers.Lambda(
+      lambda image: tf.image.rgb_to_grayscale(image), name='grey_scale'
[…]

### 2023-10-09 scripts/training/phn_extended_residual_plots_64.py

@@ -1,0 +1,205 @@
+import sys
+import subprocess
+import os
+import tensorflow as tf
+from tensorflow import keras
+import keras_tuner
+
+# Global setttings
+DATA_CLASS = 'phn_extended'
+INPUT_TYPE = 'residual_plots'
+RES = 64
+COLOR_MODE = 'rgb'
+CHANNELS = 3
+CLASS_NAMES = ('not_null', "null")
+BATCH_SIZE = 32
+
+project_dir = subprocess.run(['Rscript', '-e', 'cat(here::here())'],
+                             check=True,
+                             capture_output=True,
+                             text=True).stdout
+
+train_dir = os.path.join(project_dir, 
+                         'data', 
+                         f'{DATA_CLASS}',
+                         f'{INPUT_TYPE}',
+                         f'{RES}',
+                         'mixed',
+                         'train')
+                         
+                         
+data_gen = keras.preprocessing.image.ImageDataGenerator(validation_split=0.2)
+train_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="training")
+val_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="validation")
+                             
+def build_model(hp):
+    
+    # Preprocess the input image
+    model_input = keras.layers.Input(shape=(RES, RES, CHANNELS))
+    processed_input = keras.applications.vgg16.preprocess_input(model_input)
+    grey_scale_input = keras.layers.Lambda(
+      lambda image: tf.image.rgb_to_grayscale(image), name='grey_scale'
[…]

### 2023-10-09 scripts/training/experiment_factor_residual_plots_128.py
mkdir -> makedirs

@@ -172,7 +172,7 @@ callbacks.append(keras.callbacks.EarlyStopping(
                  verbose=1))
                  
 if os.path.exists(os.path.dirname(log_dir)) is False:
-    os.mkdir(os.path.dirname(log_dir))
+    os.makedirs(os.path.dirname(log_dir))
 callbacks.append(keras.callbacks.TensorBoard(
                  log_dir=log_dir,
                  histogram_freq=0,

### 2023-10-09 scripts/training/experiment_factor_vector_32.py

@@ -144,7 +144,7 @@ callbacks.append(keras.callbacks.EarlyStopping(
                  verbose=1))
                  
 if os.path.exists(os.path.dirname(log_dir)) is False:
-    os.mkdir(os.path.dirname(log_dir))
+    os.makedirs(os.path.dirname(log_dir))
 callbacks.append(keras.callbacks.TensorBoard(
                  log_dir=log_dir,
                  histogram_freq=0,

### 2023-10-10 scripts/evaluation/experiment_factor.Rmd
update eval

--- 2023-10-08 scripts/evaluation/experiment_factor.Rmd
@@
 title: "Analysis of residual plot model"
-output: html_document
+output: 
+  html_document:
+    toc: true
+    theme: united
+    toc_float: true
 date: "2023-10-08"
@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+```
+
+```{css}
+body {
+  font-size: 18px;
+}
 ```
@@
                                             target_size = c(res, res),
-                                            shuffle = FALSE)
+                                            shuffle = FALSE,
+                                            batch_size = 1000L)
 }
@@
   group_by(shape, e_sigma, a, b, n, x_dist) %>%
-  summarise(effect_size = first(effect_size))
+  summarise(effect_size = first(effect_size)) %>%
+  ungroup()
 ```
@@
   pred[[res]] <- pred[[res]] %>%
-    left_join(meta) %>%
-    left_join(effect_size_ref) %>%
+    left_join(meta)
+
+  pred[[res]] <- bind_rows(pred[[res]] %>%
+     filter(!is.na(shape)) %>%
+     left_join(select(filter(effect_size_ref, !is.na(shape)), 
+                      shape, e_sigma, x_dist, n, effect_size)),
+     pred[[res]] %>%
+     filter(!is.na(a)) %>%
+     left_join(select(filter(effect_size_ref, !is.na(a)), 
+                      a, b, x_dist, n, effect_size))) %>%
     mutate(type = ifelse(!is.na(shape), "polynomial", "heteroskedasticity")) %>%
-    mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete", x_dist))
+    mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete", x_dist)) %>%
+    mutate(effect_size = ifelse(response == "null", 0, effect_size))
 }
@@
 
+# Introduction
+
+This report presents an analysis of computer vision models designed to take as input a residual plot image and produce a numeric value between 0 and 1, indicating whether the input residual plot represents a null plot or a non-null plot.
+
+Currently, we have trained three such models, each with a distinct input size: 32, 64, and 128.
+
+## Training data
+
+The training data is generated using the same discrete factors utilized in the human subject experiment.
+This dataset encompasses two specific violations: non-linearity and heteroskedasticity.
+For each violation, we sample 2000 sets of experimental factors, which are subsequently employed to construct linear models.
+For each set of experimental factors, we fit 11 null linear models and 11 non-null linear models.
+Out of these, 10 are used to generate residual plots for the training dataset, while 1 is used to generate residual plots for the test dataset.
+Consequently, we have a total of 80,000 training images (2 violations x 2,000 sets x 2 types x 10 models) and 8,000 test images (2 violations x 2,000 sets x 2 types x 1 model).
+
+## Model architecture
+
+For the model architecture, we have adopted the existing VGG16 framework provided by `keras`, and we've employed the `keras-tuner` package for hyperparameter tuning.
+The hyperparameters under consideration for tuning include:
+
+1. The number of convolutional blocks. (1 to 5)
+2. The base factor determining the number of convolutional filters. (4, 8, 16, 32, 64)
+3. The choice between max pooling and average pooling for processing the output of the final convolutional layers.
+4. The number of units in the single-layer classifier. (8, 16, 32, 64, 128, 256, 512, 1024)
+5. Values for L1 and L2 regularization for the classifier. ($1^{-6}$ to $1^{-1}$)
+6. The dropout rate applied to the classifier. (0.1 to 0.8)
+7. Additionally, we are also exploring the tuning of the learning rate during the training process. ($1^{-6}$ to $1^{-2}$)
+
+The input for our models consists of RGB images with three color channels.
+To prepare the data for processing, we add a `grey_scale` layer, which transforms the input image into a grayscale image.
+This transformation is achieved using the luma formula, adhering to the CCIR 601 standard.
+
+Subsequently, we incorporate multiple convolutional blocks into the model architecture.
+Increasing the number of blocks and filters within the convolutional layers leads to a higher number of tunable parameters in the model, allowing for more complex feature extraction.
+
+The output from the final convolutional block serves as the input to a 2D global pooling layer, which is responsible for reshaping the tensor and reducing its dimensionality.
+
+Additionally, we employ a batch normalization layer to adjust the output of the dense layer.
+This normalization process helps address issues such as feature shift and can contribute to improved model performance.
+
+## Model training
+
+We employ Bayesian optimization tuning from `keras-tuner` with 30 trials for each model to find the best-performing model.
+The best epoch from the best trial will be restored at the end of the tuning process.
+
+In the validation process, we allocate 20 percent of the training data, which amounts to $20\% \times 80,000 = 16,000$ images.
+
+Furthermore, we implement early stopping for each trial.
+If the validation loss fails to improve over a span of 10 epochs, the training process is halted.
+Additionally, we incorporate a learning rate reduction strategy: if the validation loss shows no improvement for 3 consecutive epochs, the learning rate is reduced to half of its original value.
+
+To control training time and avoid lengthy computations, we cap the maximum number of epochs for each trial at 100.
+It's noteworthy that none of the trials actually require this many epochs to converge effectively.
+
+# Summary of the architectures of best models
+
+The models with input sizes of 32 and 64 are configured with 4 convolutional blocks, whereas the model with an input size of 128 utilizes 5 convolutional blocks.
+Across all three models, there are 16 base convolutional filters applied.
+
+In terms of the dense layer, the model with an input size of 32 and the model with an input size of 128 both incorporate 64 units.
+However, the model with an input size of 64 employs a denser layer with 512 units.
+
+```{r}
+for (res in ALL_RES) {
+  cat(glue("Input size: {res}\n\n"))
+  print(mod[[res]])
+  cat("\n\n\n")
+}
+```
+
+Among the top five best models, it's notable that all of them consistently feature a minimum of 4 convolutional blocks.
+Additionally, they tend to share the use of 16 filters in their convolutional layers, suggesting this as a common and effective choice.
+
+Interestingly, the number of units in the dense layers appears to have less impact on performance, as it varies significantly across these top models.
+In most cases, a range of 64 to 256 units seems to be sufficient.
+
+Furthermore, it's observed that dropout rates between 0.1 to 0.4 are commonly chosen within these high-performing models.
+This indicates that moderate dropout rates within this range can effectively contribute to model robustness and generalization.
+
+Input size 32 five best models:
+
+- Blocks: 4, 5, 4, 5, 5
+- Filters: 16, 16, 8, 16, 4
+- Units: 64, 256, 8, 32, 8
+- Dropout: 0.4, 0.5, 0.2, 0.3, 0.1
+- Learning rate: 0.000512, 0.001024, 0.001024, 0.000512, 0.002048
+
+Input size 64 five best models:
+
+- Blocks: 4, 5, 4, 4, 5
+- Filters: 16, 4, 64, 32, 8
+- Units: 512, 16, 8, 64, 32
+- Dropout: 0.1, 0.1, 0.4, 0.4, 0.4
+- Learning rate: 0.002048, 0.002048, 0.000064, 0.001024, 0.008192
+
+Input size 128 five best models:
+
+- Blocks: 5, 5, 5, 5, 5
+- Filters: 16, 16, 8, 16, 64
+- Units: 64, 8, 128, 1024, 8
+- Dropout: 0.4, 0.1, 0.4, 0.5, 0.2
+- Learning rate: 0.000016, 0.008192, 0.000064, 0.000512, 0.000256
+
 # Basic summary
 
-```{r basic_summary}
+The models with varying input sizes exhibit similar performance, with a balanced accuracy score of approximately 0.83.
+However, the model with an input size of 64 performs better at predicting null plots but performs worse in detecting both heteroskedasticity and polynomial structures.
 
-basic_summary <- function(dat, res = 32L) {
-  dat %>%
-  ggplot() +
-  geom_bar(aes(pred, fill = type), position = "dodge") +
-  geom_text(data = dat %>%
-              group_by(truth, type, pred) %>%
+```{r}
+pred %>%
+  reduce(bind_rows) %>%
+  group_by(res, truth, type, pred) %>%
   summarise(prop = n()/2000)  %>%
-              filter(type == "heteroskedasticity"),
-  aes(pred, prop * 2000, label = format(prop, digits = 3)),
-  nudge_x = -0.25, nudge_y = 30) +
-  geom_text(data = dat %>%
-              group_by(truth, type, pred) %>%
-              summarise(prop = n()/2000) %>%
-              filter(type == "polynomial"),
-  aes(pred, prop * 2000, label = format(prop, digits = 3)),
-  nudge_x = 0.25, nudge_y = 30) +
-  xlab("Prediction") +
-  theme_light() +
-  facet_wrap(~truth, labeller = label_both) +
-  ylab("Count") +
-  scale_fill_brewer(palette = "Set2") +
-  ggtitle("Summary of prediction", subtitle = glue("Image size {res} * {res}"))
-}
-
-for (res in ALL_RES) {
-  print(basic_summary(pred[[res]], res))
-}
-
+  filter(truth == pred) %>%
+  select(-pred) %>%
+  pivot_wider(names_from = truth, values_from = prop) %>%
+  arrange(type, res) %>%
+  ungroup() %>%
+  select(-type) %>%
+  kableExtra::kable(col.names = c("Input size", "Accuracy: not_null", "Accuracy: null")) %>%
+  kableExtra::kable_material() %>%
+  kableExtra::pack_rows("Heteroskedasticity", 1, 3, 
+                        label_row_css = "background-color: #666; color: #fff;") %>%
+  kableExtra::pack_rows("Polynomial", 4, 6, 
+                        label_row_css = "background-color: #666; color: #fff;")
 ```
@@
 ```
+
+This outcome may come as somewhat surprising, as one might expect that a smaller input size could have a negative impact on the model's performance.
+However, it appears that even the smallest resolution we used in the test set is still sufficient for the model to perceive the visual patterns within the plots.
+This can be seen in one of the example test plots we provided, where, despite the image being only 32x32 pixels in size, the non-linear patterns are still easily noticeable.
+
+The situation may vary when plot details play a critical role in distinguishing between null and non-null plots.
+This can be especially relevant when dealing with plots containing a significant number of observations.
+
+```{r out.height="300px", out.width="300px"}
+knitr::include_graphics(here::here("data/experiment_factor/residual_plots/32/mixed/test/not_null/6743.png"))
+```
+
+# ROC
+
+The receiver operating characteristic (ROC) curve indicates that all three models perform almost identically when making predictions on the test set.
+
+```{r roc}
+pred %>%
+  reduce(bind_rows) %>%
+  group_by(res) %>%
+  roc_curve(truth, V1) %>%
+  autoplot() +
+  labs(col = "Input size") +
+  ggtitle("ROC")
+```
+
+# Distribution of predicted probability
+
+Based on the distribution of predicted probabilities, it is evident that in the majority of cases, all three models express high confidence in their predictions regarding non-null plots, with predicted probabilities very close to 1.
+Conversely, when it comes to null plots, all three models exhibit long-tail empirical distributions, with the peak of these distributions centered around 0.1.
+Additionally, it's worth noting that the distribution of predicted probabilities for the model with an input size of 32 shows a notable rightward shift.
+
+```{r p_hat_distribution}
+p_hat_distribution <- function(dat, res = 32) {
+  dat %>%
+    ggplot() +
+    geom_vline(xintercept = 0.5, col = "red") +
+    geom_histogram(aes(V1)) +
+    ggtitle(quote("Distribution of " ~ hat(P)(y == not_null) ~ "on the test set"), 
+            subtitle = glue("Input size {res} * {res}")) +
+    xlab(quote(hat(P)(y == not_null))) +
+    facet_grid(type ~ truth, labeller = label_both) +
+    theme_light() +
+    ylim(0, 1200)
+}
+
+for (res in ALL_RES) {
+  p_hat_distribution(pred[[res]], res) %>%
+    print()
+}
+```
+
+# Distribution of predicted probability on null set
+
+It's evident that all three models exhibit a higher likelihood of being triggered by lognormal and normal fitted value distributions.
+
+```{r p_hat_null_shape}
+pred %>%
+  reduce(bind_rows) %>%
+  filter(truth == "null") %>%
+  mutate(`Input size` = res) %>%
+  ggplot() +
+  geom_boxplot(aes(factor(x_dist), V1)) +
+  facet_wrap(~`Input size`, labeller = label_both) +
+  ylab(quote(hat(P)(y == not_null))) +
+  xlab("Distribution of fitted values") +
+  scale_x_discrete(labels = c("discrete", "lognormal", "normal", "uniform")) +
+  ggtitle(quote("Distribution of " ~ hat(P)(y == not_null ~"|"~ truth == null) ~ "on the test set")) +
+  theme_light()
+```
+
+With fewer observations in a plot, the models are easier to incorrectly reject null residual plots.
+
+```{r}
+pred %>%
+  reduce(bind_rows) %>%
+  filter(truth == "null") %>%
+  mutate(`Input size` = res) %>%
+  ggplot() +
+  geom_boxplot(aes(factor(n), V1)) +
+  facet_wrap(~`Input size`, labeller = label_both) +
+  xlab("Number of observations") +
+  ggtitle(quote("Distribution of " ~ hat(P)(y == not_null ~"|"~ truth == null) ~ "on the test set")) +
+  theme_light()
+
+```
+
+# Null plot examples
+
+Let's have a look at those null plots that being predicted as not null plots with small effect size.
+
+```{r}
+
+shap_plot <- function(plot_uids, save_path, res = 32L, response = "not_null", nsamples = 1000L, label = TRUE, width = 300) {
+  shap <- reticulate::import("shap")
+  np <- reticulate::import("numpy")
+  matplotlib <- reticulate::import("matplotlib")
+  matplotlib$pyplot$clf()
+  matplotlib$pyplot$cla()
+  matplotlib$pyplot$close()
+
+  background <- reticulate::iter_next(test[[res]])[[1]]
+
+  e <- shap$GradientExplainer(mod[[res]], background)
+
+  images <- list()
+  j <- 0
+  for (plot_uid in plot_uids) {
+    j <- j + 1
+    images[[j]] <- image_load(here(glue("data/experiment_factor/residual_plots/{res}/mixed/test/{response}/{plot_uid}.png")),
+                   target_size = c(res, res)) %>%
+      image_to_array()
+  }
+
+  if (length(images) == 1) 
+    im_arr <- array_reshape(images[[1]], c(1L, res, res, 3L))
+  else
+    im_arr <- np$stack(images)
+
+  shap_values <- e$shap_values(im_arr, nsamples = nsamples)
+
+  true_labels <- map(1:length(plot_uids), ~response)
+  if (length(images) == 1)
+    labels <- list("not_null", "null")
+  else
+    labels <- np$array(rep(c("not_null", "null"), length(images))) %>%
+      array_reshape(c(length(images), 2L))
+
+  if (label)
+    shap$image_plot(shap_values, 
+                    im_arr/255, 
+                    true_labels = true_labels, 
+                    labels = labels, 
+                    width = width, 
+                    show = FALSE)
+  else
+    shap$image_plot(shap_values, 
+                    im_arr/255, 
+                    width = width,
+                    show = FALSE)
+
+  if (!dir.exists(dirname(save_path))) 
+    dir.create(dirname(save_path), recursive = TRUE, showWarnings = FALSE)
+
+  matplotlib$pyplot$savefig(save_path, dpi = 300L)
+  matplotlib$pyplot$clf()
+  matplotlib$pyplot$cla()
+  matplotlib$pyplot$close()
+}
+
+```
+
+Let's have a look at those null plots that being predicted as not null plots with high confidence.
+
+```{r eval=FALSE}
+for (res in ALL_RES) {
+  pred[[res]] %>%
+    filter(truth == "null", pred == "not_null") %>%
+    slice_max(order_by = V1, n = 3) %>%
+    pull(plot_uid) %>%
+    map(~shap_plot(.x, here(glue("scripts/evaluation/temp_plots/{res}/{.x}.png")), res = res, response = "null", label = TRUE, nsamples = 100L))
+}
+```
+
+- Input size 32
+
+```{r}
+pred[[32]] %>% 
+  filter(truth == "null", pred == "not_null") %>% 
+  slice_max(order_by = V1, n = 3) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  rename(prob = V1) %>%
+  select(prob, truth, plot_uid, pred, x_dist, n, log_effect_size) %>%
+  kableExtra::kable() %>%
+  kableExtra::kable_material()
+```
+
+```{r}
+pred[[32]] %>%
+  filter(truth == "null", pred == "not_null") %>%
+  slice_max(order_by = V1, n = 3) %>%
+  pull(plot_uid) %>%
+  {here(glue("scripts/evaluation/temp_plots/32/{.}.png"))} %>%
+  knitr::include_graphics()
+```
+
+- Input size 64
+
+```{r}
+pred[[64]] %>% 
+  filter(truth == "null", pred == "not_null") %>% 
+  slice_max(order_by = V1, n = 3) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  rename(prob = V1) %>%
+  select(prob, truth, plot_uid, pred, x_dist, n, log_effect_size) %>%
+  kableExtra::kable() %>%
+  kableExtra::kable_material()
+```
+
+```{r}
+pred[[64]] %>%
+  filter(truth == "null", pred == "not_null") %>%
+  slice_max(order_by = V1, n = 3) %>%
+  pull(plot_uid) %>%
+  {here(glue("scripts/evaluation/temp_plots/64/{.}.png"))} %>%
+  knitr::include_graphics()
+```
+
+- Input size 128
+
+```{r}
+pred[[128]] %>% 
+  filter(truth == "null", pred == "not_null") %>% 
+  slice_max(order_by = V1, n = 3) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  rename(prob = V1) %>%
+  select(prob, truth, plot_uid, pred, x_dist, n, log_effect_size) %>%
+  kableExtra::kable() %>%
+  kableExtra::kable_material()
+```
+
+```{r}
+pred[[128]] %>%
+  filter(truth == "null", pred == "not_null") %>%
+  slice_max(order_by = V1, n = 3) %>%
+  pull(plot_uid) %>%
+  {here(glue("scripts/evaluation/temp_plots/128/{.}.png"))} %>%
+  knitr::include_graphics()
+```
+
+# Distribution of predicted probability on not null set
+
+Conditioning on the `shape`, `n` and `e_sigma` parameter.
+
+```{r}
+for (res in ALL_RES) {
+  p <- pred[[res]] %>%
+    filter(truth == "not_null") %>%
+    filter(type == "polynomial") %>%
+    mutate(e_sigma = factor(e_sigma, levels = c(4, 2, 1, 0.5))) %>%
+    ggplot() +
+    geom_hline(yintercept = 0.5, col = "red") +
+    geom_boxplot(aes(factor(shape), V1, fill = factor(n))) +
+    ylab(quote(hat(P)(y == not_null))) +
+    xlab("shape") +
+    scale_x_discrete(labels = c("U", "S", "M", "Triple-U")) +
+    ggtitle(quote("Distribution of " ~ hat(P)(y == not_null ~"|"~ truth == not_null) ~ "on the polynomial test set"), 
+            subtitle = glue("Input size {res} * {res}")) +
+    facet_grid(~e_sigma, labeller = label_both) +
+    scale_fill_brewer(palette = "YlOrRd") +
+    labs(fill = "n") +
+    theme_light()
+  print(p)
+}
+
+```
+
+Conditioning on the `shape`, `x_dist` and `e_sigma` parameter.
+
+```{r}
+for (res in ALL_RES) {
+  p <- pred[[32]] %>%
+    filter(truth == "not_null") %>%
+    filter(type == "polynomial") %>%
+    mutate(e_sigma = factor(e_sigma, levels = c(4, 2, 1, 0.5))) %>%
+    ggplot() +
+    geom_hline(yintercept = 0.5, col = "red") +
+    geom_boxplot(aes(factor(shape), V1, fill = factor(e_sigma))) +
+    ylab(quote(hat(P)(y == not_null))) +
+    xlab("Shape") +
+    scale_x_discrete(labels = c("U", "S", "M", "Triple-U")) +
+    ggtitle(quote("Distribution of " ~ hat(P)(y == not_null ~"|"~ truth == not_null) ~ "on the polynomial test set"), 
+            subtitle = glue("Input size {res} * {res}")) +
+    facet_grid(~x_dist, labeller = label_both) +
+    scale_fill_brewer(palette = "YlOrRd") +
+    labs(fill = quote(sigma)) +
+    theme_light()
+  print(p)
+}  
+```
+
+```{r}
+pred[[32]] %>%
+  filter(truth == "not_null") %>%
+  filter(type == "heteroskedasticity") %>%
+  ggplot() +
+  geom_hline(yintercept = 0.5, color = "red") +
+  geom_boxplot(aes(factor(a), V1, fill = factor(n))) +
+  facet_grid(x_dist~b, labeller = label_both) +
+  xlab("a") +
+  ggtitle(quote("Distribution of " ~ hat(P)(y == not_null ~"|"~ truth == null) ~ "on the heteroskedasticity test set")) +
+  theme_light() +
+  scale_fill_brewer(palette = "YlOrRd")
+```
+
+# Not null examples
+
+We are surprised by the model's performance in detecting violations. Here, we present residual with smallest effect size but get rejected.
+
+```{r eval=FALSE}
+for (res in ALL_RES) {
+  pred[[res]] %>%
+    filter(truth == "not_null", pred == "not_null") %>%
+    slice_min(order_by = effect_size, n = 3) %>%
+    pull(plot_uid) %>%
+    map(~shap_plot(.x, here(glue("scripts/evaluation/temp_plots/{res}/{.x}.png")), res = res, response = "not_null", label = TRUE, nsamples = 100L))
+}
+```
+
+- Input size 32
+
+```{r}
+pred[[32]] %>% 
+  filter(truth == "not_null", pred == "not_null") %>% 
+  slice_min(order_by = effect_size, n = 3) %>%
+  filter(V1 > 0.6) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  rename(prob = V1) %>%
+  select(prob, truth, plot_uid, pred, x_dist, n, type, log_effect_size) %>%
+  kableExtra::kable() %>%
+  kableExtra::kable_material()
+```
+
+```{r}
+pred[[32]] %>%
+  filter(truth == "not_null", pred == "not_null") %>% 
+  slice_min(order_by = effect_size, n = 3) %>%
+  filter(V1 > 0.6) %>%
+  pull(plot_uid) %>%
+  {here(glue("scripts/evaluation/temp_plots/32/{.}.png"))} %>%
+  knitr::include_graphics()
+```
+
+- Input size 64
+
+```{r}
+pred[[64]] %>% 
+  filter(truth == "not_null", pred == "not_null") %>% 
+  slice_min(order_by = effect_size, n = 3) %>%
+  filter(V1 > 0.6) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  rename(prob = V1) %>%
+  select(prob, truth, plot_uid, pred, x_dist, n, type, log_effect_size) %>%
+  kableExtra::kable() %>%
+  kableExtra::kable_material()
+```
+
+```{r}
+pred[[64]] %>%
+  filter(truth == "not_null", pred == "not_null") %>% 
+  slice_min(order_by = effect_size, n = 3) %>%
+  filter(V1 > 0.6) %>%
+  pull(plot_uid) %>%
+  {here(glue("scripts/evaluation/temp_plots/64/{.}.png"))} %>%
+  knitr::include_graphics()
+```
+
+- Input size 128
+
+```{r}
+pred[[128]] %>% 
+  filter(truth == "not_null", pred == "not_null") %>% 
+  slice_min(order_by = effect_size, n = 3) %>%
+  filter(V1 > 0.6) %>%
+  mutate(log_effect_size = log(effect_size)) %>%
+  rename(prob = V1) %>%
+  select(prob, truth, plot_uid, pred, x_dist, n, type, log_effect_size) %>%
+  kableExtra::kable() %>%
+  kableExtra::kable_material()
+```
+
+```{r}
+pred[[128]] %>%
+  filter(truth == "not_null", pred == "not_null") %>% 
+  slice_min(order_by = effect_size, n = 3) %>%
+  filter(V1 > 0.6) %>%
+  pull(plot_uid) %>%
+  {here(glue("scripts/evaluation/temp_plots/128/{.}.png"))} %>%
+  knitr::include_graphics()
+```
+
+# Performance on the extended test set
+
+```{r}
+ex_test <- list()
+for (res in ALL_RES) {
+  ex_test[[res]] <- flow_images_from_directory(here(glue("data/phn_extended/residual_plots/{res}/mixed/test")),
+                                            target_size = c(res, res),
+                                            shuffle = FALSE,
+                                            batch_size = 1000L)
+}
+```
+
+```{r}
+ex_meta <- readRDS(here(glue("data/phn_extended/residual_plots/meta.rds")))
+```
+
+```{r}
+ex_pred <- list()
+for (res in ALL_RES) {
+  ex_pred[[res]] <- mod[[res]]$predict(ex_test[[res]]) %>%
+    as.data.frame() %>%
+    mutate(truth = ex_test[[res]]$classes + 1) %>%
+    mutate(truth = names(ex_test[[res]]$class_indices)[truth]) %>%
+    mutate(filenames = ex_test[[res]]$filenames) %>%
+    mutate(plot_uid = gsub(".*/(.*).png", "\\1", filenames)) %>%
+    mutate(plot_uid = as.integer(plot_uid)) %>%
+    mutate(pred = V1 > 0.5) %>%
+    mutate(pred = ifelse(pred, "not_null", "null")) %>%
+    mutate(pred = factor(pred)) %>%
+    mutate(truth = factor(truth)) %>%
+    mutate(res = res)
+}
+```
+
+```{r}
+for (res in ALL_RES) {
+  ex_pred[[res]] <- ex_pred[[res]] %>%
+    left_join(ex_meta)
+
+  ex_pred[[res]] <- ex_pred[[res]] %>%
+    mutate(type = ifelse(!is.na(shape), "polynomial", "heteroskedasticity")) %>%
+    mutate(type = ifelse(!is.na(a), "heteroskedasticity", type)) %>%
+    mutate(type = ifelse(!is.na(e_dist), "non-normal", type)) %>%
+    mutate(x_dist = ifelse(x_dist == "even_discrete", "discrete", x_dist))
+}
+```
+
+```{r}
+ex_pred %>%
+  reduce(bind_rows) %>%
+  group_by(res, truth, type, pred) %>%
+  summarise(prop = n()/2000)  %>%
+  filter(truth == pred) %>%
+  select(-pred) %>%
+  pivot_wider(names_from = truth, values_from = prop) %>%
+  arrange(type, res) %>%
+  ungroup() %>%
+  select(-type) %>%
+  kableExtra::kable(col.names = c("Input size", "Accuracy: not_null", "Accuracy: null")) %>%
+  kableExtra::kable_material() %>%
+  kableExtra::pack_rows("Heteroskedasticity", 1, 3, 
+                        label_row_css = "background-color: #666; color: #fff;") %>%
+  kableExtra::pack_rows("Non-normal", 4, 6, 
+                        label_row_css = "background-color: #666; color: #fff;") %>%
+  kableExtra::pack_rows("Polynomial", 7, 8, 
+                        label_row_css = "background-color: #666; color: #fff;")
+```
+
+```{r}
+ex_pred %>%
+  reduce(bind_rows) %>%
+  group_by(res) %>%
+  bal_accuracy(truth, pred) %>%
+  kableExtra::kable(col.names = c("Input size", "Metric", "Estimator", "Estimate")) %>%
+  kableExtra::kable_material()
+```
+
+# Not null examples on the extened test set
+
+## Polynomial
+
+We check residual plots that are predicted as "null plots" with smallest $\sigma$ and ($\hat{p}(y=\text{null}|\text{not_null}) > 0.7$).
+
+```{r}
+
+ex_shap_plot <- function(plot_uids, save_path, res = 32L, response = "not_null", nsamples = 1000L, label = TRUE, width = 300) {
+  shap <- reticulate::import("shap")
+  np <- reticulate::import("numpy")
+  matplotlib <- reticulate::import("matplotlib")
+  matplotlib$pyplot$clf()
+  matplotlib$pyplot$cla()
+  matplotlib$pyplot$close()
+
+  background <- reticulate::iter_next(test[[res]])[[1]]
+
+  e <- shap$GradientExplainer(mod[[res]], background)
+
+  images <- list()
+  j <- 0
+  for (plot_uid in plot_uids) {
+    j <- j + 1
+    images[[j]] <- image_load(here(glue("data/phn_extended/residual_plots/{res}/mixed/test/{response}/{plot_uid}.png")),
+                   target_size = c(res, res)) %>%
+      image_to_array()
+  }
+
+  if (length(images) == 1) 
+    im_arr <- array_reshape(images[[1]], c(1L, res, res, 3L))
+  else
+    im_arr <- np$stack(images)
+
+  shap_values <- e$shap_values(im_arr, nsamples = nsamples)
+
+  true_labels <- map(1:length(plot_uids), ~response)
+  if (length(images) == 1)
+    labels <- list("not_null", "null")
+  else
+    labels <- np$array(rep(c("not_null", "null"), length(images))) %>%
+      array_reshape(c(length(images), 2L))
+
+  if (label)
+    shap$image_plot(shap_values, 
+                    im_arr/255, 
+                    true_labels = true_labels, 
+                    labels = labels, 
+                    width = width, 
+                    show = FALSE)
+  else
+    shap$image_plot(shap_values, 
+                    im_arr/255, 
+                    width = width,
+                    show = FALSE)
+
+  if (!dir.exists(dirname(save_path))) 
+    dir.create(dirname(save_path), recursive = TRUE, showWarnings = FALSE)
+
+  matplotlib$pyplot$savefig(save_path, dpi = 300L)
+  matplotlib$pyplot$clf()
+  matplotlib$pyplot$cla()
+  matplotlib$pyplot$close()
+}
+
+```
+
+```{r eval=FALSE}
+for (res in ALL_RES) {
+  ex_pred[[res]] %>%
+    filter(type == "polynomial") %>%
+    filter(truth == "not_null", pred == "null") %>%
+    filter(V2 > 0.7) %>%
+    slice_min(order_by = e_sigma, n = 3) %>%
+    pull(plot_uid) %>%
+    map(~ex_shap_plot(.x, here(glue("scripts/evaluation/temp_plots/{res}/ex/{.x}.png")), res = res, response = "not_null", label = TRUE, nsamples = 100L))
+}
+```
+
+- Input size 32
+
+```{r}
+ex_pred[[32]] %>% 
+  filter(type == "polynomial") %>%
+  filter(truth == "not_null", pred == "null") %>%
+  filter(V2 > 0.7) %>%
+  slice_min(order_by = e_sigma, n = 3) %>%
+  rename(prob = V2) %>%
+  select(prob, truth, plot_uid, pred, j, e_sigma, x_dist, x_sigma, x_k, x_even, n, type) %>%
+  kableExtra::kable() %>%
+  kableExtra::kable_material()
+```
+
+```{r}
+ex_pred[[32]] %>% 
+  filter(type == "polynomial") %>%
+  filter(truth == "not_null", pred == "null") %>%
+  filter(V2 > 0.7) %>%
+  slice_min(order_by = e_sigma, n = 3) %>%
+  pull(plot_uid) %>%
+  {here(glue("scripts/evaluation/temp_plots/32/ex/{.}.png"))} %>%
+  knitr::include_graphics()
+```
+
+- Input size 64
+
+```{r}
+ex_pred[[64]] %>% 
+  filter(type == "polynomial") %>%
+  filter(truth == "not_null", pred == "null") %>%
+  filter(V2 > 0.7) %>%
+  slice_min(order_by = e_sigma, n = 3) %>%
+  rename(prob = V2) %>%
+  select(prob, truth, plot_uid, pred, j, e_sigma, x_dist, x_sigma, x_k, x_even, n, type) %>%
+  kableExtra::kable() %>%
+  kableExtra::kable_material()
+```
+
+```{r}
+ex_pred[[64]] %>% 
+  filter(type == "polynomial") %>%
+  filter(truth == "not_null", pred == "null") %>%
+  filter(V2 > 0.7) %>%
+  slice_min(order_by = e_sigma, n = 3) %>%
+  pull(plot_uid) %>%
+  {here(glue("scripts/evaluation/temp_plots/64/ex/{.}.png"))} %>%
+  knitr::include_graphics()
+```
+
+- Input size 128
+
+```{r}
+ex_pred[[128]] %>% 
+  filter(type == "polynomial") %>%
+  filter(truth == "not_null", pred == "null") %>%
+  filter(V2 > 0.7) %>%
+  slice_min(order_by = e_sigma, n = 3) %>%
+  rename(prob = V2) %>%
+  select(prob, truth, plot_uid, pred, j, e_sigma, x_dist, x_sigma, x_k, x_even, n, type) %>%
+  kableExtra::kable() %>%
+  kableExtra::kable_material()
+```
+
+```{r}
+ex_pred[[128]] %>% 
+  filter(type == "polynomial") %>%
+  filter(truth == "not_null", pred == "null") %>%
+  filter(V2 > 0.7) %>%
+  slice_min(order_by = e_sigma, n = 3) %>%
+  pull(plot_uid) %>%
+  {here(glue("scripts/evaluation/temp_plots/128/ex/{.}.png"))} %>%
+  knitr::include_graphics()
+```
+
+## Heteroskedasticity
+
+We check residual plots that are predicted as "null plots" with greatest $b$ and ($\hat{p}(y=\text{null}|\text{not_null}) > 0.7$).
+
+```{r eval=FALSE}
+for (res in ALL_RES) {
+  ex_pred[[res]] %>%
+    filter(type == "heteroskedasticity") %>%
+    filter(truth == "not_null", pred == "null") %>%
+    filter(V2 > 0.7) %>%
+    slice_max(order_by = b, n = 3) %>%
+    pull(plot_uid) %>%
+    map(~ex_shap_plot(.x, here(glue("scripts/evaluation/temp_plots/{res}/ex/{.x}.png")), res = res, response = "not_null", label = TRUE, nsamples = 100L))
+}
+```
+
+- Input size 32
+
+```{r}
+ex_pred[[32]] %>% 
+  filter(type == "heteroskedasticity") %>%
+  filter(truth == "not_null", pred == "null") %>%
+  filter(V2 > 0.7) %>%
+  slice_max(order_by = b, n = 3) %>%
+  rename(prob = V2) %>%
+  select(prob, truth, plot_uid, pred, , x_dist, x_sigma, x_k, x_even, a, b, n, type) %>%
+  kableExtra::kable() %>%
+  kableExtra::kable_material()
+```
+
+```{r}
+ex_pred[[32]] %>% 
+  filter(type == "heteroskedasticity") %>%
+  filter(truth == "not_null", pred == "null") %>%
+  filter(V2 > 0.7) %>%
+  slice_max(order_by = b, n = 3) %>%
+  pull(plot_uid) %>%
+  {here(glue("scripts/evaluation/temp_plots/32/ex/{.}.png"))} %>%
+  knitr::include_graphics()
+```
+
+- Input size 64
+
+```{r}
+ex_pred[[64]] %>% 
+  filter(type == "heteroskedasticity") %>%
+  filter(truth == "not_null", pred == "null") %>%
+  filter(V2 > 0.7) %>%
+  slice_max(order_by = b, n = 3) %>%
+  rename(prob = V2) %>%
+  select(prob, truth, plot_uid, pred, , x_dist, x_sigma, x_k, x_even, a, b, n, type) %>%
+  kableExtra::kable() %>%
+  kableExtra::kable_material()
+```
+
+```{r}
+ex_pred[[64]] %>% 
+  filter(type == "heteroskedasticity") %>%
+  filter(truth == "not_null", pred == "null") %>%
+  filter(V2 > 0.7) %>%
+  slice_max(order_by = b, n = 3) %>%
+  pull(plot_uid) %>%
+  {here(glue("scripts/evaluation/temp_plots/64/ex/{.}.png"))} %>%
+  knitr::include_graphics()
+```
+
+- Input size 128
+
+```{r}
+ex_pred[[128]] %>% 
+  filter(type == "heteroskedasticity") %>%
+  filter(truth == "not_null", pred == "null") %>%
+  filter(V2 > 0.7) %>%
+  slice_max(order_by = b, n = 3) %>%
+  rename(prob = V2) %>%
+  select(prob, truth, plot_uid, pred, , x_dist, x_sigma, x_k, x_even, a, b, n, type) %>%
+  kableExtra::kable() %>%
+  kableExtra::kable_material()
+```
+
+```{r}
+ex_pred[[128]] %>% 
+  filter(type == "heteroskedasticity") %>%
+  filter(truth == "not_null", pred == "null") %>%
+  filter(V2 > 0.7) %>%
+  slice_max(order_by = b, n = 3) %>%
+  pull(plot_uid) %>%
+  {here(glue("scripts/evaluation/temp_plots/128/ex/{.}.png"))} %>%
+  knitr::include_graphics()
+```

### 2023-10-11 scripts/data_preparation/experiment_factor_pair.R
update pair data

@@ -1,0 +1,219 @@
+
+# library -----------------------------------------------------------------
+
+library(tidyverse)
+library(visage)
+library(here)
+library(glue)
+
+
+# multicore ---------------------------------------------------------------
+
+library(doMC)
+library(foreach)
+
+registerDoMC()
+
+cat(glue("{getDoParWorkers()} workers is used by `doMC`!"))
+
+set.seed(10086)
+
+# Global setting ----------------------------------------------------------
+
+# Number of samples per parameter
+SAMPLE_PER_PARAMETER <- list(train = 10, test = 1)
+
+# Number of parameter sets per model
+TOTAL_NUM_PARAMETER <- 2000
+
+# Data folder for saving plots
+DATA_FOLDER <- "data/experiment_factor/pair"
+
+# draw_plots --------------------------------------------------------------
+
+# The global uid for plots
+PLOT_UID <- 0
+
+# The global meta data for plots
+PLOT_META <- data.frame()
+
+# Draw plots for a violation model
+draw_plots <- function(violation, not_null, null, n, meta_vector) {
+  mod <- list()
+  mod$not_null <- not_null
+  mod$null <- null
+  
+  for (response in c("not_null", "null")) {
+    for (data_type in c("train", "test")) {
+      plot_dat <- map(1:SAMPLE_PER_PARAMETER[[data_type]], 
+                      ~mod[[response]]$gen_lineup(n, k = 2, pos = 2))
+      
+      # Speed up the plot drawing
+      num_plots <- length(plot_dat)
+      foreach(this_dat = plot_dat, 
+              this_plot_id = (PLOT_UID + 1):(PLOT_UID + num_plots)) %do% {
+                this_plot <- this_dat %>%
+                  VI_MODEL$plot_lineup(theme = theme_light(), 
+                                       remove_axis = TRUE, 
+                                       remove_legend = TRUE, 
+                                       remove_grid_line = TRUE) +
+                  theme(strip.text = element_blank())
+                
+                # The lineup layout contains 4 rows and 5 cols
+                ggsave(glue(here("{DATA_FOLDER}/native/{violation}/{data_type}/{response}/{this_plot_id}.png")), 
+                       this_plot, 
+                       width = 7/5 * 2, 
+                       height = 7/4)
+              }
+      
+      for (.unused in 1:num_plots) {
[…]

### 2023-10-12 scripts/training/experiment_factor_pair_32.py
update pair training script

@@ -1,0 +1,205 @@
+import sys
+import subprocess
+import os
+import tensorflow as tf
+from tensorflow import keras
+import keras_tuner
+
+# Global setttings
+DATA_CLASS = 'experiment_factor'
+INPUT_TYPE = 'pair'
+RES = 32
+COLOR_MODE = 'rgb'
+CHANNELS = 3
+CLASS_NAMES = ('not_null', "null")
+BATCH_SIZE = 32
+
+project_dir = subprocess.run(['Rscript', '-e', 'cat(here::here())'],
+                             check=True,
+                             capture_output=True,
+                             text=True).stdout
+
+train_dir = os.path.join(project_dir, 
+                         'data', 
+                         f'{DATA_CLASS}',
+                         f'{INPUT_TYPE}',
+                         f'{RES}',
+                         'mixed',
+                         'train')
+                         
+                         
+data_gen = keras.preprocessing.image.ImageDataGenerator(validation_split=0.2)
+train_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES * 2, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="training")
+val_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES * 2, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="validation")
+                             
+def build_model(hp):
+    
+    # Preprocess the input image
+    model_input = keras.layers.Input(shape=(RES * 2, RES, CHANNELS))
+    processed_input = keras.applications.vgg16.preprocess_input(model_input)
+    grey_scale_input = keras.layers.Lambda(
[…]

### 2023-10-13 scripts/data_preparation/experiment_factor_residual_plots.R
upload slurm script

@@ -43,10 +43,16 @@ draw_plots <- function(violation, not_null, null, n, meta_vector) {
   mod$not_null <- not_null
   mod$null <- null
   
+  result <- vector("list", length = 4)
+  result_counter <- 1
+  
   for (response in c("not_null", "null")) {
     for (data_type in c("train", "test")) {
       plot_dat <- map(1:SAMPLE_PER_PARAMETER[[data_type]], 
                       ~mod[[response]]$gen(n))
+      
+      result[[result_counter]] <- plot_dat
+      result_counter <- result_counter + 1
       
       # Speed up the plot drawing
       num_plots <- length(plot_dat)
@@ -75,6 +81,8 @@ draw_plots <- function(violation, not_null, null, n, meta_vector) {
       }
     }
   }
+  
+  return(list_flatten(result))
 }
 
 # get_x_var ---------------------------------------------------------------
@@ -135,21 +143,26 @@ poly_parameter_range <- list(shape = parameter_discrete(1, 4),
                              x_dist = parameter_choice(c("uniform", "normal", "lognormal", "even_discrete")),
                              n = parameter_choice(c(50, 100, 300)))
 
+poly_result <- vector("list", length = TOTAL_NUM_PARAMETER)
+
 for (i in 1:TOTAL_NUM_PARAMETER) {
   
   this_parameter <- map(poly_parameter_range, ~.x())
   
-  draw_plots(violation = "poly",
-             not_null = poly_model(shape = this_parameter$shape,
-                                   x = get_x_var(this_parameter$x_dist),
-                                   sigma = this_parameter$e_sigma),
-             null = poly_model(shape = this_parameter$shape,
-                               x = get_x_var(this_parameter$x_dist),
-                               include_z = FALSE,
-                               sigma = this_parameter$e_sigma),
-             n = this_parameter$n,
-             meta_vector = this_parameter)
+  poly_result[[i]] <- gen_plot_dat(violation = "poly",
+                                   not_null = poly_model(shape = this_parameter$shape,
+                                                         x = get_x_var(this_parameter$x_dist),
+                                                         sigma = this_parameter$e_sigma),
+                                   null = poly_model(shape = this_parameter$shape,
+                                                     x = get_x_var(this_parameter$x_dist),
+                                                     include_z = FALSE,
+                                                     sigma = this_parameter$e_sigma),
+                                   n = this_parameter$n,
+                                   meta_vector = this_parameter)
 }
+
[…]

### 2023-10-13 scripts/training/phn_extended_residual_plots_128.py
upload phn residual plots 128

@@ -1,0 +1,205 @@
+import sys
+import subprocess
+import os
+import tensorflow as tf
+from tensorflow import keras
+import keras_tuner
+
+# Global setttings
+DATA_CLASS = 'phn_extended'
+INPUT_TYPE = 'residual_plots'
+RES = 128
+COLOR_MODE = 'rgb'
+CHANNELS = 3
+CLASS_NAMES = ('not_null', "null")
+BATCH_SIZE = 32
+
+project_dir = subprocess.run(['Rscript', '-e', 'cat(here::here())'],
+                             check=True,
+                             capture_output=True,
+                             text=True).stdout
+
+train_dir = os.path.join(project_dir, 
+                         'data', 
+                         f'{DATA_CLASS}',
+                         f'{INPUT_TYPE}',
+                         f'{RES}',
+                         'mixed',
+                         'train')
+                         
+                         
+data_gen = keras.preprocessing.image.ImageDataGenerator(validation_split=0.2)
+train_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="training")
+val_dat = data_gen.flow_from_directory(directory=train_dir,
+                                         target_size=(RES, RES),
+                                         color_mode=COLOR_MODE,
+                                         classes=CLASS_NAMES,
+                                         class_mode='categorical',
+                                         batch_size=BATCH_SIZE,
+                                         shuffle=True,
+                                         seed=10086,
+                                         subset="validation")
+                             
+def build_model(hp):
+    
+    # Preprocess the input image
+    model_input = keras.layers.Input(shape=(RES, RES, CHANNELS))
+    processed_input = keras.applications.vgg16.preprocess_input(model_input)
+    grey_scale_input = keras.layers.Lambda(
+      lambda image: tf.image.rgb_to_grayscale(image), name='grey_scale'
[…]

### 2023-11-13 paper/paper.Rmd
update paper

@@
 title: |
-  Automated reading of residual plot with computer vision models
+  Automated reading of residual plots with computer vision models
 type: ARTICLE TEMPLATE
@@
       Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
-# bibliography: ref.bib
+bibliography: ref.bib
 abstract: |
@@
 ```{r}
-# OOP supports needed by `visage`
-# remotes::install_github("TengMCing/bandicoot")
-# 
 # Visual inference models and p-value calculation
-# remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.0.tar.gz")
+# remotes::install_github("TengMCing/visage)
 
@@
 
-# Methods for training computer vision models to analyze residual plots
+Residuals, within regression analysis, represent the differences between fitted values and observed data points, capturing the unexplained elements in the regression model.
+The practice of plotting residuals, advocated by influential regression literature [@cook1982residuals; @draper1998applied; @belsley1980regression; @montgomery1982introduction], serves as a standard procedure in regression diagnostics.
+This visual examination is crucial for identifying deviations from the model assumptions like linearity, homoscedasticity, and normality.
 
-In the field of statistical modeling and data analysis, interpreting residual plots plays a pivotal role in assessing the validity of our models.
-These plots can reveal deviations from modeling assumptions, including non-linearity, heteroskedasticity, and non-normality.
-The automation of residual plot analysis through computer vision provides a valuable tool for model validation and informed decision-making.
+Generating a residual plot in most statistical software is often as straightforward as executing a line of code or clicking a button.
+However, accurately interpreting a residual plot can be challenging.
+Consider Figure \ref{fig:false-finding} as an example, the residuals display a triangular shape pointing to the left.
+While this might suggest heteroskedasticity, it's important to avoid over-interpreting the visual pattern.
+In this case, the fitted model is correctly specified, and the triangular shape is actually a result of the skewed distribution of the predictors, rather than indicating a flaw in the model.
 
-This section discusses various methodologies for training computer vision models to effectively read and interpret residual plots.
-We explore three distinct approaches, each with its unique characteristics and strengths.
+A residual plot can exhibit various visual features, but it's crucial to recognize that some may arise from the characteristics of predictors and the inherent randomness of the error, rather than indicating a violation of model assumptions [@li2023plot].
+The concept of visual inference, as proposed by @buja2009statistical, provides an inferential framework to assess whether residual plots indeed contain visual patterns inconsistent with the model assumptions.
+The fundamental idea involves testing whether the actual residual plot significantly differs visually from null plots, which are created using residuals generated from the null distribution.
+Typically, this is accomplished through the lineup protocol.
+In this approach, the real residual plot is embedded within a lineup alongside several null plots.
+If the real residual plot can be distinguished from the lineup, it provides evidence for rejecting the null hypothesis.
 
-## Method 1: Single Plot Model
+Delivering a residual plot as a lineup is generally considered a good practice.
+However, as pointed out by @li2023plot, a primary limitation of the lineup protocol is its reliance on human judgments.
+Unlike conventional statistical tests that can be performed numerically and automatically in statistical software, the lineup protocol requires human evaluation of images.
+This characteristic makes it less suitable for large-scale applications, given the associated high labour costs and time requirements.
 
-The first method involves training a computer vision model to analyze a single residual plot and determine the probability that it was generated from a model without any violations.
-This method aligns with the principles of traditional deep learning classification problem, making it an intuitive and straightforward approach.
+In this paper, we attempt to address this issue by integrating computer vision models into residual plot diagnostics. The paper is structured as follows: ...
 
-To implement the Single Plot Model effectively, we utilize established neural network architectures like VGG16.
-The model is trained on a diverse dataset containing both "not null" plots (generated from model with violations) and "null" plots (generated from model without violations).
+```{r false-finding, fig.cap="An example residual vs fitted values plot (red line indicates 0).
+The vertical spread of the data points varies with the fitted values.
+This often indicates the existence of heteroskedasticity."}
+set.seed(10131)
+ori_x <- rand_lognormal()
+mod <- heter_model(b = 0, x = closed_form(~-ori_x))
+ori_dat <- mod$gen(300)
 
-This approach provides the model with certain capabilities:
+ori_dat %>%
+  VI_MODEL$plot(theme = theme_light(base_size = 12), size = 1, remove_grid_line = TRUE, ) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
 
-1.
-**Distinguishing null and not null plots**: Through training, the model is expected to be able to distinguish between null plots and not null plots.
-This distinction forms the foundation for identifying model violations in new residual plots.
+<!-- # Related works? -->
 
-2.
-**Capturing variance in null plots**: Null plots often exhibit varying shapes and outliers due to the randomness of the error term and the distribution of fitted values.
-The model learns to recognize and characterize this variance, aiding in the detection of violations.
+<!-- # Literature review -->
 
-3. **Recognizing patterns in not null plots**: The model would develop the ability to identify common patterns in not null plots generated through simulations.
+<!-- ## Visual inference in residual diagnostics -->
 
-4.
-**Analyzing Aesthetic Elements**: Beyond content, the model takes into account aesthetic elements within the residual plot, including point size, auxiliary lines, layers, background colors, and other visual attributes.
-This is an undesired capability but it is unavoidable if only a single design of plot is used as inputs.
+<!-- ## Existing approaches to automated plot analysis -->
 
-training sample:
+<!-- ### Computer vision approaches -->
 
-improvements:
-1. enrich the factors in simulation
-2. sigma same, shape same, same fitted value distribution, stack residual plots  
-3. test other residual plots generated from factors not seen by the model
-4. replace the output with the effect size
+<!-- ### Other approaches -->
 
-what the function tries to approximate?
+# Methodogology
 
-## Lineup model
+## Generation of simulated training data
 
-The second method involves training a computer vision model to analyze a lineup of residual plots, typically consisting of around 20 plots.
-It then determines whether the data plot within the lineup is a null plot or a not null plot.
-This method closely resembles a visual test, where the input is a lineup, and the output is a decision indicating whether the data plot exhibits visual features corresponding to model violations.
-This method still adheres to the standard classification scheme, allowing for the utilization of the same model architecture.
+## Architecture of the computer vision model
 
-This approach enable the model to learn all four points detailed in subsection 1 and three additional key points:
+## Training process and hyperparameter tuning
 
-1. **Locating the data plot**: It learns to identify and locate the data plot within a lineup of residual plots.
+# Results
 
-2. **Comparing data plot and null plots**: The model may potentially compare the data plot with null plots within the lineup. This provide additional information for decision making.
+## Model evaluation
 
-3. **Understanding variance within a lineup**: It may also understand the variance among null plots within a lineup.
+- Metrics for model performance
+- Shap values
+- Heatmap
 
-training sample:
+## Comparison with human visual inference
 
-improvement:
-1. similarly, other lineups generated from factors not seen by the model
-2. multiple lineups from each data plot, different positions
+### Overview of the human subject experiment
 
-## Position model
+### Comparison
 
-The third method shares similarities to the second method, but with a different output objective: determining the position of the data plot within the lineup.
-Essentially, this approach asks the model to locate the data plot within a set of residual plots.
-It resembles a visual testing procedure where an individual is asked to identify the most distinct plot from a lineup.
-Even though the data plot may not always be the most distinct, training the model exclusively on lineups with substantial effect sizes can yield effective results.
+- power comparison
+- decisions
 
-The Position Model can potentially learn all the aspects mentioned in method 2 while providing extra predictive information: the probability that the data plot resides in a position different from the true one.
-This information is valuable as it can be utilized to calculate the p-value of a visual test.
+## Case study 1: ...
 
-training sample:
+## Case study 2: ...
 
-## Effect size model
+## Case study 3: datasaurus
 
-##
+# Dicussion
+
+# Conclusion
+
+- Summary of findings
+- Contributions to the field
+- Future directions for research
+
+# Acknowledgements

### 2023-11-13 paper/paper.tex
update paper

@@
 
-\title{Automated reading of residual plot with computer vision models}
+\title{Automated reading of residual plots with computer vision models}
 
@@
 
-\hypertarget{methods-for-training-computer-vision-models-to-analyze-residual-plots}{%
-\section{Methods for training computer vision models to analyze residual
-plots}\label{methods-for-training-computer-vision-models-to-analyze-residual-plots}}
+Residuals, within regression analysis, represent the differences between
+fitted values and observed data points, capturing the unexplained
+elements in the regression model. The practice of plotting residuals,
+advocated by influential regression literature
+\citep{cook1982residuals, draper1998applied, belsley1980regression, montgomery1982introduction},
+serves as a standard procedure in regression diagnostics. This visual
+examination is crucial for identifying deviations from the model
+assumptions like linearity, homoscedasticity, and normality.
 
-In the field of statistical modeling and data analysis, interpreting
-residual plots plays a pivotal role in assessing the validity of our
-models. These plots can reveal deviations from modeling assumptions,
-including non-linearity, heteroskedasticity, and non-normality. The
-automation of residual plot analysis through computer vision provides a
-valuable tool for model validation and informed decision-making.
+Generating a residual plot in most statistical software is often as
+straightforward as executing a line of code or clicking a button.
+However, accurately interpreting a residual plot can be challenging.
+Consider Figure \ref{fig:false-finding} as an example, the residuals
+display a triangular shape pointing to the left. While this might
+suggest heteroskedasticity, it's important to avoid over-interpreting
+the visual pattern. In this case, the fitted model is correctly
+specified, and the triangular shape is actually a result of the skewed
+distribution of the predictors, rather than indicating a flaw in the
+model.
 
-This section discusses various methodologies for training computer
-vision models to effectively read and interpret residual plots. We
-explore three distinct approaches, each with its unique characteristics
-and strengths.
+A residual plot can exhibit various visual features, but it's crucial to
+recognize that some may arise from the characteristics of predictors and
+the inherent randomness of the error, rather than indicating a violation
+of model assumptions \citep{li2023plot}. The concept of visual
+inference, as proposed by \citet{buja2009statistical}, provides an
+inferential framework to assess whether residual plots indeed contain
+visual patterns inconsistent with the model assumptions. The fundamental
+idea involves testing whether the actual residual plot significantly
+differs visually from null plots, which are created using residuals
+generated from the null distribution. Typically, this is accomplished
+through the lineup protocol. In this approach, the real residual plot is
+embedded within a lineup alongside several null plots. If the real
+residual plot can be distinguished from the lineup, it provides evidence
+for rejecting the null hypothesis.
 
-\hypertarget{method-1-single-plot-model}{%
-\subsection{Method 1: Single Plot
-Model}\label{method-1-single-plot-model}}
+Delivering a residual plot as a lineup is generally considered a good
+practice. However, as pointed out by \citet{li2023plot}, a primary
+limitation of the lineup protocol is its reliance on human judgments.
+Unlike conventional statistical tests that can be performed numerically
+and automatically in statistical software, the lineup protocol requires
+human evaluation of images. This characteristic makes it less suitable
+for large-scale applications, given the associated high labour costs and
+time requirements.
 
-The first method involves training a computer vision model to analyze a
-single residual plot and determine the probability that it was generated
-from a model without any violations. This method aligns with the
-principles of traditional deep learning classification problem, making
-it an intuitive and straightforward approach.
+In this paper, we attempt to address this issue by integrating computer
+vision models into residual plot diagnostics. The paper is structured as
+follows: \ldots{}
 
-To implement the Single Plot Model effectively, we utilize established
-neural network architectures like VGG16. The model is trained on a
-diverse dataset containing both ``not null'' plots (generated from model
-with violations) and ``null'' plots (generated from model without
-violations).
+\begin{figure}
 
-This approach provides the model with certain capabilities:
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/false-finding-1}
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
+}
+
+\caption{An example residual vs fitted values plot (red line indicates 0).
+The vertical spread of the data points varies with the fitted values.
+This often indicates the existence of heteroskedasticity.}\label{fig:false-finding}
+\end{figure}
+
+\hypertarget{methodogology}{%
+\section{Methodogology}\label{methodogology}}
+
+\hypertarget{generation-of-simulated-training-data}{%
+\subsection{Generation of simulated training
+data}\label{generation-of-simulated-training-data}}
+
+\hypertarget{architecture-of-the-computer-vision-model}{%
+\subsection{Architecture of the computer vision
+model}\label{architecture-of-the-computer-vision-model}}
+
+\hypertarget{training-process-and-hyperparameter-tuning}{%
+\subsection{Training process and hyperparameter
+tuning}\label{training-process-and-hyperparameter-tuning}}
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{model-evaluation}{%
+\subsection{Model evaluation}\label{model-evaluation}}
+
+\begin{itemize}
+\tightlist
 \item
-  \textbf{Distinguishing null and not null plots}: Through training, the
-  model is expected to be able to distinguish between null plots and not
-  null plots. This distinction forms the foundation for identifying
-  model violations in new residual plots.
+  Metrics for model performance
 \item
-  \textbf{Capturing variance in null plots}: Null plots often exhibit
-  varying shapes and outliers due to the randomness of the error term
-  and the distribution of fitted values. The model learns to recognize
-  and characterize this variance, aiding in the detection of violations.
+  Shap values
 \item
-  \textbf{Recognizing patterns in not null plots}: The model would
-  develop the ability to identify common patterns in not null plots
-  generated through simulations.
+  Heatmap
+\end{itemize}
+
+\hypertarget{comparison-with-human-visual-inference}{%
+\subsection{Comparison with human visual
+inference}\label{comparison-with-human-visual-inference}}
+
+\hypertarget{overview-of-the-human-subject-experiment}{%
+\subsubsection{Overview of the human subject
+experiment}\label{overview-of-the-human-subject-experiment}}
+
+\hypertarget{comparison}{%
+\subsubsection{Comparison}\label{comparison}}
+
+\begin{itemize}
+\tightlist
 \item
-  \textbf{Analyzing Aesthetic Elements}: Beyond content, the model takes
-  into account aesthetic elements within the residual plot, including
-  point size, auxiliary lines, layers, background colors, and other
-  visual attributes. This is an undesired capability but it is
-  unavoidable if only a single design of plot is used as inputs.
-\end{enumerate}
+  power comparison
+\item
+  decisions
+\end{itemize}
 
-training sample:
+\hypertarget{case-study-1}{%
+\subsection{Case study 1: \ldots{}}\label{case-study-1}}
 
-improvements: 1. enrich the factors in simulation 2. sigma same, shape
-same, same fitted value distribution, stack residual plots\\
-3. test other residual plots generated from factors not seen by the
-model 4. replace the output with the effect size
+\hypertarget{case-study-2}{%
+\subsection{Case study 2: \ldots{}}\label{case-study-2}}
 
-what the function tries to approximate?
+\hypertarget{case-study-3-datasaurus}{%
+\subsection{Case study 3: datasaurus}\label{case-study-3-datasaurus}}
 
-\hypertarget{lineup-model}{%
-\subsection{Lineup model}\label{lineup-model}}
+\hypertarget{dicussion}{%
+\section{Dicussion}\label{dicussion}}
 
-The second method involves training a computer vision model to analyze a
-lineup of residual plots, typically consisting of around 20 plots. It
-then determines whether the data plot within the lineup is a null plot
-or a not null plot. This method closely resembles a visual test, where
-the input is a lineup, and the output is a decision indicating whether
-the data plot exhibits visual features corresponding to model
-violations. This method still adheres to the standard classification
-scheme, allowing for the utilization of the same model architecture.
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 
-This approach enable the model to learn all four points detailed in
-subsection 1 and three additional key points:
+\begin{itemize}
+\tightlist
+\item
+  Summary of findings
+\item
+  Contributions to the field
+\item
+  Future directions for research
+\end{itemize}
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  \textbf{Locating the data plot}: It learns to identify and locate the
-  data plot within a lineup of residual plots.
-\item
-  \textbf{Comparing data plot and null plots}: The model may potentially
-  compare the data plot with null plots within the lineup. This provide
-  additional information for decision making.
-\item
-  \textbf{Understanding variance within a lineup}: It may also
-  understand the variance among null plots within a lineup.
-\end{enumerate}
-
-training sample:
-
-improvement: 1. similarly, other lineups generated from factors not seen
-by the model 2. multiple lineups from each data plot, different
-positions
-
-\hypertarget{position-model}{%
-\subsection{Position model}\label{position-model}}
-
-The third method shares similarities to the second method, but with a
-different output objective: determining the position of the data plot
-within the lineup. Essentially, this approach asks the model to locate
-the data plot within a set of residual plots. It resembles a visual
-testing procedure where an individual is asked to identify the most
-distinct plot from a lineup. Even though the data plot may not always be
-the most distinct, training the model exclusively on lineups with
-substantial effect sizes can yield effective results.
-
-The Position Model can potentially learn all the aspects mentioned in
-method 2 while providing extra predictive information: the probability
-that the data plot resides in a position different from the true one.
-This information is valuable as it can be utilized to calculate the
-p-value of a visual test.
-
-training sample:
-
-\hypertarget{effect-size-model}{%
-\subsection{Effect size model}\label{effect-size-model}}
+\bibliographystyle{tfcad}
+\bibliography{ref.bib}
 

### 2023-11-20 paper/paper.Rmd
update paper

@@
   \usepackage{setspace}
+  \doublespacing
 output: rticles::tf_article
@@
 
-Delivering a residual plot as a lineup is generally considered a good practice.
-However, as pointed out by @li2023plot, a primary limitation of the lineup protocol is its reliance on human judgments.
+The practice of delivering a residual plot as a lineup is generally regarded as a valuable approach.
+Beyond its application in residual diagnostics, the lineup protocol has  integrated into the analysis of diverse subjects.
+For instance, \cite{loy2013diagnostic, loy2014hlmdiag, loy2015you} illustrate its applicability in diagnosing hierarchical linear models.
+Additionally, @widen2016graphical demonstrates its utility in geographical research, while @krishnan2021hierarchical explores its effectiveness in forensic examinations.
+
+However, as pointed out by @li2023plot, a primary limitation of the lineup protocol lies in its reliance on human judgments.
 Unlike conventional statistical tests that can be performed numerically and automatically in statistical software, the lineup protocol requires human evaluation of images.
@@
 
-In this paper, we attempt to address this issue by integrating computer vision models into residual plot diagnostics. The paper is structured as follows: ...
+There is a compelling need to develop an approach that alleviates people's workload by automating repetitive tasks and providing standardized results in a controlled environment.
+The large-scale evaluation of lineups is impractical without the use of technology and machines.
+
+The utilization of computers to interpret data plots has a rich history, with early efforts such as "Scagnostics" by @tukey1985computer, focusing on scatterplot diagnostics.
+@wilkinson2005graph expanded on this work in 2005, introducing graph theoretic scagnostics, which encompassed nine computable measures applied to planar proximity graphs.
+These measures, including "Outlying," "Skinny," "Stringy," "Straight," "Monotonic," "Skewed," "Clumpy," and "Striated," aimed to characterize outliers, shape, density, trend, and coherence of the data.
+
+While this approach has been inspiring, there is a recognition that it may not capture all the necessary visual features distinguishing actual residual plots from null plots.
+An alternative, more promising approach involves allowing computers to learn the visual feature extraction function.
+In essence, this means enabling computers to discern which visual features are significant for residual diagnostics and how to extract them.
+Modern computer vision models, a subset of artificial intelligence, present a compelling solution to this challenge.
+
+Modern computer vision models often rely on deep neural networks with convolutional layers [@fukushima1982neocognitron].
+These layers leverage hierarchical patterns in data, downsizing and transforming images by summarizing information in a small space.
+Numerous studies have demonstrated the efficacy of convolutional layers in addressing various vision tasks, including image recognition [@rawat2017deep].
+
+Despite the widespread use of computer vision models in fields like computer-aided diagnosis [@lee2015image], pedestrian detection [@brunetti2018computer], and facial recognition [@emami2012facial], their application in reading data plots remains limited.
+While some studies have explored the use of computer vision models for tasks such as reading recurrence plots for time series regression [@ojeda2020multivariate], time series classification [@chu2019automatic; @hailesilassie2019financial; @hatami2018classification; @zhang2020encoding], anomaly detection [@chen2020convolutional], and pairwise causality analysis [@singh2017deep], the application of reading residual plots with computer vision models represents a relatively new field of study.
+
+In this paper, we develop computer vision models and integrate them into the residual plots diagnostics workflow, filling the gap of.... The paper is structured as follows: ...
 
@@
 ```
-
-<!-- # Related works? -->
-
-<!-- # Literature review -->
-
-<!-- ## Visual inference in residual diagnostics -->
-
-<!-- ## Existing approaches to automated plot analysis -->
-
-<!-- ### Computer vision approaches -->
-
-<!-- ### Other approaches -->
 

### 2023-11-20 paper/paper.tex
update paper

@@
 \usepackage{setspace}
+\doublespacing
 
@@
 
-Delivering a residual plot as a lineup is generally considered a good
-practice. However, as pointed out by \citet{li2023plot}, a primary
-limitation of the lineup protocol is its reliance on human judgments.
-Unlike conventional statistical tests that can be performed numerically
-and automatically in statistical software, the lineup protocol requires
+The practice of delivering a residual plot as a lineup is generally
+regarded as a valuable approach. Beyond its application in residual
+diagnostics, the lineup protocol has integrated into the analysis of
+diverse subjects. For instance,
+\cite{loy2013diagnostic, loy2014hlmdiag, loy2015you} illustrate its
+applicability in diagnosing hierarchical linear models. Additionally,
+\citet{widen2016graphical} demonstrates its utility in geographical
+research, while \citet{krishnan2021hierarchical} explores its
+effectiveness in forensic examinations.
+
+However, as pointed out by \citet{li2023plot}, a primary limitation of
+the lineup protocol lies in its reliance on human judgments. Unlike
+conventional statistical tests that can be performed numerically and
+automatically in statistical software, the lineup protocol requires
 human evaluation of images. This characteristic makes it less suitable
@@
 
-In this paper, we attempt to address this issue by integrating computer
-vision models into residual plot diagnostics. The paper is structured as
-follows: \ldots{}
+There is a compelling need to develop an approach that alleviates
+people's workload by automating repetitive tasks and providing
+standardized results in a controlled environment. The large-scale
+evaluation of lineups is impractical without the use of technology and
+machines.
+
+The utilization of computers to interpret data plots has a rich history,
+with early efforts such as ``Scagnostics'' by \citet{tukey1985computer},
+focusing on scatterplot diagnostics. \citet{wilkinson2005graph} expanded
+on this work in 2005, introducing graph theoretic scagnostics, which
+encompassed nine computable measures applied to planar proximity graphs.
+These measures, including ``Outlying,'' ``Skinny,'' ``Stringy,''
+``Straight,'' ``Monotonic,'' ``Skewed,'' ``Clumpy,'' and ``Striated,''
+aimed to characterize outliers, shape, density, trend, and coherence of
+the data.
+
+While this approach has been inspiring, there is a recognition that it
+may not capture all the necessary visual features distinguishing actual
+residual plots from null plots. An alternative, more promising approach
+involves allowing computers to learn the visual feature extraction
+function. In essence, this means enabling computers to discern which
+visual features are significant for residual diagnostics and how to
+extract them. Modern computer vision models, a subset of artificial
+intelligence, present a compelling solution to this challenge.
+
+Modern computer vision models often rely on deep neural networks with
+convolutional layers \citep{fukushima1982neocognitron}. These layers
+leverage hierarchical patterns in data, downsizing and transforming
+images by summarizing information in a small space. Numerous studies
+have demonstrated the efficacy of convolutional layers in addressing
+various vision tasks, including image recognition \citep{rawat2017deep}.
+
+Despite the widespread use of computer vision models in fields like
+computer-aided diagnosis \citep{lee2015image}, pedestrian detection
+\citep{brunetti2018computer}, and facial recognition
+\citep{emami2012facial}, their application in reading data plots remains
+limited. While some studies have explored the use of computer vision
+models for tasks such as reading recurrence plots for time series
+regression \citep{ojeda2020multivariate}, time series classification
+\citep{chu2019automatic, hailesilassie2019financial, hatami2018classification, zhang2020encoding},
+anomaly detection \citep{chen2020convolutional}, and pairwise causality
+analysis \citep{singh2017deep}, the application of reading residual
+plots with computer vision models represents a relatively new field of
+study.
+
+In this paper, we develop computer vision models and integrate them into
+the residual plots diagnostics workflow, filling the gap of\ldots. The
+paper is structured as follows: \ldots{}
 

### 2024-01-15 paper/paper.Rmd
finish distance desc

@@
 
+<!-- Who is the audience? -->
+
+<!-- someone cares about linear regression -> statistics journals -->
+
+<!-- -> explain CNN design -->
+<!-- ->  -->
+
 # Introduction
 
-Residuals, within regression analysis, represent the differences between fitted values and observed data points, capturing the unexplained elements in the regression model.
-The practice of plotting residuals, advocated by influential regression literature [@cook1982residuals; @draper1998applied; @belsley1980regression; @montgomery1982introduction], serves as a standard procedure in regression diagnostics.
-This visual examination is crucial for identifying deviations from the model assumptions like linearity, homoscedasticity, and normality.
+The practice of plotting residuals is commonly regarded as a standard procedure in linear regression diagnostics \citep[see][]{cook1982residuals, belsley1980regression}.
+This visual assessment plays a crucial role in identifying deviations from model assumptions, such as linearity, homoscedasticity, and normality.
+It also helps in understanding the goodness of fit and various characteristics of the model.
 
@@
 Consider Figure \ref{fig:false-finding} as an example, the residuals display a triangular shape pointing to the left.
-While this might suggest heteroskedasticity, it's important to avoid over-interpreting the visual pattern.
+While this might suggest heteroskedasticity, it is important to avoid over-interpreting the visual pattern.
 In this case, the fitted model is correctly specified, and the triangular shape is actually a result of the skewed distribution of the predictors, rather than indicating a flaw in the model.
 
-A residual plot can exhibit various visual features, but it's crucial to recognize that some may arise from the characteristics of predictors and the inherent randomness of the error, rather than indicating a violation of model assumptions [@li2023plot].
+A residual plot can exhibit various visual features, but it is crucial to recognize that some may arise from the characteristics of predictors and the inherent randomness of the error, rather than indicating a violation of model assumptions [@li2023plot].
 The concept of visual inference, as proposed by @buja2009statistical, provides an inferential framework to assess whether residual plots indeed contain visual patterns inconsistent with the model assumptions.
-The fundamental idea involves testing whether the actual residual plot significantly differs visually from null plots, which are created using residuals generated from the null distribution.
+The fundamental idea involves testing whether the actual residual plot visually differs significantly from null plots, which are created using residuals generated from the null distribution.
 Typically, this is accomplished through the lineup protocol.
@@
 
-There is a compelling need to develop an approach that alleviates people's workload by automating repetitive tasks and providing standardized results in a controlled environment.
+There is a substantial need to develop an approach that alleviates people's workload by automating repetitive tasks and providing standardized results in a controlled environment.
 The large-scale evaluation of lineups is impractical without the use of technology and machines.
@@
 The utilization of computers to interpret data plots has a rich history, with early efforts such as "Scagnostics" by @tukey1985computer, focusing on scatterplot diagnostics.
-@wilkinson2005graph expanded on this work in 2005, introducing graph theoretic scagnostics, which encompassed nine computable measures applied to planar proximity graphs.
-These measures, including "Outlying," "Skinny," "Stringy," "Straight," "Monotonic," "Skewed," "Clumpy," and "Striated," aimed to characterize outliers, shape, density, trend, and coherence of the data.
-
-While this approach has been inspiring, there is a recognition that it may not capture all the necessary visual features distinguishing actual residual plots from null plots.
-An alternative, more promising approach involves allowing computers to learn the visual feature extraction function.
-In essence, this means enabling computers to discern which visual features are significant for residual diagnostics and how to extract them.
-Modern computer vision models, a subset of artificial intelligence, present a compelling solution to this challenge.
+@wilkinson2005graph expanded on this work, introducing graph theoretic scagnostics, which encompassed nine computable measures applied to planar proximity graphs.
+These measures, including "Outlying", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy", and "Striated" aimed to characterize outliers, shape, density, trend, and coherence of the data.
+While this approach has been inspiring, there is a recognition [@buja2009statistical] that it may not capture all the necessary visual features that differentiate actual residual plots from null plots.
+A more promising alternative entails enabling computers to learn the function for extracting visual features from residual plots.
+Essentially, this means empowering computers to discern the crucial visual features for residual diagnostics and determining the method to extract them.
+Modern computer vision models are well-suited for addressing this challenge.
 
@@
 Numerous studies have demonstrated the efficacy of convolutional layers in addressing various vision tasks, including image recognition [@rawat2017deep].
-
 Despite the widespread use of computer vision models in fields like computer-aided diagnosis [@lee2015image], pedestrian detection [@brunetti2018computer], and facial recognition [@emami2012facial], their application in reading data plots remains limited.
@@
 
+## Distance from the good residual plots
+
+In a visual test, the observer will be asked to choose one or more plots that stand out as most distinct from others in a given lineup.
+To develop a computer vision model for evaluating residual plots within the visual inference framework, it is important to precisely define a numerical measure of "difference" or "distance" between plots.
+This distance can take the form of a basic statistical operation on pixels, such as the sum of square differences.
+Alternatively, it could involve established image similarity metrics like the Structural Similarity Index Measure (SSIM) (ref here).
+The challenge lies in the fact that metrics tailored for image comparison may not be suitable for evaluating data plots, where only essential plot elements require assessment [@chowdhury2018measuring].
+
+### Residual distribution
+
+The distance metrics proposed in this paper takes into account the fact that we try to measure how different a residual plot is from a good residual plot, or in other words, how different a given fitted model is from a correctly specified model.
+For classical normal linear regression model, residuals are derived from the fitted values $\hat{\boldsymbol{y}}$ and observed values $\boldsymbol{y}$.
+However, if the data generating process is known and the model is correctly specified, by the Frisch-Waugh-Lowell theorem (ref here), residuals $\boldsymbol{e}$ can be written as a linear transformation of the error $\boldsymbol{\varepsilon}$:
+
+\begin{equation}
+\label{eq:residual-1}
+\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon},
+\end{equation}
+
+\noindent where $\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the residual operator, $\boldsymbol{X}$ is the design matrix, $\boldsymbol{I}_n$ is a $n$ by $n$ identity matrix, and $n$ is the number of observations.
+
+One of the assumptions of the classical normal linear regression model is the error $\boldsymbol{\varepsilon}$ follows a multivariate normal distribution with zero mean and constant variance, i.e., $N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$.
+Using equation \ref{eq:residual-1}, it can be known that residuals $\boldsymbol{e}$ also follow a certain probability distribution, which will be denoted as $Q$.
+This reference distribution $Q$ summarises what good residuals should follow given the predictors are known and fixed.
+
+Probability distribution $Q$ is actually a degenerate multivariate distribution due to the fact that $rank(\boldsymbol{R}) = n - 1 < n$.
+This means the $n$th residual value will be known given the $n - 1$ residuals.
+To capture the exact characteristics of $Q$, such as moments, we can simulate a large numbers of $\boldsymbol{\varepsilon}$ and transform it to $\boldsymbol{e}$.
+For simplicity, we replaced $\boldsymbol{R}$ with a diagonal matrix $diag(\boldsymbol{R})$.
+The resulting distribution for $\boldsymbol{Q}$ is $N(\boldsymbol{0}_n, diag(\boldsymbol{R}\sigma^2))$.
+
+If the model is misspecified, then the actual distribution of residuals denoted as $P$, will be different to $Q$.
+For example, if the model formula is correct, but the error $\boldsymbol{\varepsilon}$ follows a multivariate lognormal distribution, $P$ will not be the same as $Q$ and we would expect a skewed empirical distribution for residuals.
+And if there is an omitted variable problem, then the residual operator obtained from the fitted model will not be the same as $\boldsymbol{R}$, and again, $P$ will be different to $Q$.
+
+### Kullback-Leibler divergence of $P$ from $Q$
+
+Define a proper distance between distributions is usually easier than define a proper distance between data plots.
+Given the actual residual distribution $Q$ and the reference residual distribution $P$, we used a distance metric based on Kullback-Leibler divergence (ref here) to quantify the difference between two distributions
+
+\begin{align}
+D &= log\left(1 + KL\right), \\
+\label{eq:kl-1}
+KL &= \int_{\mathbb{R}^{n}}log\frac{p(\boldsymbol{e})}{q(\boldsymbol{e})}p(\boldsymbol{e})d\boldsymbol{e},
+\end{align}
+
+\noindent where $p(\boldsymbol{e})$ is the probability density function for distribution $P$, and $q(\boldsymbol{e})$ is the probability density function for distribution $Q$.
+
+This distance metric was first proposed in @li2023plot.
+It was mainly designed for measuring the effect size of non-linearity and heteroskedasticity in a residual plot.
+For a linear regression model that has omitted higher-order predictors $\boldsymbol{Z}$ along with coefficients $\boldsymbol{\beta}_z$, and an error distribution $N(\boldsymbol{0}_n, \boldsymbol{V})$, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))$.
+Note that the variance-covariance matrix is replaced with its diagonal analogy to ensure it is a full rank matrix.
+
+Since both $P$ and $Q$ are adjusted to be multivariate normal distributions, equation \ref{eq:kl-1} can be further expanded to
+
+\begin{align}
+\label{eq:kl-2}
+KL &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(diag(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
+\end{align}
+
+\noindent where $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$, and $\boldsymbol{W} = \boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'$.
+The assumed error variance $\sigma^2$ is set to be $tr(\boldsymbol{V})/n$, which is the expectation of the estimated variance.
+
+### Evaluation of Kullback-Leibler divergence for non-normal $P$
+
+For non-normal error $\boldsymbol{\varepsilon}$, the actual residual distribution $P$ is unlikely to be a multivariate normal distribution.
+Thus, equation \ref{eq:kl-2} will not be applicable to linear regression model with non-normality issue.
+
+To evaluate the Kullback-Leibler divergence of non-normal $P$ from $Q$, we need to solve equation \ref{eq:kl-1}.
+However, since $\boldsymbol{e}$ is a linear transformation of non-normal random variables, it is very common that the general form of $P$ is unknown, meaning that we can not easily compute $p(\boldsymbol{e})$ using a well-known probability density function.
+
+Additionally, even if $p(\boldsymbol{e})$ can be calculated for any $\boldsymbol{e} \in \mathbb{R}^n$, it will be very difficult to do numerical integration over the $n$ dimensional space, because $n$ could be potentially very large.
+
+In order to evaluate equation \ref{eq:kl-1} in a practically computable manner, the elements of $\boldsymbol{e}$ are assumed to be independent of each other.
+This assumption solves both of the issues mentioned above.
+First, we no longer need to integrate over $n$ random variables.
+The result is now the sum of the Kullback-Leibler divergence evaluated for each individual residual thanks for the independence assumption.
+Second, it is not required to know the joint probability density $p(\boldsymbol{e})$ any more.
+Instead, the evaluation of Kullback-Leibler divergence for a single residual relies on the knowledge of $p(e_i)$, where $e_i$ is the $i$th residual for $i = 1, ..., n$.
+This is much easier to estimate through simulation.
+
+The algorithm to compute equation \ref{eq:kl-1} starts from simulating $m$ sets of $\boldsymbol{\varepsilon}$ according to the error distribution.
+The simulated errors are stored in a matrix $\boldsymbol{A}$ with $n$ rows and $m$ columns.
+So each column of $\boldsymbol{A}$ is a set of realization values of $\boldsymbol{\varepsilon}$.
+Then, we can get $m$ sets of $\boldsymbol{e}$ stored in the matrix $\boldsymbol{B}$ by applying the residual rotation operator $\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}$.
+Furthermore, kernel density estimation (ref here) is applied on each row of $B$ to estimate $p(e_i)$ for $i = 1, ..., n$.
+
+Since the Kullback-Leibler divergence can be viewed as the expectation of the log-likelihood ratio between distribution $P$ and distribution $Q$ evaluated on distribution $P$, we can reuse the simulated residuals in matrix $\boldsymbol{B}$ to compute the expectation.
+Therefore, the Kullback-Leibler divergence can be formulated as
+
+\begin{align}
+\label{eq:kl-3}
+KL &= \sum_{i = 1}^{n} KL_i, \\
+KL_i &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{p(B_{ij})}{q(B_{ij})},
+\end{align}
+
+\noindent where $KL_i$ is the Kullback-Leibler divergence for an individual residual $e_i$, $B_{ij}$ is the $i$th row and $j$th column entry of the matrix $B$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\hat{\sigma^2} = var(flatten(B))$, and $flatten(.)$ unfolds a matrix into a vector.
+
+### Approximation of the distance metric with a residual plot
+
+In the previous sections, we have defined a distance metric for quantifying the difference between the actual residual distribution and an ideal reference distribution.
+This distance metric can only be computed when the data generating process is known.
+In reality, we often have no knowledge about the data generating process, otherwise we do not need to fit a linear regression model in the first place.
+
+What we proposed in this paper is a method to approximate this distance with a residual plot.
+This is done by training computer vision models with a large number of pairs of residual plots and distance.
+With the approximated distance, we will be able to know how different the underlying distribution of the residuals is from a good residual distribution.
+This is meaningful way to check if a residual plot is a good residual plot, and to know the strength of the visual signal indicating model violations.
+
+The approximated distance is not expected to be the same as the original distance.
+This is largely because information embedded in a single residual plot is limited and it may not be able to summarise the characteristics of the residual distribution.
+For a given residual distribution $P$, we can generate many different residual plots.
+Some of them share similar visual patterns, but some of them could be visually very different from the rest, especially for models with small $n$.
+This suggests the approximated distance will vary depends on whether the observed residual plot is representative or not.
+
 ## Generation of simulated training data
+
+The computer vision models we developed in this study were trained with synthetic data.
+This means the training data and the test data were simulated from a known data generating process under different parameter settings.
+
+We have incorporated three types of residual departures of linear regression model in this study, including non-linearity, heteroskedasticity and non-normality.
+
+<!-- The training data and test data was simulated using the following data   -->
+
+<!-- The data was simulated using the following data generating process -->
+
+\begin{align*}
+\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 + \beta_2(\boldsymbol{z} + \beta_1\boldsymbol{w}) + \boldsymbol{k} \odot \boldsymbol{\varepsilon}, \\
+\boldsymbol{z} &= He_j(g(\boldsymbol{x}_1, 2)), \\
+\boldsymbol{w} &= He_j(g(\boldsymbol{x}_2, 2)), \\
+\boldsymbol{k} &= \sqrt{\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^2}, \\
+\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{I}_n),
+\end{align*}
+
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{z}$, $\boldsymbol{w}$, $\boldsymbol{k}$ and $\boldsymbol{\varepsilon}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $He_j(.)$ is the $j$th-order probabilist's Hermite polynomials (ref), the $\sqrt{(.)}$ and $(.)^2$ operator are element-wise operators, $\odot$ is the Hadamard product, and $g(., k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+
+$$g(\boldsymbol{x}, k) = 2k \cdot \frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k\boldsymbol{1}_n,~for~k > 0,$$
+\noindent where $x_{min} = \underset{i \in \{ 1,...,n\}}{min} x_i$, $x_{max} = \underset{i \in \{ 1,...,n\}}{max} x_i$ and $x_i$ is the $i$-th entry of $\boldsymbol{x}$.
+
+## Different configurations of the model formula
+
+discuss different types of inputs and output choices, and why we choose the current design
 
@@
 
+## Model evaluation methods
+
 # Results
 
-## Model evaluation
+## Best model performance
 
@@
 
-## Case study 1: ...
+## When the model works
 
-## Case study 2: ...
+- simple examples (non-linearity, heteroskedasticity, ...)
+- datasaurus
 
-## Case study 3: datasaurus
+## When the model does not work
+
+- human detect but model does not
+- cartoon residuals?
+
+## Workflow: how one use this model? (small showcase)
 

### 2024-01-15 paper/paper.tex
finish distance desc

@@
 
-Residuals, within regression analysis, represent the differences between
-fitted values and observed data points, capturing the unexplained
-elements in the regression model. The practice of plotting residuals,
-advocated by influential regression literature
-\citep{cook1982residuals, draper1998applied, belsley1980regression, montgomery1982introduction},
-serves as a standard procedure in regression diagnostics. This visual
-examination is crucial for identifying deviations from the model
-assumptions like linearity, homoscedasticity, and normality.
+The practice of plotting residuals is commonly regarded as a standard
+procedure in linear regression diagnostics
+\citep[see][]{cook1982residuals, belsley1980regression}. This visual
+assessment plays a crucial role in identifying deviations from model
+assumptions, such as linearity, homoscedasticity, and normality. It also
+helps in understanding the goodness of fit and various characteristics
+of the model.
 
@@
 display a triangular shape pointing to the left. While this might
-suggest heteroskedasticity, it's important to avoid over-interpreting
+suggest heteroskedasticity, it is important to avoid over-interpreting
 the visual pattern. In this case, the fitted model is correctly
@@
 
-A residual plot can exhibit various visual features, but it's crucial to
-recognize that some may arise from the characteristics of predictors and
-the inherent randomness of the error, rather than indicating a violation
-of model assumptions \citep{li2023plot}. The concept of visual
+A residual plot can exhibit various visual features, but it is crucial
+to recognize that some may arise from the characteristics of predictors
+and the inherent randomness of the error, rather than indicating a
+violation of model assumptions \citep{li2023plot}. The concept of visual
 inference, as proposed by \citet{buja2009statistical}, provides an
@@
 visual patterns inconsistent with the model assumptions. The fundamental
-idea involves testing whether the actual residual plot significantly
-differs visually from null plots, which are created using residuals
+idea involves testing whether the actual residual plot visually differs
+significantly from null plots, which are created using residuals
 generated from the null distribution. Typically, this is accomplished
@@
 
-There is a compelling need to develop an approach that alleviates
+There is a substantial need to develop an approach that alleviates
 people's workload by automating repetitive tasks and providing
@@
 focusing on scatterplot diagnostics. \citet{wilkinson2005graph} expanded
-on this work in 2005, introducing graph theoretic scagnostics, which
-encompassed nine computable measures applied to planar proximity graphs.
-These measures, including ``Outlying,'' ``Skinny,'' ``Stringy,''
-``Straight,'' ``Monotonic,'' ``Skewed,'' ``Clumpy,'' and ``Striated,''
-aimed to characterize outliers, shape, density, trend, and coherence of
-the data.
-
-While this approach has been inspiring, there is a recognition that it
-may not capture all the necessary visual features distinguishing actual
-residual plots from null plots. An alternative, more promising approach
-involves allowing computers to learn the visual feature extraction
-function. In essence, this means enabling computers to discern which
-visual features are significant for residual diagnostics and how to
-extract them. Modern computer vision models, a subset of artificial
-intelligence, present a compelling solution to this challenge.
+on this work, introducing graph theoretic scagnostics, which encompassed
+nine computable measures applied to planar proximity graphs. These
+measures, including ``Outlying'', ``Skinny'', ``Stringy'', ``Straight'',
+``Monotonic'', ``Skewed'', ``Clumpy'', and ``Striated'' aimed to
+characterize outliers, shape, density, trend, and coherence of the data.
+While this approach has been inspiring, there is a recognition
+\citep{buja2009statistical} that it may not capture all the necessary
+visual features that differentiate actual residual plots from null
+plots. A more promising alternative entails enabling computers to learn
+the function for extracting visual features from residual plots.
+Essentially, this means empowering computers to discern the crucial
+visual features for residual diagnostics and determining the method to
+extract them. Modern computer vision models are well-suited for
+addressing this challenge.
 
@@
 various vision tasks, including image recognition \citep{rawat2017deep}.
-
 Despite the widespread use of computer vision models in fields like
@@
 
+\hypertarget{distance-from-the-good-residual-plots}{%
+\subsection{Distance from the good residual
+plots}\label{distance-from-the-good-residual-plots}}
+
+In a visual test, the observer will be asked to choose one or more plots
+that stand out as most distinct from others in a given lineup. To
+develop a computer vision model for evaluating residual plots within the
+visual inference framework, it is important to precisely define a
+numerical measure of ``difference'' or ``distance'' between plots. This
+distance can take the form of a basic statistical operation on pixels,
+such as the sum of square differences. Alternatively, it could involve
+established image similarity metrics like the Structural Similarity
+Index Measure (SSIM) (ref here). The challenge lies in the fact that
+metrics tailored for image comparison may not be suitable for evaluating
+data plots, where only essential plot elements require assessment
+\citep{chowdhury2018measuring}.
+
+\hypertarget{residual-distribution}{%
+\subsubsection{Residual distribution}\label{residual-distribution}}
+
+The distance metrics proposed in this paper takes into account the fact
+that we try to measure how different a residual plot is from a good
+residual plot, or in other words, how different a given fitted model is
+from a correctly specified model. For classical normal linear regression
+model, residuals are derived from the fitted values
+\(\hat{\boldsymbol{y}}\) and observed values \(\boldsymbol{y}\).
+However, if the data generating process is known and the model is
+correctly specified, by the Frisch-Waugh-Lowell theorem (ref here),
+residuals \(\boldsymbol{e}\) can be written as a linear transformation
+of the error \(\boldsymbol{\varepsilon}\):
+
+\begin{equation}
+\label{eq:residual-1}
+\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon},
+\end{equation}
+
+\noindent where
+\(\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\)
+is the residual operator, \(\boldsymbol{X}\) is the design matrix,
+\(\boldsymbol{I}_n\) is a \(n\) by \(n\) identity matrix, and \(n\) is
+the number of observations.
+
+One of the assumptions of the classical normal linear regression model
+is the error \(\boldsymbol{\varepsilon}\) follows a multivariate normal
+distribution with zero mean and constant variance, i.e.,
+\(N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)\). Using equation
+\ref{eq:residual-1}, it can be known that residuals \(\boldsymbol{e}\)
+also follow a certain probability distribution, which will be denoted as
+\(Q\). This reference distribution \(Q\) summarises what good residuals
+should follow given the predictors are known and fixed.
+
+Probability distribution \(Q\) is actually a degenerate multivariate
+distribution due to the fact that \(rank(\boldsymbol{R}) = n - 1 < n\).
+This means the \(n\)th residual value will be known given the \(n - 1\)
+residuals. To capture the exact characteristics of \(Q\), such as
+moments, we can simulate a large numbers of \(\boldsymbol{\varepsilon}\)
+and transform it to \(\boldsymbol{e}\). For simplicity, we replaced
+\(\boldsymbol{R}\) with a diagonal matrix \(diag(\boldsymbol{R})\). The
+resulting distribution for \(\boldsymbol{Q}\) is
+\(N(\boldsymbol{0}_n, diag(\boldsymbol{R}\sigma^2))\).
+
+If the model is misspecified, then the actual distribution of residuals
+denoted as \(P\), will be different to \(Q\). For example, if the model
+formula is correct, but the error \(\boldsymbol{\varepsilon}\) follows a
+multivariate lognormal distribution, \(P\) will not be the same as \(Q\)
+and we would expect a skewed empirical distribution for residuals. And
+if there is an omitted variable problem, then the residual operator
+obtained from the fitted model will not be the same as
+\(\boldsymbol{R}\), and again, \(P\) will be different to \(Q\).
+
+\hypertarget{kullback-leibler-divergence-of-p-from-q}{%
+\subsubsection{\texorpdfstring{Kullback-Leibler divergence of \(P\) from
+\(Q\)}{Kullback-Leibler divergence of P from Q}}\label{kullback-leibler-divergence-of-p-from-q}}
+
+Define a proper distance between distributions is usually easier than
+define a proper distance between data plots. Given the actual residual
+distribution \(Q\) and the reference residual distribution \(P\), we
+used a distance metric based on Kullback-Leibler divergence (ref here)
+to quantify the difference between two distributions
+
+\begin{align}
+D &= log\left(1 + KL\right), \\
+\label{eq:kl-1}
+KL &= \int_{\mathbb{R}^{n}}log\frac{p(\boldsymbol{e})}{q(\boldsymbol{e})}p(\boldsymbol{e})d\boldsymbol{e},
+\end{align}
+
+\noindent where \(p(\boldsymbol{e})\) is the probability density
+function for distribution \(P\), and \(q(\boldsymbol{e})\) is the
+probability density function for distribution \(Q\).
+
+This distance metric was first proposed in \citet{li2023plot}. It was
+mainly designed for measuring the effect size of non-linearity and
+heteroskedasticity in a residual plot. For a linear regression model
+that has omitted higher-order predictors \(\boldsymbol{Z}\) along with
+coefficients \(\boldsymbol{\beta}_z\), and an error distribution
+\(N(\boldsymbol{0}_n, \boldsymbol{V})\), \(Q\) can be represented as
+\(N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))\).
+Note that the variance-covariance matrix is replaced with its diagonal
+analogy to ensure it is a full rank matrix.
+
+Since both \(P\) and \(Q\) are adjusted to be multivariate normal
+distributions, equation \ref{eq:kl-1} can be further expanded to
+
+\begin{align}
+\label{eq:kl-2}
+KL &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(diag(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
+\end{align}
+
+\noindent where
+\(\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z\),
+and \(\boldsymbol{W} = \boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'\).
+The assumed error variance \(\sigma^2\) is set to be
+\(tr(\boldsymbol{V})/n\), which is the expectation of the estimated
+variance.
+
+\hypertarget{evaluation-of-kullback-leibler-divergence-for-non-normal-p}{%
+\subsubsection{\texorpdfstring{Evaluation of Kullback-Leibler divergence
+for non-normal
+\(P\)}{Evaluation of Kullback-Leibler divergence for non-normal P}}\label{evaluation-of-kullback-leibler-divergence-for-non-normal-p}}
+
+For non-normal error \(\boldsymbol{\varepsilon}\), the actual residual
+distribution \(P\) is unlikely to be a multivariate normal distribution.
+Thus, equation \ref{eq:kl-2} will not be applicable to linear regression
+model with non-normality issue.
+
+To evaluate the Kullback-Leibler divergence of non-normal \(P\) from
+\(Q\), we need to solve equation \ref{eq:kl-1}. However, since
+\(\boldsymbol{e}\) is a linear transformation of non-normal random
+variables, it is very common that the general form of \(P\) is unknown,
+meaning that we can not easily compute \(p(\boldsymbol{e})\) using a
+well-known probability density function.
+
+Additionally, even if \(p(\boldsymbol{e})\) can be calculated for any
+\(\boldsymbol{e} \in \mathbb{R}^n\), it will be very difficult to do
+numerical integration over the \(n\) dimensional space, because \(n\)
+could be potentially very large.
+
+In order to evaluate equation \ref{eq:kl-1} in a practically computable
+manner, the elements of \(\boldsymbol{e}\) are assumed to be independent
+of each other. This assumption solves both of the issues mentioned
+above. First, we no longer need to integrate over \(n\) random
+variables. The result is now the sum of the Kullback-Leibler divergence
+evaluated for each individual residual thanks for the independence
+assumption. Second, it is not required to know the joint probability
+density \(p(\boldsymbol{e})\) any more. Instead, the evaluation of
+Kullback-Leibler divergence for a single residual relies on the
+knowledge of \(p(e_i)\), where \(e_i\) is the \(i\)th residual for
+\(i = 1, ..., n\). This is much easier to estimate through simulation.
+
+The algorithm to compute equation \ref{eq:kl-1} starts from simulating
+\(m\) sets of \(\boldsymbol{\varepsilon}\) according to the error
+distribution. The simulated errors are stored in a matrix
+\(\boldsymbol{A}\) with \(n\) rows and \(m\) columns. So each column of
+\(\boldsymbol{A}\) is a set of realization values of
+\(\boldsymbol{\varepsilon}\). Then, we can get \(m\) sets of
+\(\boldsymbol{e}\) stored in the matrix \(\boldsymbol{B}\) by applying
+the residual rotation operator
+\(\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}\). Furthermore, kernel
+density estimation (ref here) is applied on each row of \(B\) to
+estimate \(p(e_i)\) for \(i = 1, ..., n\).
+
+Since the Kullback-Leibler divergence can be viewed as the expectation
+of the log-likelihood ratio between distribution \(P\) and distribution
+\(Q\) evaluated on distribution \(P\), we can reuse the simulated
+residuals in matrix \(\boldsymbol{B}\) to compute the expectation.
+Therefore, the Kullback-Leibler divergence can be formulated as
+
+\begin{align}
+\label{eq:kl-3}
+KL &= \sum_{i = 1}^{n} KL_i, \\
+KL_i &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{p(B_{ij})}{q(B_{ij})},
+\end{align}
+
+\noindent where \(KL_i\) is the Kullback-Leibler divergence for an
+individual residual \(e_i\), \(B_{ij}\) is the \(i\)th row and \(j\)th
+column entry of the matrix \(B\), \(q(.)\) is the normal density
+function with mean zero and an assumed variance estimated as
+\(\hat{\sigma^2} = var(flatten(B))\), and \(flatten(.)\) unfolds a
+matrix into a vector.
+
+\hypertarget{approximation-of-the-distance-metric-with-a-residual-plot}{%
+\subsubsection{Approximation of the distance metric with a residual
+plot}\label{approximation-of-the-distance-metric-with-a-residual-plot}}
+
+In the previous sections, we have defined a distance metric for
+quantifying the difference between the actual residual distribution and
+an ideal reference distribution. This distance metric can only be
+computed when the data generating process is known. In reality, we often
+have no knowledge about the data generating process, otherwise we do not
+need to fit a linear regression model in the first place.
+
+What we proposed in this paper is a method to approximate this distance
+with a residual plot. This is done by training computer vision models
+with a large number of pairs of residual plots and distance. With the
+approximated distance, we will be able to know how different the
+underlying distribution of the residuals is from a good residual
+distribution. This is meaningful way to check if a residual plot is a
+good residual plot, and to know the strength of the visual signal
+indicating model violations.
+
+The approximated distance is not expected to be the same as the original
+distance. This is largely because information embedded in a single
+residual plot is limited and it may not be able to summarise the
+characteristics of the residual distribution. For a given residual
+distribution \(P\), we can generate many different residual plots. Some
+of them share similar visual patterns, but some of them could be
+visually very different from the rest, especially for models with small
+\(n\). This suggests the approximated distance will vary depends on
+whether the observed residual plot is representative or not.
+
 \hypertarget{generation-of-simulated-training-data}{%
@@
 data}\label{generation-of-simulated-training-data}}
+
+The computer vision models we developed in this study were trained with
+synthetic data. This means the training data and the test data were
+simulated from a known data generating process under different parameter
+settings.
+
+We have incorporated three types of residual departures of linear
+regression model in this study, including non-linearity,
+heteroskedasticity and non-normality.
+
+\begin{align*}
+\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 + \beta_2(\boldsymbol{z} + \beta_1\boldsymbol{w}) + \boldsymbol{k} \odot \boldsymbol{\varepsilon}, \\
+\boldsymbol{z} &= He_j(g(\boldsymbol{x}_1, 2)), \\
+\boldsymbol{w} &= He_j(g(\boldsymbol{x}_2, 2)), \\
+\boldsymbol{k} &= \sqrt{\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^2}, \\
+\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{I}_n),
+\end{align*}
+
+\noindent where \(\boldsymbol{y}\), \(\boldsymbol{x}_1\),
+\(\boldsymbol{x}_2\), \(\boldsymbol{z}\), \(\boldsymbol{w}\),
+\(\boldsymbol{k}\) and \(\boldsymbol{\varepsilon}\) are vectors of size
+\(n\), \(\boldsymbol{1}_n\) is a vector of ones of size \(n\),
+\(He_j(.)\) is the \(j\)th-order probabilist's Hermite polynomials
+(ref), the \(\sqrt{(.)}\) and \((.)^2\) operator are element-wise
+operators, \(\odot\) is the Hadamard product, and \(g(., k)\) is a
+scaling function to enforce the support of the random vector to be
+\([-k, k]^n\) defined as
+
+\[g(\boldsymbol{x}, k) = 2k \cdot \frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k\boldsymbol{1}_n,~for~k > 0,\]
+\noindent where \(x_{min} = \underset{i \in \{ 1,...,n\}}{min} x_i\),
+\(x_{max} = \underset{i \in \{ 1,...,n\}}{max} x_i\) and \(x_i\) is the
+\(i\)-th entry of \(\boldsymbol{x}\).
+
+\hypertarget{different-configurations-of-the-model-formula}{%
+\subsection{Different configurations of the model
+formula}\label{different-configurations-of-the-model-formula}}
+
+discuss different types of inputs and output choices, and why we choose
+the current design
 
@@
 
+\hypertarget{model-evaluation-methods}{%
+\subsection{Model evaluation methods}\label{model-evaluation-methods}}
+
 \hypertarget{results}{%
@@
 
-\hypertarget{model-evaluation}{%
-\subsection{Model evaluation}\label{model-evaluation}}
+\hypertarget{best-model-performance}{%
+\subsection{Best model performance}\label{best-model-performance}}
 
@@
 
-\hypertarget{case-study-1}{%
-\subsection{Case study 1: \ldots{}}\label{case-study-1}}
+\hypertarget{when-the-model-works}{%
+\subsection{When the model works}\label{when-the-model-works}}
 
-\hypertarget{case-study-2}{%
-\subsection{Case study 2: \ldots{}}\label{case-study-2}}
+\begin{itemize}
+\tightlist
+\item
+  simple examples (non-linearity, heteroskedasticity, \ldots)
+\item
+  datasaurus
+\end{itemize}
 
-\hypertarget{case-study-3-datasaurus}{%
-\subsection{Case study 3: datasaurus}\label{case-study-3-datasaurus}}
+\hypertarget{when-the-model-does-not-work}{%
+\subsection{When the model does not
+work}\label{when-the-model-does-not-work}}
+
+\begin{itemize}
+\tightlist
+\item
+  human detect but model does not
+\item
+  cartoon residuals?
+\end{itemize}
+
+\hypertarget{workflow-how-one-use-this-model-small-showcase}{%
+\subsection{Workflow: how one use this model? (small
+showcase)}\label{workflow-how-one-use-this-model-small-showcase}}
 

### 2024-01-22 paper/paper.Rmd
update paper methodology and toc

@@
 
-<!-- Who is the audience? -->
-
-<!-- someone cares about linear regression -> statistics journals -->
-
-<!-- -> explain CNN design -->
-<!-- ->  -->
+\newpage
+\tableofcontents
+\newpage
 
@@
 There is a substantial need to develop an approach that alleviates people's workload by automating repetitive tasks and providing standardized results in a controlled environment.
-The large-scale evaluation of lineups is impractical without the use of technology and machines.
+The large-scale evaluation of lineups ihs impractical without the use of technology and machines.
 
@@
 @wilkinson2005graph expanded on this work, introducing graph theoretic scagnostics, which encompassed nine computable measures applied to planar proximity graphs.
-These measures, including "Outlying", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy", and "Striated" aimed to characterize outliers, shape, density, trend, and coherence of the data.
+These measures, including, but not limited to, "Outlying", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy", and "Striated" aimed to characterize outliers, shape, density, trend, coherence and other characteristics of the data.
 While this approach has been inspiring, there is a recognition [@buja2009statistical] that it may not capture all the necessary visual features that differentiate actual residual plots from null plots.
@@
 
-```{r false-finding, fig.cap="An example residual vs fitted values plot (red line indicates 0).
+```{r false-finding, fig.width = 8, fig.height = 6, fig.cap="An example residual vs fitted values plot (red line indicates 0).
 The vertical spread of the data points varies with the fitted values.
@@
 ori_dat %>%
-  VI_MODEL$plot(theme = theme_light(base_size = 12), size = 1, remove_grid_line = TRUE, ) +
+  VI_MODEL$plot(theme = theme_light(), remove_grid_line = TRUE) +
   xlab("Fitted values") +
@@
 
+## Different possible configurations of the model formula
+
+<!-- - discuss different types of inputs and output choices, and why we choose the current design -->
+
+In addressing a problem, there exist multiple avenues for solutions.
+Various configurations of the computer vision model formula can effectively read residual plots, and these configurations can be categorized based on two key components of the model formula: the input and the output format.
+
+### Different input formats
+
+In computer vision models, the input plays a role similar to predictors in a linear regression model.
+The quality and relevance of the input data greatly influence the model's capacity to generate insightful and meaningful results.
+A straightforward approach involves feeding the model a vector of residuals along with a vector of fitted values, essentially providing all the necessary details for creating a residuals vs fitted values plot.
+However, a drawback of this method is the dynamic input size, as different fitted regression models may have varying numbers of observations.
+In modern computer vision models created with mainstream software like TensorFlow (ref here), the input is typically a fixed-shape tensor.
+One solution is to pad the input vector with leading or trailing zeros if the input tensor expects a longer vector.
+However, this approach may fail if the input vector surpasses the designed length.
+Another strategy is to summarize the residuals and fitted values separately using histograms and utilize the counts as the input.
+By controlling the number of bins in the histograms, it becomes possible to provide a fixed-length input vector.
+
+The primary advantage of using an image as input, as opposed to a vector input format, relates to the development of sophisticated image processing architectures over the years (see ref here).
+These architectures effectively capture and summarize spatial information from nearby pixels, which is less straightforward with vector input.
+The only considerations are the image resolution and the aesthetics of the residual plot.
+In general, higher resolution provides more information to the model but comes with the trade-off of increased complexity and greater difficulty in training.
+As for the aesthetics of the residual plot, a practical solution is to consistently present residual plots in the same style to the model.
+This implies that the model can not accept arbitrary images as input but requires the use of the same preprocessing software to convert provided residuals and fitted values into a standardized style for the residual plot.
+
+Other possible input formats include, but are not limited to, a pair of residual plots, a triplet, and a lineup.
+Existing computer vision models designed for image comparison can assess whether a pair of images are similar or dissimilar (as seen in LeCun's 2005 paper).
+Applied to our specific problem, we can define null plots of a fitted model to be similar to each other, while considering actual residual plots to be distinct from null plots of any fitted model.
+A triplet constitutes a set of three images, denoted as $image_1$, $image_2$ and $image_3$.
+In computer vision models, triplets are often used to predict whether $image_2$ or $image_3$ is more similar to $image_1$, proving particularly useful for establishing rankings between samples.
+In the context of our problem, we can apply the same criteria to define similarity between images.
+However, it's important to note that these two approaches usually require additional considerations regarding the loss function and, at times, non-standard training processes due to shared weights between different convolutional blocks.
+
+Presenting a lineup to a model and asking it to predict which residual plot is the most different one aligns closely with the lineup protocol.
+However, if the lineup consists of around 20 plots, similar to previous human subject experiments (reference here), the resolution of the input image could become quite large, posing challenges in training the model.
+In fact, we experimented with this approach in a pilot study, and the model's performance was suboptimal.
+
+We did not explore all the mentioned input formats due to the considerable costs associated with data preparation and model training.
+Considering the implementation cost and the interpretability of the model, we settled on the single residual plot input format.
+
+### Different output formats
+
+Given that the input is a single residual plot represented as a fixed-resolution image, the computer vision model's output can take one of two formats: binary or numeric.
+This choice determines whether the model is a classification model or a regression model.
+The binary outcome might indicates whether the input image is a null plot or not, or whether the input image would be rejected in a visual test conducted by humans.
+The latter option requires data from prior human subject experiments, posing challenges in controlling the quality and quantity of data due to variations in experimental settings across different studies.
+Additionally, some visual inference experiments are unrelated to linear regression models and residual plots, resulting in a limited amount of available training data.
+
+Alternatively, the output could be a meaningful numeric measure, such as the average distance to good residual plots.
+This approach necessitates defining a distance measure between residual plots, which may be a non-trivial task.
+However, once a meaningful distance measure is established, the computer vision model's prediction becomes an interpretable value with tangible significance.
+Studies have demonstrated that defining a proper distance between images can enhance the matching accuracy in image search compared to a binary outcome model (ref here).
+
+The following section will propose the distance measure used for training the computer vision model.
+
 ## Distance from the good residual plots
@@
 This distance can take the form of a basic statistical operation on pixels, such as the sum of square differences.
-Alternatively, it could involve established image similarity metrics like the Structural Similarity Index Measure (SSIM) (ref here).
+Alternatively, it could involve established image similarity metrics like the Structural Similarity Index Measure [@wang2004image].
 The challenge lies in the fact that metrics tailored for image comparison may not be suitable for evaluating data plots, where only essential plot elements require assessment [@chowdhury2018measuring].
+Furthermore, scagnostics mentioned in Section \ref{introduction} could be used to construct distance metrics for data plots comparison, but the functional form still needs to be carefully refined to accurately reflect the extent of the violations.
 
@@
 The distance metrics proposed in this paper takes into account the fact that we try to measure how different a residual plot is from a good residual plot, or in other words, how different a given fitted model is from a correctly specified model.
-For classical normal linear regression model, residuals are derived from the fitted values $\hat{\boldsymbol{y}}$ and observed values $\boldsymbol{y}$.
-However, if the data generating process is known and the model is correctly specified, by the Frisch-Waugh-Lowell theorem (ref here), residuals $\boldsymbol{e}$ can be written as a linear transformation of the error $\boldsymbol{\varepsilon}$:
+For the classical normal linear regression model, residuals are derived from the fitted values $\hat{\boldsymbol{y}}$ and observed values $\boldsymbol{y}$.
+Suppose the data generating process is known and the model is correctly specified, by the Frisch-Waugh-Lowell theorem [@frisch1933partial], residuals $\boldsymbol{e}$ can also be written as a linear transformation of the error $\boldsymbol{\varepsilon}$ formulated as $\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon}$, where $\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the residual operator, $\boldsymbol{X}$ is the design matrix, $\boldsymbol{I}_n$ is a $n$ by $n$ identity matrix, and $n$ is the number of observations.
 
-\begin{equation}
-\label{eq:residual-1}
-\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon},
-\end{equation}
-
-\noindent where $\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the residual operator, $\boldsymbol{X}$ is the design matrix, $\boldsymbol{I}_n$ is a $n$ by $n$ identity matrix, and $n$ is the number of observations.
-
-One of the assumptions of the classical normal linear regression model is the error $\boldsymbol{\varepsilon}$ follows a multivariate normal distribution with zero mean and constant variance, i.e., $N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$.
-Using equation \ref{eq:residual-1}, it can be known that residuals $\boldsymbol{e}$ also follow a certain probability distribution, which will be denoted as $Q$.
+One of the assumptions of the classical normal linear regression model is the error $\boldsymbol{\varepsilon}$ follows a multivariate normal distribution with zero mean and constant variance, i.e., $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$.
+It can be known that residuals $\boldsymbol{e}$ also follow a certain probability distribution transformed from the multivariate normal distribution, which will be denoted as $Q$.
 This reference distribution $Q$ summarises what good residuals should follow given the predictors are known and fixed.
 
-Probability distribution $Q$ is actually a degenerate multivariate distribution due to the fact that $rank(\boldsymbol{R}) = n - 1 < n$.
-This means the $n$th residual value will be known given the $n - 1$ residuals.
-To capture the exact characteristics of $Q$, such as moments, we can simulate a large numbers of $\boldsymbol{\varepsilon}$ and transform it to $\boldsymbol{e}$.
-For simplicity, we replaced $\boldsymbol{R}$ with a diagonal matrix $diag(\boldsymbol{R})$.
+When fitting a linear regression model, the solver will force $\sum_{i=1}^{n} e_i = 0$, making any residual value to be a linear combination of the remaining $n - 1$ residuals.
+This effectively means $rank(\boldsymbol{R}) = n - 1 < n$ and $Q$ becomes a degenerate multivariate distribution.
+To capture the characteristics of $Q$, such as moments, we can simulate a large numbers of $\boldsymbol{\varepsilon}$ and transform it to $\boldsymbol{e}$ to get the empirical estimates.
+For simplicity, we replaced $\boldsymbol{R}$ with a full-rank diagonal matrix $diag(\boldsymbol{R})$, where $diag(.)$ set the non-diagonal entries of a matrix to zeros.
 The resulting distribution for $\boldsymbol{Q}$ is $N(\boldsymbol{0}_n, diag(\boldsymbol{R}\sigma^2))$.
 
-If the model is misspecified, then the actual distribution of residuals denoted as $P$, will be different to $Q$.
-For example, if the model formula is correct, but the error $\boldsymbol{\varepsilon}$ follows a multivariate lognormal distribution, $P$ will not be the same as $Q$ and we would expect a skewed empirical distribution for residuals.
-And if there is an omitted variable problem, then the residual operator obtained from the fitted model will not be the same as $\boldsymbol{R}$, and again, $P$ will be different to $Q$.
+Distribution $Q$ is derived from the correctly specified model.
+However, if the model is misspecified, then the actual distribution of residuals denoted as $P$, will be different to $Q$.
+For example, if the data generating process contains variables correlated with any column of $\boldsymbol{X}$ but not included in $\boldsymbol{X}$, causing an omitted variable problem, $P$ will be different to $Q$ because the residual operator obtained from the fitted model will not be the same as $\boldsymbol{R}$.
+Besides, if the $\boldsymbol{\varepsilon}$ follows a non-normal distribution such as a multivariate lognormal distribution, the empirical residual distribution will usually be skewed and has a long tail.
 
@@
 Define a proper distance between distributions is usually easier than define a proper distance between data plots.
-Given the actual residual distribution $Q$ and the reference residual distribution $P$, we used a distance metric based on Kullback-Leibler divergence (ref here) to quantify the difference between two distributions
+Given the actual residual distribution $Q$ and the reference residual distribution $P$, we used a distance metric based on Kullback-Leibler divergence [@kullback1951information] to quantify the difference between two distributions
 
 \begin{align}
+\label{eq:kl-0}
 D &= log\left(1 + KL\right), \\
@@
 
-\noindent where $p(\boldsymbol{e})$ is the probability density function for distribution $P$, and $q(\boldsymbol{e})$ is the probability density function for distribution $Q$.
+\noindent where $p(.)$ is the probability density function for distribution $P$, and $q(.)$ is the probability density function for distribution $Q$.
 
@@
 It was mainly designed for measuring the effect size of non-linearity and heteroskedasticity in a residual plot.
-For a linear regression model that has omitted higher-order predictors $\boldsymbol{Z}$ along with coefficients $\boldsymbol{\beta}_z$, and an error distribution $N(\boldsymbol{0}_n, \boldsymbol{V})$, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))$.
-Note that the variance-covariance matrix is replaced with its diagonal analogy to ensure it is a full rank matrix.
+@li2023plot showed that, for a classical normal linear regression model that omits a necessary higher-order predictors $\boldsymbol{Z}$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))$.
+Note that the variance-covariance matrix is replaced with the diagonal matrix to ensure it is a full-rank matrix.
 
@@
 For non-normal error $\boldsymbol{\varepsilon}$, the actual residual distribution $P$ is unlikely to be a multivariate normal distribution.
-Thus, equation \ref{eq:kl-2} will not be applicable to linear regression model with non-normality issue.
+Thus, equation \ref{eq:kl-2} given in @li2023plot will not be applicable to models with non-normality violations.
 
-To evaluate the Kullback-Leibler divergence of non-normal $P$ from $Q$, we need to solve equation \ref{eq:kl-1}.
+To evaluate the Kullback-Leibler divergence of non-normal $P$ from $Q$, the fallback is to solve equation \ref{eq:kl-1} numerically.
 However, since $\boldsymbol{e}$ is a linear transformation of non-normal random variables, it is very common that the general form of $P$ is unknown, meaning that we can not easily compute $p(\boldsymbol{e})$ using a well-known probability density function.
-
 Additionally, even if $p(\boldsymbol{e})$ can be calculated for any $\boldsymbol{e} \in \mathbb{R}^n$, it will be very difficult to do numerical integration over the $n$ dimensional space, because $n$ could be potentially very large.
@@
 First, we no longer need to integrate over $n$ random variables.
-The result is now the sum of the Kullback-Leibler divergence evaluated for each individual residual thanks for the independence assumption.
+The result of equation \ref{eq:kl-1} is now the sum of the Kullback-Leibler divergence evaluated for each individual residual thanks for the independence assumption.
 Second, it is not required to know the joint probability density $p(\boldsymbol{e})$ any more.
-Instead, the evaluation of Kullback-Leibler divergence for a single residual relies on the knowledge of $p(e_i)$, where $e_i$ is the $i$th residual for $i = 1, ..., n$.
+Instead, the evaluation of Kullback-Leibler divergence for an individual residual relies on the knowledge of the marginal density $p_i(e_i)$, where $e_i$ is the $i$-th residual for $i = 1, ..., n$.
 This is much easier to estimate through simulation.
 
-The algorithm to compute equation \ref{eq:kl-1} starts from simulating $m$ sets of $\boldsymbol{\varepsilon}$ according to the error distribution.
+The algorithm for computing equation \ref{eq:kl-1} starts from simulating $m$ sets of $\boldsymbol{\varepsilon}$ according to the error distribution.
 The simulated errors are stored in a matrix $\boldsymbol{A}$ with $n$ rows and $m$ columns.
@@
 Then, we can get $m$ sets of $\boldsymbol{e}$ stored in the matrix $\boldsymbol{B}$ by applying the residual rotation operator $\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}$.
-Furthermore, kernel density estimation (ref here) is applied on each row of $B$ to estimate $p(e_i)$ for $i = 1, ..., n$.
+Furthermore, kernel density estimation (KDE) with Gaussian kernel and optimal bandwidth selected by the Sliverman's rule of thumb [@silverman2018density] is applied on each row of $B$ to estimate $p_i(e_i)$ for $i = 1, ..., n$.
+The KDE computation is done by the `density` function in R.
 
-Since the Kullback-Leibler divergence can be viewed as the expectation of the log-likelihood ratio between distribution $P$ and distribution $Q$ evaluated on distribution $P$, we can reuse the simulated residuals in matrix $\boldsymbol{B}$ to compute the expectation.
-Therefore, the Kullback-Leibler divergence can be formulated as
+Since the Kullback-Leibler divergence can be viewed as the expectation of the log-likelihood ratio between distribution $P$ and distribution $Q$ evaluated on distribution $P$, we can reuse the simulated residuals in matrix $\boldsymbol{B}$ to estimate the expectation by the sample mean.
+With the independence assumption, for non-normal $P$, $KL$ can be estimated by
 
@@
 \label{eq:kl-3}
-KL &= \sum_{i = 1}^{n} KL_i, \\
-KL_i &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{p(B_{ij})}{q(B_{ij})},
+\hat{KL} &= \sum_{i = 1}^{n} \hat{KL_i}, \\
+\hat{KL_i} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
 \end{align}
 
-\noindent where $KL_i$ is the Kullback-Leibler divergence for an individual residual $e_i$, $B_{ij}$ is the $i$th row and $j$th column entry of the matrix $B$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\hat{\sigma^2} = var(flatten(B))$, and $flatten(.)$ unfolds a matrix into a vector.
+\noindent where $\hat{KL_i}$ is the estimator of the Kullback-Leibler divergence for an individual residual $e_i$, $B_{ij}$ is the $i$-th row and $j$-th column entry of the matrix $B$, $\hat{p_i}(.)$ is the kernel density estimator of $p_i(.)$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\hat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)$, and $vec(.)$ is the vectorization operator which turns a $n \times m$ matrix into a $nm \times 1$ column vector by stacking the columns of the matrix on top of each other.
 
-### Approximation of the distance metric with a residual plot
+### Approximation of the distance metric
 
-In the previous sections, we have defined a distance metric for quantifying the difference between the actual residual distribution and an ideal reference distribution.
-This distance metric can only be computed when the data generating process is known.
+In the previous sections, we have defined a distance metric given in equation \ref{eq:kl-0} for quantifying the difference between the actual residual distribution $P$ and an ideal reference distribution $Q$.
+You may have noticed that this distance metric can only be computed when the data generating process is known.
 In reality, we often have no knowledge about the data generating process, otherwise we do not need to fit a linear regression model in the first place.
@@
 What we proposed in this paper is a method to approximate this distance with a residual plot.
-This is done by training computer vision models with a large number of pairs of residual plots and distance.
-With the approximated distance, we will be able to know how different the underlying distribution of the residuals is from a good residual distribution.
-This is meaningful way to check if a residual plot is a good residual plot, and to know the strength of the visual signal indicating model violations.
+Let $D$ be the result of equation \ref{eq:kl-0} indicating the extent of the model violations, and our estimator $\hat{D}$ is formulated as
 
-The approximated distance is not expected to be the same as the original distance.
-This is largely because information embedded in a single residual plot is limited and it may not be able to summarise the characteristics of the residual distribution.
+\begin{equation}
+\label{eq:d-approx}
+\hat{D} = cv(plot_{h\times w}(\boldsymbol{e}, \hat{\boldsymbol{y}})),
+\end{equation}
+
+\noindent where $\boldsymbol{e}$ is a vector of residuals obtained from the fitted model rather than a vector of random variables, $\hat{\boldsymbol{y}}$ is the fitted values also obtained from the fitted model, $plot(.)_{h \times w}$ is a plotting function that generates a residuals vs fitted values plot with fixed aesthetic and then saves it as an image with $h \times w$ pixels and three colour channels, and $cv(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance.
+
+With the approximated distance $\hat{D}$, we will be able to know how different the underlying distribution of the residuals is from a good residual distribution.
+This is meaningful way to check if a residual plot is a good residual plot, and to know the strength of the visual signal embedded in the residual plot.
+
+The approximated distance $\hat{D}$ is not expected to be the same as the original distance $D$.
+This is largely because information contained in a single residual plot is limited and it may not be able to summarise all the characteristics of the residual distribution.
 For a given residual distribution $P$, we can generate many different residual plots.
 Some of them share similar visual patterns, but some of them could be visually very different from the rest, especially for models with small $n$.
-This suggests the approximated distance will vary depends on whether the observed residual plot is representative or not.
+This suggests the error of the approximation will vary depends on whether the observed residual plot is representative or not.
 
-## Generation of simulated training data
+## Statistical testing based on the approximated distance
 
-The computer vision models we developed in this study were trained with synthetic data.
-This means the training data and the test data were simulated from a known data generating process under different parameter settings.
+### Null distribution of the approximated distance
 
-We have incorporated three types of residual departures of linear regression model in this study, including non-linearity, heteroskedasticity and non-normality.
+### Estimation of quantiles of the null distribution
 
-<!-- The training data and test data was simulated using the following data   -->
+### Bootstrapping the approximated distance
 
-<!-- The data was simulated using the following data generating process -->
+## Generation of training data
 
-\begin{align*}
+### Data generating process
+
+While observational data is frequently employed in training models for real-world applications, the data generating process of observational data often remains unknown, making computation for our target variable $D$ unattainable.
+Consequently, the computer vision models developed in this study were trained using synthetic data.
+This approach provides us with precise label annotations.
+Additionally, it ensures a large and diverse training dataset, as we have control over the data generating process, and the simulation of the training data is relatively cost-effective.
+
+<!-- The computer vision model is trained with $n_{train}$ pairs of residual plots and distance $D$, and the trainable parameters $\boldsymbol{\theta}$ are optimized using the mean square error loss function written as -->
+
+<!-- $$\hat{\boldsymbol{\theta}} = \underset{\boldsymbol{\theta}}{argmin}\frac{1}{n_{train}}\sum_{i=1}^{n_{train}}(D_i - vss_{\boldsymbol{\theta}}(plot_i))^2,$$ -->
+<!-- \noindent where $plot_{i}$ is the i$th$ residual plot, and $D_i$ is the computed distance for the underlying residual distribution of the $i$th residual plot.
+The detailed training process is described in Section ....
+-->
+
+We have incorporated three types of residual departures of linear regression model in the training data, including non-linearity, heteroskedasticity and non-normality.
+All three departures can be summarised by the data generating process formulated as
+
+\begin{align}
+\label{eq:data-sim}
 \boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 + \beta_2(\boldsymbol{z} + \beta_1\boldsymbol{w}) + \boldsymbol{k} \odot \boldsymbol{\varepsilon}, \\
@@
 \boldsymbol{w} &= He_j(g(\boldsymbol{x}_2, 2)), \\
-\boldsymbol{k} &= \sqrt{\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^2}, \\
-\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{I}_n),
-\end{align*}
+\boldsymbol{k} &= \sqrt{\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^2},
+\end{align}
 
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{z}$, $\boldsymbol{w}$, $\boldsymbol{k}$ and $\boldsymbol{\varepsilon}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $He_j(.)$ is the $j$th-order probabilist's Hermite polynomials (ref), the $\sqrt{(.)}$ and $(.)^2$ operator are element-wise operators, $\odot$ is the Hadamard product, and $g(., k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{z}$, $\boldsymbol{w}$, $\boldsymbol{k}$ and $\boldsymbol{\varepsilon}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$ are two independent predictors, $He_j(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau], the $\sqrt{(.)}$ and $(.)^2$ operators are element-wise operators, $\odot$ is the Hadamard product, and $g(., k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
 
@@
 
-## Different configurations of the model formula
+The residuals and fitted values of the fitted model is obtained by regressing $\boldsymbol{y}$ on $\boldsymbol{x}_1$.
+This data generating process is adopted from @li2023plot where it was used to simulate residual plots with non-linearity and heteroskedasticity visual patterns for human subject experiments.
 
-discuss different types of inputs and output choices, and why we choose the current design
+In equation \ref{eq:data-sim}, predictor $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$ are independent of each other, so excluding $\boldsymbol{x}_2$ from the regression formula will not cause any problem.
+However, $\boldsymbol{z}$ and $\boldsymbol{w}$ are higher-order terms of $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$.
+If $\beta_2 \neq 0$, the regression model will suffer from non-linearity issues.
+Parameter $j$ is a shape parameter controlling the number of tuning points of the non-linearity pattern.
+Generally, greater values of $j$ will result in more tuning points.
+
+Additionally, Scaling factor $\boldsymbol{k}$ directly affects the error distribution and it is correlated with $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$.
+If $b \neq 0$ and $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$, the constant variance assumption will be violated.
+Parameter $a$ is a shape parameter controlling the location of the smallest variance in a residual plot.
+
+Non-normality violations are introduced by defining a non-normal distribution for $\boldsymbol{\varepsilon}$. 
+- distribution of x1 and x2
+- summary table of all the factor values
+- example plots for the violations
+
+### Computation of scagnostics
+
+We have mentioned in Section \ref{introduction} that scagnostics summarise some useful characteristics of the scatter plot.
+Although the computer vision model will learn the feature extraction function during the training process, it will be beneficial to utilize scagnostics to help computer vision model making more accurate predictions as these measure could provide additional useful information.
+
+For each generated residual plot, four scagnostics, namely, "Monotonic", "Sparse", "Splines" and "Striped" are computed using the `cassowaryr` R package (ref here).
+The computed measures along with the number of observations of the fitted model are provided as the second input of the computer vision model.
+
+Other scagnostics are also informative, but unfortunately, they are unavailable due to a fatal bug in the compiled C program of the `interp` R package (ref here) that will crash the process in a unpredictable manner.
+For reproducibility, we do not include these scagnostics in the training data.
+
+### Crafting a balanced training set
 
@@
 
+- optimization (MSE loss function, with equation)
+- a diagram
+- explain the existence of removable layers 
+- explain three different resolutions
+
 ## Training process and hyperparameter tuning
 
+- software environment (tf and keras)
+- tuner (keras_tuner)
+- 80% validation data, 20% training data
+- 100 trials (Bayesian optimizer)
+- monitor RMSE
+- early stopping (50 epochs)
+- learning rate reduction (10 epochs, 0.5 factor)
+- max 2000 epochs (no one reach)
+- restore best epoch of the best trial
+
 ## Model evaluation methods
+
+- RMSE for the test set
+- R^2
+- Mean bias deviation to understand the overall bias
+- quantile loss to understand how well the model captures the entire distribution of D
+- Percentage of predictions within a tolerance interval (like 0.1)
 
@@
 
+There are other kinds of residual departures like autocorrelation that are not considered in this study.
+The primary goal of this paper is to establish a new way of evaluating residual plot and conducting visual test with computer vision models.
+Building computer vision models for other model violations and other types of diagnostic plots could be future directions of this field.
+
 # Conclusion

### 2024-01-22 paper/paper.tex
update paper methodology and toc

@@
 
+\newpage
+\tableofcontents
+\newpage
+
 \hypertarget{introduction}{%
@@
 standardized results in a controlled environment. The large-scale
-evaluation of lineups is impractical without the use of technology and
+evaluation of lineups ihs impractical without the use of technology and
 machines.
@@
 nine computable measures applied to planar proximity graphs. These
-measures, including ``Outlying'', ``Skinny'', ``Stringy'', ``Straight'',
-``Monotonic'', ``Skewed'', ``Clumpy'', and ``Striated'' aimed to
-characterize outliers, shape, density, trend, and coherence of the data.
-While this approach has been inspiring, there is a recognition
-\citep{buja2009statistical} that it may not capture all the necessary
-visual features that differentiate actual residual plots from null
-plots. A more promising alternative entails enabling computers to learn
-the function for extracting visual features from residual plots.
-Essentially, this means empowering computers to discern the crucial
-visual features for residual diagnostics and determining the method to
-extract them. Modern computer vision models are well-suited for
-addressing this challenge.
+measures, including, but not limited to, ``Outlying'', ``Skinny'',
+``Stringy'', ``Straight'', ``Monotonic'', ``Skewed'', ``Clumpy'', and
+``Striated'' aimed to characterize outliers, shape, density, trend,
+coherence and other characteristics of the data. While this approach has
+been inspiring, there is a recognition \citep{buja2009statistical} that
+it may not capture all the necessary visual features that differentiate
+actual residual plots from null plots. A more promising alternative
+entails enabling computers to learn the function for extracting visual
+features from residual plots. Essentially, this means empowering
+computers to discern the crucial visual features for residual
+diagnostics and determining the method to extract them. Modern computer
+vision models are well-suited for addressing this challenge.
 
@@
 
+\hypertarget{different-possible-configurations-of-the-model-formula}{%
+\subsection{Different possible configurations of the model
+formula}\label{different-possible-configurations-of-the-model-formula}}
+
+In addressing a problem, there exist multiple avenues for solutions.
+Various configurations of the computer vision model formula can
+effectively read residual plots, and these configurations can be
+categorized based on two key components of the model formula: the input
+and the output format.
+
+\hypertarget{different-input-formats}{%
+\subsubsection{Different input formats}\label{different-input-formats}}
+
+In computer vision models, the input plays a role similar to predictors
+in a linear regression model. The quality and relevance of the input
+data greatly influence the model's capacity to generate insightful and
+meaningful results. A straightforward approach involves feeding the
+model a vector of residuals along with a vector of fitted values,
+essentially providing all the necessary details for creating a residuals
+vs fitted values plot. However, a drawback of this method is the dynamic
+input size, as different fitted regression models may have varying
+numbers of observations. In modern computer vision models created with
+mainstream software like TensorFlow (ref here), the input is typically a
+fixed-shape tensor. One solution is to pad the input vector with leading
+or trailing zeros if the input tensor expects a longer vector. However,
+this approach may fail if the input vector surpasses the designed
+length. Another strategy is to summarize the residuals and fitted values
+separately using histograms and utilize the counts as the input. By
+controlling the number of bins in the histograms, it becomes possible to
+provide a fixed-length input vector.
+
+The primary advantage of using an image as input, as opposed to a vector
+input format, relates to the development of sophisticated image
+processing architectures over the years (see ref here). These
+architectures effectively capture and summarize spatial information from
+nearby pixels, which is less straightforward with vector input. The only
+considerations are the image resolution and the aesthetics of the
+residual plot. In general, higher resolution provides more information
+to the model but comes with the trade-off of increased complexity and
+greater difficulty in training. As for the aesthetics of the residual
+plot, a practical solution is to consistently present residual plots in
+the same style to the model. This implies that the model can not accept
+arbitrary images as input but requires the use of the same preprocessing
+software to convert provided residuals and fitted values into a
+standardized style for the residual plot.
+
+Other possible input formats include, but are not limited to, a pair of
+residual plots, a triplet, and a lineup. Existing computer vision models
+designed for image comparison can assess whether a pair of images are
+similar or dissimilar (as seen in LeCun's 2005 paper). Applied to our
+specific problem, we can define null plots of a fitted model to be
+similar to each other, while considering actual residual plots to be
+distinct from null plots of any fitted model. A triplet constitutes a
+set of three images, denoted as \(image_1\), \(image_2\) and
+\(image_3\). In computer vision models, triplets are often used to
+predict whether \(image_2\) or \(image_3\) is more similar to
+\(image_1\), proving particularly useful for establishing rankings
+between samples. In the context of our problem, we can apply the same
+criteria to define similarity between images. However, it's important to
+note that these two approaches usually require additional considerations
+regarding the loss function and, at times, non-standard training
+processes due to shared weights between different convolutional blocks.
+
+Presenting a lineup to a model and asking it to predict which residual
+plot is the most different one aligns closely with the lineup protocol.
+However, if the lineup consists of around 20 plots, similar to previous
+human subject experiments (reference here), the resolution of the input
+image could become quite large, posing challenges in training the model.
+In fact, we experimented with this approach in a pilot study, and the
+model's performance was suboptimal.
+
+We did not explore all the mentioned input formats due to the
+considerable costs associated with data preparation and model training.
+Considering the implementation cost and the interpretability of the
+model, we settled on the single residual plot input format.
+
+\hypertarget{different-output-formats}{%
+\subsubsection{Different output
+formats}\label{different-output-formats}}
+
+Given that the input is a single residual plot represented as a
+fixed-resolution image, the computer vision model's output can take one
+of two formats: binary or numeric. This choice determines whether the
+model is a classification model or a regression model. The binary
+outcome might indicates whether the input image is a null plot or not,
+or whether the input image would be rejected in a visual test conducted
+by humans. The latter option requires data from prior human subject
+experiments, posing challenges in controlling the quality and quantity
+of data due to variations in experimental settings across different
+studies. Additionally, some visual inference experiments are unrelated
+to linear regression models and residual plots, resulting in a limited
+amount of available training data.
+
+Alternatively, the output could be a meaningful numeric measure, such as
+the average distance to good residual plots. This approach necessitates
+defining a distance measure between residual plots, which may be a
+non-trivial task. However, once a meaningful distance measure is
+established, the computer vision model's prediction becomes an
+interpretable value with tangible significance. Studies have
+demonstrated that defining a proper distance between images can enhance
+the matching accuracy in image search compared to a binary outcome model
+(ref here).
+
+The following section will propose the distance measure used for
+training the computer vision model.
+
 \hypertarget{distance-from-the-good-residual-plots}{%
@@
 established image similarity metrics like the Structural Similarity
-Index Measure (SSIM) (ref here). The challenge lies in the fact that
+Index Measure \citep{wang2004image}. The challenge lies in the fact that
 metrics tailored for image comparison may not be suitable for evaluating
 data plots, where only essential plot elements require assessment
-\citep{chowdhury2018measuring}.
+\citep{chowdhury2018measuring}. Furthermore, scagnostics mentioned in
+Section \ref{introduction} could be used to construct distance metrics
+for data plots comparison, but the functional form still needs to be
+carefully refined to accurately reflect the extent of the violations.
 
@@
 residual plot, or in other words, how different a given fitted model is
-from a correctly specified model. For classical normal linear regression
-model, residuals are derived from the fitted values
-\(\hat{\boldsymbol{y}}\) and observed values \(\boldsymbol{y}\).
-However, if the data generating process is known and the model is
-correctly specified, by the Frisch-Waugh-Lowell theorem (ref here),
-residuals \(\boldsymbol{e}\) can be written as a linear transformation
-of the error \(\boldsymbol{\varepsilon}\):
-
-\begin{equation}
-\label{eq:residual-1}
-\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon},
-\end{equation}
-
-\noindent where
+from a correctly specified model. For the classical normal linear
+regression model, residuals are derived from the fitted values
+\(\hat{\boldsymbol{y}}\) and observed values \(\boldsymbol{y}\). Suppose
+the data generating process is known and the model is correctly
+specified, by the Frisch-Waugh-Lowell theorem \citep{frisch1933partial},
+residuals \(\boldsymbol{e}\) can also be written as a linear
+transformation of the error \(\boldsymbol{\varepsilon}\) formulated as
+\(\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon}\), where
 \(\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\)
@@
 distribution with zero mean and constant variance, i.e.,
-\(N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)\). Using equation
-\ref{eq:residual-1}, it can be known that residuals \(\boldsymbol{e}\)
-also follow a certain probability distribution, which will be denoted as
-\(Q\). This reference distribution \(Q\) summarises what good residuals
-should follow given the predictors are known and fixed.
+\(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)\).
+It can be known that residuals \(\boldsymbol{e}\) also follow a certain
+probability distribution transformed from the multivariate normal
+distribution, which will be denoted as \(Q\). This reference
+distribution \(Q\) summarises what good residuals should follow given
+the predictors are known and fixed.
 
-Probability distribution \(Q\) is actually a degenerate multivariate
-distribution due to the fact that \(rank(\boldsymbol{R}) = n - 1 < n\).
-This means the \(n\)th residual value will be known given the \(n - 1\)
-residuals. To capture the exact characteristics of \(Q\), such as
-moments, we can simulate a large numbers of \(\boldsymbol{\varepsilon}\)
-and transform it to \(\boldsymbol{e}\). For simplicity, we replaced
-\(\boldsymbol{R}\) with a diagonal matrix \(diag(\boldsymbol{R})\). The
-resulting distribution for \(\boldsymbol{Q}\) is
-\(N(\boldsymbol{0}_n, diag(\boldsymbol{R}\sigma^2))\).
+When fitting a linear regression model, the solver will force
+\(\sum_{i=1}^{n} e_i = 0\), making any residual value to be a linear
+combination of the remaining \(n - 1\) residuals. This effectively means
+\(rank(\boldsymbol{R}) = n - 1 < n\) and \(Q\) becomes a degenerate
+multivariate distribution. To capture the characteristics of \(Q\), such
+as moments, we can simulate a large numbers of
+\(\boldsymbol{\varepsilon}\) and transform it to \(\boldsymbol{e}\) to
+get the empirical estimates. For simplicity, we replaced
+\(\boldsymbol{R}\) with a full-rank diagonal matrix
+\(diag(\boldsymbol{R})\), where \(diag(.)\) set the non-diagonal entries
+of a matrix to zeros. The resulting distribution for \(\boldsymbol{Q}\)
+is \(N(\boldsymbol{0}_n, diag(\boldsymbol{R}\sigma^2))\).
 
-If the model is misspecified, then the actual distribution of residuals
-denoted as \(P\), will be different to \(Q\). For example, if the model
-formula is correct, but the error \(\boldsymbol{\varepsilon}\) follows a
-multivariate lognormal distribution, \(P\) will not be the same as \(Q\)
-and we would expect a skewed empirical distribution for residuals. And
-if there is an omitted variable problem, then the residual operator
-obtained from the fitted model will not be the same as
-\(\boldsymbol{R}\), and again, \(P\) will be different to \(Q\).
+Distribution \(Q\) is derived from the correctly specified model.
+However, if the model is misspecified, then the actual distribution of
+residuals denoted as \(P\), will be different to \(Q\). For example, if
+the data generating process contains variables correlated with any
+column of \(\boldsymbol{X}\) but not included in \(\boldsymbol{X}\),
+causing an omitted variable problem, \(P\) will be different to \(Q\)
+because the residual operator obtained from the fitted model will not be
+the same as \(\boldsymbol{R}\). Besides, if the
+\(\boldsymbol{\varepsilon}\) follows a non-normal distribution such as a
+multivariate lognormal distribution, the empirical residual distribution
+will usually be skewed and has a long tail.
 
@@
 distribution \(Q\) and the reference residual distribution \(P\), we
-used a distance metric based on Kullback-Leibler divergence (ref here)
-to quantify the difference between two distributions
+used a distance metric based on Kullback-Leibler divergence
+\citep{kullback1951information} to quantify the difference between two
+distributions
 
 \begin{align}
+\label{eq:kl-0}
 D &= log\left(1 + KL\right), \\
@@
 
-\noindent where \(p(\boldsymbol{e})\) is the probability density
-function for distribution \(P\), and \(q(\boldsymbol{e})\) is the
-probability density function for distribution \(Q\).
+\noindent where \(p(.)\) is the probability density function for
+distribution \(P\), and \(q(.)\) is the probability density function for
+distribution \(Q\).
 
@@
 mainly designed for measuring the effect size of non-linearity and
-heteroskedasticity in a residual plot. For a linear regression model
-that has omitted higher-order predictors \(\boldsymbol{Z}\) along with
-coefficients \(\boldsymbol{\beta}_z\), and an error distribution
-\(N(\boldsymbol{0}_n, \boldsymbol{V})\), \(Q\) can be represented as
+heteroskedasticity in a residual plot. \citet{li2023plot} showed that,
+for a classical normal linear regression model that omits a necessary
+higher-order predictors \(\boldsymbol{Z}\), and incorrectly assumes
+\(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)\)
+while in fact
+\(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})\),
+\(Q\) can be represented as
 \(N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))\).
-Note that the variance-covariance matrix is replaced with its diagonal
-analogy to ensure it is a full rank matrix.
+Note that the variance-covariance matrix is replaced with the diagonal
+matrix to ensure it is a full-rank matrix.
 
@@
 distribution \(P\) is unlikely to be a multivariate normal distribution.
-Thus, equation \ref{eq:kl-2} will not be applicable to linear regression
-model with non-normality issue.
+Thus, equation \ref{eq:kl-2} given in \citet{li2023plot} will not be
+applicable to models with non-normality violations.
 
 To evaluate the Kullback-Leibler divergence of non-normal \(P\) from
-\(Q\), we need to solve equation \ref{eq:kl-1}. However, since
-\(\boldsymbol{e}\) is a linear transformation of non-normal random
-variables, it is very common that the general form of \(P\) is unknown,
-meaning that we can not easily compute \(p(\boldsymbol{e})\) using a
-well-known probability density function.
-
+\(Q\), the fallback is to solve equation \ref{eq:kl-1} numerically.
+However, since \(\boldsymbol{e}\) is a linear transformation of
+non-normal random variables, it is very common that the general form of
+\(P\) is unknown, meaning that we can not easily compute
+\(p(\boldsymbol{e})\) using a well-known probability density function.
 Additionally, even if \(p(\boldsymbol{e})\) can be calculated for any
@@
 above. First, we no longer need to integrate over \(n\) random
-variables. The result is now the sum of the Kullback-Leibler divergence
-evaluated for each individual residual thanks for the independence
-assumption. Second, it is not required to know the joint probability
-density \(p(\boldsymbol{e})\) any more. Instead, the evaluation of
-Kullback-Leibler divergence for a single residual relies on the
-knowledge of \(p(e_i)\), where \(e_i\) is the \(i\)th residual for
-\(i = 1, ..., n\). This is much easier to estimate through simulation.
+variables. The result of equation \ref{eq:kl-1} is now the sum of the
+Kullback-Leibler divergence evaluated for each individual residual
+thanks for the independence assumption. Second, it is not required to
+know the joint probability density \(p(\boldsymbol{e})\) any more.
+Instead, the evaluation of Kullback-Leibler divergence for an individual
+residual relies on the knowledge of the marginal density \(p_i(e_i)\),
+where \(e_i\) is the \(i\)-th residual for \(i = 1, ..., n\). This is
+much easier to estimate through simulation.
 
-The algorithm to compute equation \ref{eq:kl-1} starts from simulating
-\(m\) sets of \(\boldsymbol{\varepsilon}\) according to the error
-distribution. The simulated errors are stored in a matrix
+The algorithm for computing equation \ref{eq:kl-1} starts from
+simulating \(m\) sets of \(\boldsymbol{\varepsilon}\) according to the
+error distribution. The simulated errors are stored in a matrix
 \(\boldsymbol{A}\) with \(n\) rows and \(m\) columns. So each column of
@@
 \(\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}\). Furthermore, kernel
-density estimation (ref here) is applied on each row of \(B\) to
-estimate \(p(e_i)\) for \(i = 1, ..., n\).
+density estimation (KDE) with Gaussian kernel and optimal bandwidth
+selected by the Sliverman's rule of thumb \citep{silverman2018density}
+is applied on each row of \(B\) to estimate \(p_i(e_i)\) for
+\(i = 1, ..., n\). The KDE computation is done by the \texttt{density}
+function in R.
 
@@
 \(Q\) evaluated on distribution \(P\), we can reuse the simulated
-residuals in matrix \(\boldsymbol{B}\) to compute the expectation.
-Therefore, the Kullback-Leibler divergence can be formulated as
+residuals in matrix \(\boldsymbol{B}\) to estimate the expectation by
+the sample mean. With the independence assumption, for non-normal \(P\),
+\(KL\) can be estimated by
 
@@
 \label{eq:kl-3}
-KL &= \sum_{i = 1}^{n} KL_i, \\
-KL_i &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{p(B_{ij})}{q(B_{ij})},
+\hat{KL} &= \sum_{i = 1}^{n} \hat{KL_i}, \\
+\hat{KL_i} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
 \end{align}
 
-\noindent where \(KL_i\) is the Kullback-Leibler divergence for an
-individual residual \(e_i\), \(B_{ij}\) is the \(i\)th row and \(j\)th
-column entry of the matrix \(B\), \(q(.)\) is the normal density
-function with mean zero and an assumed variance estimated as
-\(\hat{\sigma^2} = var(flatten(B))\), and \(flatten(.)\) unfolds a
-matrix into a vector.
+\noindent where \(\hat{KL_i}\) is the estimator of the Kullback-Leibler
+divergence for an individual residual \(e_i\), \(B_{ij}\) is the
+\(i\)-th row and \(j\)-th column entry of the matrix \(B\),
+\(\hat{p_i}(.)\) is the kernel density estimator of \(p_i(.)\), \(q(.)\)
+is the normal density function with mean zero and an assumed variance
+estimated as
+\(\hat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)\),
+and \(vec(.)\) is the vectorization operator which turns a
+\(n \times m\) matrix into a \(nm \times 1\) column vector by stacking
+the columns of the matrix on top of each other.
 
-\hypertarget{approximation-of-the-distance-metric-with-a-residual-plot}{%
-\subsubsection{Approximation of the distance metric with a residual
-plot}\label{approximation-of-the-distance-metric-with-a-residual-plot}}
+\hypertarget{approximation-of-the-distance-metric}{%
+\subsubsection{Approximation of the distance
+metric}\label{approximation-of-the-distance-metric}}
 
-In the previous sections, we have defined a distance metric for
-quantifying the difference between the actual residual distribution and
-an ideal reference distribution. This distance metric can only be
-computed when the data generating process is known. In reality, we often
-have no knowledge about the data generating process, otherwise we do not
-need to fit a linear regression model in the first place.
+In the previous sections, we have defined a distance metric given in
+equation \ref{eq:kl-0} for quantifying the difference between the actual
+residual distribution \(P\) and an ideal reference distribution \(Q\).
+You may have noticed that this distance metric can only be computed when
+the data generating process is known. In reality, we often have no
+knowledge about the data generating process, otherwise we do not need to
+fit a linear regression model in the first place.
 
 What we proposed in this paper is a method to approximate this distance
-with a residual plot. This is done by training computer vision models
-with a large number of pairs of residual plots and distance. With the
-approximated distance, we will be able to know how different the
-underlying distribution of the residuals is from a good residual
-distribution. This is meaningful way to check if a residual plot is a
-good residual plot, and to know the strength of the visual signal
-indicating model violations.
+with a residual plot. Let \(D\) be the result of equation \ref{eq:kl-0}
+indicating the extent of the model violations, and our estimator
+\(\hat{D}\) is formulated as
 
-The approximated distance is not expected to be the same as the original
-distance. This is largely because information embedded in a single
-residual plot is limited and it may not be able to summarise the
-characteristics of the residual distribution. For a given residual
-distribution \(P\), we can generate many different residual plots. Some
-of them share similar visual patterns, but some of them could be
-visually very different from the rest, especially for models with small
-\(n\). This suggests the approximated distance will vary depends on
-whether the observed residual plot is representative or not.
+\begin{equation}
+\label{eq:d-approx}
+\hat{D} = cv(plot_{h\times w}(\boldsymbol{e}, \hat{\boldsymbol{y}})),
+\end{equation}
 
-\hypertarget{generation-of-simulated-training-data}{%
-\subsection{Generation of simulated training
-data}\label{generation-of-simulated-training-data}}
+\noindent where \(\boldsymbol{e}\) is a vector of residuals obtained
+from the fitted model rather than a vector of random variables,
+\(\hat{\boldsymbol{y}}\) is the fitted values also obtained from the
+fitted model, \(plot(.)_{h \times w}\) is a plotting function that
+generates a residuals vs fitted values plot with fixed aesthetic and
+then saves it as an image with \(h \times w\) pixels and three colour
+channels, and \(cv(.)\) is a computer vision model which takes an
+\(h \times w\) image as input and predicts the distance.
 
-The computer vision models we developed in this study were trained with
-synthetic data. This means the training data and the test data were
-simulated from a known data generating process under different parameter
-settings.
+With the approximated distance \(\hat{D}\), we will be able to know how
+different the underlying distribution of the residuals is from a good
+residual distribution. This is meaningful way to check if a residual
+plot is a good residual plot, and to know the strength of the visual
+signal embedded in the residual plot.
+
+The approximated distance \(\hat{D}\) is not expected to be the same as
+the original distance \(D\). This is largely because information
+contained in a single residual plot is limited and it may not be able to
+summarise all the characteristics of the residual distribution. For a
+given residual distribution \(P\), we can generate many different
+residual plots. Some of them share similar visual patterns, but some of
+them could be visually very different from the rest, especially for
+models with small \(n\). This suggests the error of the approximation
+will vary depends on whether the observed residual plot is
+representative or not.
+
+\hypertarget{statistical-testing-based-on-the-approximated-distance}{%
+\subsection{Statistical testing based on the approximated
+distance}\label{statistical-testing-based-on-the-approximated-distance}}
+
+\hypertarget{null-distribution-of-the-approximated-distance}{%
+\subsubsection{Null distribution of the approximated
+distance}\label{null-distribution-of-the-approximated-distance}}
+
+\hypertarget{estimation-of-quantiles-of-the-null-distribution}{%
+\subsubsection{Estimation of quantiles of the null
+distribution}\label{estimation-of-quantiles-of-the-null-distribution}}
+
+\hypertarget{bootstrapping-the-approximated-distance}{%
+\subsubsection{Bootstrapping the approximated
+distance}\label{bootstrapping-the-approximated-distance}}
+
+\hypertarget{generation-of-training-data}{%
+\subsection{Generation of training
+data}\label{generation-of-training-data}}
+
+\hypertarget{data-generating-process}{%
+\subsubsection{Data generating process}\label{data-generating-process}}
+
+While observational data is frequently employed in training models for
+real-world applications, the data generating process of observational
+data often remains unknown, making computation for our target variable
+\(D\) unattainable. Consequently, the computer vision models developed
+in this study were trained using synthetic data. This approach provides
+us with precise label annotations. Additionally, it ensures a large and
+diverse training dataset, as we have control over the data generating
+process, and the simulation of the training data is relatively
+cost-effective.
 
 We have incorporated three types of residual departures of linear
-regression model in this study, including non-linearity,
-heteroskedasticity and non-normality.
+regression model in the training data, including non-linearity,
+heteroskedasticity and non-normality. All three departures can be
+summarised by the data generating process formulated as
 
-\begin{align*}
+\begin{align}
+\label{eq:data-sim}
 \boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 + \beta_2(\boldsymbol{z} + \beta_1\boldsymbol{w}) + \boldsymbol{k} \odot \boldsymbol{\varepsilon}, \\
@@
 \boldsymbol{w} &= He_j(g(\boldsymbol{x}_2, 2)), \\
-\boldsymbol{k} &= \sqrt{\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^2}, \\
-\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}_n, \boldsymbol{I}_n),
-\end{align*}
+\boldsymbol{k} &= \sqrt{\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^2},
+\end{align}
 
@@
 \(n\), \(\boldsymbol{1}_n\) is a vector of ones of size \(n\),
-\(He_j(.)\) is the \(j\)th-order probabilist's Hermite polynomials
-(ref), the \(\sqrt{(.)}\) and \((.)^2\) operator are element-wise
-operators, \(\odot\) is the Hadamard product, and \(g(., k)\) is a
-scaling function to enforce the support of the random vector to be
-\([-k, k]^n\) defined as
+\(\boldsymbol{x}_1\) and \(\boldsymbol{x}_2\) are two independent
+predictors, \(He_j(.)\) is the \(j\)th-order probabilist's Hermite
+polynomials \citep{hermite1864nouveau}, the \(\sqrt{(.)}\) and \((.)^2\)
+operators are element-wise operators, \(\odot\) is the Hadamard product,
+and \(g(., k)\) is a scaling function to enforce the support of the
+random vector to be \([-k, k]^n\) defined as
 
@@
 
-\hypertarget{different-configurations-of-the-model-formula}{%
-\subsection{Different configurations of the model
-formula}\label{different-configurations-of-the-model-formula}}
+The residuals and fitted values of the fitted model is obtained by
+regressing \(\boldsymbol{y}\) on \(\boldsymbol{x}_1\). This data
+generating process is adopted from \citet{li2023plot} where it was used
+to simulate residual plots with non-linearity and heteroskedasticity
+visual patterns for human subject experiments.
 
-discuss different types of inputs and output choices, and why we choose
-the current design
+In equation \ref{eq:data-sim}, predictor \(\boldsymbol{x}_1\) and
+\(\boldsymbol{x}_2\) are independent of each other, so excluding
+\(\boldsymbol{x}_2\) from the regression formula will not cause any
+problem. However, \(\boldsymbol{z}\) and \(\boldsymbol{w}\) are
+higher-order terms of \(\boldsymbol{x}_1\) and \(\boldsymbol{x}_2\). If
+\(\beta_2 \neq 0\), the regression model will suffer from non-linearity
+issues. Parameter \(j\) is a shape parameter controlling the number of
+tuning points of the non-linearity pattern. Generally, greater values of
+\(j\) will result in more tuning points.
+
+Additionally, Scaling factor \(\boldsymbol{k}\) directly affects the
+error distribution and it is correlated with \(\boldsymbol{x}_1\) and
+\(\boldsymbol{x}_2\). If \(b \neq 0\) and
+\(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)\),
+the constant variance assumption will be violated. Parameter \(a\) is a
+shape parameter controlling the location of the smallest variance in a
+residual plot.
+
+Non-normality violations are introduced by defining a non-normal
+distribution for \(\boldsymbol{\varepsilon}\). - distribution of x1 and
+x2 - summary table of all the factor values - example plots for the
+violations
+
+\hypertarget{computation-of-scagnostics}{%
+\subsubsection{Computation of
+scagnostics}\label{computation-of-scagnostics}}
+
+We have mentioned in Section \ref{introduction} that scagnostics
+summarise some useful characteristics of the scatter plot. Although the
+computer vision model will learn the feature extraction function during
+the training process, it will be beneficial to utilize scagnostics to
+help computer vision model making more accurate predictions as these
+measure could provide additional useful information.
+
+For each generated residual plot, four scagnostics, namely,
+``Monotonic'', ``Sparse'', ``Splines'' and ``Striped'' are computed
+using the \texttt{cassowaryr} R package (ref here). The computed
+measures along with the number of observations of the fitted model are
+provided as the second input of the computer vision model.
+
+Other scagnostics are also informative, but unfortunately, they are
+unavailable due to a fatal bug in the compiled C program of the
+\texttt{interp} R package (ref here) that will crash the process in a
+unpredictable manner. For reproducibility, we do not include these
+scagnostics in the training data.
+
+\hypertarget{crafting-a-balanced-training-set}{%
+\subsubsection{Crafting a balanced training
+set}\label{crafting-a-balanced-training-set}}
 
@@
 
+\begin{itemize}
+\tightlist
+\item
+  optimization (MSE loss function, with equation)
+\item
+  a diagram
+\item
+  explain the existence of removable layers
+\item
+  explain three different resolutions
+\end{itemize}
+
 \hypertarget{training-process-and-hyperparameter-tuning}{%
@@
 
+\begin{itemize}
+\tightlist
+\item
+  software environment (tf and keras)
+\item
+  tuner (keras\_tuner)
+\item
+  80\% validation data, 20\% training data
+\item
+  100 trials (Bayesian optimizer)
+\item
+  monitor RMSE
+\item
+  early stopping (50 epochs)
+\item
+  learning rate reduction (10 epochs, 0.5 factor)
+\item
+  max 2000 epochs (no one reach)
+\item
+  restore best epoch of the best trial
+\end{itemize}
+
 \hypertarget{model-evaluation-methods}{%
 \subsection{Model evaluation methods}\label{model-evaluation-methods}}
+
+\begin{itemize}
+\tightlist
+\item
+  RMSE for the test set
+\item
+  R\^{}2
+\item
+  Mean bias deviation to understand the overall bias
+\item
+  quantile loss to understand how well the model captures the entire
+  distribution of D
+\item
+  Percentage of predictions within a tolerance interval (like 0.1)
+\end{itemize}
 
@@
 
+There are other kinds of residual departures like autocorrelation that
+are not considered in this study. The primary goal of this paper is to
+establish a new way of evaluating residual plot and conducting visual
+test with computer vision models. Building computer vision models for
+other model violations and other types of diagnostic plots could be
+future directions of this field.
+
 \hypertarget{conclusion}{%

### 2024-01-22 paper/paper.Rmd
put some points down

@@
 So each column of $\boldsymbol{A}$ is a set of realization values of $\boldsymbol{\varepsilon}$.
-Then, we can get $m$ sets of $\boldsymbol{e}$ stored in the matrix $\boldsymbol{B}$ by applying the residual rotation operator $\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}$.
+Then, we can get $m$ sets of $\boldsymbol{e}$ stored in the matrix $\boldsymbol{B}$ by applying the residual operator $\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}$.
 Furthermore, kernel density estimation (KDE) with Gaussian kernel and optimal bandwidth selected by the Sliverman's rule of thumb [@silverman2018density] is applied on each row of $B$ to estimate $p_i(e_i)$ for $i = 1, ..., n$.
@@
 
-\noindent where $\boldsymbol{e}$ is a vector of residuals obtained from the fitted model rather than a vector of random variables, $\hat{\boldsymbol{y}}$ is the fitted values also obtained from the fitted model, $plot(.)_{h \times w}$ is a plotting function that generates a residuals vs fitted values plot with fixed aesthetic and then saves it as an image with $h \times w$ pixels and three colour channels, and $cv(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance.
+\noindent where $\boldsymbol{e}$ is a vector of residuals obtained from the fitted model rather than a vector of random variables, $\hat{\boldsymbol{y}}$ is the fitted values also obtained from the fitted model, $plot(.)_{h \times w}$ is a plotting function that generates a residuals vs fitted values plot with fixed aesthetic and then saves it as an image with $h \times w$ pixels and three colour channels, and $cv(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance in the domain $[0, +\infty)$.
 
@@
 
+Having a computer vision model $cv$ which gives an approximated distance between $0$ and $+\infty$ is not enough to determine if the null hypothesis that the model is correctly specified should be rejected.
+Theoretically, the distance $D$ for a correctly specified model is $0$, because $P$ will be the same as $Q$, and $log(1 + 0) = 0$.
+The computer vision model may not necessary predict $0$ for a null plot.
+Using \ref{fig:false-finding} as an example, it contains a visual pattern which is an indication of heteroskedasticity.
+We would not expect the model to be able to magically tell if the suspicious pattern is caused by the skewed distribution of the fitted values or the existence of heteroskedasticity.
+Assume we only train the model with this single image, while $50$% of the time the label is $0$, and $50$% of the time the label is some constant greater than $0$.
+Then, a perfect model will predict a value between $0$ and that constant to achieve a minimum loss.
+So, it is expected that our computer vision models will also predict value greater than $0$ for some null plots.
+It will make more sense if we interpret the model prediction as the visual signal strength of the residual plot.
+Some null plots could have outliers or strong visual patterns due to randomness, and the model will try to summarise those information into the prediction.
+
+To align with the principle of visual inference, we need to provide a valid null distribution for the test statistic.
+If we treat the approximated distance $\hat{D}$ as a test statistic, then the null distribution of this statistic can be approximated by the empirical distribution of $\hat{D}$ for a large amount of null plots of given a fitted model.
+The procedure involves applying the residual rotation technique [@buja2009statistical] on the fitted model to obtain null residuals.
+The null residuals are then used to make null plots and fed into the model.
+The results are used to construct an empirical distribution.
+
 ### Estimation of quantiles of the null distribution
 
+With the null distribution of the test statistic,
+
+- How many null samples are needed to get stable 95% quantile
+- How do we make decision based on the quantile (reject or not reject)
+- p-value calculation (percentage of null vss > vss)
+
 ### Bootstrapping the approximated distance
+
+- What is the assumption of bootstrapping
+- Any drawback: loss of information
+- For each bootstrapped fitted model, we can get a bootstrapped vss and construct a new 95% quantile for the null distribution, and check if H_0 should be rejected
+- However, it is very expensive to construct a new 95% quantile for each bootstrapped fitted model
+- The 95% quantile constructed using the original fitted model can be considered as an estimate of those new 95% quantiles
+- We borrow information from the original fitted model
+- We can check how many bootstrapped fitted model should be rejected (like the percentage)
+- This gives an overall estimate of how often the assumed regression model are considered to be incorrect if the data can be obtained repetitively from the same data generating process
+- If this percentage is relatively large, it gives the analyst an warning that the regression model is inappropriate to the data
 
@@
 Non-normality violations are introduced by defining a non-normal distribution for $\boldsymbol{\varepsilon}$.
+
 - distribution of x1 and x2
@@
 ### Crafting a balanced training set
+
+In order to train a robust computer vision model, we intentionally control the distribution of the target variable $D$ of the training data.
+Making it uniform between $0$ and $7$.
+This is accomplished by preparing $50$ buckets each only accepts training samples with $D$  between $[7(i - 1)/49, 7i/49)$ for $i < 50$, where $i$ is the index of the $i$-th bucket.
+For the $50$-th bucket, any training samples with $D >= 7$ will be accepted.
+We have prepared 80000 training images, so each bucket can only contain $80000 \div 50 = 1600$ training samples.
+The simulator will repeatedly sample parameter values from the parameter sample, generate residuals and fitted values using the data generating process, compute the distance, and check if the sample can be accepted by the corresponding bucket, until all the buckets are full.
 

### 2024-01-22 paper/paper.tex
put some points down

@@
 \(\boldsymbol{e}\) stored in the matrix \(\boldsymbol{B}\) by applying
-the residual rotation operator
-\(\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}\). Furthermore, kernel
-density estimation (KDE) with Gaussian kernel and optimal bandwidth
-selected by the Sliverman's rule of thumb \citep{silverman2018density}
-is applied on each row of \(B\) to estimate \(p_i(e_i)\) for
-\(i = 1, ..., n\). The KDE computation is done by the \texttt{density}
-function in R.
+the residual operator \(\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}\).
+Furthermore, kernel density estimation (KDE) with Gaussian kernel and
+optimal bandwidth selected by the Sliverman's rule of thumb
+\citep{silverman2018density} is applied on each row of \(B\) to estimate
+\(p_i(e_i)\) for \(i = 1, ..., n\). The KDE computation is done by the
+\texttt{density} function in R.
 
@@
 channels, and \(cv(.)\) is a computer vision model which takes an
-\(h \times w\) image as input and predicts the distance.
+\(h \times w\) image as input and predicts the distance in the domain
+\([0, +\infty)\).
 
@@
 
+Having a computer vision model \(cv\) which gives an approximated
+distance between \(0\) and \(+\infty\) is not enough to determine if the
+null hypothesis that the model is correctly specified should be
+rejected. Theoretically, the distance \(D\) for a correctly specified
+model is \(0\), because \(P\) will be the same as \(Q\), and
+\(log(1 + 0) = 0\). The computer vision model may not necessary predict
+\(0\) for a null plot. Using \ref{fig:false-finding} as an example, it
+contains a visual pattern which is an indication of heteroskedasticity.
+We would not expect the model to be able to magically tell if the
+suspicious pattern is caused by the skewed distribution of the fitted
+values or the existence of heteroskedasticity. Assume we only train the
+model with this single image, while \(50\)\% of the time the label is
+\(0\), and \(50\)\% of the time the label is some constant greater than
+\(0\). Then, a perfect model will predict a value between \(0\) and that
+constant to achieve a minimum loss. So, it is expected that our computer
+vision models will also predict value greater than \(0\) for some null
+plots. It will make more sense if we interpret the model prediction as
+the visual signal strength of the residual plot. Some null plots could
+have outliers or strong visual patterns due to randomness, and the model
+will try to summarise those information into the prediction.
+
+To align with the principle of visual inference, we need to provide a
+valid null distribution for the test statistic. If we treat the
+approximated distance \(\hat{D}\) as a test statistic, then the null
+distribution of this statistic can be approximated by the empirical
+distribution of \(\hat{D}\) for a large amount of null plots of given a
+fitted model. The procedure involves applying the residual rotation
+technique \citep{buja2009statistical} on the fitted model to obtain null
+residuals. The null residuals are then used to make null plots and fed
+into the model. The results are used to construct an empirical
+distribution.
+
 \hypertarget{estimation-of-quantiles-of-the-null-distribution}{%
@@
 
+With the null distribution of the test statistic,
+
+\begin{itemize}
+\tightlist
+\item
+  How many null samples are needed to get stable 95\% quantile
+\item
+  How do we make decision based on the quantile (reject or not reject)
+\item
+  p-value calculation (percentage of null vss \textgreater{} vss)
+\end{itemize}
+
 \hypertarget{bootstrapping-the-approximated-distance}{%
@@
 distance}\label{bootstrapping-the-approximated-distance}}
+
+\begin{itemize}
+\tightlist
+\item
+  What is the assumption of bootstrapping
+\item
+  Any drawback: loss of information
+\item
+  For each bootstrapped fitted model, we can get a bootstrapped vss and
+  construct a new 95\% quantile for the null distribution, and check if
+  H\_0 should be rejected
+\item
+  However, it is very expensive to construct a new 95\% quantile for
+  each bootstrapped fitted model
+\item
+  The 95\% quantile constructed using the original fitted model can be
+  considered as an estimate of those new 95\% quantiles
+\item
+  We borrow information from the original fitted model
+\item
+  We can check how many bootstrapped fitted model should be rejected
+  (like the percentage)
+\item
+  This gives an overall estimate of how often the assumed regression
+  model are considered to be incorrect if the data can be obtained
+  repetitively from the same data generating process
+\item
+  If this percentage is relatively large, it gives the analyst an
+  warning that the regression model is inappropriate to the data
+\end{itemize}
 
@@
 Non-normality violations are introduced by defining a non-normal
-distribution for \(\boldsymbol{\varepsilon}\). - distribution of x1 and
-x2 - summary table of all the factor values - example plots for the
-violations
+distribution for \(\boldsymbol{\varepsilon}\).
+
+\begin{itemize}
+\tightlist
+\item
+  distribution of x1 and x2
+\item
+  summary table of all the factor values
+\item
+  example plots for the violations
+\end{itemize}
 
@@
 set}\label{crafting-a-balanced-training-set}}
+
+In order to train a robust computer vision model, we intentionally
+control the distribution of the target variable \(D\) of the training
+data. Making it uniform between \(0\) and \(7\). This is accomplished by
+preparing \(50\) buckets each only accepts training samples with \(D\)
+between \([7(i - 1)/49, 7i/49)\) for \(i < 50\), where \(i\) is the
+index of the \(i\)-th bucket. For the \(50\)-th bucket, any training
+samples with \(D >= 7\) will be accepted. We have prepared 80000
+training images, so each bucket can only contain
+\(80000 \div 50 = 1600\) training samples. The simulator will repeatedly
+sample parameter values from the parameter sample, generate residuals
+and fitted values using the data generating process, compute the
+distance, and check if the sample can be accepted by the corresponding
+bucket, until all the buckets are full.
 

### 2024-01-22 paper/paper.Rmd
spell check

@@
 Unlike conventional statistical tests that can be performed numerically and automatically in statistical software, the lineup protocol requires human evaluation of images.
-This characteristic makes it less suitable for large-scale applications, given the associated high labour costs and time requirements.
+This characteristic makes it less suitable for large-scale applications, given the associated high labor costs and time requirements.
 
 There is a substantial need to develop an approach that alleviates people's workload by automating repetitive tasks and providing standardized results in a controlled environment.
-The large-scale evaluation of lineups ihs impractical without the use of technology and machines.
+The large-scale evaluation of lineups is impractical without the use of technology and machines.
 
@@
 
-# Methodogology
+# Methodology
 
@@
 However, if the lineup consists of around 20 plots, similar to previous human subject experiments (reference here), the resolution of the input image could become quite large, posing challenges in training the model.
-In fact, we experimented with this approach in a pilot study, and the model's performance was suboptimal.
+In fact, we experimented with this approach in a pilot study, and the model's performance was sub-optimal.
 
@@
 It can be known that residuals $\boldsymbol{e}$ also follow a certain probability distribution transformed from the multivariate normal distribution, which will be denoted as $Q$.
-This reference distribution $Q$ summarises what good residuals should follow given the predictors are known and fixed.
+This reference distribution $Q$ summarizes what good residuals should follow given the predictors are known and fixed.
 
@@
 Then, we can get $m$ sets of $\boldsymbol{e}$ stored in the matrix $\boldsymbol{B}$ by applying the residual operator $\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}$.
-Furthermore, kernel density estimation (KDE) with Gaussian kernel and optimal bandwidth selected by the Sliverman's rule of thumb [@silverman2018density] is applied on each row of $B$ to estimate $p_i(e_i)$ for $i = 1, ..., n$.
+Furthermore, kernel density estimation (KDE) with Gaussian kernel and optimal bandwidth selected by the Silverman's rule of thumb [@silverman2018density] is applied on each row of $B$ to estimate $p_i(e_i)$ for $i = 1, ..., n$.
 The KDE computation is done by the `density` function in R.

### 2024-01-22 paper/paper.tex
spell check

@@
 human evaluation of images. This characteristic makes it less suitable
-for large-scale applications, given the associated high labour costs and
+for large-scale applications, given the associated high labor costs and
 time requirements.
@@
 standardized results in a controlled environment. The large-scale
-evaluation of lineups ihs impractical without the use of technology and
+evaluation of lineups is impractical without the use of technology and
 machines.
@@
 
-\hypertarget{methodogology}{%
-\section{Methodogology}\label{methodogology}}
+\hypertarget{methodology}{%
+\section{Methodology}\label{methodology}}
 
@@
 In fact, we experimented with this approach in a pilot study, and the
-model's performance was suboptimal.
+model's performance was sub-optimal.
 
@@
 distribution, which will be denoted as \(Q\). This reference
-distribution \(Q\) summarises what good residuals should follow given
+distribution \(Q\) summarizes what good residuals should follow given
 the predictors are known and fixed.
@@
 Furthermore, kernel density estimation (KDE) with Gaussian kernel and
-optimal bandwidth selected by the Sliverman's rule of thumb
+optimal bandwidth selected by the Silverman's rule of thumb
 \citep{silverman2018density} is applied on each row of \(B\) to estimate

### 2024-01-29 paper/paper.Rmd
Auto stash before merge of "master" and "origin/master"

@@
 
-In addressing a problem, there exist multiple avenues for solutions.
-Various configurations of the computer vision model formula can effectively read residual plots, and these configurations can be categorized based on two key components of the model formula: the input and the output format.
+In addressing the residual reading problem, there exist multiple configurations of the computer vision model formula.
+These configurations can be categorized by two key components of the model formula: the input and the output format.
 
@@
 
-In computer vision models, the input plays a role similar to predictors in a linear regression model.
-The quality and relevance of the input data greatly influence the model's capacity to generate insightful and meaningful results.
-A straightforward approach involves feeding the model a vector of residuals along with a vector of fitted values, essentially providing all the necessary details for creating a residuals vs fitted values plot.
+In computer vision models, the quality and relevance of the input data greatly influence the model's capacity to generate insightful and meaningful results.
+The input of the model can be designed in different formats.
+
+A straightforward approach involves feeding the model a vector of residuals along with a vector of fitted values, essentially providing all the necessary information for creating a residuals vs fitted values plot.
 However, a drawback of this method is the dynamic input size, as different fitted regression models may have varying numbers of observations.
-In modern computer vision models created with mainstream software like TensorFlow (ref here), the input is typically a fixed-shape tensor.
-One solution is to pad the input vector with leading or trailing zeros if the input tensor expects a longer vector.
-However, this approach may fail if the input vector surpasses the designed length.
+For modern computer vision models implemented with mainstream software like TensorFlow [@abadi2016tensorflow], the input shape is typically fixed.
+One solution is to pad the input vectors with leading or trailing zeros if the input tensor expects longer vectors, but it may fail if the input vector surpasses the designed length.
 Another strategy is to summarize the residuals and fitted values separately using histograms and utilize the counts as the input.
-By controlling the number of bins in the histograms, it becomes possible to provide a fixed-length input vector.
+By controlling the number of bins in the histograms, it becomes possible to provide fixed-length input vectors.
 
-The primary advantage of using an image as input, as opposed to a vector input format, relates to the development of sophisticated image processing architectures over the years (see ref here).
-These architectures effectively capture and summarize spatial information from nearby pixels, which is less straightforward with vector input.
+Another approach is to use an image as input.
+The primary advantage of using an image, as opposed to the vector format, is the utilization of the sophisticated image processing architectures developed over the years, such as the VGG16 architecture proposed in @simonyan2014very.
+These architectures can effectively capture and summarize spatial information from nearby pixels, which is less straightforward with vector input.
 The only considerations are the image resolution and the aesthetics of the residual plot.
@@
 Other possible input formats include, but are not limited to, a pair of residual plots, a triplet, and a lineup.
-Existing computer vision models designed for image comparison can assess whether a pair of images are similar or dissimilar (as seen in LeCun's 2005 paper).
+@chopra2005learning has shown that computer vision models designed for image comparison can assess whether a pair of images are similar or dissimilar.
 Applied to our specific problem, we can define null plots of a fitted model to be similar to each other, while considering actual residual plots to be distinct from null plots of any fitted model.
 A triplet constitutes a set of three images, denoted as $image_1$, $image_2$ and $image_3$.
-In computer vision models, triplets are often used to predict whether $image_2$ or $image_3$ is more similar to $image_1$, proving particularly useful for establishing rankings between samples.
-In the context of our problem, we can apply the same criteria to define similarity between images.
+It is often used to predict whether $image_2$ or $image_3$ is more similar to $image_1$, proving particularly useful for establishing rankings between samples.
 However, it's important to note that these two approaches usually require additional considerations regarding the loss function and, at times, non-standard training processes due to shared weights between different convolutional blocks.
@@
 Presenting a lineup to a model and asking it to predict which residual plot is the most different one aligns closely with the lineup protocol.
-However, if the lineup consists of around 20 plots, similar to previous human subject experiments (reference here), the resolution of the input image could become quite large, posing challenges in training the model.
-In fact, we experimented with this approach in a pilot study, and the model's performance was sub-optimal.
+However, if a lineup consists of a large amount of residual plots, the resolution of the input image could become quite large, posing challenges in training the model.
+This approach was experimented in a pilot study conducted by us but the performance of the trained model was sub-optimal.
 
-We did not explore all the mentioned input formats due to the considerable costs associated with data preparation and model training.
-Considering the implementation cost and the interpretability of the model, we settled on the single residual plot input format.
+Not all the mentioned input formats were explored and tested in this study due to the considerable costs associated with data preparation and model training.
+The single residual plot input format was chosen for its relatively low implementation cost and high interpretability.
 
@@
 Given that the input is a single residual plot represented as a fixed-resolution image, the computer vision model's output can take one of two formats: binary or numeric.
-This choice determines whether the model is a classification model or a regression model.
+This choice determines whether the model belongs to a classification model or a regression model.
 The binary outcome might indicates whether the input image is a null plot or not, or whether the input image would be rejected in a visual test conducted by humans.
-The latter option requires data from prior human subject experiments, posing challenges in controlling the quality and quantity of data due to variations in experimental settings across different studies.
+The latter option requires data from prior human subject experiments, presenting difficulties in controlling the quality of data due to variations in experimental settings across different studies.
 Additionally, some visual inference experiments are unrelated to linear regression models and residual plots, resulting in a limited amount of available training data.
@@
 Studies have demonstrated that defining a proper distance between images can enhance the matching accuracy in image search compared to a binary outcome model (ref here).
-
-The following section will propose the distance measure used for training the computer vision model.
 
@@
 \label{eq:kl-0}
-D &= log\left(1 + KL\right), \\
+D &= log\left(1 + D_{KL}\right), \\
 \label{eq:kl-1}
-KL &= \int_{\mathbb{R}^{n}}log\frac{p(\boldsymbol{e})}{q(\boldsymbol{e})}p(\boldsymbol{e})d\boldsymbol{e},
+D_{KL} &= \int_{\mathbb{R}^{n}}log\frac{p(\boldsymbol{e})}{q(\boldsymbol{e})}p(\boldsymbol{e})d\boldsymbol{e},
 \end{align}
@@
 \label{eq:kl-2}
-KL &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(diag(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
+D_{KL} &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(diag(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
 \end{align}
@@
 Since the Kullback-Leibler divergence can be viewed as the expectation of the log-likelihood ratio between distribution $P$ and distribution $Q$ evaluated on distribution $P$, we can reuse the simulated residuals in matrix $\boldsymbol{B}$ to estimate the expectation by the sample mean.
-With the independence assumption, for non-normal $P$, $KL$ can be estimated by
+With the independence assumption, for non-normal $P$, $D_{KL}$ can be estimated by
 
@@
 \label{eq:kl-3}
-\hat{KL} &= \sum_{i = 1}^{n} \hat{KL_i}, \\
-\hat{KL_i} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
+\widehat{D_{KL}} &= \sum_{i = 1}^{n} \widehat{D_{KL}^{(i)}}, \\
+\widehat{D_{KL}^{(i)}} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
 \end{align}
 
-\noindent where $\hat{KL_i}$ is the estimator of the Kullback-Leibler divergence for an individual residual $e_i$, $B_{ij}$ is the $i$-th row and $j$-th column entry of the matrix $B$, $\hat{p_i}(.)$ is the kernel density estimator of $p_i(.)$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\hat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)$, and $vec(.)$ is the vectorization operator which turns a $n \times m$ matrix into a $nm \times 1$ column vector by stacking the columns of the matrix on top of each other.
+\noindent where $\widehat{D_{KL}^{(i)}}$ is the estimator of the Kullback-Leibler divergence for an individual residual $e_i$, $B_{ij}$ is the $i$-th row and $j$-th column entry of the matrix $B$, $\hat{p_i}(.)$ is the kernel density estimator of $p_i(.)$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\widehat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)$, and $vec(.)$ is the vectorization operator which turns a $n \times m$ matrix into a $nm \times 1$ column vector by stacking the columns of the matrix on top of each other.
 
@@
 \label{eq:d-approx}
-\hat{D} = cv(plot_{h\times w}(\boldsymbol{e}, \hat{\boldsymbol{y}})),
+\hat{D} = f(V(\hat{\boldsymbol{e}}, \hat{\boldsymbol{y}})),
 \end{equation}
 
-\noindent where $\boldsymbol{e}$ is a vector of residuals obtained from the fitted model rather than a vector of random variables, $\hat{\boldsymbol{y}}$ is the fitted values also obtained from the fitted model, $plot(.)_{h \times w}$ is a plotting function that generates a residuals vs fitted values plot with fixed aesthetic and then saves it as an image with $h \times w$ pixels and three colour channels, and $cv(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance in the domain $[0, +\infty)$.
+\noindent where $\hat{\boldsymbol{e}}$ is a vector of residuals obtained from the fitted model rather than a vector of random variables, $\hat{\boldsymbol{y}}$ is the fitted values also obtained from the fitted model, $V(.)$ is a plotting function that generates a residuals vs fitted values plot with fixed aesthetic and then saves it as an image with $h \times w$ pixels and three colour channels, and $f(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance in the domain $[0, +\infty)$.
 
@@
 
-Having a computer vision model $cv$ which gives an approximated distance between $0$ and $+\infty$ is not enough to determine if the null hypothesis that the model is correctly specified should be rejected.
-Theoretically, the distance $D$ for a correctly specified model is $0$, because $P$ will be the same as $Q$, and $log(1 + 0) = 0$.
-The computer vision model may not necessary predict $0$ for a null plot.
-Using \ref{fig:false-finding} as an example, it contains a visual pattern which is an indication of heteroskedasticity.
+Theoretically, the distance $D$ for a correctly specified model is $0$, because $P$ will be the same as $Q$.
+However, the computer vision model may not necessary predict $0$ for a null plot.
+Using Figure \ref{fig:false-finding} as an example, it contains a visual pattern which is an indication of heteroskedasticity.
 We would not expect the model to be able to magically tell if the suspicious pattern is caused by the skewed distribution of the fitted values or the existence of heteroskedasticity.
-Assume we only train the model with this single image, while $50$% of the time the label is $0$, and $50$% of the time the label is some constant greater than $0$.
-Then, a perfect model will predict a value between $0$ and that constant to achieve a minimum loss.
-So, it is expected that our computer vision models will also predict value greater than $0$ for some null plots.
-It will make more sense if we interpret the model prediction as the visual signal strength of the residual plot.
-Some null plots could have outliers or strong visual patterns due to randomness, and the model will try to summarise those information into the prediction.
+Assume we only train the model with this single image, while $50$% of the time the label is $0$, and $50$% of the time the label is some constant $c > 0$.
+Then, a perfect model will predict $0 < \hat{D} < c$ to achieve a minimum loss.
+Some null plots could have outliers or strong visual patterns due to randomness, and a reasonable model will try to summarise those information into the prediction, resulting in $\hat{D} > 0$.
 
-To align with the principle of visual inference, we need to provide a valid null distribution for the test statistic.
-If we treat the approximated distance $\hat{D}$ as a test statistic, then the null distribution of this statistic can be approximated by the empirical distribution of $\hat{D}$ for a large amount of null plots of given a fitted model.
+This property is not an issue if $\hat{D} \gg 0$ for which the visual signal of the residual plot is very strong, and we usually do not need any further examination of the significance of the result.
+However, if the visual pattern is weak or moderate, having $\hat{D}$ will not be sufficient to determine if the null hypothesis that the model is correctly specified should be rejected.
+
+To solve this issue while aligning with the principle of visual inference, the approximated distance $\hat{D}$ can be defined as a test statistic.
+And the null distribution of this statistic can be approximated by the empirical distribution of $\widehat{D_{null}}$ for a large amount of null plots of a fitted model.
 The procedure involves applying the residual rotation technique [@buja2009statistical] on the fitted model to obtain null residuals.
 The null residuals are then used to make null plots and fed into the model.
-The results are used to construct an empirical distribution.
+The predicted distance are used to construct an empirical distribution.
 
@@
 
-With the null distribution of the test statistic,
-
-- How many null samples are needed to get stable 95% quantile
-- How do we make decision based on the quantile (reject or not reject)
-- p-value calculation (percentage of null vss > vss)
+Let $\widehat{D_{null}^{(i)}}$ be the predicted distance of the $i$-th null plots, where $i = 1,...,m$ and $m \in \mathbb{N}^+$ is a sufficiently large number.
+Quantiles of the null distribution can then be estimated using the sample quantiles available in statistical software such as R.
+The details of the sample quantile computation can be found in @hyndman1996sample.
+In statistical testing, analysts often care about certain quantiles of the null distribution, such as 90% quantile, 95% quantile and 99% quantile.
+These quantiles are used as thresholds to decide if $H_0$ needs to be rejected.
+For example, if $\hat{D}$ is greater than and equal to the 95% sample quantile of the null distribution, we could say the predicted distance for the actual residual plot is significantly different from the predicted distance for null plots with 95% significance level.
+Based on our experience, in order to get a stable estimate of the 95% quantile, $m$ usually needs to be at least $100$.
+And if the null distribution has a long tail, more null plots will be needed.
+Alternatively, a p-value can be used to represents the probability of observing an event equally or more extreme than the given event under $H_0$, and it can be estimated by $1/m\sum_{i=1}^{m}I\left(\widehat{D_{null}^{(i)}} \geq \hat{D}\right)$.
 
@@
 
-We have mentioned in Section \ref{introduction} that scagnostics summarise some useful characteristics of the scatter plot.
-Although the computer vision model will learn the feature extraction function during the training process, it will be beneficial to utilize scagnostics to help computer vision model making more accurate predictions as these measure could provide additional useful information.
+We have mentioned in Section \ref{introduction} that scagnostics are a set of manually designed visual feature extraction functions.
+Although our computer vision model will learn its own feature extraction function during the training process, it will be beneficial to utilize additional information provided by scagnostics to help computer vision model making more accurate predictions.
 
@@
 The computed measures along with the number of observations of the fitted model are provided as the second input of the computer vision model.
-
-Other scagnostics are also informative, but unfortunately, they are unavailable due to a fatal bug in the compiled C program of the `interp` R package (ref here) that will crash the process in a unpredictable manner.
+Other scagnositcs are also informative, but unfortunately, they are unavailable due to a fatal bug in the compiled C program of the `interp` R package (ref here) that will crash the process in a unpredictable manner.
 For reproducibility, we do not include these scagnostics in the training data.
@@
 
-- optimization (MSE loss function, with equation)
-- a diagram
-- explain the existence of removable layers 
-- explain three different resolutions
+The architecture of the computer vision model is adopted from a existing architecture called VGG16 that proven to be performant in image classification [@simonyan2014very].
+The first layer is the input layer with shape $n \times h \times w \times 3$ which could accept $n$ RGB images, followed by a grayscale conversion layer using the luma formula under the CCIR 601 standard (ref here) to convert the colour image to an grey scale image.
+Grayscale is enough for our task since data points are plotted in black.
+Three different combinations of $h$ and $w$ are chosen to seek a sufficient large image resolution for this problem.
+This includes $32 \times 32$, $64 \times 64$ and $128 \times 128$.
 
-## Training process and hyperparameter tuning
+The processed image is then used as the input of the first convoluational block.
+The model has at most five consecutive convolutional blocks which is similar to the original VGG16 architecture.
+In each block, there are two 2D convoluational layers followed by two activation layers respectively, and there is a 2D max pooling layer follows the second activation layer.
+The 2D convolutional layer convolves the input with a fixed number of $3 \times 3$ convolution filters, and the 2D max pooling layer downsamples the input along its spatial dimensions by taking the maximum value over a $2 \times 2$ window for each channel of the input.
+The activation layer is set to be the rectified linear unit activation function (ReLu), which is a standard practice in deep learning.
+This introduces a non-linear transformation of the output of the 2D convolutional layer.
+Furthermore, to regularize the training, a batch normalization layer is added after each of the 2D convolutional layer and before the activation layer.
+A dropout layer is also appended at the end of each convolutional block to randomly set some inputs to zeros during training.
 
-- software environment (tf and keras)
-- tuner (keras_tuner)
-- 80% validation data, 20% training data
-- 100 trials (Bayesian optimizer)
-- monitor RMSE
-- early stopping (50 epochs)
-- learning rate reduction (10 epochs, 0.5 factor)
-- max 2000 epochs (no one reach)
-- restore best epoch of the best trial
+The output of the last convolutional block will be summarised by a global max pooling layer or a global average pooling layer to gain a two-dimensional tensor.
+Since we want to utilize the information contained in scagnostics, the two-dimensional tensor is concatenated with an additional $n \times 5$ tensor, which contains the "Monotonic", "Sparse", "Splines" and "Striped" measures and the number of observations for $n$ residual plots.
+
+The concatenated tensor will be fed into the final prediction block.
+This block contains two fully-connected layers, the first one has at least $128$ units followed by a dropout layer.
+Sometimes, a batch normalization layer will be inserted between the fully-connected layer and the dropout layer.
+The second fully-connected layer has only one unit, serves as the output of the model.
+
+The model weights are randomly initialized and they are optimized by the Adam optimizer with the mean square error loss function.
+Therefore, we try to find a set of weights $\hat{\boldsymbol{\theta}}$ by
+
+$$\hat{\boldsymbol{\theta}} = \underset{\boldsymbol{\theta}}{argmin}\frac{1}{n_{train}}\sum_{i=1}^{n_{train}}(D_i - f_{\boldsymbol{\theta}}(V_i, S_i))^2,$$
+
+\noindent where $n_{train}$ is the number of training samples, $V_i$ is the $i$-th residual plot and $S_i$ is the additional information about the $i$-th residual plot including four scagnostics and the number of observations.
+
+- a diagram for the model
+
+## Training and hyperparameter tuning
+
+To obtain an optimal deep learning model, hyperparameters such as learning rate often needs to be tuned via a tuner.
+In this study, we used the Bayesian optimization tuner from the `keras-tuner` Python library (ref here) to tune the hyperparameters.
+The complete list of hyperparameters is provided in Table \ref{tab:hyperparameter}.
+
+The number of base filters determines the number of filters for the first 2D convolutional layer.
+In VGG16, the number of filters for the 2D convolutional layer in a block is 2 times the number in the previous block, except for the last block which has the same number of convolution filters as the previous one.
+This hyperparameter helps to control the complexity of the computer vision model.
+More base filters will result in more trainable parameters.
+
+Similarly, the number of units for the first fully-connected layer determines the complexity of the final prediction block. And more units will lead to more trainable parameters.
+
+Dropout rate and batch normalization are flexible hyperparameters which works jointly with other hyperparameters to smooth the training process.
+A high dropout rate is needed when the model learns too much noise from the training dataset.
+And a low dropout rate is needed when the model is complex and difficult to converge.
+Batch normalization on the other hand may be needed for solving the internal covariate shift problem due to randomness in the weight initialization and randomness in the input data.
+
+Having additional inputs including the scagnostics and the number of observations could be beneficial for making more accurate prediction.
+So we let the tuner to decide if these inputs are actually needed.
+
+Learning rate is one of the most important hyperparameters, and it determines the step size of the optimization algorithm.
+A high learning rate could help the model to avoid getting stuck in a local minimum, but it could also ruin the optimization process by jumping over the global minimum.
+A low learning rate could smooth the training process but it increases the time to converge and the chance of getting stuck in local minimum.
+
+The model is trained on a high-performance computing platform owned by Monash university with TensorFlow (ref here) and Keras (ref here).
+During the training process, 80% of the training data are used as the actual training data, and 20% of the training data are reserved as validation data.
+We let the Bayesian optimization tuner runs 100 trials to find the best hyperparameter values according to the validation root mean square error.
+The tuner will restore the best epoch of the best model from the trials.
+Early stopping is also applied such that the training process will halt if the validation root mean square error does not improve for 50 epochs.
+The maximum allowed epochs is 2000 where no models reach.
+
+```{r hyperparameter}
+data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
+           Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[1e-8, 1e-1]")) %>%
+  kableExtra::kable("latex", booktabs = TRUE, caption = "Name of hyperparameters and their correspoding domain for the computer vision model.")
+```
 
@@
 
+- comparison with human visual inference
+- overview of the human subject experiment
+- metrics
+- agreement rate
+- check cases that are aggreed
+- check cases that are disaggreed
+
 # Results
 
-## Best model performance
+## Best hyperparameters
 
-- Metrics for model performance
-- Shap values
-- Heatmap
+Based on the tuning process described in Section \ref{training-and-hyperparameter-tuning}, the best hyperparameter values are provided in Table \ref{tab:best-hyperparameter}.
+It can be observed that at least $32$ base filters are needed, and the preferable choice for the $64 \times 64$ model and the $128 \times 128$ model is $64$ base filters, which is the same as the original VGG16 architecture.
+The best values for dropout rate for convolutional blocks is around $0.4$.
+And applying batch normalization for convolutional blocks has a positive impact on the performance.
+All best models choose to preserve the additional inputs to reduce the validation error.
+The number of units needed for the fully-connected layer is $256$, which is not a very large number comparing to the VGG16 classifier, suggesting that the problem we are trying to solve is relatively easy.
+The best learning rates are usually smaller than the default value recommended by Keras which is 0.001.
+
+```{r best-hyperparameter}
+data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
+           `32` = c("32", "0.4", "true", "max", "false", "256", "false", "0.2", "0.0003"),
+           `64` = c("64", "0.3", "true", "average", "false", "256", "true", "0.4", "0.0006"),
+           `128` = c("64", "0.4", "true", "average", "false", "256", "true", "0.1", "0.0052")) %>%
+  kableExtra::kable("latex", 
+                    booktabs = TRUE, 
+                    caption = "Hyperparameters values for the best computer vision models with differnet input sizes.", 
+                    col.names = c("Hyperparameter", "$32 \\times 32$", "$64 \\times 64$", "$128 \\times 128$"),
+                    escape = FALSE) %>%
+  kableExtra::kable_styling(latex_options = c("scale_down"))
+```
+
+## Model performance
+
+<!-- - Metrics for model performance -->
+<!-- - Shap values -->
+<!-- - Heatmap -->
+
+The training and test performance of models with three different input sizes are provided in Table \ref{tab:performance}.
+Among these models, the $64 \times 64$ model and the $32 \times 32$ model are consistently having the best metrics on the training set and the test set respectively.
+The mean absolute error shows that the difference between $\hat{D}$ and $D$ is around $0.43$ on the test set, which is very small considering the range of $D$ is normally between $0$ and $7$.
+The high $R^2$ also suggests the prediction is mostly linearly correlated with the target.
+
+Figure \ref{fig:model-performance} displays a scatter plot for $D - \hat{D}$ vs $D$.
+For residual plots with $D$ close to zero, all the models tends to over-predict $\hat{D}$, leading to negative residuals.
+When $D$ is between $1.5$ and $6$, the smooth curves indicate that $\hat{D}$ is not heavily biased, and there are some instances where $\hat{D}$ is severely under-predicted.
+And when $D$ is greater than $6$, there are more instances being under-predicted than over-predicted.
+It can be observed from Figure \ref{fig:model-performance-density} that the models produced much more $\hat{D}$ at around $1$ than expected.
+
+Given the model performance metrics, only the best model evaluated on the test set, i.e. the $32 \times 32$ model, will be used in the following analysis.
+
+```{r performance}
+test_pred <- readr::read_csv(here::here("paper/data/test_pred.csv"))
+train_pred <- readr::read_csv(here::here("paper/data/train_pred.csv"))
+meta <- readr::read_csv(here::here("paper/data/meta.csv"))
+
+test_summary <- test_pred %>% 
+  left_join(meta) %>%
+  group_by(res) %>%
+  summarise(RMSE = yardstick::rmse_vec(effect_size, vss),
+            R2 = yardstick::rsq_vec(effect_size, vss),
+            MAE = yardstick::mae_vec(effect_size, vss),
+            HUBER = yardstick::huber_loss_vec(effect_size, vss)) %>%
+  mutate(type = "test")
+
+train_summary <- train_pred %>% 
+  left_join(meta) %>%
+  group_by(res) %>%
+  summarise(RMSE = yardstick::rmse_vec(effect_size, vss),
+            R2 = yardstick::rsq_vec(effect_size, vss),
+            MAE = yardstick::mae_vec(effect_size, vss),
+            HUBER = yardstick::huber_loss_vec(effect_size, vss)) %>%
+  mutate(type = "train")
+
+train_summary %>%
+  bind_rows(test_summary) %>%
+  select(res, RMSE, R2, MAE, HUBER) %>%
+  mutate(res = glue::glue("${res} \\times {res}$")) %>%
+  mutate(across(RMSE:MAE, ~format(.x, digits = 3))) %>%
+  mutate(across(HUBER, ~format(.x, digits = 2))) %>%
+  kableExtra::kable(format = "latex", 
+                    booktabs = TRUE, 
+                    escape = FALSE,
+                    col.names = c("", "RMSE", "$R^2$", "MAE", "Huber loss"),
+                    caption = "The training and test performance of three models with different input sizes.") %>%
+  kableExtra::pack_rows("Training set", 1, 3) %>%
+  kableExtra::pack_rows("Test set", 4, 6)
+```
+
+```{r model-performance, fig.cap = "Residuals vs distance plot on training and test data for three best models with different input sizes.
+The blue lines are smoothing curves produced by fitting gnealized additive models.", dev = 'png', dpi = 300}
+train_pred %>% 
+  left_join(meta) %>%
+  mutate(type = "train") %>%
+  bind_rows(test_pred %>% 
+              left_join(meta) %>%
+              mutate(type = "test")) %>%
+  ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(effect_size, effect_size - vss), alpha = 0.03) +
+  geom_smooth(aes(effect_size, effect_size - vss), se = FALSE) +
+  facet_grid(type ~ res) +
+  coord_fixed() +
+  ylab(expression(D - hat(D))) +
+  xlab("D") +
+  theme_light()
+```
+
+```{r model-performance-density, fig.cap = "Density plot of predicted distance on training and test data for three best models with different input sizes.
+The density drawn in red is the density of the distance."}
+train_pred %>% 
+  left_join(meta) %>%
+  mutate(type = "train") %>%
+  bind_rows(test_pred %>% 
+              left_join(meta) %>%
+              mutate(type = "test")) %>%
+  ggplot() +
+  geom_density(aes(effect_size, col = "D")) +
+  geom_density(aes(vss, col = "hat_D")) +
+  facet_grid(type ~ res) +
+  scale_colour_manual(values = c("red", "blue"), labels = expression(D, hat(D))) +
+  labs(col = "") +
+  theme_light() +
+  xlab(expression(hat(D)))
+```
 
@@
 
+- basic background
+- highlights the same and the different settings
+
 ### Comparison
 
-- power comparison
-- decisions
+<!-- - power comparison -->
+<!-- - decisions -->
+
+```{r fig.cap = "Residuals vs distance plot on data used in the human subject experiment. The blue line is a smoothing curve produced by fitting a gnealized additive model."}
+vss_32 <- readRDS(here::here("paper/data/vss_32.rds"))
+
+experiment <- vi_survey %>%
+  group_by(unique_lineup_id) %>%
+  summarise(across(everything(), first)) %>%
+  select(unique_lineup_id, attention_check, null_lineup,
+         answer, effect_size, conventional_p_value,
+         p_value, type, shape, a, b, x_dist, e_dist,
+         e_sigma, include_z, k, n)
+
+experiment <- vss_32 %>%
+  left_join(experiment) %>%
+  mutate(conventional_reject = conventional_p_value <= 0.05) %>%
+  mutate(reject = p_value <= 0.05) %>%
+  mutate(model_reject = vss_p_value <= 0.05) %>%
+  mutate(model_boot_reject = vss_boot_p_value <= 0.05)
+
+experiment <- experiment %>%
+  left_join(readRDS(here::here("paper/data/actual_ss.rds")))
+
+experiment %>%
+  ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(actual_ss, actual_ss - vss)) +
+  geom_smooth(aes(actual_ss, actual_ss - vss), se = FALSE) +
+  coord_fixed() +
+  ylab(expression(D - hat(D))) +
+  xlab("D") +
+  theme_light()
+```
+
+```{r experiment-performance}
+experiment %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  group_by(type) %>%
+  summarise(RMSE = yardstick::rmse_vec(actual_ss, vss),
+            R2 = yardstick::rsq_vec(actual_ss, vss),
+            MAE = yardstick::mae_vec(actual_ss, vss),
+            HUBER = yardstick::huber_loss_vec(actual_ss, vss)) %>%
+  mutate(across(RMSE:HUBER, ~format(.x, digits = 3))) %>%
+  kableExtra::kable(format = "latex", 
+                    booktabs = TRUE, 
+                    escape = FALSE,
+                    col.names = c("Type", "RMSE", "$R^2$", "MAE", "Huber loss"),
+                    caption = "The performance of the $32 \\times 32$ model on the data used in the human subject experiment.")
+```
+
+It can be observed from Table \ref{tab:experiment-performance} that all the performance metrics are lower than the metrics evaluated on the test data.
 

### 2024-01-29 paper/paper.tex
Auto stash before merge of "master" and "origin/master"

@@
 \doublespacing
+
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 
-In addressing a problem, there exist multiple avenues for solutions.
-Various configurations of the computer vision model formula can
-effectively read residual plots, and these configurations can be
-categorized based on two key components of the model formula: the input
-and the output format.
+In addressing the residual reading problem, there exist multiple
+configurations of the computer vision model formula. These
+configurations can be categorized by two key components of the model
+formula: the input and the output format.
 
@@
 
-In computer vision models, the input plays a role similar to predictors
-in a linear regression model. The quality and relevance of the input
-data greatly influence the model's capacity to generate insightful and
-meaningful results. A straightforward approach involves feeding the
-model a vector of residuals along with a vector of fitted values,
-essentially providing all the necessary details for creating a residuals
-vs fitted values plot. However, a drawback of this method is the dynamic
-input size, as different fitted regression models may have varying
-numbers of observations. In modern computer vision models created with
-mainstream software like TensorFlow (ref here), the input is typically a
-fixed-shape tensor. One solution is to pad the input vector with leading
-or trailing zeros if the input tensor expects a longer vector. However,
-this approach may fail if the input vector surpasses the designed
+In computer vision models, the quality and relevance of the input data
+greatly influence the model's capacity to generate insightful and
+meaningful results. The input of the model can be designed in different
+formats.
+
+A straightforward approach involves feeding the model a vector of
+residuals along with a vector of fitted values, essentially providing
+all the necessary information for creating a residuals vs fitted values
+plot. However, a drawback of this method is the dynamic input size, as
+different fitted regression models may have varying numbers of
+observations. For modern computer vision models implemented with
+mainstream software like TensorFlow \citep{abadi2016tensorflow}, the
+input shape is typically fixed. One solution is to pad the input vectors
+with leading or trailing zeros if the input tensor expects longer
+vectors, but it may fail if the input vector surpasses the designed
 length. Another strategy is to summarize the residuals and fitted values
@@
 controlling the number of bins in the histograms, it becomes possible to
-provide a fixed-length input vector.
-
-The primary advantage of using an image as input, as opposed to a vector
-input format, relates to the development of sophisticated image
-processing architectures over the years (see ref here). These
-architectures effectively capture and summarize spatial information from
-nearby pixels, which is less straightforward with vector input. The only
-considerations are the image resolution and the aesthetics of the
-residual plot. In general, higher resolution provides more information
-to the model but comes with the trade-off of increased complexity and
-greater difficulty in training. As for the aesthetics of the residual
-plot, a practical solution is to consistently present residual plots in
-the same style to the model. This implies that the model can not accept
-arbitrary images as input but requires the use of the same preprocessing
-software to convert provided residuals and fitted values into a
-standardized style for the residual plot.
+provide fixed-length input vectors.
+
+Another approach is to use an image as input. The primary advantage of
+using an image, as opposed to the vector format, is the utilization of
+the sophisticated image processing architectures developed over the
+years, such as the VGG16 architecture proposed in
+\citet{simonyan2014very}. These architectures can effectively capture
+and summarize spatial information from nearby pixels, which is less
+straightforward with vector input. The only considerations are the image
+resolution and the aesthetics of the residual plot. In general, higher
+resolution provides more information to the model but comes with the
+trade-off of increased complexity and greater difficulty in training. As
+for the aesthetics of the residual plot, a practical solution is to
+consistently present residual plots in the same style to the model. This
+implies that the model can not accept arbitrary images as input but
+requires the use of the same preprocessing software to convert provided
+residuals and fitted values into a standardized style for the residual
+plot.
 
 Other possible input formats include, but are not limited to, a pair of
-residual plots, a triplet, and a lineup. Existing computer vision models
-designed for image comparison can assess whether a pair of images are
-similar or dissimilar (as seen in LeCun's 2005 paper). Applied to our
-specific problem, we can define null plots of a fitted model to be
+residual plots, a triplet, and a lineup. \citet{chopra2005learning} has
+shown that computer vision models designed for image comparison can
+assess whether a pair of images are similar or dissimilar. Applied to
+our specific problem, we can define null plots of a fitted model to be
 similar to each other, while considering actual residual plots to be
@@
 set of three images, denoted as \(image_1\), \(image_2\) and
-\(image_3\). In computer vision models, triplets are often used to
-predict whether \(image_2\) or \(image_3\) is more similar to
-\(image_1\), proving particularly useful for establishing rankings
-between samples. In the context of our problem, we can apply the same
-criteria to define similarity between images. However, it's important to
+\(image_3\). It is often used to predict whether \(image_2\) or
+\(image_3\) is more similar to \(image_1\), proving particularly useful
+for establishing rankings between samples. However, it's important to
 note that these two approaches usually require additional considerations
@@
 plot is the most different one aligns closely with the lineup protocol.
-However, if the lineup consists of around 20 plots, similar to previous
-human subject experiments (reference here), the resolution of the input
-image could become quite large, posing challenges in training the model.
-In fact, we experimented with this approach in a pilot study, and the
-model's performance was sub-optimal.
-
-We did not explore all the mentioned input formats due to the
-considerable costs associated with data preparation and model training.
-Considering the implementation cost and the interpretability of the
-model, we settled on the single residual plot input format.
+However, if a lineup consists of a large amount of residual plots, the
+resolution of the input image could become quite large, posing
+challenges in training the model. This approach was experimented in a
+pilot study conducted by us but the performance of the trained model was
+sub-optimal.
+
+Not all the mentioned input formats were explored and tested in this
+study due to the considerable costs associated with data preparation and
+model training. The single residual plot input format was chosen for its
+relatively low implementation cost and high interpretability.
 
@@
 of two formats: binary or numeric. This choice determines whether the
-model is a classification model or a regression model. The binary
-outcome might indicates whether the input image is a null plot or not,
-or whether the input image would be rejected in a visual test conducted
-by humans. The latter option requires data from prior human subject
-experiments, posing challenges in controlling the quality and quantity
+model belongs to a classification model or a regression model. The
+binary outcome might indicates whether the input image is a null plot or
+not, or whether the input image would be rejected in a visual test
+conducted by humans. The latter option requires data from prior human
+subject experiments, presenting difficulties in controlling the quality
 of data due to variations in experimental settings across different
@@
 (ref here).
-
-The following section will propose the distance measure used for
-training the computer vision model.
 
@@
 the sample mean. With the independence assumption, for non-normal \(P\),
-\(KL\) can be estimated by
+\(D_{KL}\) can be estimated by
 
@@
 \label{eq:kl-3}
-\hat{KL} &= \sum_{i = 1}^{n} \hat{KL_i}, \\
-\hat{KL_i} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
+\widehat{D_{KL}} &= \sum_{i = 1}^{n} \widehat{D_{KL}^{(i)}}, \\
+\widehat{D_{KL}^{(i)}} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
 \end{align}
 
-\noindent where \(\hat{KL_i}\) is the estimator of the Kullback-Leibler
-divergence for an individual residual \(e_i\), \(B_{ij}\) is the
-\(i\)-th row and \(j\)-th column entry of the matrix \(B\),
-\(\hat{p_i}(.)\) is the kernel density estimator of \(p_i(.)\), \(q(.)\)
-is the normal density function with mean zero and an assumed variance
-estimated as
-\(\hat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)\),
+\noindent where \(\widehat{D_{KL}^{(i)}}\) is the estimator of the
+Kullback-Leibler divergence for an individual residual \(e_i\),
+\(B_{ij}\) is the \(i\)-th row and \(j\)-th column entry of the matrix
+\(B\), \(\hat{p_i}(.)\) is the kernel density estimator of \(p_i(.)\),
+\(q(.)\) is the normal density function with mean zero and an assumed
+variance estimated as
+\(\widehat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)\),
 and \(vec(.)\) is the vectorization operator which turns a
@@
 \label{eq:d-approx}
-\hat{D} = cv(plot_{h\times w}(\boldsymbol{e}, \hat{\boldsymbol{y}})),
+\hat{D} = f(V(\hat{\boldsymbol{e}}, \hat{\boldsymbol{y}})),
 \end{equation}
 
-\noindent where \(\boldsymbol{e}\) is a vector of residuals obtained
-from the fitted model rather than a vector of random variables,
+\noindent where \(\hat{\boldsymbol{e}}\) is a vector of residuals
+obtained from the fitted model rather than a vector of random variables,
 \(\hat{\boldsymbol{y}}\) is the fitted values also obtained from the
-fitted model, \(plot(.)_{h \times w}\) is a plotting function that
-generates a residuals vs fitted values plot with fixed aesthetic and
-then saves it as an image with \(h \times w\) pixels and three colour
-channels, and \(cv(.)\) is a computer vision model which takes an
-\(h \times w\) image as input and predicts the distance in the domain
-\([0, +\infty)\).
+fitted model, \(V(.)\) is a plotting function that generates a residuals
+vs fitted values plot with fixed aesthetic and then saves it as an image
+with \(h \times w\) pixels and three colour channels, and \(f(.)\) is a
+computer vision model which takes an \(h \times w\) image as input and
+predicts the distance in the domain \([0, +\infty)\).
 
@@
 
-Having a computer vision model \(cv\) which gives an approximated
-distance between \(0\) and \(+\infty\) is not enough to determine if the
-null hypothesis that the model is correctly specified should be
-rejected. Theoretically, the distance \(D\) for a correctly specified
-model is \(0\), because \(P\) will be the same as \(Q\), and
-\(log(1 + 0) = 0\). The computer vision model may not necessary predict
-\(0\) for a null plot. Using \ref{fig:false-finding} as an example, it
-contains a visual pattern which is an indication of heteroskedasticity.
-We would not expect the model to be able to magically tell if the
-suspicious pattern is caused by the skewed distribution of the fitted
-values or the existence of heteroskedasticity. Assume we only train the
-model with this single image, while \(50\)\% of the time the label is
-\(0\), and \(50\)\% of the time the label is some constant greater than
-\(0\). Then, a perfect model will predict a value between \(0\) and that
-constant to achieve a minimum loss. So, it is expected that our computer
-vision models will also predict value greater than \(0\) for some null
-plots. It will make more sense if we interpret the model prediction as
-the visual signal strength of the residual plot. Some null plots could
-have outliers or strong visual patterns due to randomness, and the model
-will try to summarise those information into the prediction.
-
-To align with the principle of visual inference, we need to provide a
-valid null distribution for the test statistic. If we treat the
-approximated distance \(\hat{D}\) as a test statistic, then the null
-distribution of this statistic can be approximated by the empirical
-distribution of \(\hat{D}\) for a large amount of null plots of given a
-fitted model. The procedure involves applying the residual rotation
-technique \citep{buja2009statistical} on the fitted model to obtain null
-residuals. The null residuals are then used to make null plots and fed
-into the model. The results are used to construct an empirical
-distribution.
+Theoretically, the distance \(D\) for a correctly specified model is
+\(0\), because \(P\) will be the same as \(Q\). However, the computer
+vision model may not necessary predict \(0\) for a null plot. Using
+Figure \ref{fig:false-finding} as an example, it contains a visual
+pattern which is an indication of heteroskedasticity. We would not
+expect the model to be able to magically tell if the suspicious pattern
+is caused by the skewed distribution of the fitted values or the
+existence of heteroskedasticity. Assume we only train the model with
+this single image, while \(50\)\% of the time the label is \(0\), and
+\(50\)\% of the time the label is some constant \(c > 0\). Then, a
+perfect model will predict \(0 < \hat{D} < c\) to achieve a minimum
+loss. Some null plots could have outliers or strong visual patterns due
+to randomness, and a reasonable model will try to summarise those
+information into the prediction, resulting in \(\hat{D} > 0\).
+
+This property is not an issue if \(\hat{D} \gg 0\) for which the visual
+signal of the residual plot is very strong, and we usually do not need
+any further examination of the significance of the result. However, if
+the visual pattern is weak or moderate, having \(\hat{D}\) will not be
+sufficient to determine if the null hypothesis that the model is
+correctly specified should be rejected.
+
+To solve this issue while aligning with the principle of visual
+inference, the approximated distance \(\hat{D}\) can be defined as a
+test statistic. And the null distribution of this statistic can be
+approximated by the empirical distribution of \(\widehat{D_{null}}\) for
+a large amount of null plots of a fitted model. The procedure involves
+applying the residual rotation technique \citep{buja2009statistical} on
+the fitted model to obtain null residuals. The null residuals are then
+used to make null plots and fed into the model. The predicted distance
+are used to construct an empirical distribution.
 
@@
 
-With the null distribution of the test statistic,
-
-\begin{itemize}
-\tightlist
-\item
-  How many null samples are needed to get stable 95\% quantile
-\item
-  How do we make decision based on the quantile (reject or not reject)
-\item
-  p-value calculation (percentage of null vss \textgreater{} vss)
-\end{itemize}
+Let \(\widehat{D_{null}^{(i)}}\) be the predicted distance of the
+\(i\)-th null plots, where \(i = 1,...,m\) and \(m \in \mathbb{N}^+\) is
+a sufficiently large number. Quantiles of the null distribution can then
+be estimated using the sample quantiles available in statistical
+software such as R. The details of the sample quantile computation can
+be found in \citet{hyndman1996sample}. In statistical testing, analysts
+often care about certain quantiles of the null distribution, such as
+90\% quantile, 95\% quantile and 99\% quantile. These quantiles are used
+as thresholds to decide if \(H_0\) needs to be rejected. For example, if
+\(\hat{D}\) is greater than and equal to the 95\% sample quantile of the
+null distribution, we could say the predicted distance for the actual
+residual plot is significantly different from the predicted distance for
+null plots with 95\% significance level. Based on our experience, in
+order to get a stable estimate of the 95\% quantile, \(m\) usually needs
+to be at least \(100\). And if the null distribution has a long tail,
+more null plots will be needed. Alternatively, a p-value can be used to
+represents the probability of observing an event equally or more extreme
+than the given event under \(H_0\), and it can be estimated by
+\(1/m\sum_{i=1}^{m}I\left(\widehat{D_{null}^{(i)}} \geq \hat{D}\right)\).
 
@@
 
-We have mentioned in Section \ref{introduction} that scagnostics
-summarise some useful characteristics of the scatter plot. Although the
-computer vision model will learn the feature extraction function during
-the training process, it will be beneficial to utilize scagnostics to
-help computer vision model making more accurate predictions as these
-measure could provide additional useful information.
+We have mentioned in Section \ref{introduction} that scagnostics are a
+set of manually designed visual feature extraction functions. Although
+our computer vision model will learn its own feature extraction function
+during the training process, it will be beneficial to utilize additional
+information provided by scagnostics to help computer vision model making
+more accurate predictions.
 
@@
 measures along with the number of observations of the fitted model are
-provided as the second input of the computer vision model.
-
-Other scagnostics are also informative, but unfortunately, they are
+provided as the second input of the computer vision model. Other
+scagnositcs are also informative, but unfortunately, they are
 unavailable due to a fatal bug in the compiled C program of the
@@
 
+The architecture of the computer vision model is adopted from a existing
+architecture called VGG16 that proven to be performant in image
+classification \citep{simonyan2014very}. The first layer is the input
+layer with shape \(n \times h \times w \times 3\) which could accept
+\(n\) RGB images, followed by a grayscale conversion layer using the
+luma formula under the CCIR 601 standard (ref here) to convert the
+colour image to an grey scale image. Grayscale is enough for our task
+since data points are plotted in black. Three different combinations of
+\(h\) and \(w\) are chosen to seek a sufficient large image resolution
+for this problem. This includes \(32 \times 32\), \(64 \times 64\) and
+\(128 \times 128\).
+
+The processed image is then used as the input of the first
+convoluational block. The model has at most five consecutive
+convolutional blocks which is similar to the original VGG16
+architecture. In each block, there are two 2D convoluational layers
+followed by two activation layers respectively, and there is a 2D max
+pooling layer follows the second activation layer. The 2D convolutional
+layer convolves the input with a fixed number of \(3 \times 3\)
+convolution filters, and the 2D max pooling layer downsamples the input
+along its spatial dimensions by taking the maximum value over a
+\(2 \times 2\) window for each channel of the input. The activation
+layer is set to be the rectified linear unit activation function (ReLu),
+which is a standard practice in deep learning. This introduces a
+non-linear transformation of the output of the 2D convolutional layer.
+Furthermore, to regularize the training, a batch normalization layer is
+added after each of the 2D convolutional layer and before the activation
+layer. A dropout layer is also appended at the end of each convolutional
+block to randomly set some inputs to zeros during training.
+
+The output of the last convolutional block will be summarised by a
+global max pooling layer or a global average pooling layer to gain a
+two-dimensional tensor. Since we want to utilize the information
+contained in scagnostics, the two-dimensional tensor is concatenated
+with an additional \(n \times 5\) tensor, which contains the
+``Monotonic'', ``Sparse'', ``Splines'' and ``Striped'' measures and the
+number of observations for \(n\) residual plots.
+
+The concatenated tensor will be fed into the final prediction block.
+This block contains two fully-connected layers, the first one has at
+least \(128\) units followed by a dropout layer. Sometimes, a batch
+normalization layer will be inserted between the fully-connected layer
+and the dropout layer. The second fully-connected layer has only one
+unit, serves as the output of the model.
+
+The model weights are randomly initialized and they are optimized by the
+Adam optimizer with the mean square error loss function. Therefore, we
+try to find a set of weights \(\hat{\boldsymbol{\theta}}\) by
+
+\[\hat{\boldsymbol{\theta}} = \underset{\boldsymbol{\theta}}{argmin}\frac{1}{n_{train}}\sum_{i=1}^{n_{train}}(D_i - f_{\boldsymbol{\theta}}(V_i, S_i))^2,\]
+
+\noindent where \(n_{train}\) is the number of training samples, \(V_i\)
+is the \(i\)-th residual plot and \(S_i\) is the additional information
+about the \(i\)-th residual plot including four scagnostics and the
+number of observations.
+
 \begin{itemize}
@@
 \item
-  optimization (MSE loss function, with equation)
-\item
-  a diagram
-\item
-  explain the existence of removable layers
-\item
-  explain three different resolutions
+  a diagram for the model
 \end{itemize}
 
-\hypertarget{training-process-and-hyperparameter-tuning}{%
-\subsection{Training process and hyperparameter
-tuning}\label{training-process-and-hyperparameter-tuning}}
-
-\begin{itemize}
-\tightlist
-\item
-  software environment (tf and keras)
-\item
-  tuner (keras\_tuner)
-\item
-  80\% validation data, 20\% training data
-\item
-  100 trials (Bayesian optimizer)
-\item
-  monitor RMSE
-\item
-  early stopping (50 epochs)
-\item
-  learning rate reduction (10 epochs, 0.5 factor)
-\item
-  max 2000 epochs (no one reach)
-\item
-  restore best epoch of the best trial
-\end{itemize}
+\hypertarget{training-and-hyperparameter-tuning}{%
+\subsection{Training and hyperparameter
+tuning}\label{training-and-hyperparameter-tuning}}
+
+To obtain an optimal deep learning model, hyperparameters such as
+learning rate often needs to be tuned via a tuner. In this study, we
+used the Bayesian optimization tuner from the \texttt{keras-tuner}
+Python library (ref here) to tune the hyperparameters. The complete list
+of hyperparameters is provided in Table \ref{tab:hyperparameter}.
+
+The number of base filters determines the number of filters for the
+first 2D convolutional layer. In VGG16, the number of filters for the 2D
+convolutional layer in a block is 2 times the number in the previous
+block, except for the last block which has the same number of
+convolution filters as the previous one. This hyperparameter helps to
+control the complexity of the computer vision model. More base filters
+will result in more trainable parameters.
+
+Similarly, the number of units for the first fully-connected layer
+determines the complexity of the final prediction block. And more units
+will lead to more trainable parameters.
+
+Dropout rate and batch normalization are flexible hyperparameters which
+works jointly with other hyperparameters to smooth the training process.
+A high dropout rate is needed when the model learns too much noise from
+the training dataset. And a low dropout rate is needed when the model is
+complex and difficult to converge. Batch normalization on the other hand
+may be needed for solving the internal covariate shift problem due to
+randomness in the weight initialization and randomness in the input
+data.
+
+Having additional inputs including the scagnostics and the number of
+observations could be beneficial for making more accurate prediction. So
+we let the tuner to decide if these inputs are actually needed.
+
+Learning rate is one of the most important hyperparameters, and it
+determines the step size of the optimization algorithm. A high learning
+rate could help the model to avoid getting stuck in a local minimum, but
+it could also ruin the optimization process by jumping over the global
+minimum. A low learning rate could smooth the training process but it
+increases the time to converge and the chance of getting stuck in local
+minimum.
+
+The model is trained on a high-performance computing platform owned by
+Monash university with TensorFlow (ref here) and Keras (ref here).
+During the training process, 80\% of the training data are used as the
+actual training data, and 20\% of the training data are reserved as
+validation data. We let the Bayesian optimization tuner runs 100 trials
+to find the best hyperparameter values according to the validation root
+mean square error. The tuner will restore the best epoch of the best
+model from the trials. Early stopping is also applied such that the
+training process will halt if the validation root mean square error does
+not improve for 50 epochs. The maximum allowed epochs is 2000 where no
+models reach.
+
+\begin{table}
+
+\caption{\label{tab:hyperparameter}Name of hyperparameters and their correspoding domain for the computer vision model.}
+\centering
+\begin{tabular}[t]{ll}
+\toprule
+Hyperparameter & Domain\\
+\midrule
+Number of base filters & \{4, 8, 16, 32, 64\}\\
+Dropout rate for convolutional blocks & {}[0.1, 0.6]\\
+Batch normalization for convolutional blocks & \{false, true\}\\
+Type of global pooling & \{max, average\}\\
+Ignore additional inputs & \{false, true\}\\
+\addlinespace
+Number of units for the fully-connected layer & \{128, 256, 512, 1024, 2048\}\\
+Batch normalization for the fully-connected layer & \{false, true\}\\
+Dropout rate for the fully-connected layer & {}[0.1, 0.6]\\
+Learning rate & {}[1e-8, 1e-1]\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 \begin{itemize}
-\tightlist
 \item
@@
   Percentage of predictions within a tolerance interval (like 0.1)
+\item
+  comparison with human visual inference
+\item
+  overview of the human subject experiment
+\item
+  metrics
+\item
+  agreement rate
+\item
+  check cases that are aggreed
+\item
+  check cases that are disaggreed
 \end{itemize}
@@
 
-\hypertarget{best-model-performance}{%
-\subsection{Best model performance}\label{best-model-performance}}
-
-\begin{itemize}
-\tightlist
-\item
-  Metrics for model performance
-\item
-  Shap values
-\item
-  Heatmap
-\end{itemize}
+\hypertarget{best-hyperparameters}{%
+\subsection{Best hyperparameters}\label{best-hyperparameters}}
+
+Based on the tuning process described in Section
+\ref{training-and-hyperparameter-tuning}, the best hyperparameter values
+are provided in Table \ref{tab:best-hyperparameter}. It can be observed
+that at least \(32\) base filters are needed, and the preferable choice
+for the \(64 \times 64\) model and the \(128 \times 128\) model is
+\(64\) base filters, which is the same as the original VGG16
+architecture. The best values for dropout rate for convolutional blocks
+is around \(0.4\). And applying batch normalization for convolutional
+blocks has a positive impact on the performance. All best models choose
+to preserve the additional inputs to reduce the validation error. The
+number of units needed for the fully-connected layer is \(256\), which
+is not a very large number comparing to the VGG16 classifier, suggesting
+that the problem we are trying to solve is relatively easy. The best
+learning rates are usually smaller than the default value recommended by
+Keras which is 0.001.
+
+\begin{table}
+
+\caption{\label{tab:best-hyperparameter}Hyperparameters values for the best computer vision models with differnet input sizes.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{llll}
+\toprule
+Hyperparameter & $32 \times 32$ & $64 \times 64$ & $128 \times 128$\\
+\midrule
+Number of base filters & 32 & 64 & 64\\
+Dropout rate for convolutional blocks & 0.4 & 0.3 & 0.4\\
+Batch normalization for convolutional blocks & true & true & true\\
+Type of global pooling & max & average & average\\
+Ignore additional inputs & false & false & false\\
+\addlinespace
+Number of units for the fully-connected layer & 256 & 256 & 256\\
+Batch normalization for the fully-connected layer & false & true & true\\
+Dropout rate for the fully-connected layer & 0.2 & 0.4 & 0.1\\
+Learning rate & 0.0003 & 0.0006 & 0.0052\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\hypertarget{model-performance}{%
+\subsection{Model performance}\label{model-performance}}
+
+The training and test performance of models with three different input
+sizes are provided in Table \ref{tab:performance}. Among these models,
+the \(64 \times 64\) model and the \(32 \times 32\) model are
+consistently having the best metrics on the training set and the test
+set respectively. The mean absolute error shows that the difference
+between \(\hat{D}\) and \(D\) is around \(0.43\) on the test set, which
+is very small considering the range of \(D\) is normally between \(0\)
+and \(7\). The high \(R^2\) also suggests the prediction is mostly
+linearly correlated with the target.
+
+Figure \ref{fig:model-performance} displays a scatter plot for
+\(D - \hat{D}\) vs \(D\). For residual plots with \(D\) close to zero,
+all the models tends to over-predict \(\hat{D}\), leading to negative
+residuals. When \(D\) is between \(1.5\) and \(6\), the smooth curves
+indicate that \(\hat{D}\) is not heavily biased, and there are some
+instances where \(\hat{D}\) is severely under-predicted. And when \(D\)
+is greater than \(6\), there are more instances being under-predicted
+than over-predicted. It can be observed from Figure
+\ref{fig:model-performance-density} that the models produced much more
+\(\hat{D}\) at around \(1\) than expected.
+
+Given the model performance metrics, only the best model evaluated on
+the test set, i.e.~the \(32 \times 32\) model, will be used in the
+following analysis.
+
+\begin{table}
+
+\caption{\label{tab:performance}The training and test performance of three models with different input sizes.}
+\centering
+\begin{tabular}[t]{lllll}
+\toprule
+ & RMSE & $R^2$ & MAE & Huber loss\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{5}{l}{\textbf{Training set}}\\
+\hspace{1em}$32 \times 32$ & 0.531 & 0.937 & 0.364 & 0.126\\
+\hspace{1em}$64 \times 64$ & 0.405 & 0.963 & 0.260 & 0.072\\
+\hspace{1em}$128 \times 128$ & 0.432 & 0.959 & 0.290 & 0.084\\
+\addlinespace[0.3em]
+\multicolumn{5}{l}{\textbf{Test set}}\\
+\hspace{1em}$32 \times 32$ & 0.660 & 0.901 & 0.434 & 0.181\\
+\hspace{1em}$64 \times 64$ & 0.674 & 0.897 & 0.438 & 0.186\\
+\hspace{1em}$128 \times 128$ & 0.692 & 0.892 & 0.460 & 0.199\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/model-performance-1}
+
+}
+
+\caption{Residuals vs distance plot on training and test data for three best models with different input sizes.
+The blue lines are smoothing curves produced by fitting gnealized additive models.}\label{fig:model-performance}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/model-performance-density-1}
+
+}
+
+\caption{Density plot of predicted distance on training and test data for three best models with different input sizes.
+The density drawn in red is the density of the distance.}\label{fig:model-performance-density}
+\end{figure}
 
@@
 
+\begin{itemize}
+\tightlist
+\item
+  basic background
+\item
+  highlights the same and the different settings
+\end{itemize}
+
 \hypertarget{comparison}{%
@@
 
-\begin{itemize}
-\tightlist
-\item
-  power comparison
-\item
-  decisions
-\end{itemize}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-2-1}
+
+}
+
+\caption{Residuals vs distance plot on data used in the human subject experiment. The blue line is a smoothing curve produced by fitting a gnealized additive model.}\label{fig:unnamed-chunk-2}
+\end{figure}
+
+\begin{table}
+
+\caption{\label{tab:experiment-performance}The performance of the $32 \times 32$ model on the data used in the human subject experiment.}
+\centering
+\begin{tabular}[t]{lllll}
+\toprule
+Type & RMSE & $R^2$ & MAE & Huber loss\\
+\midrule
+heteroskedasticity & 0.768 & 0.863 & 0.594 & 0.263\\
+non-linearity & 0.731 & 0.817 & 0.560 & 0.242\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+It can be observed from Table \ref{tab:experiment-performance} that all
+the performance metrics are lower than the metrics evaluated on the test
+data.
 

### 2024-01-29 paper/paper.Rmd
add delta and gamma metrics

@@
   summarise(across(everything(), first)) %>%
-  select(unique_lineup_id, attention_check, null_lineup,
+  select(unique_lineup_id, attention_check, null_lineup, prop_detect,
          answer, effect_size, conventional_p_value,
@@
 
+```{r fig.cap = "A weighted detection rate vs $\\delta$-differnence plot."}
+lineup_vss <- readRDS(here::here("paper/data/lineup_vss.rds"))
+lineup_vss <- lineup_vss %>%
+  group_by(unique_lineup_id) %>%
+  mutate(delta_diff = vss[!null] - max(vss[null])) %>%
+  mutate(gamma_diff = sum(vss[null] > vss[!null]))
+
+experiment %>%
+  left_join(lineup_vss %>%
+              group_by(unique_lineup_id) %>%
+              summarise(delta_diff = first(delta_diff), gamma_diff = first(gamma_diff))) %>%
+  ggplot() +
+  geom_point(aes(delta_diff, prop_detect)) +
+  geom_smooth(aes(delta_diff, prop_detect), se = FALSE) +
+  ylab("Weighted detection rate") +
+  xlab(expression(gamma-difference)) +
+  theme_light()
+```
+
+```{r fig.cap = "A weighted detection rate vs $\\gamma$-number of extreme nulls plot."}
+experiment %>%
+  left_join(lineup_vss %>%
+              group_by(unique_lineup_id) %>%
+              summarise(delta_diff = first(delta_diff), gamma_diff = first(gamma_diff))) %>%
+  ggplot() +
+  geom_point(aes(gamma_diff, prop_detect)) +
+  ylab("Weighted detection rate") +
+  xlab(expression(gamma-number~of~extreme~nulls)) +
+  theme_light()
+```
+
+```{r fig.cap = "A weighted detection rate vs $\\delta$-differnence plot for lineups evaluated by 11 people."}
+experiment %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  filter(x_dist == "uniform") %>%
+  left_join(lineup_vss %>%
+              group_by(unique_lineup_id) %>%
+              summarise(delta_diff = first(delta_diff), gamma_diff = first(gamma_diff))) %>%
+  ggplot() +
+  geom_point(aes(delta_diff, prop_detect)) +
+  geom_smooth(aes(delta_diff, prop_detect), se = FALSE) +
+  ylab("Weighted detection rate") +
+  xlab(expression(delta-difference)) +
+  theme_light()
+```
+
+```{r fig.cap = "A weighted detection rate vs $\\gamma$-number of extreme nulls plot for lineups evaluated by 11 people."}
+experiment %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  filter(x_dist == "uniform") %>%
+  left_join(lineup_vss %>%
+              group_by(unique_lineup_id) %>%
+              summarise(delta_diff = first(delta_diff), gamma_diff = first(gamma_diff))) %>%
+  ggplot() +
+  geom_point(aes(gamma_diff, prop_detect)) +
+  ylab("Weighted detection rate") +
+  xlab(expression(gamma-number~of~extreme~nulls)) +
+  theme_light()
+```
+
 ## When the model works

### 2024-01-29 paper/paper.tex
add delta and gamma metrics

@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-3-1}
+
+}
+
+\caption{A weighted detection rate vs $\delta$-differnence plot.}\label{fig:unnamed-chunk-3}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-4-1}
+
+}
+
+\caption{A weighted detection rate vs $\gamma$-number of extreme nulls plot.}\label{fig:unnamed-chunk-4}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-5-1}
+
+}
+
+\caption{A weighted detection rate vs $\delta$-differnence plot for lineups evaluated by 11 people.}\label{fig:unnamed-chunk-5}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-6-1}
+
+}
+
+\caption{A weighted detection rate vs $\gamma$-number of extreme nulls plot for lineups evaluated by 11 people.}\label{fig:unnamed-chunk-6}
+\end{figure}
+
 \hypertarget{when-the-model-works}{%

### 2024-02-05 paper/paper.Rmd
update the paper

@@
 title: |
-  Automated reading of residual plots with computer vision models
+  Automated assessment of residual plots with computer vision models
 type: ARTICLE TEMPLATE
@@
 
-The practice of plotting residuals is commonly regarded as a standard procedure in linear regression diagnostics \citep[see][]{cook1982residuals, belsley1980regression}.
+Plotting residuals is commonly regarded as a standard practice in linear regression diagnostics \citep[see][]{cook1982residuals, belsley1980regression}.
 This visual assessment plays a crucial role in identifying deviations from model assumptions, such as linearity, homoscedasticity, and normality.
@@
 Consider Figure \ref{fig:false-finding} as an example, the residuals display a triangular shape pointing to the left.
-While this might suggest heteroskedasticity, it is important to avoid over-interpreting the visual pattern.
+While this might suggest heteroskedasticity, it is important to avoid over-interpreting this visual pattern.
 In this case, the fitted model is correctly specified, and the triangular shape is actually a result of the skewed distribution of the predictors, rather than indicating a flaw in the model.
@@
 However, as pointed out by @li2023plot, a primary limitation of the lineup protocol lies in its reliance on human judgments.
-Unlike conventional statistical tests that can be performed numerically and automatically in statistical software, the lineup protocol requires human evaluation of images.
+Unlike conventional statistical tests that can be performed computationally in statistical software, the lineup protocol requires human evaluation of images.
 This characteristic makes it less suitable for large-scale applications, given the associated high labor costs and time requirements.
 
-There is a substantial need to develop an approach that alleviates people's workload by automating repetitive tasks and providing standardized results in a controlled environment.
+There is a substantial need to develop an approach that alleviates analysts' workload by automating repetitive tasks and providing standardized results in a controlled environment.
 The large-scale evaluation of lineups is impractical without the use of technology and machines.
@@
 
-<!-- - discuss different types of inputs and output choices, and why we choose the current design -->
-
-In addressing the residual reading problem, there exist multiple configurations of the computer vision model formula.
-These configurations can be categorized by two key components of the model formula: the input and the output format.
+There are various configurations of the computer vision model that can be used to assess residual plots.
+We discuss these configurations below based on two key components of the model formula: the input and the output format.
 
@@
 
-In computer vision models, the quality and relevance of the input data greatly influence the model's capacity to generate insightful and meaningful results.
-The input of the model can be designed in different formats.
+Deep learning models are in general very sensitive to the input data. The quality and relevance of the input data greatly influence the model's capacity to generate insightful and meaningful results.
 
-A straightforward approach involves feeding the model a vector of residuals along with a vector of fitted values, essentially providing all the necessary information for creating a residuals vs fitted values plot.
-However, a drawback of this method is the dynamic input size, as different fitted regression models may have varying numbers of observations.
-For modern computer vision models implemented with mainstream software like TensorFlow [@abadi2016tensorflow], the input shape is typically fixed.
-One solution is to pad the input vectors with leading or trailing zeros if the input tensor expects longer vectors, but it may fail if the input vector surpasses the designed length.
+A straightforward approach involves feeding a vector of residuals along with a vector of fitted values, essentially providing all the necessary information for creating a residuals vs fitted values plot.
+However, a drawback of this method is the dynamic input size, which changes based on the number of observations.
+For modern computer vision models implemented in mainstream software like TensorFlow [@abadi2016tensorflow], the input shape is typically fixed.
+One solution is to pad the input vectors with leading or trailing zeros when the input tensor expects longer vectors, but it may fail if the input vector surpasses the designed length.
 Another strategy is to summarize the residuals and fitted values separately using histograms and utilize the counts as the input.
 By controlling the number of bins in the histograms, it becomes possible to provide fixed-length input vectors.
+Still, since the histogram only captures the marginal distribution of residuals and fitted values respectively, it can not be used to differentiate visual patterns with same marginal distributions but different joint distributions.
 
@@
 
-Other possible input formats include, but are not limited to, a pair of residual plots, a triplet, and a lineup.
+Providing multiple residual plots to the model, such as a pair of plots, a triplet or a lineup is also a possible option.
 @chopra2005learning has shown that computer vision models designed for image comparison can assess whether a pair of images are similar or dissimilar.
@@
 It is often used to predict whether $image_2$ or $image_3$ is more similar to $image_1$, proving particularly useful for establishing rankings between samples.
+For this setup, we can apply the same criteria to define similarity between images.
 However, it's important to note that these two approaches usually require additional considerations regarding the loss function and, at times, non-standard training processes due to shared weights between different convolutional blocks.
@@
 
-Not all the mentioned input formats were explored and tested in this study due to the considerable costs associated with data preparation and model training.
-The single residual plot input format was chosen for its relatively low implementation cost and high interpretability.
+We did not explore all the mentioned input formats due the considerable costs associated with data preparation and model training.
+Considering the implementation cost and the interpretability of the model, we settled on the single residual plot input format.
 
@@
 
-Given that the input is a single residual plot represented as a fixed-resolution image, the computer vision model's output can take one of two formats: binary or numeric.
+Given that the input is a single residual plot represented as a fixed-resolution image, the output from the computer vision model can take one of two formats: binary or numeric.
 This choice determines whether the model belongs to a classification model or a regression model.
-The binary outcome might indicates whether the input image is a null plot or not, or whether the input image would be rejected in a visual test conducted by humans.
+The binary outcome could be designed to indicate whether the input image is a null plot, or whether the input image would be rejected in a visual test conducted by humans.
 The latter option requires data from prior human subject experiments, presenting difficulties in controlling the quality of data due to variations in experimental settings across different studies.
-Additionally, some visual inference experiments are unrelated to linear regression models and residual plots, resulting in a limited amount of available training data.
+Additionally, some visual inference experiments are unrelated to linear regression models and residual plot diagnostics, resulting in a limited amount of available training data.
 
-Alternatively, the output could be a meaningful numeric measure, such as the average distance to good residual plots.
+Alternatively, the output could be designed as a meaningful numeric measure, such as the average distance to good residual plots.
 This approach necessitates defining a distance measure between residual plots, which may be a non-trivial task.
-However, once a meaningful distance measure is established, the computer vision model's prediction becomes an interpretable value with tangible significance.
+However, once a meaningful distance measure is established, the prediction from the computer vision model becomes an interpretable value with tangible significance.
 Studies have demonstrated that defining a proper distance between images can enhance the matching accuracy in image search compared to a binary outcome model (ref here).
@@
 In a visual test, the observer will be asked to choose one or more plots that stand out as most distinct from others in a given lineup.
-To develop a computer vision model for evaluating residual plots within the visual inference framework, it is important to precisely define a numerical measure of "difference" or "distance" between plots.
+To develop a computer vision model for assessing residual plots within the visual inference framework, it is important to precisely define a numerical measure of "difference" or "distance" between plots.
 This distance can take the form of a basic statistical operation on pixels, such as the sum of square differences.
@@
 The distance metrics proposed in this paper takes into account the fact that we try to measure how different a residual plot is from a good residual plot, or in other words, how different a given fitted model is from a correctly specified model.
-For the classical normal linear regression model, residuals are derived from the fitted values $\hat{\boldsymbol{y}}$ and observed values $\boldsymbol{y}$.
-Suppose the data generating process is known and the model is correctly specified, by the Frisch-Waugh-Lowell theorem [@frisch1933partial], residuals $\boldsymbol{e}$ can also be written as a linear transformation of the error $\boldsymbol{\varepsilon}$ formulated as $\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon}$, where $\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the residual operator, $\boldsymbol{X}$ is the design matrix, $\boldsymbol{I}_n$ is a $n$ by $n$ identity matrix, and $n$ is the number of observations.
+For the classical normal linear regression model, residuals $\boldsymbol{e}$ are derived from the fitted values $\hat{\boldsymbol{y}}$ and observed values $\boldsymbol{y}$.
+Suppose the data generating process is known and the regression model is correctly specified, by the Frisch-Waugh-Lowell theorem [@frisch1933partial], residuals $\boldsymbol{e}$ can also be treated as random variables and written as a linear transformation of the error $\boldsymbol{\varepsilon}$ formulated as $\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon}$, where $\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the residual operator and an idempotent matrix, $\boldsymbol{X}$ is the design matrix, $\boldsymbol{I}_n$ is a $n$ by $n$ identity matrix, and $n$ is the number of observations.
 
@@
 
-When fitting a linear regression model, the solver will force $\sum_{i=1}^{n} e_i = 0$, making any residual value to be a linear combination of the remaining $n - 1$ residuals.
-This effectively means $rank(\boldsymbol{R}) = n - 1 < n$ and $Q$ becomes a degenerate multivariate distribution.
+In ordinary least square, the minimization of the sum of square residuals implies  $\sum_{i=1}^{n} e_i = 0$, making any residual value to be a linear combination of the remaining $n - 1$ residuals.
+This effectively means $\text{rank}(\boldsymbol{R}) = n - 1 < n$ and $Q$ is a degenerate multivariate distribution.
 To capture the characteristics of $Q$, such as moments, we can simulate a large numbers of $\boldsymbol{\varepsilon}$ and transform it to $\boldsymbol{e}$ to get the empirical estimates.
-For simplicity, we replaced $\boldsymbol{R}$ with a full-rank diagonal matrix $diag(\boldsymbol{R})$, where $diag(.)$ set the non-diagonal entries of a matrix to zeros.
-The resulting distribution for $\boldsymbol{Q}$ is $N(\boldsymbol{0}_n, diag(\boldsymbol{R}\sigma^2))$.
+For simplicity, we replaced the variance-covariance matrix of residuals $\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}' = \boldsymbol{R}\sigma^2$ with a full-rank diagonal matrix $\text{diag}(\boldsymbol{R}\sigma^2)$, where $\text{diag}(.)$ sets the non-diagonal entries of a matrix to zeros.
+The resulting distribution for $\boldsymbol{Q}$ is $N(\boldsymbol{0}_n, \text{diag}(\boldsymbol{R}\sigma^2))$.
 
 Distribution $Q$ is derived from the correctly specified model.
-However, if the model is misspecified, then the actual distribution of residuals denoted as $P$, will be different to $Q$.
-For example, if the data generating process contains variables correlated with any column of $\boldsymbol{X}$ but not included in $\boldsymbol{X}$, causing an omitted variable problem, $P$ will be different to $Q$ because the residual operator obtained from the fitted model will not be the same as $\boldsymbol{R}$.
-Besides, if the $\boldsymbol{\varepsilon}$ follows a non-normal distribution such as a multivariate lognormal distribution, the empirical residual distribution will usually be skewed and has a long tail.
+However, if the model is misspecified, then the actual distribution of residuals denoted as $P$, will be different from $Q$.
+For example, if the data generating process contains variables correlated with any column of $\boldsymbol{X}$ but not included in $\boldsymbol{X}$, causing an omitted variable problem, $P$ will be different from $Q$ because the residual operator obtained from the fitted model will not be the same as $\boldsymbol{R}$.
+Besides, if the $\boldsymbol{\varepsilon}$ follows a non-normal distribution such as a multivariate lognormal distribution, $P$ will usually be skewed and has a long tail.
 
@@
 \label{eq:kl-0}
-D &= log\left(1 + D_{KL}\right), \\
+D &= \log\left(1 + D_{KL}\right), \\
 \label{eq:kl-1}
-D_{KL} &= \int_{\mathbb{R}^{n}}log\frac{p(\boldsymbol{e})}{q(\boldsymbol{e})}p(\boldsymbol{e})d\boldsymbol{e},
+D_{KL} &= \int_{\mathbb{R}^{n}}\log\frac{p(\boldsymbol{e})}{q(\boldsymbol{e})}p(\boldsymbol{e})d\boldsymbol{e},
 \end{align}
@@
 It was mainly designed for measuring the effect size of non-linearity and heteroskedasticity in a residual plot.
-@li2023plot showed that, for a classical normal linear regression model that omits a necessary higher-order predictors $\boldsymbol{Z}$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))$.
+@li2023plot showed that, for a classical normal linear regression model that omits a necessary higher-order predictors $\boldsymbol{Z}$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$ with $\boldsymbol{V}$ being an arbitrary symmetric positive semi-definite matrix, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}))$.
 Note that the variance-covariance matrix is replaced with the diagonal matrix to ensure it is a full-rank matrix.
@@
 \label{eq:kl-2}
-D_{KL} &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(diag(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
+D_{KL} &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
 \end{align}
 
-\noindent where $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$, and $\boldsymbol{W} = \boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'$.
-The assumed error variance $\sigma^2$ is set to be $tr(\boldsymbol{V})/n$, which is the expectation of the estimated variance.
+\noindent where $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$, and $\boldsymbol{W} = \boldsymbol{R}\boldsymbol{V}\boldsymbol{R}$.
+The assumed error variance $\sigma^2$ is set to be $\text{tr}(\boldsymbol{V})/n$, which is the expectation of the estimated variance.
 
@@
 
-In order to evaluate equation \ref{eq:kl-1} in a practically computable manner, the elements of $\boldsymbol{e}$ are assumed to be independent of each other.
+In order to approximate $D_{KL}$ in a practically computable manner, the elements of $\boldsymbol{e}$ are assumed to be independent of each other.
 This assumption solves both of the issues mentioned above.
@@
 Instead, the evaluation of Kullback-Leibler divergence for an individual residual relies on the knowledge of the marginal density $p_i(e_i)$, where $e_i$ is the $i$-th residual for $i = 1, ..., n$.
-This is much easier to estimate through simulation.
+This is much easier to approximate through simulation.
+It is also worth mentioning that this independence assumption generally will not hold, since $\text{cov}(e_i, e_j) \neq 0$ if $\boldsymbol{R}_{ij} \neq 0$ for any $1 \leq i < j \leq n$, but its existence is essential for reducing the computational cost.
 
-The algorithm for computing equation \ref{eq:kl-1} starts from simulating $m$ sets of $\boldsymbol{\varepsilon}$ according to the error distribution.
-The simulated errors are stored in a matrix $\boldsymbol{A}$ with $n$ rows and $m$ columns.
-So each column of $\boldsymbol{A}$ is a set of realization values of $\boldsymbol{\varepsilon}$.
-Then, we can get $m$ sets of $\boldsymbol{e}$ stored in the matrix $\boldsymbol{B}$ by applying the residual operator $\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}$.
+The algorithm for approximating equation \ref{eq:kl-1} starts from simulating $m$ sets of $\boldsymbol{\varepsilon}$ according to the error distribution.
+The simulated errors are stored in a matrix $\boldsymbol{A}$ with $n$ rows and $m$ columns, where each column of $\boldsymbol{A}$ is a set of realized values of $\boldsymbol{\varepsilon}$.
+Then, we can get $m$ sets of realized values of $\boldsymbol{e}$ stored in the matrix $\boldsymbol{B}$ by applying the residual operator $\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}$.
 Furthermore, kernel density estimation (KDE) with Gaussian kernel and optimal bandwidth selected by the Silverman's rule of thumb [@silverman2018density] is applied on each row of $B$ to estimate $p_i(e_i)$ for $i = 1, ..., n$.
@@
 Since the Kullback-Leibler divergence can be viewed as the expectation of the log-likelihood ratio between distribution $P$ and distribution $Q$ evaluated on distribution $P$, we can reuse the simulated residuals in matrix $\boldsymbol{B}$ to estimate the expectation by the sample mean.
-With the independence assumption, for non-normal $P$, $D_{KL}$ can be estimated by
+With the independence assumption, for non-normal $P$, $D_{KL}$ can be approximated by
 
@@
 \label{eq:kl-3}
-\widehat{D_{KL}} &= \sum_{i = 1}^{n} \widehat{D_{KL}^{(i)}}, \\
-\widehat{D_{KL}^{(i)}} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
+D_{KL} &\approx \sum_{i = 1}^{n} \hat{D}_{KL}^{(i)}, \\
+\hat{D}_{KL}^{(i)} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
 \end{align}
 
-\noindent where $\widehat{D_{KL}^{(i)}}$ is the estimator of the Kullback-Leibler divergence for an individual residual $e_i$, $B_{ij}$ is the $i$-th row and $j$-th column entry of the matrix $B$, $\hat{p_i}(.)$ is the kernel density estimator of $p_i(.)$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\widehat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)$, and $vec(.)$ is the vectorization operator which turns a $n \times m$ matrix into a $nm \times 1$ column vector by stacking the columns of the matrix on top of each other.
+\noindent where $\hat{D}_{KL}^{(i)}$ is the estimator of the Kullback-Leibler divergence for an individual residual $e_i$, $B_{ij}$ is the $i$-th row and $j$-th column entry of the matrix $B$, $\hat{p_i}(.)$ is the kernel density estimator of $p_i(.)$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\widehat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)$, and $vec(.)$ is the vectorization operator which turns a $n \times m$ matrix into a $nm \times 1$ column vector by stacking the columns of the matrix on top of each other.
 
@@
 \label{eq:d-approx}
-\hat{D} = f(V(\hat{\boldsymbol{e}}, \hat{\boldsymbol{y}})),
+\hat{D} = f_{CV}(V_{h \times w}(\boldsymbol{e}, \hat{\boldsymbol{y}})),
 \end{equation}
 
-\noindent where $\hat{\boldsymbol{e}}$ is a vector of residuals obtained from the fitted model rather than a vector of random variables, $\hat{\boldsymbol{y}}$ is the fitted values also obtained from the fitted model, $V(.)$ is a plotting function that generates a residuals vs fitted values plot with fixed aesthetic and then saves it as an image with $h \times w$ pixels and three colour channels, and $f(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance in the domain $[0, +\infty)$.
+\noindent where $V_{h \times w}(.)$ is a plotting function that saves a residuals vs fitted values plot with fixed aesthetic as an image with $h \times w$ pixels and three colour channels, $f_{CV}(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance in the domain $[0, +\infty)$.
 
@@
 We would not expect the model to be able to magically tell if the suspicious pattern is caused by the skewed distribution of the fitted values or the existence of heteroskedasticity.
-Assume we only train the model with this single image, while $50$% of the time the label is $0$, and $50$% of the time the label is some constant $c > 0$.
-Then, a perfect model will predict $0 < \hat{D} < c$ to achieve a minimum loss.
-Some null plots could have outliers or strong visual patterns due to randomness, and a reasonable model will try to summarise those information into the prediction, resulting in $\hat{D} > 0$.
+Additionally, some null plots could have outliers or strong visual patterns due to randomness, and a reasonable model will try to summarise those information into the prediction, resulting in $\hat{D} > 0$.
 
 This property is not an issue if $\hat{D} \gg 0$ for which the visual signal of the residual plot is very strong, and we usually do not need any further examination of the significance of the result.
-However, if the visual pattern is weak or moderate, having $\hat{D}$ will not be sufficient to determine if the null hypothesis that the model is correctly specified should be rejected.
+However, if the visual pattern is weak or moderate, having $\hat{D}$ will not be sufficient to determine if $H_0$ that the model is correctly specified should be rejected.
 
-To solve this issue while aligning with the principle of visual inference, the approximated distance $\hat{D}$ can be defined as a test statistic.
-And the null distribution of this statistic can be approximated by the empirical distribution of $\widehat{D_{null}}$ for a large amount of null plots of a fitted model.
+To solve this issue while aligning with the principle of visual inference, the approximated distance $\hat{D}$ can be viewed as a test statistic.
+And the null distribution of this statistic can be approximated by the empirical distribution of $\hat{D}_{null}$ for a large amount of null plots of a fitted model.
 The procedure involves applying the residual rotation technique [@buja2009statistical] on the fitted model to obtain null residuals.
-The null residuals are then used to make null plots and fed into the model.
-The predicted distance are used to construct an empirical distribution.
+The null residuals are then used to make null plots and fed into the model to get predictions.
+The empirical distribution is constructed with the predicted distance.
+
+There are two types of error associated with the approximation of the null distribution.
+These include the sampling error and the estimation error for model parameters.
+Increasing the number of null plots within the distribution approximation can effectively reduce sampling error.
+However, the estimation error for model parameters remains irreducible.
+The residual rotation technique operates under the assumption of a correct fitted model, using the estimated variance instead of the unknown true variance to rescale rotated residuals.
+Consequently, under $H_0$, the true distribution of $\hat{D}_{null}$ will slightly deviate from the approximated null distribution.
 
@@
 
-Let $\widehat{D_{null}^{(i)}}$ be the predicted distance of the $i$-th null plots, where $i = 1,...,m$ and $m \in \mathbb{N}^+$ is a sufficiently large number.
+Let $\hat{D}_{null}^{(i)}$ be the predicted distance of the $i$-th null plots, where $i = 1,...,n_{null}$ and $n_{null} \in \mathbb{N}^+$ is a sufficiently large number.
 Quantiles of the null distribution can then be estimated using the sample quantiles available in statistical software such as R.
 The details of the sample quantile computation can be found in @hyndman1996sample.
-In statistical testing, analysts often care about certain quantiles of the null distribution, such as 90% quantile, 95% quantile and 99% quantile.
+In statistical testing, analysts often care about certain quantiles of the null distribution, such as the 90% quantile, the 95% quantile and the 99% quantile.
 These quantiles are used as thresholds to decide if $H_0$ needs to be rejected.
-For example, if $\hat{D}$ is greater than and equal to the 95% sample quantile of the null distribution, we could say the predicted distance for the actual residual plot is significantly different from the predicted distance for null plots with 95% significance level.
-Based on our experience, in order to get a stable estimate of the 95% quantile, $m$ usually needs to be at least $100$.
+For example, if $\hat{D}$ is greater than and equal to the 95% sample quantile of the null distribution $Q_{null}(0.95)$, we could say the predicted distance for the actual residual plot is significantly different from the predicted distance for null plots with 95% significance level.
+Based on our experience, in order to get a stable estimate of the 95% quantile, $n_{null}$ usually needs to be at least $100$.
 And if the null distribution has a long tail, more null plots will be needed.
-Alternatively, a p-value can be used to represents the probability of observing an event equally or more extreme than the given event under $H_0$, and it can be estimated by $1/m\sum_{i=1}^{m}I\left(\widehat{D_{null}^{(i)}} \geq \hat{D}\right)$.
+Alternatively, a p-value can be used to represents the probability of observing an event equally or more extreme than the given event under $H_0$, and it can be estimated by $1/m\sum_{i=1}^{n_{null}}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)$.
 
@@
 
-- What is the assumption of bootstrapping
-- Any drawback: loss of information
-- For each bootstrapped fitted model, we can get a bootstrapped vss and construct a new 95% quantile for the null distribution, and check if H_0 should be rejected
-- However, it is very expensive to construct a new 95% quantile for each bootstrapped fitted model
-- The 95% quantile constructed using the original fitted model can be considered as an estimate of those new 95% quantiles
-- We borrow information from the original fitted model
-- We can check how many bootstrapped fitted model should be rejected (like the percentage)
-- This gives an overall estimate of how often the assumed regression model are considered to be incorrect if the data can be obtained repetitively from the same data generating process
-- If this percentage is relatively large, it gives the analyst an warning that the regression model is inappropriate to the data
+Bootstrap is often employed in linear regression when conducting inference for estimated parameters.
+It is typically done by sampling individual cases with replacement and refitting the regression model.
+If the observed data accurately reflects the true distribution of the population, the bootstrapped estimates can be used to measure the variability of the parameter estimate without making strong distributional assumptions about the data generating process.
+
+Similarly, bootstrap can be applied on the approximated distance $\hat{D}$.
+For each refitted model $M_{boot}^{(i)}$, there will be an associated residual plot $V_{boot}^{(i)}$ which can be fed into the computer vision model to obtain $\hat{D}_{boot}^{(i)}$, where $i = 1,...,n_{boot}$, and $n_{boot}$ is the number of bootstrapped samples.
+If we are interested in the variation of $\hat{D}$, we can use $\hat{D}_{boot}^{(i)}$ to estimate a confidence interval.
+
+Alternatively, since each $M_{boot}^{(i)}$ has a set of estimated coefficients $\hat{\boldsymbol{\beta}}_{boot}^{(i)}$ and an estimated variance $\hat{\sigma^2}_{boot}^{(i)}$, a new approximated null distribution can be construed and the corresponding 95% sample quantile $Q_{boot}^{(i)}(0.95)$ can be computed.
+Then, if $\hat{D}_{boot}^{(i)} \geq Q_{boot}^{(i)}(0.95)$, $H_0$ will be rejected for $M_{boot}^{(i)}$.
+The ratio of rejected $M_{boot}^{(i)}$ among all the refitted models provides an indication of how often the assumed regression model are considered to be incorrect if the data can be obtained repetitively from the same data generating process.
+But this approach is computationally very expensive because it requires $n_{boot} \times n_{null}$ times of residual plot assessment.
+In practice, $Q_{null}(0.95)$ will be used to replace $Q_{boot}^{(i)}(0.95)$ in the computation.
 
@@
 
-<!-- The computer vision model is trained with $n_{train}$ pairs of residual plots and distance $D$, and the trainable parameters $\boldsymbol{\theta}$ are optimized using the mean square error loss function written as -->
-
-<!-- $$\hat{\boldsymbol{\theta}} = \underset{\boldsymbol{\theta}}{argmin}\frac{1}{n_{train}}\sum_{i=1}^{n_{train}}(D_i - vss_{\boldsymbol{\theta}}(plot_i))^2,$$ -->
-<!-- \noindent where $plot_{i}$ is the i$th$ residual plot, and $D_i$ is the computed distance for the underlying residual distribution of the $i$th residual plot.
-The detailed training process is described in Section ....
--->
-
 We have incorporated three types of residual departures of linear regression model in the training data, including non-linearity, heteroskedasticity and non-normality.
@@
 \boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 + \beta_2(\boldsymbol{z} + \beta_1\boldsymbol{w}) + \boldsymbol{k} \odot \boldsymbol{\varepsilon}, \\
-\boldsymbol{z} &= He_j(g(\boldsymbol{x}_1, 2)), \\
-\boldsymbol{w} &= He_j(g(\boldsymbol{x}_2, 2)), \\
+\boldsymbol{z} &= \text{He}_j(g(\boldsymbol{x}_1, 2)), \\
+\boldsymbol{w} &= \text{He}_j(g(\boldsymbol{x}_2, 2)), \\
 \boldsymbol{k} &= \sqrt{\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^2},
@@
 
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{z}$, $\boldsymbol{w}$, $\boldsymbol{k}$ and $\boldsymbol{\varepsilon}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$ are two independent predictors, $He_j(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau], the $\sqrt{(.)}$ and $(.)^2$ operators are element-wise operators, $\odot$ is the Hadamard product, and $g(., k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{z}$, $\boldsymbol{w}$, $\boldsymbol{k}$ and $\boldsymbol{\varepsilon}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$ are two independent predictors, $\text{He}_j(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau], the $\sqrt{(.)}$ and $(.)^2$ operators are element-wise operators, $\odot$ is the Hadamard product, and $g(., k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
 
@@
 
-The model is trained on a high-performance computing platform owned by Monash university with TensorFlow (ref here) and Keras (ref here).
+The model is trained on the M3 high-performance computing platform (ref here) with TensorFlow (ref here) and Keras (ref here).
 During the training process, 80% of the training data are used as the actual training data, and 20% of the training data are reserved as validation data.
@@
 
-## Best hyperparameters
+## Optimized hyperparameters
 
-Based on the tuning process described in Section \ref{training-and-hyperparameter-tuning}, the best hyperparameter values are provided in Table \ref{tab:best-hyperparameter}.
+Based on the tuning process described in Section \ref{training-and-hyperparameter-tuning}, the optimized hyperparameter values are provided in Table \ref{tab:best-hyperparameter}.
 It can be observed that at least $32$ base filters are needed, and the preferable choice for the $64 \times 64$ model and the $128 \times 128$ model is $64$ base filters, which is the same as the original VGG16 architecture.
-The best values for dropout rate for convolutional blocks is around $0.4$.
+The optimized values for dropout rate for convolutional blocks is around $0.4$.
 And applying batch normalization for convolutional blocks has a positive impact on the performance.
-All best models choose to preserve the additional inputs to reduce the validation error.
+All optimized models choose to preserve the additional inputs to reduce the validation error.
 The number of units needed for the fully-connected layer is $256$, which is not a very large number comparing to the VGG16 classifier, suggesting that the problem we are trying to solve is relatively easy.
-The best learning rates are usually smaller than the default value recommended by Keras which is 0.001.
+The optimized learning rates are usually smaller than the default value recommended by Keras which is 0.001.
 
@@
                     booktabs = TRUE, 
-                    caption = "Hyperparameters values for the best computer vision models with differnet input sizes.", 
+                    caption = "Hyperparameters values for the optimized computer vision models with different input sizes.", 
                     col.names = c("Hyperparameter", "$32 \\times 32$", "$64 \\times 64$", "$128 \\times 128$"),
@@
 
-<!-- - Metrics for model performance -->
-<!-- - Shap values -->
-<!-- - Heatmap -->
-
-The training and test performance of models with three different input sizes are provided in Table \ref{tab:performance}.
-Among these models, the $64 \times 64$ model and the $32 \times 32$ model are consistently having the best metrics on the training set and the test set respectively.
+The training and test performance of optimized models with three different input sizes are provided in Table \ref{tab:performance}.
+Among these models, the $64 \times 64$ model and the $32 \times 32$ model are consistently having the best training and test performance respectively.
 The mean absolute error shows that the difference between $\hat{D}$ and $D$ is around $0.43$ on the test set, which is very small considering the range of $D$ is normally between $0$ and $7$.
@@
 
-Figure \ref{fig:model-performance} displays a scatter plot for $D - \hat{D}$ vs $D$.
-For residual plots with $D$ close to zero, all the models tends to over-predict $\hat{D}$, leading to negative residuals.
+Figure \ref{fig:model-performance} displays a hexagonal heatmap for $D - \hat{D}$ vs $D$.
+For residual plots with $D$ close to zero, all the optimized models tends to over-predict $\hat{D}$, leading to negative residuals.
 When $D$ is between $1.5$ and $6$, the smooth curves indicate that $\hat{D}$ is not heavily biased, and there are some instances where $\hat{D}$ is severely under-predicted.
 And when $D$ is greater than $6$, there are more instances being under-predicted than over-predicted.
-It can be observed from Figure \ref{fig:model-performance-density} that the models produced much more $\hat{D}$ at around $1$ than expected.
+
+<!-- It can be observed from Figure \ref{fig:model-performance-density} that the optimized models produced much more $\hat{D}$ at around $1$ than expected.  -->
 
@@
                     col.names = c("", "RMSE", "$R^2$", "MAE", "Huber loss"),
-                    caption = "The training and test performance of three models with different input sizes.") %>%
+                    caption = "The training and test performance of three optimized models with different input sizes.") %>%
   kableExtra::pack_rows("Training set", 1, 3) %>%
@@
 
-```{r model-performance, fig.cap = "Residuals vs distance plot on training and test data for three best models with different input sizes.
+```{r model-performance, fig.cap = "Hexagonal heatmap for residuals vs distance on training and test data for three optimized models with different input sizes.
 The blue lines are smoothing curves produced by fitting gnealized additive models.", dev = 'png', dpi = 300}
@@
   geom_hline(yintercept = 0) +
-  geom_point(aes(effect_size, effect_size - vss), alpha = 0.03) +
+  # geom_point(aes(effect_size, effect_size - vss), alpha = 0.03) +
+  geom_hex(aes(effect_size, effect_size - vss), bins = 15) +
   geom_smooth(aes(effect_size, effect_size - vss), se = FALSE) +
@@
   xlab("D") +
-  theme_light()
+  theme_light() +
+  scale_fill_continuous(limits = c(0, 400))
 ```
 
-```{r model-performance-density, fig.cap = "Density plot of predicted distance on training and test data for three best models with different input sizes.
-The density drawn in red is the density of the distance."}
+```{r model-performance-density, fig.cap = "Density plot of predicted distance on training and test data for three optimized models with different input sizes.
+The density drawn in red is the density of the distance.", eval = FALSE}
 train_pred %>% 
@@
 
-- basic background
-- highlights the same and the different settings
+In order to check the validity of the proposed computer vision model, residual plots presented in the human subject experiment conducted by @li2023plot will be assessed.
+This study has collected $7974$ human responses to $1152$ lineups.
+Each lineup contains one randomly placed actual residual plot and 19 null plots.
+Among the $1152$ lineups, $24$ are attention check lineups in which the visual patterns are designed to be extremely obvious and very different from the corresponding to null plots, $36$ are null lineups where all the lineups consist of only null plots, $279$ are lineups with uniform predictor distribution evaluated by $11$ participants, and the remaining $813$ are lineups with discrete, skewed or normal predictor distribution evaluated by $5$ participants.
+Attention check lineups and null lineups will not be assessed in the following analysis.
 
-### Comparison
+In @li2023plot, the residual plots are simulated from a data generating process which is a special case of Equation \ref{eq:data-sim}.
+The main characteristic is the model violations are introduced separately, meaning non-linearity and heteroskedasticity will not co-exist in one lineup but assigned uniformly to all lineups.
+Additionally, non-normality and multiple predictors are not considered in their experimental design.
 
-<!-- - power comparison -->
-<!-- - decisions -->
+### Model performance on the human data
 
-```{r fig.cap = "Residuals vs distance plot on data used in the human subject experiment. The blue line is a smoothing curve produced by fitting a gnealized additive model."}
+For each lineup used in @li2023plot, there are one actual residual plot and $19$ null plots.
+While the distance $D$ for the actual residual plot depends on the underlying data generating process, the distance $D$ for the null plots are zeros.
+We have used our optimized computer vision model to predict $\hat{D}$ for both the actual residual plots and the null plots.
+
+The performance metrics of $\hat{D}$ for actual residual plots are provided in Table \ref{tab:experiment-performance}.
+It can be observed that all the performance metrics are slightly worse than the metrics evaluated on the test data.
+Still, the mean absolute error remains at a low level, and the linearly correlation between the prediction and the true value is still very high.
+Lineups with non-linearity issues are more difficult to be predicted than those with heteroskedasticity issues.
+
+Figure \ref{fig:rd-human} displays the $D - \hat{D}$ vs $D$ plot for actual residual plots used in the experiment.
+The behaviour of the prediction is roughly the same as what we can be observed from the evaluation of the test data.
+The model over-predicts $\hat{D}$ when $D$ is close to zero, and under-predicts $\hat{D}$ when $D$ is greater than $6$.
+However, for residual plots with non-linearity issues, we do not observe the over-prediction for $D > 6$ because there are not many cases in that area.
+And the over-prediction happens at around $2 < D < 4$.
+
+For null plots, a histogram of the prediction is provided in Figure \ref{fig:hist-null-human}.
+In most cases, $\hat{D}$ will be predicted at around $1$.
+And as $\hat{D}$ increases, the number of predictions decreases.
+There are some outliers with $\hat{D} > 5$.
+
+```{r rd-human, fig.cap = "Residuals vs distance plot on lineups used in the human subject experiment. The blue line is a smoothing curve produced by fitting a gnealized additive model."}
 vss_32 <- readRDS(here::here("paper/data/vss_32.rds"))
@@
 experiment %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
   ggplot() +
   geom_hline(yintercept = 0) +
-  geom_point(aes(actual_ss, actual_ss - vss)) +
+  geom_point(aes(actual_ss, actual_ss - vss), alpha = 0.3) +
   geom_smooth(aes(actual_ss, actual_ss - vss), se = FALSE) +
@@
   xlab("D") +
-  theme_light()
+  theme_light() +
+  facet_wrap(~type)
 ```
@@
 experiment %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
   mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
@@
                     escape = FALSE,
-                    col.names = c("Type", "RMSE", "$R^2$", "MAE", "Huber loss"),
+                    col.names = c("Voilation", "RMSE", "$R^2$", "MAE", "Huber loss"),
                     caption = "The performance of the $32 \\times 32$ model on the data used in the human subject experiment.")
@@
 
-It can be observed from Table \ref{tab:experiment-performance} that all the performance metrics are lower than the metrics evaluated on the test data.
+```{r rd-uniform-human, fig.cap = "Residuals vs distance plot on lineups with uniform predictor distribution used in the human subject experiment.
+The blue line is a smoothing curve produced by fitting a gnealized additive model.", eval = FALSE}
+experiment %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  filter(x_dist == "uniform") %>%
+  ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(actual_ss, actual_ss - vss), alpha = 0.6) +
+  geom_smooth(aes(actual_ss, actual_ss - vss), se = FALSE) +
+  coord_fixed() +
+  ylab(expression(D - hat(D))) +
+  xlab("D") +
+  theme_light() +
+  facet_wrap(~type)
+```
 
-```{r fig.cap = "A weighted detection rate vs $\\delta$-differnence plot."}
+```{r hist-null-human, fig.cap = "Histogram of predicted distance on null plots used in the human subject experiment."}
 lineup_vss <- readRDS(here::here("paper/data/lineup_vss.rds"))
@@
   mutate(gamma_diff = sum(vss[null] > vss[!null]))
+
+lineup_vss %>%
+  filter(null) %>%
+  left_join(vi_survey %>%
+              group_by(unique_lineup_id) %>% 
+              summarise(type = first(type), 
+                        attention_check = first(attention_check),
+                        null_lineup = first(null_lineup))) %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  ggplot() +
+  geom_histogram(aes(vss)) +
+  xlab(expression(hat(D))) +
+  theme_light() +
+  facet_wrap(~type)
+```
+
+```{r cache = TRUE}
+vi_lineup <- get_vi_lineup()
+```
+
+```{r fig.height = 7/4*2, fig.width = 7/5*3}
+lineup_vss %>%
+  filter(null) %>%
+  left_join(vi_survey %>%
+              group_by(unique_lineup_id) %>% 
+              summarise(type = first(type), 
+                        attention_check = first(attention_check),
+                        null_lineup = first(null_lineup))) %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  filter(vss > 5) -> human_null_outliers
+
+vi_lineup[human_null_outliers$unique_lineup_id] %>%
+  map2_df(human_null_outliers$pos, 
+          ~.x$data %>% 
+            filter(k == .y) %>% 
+            mutate(unique_lineup_id = .x$metadata$name)) %>%
+  ggplot() +
+  geom_hline(yintercept = 0, col = "red") +
+  geom_point(aes(.fitted, .resid), size = 0.5, stroke = 0.5) +
+  theme_light(base_size = 11/5) +
+  theme(axis.ticks = element_blank(), 
+        axis.text.x = element_blank(),
+        axis.text.y = element_blank(), 
+        axis.title.x = element_blank(), 
+        axis.title.y = element_blank(),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank()) +
+  facet_wrap(~unique_lineup_id, scales = "free")
+
+```
+
+```{r fig.cap = "A weighted detection rate vs $\\delta$-differnence plot."}
 

### 2024-02-05 paper/paper.tex
update the paper

--- 2024-01-29 paper/paper.tex
@@
 
-\title{Automated reading of residual plots with computer vision models}
+\title{Automated assessment of residual plots with computer vision
+models}
 
@@
 
-The practice of plotting residuals is commonly regarded as a standard
-procedure in linear regression diagnostics
+Plotting residuals is commonly regarded as a standard practice in linear
+regression diagnostics
 \citep[see][]{cook1982residuals, belsley1980regression}. This visual
@@
 suggest heteroskedasticity, it is important to avoid over-interpreting
-the visual pattern. In this case, the fitted model is correctly
+this visual pattern. In this case, the fitted model is correctly
 specified, and the triangular shape is actually a result of the skewed
@@
 the lineup protocol lies in its reliance on human judgments. Unlike
-conventional statistical tests that can be performed numerically and
-automatically in statistical software, the lineup protocol requires
-human evaluation of images. This characteristic makes it less suitable
-for large-scale applications, given the associated high labor costs and
-time requirements.
+conventional statistical tests that can be performed computationally in
+statistical software, the lineup protocol requires human evaluation of
+images. This characteristic makes it less suitable for large-scale
+applications, given the associated high labor costs and time
+requirements.
 
 There is a substantial need to develop an approach that alleviates
-people's workload by automating repetitive tasks and providing
+analysts' workload by automating repetitive tasks and providing
 standardized results in a controlled environment. The large-scale
@@
 
-In addressing the residual reading problem, there exist multiple
-configurations of the computer vision model formula. These
-configurations can be categorized by two key components of the model
-formula: the input and the output format.
+There are various configurations of the computer vision model that can
+be used to assess residual plots. We discuss these configurations below
+based on two key components of the model formula: the input and the
+output format.
 
@@
 
-In computer vision models, the quality and relevance of the input data
-greatly influence the model's capacity to generate insightful and
-meaningful results. The input of the model can be designed in different
-formats.
-
-A straightforward approach involves feeding the model a vector of
-residuals along with a vector of fitted values, essentially providing
-all the necessary information for creating a residuals vs fitted values
-plot. However, a drawback of this method is the dynamic input size, as
-different fitted regression models may have varying numbers of
-observations. For modern computer vision models implemented with
-mainstream software like TensorFlow \citep{abadi2016tensorflow}, the
-input shape is typically fixed. One solution is to pad the input vectors
-with leading or trailing zeros if the input tensor expects longer
-vectors, but it may fail if the input vector surpasses the designed
-length. Another strategy is to summarize the residuals and fitted values
-separately using histograms and utilize the counts as the input. By
-controlling the number of bins in the histograms, it becomes possible to
-provide fixed-length input vectors.
+Deep learning models are in general very sensitive to the input data.
+The quality and relevance of the input data greatly influence the
+model's capacity to generate insightful and meaningful results.
+
+A straightforward approach involves feeding a vector of residuals along
+with a vector of fitted values, essentially providing all the necessary
+information for creating a residuals vs fitted values plot. However, a
+drawback of this method is the dynamic input size, which changes based
+on the number of observations. For modern computer vision models
+implemented in mainstream software like TensorFlow
+\citep{abadi2016tensorflow}, the input shape is typically fixed. One
+solution is to pad the input vectors with leading or trailing zeros when
+the input tensor expects longer vectors, but it may fail if the input
+vector surpasses the designed length. Another strategy is to summarize
+the residuals and fitted values separately using histograms and utilize
+the counts as the input. By controlling the number of bins in the
+histograms, it becomes possible to provide fixed-length input vectors.
+Still, since the histogram only captures the marginal distribution of
+residuals and fitted values respectively, it can not be used to
+differentiate visual patterns with same marginal distributions but
+different joint distributions.
 
@@
 
-Other possible input formats include, but are not limited to, a pair of
-residual plots, a triplet, and a lineup. \citet{chopra2005learning} has
-shown that computer vision models designed for image comparison can
-assess whether a pair of images are similar or dissimilar. Applied to
-our specific problem, we can define null plots of a fitted model to be
-similar to each other, while considering actual residual plots to be
-distinct from null plots of any fitted model. A triplet constitutes a
-set of three images, denoted as \(image_1\), \(image_2\) and
-\(image_3\). It is often used to predict whether \(image_2\) or
-\(image_3\) is more similar to \(image_1\), proving particularly useful
-for establishing rankings between samples. However, it's important to
-note that these two approaches usually require additional considerations
-regarding the loss function and, at times, non-standard training
-processes due to shared weights between different convolutional blocks.
+Providing multiple residual plots to the model, such as a pair of plots,
+a triplet or a lineup is also a possible option.
+\citet{chopra2005learning} has shown that computer vision models
+designed for image comparison can assess whether a pair of images are
+similar or dissimilar. Applied to our specific problem, we can define
+null plots of a fitted model to be similar to each other, while
+considering actual residual plots to be distinct from null plots of any
+fitted model. A triplet constitutes a set of three images, denoted as
+\(image_1\), \(image_2\) and \(image_3\). It is often used to predict
+whether \(image_2\) or \(image_3\) is more similar to \(image_1\),
+proving particularly useful for establishing rankings between samples.
+For this setup, we can apply the same criteria to define similarity
+between images. However, it's important to note that these two
+approaches usually require additional considerations regarding the loss
+function and, at times, non-standard training processes due to shared
+weights between different convolutional blocks.
 
@@
 
-Not all the mentioned input formats were explored and tested in this
-study due to the considerable costs associated with data preparation and
-model training. The single residual plot input format was chosen for its
-relatively low implementation cost and high interpretability.
+We did not explore all the mentioned input formats due the considerable
+costs associated with data preparation and model training. Considering
+the implementation cost and the interpretability of the model, we
+settled on the single residual plot input format.
 
@@
 Given that the input is a single residual plot represented as a
-fixed-resolution image, the computer vision model's output can take one
-of two formats: binary or numeric. This choice determines whether the
-model belongs to a classification model or a regression model. The
-binary outcome might indicates whether the input image is a null plot or
-not, or whether the input image would be rejected in a visual test
-conducted by humans. The latter option requires data from prior human
-subject experiments, presenting difficulties in controlling the quality
-of data due to variations in experimental settings across different
-studies. Additionally, some visual inference experiments are unrelated
-to linear regression models and residual plots, resulting in a limited
-amount of available training data.
-
-Alternatively, the output could be a meaningful numeric measure, such as
-the average distance to good residual plots. This approach necessitates
-defining a distance measure between residual plots, which may be a
-non-trivial task. However, once a meaningful distance measure is
-established, the computer vision model's prediction becomes an
-interpretable value with tangible significance. Studies have
-demonstrated that defining a proper distance between images can enhance
-the matching accuracy in image search compared to a binary outcome model
-(ref here).
+fixed-resolution image, the output from the computer vision model can
+take one of two formats: binary or numeric. This choice determines
+whether the model belongs to a classification model or a regression
+model. The binary outcome could be designed to indicate whether the
+input image is a null plot, or whether the input image would be rejected
+in a visual test conducted by humans. The latter option requires data
+from prior human subject experiments, presenting difficulties in
+controlling the quality of data due to variations in experimental
+settings across different studies. Additionally, some visual inference
+experiments are unrelated to linear regression models and residual plot
+diagnostics, resulting in a limited amount of available training data.
+
+Alternatively, the output could be designed as a meaningful numeric
+measure, such as the average distance to good residual plots. This
+approach necessitates defining a distance measure between residual
+plots, which may be a non-trivial task. However, once a meaningful
+distance measure is established, the prediction from the computer vision
+model becomes an interpretable value with tangible significance. Studies
+have demonstrated that defining a proper distance between images can
+enhance the matching accuracy in image search compared to a binary
+outcome model (ref here).
 
@@
 that stand out as most distinct from others in a given lineup. To
-develop a computer vision model for evaluating residual plots within the
+develop a computer vision model for assessing residual plots within the
 visual inference framework, it is important to precisely define a
@@
 from a correctly specified model. For the classical normal linear
-regression model, residuals are derived from the fitted values
-\(\hat{\boldsymbol{y}}\) and observed values \(\boldsymbol{y}\). Suppose
-the data generating process is known and the model is correctly
-specified, by the Frisch-Waugh-Lowell theorem \citep{frisch1933partial},
-residuals \(\boldsymbol{e}\) can also be written as a linear
-transformation of the error \(\boldsymbol{\varepsilon}\) formulated as
+regression model, residuals \(\boldsymbol{e}\) are derived from the
+fitted values \(\hat{\boldsymbol{y}}\) and observed values
+\(\boldsymbol{y}\). Suppose the data generating process is known and the
+regression model is correctly specified, by the Frisch-Waugh-Lowell
+theorem \citep{frisch1933partial}, residuals \(\boldsymbol{e}\) can also
+be treated as random variables and written as a linear transformation of
+the error \(\boldsymbol{\varepsilon}\) formulated as
 \(\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon}\), where
 \(\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\)
-is the residual operator, \(\boldsymbol{X}\) is the design matrix,
-\(\boldsymbol{I}_n\) is a \(n\) by \(n\) identity matrix, and \(n\) is
-the number of observations.
+is the residual operator and an idempotent matrix, \(\boldsymbol{X}\) is
+the design matrix, \(\boldsymbol{I}_n\) is a \(n\) by \(n\) identity
+matrix, and \(n\) is the number of observations.
 
@@
 
-When fitting a linear regression model, the solver will force
-\(\sum_{i=1}^{n} e_i = 0\), making any residual value to be a linear
-combination of the remaining \(n - 1\) residuals. This effectively means
-\(rank(\boldsymbol{R}) = n - 1 < n\) and \(Q\) becomes a degenerate
-multivariate distribution. To capture the characteristics of \(Q\), such
-as moments, we can simulate a large numbers of
-\(\boldsymbol{\varepsilon}\) and transform it to \(\boldsymbol{e}\) to
-get the empirical estimates. For simplicity, we replaced
-\(\boldsymbol{R}\) with a full-rank diagonal matrix
-\(diag(\boldsymbol{R})\), where \(diag(.)\) set the non-diagonal entries
-of a matrix to zeros. The resulting distribution for \(\boldsymbol{Q}\)
-is \(N(\boldsymbol{0}_n, diag(\boldsymbol{R}\sigma^2))\).
+In ordinary least square, the minimization of the sum of square
+residuals implies \(\sum_{i=1}^{n} e_i = 0\), making any residual value
+to be a linear combination of the remaining \(n - 1\) residuals. This
+effectively means \(\text{rank}(\boldsymbol{R}) = n - 1 < n\) and \(Q\)
+is a degenerate multivariate distribution. To capture the
+characteristics of \(Q\), such as moments, we can simulate a large
+numbers of \(\boldsymbol{\varepsilon}\) and transform it to
+\(\boldsymbol{e}\) to get the empirical estimates. For simplicity, we
+replaced the variance-covariance matrix of residuals
+\(\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}' = \boldsymbol{R}\sigma^2\)
+with a full-rank diagonal matrix
+\(\text{diag}(\boldsymbol{R}\sigma^2)\), where \(\text{diag}(.)\) sets
+the non-diagonal entries of a matrix to zeros. The resulting
+distribution for \(\boldsymbol{Q}\) is
+\(N(\boldsymbol{0}_n, \text{diag}(\boldsymbol{R}\sigma^2))\).
 
@@
 However, if the model is misspecified, then the actual distribution of
-residuals denoted as \(P\), will be different to \(Q\). For example, if
-the data generating process contains variables correlated with any
+residuals denoted as \(P\), will be different from \(Q\). For example,
+if the data generating process contains variables correlated with any
 column of \(\boldsymbol{X}\) but not included in \(\boldsymbol{X}\),
-causing an omitted variable problem, \(P\) will be different to \(Q\)
+causing an omitted variable problem, \(P\) will be different from \(Q\)
 because the residual operator obtained from the fitted model will not be
@@
 \(\boldsymbol{\varepsilon}\) follows a non-normal distribution such as a
-multivariate lognormal distribution, the empirical residual distribution
-will usually be skewed and has a long tail.
+multivariate lognormal distribution, \(P\) will usually be skewed and
+has a long tail.
 
@@
 while in fact
-\(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})\),
-\(Q\) can be represented as
-\(N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, diag(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'))\).
+\(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})\)
+with \(\boldsymbol{V}\) being an arbitrary symmetric positive
+semi-definite matrix, \(Q\) can be represented as
+\(N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}))\).
 Note that the variance-covariance matrix is replaced with the diagonal
@@
 \label{eq:kl-2}
-D_{KL} &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(diag(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
+D_{KL} &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
 \end{align}
@@
 \(\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z\),
-and \(\boldsymbol{W} = \boldsymbol{R}\boldsymbol{V}\boldsymbol{R}'\).
-The assumed error variance \(\sigma^2\) is set to be
-\(tr(\boldsymbol{V})/n\), which is the expectation of the estimated
-variance.
+and \(\boldsymbol{W} = \boldsymbol{R}\boldsymbol{V}\boldsymbol{R}\). The
+assumed error variance \(\sigma^2\) is set to be
+\(\text{tr}(\boldsymbol{V})/n\), which is the expectation of the
+estimated variance.
 
@@
 
-In order to evaluate equation \ref{eq:kl-1} in a practically computable
-manner, the elements of \(\boldsymbol{e}\) are assumed to be independent
-of each other. This assumption solves both of the issues mentioned
-above. First, we no longer need to integrate over \(n\) random
-variables. The result of equation \ref{eq:kl-1} is now the sum of the
-Kullback-Leibler divergence evaluated for each individual residual
-thanks for the independence assumption. Second, it is not required to
-know the joint probability density \(p(\boldsymbol{e})\) any more.
-Instead, the evaluation of Kullback-Leibler divergence for an individual
-residual relies on the knowledge of the marginal density \(p_i(e_i)\),
-where \(e_i\) is the \(i\)-th residual for \(i = 1, ..., n\). This is
-much easier to estimate through simulation.
-
-The algorithm for computing equation \ref{eq:kl-1} starts from
+In order to approximate \(D_{KL}\) in a practically computable manner,
+the elements of \(\boldsymbol{e}\) are assumed to be independent of each
+other. This assumption solves both of the issues mentioned above. First,
+we no longer need to integrate over \(n\) random variables. The result
+of equation \ref{eq:kl-1} is now the sum of the Kullback-Leibler
+divergence evaluated for each individual residual thanks for the
+independence assumption. Second, it is not required to know the joint
+probability density \(p(\boldsymbol{e})\) any more. Instead, the
+evaluation of Kullback-Leibler divergence for an individual residual
+relies on the knowledge of the marginal density \(p_i(e_i)\), where
+\(e_i\) is the \(i\)-th residual for \(i = 1, ..., n\). This is much
+easier to approximate through simulation. It is also worth mentioning
+that this independence assumption generally will not hold, since
+\(\text{cov}(e_i, e_j) \neq 0\) if \(\boldsymbol{R}_{ij} \neq 0\) for
+any \(1 \leq i < j \leq n\), but its existence is essential for reducing
+the computational cost.
+
+The algorithm for approximating equation \ref{eq:kl-1} starts from
 simulating \(m\) sets of \(\boldsymbol{\varepsilon}\) according to the
 error distribution. The simulated errors are stored in a matrix
-\(\boldsymbol{A}\) with \(n\) rows and \(m\) columns. So each column of
-\(\boldsymbol{A}\) is a set of realization values of
-\(\boldsymbol{\varepsilon}\). Then, we can get \(m\) sets of
-\(\boldsymbol{e}\) stored in the matrix \(\boldsymbol{B}\) by applying
-the residual operator \(\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}\).
-Furthermore, kernel density estimation (KDE) with Gaussian kernel and
-optimal bandwidth selected by the Silverman's rule of thumb
-\citep{silverman2018density} is applied on each row of \(B\) to estimate
-\(p_i(e_i)\) for \(i = 1, ..., n\). The KDE computation is done by the
-\texttt{density} function in R.
+\(\boldsymbol{A}\) with \(n\) rows and \(m\) columns, where each column
+of \(\boldsymbol{A}\) is a set of realized values of
+\(\boldsymbol{\varepsilon}\). Then, we can get \(m\) sets of realized
+values of \(\boldsymbol{e}\) stored in the matrix \(\boldsymbol{B}\) by
+applying the residual operator
+\(\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}\). Furthermore, kernel
+density estimation (KDE) with Gaussian kernel and optimal bandwidth
+selected by the Silverman's rule of thumb \citep{silverman2018density}
+is applied on each row of \(B\) to estimate \(p_i(e_i)\) for
+\(i = 1, ..., n\). The KDE computation is done by the \texttt{density}
+function in R.
 
@@
 the sample mean. With the independence assumption, for non-normal \(P\),
-\(D_{KL}\) can be estimated by
+\(D_{KL}\) can be approximated by
 
@@
 \label{eq:kl-3}
-\widehat{D_{KL}} &= \sum_{i = 1}^{n} \widehat{D_{KL}^{(i)}}, \\
-\widehat{D_{KL}^{(i)}} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
+D_{KL} &\approx \sum_{i = 1}^{n} \hat{D}_{KL}^{(i)}, \\
+\hat{D}_{KL}^{(i)} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
 \end{align}
 
-\noindent where \(\widehat{D_{KL}^{(i)}}\) is the estimator of the
+\noindent where \(\hat{D}_{KL}^{(i)}\) is the estimator of the
 Kullback-Leibler divergence for an individual residual \(e_i\),
@@
 \label{eq:d-approx}
-\hat{D} = f(V(\hat{\boldsymbol{e}}, \hat{\boldsymbol{y}})),
+\hat{D} = f_{CV}(V_{h \times w}(\boldsymbol{e}, \hat{\boldsymbol{y}})),
 \end{equation}
 
-\noindent where \(\hat{\boldsymbol{e}}\) is a vector of residuals
-obtained from the fitted model rather than a vector of random variables,
-\(\hat{\boldsymbol{y}}\) is the fitted values also obtained from the
-fitted model, \(V(.)\) is a plotting function that generates a residuals
-vs fitted values plot with fixed aesthetic and then saves it as an image
-with \(h \times w\) pixels and three colour channels, and \(f(.)\) is a
+\noindent where \(V_{h \times w}(.)\) is a plotting function that saves
+a residuals vs fitted values plot with fixed aesthetic as an image with
+\(h \times w\) pixels and three colour channels, \(f_{CV}(.)\) is a
 computer vision model which takes an \(h \times w\) image as input and
@@
 is caused by the skewed distribution of the fitted values or the
-existence of heteroskedasticity. Assume we only train the model with
-this single image, while \(50\)\% of the time the label is \(0\), and
-\(50\)\% of the time the label is some constant \(c > 0\). Then, a
-perfect model will predict \(0 < \hat{D} < c\) to achieve a minimum
-loss. Some null plots could have outliers or strong visual patterns due
-to randomness, and a reasonable model will try to summarise those
-information into the prediction, resulting in \(\hat{D} > 0\).
+existence of heteroskedasticity. Additionally, some null plots could
+have outliers or strong visual patterns due to randomness, and a
+reasonable model will try to summarise those information into the
+prediction, resulting in \(\hat{D} > 0\).
 
@@
 the visual pattern is weak or moderate, having \(\hat{D}\) will not be
-sufficient to determine if the null hypothesis that the model is
-correctly specified should be rejected.
+sufficient to determine if \(H_0\) that the model is correctly specified
+should be rejected.
 
 To solve this issue while aligning with the principle of visual
-inference, the approximated distance \(\hat{D}\) can be defined as a
-test statistic. And the null distribution of this statistic can be
-approximated by the empirical distribution of \(\widehat{D_{null}}\) for
-a large amount of null plots of a fitted model. The procedure involves
+inference, the approximated distance \(\hat{D}\) can be viewed as a test
+statistic. And the null distribution of this statistic can be
+approximated by the empirical distribution of \(\hat{D}_{null}\) for a
+large amount of null plots of a fitted model. The procedure involves
 applying the residual rotation technique \citep{buja2009statistical} on
 the fitted model to obtain null residuals. The null residuals are then
-used to make null plots and fed into the model. The predicted distance
-are used to construct an empirical distribution.
+used to make null plots and fed into the model to get predictions. The
+empirical distribution is constructed with the predicted distance.
+
+There are two types of error associated with the approximation of the
+null distribution. These include the sampling error and the estimation
+error for model parameters. Increasing the number of null plots within
+the distribution approximation can effectively reduce sampling error.
+However, the estimation error for model parameters remains irreducible.
+The residual rotation technique operates under the assumption of a
+correct fitted model, using the estimated variance instead of the
+unknown true variance to rescale rotated residuals. Consequently, under
+\(H_0\), the true distribution of \(\hat{D}_{null}\) will slightly
+deviate from the approximated null distribution.
 
@@
 
-Let \(\widehat{D_{null}^{(i)}}\) be the predicted distance of the
-\(i\)-th null plots, where \(i = 1,...,m\) and \(m \in \mathbb{N}^+\) is
-a sufficiently large number. Quantiles of the null distribution can then
-be estimated using the sample quantiles available in statistical
-software such as R. The details of the sample quantile computation can
-be found in \citet{hyndman1996sample}. In statistical testing, analysts
-often care about certain quantiles of the null distribution, such as
-90\% quantile, 95\% quantile and 99\% quantile. These quantiles are used
-as thresholds to decide if \(H_0\) needs to be rejected. For example, if
-\(\hat{D}\) is greater than and equal to the 95\% sample quantile of the
-null distribution, we could say the predicted distance for the actual
-residual plot is significantly different from the predicted distance for
-null plots with 95\% significance level. Based on our experience, in
-order to get a stable estimate of the 95\% quantile, \(m\) usually needs
-to be at least \(100\). And if the null distribution has a long tail,
-more null plots will be needed. Alternatively, a p-value can be used to
-represents the probability of observing an event equally or more extreme
-than the given event under \(H_0\), and it can be estimated by
-\(1/m\sum_{i=1}^{m}I\left(\widehat{D_{null}^{(i)}} \geq \hat{D}\right)\).
+Let \(\hat{D}_{null}^{(i)}\) be the predicted distance of the \(i\)-th
+null plots, where \(i = 1,...,n_{null}\) and
+\(n_{null} \in \mathbb{N}^+\) is a sufficiently large number. Quantiles
+of the null distribution can then be estimated using the sample
+quantiles available in statistical software such as R. The details of
+the sample quantile computation can be found in
+\citet{hyndman1996sample}. In statistical testing, analysts often care
+about certain quantiles of the null distribution, such as the 90\%
+quantile, the 95\% quantile and the 99\% quantile. These quantiles are
+used as thresholds to decide if \(H_0\) needs to be rejected. For
+example, if \(\hat{D}\) is greater than and equal to the 95\% sample
+quantile of the null distribution \(Q_{null}(0.95)\), we could say the
+predicted distance for the actual residual plot is significantly
+different from the predicted distance for null plots with 95\%
+significance level. Based on our experience, in order to get a stable
+estimate of the 95\% quantile, \(n_{null}\) usually needs to be at least
+\(100\). And if the null distribution has a long tail, more null plots
+will be needed. Alternatively, a p-value can be used to represents the
+probability of observing an event equally or more extreme than the given
+event under \(H_0\), and it can be estimated by
+\(1/m\sum_{i=1}^{n_{null}}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)\).
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  What is the assumption of bootstrapping
-\item
-  Any drawback: loss of information
-\item
-  For each bootstrapped fitted model, we can get a bootstrapped vss and
-  construct a new 95\% quantile for the null distribution, and check if
-  H\_0 should be rejected
-\item
-  However, it is very expensive to construct a new 95\% quantile for
-  each bootstrapped fitted model
-\item
-  The 95\% quantile constructed using the original fitted model can be
-  considered as an estimate of those new 95\% quantiles
-\item
-  We borrow information from the original fitted model
-\item
-  We can check how many bootstrapped fitted model should be rejected
-  (like the percentage)
-\item
-  This gives an overall estimate of how often the assumed regression
-  model are considered to be incorrect if the data can be obtained
-  repetitively from the same data generating process
-\item
-  If this percentage is relatively large, it gives the analyst an
-  warning that the regression model is inappropriate to the data
-\end{itemize}
+Bootstrap is often employed in linear regression when conducting
+inference for estimated parameters. It is typically done by sampling
+individual cases with replacement and refitting the regression model. If
+the observed data accurately reflects the true distribution of the
+population, the bootstrapped estimates can be used to measure the
+variability of the parameter estimate without making strong
+distributional assumptions about the data generating process.
+
+Similarly, bootstrap can be applied on the approximated distance
+\(\hat{D}\). For each refitted model \(M_{boot}^{(i)}\), there will be
+an associated residual plot \(V_{boot}^{(i)}\) which can be fed into the
+computer vision model to obtain \(\hat{D}_{boot}^{(i)}\), where
+\(i = 1,...,n_{boot}\), and \(n_{boot}\) is the number of bootstrapped
+samples. If we are interested in the variation of \(\hat{D}\), we can
+use \(\hat{D}_{boot}^{(i)}\) to estimate a confidence interval.
+
+Alternatively, since each \(M_{boot}^{(i)}\) has a set of estimated
+coefficients \(\hat{\boldsymbol{\beta}}_{boot}^{(i)}\) and an estimated
+variance \(\hat{\sigma^2}_{boot}^{(i)}\), a new approximated null
+distribution can be construed and the corresponding 95\% sample quantile
+\(Q_{boot}^{(i)}(0.95)\) can be computed. Then, if
+\(\hat{D}_{boot}^{(i)} \geq Q_{boot}^{(i)}(0.95)\), \(H_0\) will be
+rejected for \(M_{boot}^{(i)}\). The ratio of rejected
+\(M_{boot}^{(i)}\) among all the refitted models provides an indication
+of how often the assumed regression model are considered to be incorrect
+if the data can be obtained repetitively from the same data generating
+process. But this approach is computationally very expensive because it
+requires \(n_{boot} \times n_{null}\) times of residual plot assessment.
+In practice, \(Q_{null}(0.95)\) will be used to replace
+\(Q_{boot}^{(i)}(0.95)\) in the computation.
 
@@
 \(\boldsymbol{x}_1\) and \(\boldsymbol{x}_2\) are two independent
-predictors, \(He_j(.)\) is the \(j\)th-order probabilist's Hermite
-polynomials \citep{hermite1864nouveau}, the \(\sqrt{(.)}\) and \((.)^2\)
-operators are element-wise operators, \(\odot\) is the Hadamard product,
-and \(g(., k)\) is a scaling function to enforce the support of the
-random vector to be \([-k, k]^n\) defined as
+predictors, \(\text{He}_j(.)\) is the \(j\)th-order probabilist's
+Hermite polynomials \citep{hermite1864nouveau}, the \(\sqrt{(.)}\) and
+\((.)^2\) operators are element-wise operators, \(\odot\) is the
+Hadamard product, and \(g(., k)\) is a scaling function to enforce the
+support of the random vector to be \([-k, k]^n\) defined as
 
@@
 
-The model is trained on a high-performance computing platform owned by
-Monash university with TensorFlow (ref here) and Keras (ref here).
-During the training process, 80\% of the training data are used as the
-actual training data, and 20\% of the training data are reserved as
-validation data. We let the Bayesian optimization tuner runs 100 trials
-to find the best hyperparameter values according to the validation root
-mean square error. The tuner will restore the best epoch of the best
-model from the trials. Early stopping is also applied such that the
-training process will halt if the validation root mean square error does
-not improve for 50 epochs. The maximum allowed epochs is 2000 where no
-models reach.
+The model is trained on the M3 high-performance computing platform (ref
+here) with TensorFlow (ref here) and Keras (ref here). During the
+training process, 80\% of the training data are used as the actual
+training data, and 20\% of the training data are reserved as validation
+data. We let the Bayesian optimization tuner runs 100 trials to find the
+best hyperparameter values according to the validation root mean square
+error. The tuner will restore the best epoch of the best model from the
+trials. Early stopping is also applied such that the training process
+will halt if the validation root mean square error does not improve for
+50 epochs. The maximum allowed epochs is 2000 where no models reach.
 
@@
 
-\hypertarget{best-hyperparameters}{%
-\subsection{Best hyperparameters}\label{best-hyperparameters}}
+\hypertarget{optimized-hyperparameters}{%
+\subsection{Optimized hyperparameters}\label{optimized-hyperparameters}}
 
 Based on the tuning process described in Section
-\ref{training-and-hyperparameter-tuning}, the best hyperparameter values
-are provided in Table \ref{tab:best-hyperparameter}. It can be observed
-that at least \(32\) base filters are needed, and the preferable choice
-for the \(64 \times 64\) model and the \(128 \times 128\) model is
-\(64\) base filters, which is the same as the original VGG16
-architecture. The best values for dropout rate for convolutional blocks
-is around \(0.4\). And applying batch normalization for convolutional
-blocks has a positive impact on the performance. All best models choose
-to preserve the additional inputs to reduce the validation error. The
-number of units needed for the fully-connected layer is \(256\), which
-is not a very large number comparing to the VGG16 classifier, suggesting
-that the problem we are trying to solve is relatively easy. The best
-learning rates are usually smaller than the default value recommended by
-Keras which is 0.001.
+\ref{training-and-hyperparameter-tuning}, the optimized hyperparameter
+values are provided in Table \ref{tab:best-hyperparameter}. It can be
+observed that at least \(32\) base filters are needed, and the
+preferable choice for the \(64 \times 64\) model and the
+\(128 \times 128\) model is \(64\) base filters, which is the same as
+the original VGG16 architecture. The optimized values for dropout rate
+for convolutional blocks is around \(0.4\). And applying batch
+normalization for convolutional blocks has a positive impact on the
+performance. All optimized models choose to preserve the additional
+inputs to reduce the validation error. The number of units needed for
+the fully-connected layer is \(256\), which is not a very large number
+comparing to the VGG16 classifier, suggesting that the problem we are
+trying to solve is relatively easy. The optimized learning rates are
+usually smaller than the default value recommended by Keras which is
+0.001.
 
@@
 
-\caption{\label{tab:best-hyperparameter}Hyperparameters values for the best computer vision models with differnet input sizes.}
+\caption{\label{tab:best-hyperparameter}Hyperparameters values for the optimized computer vision models with different input sizes.}
 \centering
@@
 
-The training and test performance of models with three different input
-sizes are provided in Table \ref{tab:performance}. Among these models,
-the \(64 \times 64\) model and the \(32 \times 32\) model are
-consistently having the best metrics on the training set and the test
-set respectively. The mean absolute error shows that the difference
-between \(\hat{D}\) and \(D\) is around \(0.43\) on the test set, which
-is very small considering the range of \(D\) is normally between \(0\)
-and \(7\). The high \(R^2\) also suggests the prediction is mostly
-linearly correlated with the target.
-
-Figure \ref{fig:model-performance} displays a scatter plot for
+The training and test performance of optimized models with three
+different input sizes are provided in Table \ref{tab:performance}. Among
+these models, the \(64 \times 64\) model and the \(32 \times 32\) model
+are consistently having the best training and test performance
+respectively. The mean absolute error shows that the difference between
+\(\hat{D}\) and \(D\) is around \(0.43\) on the test set, which is very
+small considering the range of \(D\) is normally between \(0\) and
+\(7\). The high \(R^2\) also suggests the prediction is mostly linearly
+correlated with the target.
+
+Figure \ref{fig:model-performance} displays a hexagonal heatmap for
 \(D - \hat{D}\) vs \(D\). For residual plots with \(D\) close to zero,
-all the models tends to over-predict \(\hat{D}\), leading to negative
-residuals. When \(D\) is between \(1.5\) and \(6\), the smooth curves
-indicate that \(\hat{D}\) is not heavily biased, and there are some
-instances where \(\hat{D}\) is severely under-predicted. And when \(D\)
-is greater than \(6\), there are more instances being under-predicted
-than over-predicted. It can be observed from Figure
-\ref{fig:model-performance-density} that the models produced much more
-\(\hat{D}\) at around \(1\) than expected.
+all the optimized models tends to over-predict \(\hat{D}\), leading to
+negative residuals. When \(D\) is between \(1.5\) and \(6\), the smooth
+curves indicate that \(\hat{D}\) is not heavily biased, and there are
+some instances where \(\hat{D}\) is severely under-predicted. And when
+\(D\) is greater than \(6\), there are more instances being
+under-predicted than over-predicted.
 
@@
 
-\caption{\label{tab:performance}The training and test performance of three models with different input sizes.}
+\caption{\label{tab:performance}The training and test performance of three optimized models with different input sizes.}
 \centering
@@
 
-\caption{Residuals vs distance plot on training and test data for three best models with different input sizes.
+\caption{Hexagonal heatmap for residuals vs distance on training and test data for three optimized models with different input sizes.
 The blue lines are smoothing curves produced by fitting gnealized additive models.}\label{fig:model-performance}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/model-performance-density-1}
-
-}
-
-\caption{Density plot of predicted distance on training and test data for three best models with different input sizes.
-The density drawn in red is the density of the distance.}\label{fig:model-performance-density}
 \end{figure}
@@
 
-\begin{itemize}
-\tightlist
-\item
-  basic background
-\item
-  highlights the same and the different settings
-\end{itemize}
-
-\hypertarget{comparison}{%
-\subsubsection{Comparison}\label{comparison}}
+In order to check the validity of the proposed computer vision model,
+residual plots presented in the human subject experiment conducted by
+\citet{li2023plot} will be assessed. This study has collected \(7974\)
+human responses to \(1152\) lineups. Each lineup contains one randomly
+placed actual residual plot and 19 null plots. Among the \(1152\)
+lineups, \(24\) are attention check lineups in which the visual patterns
+are designed to be extremely obvious and very different from the
+corresponding to null plots, \(36\) are null lineups where all the
+lineups consist of only null plots, \(279\) are lineups with uniform
+predictor distribution evaluated by \(11\) participants, and the
+remaining \(813\) are lineups with discrete, skewed or normal predictor
+distribution evaluated by \(5\) participants. Attention check lineups
+and null lineups will not be assessed in the following analysis.
+
+In \citet{li2023plot}, the residual plots are simulated from a data
+generating process which is a special case of Equation
+\ref{eq:data-sim}. The main characteristic is the model violations are
+introduced separately, meaning non-linearity and heteroskedasticity will
+not co-exist in one lineup but assigned uniformly to all lineups.
+Additionally, non-normality and multiple predictors are not considered
+in their experimental design.
+
+\hypertarget{model-performance-on-the-human-data}{%
+\subsubsection{Model performance on the human
+data}\label{model-performance-on-the-human-data}}
+
+For each lineup used in \citet{li2023plot}, there are one actual
+residual plot and \(19\) null plots. While the distance \(D\) for the
+actual residual plot depends on the underlying data generating process,
+the distance \(D\) for the null plots are zeros. We have used our
+optimized computer vision model to predict \(\hat{D}\) for both the
+actual residual plots and the null plots.
+
+The performance metrics of \(\hat{D}\) for actual residual plots are
+provided in Table \ref{tab:experiment-performance}. It can be observed
+that all the performance metrics are slightly worse than the metrics
+evaluated on the test data. Still, the mean absolute error remains at a
+low level, and the linearly correlation between the prediction and the
+true value is still very high. Lineups with non-linearity issues are
+more difficult to be predicted than those with heteroskedasticity
+issues.
+
+Figure \ref{fig:rd-human} displays the \(D - \hat{D}\) vs \(D\) plot for
+actual residual plots used in the experiment. The behaviour of the
+prediction is roughly the same as what we can be observed from the
+evaluation of the test data. The model over-predicts \(\hat{D}\) when
+\(D\) is close to zero, and under-predicts \(\hat{D}\) when \(D\) is
+greater than \(6\). However, for residual plots with non-linearity
+issues, we do not observe the over-prediction for \(D > 6\) because
+there are not many cases in that area. And the over-prediction happens
+at around \(2 < D < 4\).
+
+For null plots, a histogram of the prediction is provided in Figure
+\ref{fig:hist-null-human}. In most cases, \(\hat{D}\) will be predicted
+at around \(1\). And as \(\hat{D}\) increases, the number of predictions
+decreases. There are some outliers with \(\hat{D} > 5\).
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-2-1}
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/rd-human-1}
 
@@
 
-\caption{Residuals vs distance plot on data used in the human subject experiment. The blue line is a smoothing curve produced by fitting a gnealized additive model.}\label{fig:unnamed-chunk-2}
+\caption{Residuals vs distance plot on lineups used in the human subject experiment. The blue line is a smoothing curve produced by fitting a gnealized additive model.}\label{fig:rd-human}
 \end{figure}
@@
 \toprule
-Type & RMSE & $R^2$ & MAE & Huber loss\\
+Voilation & RMSE & $R^2$ & MAE & Huber loss\\
 \midrule
-heteroskedasticity & 0.768 & 0.863 & 0.594 & 0.263\\
-non-linearity & 0.731 & 0.817 & 0.560 & 0.242\\
+heteroskedasticity & 0.721 & 0.852 & 0.553 & 0.235\\
+non-linearity & 0.738 & 0.770 & 0.566 & 0.246\\
 \bottomrule
@@
 
-It can be observed from Table \ref{tab:experiment-performance} that all
-the performance metrics are lower than the metrics evaluated on the test
-data.
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/hist-null-human-1}
+
+}
+
+\caption{Histogram of predicted distance on null plots used in the human subject experiment.}\label{fig:hist-null-human}
+\end{figure}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-3-1} \end{center}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-4-1}
+
+}
+
+\caption{A weighted detection rate vs $\delta$-differnence plot.}\label{fig:unnamed-chunk-4}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-5-1}
+
+}
+
+\caption{A weighted detection rate vs $\gamma$-number of extreme nulls plot.}\label{fig:unnamed-chunk-5}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-6-1}
+
+}
+
+\caption{A weighted detection rate vs $\delta$-differnence plot for lineups evaluated by 11 people.}\label{fig:unnamed-chunk-6}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-7-1}
+
+}
+
+\caption{A weighted detection rate vs $\gamma$-number of extreme nulls plot for lineups evaluated by 11 people.}\label{fig:unnamed-chunk-7}
+\end{figure}
 

### 2024-02-12 paper/paper.Rmd
update paper

@@
   Automated assessment of residual plots with computer vision models
-type: ARTICLE TEMPLATE
 author:
@@
     email: weihao.li@monash.edu
+  - name: Dianne Cook
+    affil: a
+    email: dicook@monash.edu
+  - name: Emi Tanaka
+    affil: b, c
+    email: emi.tanaka@anu.edu.au
+  - name: Susan VanderPlas
+    affil: d
+    email: susan.vanderplas@unl.edu
+  - name: Klaus Ackermann
+    affil: a
+    email: Klaus.Ackermann@monash.edu
 affiliation:
@@
       Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
+  - num: b
+    address: |
+      Biological Data Science Institute, Australian National University, Acton, ACT, Australia
+  - num: c
+    address: |
+      Research School of Finance, Actuarial Studies and Statistics, Australian National University, Acton, ACT, Australia
+  - num: d
+    address: |
+      Department of Statistics, University of Nebraska, Lincoln, Nebraska, USA
 bibliography: ref.bib
@@
 library(visage)
+library(glue)
 
@@
 While this might suggest heteroskedasticity, it is important to avoid over-interpreting this visual pattern.
-In this case, the fitted model is correctly specified, and the triangular shape is actually a result of the skewed distribution of the predictors, rather than indicating a flaw in the model.
+In this case, the fitted regression model is correctly specified, and the triangular shape is actually a result of the skewed distribution of the predictors, rather than indicating a flaw in the model.
 
@@
 The concept of visual inference, as proposed by @buja2009statistical, provides an inferential framework to assess whether residual plots indeed contain visual patterns inconsistent with the model assumptions.
-The fundamental idea involves testing whether the actual residual plot visually differs significantly from null plots, which are created using residuals generated from the null distribution.
-Typically, this is accomplished through the lineup protocol.
-In this approach, the real residual plot is embedded within a lineup alongside several null plots.
-If the real residual plot can be distinguished from the lineup, it provides evidence for rejecting the null hypothesis.
+The fundamental idea involves testing whether the actual residual plot visually differs significantly from null plots, where null plots are plotted with residuals generated from the residual rotation distribution [@langsrud2005rotation], which is a distribution consistent with the null hypothesis $H_0$ that the linear regression model is correctly specified.
+Typically, the visual test is accomplished through the lineup protocol, where the real residual plot is embedded within a lineup alongside several null plots.
+If the real residual plot can be distinguished from the lineup, it provides evidence for rejecting $H_0$.
 
@@
 Beyond its application in residual diagnostics, the lineup protocol has integrated into the analysis of diverse subjects.
-For instance, \cite{loy2013diagnostic, loy2014hlmdiag, loy2015you} illustrate its applicability in diagnosing hierarchical linear models.
-Additionally, @widen2016graphical demonstrates its utility in geographical research, while @krishnan2021hierarchical explores its effectiveness in forensic examinations.
+For instance, \cite{loy2013diagnostic, loy2014hlmdiag, loy2015you} illustrated its applicability in diagnosing hierarchical linear models.
+Additionally, @widen2016graphical demonstrated its utility in geographical research, while @krishnan2021hierarchical explored its effectiveness in forensic examinations.
 
@@
 The utilization of computers to interpret data plots has a rich history, with early efforts such as "Scagnostics" by @tukey1985computer, focusing on scatterplot diagnostics.
-@wilkinson2005graph expanded on this work, introducing graph theoretic scagnostics, which encompassed nine computable measures applied to planar proximity graphs.
+@wilkinson2005graph expanded on this work, introducing graph theoretic scagnostics, which encompassed computable measures applied to planar proximity graphs.
 These measures, including, but not limited to, "Outlying", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy", and "Striated" aimed to characterize outliers, shape, density, trend, coherence and other characteristics of the data.
 While this approach has been inspiring, there is a recognition [@buja2009statistical] that it may not capture all the necessary visual features that differentiate actual residual plots from null plots.
-A more promising alternative entails enabling computers to learn the function for extracting visual features from residual plots.
+A more promising alternative entails enabling machines to learn the function for extracting visual features from residual plots.
 Essentially, this means empowering computers to discern the crucial visual features for residual diagnostics and determining the method to extract them.
+
 Modern computer vision models are well-suited for addressing this challenge.
-
-Modern computer vision models often rely on deep neural networks with convolutional layers [@fukushima1982neocognitron].
+They rely on deep neural networks with convolutional layers [@fukushima1982neocognitron].
 These layers leverage hierarchical patterns in data, downsizing and transforming images by summarizing information in a small space.
@@
 
-```{r false-finding, fig.width = 8, fig.height = 6, fig.cap="An example residual vs fitted values plot (red line indicates 0).
+```{r false-finding, fig.width = 8/5*4, fig.height = 6/4*4, fig.pos="!h", fig.cap="An example residual vs fitted values plot (red line indicates 0).
 The vertical spread of the data points varies with the fitted values.
@@
 
-### Different input formats
-
-Deep learning models are in general very sensitive to the input data. The quality and relevance of the input data greatly influence the model's capacity to generate insightful and meaningful results.
-
-A straightforward approach involves feeding a vector of residuals along with a vector of fitted values, essentially providing all the necessary information for creating a residuals vs fitted values plot.
+### Input formats
+
+Deep learning models are in general very sensitive to the input data.
+The quality and relevance of the input data greatly influence the model's capacity to generate insightful and meaningful results.
+There are several designs of the input format can be considered.
+
+A straightforward design involves feeding a vector of residuals along with a vector of fitted values, essentially providing all the necessary information for creating a residuals vs fitted values plot.
 However, a drawback of this method is the dynamic input size, which changes based on the number of observations.
@@
 By controlling the number of bins in the histograms, it becomes possible to provide fixed-length input vectors.
-Still, since the histogram only captures the marginal distribution of residuals and fitted values respectively, it can not be used to differentiate visual patterns with same marginal distributions but different joint distributions.
-
-Another approach is to use an image as input.
-The primary advantage of using an image, as opposed to the vector format, is the utilization of the sophisticated image processing architectures developed over the years, such as the VGG16 architecture proposed in @simonyan2014very.
+Still, since histograms only capture the marginal distribution of residuals and fitted values respectively, they can not be used to differentiate visual patterns with same marginal distributions but different joint distributions.
+
+Another design involves using an image as input.
+The primary advantage of this design, as opposed to the vector format, is the availability of the existing and sophisticated image processing architectures developed over the years, such as the VGG16 architecture proposed in @simonyan2014very.
 These architectures can effectively capture and summarize spatial information from nearby pixels, which is less straightforward with vector input.
-The only considerations are the image resolution and the aesthetics of the residual plot.
+The main considerations are the image resolution and the aesthetics of the residual plot.
 In general, higher resolution provides more information to the model but comes with the trade-off of increased complexity and greater difficulty in training.
 As for the aesthetics of the residual plot, a practical solution is to consistently present residual plots in the same style to the model.
-This implies that the model can not accept arbitrary images as input but requires the use of the same preprocessing software to convert provided residuals and fitted values into a standardized style for the residual plot.
+This implies that the model can not accept arbitrary images as input but requires the use of the same preprocessing pipeline to convert residuals and fitted values into a standardized-style residual plot.
 
 Providing multiple residual plots to the model, such as a pair of plots, a triplet or a lineup is also a possible option.
-@chopra2005learning has shown that computer vision models designed for image comparison can assess whether a pair of images are similar or dissimilar.
-Applied to our specific problem, we can define null plots of a fitted model to be similar to each other, while considering actual residual plots to be distinct from null plots of any fitted model.
+@chopra2005learning have shown that computer vision models designed for image comparison can assess whether a pair of images are similar or dissimilar.
+Applied to our specific problem, we can define null plots of a fitted regression model to be similar to each other, while considering actual residual plots to be distinct from null plots of any fitted regression model.
 A triplet constitutes a set of three images, denoted as $image_1$, $image_2$ and $image_3$.
@@
 For this setup, we can apply the same criteria to define similarity between images.
-However, it's important to note that these two approaches usually require additional considerations regarding the loss function and, at times, non-standard training processes due to shared weights between different convolutional blocks.
-
-Presenting a lineup to a model and asking it to predict which residual plot is the most different one aligns closely with the lineup protocol.
-However, if a lineup consists of a large amount of residual plots, the resolution of the input image could become quite large, posing challenges in training the model.
-This approach was experimented in a pilot study conducted by us but the performance of the trained model was sub-optimal.
+However, it is important to note that these two approaches usually require additional considerations regarding the loss function and, at times, non-standard training processes due to shared weights between different convolutional blocks.
+
+Presenting a lineup to a model aligns closely with the lineup protocol.
+However, as the number of residual plots in a lineup increases, the resolution of the input image grows rapidly, posing challenges in training the model.
+We experimented with this approach in a pilot study, but the performance of the trained model was sub-optimal.
 
@@
 
-### Different output formats
-
-Given that the input is a single residual plot represented as a fixed-resolution image, the output from the computer vision model can take one of two formats: binary or numeric.
+### Output formats
+
+Given that the input is a single residual plot represented as a fixed-resolution image, the output from the computer vision model can take one of two forms: binary or numeric.
 This choice determines whether the model belongs to a classification model or a regression model.
-The binary outcome could be designed to indicate whether the input image is a null plot, or whether the input image would be rejected in a visual test conducted by humans.
-The latter option requires data from prior human subject experiments, presenting difficulties in controlling the quality of data due to variations in experimental settings across different studies.
-Additionally, some visual inference experiments are unrelated to linear regression models and residual plot diagnostics, resulting in a limited amount of available training data.
-
-Alternatively, the output could be designed as a meaningful numeric measure, such as the average distance to good residual plots.
-This approach necessitates defining a distance measure between residual plots, which may be a non-trivial task.
-However, once a meaningful distance measure is established, the prediction from the computer vision model becomes an interpretable value with tangible significance.
-Studies have demonstrated that defining a proper distance between images can enhance the matching accuracy in image search compared to a binary outcome model (ref here).
+The binary outcome encoded as $0$ and $1$ could be used to represent whether the input image is a null plot, or whether the input image would be rejected in a visual test conducted by humans.
+Training a model following the latter option requires data from prior human subject experiments, presenting difficulties in controlling the quality of data due to variations in experimental settings across different studies.
+Additionally, some visual inference experiments are unrelated to linear regression models or residual plot diagnostics, resulting in a limited amount of available training data.
+
+Alternatively, the output could be a meaningful and interpretable numerical measure useful for assessing residual plots, such as the strength of suspicious visual patterns reflecting the extent of model violations and the difficulty index for identifying whether a residual plot has no issues.
+However, these numeric measures are often informally used in daily communication but are not typically formalized or rigorously defined.
+For the purpose of training a model, this numeric measure has to be quantifiable.
+
+In this study, we chose to define and use a distance measure to quantify the difference between the residual plot and the null plots.
+@vo2016localizing have also demonstrated that defining a proper distance between images can enhance the matching accuracy in image search compared to a binary outcome model.
 
@@
 The challenge lies in the fact that metrics tailored for image comparison may not be suitable for evaluating data plots, where only essential plot elements require assessment [@chowdhury2018measuring].
-Furthermore, scagnostics mentioned in Section \ref{introduction} could be used to construct distance metrics for data plots comparison, but the functional form still needs to be carefully refined to accurately reflect the extent of the violations.
+Furthermore, scagnostics mentioned in Section \ref{introduction} could be used to construct distance metrics for residual plots comparison, but the functional form still needs to be carefully refined to accurately reflect the extent of the violations.
 
@@
 
-The distance metrics proposed in this paper takes into account the fact that we try to measure how different a residual plot is from a good residual plot, or in other words, how different a given fitted model is from a correctly specified model.
+The distance measure proposed in this study takes into account the fact that we tried to measure how different a residual plot is from a good residual plot, or in other words, how different a given fitted regression model is from a correctly specified model.
 For the classical normal linear regression model, residuals $\boldsymbol{e}$ are derived from the fitted values $\hat{\boldsymbol{y}}$ and observed values $\boldsymbol{y}$.
@@
 It can be known that residuals $\boldsymbol{e}$ also follow a certain probability distribution transformed from the multivariate normal distribution, which will be denoted as $Q$.
-This reference distribution $Q$ summarizes what good residuals should follow given the predictors are known and fixed.
+This reference distribution $Q$ summarizes what good residuals should follow given the design matrix $\boldsymbol{X}$ is known and fixed.
 
@@
 To capture the characteristics of $Q$, such as moments, we can simulate a large numbers of $\boldsymbol{\varepsilon}$ and transform it to $\boldsymbol{e}$ to get the empirical estimates.
-For simplicity, we replaced the variance-covariance matrix of residuals $\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}' = \boldsymbol{R}\sigma^2$ with a full-rank diagonal matrix $\text{diag}(\boldsymbol{R}\sigma^2)$, where $\text{diag}(.)$ sets the non-diagonal entries of a matrix to zeros.
+For simplicity, in this study, we replaced the variance-covariance matrix of residuals $\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}' = \boldsymbol{R}\sigma^2$ with a full-rank diagonal matrix $\text{diag}(\boldsymbol{R}\sigma^2)$, where $\text{diag}(.)$ sets the non-diagonal entries of a matrix to zeros.
 The resulting distribution for $\boldsymbol{Q}$ is $N(\boldsymbol{0}_n, \text{diag}(\boldsymbol{R}\sigma^2))$.
@@
 However, if the model is misspecified, then the actual distribution of residuals denoted as $P$, will be different from $Q$.
-For example, if the data generating process contains variables correlated with any column of $\boldsymbol{X}$ but not included in $\boldsymbol{X}$, causing an omitted variable problem, $P$ will be different from $Q$ because the residual operator obtained from the fitted model will not be the same as $\boldsymbol{R}$.
+For example, if the data generating process contains variables correlated with any column of $\boldsymbol{X}$ but not included in $\boldsymbol{X}$, causing an omitted variable problem, $P$ will be different from $Q$ because the residual operator obtained from the fitted regression model will not be the same as $\boldsymbol{R}$.
 Besides, if the $\boldsymbol{\varepsilon}$ follows a non-normal distribution such as a multivariate lognormal distribution, $P$ will usually be skewed and has a long tail.
@@
 Define a proper distance between distributions is usually easier than define a proper distance between data plots.
-Given the actual residual distribution $Q$ and the reference residual distribution $P$, we used a distance metric based on Kullback-Leibler divergence [@kullback1951information] to quantify the difference between two distributions
+Given the actual residual distribution $Q$ and the reference residual distribution $P$, we used a distance measure based on Kullback-Leibler divergence [@kullback1951information] to quantify the difference between two distributions
 
@@
 
-This distance metric was first proposed in @li2023plot.
+This distance measure was first proposed in @li2023plot.
 It was mainly designed for measuring the effect size of non-linearity and heteroskedasticity in a residual plot.
-@li2023plot showed that, for a classical normal linear regression model that omits a necessary higher-order predictors $\boldsymbol{Z}$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$ with $\boldsymbol{V}$ being an arbitrary symmetric positive semi-definite matrix, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}))$.
+@li2023plot have showed that, for a classical normal linear regression model that omits a necessary higher-order predictors $\boldsymbol{Z}$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$ with $\boldsymbol{V}$ being an arbitrary symmetric positive semi-definite matrix, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}))$.
 Note that the variance-covariance matrix is replaced with the diagonal matrix to ensure it is a full-rank matrix.
@@
 For non-normal error $\boldsymbol{\varepsilon}$, the actual residual distribution $P$ is unlikely to be a multivariate normal distribution.
-Thus, equation \ref{eq:kl-2} given in @li2023plot will not be applicable to models with non-normality violations.
+Thus, equation \ref{eq:kl-2} given in @li2023plot will not be applicable to models violating the normality assumption.
 
@@
 
-The algorithm for approximating equation \ref{eq:kl-1} starts from simulating $m$ sets of $\boldsymbol{\varepsilon}$ according to the error distribution.
-The simulated errors are stored in a matrix $\boldsymbol{A}$ with $n$ rows and $m$ columns, where each column of $\boldsymbol{A}$ is a set of realized values of $\boldsymbol{\varepsilon}$.
+Given $\boldsymbol{X}$ and $\boldsymbol{\beta}$, the algorithm for approximating equation \ref{eq:kl-1} starts from simulating $m$ sets of observed values $\boldsymbol{y}$ according to the data generating process.
+The observed values are stored in a matrix $\boldsymbol{A}$ with $n$ rows and $m$ columns, where each column of $\boldsymbol{A}$ is a set of observed values.
 Then, we can get $m$ sets of realized values of $\boldsymbol{e}$ stored in the matrix $\boldsymbol{B}$ by applying the residual operator $\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}$.
 Furthermore, kernel density estimation (KDE) with Gaussian kernel and optimal bandwidth selected by the Silverman's rule of thumb [@silverman2018density] is applied on each row of $B$ to estimate $p_i(e_i)$ for $i = 1, ..., n$.
-The KDE computation is done by the `density` function in R.
+The KDE computation can be done by the `density` function in R.
 
@@
 
-### Approximation of the distance metric
-
-In the previous sections, we have defined a distance metric given in equation \ref{eq:kl-0} for quantifying the difference between the actual residual distribution $P$ and an ideal reference distribution $Q$.
-You may have noticed that this distance metric can only be computed when the data generating process is known.
+### Approximation of the distance measure
+
+In the previous sections, we have defined a distance measure given in equation \ref{eq:kl-0} for quantifying the difference between the actual residual distribution $P$ and an ideal reference distribution $Q$.
+You may have noticed that this distance measure can only be computed when the data generating process is known.
 In reality, we often have no knowledge about the data generating process, otherwise we do not need to fit a linear regression model in the first place.
 
-What we proposed in this paper is a method to approximate this distance with a residual plot.
+We tried to train a computer vision model to approximate this distance measure with a residual plot.
 Let $D$ be the result of equation \ref{eq:kl-0} indicating the extent of the model violations, and our estimator $\hat{D}$ is formulated as
@@
 With the approximated distance $\hat{D}$, we will be able to know how different the underlying distribution of the residuals is from a good residual distribution.
-This is meaningful way to check if a residual plot is a good residual plot, and to know the strength of the visual signal embedded in the residual plot.
+It also provides information for the strength of the visual signal embedded in the residual plot.
 
 The approximated distance $\hat{D}$ is not expected to be the same as the original distance $D$.
-This is largely because information contained in a single residual plot is limited and it may not be able to summarise all the characteristics of the residual distribution.
+This is largely because information contained in a single residual plot is limited and it may not be able to summarise all the important characteristics of the residual distribution.
 For a given residual distribution $P$, we can generate many different residual plots.
-Some of them share similar visual patterns, but some of them could be visually very different from the rest, especially for models with small $n$.
+Some of them share similar visual patterns, but some of them could be visually very different from the rest, especially for regression models with small $n$.
 This suggests the error of the approximation will vary depends on whether the observed residual plot is representative or not.
@@
 This property is not an issue if $\hat{D} \gg 0$ for which the visual signal of the residual plot is very strong, and we usually do not need any further examination of the significance of the result.
-However, if the visual pattern is weak or moderate, having $\hat{D}$ will not be sufficient to determine if $H_0$ that the model is correctly specified should be rejected.
-
-To solve this issue while aligning with the principle of visual inference, the approximated distance $\hat{D}$ can be viewed as a test statistic.
-And the null distribution of this statistic can be approximated by the empirical distribution of $\hat{D}_{null}$ for a large amount of null plots of a fitted model.
-The procedure involves applying the residual rotation technique [@buja2009statistical] on the fitted model to obtain null residuals.
-The null residuals are then used to make null plots and fed into the model to get predictions.
-The empirical distribution is constructed with the predicted distance.
+However, if the visual pattern is weak or moderate, having $\hat{D}$ will not be sufficient to determine if $H_0$ should be rejected.
+
+To solve this issue while aligning with the principle of visual inference, $\hat{D}$ can be viewed as a test statistic.
+And the null distribution of this statistic can be approximated by the empirical distribution of $\hat{D}_{null}$, where $\hat{D}_{null}$ is the approximated distance for a null plot simulated from the fitted regression model.
+The approximation of the distribution involves applying the residual rotation technique [@buja2009statistical] on the fitted regression model to obtain null residuals.
+The null residuals are then used to make null plots and fed into the computer vision model to get predictions.
+The empirical distribution is constructed with the approximated distance.
 
@@
 
-Let $\hat{D}_{null}^{(i)}$ be the predicted distance of the $i$-th null plots, where $i = 1,...,n_{null}$ and $n_{null} \in \mathbb{N}^+$ is a sufficiently large number.
+Let $\hat{D}_{null}^{(i)}$ be the approximated distance of the $i$-th null plots, where $i = 1,...,n_{null}$ and $n_{null} \in \mathbb{N}^+$ is a sufficiently large number.
 Quantiles of the null distribution can then be estimated using the sample quantiles available in statistical software such as R.
@@
 In statistical testing, analysts often care about certain quantiles of the null distribution, such as the 90% quantile, the 95% quantile and the 99% quantile.
-These quantiles are used as thresholds to decide if $H_0$ needs to be rejected.
-For example, if $\hat{D}$ is greater than and equal to the 95% sample quantile of the null distribution $Q_{null}(0.95)$, we could say the predicted distance for the actual residual plot is significantly different from the predicted distance for null plots with 95% significance level.
+These quantiles are used as critical values to decide if $H_0$ needs to be rejected.
+For example, if $\hat{D}$ is greater than and equal to the 95% sample quantile $Q_{null}(0.95)$, we could say the approximated distance for the actual residual plot is significantly different from the approximated distance for null plots with 95% significance level.
 Based on our experience, in order to get a stable estimate of the 95% quantile, $n_{null}$ usually needs to be at least $100$.
@@
 Alternatively, a p-value can be used to represents the probability of observing an event equally or more extreme than the given event under $H_0$, and it can be estimated by $1/m\sum_{i=1}^{n_{null}}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)$.
+
+If precision in sample quantiles is not the main priority, using a pre-calculated table of quantiles is an available option.
+Such a table offers pre-determined quantiles for a specified number of observations.
+It is generated by assessing numerous null plots derived from various simulated regression models and averaging them.
+Essentially, this shifts the computational burden from the user to the developer.
 
@@
 But this approach is computationally very expensive because it requires $n_{boot} \times n_{null}$ times of residual plot assessment.
-In practice, $Q_{null}(0.95)$ will be used to replace $Q_{boot}^{(i)}(0.95)$ in the computation.
+In practice, $Q_{null}(0.95)$ can be used to replace $Q_{boot}^{(i)}(0.95)$ in the computation.
 
@@
 Consequently, the computer vision models developed in this study were trained using synthetic data.
-This approach provides us with precise label annotations.
-Additionally, it ensures a large and diverse training dataset, as we have control over the data generating process, and the simulation of the training data is relatively cost-effective.
+This approach provided us with precise label annotations.
+Additionally, it ensured a large and diverse training dataset, as we had control over the data generating process, and the simulation of the training data was relatively cost-effective.
 
@@
 
-The residuals and fitted values of the fitted model is obtained by regressing $\boldsymbol{y}$ on $\boldsymbol{x}_1$.
-This data generating process is adopted from @li2023plot where it was used to simulate residual plots with non-linearity and heteroskedasticity visual patterns for human subject experiments.
-
-In equation \ref{eq:data-sim}, predictor $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$ are independent of each other, so excluding $\boldsymbol{x}_2$ from the regression formula will not cause any problem.
-However, $\boldsymbol{z}$ and $\boldsymbol{w}$ are higher-order terms of $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$.
-If $\beta_2 \neq 0$, the regression model will suffer from non-linearity issues.
-Parameter $j$ is a shape parameter controlling the number of tuning points of the non-linearity pattern.
-Generally, greater values of $j$ will result in more tuning points.
+```{r factor}
+tibble(Factor = c("j", "a", "b", "$\\beta_1$", "$\\beta_2$", " $\\text{dist}_\\varepsilon$", "$\\text{dist}_{x1}$", "$\\text{dist}_{x2}$", "$\\sigma_{\\varepsilon}$", "$\\sigma_{X1}$", "$\\sigma_{X2}$", "n"),
+       Domain = c("\\{2, 3, ..., 18\\}", "[-1, 1]", "[0, 100]", "{0, 1}", "{0, 1}", "\\{discrete, uniform, normal, lognormal\\}", "\\{discrete, uniform, normal, lognormal\\}", "\\{discrete, uniform, normal, lognormal\\}", "[0.0625, 9]", "[0.3, 0.6]", "[0.3, 0.6]", "[50, 500]")) %>%
+  kableExtra::kable("latex", 
+                    escape = FALSE,
+                    booktabs = TRUE,
+                    caption = "Factors used in the data generating process for synthetic data simulation.
+Factor $j$ and $a$ controls the non-linearity shape and the heteroskedasticity shape respectively.
+Factor $b$, $\\sigma_\\varepsilon$ and $n$ control the signal strength.
+Factor $\\text{dist}_\\varepsilon$, $\\text{dist}_{x1}$ and $\\text{dist}_{x2}$ specifies the distribution of $\\varepsilon$, $X_1$ and $X_2$ respectively.")
+```
+
+The residuals and fitted values of the fitted model were obtained by regressing $\boldsymbol{y}$ on $\boldsymbol{x}_1$.
+If $\beta_1 \neq 0$, $\boldsymbol{x}_2$ was also included in the design matrix.
+This data generation process was adapted from @li2023plot, where it was utilized to simulate residual plots exhibiting non-linearity and heteroskedasticity visual patterns for human subject experiments.
+A summary of the factors utilized in Equation \ref{eq:data-sim} is provided in Table \ref{tab:factor}.
+
+In Equation \ref{eq:data-sim}, $\boldsymbol{z}$ and $\boldsymbol{w}$ represent higher-order terms of $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$, respectively.
+If $\beta_2 \neq 0$, the regression model will encounter non-linearity issues.
+Parameter $j$ serves as a shape parameter that controls the number of tuning points in the non-linear pattern.
+Typically, higher values of $j$ lead to an increase in the number of tuning points, as illustrated in Figure \ref{fig:different-j}.
+
+```{r different-j, fig.pos="!h", fig.cap = "Non-linearity forms generated for the synthetic data simulation. The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}
+
+set.seed(10086)
+
+# Data for shape 1
+dat_shape_1 <- phn_model(j = 2, include_x2 = FALSE, sigma = 0.05)$gen(500) %>%
+  mutate(j = 2)
+
+# Generate data for shape 2, 3 and 4. Reuse x and e.
+map_df(3:18, function(j) {
+  phn_model(j = j, include_x2 = FALSE, sigma = 0.05)$
+    gen(500, computed = select(dat_shape_1, x1, e)) %>%
+  mutate(j = j)
+}) %>%
+
+  # Combined with data for shape 1
+  bind_rows(dat_shape_1) %>%
+  mutate(j = factor(j)) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
+```
 
@@
 If $b \neq 0$ and $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$, the constant variance assumption will be violated.
-Parameter $a$ is a shape parameter controlling the location of the smallest variance in a residual plot.
-
-Non-normality violations are introduced by defining a non-normal distribution for $\boldsymbol{\varepsilon}$.
-
-- distribution of x1 and x2
-- summary table of all the factor values
-- example plots for the violations
+Parameter $a$ is a shape parameter controlling the location of the smallest variance in a residual plot as shown in Figure \ref{fig:different-a}.
+
+```{r different-a, fig.width=8/5*3*2, fig.height=6/4*3*2, fig.pos="!h", fig.cap = 'Heteroskedasticity forms generated for the synthetic data simulation.
+Different shapes are controlled by the continuous factor $a$ between -1 and 1.
+For $a = -1$, the residual plot exhibits a "left-triangle" shape.
+And for $a = 1$, the residual plot exhibits a "right-triangle" shape.
+'}
+
+set.seed(10085)
+
+# Generate data for a = -1
+dat_a_n1 <- phn_model(include_z = FALSE,
+                      include_x2 = FALSE,
+                      a = -1,
+                      b = 100)$gen(500) %>%
+  mutate(a = -1)
+
+# Generate data for other a
+map(c(-0.75, -0.5, -0.25, 0, 0.25, 0.5, 0.75, 1), function(a) {
+  phn_model(include_z = FALSE,
+            include_x2 = FALSE,
+            a = a,
+            b = 100)$gen(500) %>%
+  mutate(a = a)
+}) %>%
+
+  # Combined with data for a = -1
+  bind_rows(dat_a_n1) %>%
+  mutate(a = factor(a)) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~a, scales = "free", ncol = 3) +
+  xlab("Fitted values") +
+  ylab("Residuals")
+```
+
+```{r different-e, fig.pos="!h", fig.cap = 'Non-normality forms generated for the synthetic data simulation.
+Four different error distributions including discrete, lognormal, normal and uniform are considered.'}
+
+set.seed(10086)
+
+# Data for shape 1
+dat_shape_1 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_uniform(-1.4, 1.4))$gen(500) %>%
+  mutate(e_dist = "uniform")
+
+dat_shape_2 <- phn_model(include_z = FALSE, include_x2 = FALSE, sigma = 0.8)$gen(500) %>%
+  mutate(e_dist = "normal")
+
+dat_shape_3 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_lognormal(sigma = 0.6))$gen(500) %>%
+  mutate(e_dist = "lognormal")
+
+dat_shape_4 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_uniform_d(-1.4, 1.4, even = TRUE))$gen(500) %>%
+  mutate(e_dist = "discrete")
+
+# Generate data for shape 2, 3 and 4. Reuse x and e.
+bind_rows(dat_shape_1, dat_shape_2, dat_shape_3, dat_shape_4) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~e_dist, scales = "free", labeller = label_parsed, ncol = 2)
+```
+
+Non-normality violations arise from specifying a non-normal distribution for $\boldsymbol{\varepsilon}$.
+In the synthetic data simulation, four distinct error distributions are considered, including discrete, uniform, normal, and lognormal distributions, as presented in Figure \ref{fig:different-e}.
+Each distribution imparts unique characteristics to the residuals.
+The discrete error distribution introduces discreteness in residuals, while the lognormal distribution typically yields outliers.
+Uniform error distribution may result in residuals filling the entire space of the residual plot.
+All of these distributions exhibit visual distinctions from the normal error distribution.
+
+```{r different-j-x2, fig.pos="!h", fig.cap = "Residual plots of multiple linear regression models with non-linearity issues.
+The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.
+A second predictor $\\boldsymbol{x}_2$ is introduced to the regression model to create complex shapes."}
+
+set.seed(10086)
+
+# Data for shape 1
+dat_shape_1 <- phn_model(j = 2, include_x2 = TRUE, sigma = 0.05)$gen(500) %>%
+  mutate(j = 2)
+
+# Generate data for shape 2, 3 and 4. Reuse x and e.
+map_df(3:18, function(j) {
+  phn_model(j = j, include_x2 = TRUE, sigma = 0.05)$
+    gen(500, computed = select(dat_shape_1, x1, e)) %>%
+  mutate(j = j)
+}) %>%
+
+  # Combined with data for shape 1
+  bind_rows(dat_shape_1) %>%
+  mutate(j = factor(j)) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
+```
+
+```{r different-j-heter, fig.pos="!h", fig.cap = 'Residual plots of models violating both the non-linearity and the heteroskedasticity assumptions.
+The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$, and the "left-triangle" shape is introduced by setting $a = -1$.'}
+
+set.seed(10086)
+
+# Data for shape 1
+dat_shape_1 <- phn_model(j = 2, a = -1, b = 100, include_x2 = FALSE, sigma = 0.05)$gen(500) %>%
+  mutate(j = 2)
+
+# Generate data for shape 2, 3 and 4. Reuse x and e.
+map_df(3:18, function(j) {
+  phn_model(j = j, a = -1, b = 100, include_x2 = FALSE, sigma = 0.05)$
+    gen(500, computed = select(dat_shape_1, x1, e)) %>%
+  mutate(j = j)
+}) %>%
+
+  # Combined with data for shape 1
+  bind_rows(dat_shape_1) %>%
+  mutate(j = factor(j)) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
+```
+
+```{r different-e-heter, fig.pos="!h", fig.cap = 'Residual plots of models violating both the non-normality and the heteroskedasticity assumptions.
+The four shapes are generated by using four different error distributions including discrete, lognormal, normal and uniform, and the "left-triangle" shape is introduced by setting $a = -1$.
+'}
+
+set.seed(10085)
+
+# Data for shape 1
+dat_shape_1 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_uniform(-1.4, 1.4))$gen(500) %>%
+  mutate(e_dist = "uniform")
+
+dat_shape_2 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, sigma = 0.8)$gen(500) %>%
+  mutate(e_dist = "normal")
+
+dat_shape_3 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_lognormal(sigma = 0.6))$gen(500) %>%
+  mutate(e_dist = "lognormal")
+
+dat_shape_4 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_uniform_d(-1.4, 1.4, even = TRUE))$gen(500) %>%
+  mutate(e_dist = "discrete")
+
+# Generate data for shape 2, 3 and 4. Reuse x and e.
+bind_rows(dat_shape_1, dat_shape_2, dat_shape_3, dat_shape_4) %>%
+  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
+  facet_wrap(~e_dist, scales = "free", labeller = label_parsed, ncol = 2)
+```
+
+Equation \ref{eq:data-sim} accommodates the incorporation of the second predictor $\boldsymbol{x}_2$.
+Introducing it into the data generation process by setting $\beta_1 = 1$ significantly enhances the complexity of the shapes, as illustrated in Figure \ref{fig:different-j-x2}.
+In comparison to Figure \ref{fig:different-j}, Figure \ref{fig:different-j-x2} demonstrates that the non-linear shape resembles a surface rather than a single curve.
+This augmentation can facilitate the computer vision model in learning visual patterns from residual plots of the multiple linear regression model.
+
+In real-world analysis, it's not uncommon to encounter instances where multiple model violations coexist.
+In such cases, the residual plots often exhibit a mixed pattern of visual anomalies corresponding to different types of model violations.
+Figure \ref{fig:different-j-heter} and \ref{fig:different-e-heter} show the visual patterns of models with multiple model violations.
 
@@
 
-We have mentioned in Section \ref{introduction} that scagnostics are a set of manually designed visual feature extraction functions.
-Although our computer vision model will learn its own feature extraction function during the training process, it will be beneficial to utilize additional information provided by scagnostics to help computer vision model making more accurate predictions.
-
-For each generated residual plot, four scagnostics, namely, "Monotonic", "Sparse", "Splines" and "Striped" are computed using the `cassowaryr` R package (ref here).
-The computed measures along with the number of observations of the fitted model are provided as the second input of the computer vision model.
-Other scagnositcs are also informative, but unfortunately, they are unavailable due to a fatal bug in the compiled C program of the `interp` R package (ref here) that will crash the process in a unpredictable manner.
-For reproducibility, we do not include these scagnostics in the training data.
+In Section \ref{introduction}, we mentioned that scagnostics consist of a set of manually designed visual feature extraction functions.
+While our computer vision model will learn its own feature extraction function during training, leveraging additional information from scagnostics can enhance the model's predictive accuracy.
+
+For each generated residual plot, we computed four scagnostics – "Monotonic," "Sparse," "Splines," and "Striped" – using the cassowaryr R package [@mason2022cassowaryr].
+These computed measures, along with the number of observations from the fitted model, were provided as the second input for the computer vision model.
+Although other scagnostics are informative, they are currently unavailable due to a fatal bug in the compiled C program of the interp R package [@Albrecht2023interp], which may unpredictably crash the process.
+For reproducibility, we excluded these scagnostics from the training data.
 
@@
 
-In order to train a robust computer vision model, we intentionally control the distribution of the target variable $D$ of the training data.
-Making it uniform between $0$ and $7$.
-This is accomplished by preparing $50$ buckets each only accepts training samples with $D$  between $[7(i - 1)/49, 7i/49)$ for $i < 50$, where $i$ is the index of the $i$-th bucket.
-For the $50$-th bucket, any training samples with $D >= 7$ will be accepted.
-We have prepared 80000 training images, so each bucket can only contain $80000 \div 50 = 1600$ training samples.
-The simulator will repeatedly sample parameter values from the parameter sample, generate residuals and fitted values using the data generating process, compute the distance, and check if the sample can be accepted by the corresponding bucket, until all the buckets are full.
+To train a robust computer vision model, we deliberately controlled the distribution of the target variable $D$ in the training data.
+We ensured that it followed a uniform distribution between $0$ and $7$.
+This was achieved by organizing $50$ buckets, each exclusively accepting training samples with $D$ falling within the range $[7(i - 1)/49, 7i/49)$ for $i < 50$, where $i$ represents the index of the $i$-th bucket.
+For the $50$-th bucket, any training samples with $D \geq 7$ were accepted.
+
+With 80000 training images prepared, each bucket accommodated a maximum of $80000 \div 50 = 1600$ training samples.
+The simulator iteratively sampled parameter values from the parameter space, generated residuals and fitted values using the data generation process, computed the distance, and checked if the sample fitted within the corresponding bucket.
+This process continued until all buckets were filled.
+
+Similarly, we adopted the same methodology to prepare 8000 test images for performance evaluation and model diagnostics.
 
@@
 
-The architecture of the computer vision model is adopted from a existing architecture called VGG16 that proven to be performant in image classification [@simonyan2014very].
-The first layer is the input layer with shape $n \times h \times w \times 3$ which could accept $n$ RGB images, followed by a grayscale conversion layer using the luma formula under the CCIR 601 standard (ref here) to convert the colour image to an grey scale image.
-Grayscale is enough for our task since data points are plotted in black.
-Three different combinations of $h$ and $w$ are chosen to seek a sufficient large image resolution for this problem.
-This includes $32 \times 32$, $64 \times 64$ and $128 \times 128$.
-
-The processed image is then used as the input of the first convoluational block.
-The model has at most five consecutive convolutional blocks which is similar to the original VGG16 architecture.
-In each block, there are two 2D convoluational layers followed by two activation layers respectively, and there is a 2D max pooling layer follows the second activation layer.
-The 2D convolutional layer convolves the input with a fixed number of $3 \times 3$ convolution filters, and the 2D max pooling layer downsamples the input along its spatial dimensions by taking the maximum value over a $2 \times 2$ window for each channel of the input.
-The activation layer is set to be the rectified linear unit activation function (ReLu), which is a standard practice in deep learning.
-This introduces a non-linear transformation of the output of the 2D convolutional layer.
-Furthermore, to regularize the training, a batch normalization layer is added after each of the 2D convolutional layer and before the activation layer.
-A dropout layer is also appended at the end of each convolutional block to randomly set some inputs to zeros during training.
-
-The output of the last convolutional block will be summarised by a global max pooling layer or a global average pooling layer to gain a two-dimensional tensor.
-Since we want to utilize the information contained in scagnostics, the two-dimensional tensor is concatenated with an additional $n \times 5$ tensor, which contains the "Monotonic", "Sparse", "Splines" and "Striped" measures and the number of observations for $n$ residual plots.
-
-The concatenated tensor will be fed into the final prediction block.
-This block contains two fully-connected layers, the first one has at least $128$ units followed by a dropout layer.
-Sometimes, a batch normalization layer will be inserted between the fully-connected layer and the dropout layer.
-The second fully-connected layer has only one unit, serves as the output of the model.
-
-The model weights are randomly initialized and they are optimized by the Adam optimizer with the mean square error loss function.
-Therefore, we try to find a set of weights $\hat{\boldsymbol{\theta}}$ by
+The architecture of the computer vision model is adapted from a well-established architecture known as VGG16, which has demonstrated high performance in image classification [@simonyan2014very].
+
+The model begins with an input layer of shape $n \times h \times w \times 3$, capable of handling $n$ RGB images.
+This is followed by a grayscale conversion layer utilizing the luma formula under the Rec.
+601 standard, which converts the color image to grayscale.
+Grayscale suffices for our task since data points are plotted in black.
+We experiment with three combinations of $h$ and $w$: $32 \times 32$, $64 \times 64$, and $128 \times 128$, aiming to achieve sufficiently high image resolution for the problem at hand.
+
+The processed image is used as the input for the first convolutional block.
+The model comprises at most five consecutive convolutional blocks, mirroring the original VGG16 architecture.
+Within each block, there are two 2D convolutional layers followed by two activation layers, respectively.
+Subsequently, a 2D max-pooling layer follows the second activation layer.
+The 2D convolutional layer convolves the input with a fixed number of $3 \times 3$ convolution filters, while the 2D max-pooling layer downsamples the input along its spatial dimensions by taking the maximum value over a $2 \times 2$ window for each channel of the input.
+The activation layer employs the rectified linear unit (ReLU) activation function, a standard practice in deep learning, which introduces a non-linear transformation of the output of the 2D convolutional layer.
+Additionally, to regularize training, a batch normalization layer is added after each 2D convolutional layer and before the activation layer.
+Finally, a dropout layer is appended at the end of each convolutional block to randomly set some inputs to zero during training, further aiding in regularization.
+
+The output of the last convolutional block is summarized by either a global max pooling layer or a global average pooling layer, resulting in a two-dimensional tensor.
+To leverage the information contained in scagnostics, this tensor is concatenated with an additional $n \times 5$ tensor, which contains the "Monotonic," "Sparse," "Splines," and "Striped" measures, along with the number of observations for $n$ residual plots.
+
+The concatenated tensor is then fed into the final prediction block.
+This block consists of two fully-connected layers.
+The first layer contains at least $128$ units, followed by a dropout layer.
+Occasionally, a batch normalization layer is inserted between the fully-connected layer and the dropout layer for regularization purposes.
+The second fully-connected layer consists of only one unit, serving as the output of the model.
+
+The model weights $\boldsymbol{\theta}$ were randomly initialized and they were optimized by the Adam optimizer with the mean square error loss function
 
@@
 
-- a diagram for the model
-
 ## Training and hyperparameter tuning
 
-To obtain an optimal deep learning model, hyperparameters such as learning rate often needs to be tuned via a tuner.
-In this study, we used the Bayesian optimization tuner from the `keras-tuner` Python library (ref here) to tune the hyperparameters.
-The complete list of hyperparameters is provided in Table \ref{tab:hyperparameter}.
-
-The number of base filters determines the number of filters for the first 2D convolutional layer.
-In VGG16, the number of filters for the 2D convolutional layer in a block is 2 times the number in the previous block, except for the last block which has the same number of convolution filters as the previous one.
-This hyperparameter helps to control the complexity of the computer vision model.
-More base filters will result in more trainable parameters.
-
-Similarly, the number of units for the first fully-connected layer determines the complexity of the final prediction block. And more units will lead to more trainable parameters.
-
-Dropout rate and batch normalization are flexible hyperparameters which works jointly with other hyperparameters to smooth the training process.
-A high dropout rate is needed when the model learns too much noise from the training dataset.
-And a low dropout rate is needed when the model is complex and difficult to converge.
-Batch normalization on the other hand may be needed for solving the internal covariate shift problem due to randomness in the weight initialization and randomness in the input data.
-
-Having additional inputs including the scagnostics and the number of observations could be beneficial for making more accurate prediction.
-So we let the tuner to decide if these inputs are actually needed.
-
-Learning rate is one of the most important hyperparameters, and it determines the step size of the optimization algorithm.
-A high learning rate could help the model to avoid getting stuck in a local minimum, but it could also ruin the optimization process by jumping over the global minimum.
-A low learning rate could smooth the training process but it increases the time to converge and the chance of getting stuck in local minimum.
-
-The model is trained on the M3 high-performance computing platform (ref here) with TensorFlow (ref here) and Keras (ref here).
-During the training process, 80% of the training data are used as the actual training data, and 20% of the training data are reserved as validation data.
-We let the Bayesian optimization tuner runs 100 trials to find the best hyperparameter values according to the validation root mean square error.
-The tuner will restore the best epoch of the best model from the trials.
-Early stopping is also applied such that the training process will halt if the validation root mean square error does not improve for 50 epochs.
-The maximum allowed epochs is 2000 where no models reach.
+To achieve an optimal deep learning model, hyperparameters like the learning rate often need to be fine-tuned using a tuner.
+In our study, we utilized the Bayesian optimization tuner from the KerasTuner Python library [@omalley2019kerastuner] for this purpose.
+A comprehensive list of hyperparameters is provided in Table \ref{tab:hyperparameter}.
+
+The "number of base filters" determines the number of filters for the first 2D convolutional layer.
+In the VGG16 architecture, the number of filters for the 2D convolutional layer in a block is typically twice the number in the previous block, except for the last block, which maintains the same number of convolution filters as the previous one.
+This hyperparameter aids in controlling the complexity of the computer vision model.
+Higher numbers of base filters result in more trainable parameters.
+Likewise, the "number of units for the fully-connected layer" determines the complexity of the final prediction block.
+Increasing the number of units enhances model complexity, resulting in more trainable parameters.
+
+The dropout rate and batch normalization are flexible hyperparameters that work in conjunction with other parameters to facilitate smooth training.
+A higher dropout rate is necessary when the model tends to overfit the training dataset by learning too much noise.
+Conversely, a lower dropout rate is preferred when the model is complex and challenging to converge.
+Batch normalization, on the other hand, addresses the internal covariate shift problem arising from the randomness in weight initialization and input data.
+It helps stabilize and accelerate the training process by normalizing the activations of each layer.
+
+Additionally, incorporating additional inputs such as scagnostics and the number of observations can potentially enhance prediction accuracy.
+Therefore, we allowed the tuner to determine whether these inputs were necessary for optimal model performance.
+
+The learning rate is a crucial hyperparameter, as it dictates the step size of the optimization algorithm.
+A high learning rate can help the model avoid local minima but risks overshooting and missing the global minimum.
+Conversely, a low learning rate smoothens the training process but makes the convergence time longer and increases the likelihood of getting trapped in local minima.
+
+Our model was trained on the M3 high-performance computing platform, using TensorFlow and Keras.
+During training, 80% of the training data was utilized for actual training, while the remaining 20% was used as validation data.
+The Bayesian optimization tuner conducted 100 trials to identify the best hyperparameter values based on validation root mean square error.
+The tuner then restored the best epoch of the best model from the trials.
+Additionally, we applied early stopping, terminating the training process if the validation root mean square error fails to improve for 50 epochs.
+The maximum allowed epochs was set at 2000, although no models reached this threshold.
 
@@
 
-Based on the tuning process described in Section \ref{training-and-hyperparameter-tuning}, the optimized hyperparameter values are provided in Table \ref{tab:best-hyperparameter}.
-It can be observed that at least $32$ base filters are needed, and the preferable choice for the $64 \times 64$ model and the $128 \times 128$ model is $64$ base filters, which is the same as the original VGG16 architecture.
-The optimized values for dropout rate for convolutional blocks is around $0.4$.
-And applying batch normalization for convolutional blocks has a positive impact on the performance.
-All optimized models choose to preserve the additional inputs to reduce the validation error.
-The number of units needed for the fully-connected layer is $256$, which is not a very large number comparing to the VGG16 classifier, suggesting that the problem we are trying to solve is relatively easy.
-The optimized learning rates are usually smaller than the default value recommended by Keras which is 0.001.
+Based on the tuning process described in Section \ref{training-and-hyperparameter-tuning}, the optimized hyperparameter values are presented in Table \ref{tab:best-hyperparameter}.
+It was observable that a minimum of $32$ base filters was necessary, with the preferable choice being $64$ base filters for both the $64 \times 64$ and $128 \times 128$ models, mirroring the original VGG16 architecture.
+The optimized dropout rate for convolutional blocks hovered around $0.4$, and incorporating batch normalization for convolutional blocks proved beneficial for performance.
+
+All optimized models chose to retain the additional inputs, contributing to the reduction of validation error.
+The number of units required for the fully-connected layer was $256$, a relatively modest number compared to the VGG16 classifier, suggesting that the problem at hand was relatively straightforward.
+The optimized learning rates were typically smaller than the default value recommended by Keras, which is 0.001.
 
@@
 
-The training and test performance of optimized models with three different input sizes are provided in Table \ref{tab:performance}.
-Among these models, the $64 \times 64$ model and the $32 \times 32$ model are consistently having the best training and test performance respectively.
-The mean absolute error shows that the difference between $\hat{D}$ and $D$ is around $0.43$ on the test set, which is very small considering the range of $D$ is normally between $0$ and $7$.
-The high $R^2$ also suggests the prediction is mostly linearly correlated with the target.
-
-Figure \ref{fig:model-performance} displays a hexagonal heatmap for $D - \hat{D}$ vs $D$.
-For residual plots with $D$ close to zero, all the optimized models tends to over-predict $\hat{D}$, leading to negative residuals.
-When $D$ is between $1.5$ and $6$, the smooth curves indicate that $\hat{D}$ is not heavily biased, and there are some instances where $\hat{D}$ is severely under-predicted.
-And when $D$ is greater than $6$, there are more instances being under-predicted than over-predicted.
-
-<!-- It can be observed from Figure \ref{fig:model-performance-density} that the optimized models produced much more $\hat{D}$ at around $1$ than expected.  -->
-
-Given the model performance metrics, only the best model evaluated on the test set, i.e. the $32 \times 32$ model, will be used in the following analysis.
+The training and test performance for the optimized models with three different input sizes are summarized in Table \ref{tab:performance}.
+Among these models, the $64 \times 64$ model and the $32 \times 32$ model consistently exhibited the best training and test performance, respectively.
+The mean absolute error indicated that the difference between $\hat{D}$ and $D$ was approximately $0.43$ on the test set, a negligible deviation considering the normal range of $D$ typically falls between $0$ and $7$.
+The high $R^2$ values also suggested that the predictions were largely linearly correlated with the target.
+
+Figure \ref{fig:model-performance} presents a hexagonal heatmap for $D - \hat{D}$ versus $\hat{D}$.
+The red smoothing curves, fitted by generalized additive models [@hastie2017generalized], demonstrate that all the optimized models perform admirably on both the training and test sets.
+No structural issues are noticeable in Figure \ref{fig:model-performance}, but some minor issues regarding over-prediction and under-prediction are observed.
+
+The figure highlights that most under-predictions occurred when $\hat{D} < 3$, while over-predictions occurred predominantly when $3 < \hat{D} < 6$.
+To comprehend the reasons behind these deviations, we conducted a meta-analysis on $D - \hat{D}$ for all factors used in the data generating process.
+A regression tree was constructed using the rpart R package [@Terry2022rpart].
+The results of the regression tree are provided in the Appendix.
+
+The meta-analysis revealed that most issues arose from non-linearity problems and the presence of a second predictor in the regression model.
+When the error distribution had a very small variance, all computer vision models tended to under-predict the distance.
+Conversely, when the error distribution had a large variance, all models tended to over-predict the distance.
+Therefore, Additionally, for input images representing null plots, it was expected that the models will over-predict the distance, as explained in Section \ref{null-distribution-of-the-approximated-distance}.
+
+Since most of the deviation stemmed from the presence of non-linearity violations, to further investigate this, we split the test set and re-evaluated the performance, as detailed in Table \ref{tab:performance-sub}.
+It was evident that metrics for null plots were notably worse compared to other categories.
+Furthermore, residual plots solely exhibiting non-normality issues were the easiest to predict, with very low test root mean square error (RMSE) around $0.3$, very low mean absolute error (MAE) around $0.2$, and very high $R^2$ around $0.94$.
+Residual plots with non-linearity issues were more challenging to assess than those with heteroskedasticity or non-normality issues.
+Assessing residual plots with heteroskedasticity was not as difficult as assessing those with non-linearity issues.
+When multiple violations were introduced to a residual plot, the performance metrics typically lay between the metrics for each individual violation.
+
+Based on the model performance metrics, we will only use the best-performing model evaluated on the test set, namely the $32 \times 32$ model, for the subsequent analysis.
 
@@
                     col.names = c("", "RMSE", "$R^2$", "MAE", "Huber loss"),
-                    caption = "The training and test performance of three optimized models with different input sizes.") %>%
+                    caption = "The training and test performance of three optimized models with different input sizes. The best metrics are colored in red.") %>%
   kableExtra::pack_rows("Training set", 1, 3) %>%
-  kableExtra::pack_rows("Test set", 4, 6)
-```
-
-```{r model-performance, fig.cap = "Hexagonal heatmap for residuals vs distance on training and test data for three optimized models with different input sizes.
-The blue lines are smoothing curves produced by fitting gnealized additive models.", dev = 'png', dpi = 300}
-train_pred %>% 
+  kableExtra::pack_rows("Test set", 4, 6) %>%
+  kableExtra::column_spec(2:5, color = c("black", "red", "black",
+                                       "red", "black", "black"))
+
+```
+
+```{r model-performance, fig.pos="!h", fig.cap = "Hexagonal heatmap for residuals vs predicted distance on training and test data for three optimized models with different input sizes.
+The red lines are smoothing curves produced by fitting gnealized additive models.
+The area over the zero line in light pink indicates under-prediction, and the area under the zero line in light green indicates over-prediction.", dev = 'png', dpi = 300}
+model_pred <- train_pred %>% 
   left_join(meta) %>%
@@
               left_join(meta) %>%
-              mutate(type = "test")) %>%
+              mutate(type = "test"))
+
+model_pred %>%
   ggplot() +
-  geom_hline(yintercept = 0) +
+  geom_hline(yintercept = 0, alpha = 0.5) +
   # geom_point(aes(effect_size, effect_size - vss), alpha = 0.03) +
-  geom_hex(aes(effect_size, effect_size - vss), bins = 15) +
-  geom_smooth(aes(effect_size, effect_size - vss), se = FALSE) +
+  annotate("rect", 
+           ymin = min(model_pred$effect_size - model_pred$vss),
+                ymax = 0,
+                xmin = min(model_pred$vss),
+                xmax = max(model_pred$vss),
+            fill = "#66c2a5",
+            alpha = 0.3) +
+  geom_point(data = NULL, aes(5, 0, col = "over-prediction"), shape = 15) +
+  annotate("rect", 
+           ymin = 0,
+                ymax = max(model_pred$effect_size - model_pred$vss),
+                xmin = min(model_pred$vss),
+                xmax = max(model_pred$vss),
+            fill = "#fc8d62",
+            alpha = 0.3) +
+  geom_point(data = NULL, aes(5, 0, col = "under-prediction"), shape = 15) +
+  geom_hex(aes(effect_size, effect_size - vss), bins = 25) +
+  geom_smooth(aes(vss, effect_size - vss), se = FALSE, col = "#e34a33") +
   facet_grid(type ~ res) +
@@
   ylab(expression(D - hat(D))) +
-  xlab("D") +
+  xlab(expression(hat(D))) +
   theme_light() +
-  scale_fill_continuous(limits = c(0, 400))
-```
-
-```{r model-performance-density, fig.cap = "Density plot of predicted distance on training and test data for three optimized models with different input sizes.
-The density drawn in red is the density of the distance.", eval = FALSE}
-train_pred %>% 
-  left_join(meta) %>%
-  mutate(type = "train") %>%
-  bind_rows(test_pred %>% 
-              left_join(meta) %>%
-              mutate(type = "test")) %>%
-  ggplot() +
-  geom_density(aes(effect_size, col = "D")) +
-  geom_density(aes(vss, col = "hat_D")) +
-  facet_grid(type ~ res) +
-  scale_colour_manual(values = c("red", "blue"), labels = expression(D, hat(D))) +
-  labs(col = "") +
-  theme_light() +
-  xlab(expression(hat(D)))
+  scale_fill_continuous(limits = c(1, 200), trans = "log10", low = "#56B1F7", high = "#132B43") +
+  scale_color_manual(values = c("under-prediction" = scales::alpha("#fc8d62", 0.3),
+                                "over-prediction" = scales::alpha("#66c2a5", 0.3)),
+                     breaks = c("under-prediction", "over-prediction")) +
+  guides(col = guide_legend(title = NULL, override.aes = list(size = 8), order = 1))
+```
+
+```{r eval = FALSE}
+
+ # 1) root 8000 3466.12800  0.04808905  
+ #   2) e_sigma>=4.108756 3288 1146.50300 -0.09147268  
+ #     4) include_non_normal=FALSE 1983  817.30850 -0.22151770  
+ #       8) include_heter=FALSE 1210  416.16970 -0.39866830  
+ #        16) include_z=FALSE 49   20.28332 -1.22777500 * (only non-normal and large error sigma -> large over-prediction)
+ #        17) include_z=TRUE 1161  360.78130 -0.36367590 *
+ #       9) include_heter=TRUE 773  303.72640  0.05578156 *
+ #     5) include_non_normal=TRUE 1305  244.69960  0.10613590 *
+ #   3) e_sigma< 4.108756 4712 2210.89500  0.14547420  
+ #     6) include_z=FALSE 2030  568.99490 -0.04100357 *
+ #     7) include_z=TRUE 2682 1517.87900  0.28661890  
+ #      14) include_x2=FALSE 1432  377.35510  0.07228115 *
+ #      15) include_x2=TRUE 1250  999.37090  0.53216420  
+ #        30) e_sigma>=0.1149429 1202  829.80300  0.49933640 *
+ #        31) e_sigma< 0.1149429 48  135.83470  1.35422700  (non-linearity, second predictor, small error sigma)
+ #          62) measure_striped< 0.44 40   54.18083  0.92957180 *
+ #          63) measure_striped>=0.44 8   38.37423  3.47750200 * (high discreteness -> large under-prediction)
+
+# n= 8000 
+# 
+# node), split, n, deviance, yval
+#       * denotes terminal node
+# 
+#     1) root 8000 3466.1280000  0.04808905  
+#       2) e_sigma>=4.108756 3288 1146.5030000 -0.09147268  
+#         4) include_non_normal=FALSE 1983  817.3085000 -0.22151770  
+#           8) include_heter=FALSE 1210  416.1697000 -0.39866830  
+#            16) include_z=FALSE 49   20.2833200 -1.22777500  
+#              32) measure_sparse< 0.3630358 7    8.9396360 -2.12605300 * (large e_sigma, null plots, not very sparse -> large over-prediction)
+#              33) measure_sparse>=0.3630358 42    4.7539680 -1.07806200 * (large e_sigma, null plots, sparse -> large over-prediction)
+#            17) include_z=TRUE 1161  360.7813000 -0.36367590  
+#              34) measure_splines>=0.02606246 376  134.3999000 -0.59957120  
+#                68) n< 143.5 116   39.6936700 -0.84719690  
+#                 136) measure_monotonic>=0.07456592 7    1.3762680 -1.55893700 * (non-linearity only, small n, large e_sigma -> large over-prediction)
+#                 137) measure_monotonic< 0.07456592 109   34.5436600 -0.80148880 *
+#                69) n>=143.5 260   84.4197800 -0.48909200  
+#                 138) e_sigma>=7.367384 63   15.0193100 -0.79693070 *
+#                 139) e_sigma< 7.367384 197   61.5210600 -0.39064610  
+#                   278) include_x2=FALSE 90   19.5837000 -0.53918530 *
+#                   279) include_x2=TRUE 107   38.2813500 -0.26570670  
+#                     558) measure_sparse< 0.5586777 84   27.3861700 -0.37645850 *
+#                     559) measure_sparse>=0.5586777 23    6.1018500  0.13877850 *
+#              35) measure_splines< 0.02606246 785  195.4365000 -0.25068660  
+#                70) e_sigma>=5.64113 529  100.6972000 -0.37916600  
+#                 140) n< 109 41   11.8877000 -0.73213180 *
+#                 141) n>=109 488   83.2723700 -0.34951110  
+#                   282) e_sigma>=7.90547 154   26.5350800 -0.48055150 *
+#                   283) e_sigma< 7.90547 334   52.8735800 -0.28909130 *
+#                71) e_sigma< 5.64113 256   67.9628700  0.01480427  
+#                 142) include_x2=FALSE 113   23.8143300 -0.15341150 *
+#                 143) include_x2=TRUE 143   38.4243300  0.14773000  
+#                   286) n< 72.5 7    2.6335630 -0.69382120 *
+#                   287) n>=72.5 136   30.5781400  0.19104510 *
+#           9) include_heter=TRUE 773  303.7264000  0.05578156  
+#            18) b< 32.23487 487  221.2986000 -0.09202344  
+#              36) b< 0.2107948 67   26.8944500 -0.51926710  
+#                72) n< 126 7    3.3765570 -1.24628700 * (large e_sigma, heteroskedasticity, small b, small n -> over-prediction)
+#                73) n>=126 60   19.3863300 -0.43444810 *
+#              37) b>=0.2107948 420  180.2232000 -0.02386791  
+#                74) measure_splines>=0.05482434 223   77.7820400 -0.18200180  
+#                 148) b< 7.647324 119   51.8788300 -0.32137070  
+#                   296) measure_sparse>=0.4117995 111   43.0150500 -0.38433220 *
+#                   297) measure_sparse< 0.4117995 8    2.3184780  0.55221930 *
+#                 149) b>=7.647324 104   20.9469900 -0.02253160 *
+#                75) measure_splines< 0.05482434 197   90.5523700  0.15513650  
+#                 150) b>=5.165768 63   26.5340300 -0.02749363  
+#                   300) include_x2=FALSE 35    7.9867080 -0.19247040 *
+#                   301) include_x2=TRUE 28   16.4039600  0.17872730  
+#                     602) measure_monotonic< 0.03127537 21    4.3016860 -0.07172379 *
+#                     603) measure_monotonic>=0.03127537 7    6.8333090  0.93008060 *
+#                 151) b< 5.165768 134   60.9291300  0.24099990  
+#                   302) x2_sigma< 0.5865284 126   52.7919200  0.19154170  
+#                     604) x2_k>=5.5 106   43.2640200  0.11519510  
+#                      1208) j< 17.5 93   38.5219100  0.05750563  
+#                        2416) e_even=TRUE 41   15.2536200 -0.17697420  
+#                          4832) measure_sparse>=0.5392181 29    8.8138120 -0.40043160 *
+#                          4833) measure_sparse< 0.5392181 12    1.4922590  0.36304760 *
+#                        2417) e_even=FALSE 52   19.2367200  0.24238400  
+#                          4834) measure_splines>=0.01044483 31    7.8888290 -0.02089042 *
+#                          4835) measure_splines< 0.01044483 21    6.0272570  0.63102720 *
+#                      1209) j>=17.5 13    2.2184070  0.52789690 *
+#                     605) x2_k< 5.5 20    5.6354220  0.59617870 *
+#                   303) x2_sigma>=0.5865284 8    2.9747040  1.01996600 * (large e_sigma, small spline dependence, heteroskedasticity, large x2_sigma -> large under-prediction)
+#            19) b>=32.23487 286   53.6723400  0.30746350  
+#              38) include_x2=FALSE 139   17.0423600  0.13425010 *
+#              39) include_x2=TRUE 147   28.5161400  0.47125030 *
+#         5) include_non_normal=TRUE 1305  244.6996000  0.10613590  
+#          10) measure_splines>=0.08262098 455   47.2275500  0.00622856 *
+#          11) measure_splines< 0.08262098 850  190.4994000  0.15961570  
+#            22) include_heter=FALSE 594   22.8029300  0.08838938 *
+#            23) include_heter=TRUE 256  157.6908000  0.32488300  
+#              46) measure_striped>=0.47 185   31.8984100  0.17734340 *
+#              47) measure_striped< 0.47 71  111.2723000  0.70931710  
+#                94) e_even=FALSE 32   18.0084200  0.22023240  
+#                 188) a>=-0.205314 21    5.0433870 -0.06070074 *
+#                 189) a< -0.205314 11    8.1435320  0.75655920 *
+#                95) e_even=TRUE 39   79.3286800  1.11061700  
+#                 190) b< 2.373839 11    0.2423938  0.06215229 *
+#                 191) b>=2.373839 28   62.2437600  1.52251400  (discrete e_dist,  high discreteness, heteroskedasticity, small splines dependence, non-normality, large e_sigma -> large under-prediction)
+#                   382) x2_dist=even_discrete,lognormal,uniform 19   33.0782400  1.01336800 *
+#                   383) x2_dist=normal 9   13.8421700  2.59737800 *
+#       3) e_sigma< 4.108756 4712 2210.8950000  0.14547420  
+#         6) include_z=FALSE 2030  568.9949000 -0.04100357  
+#          12) b< 28.58967 1159  316.0764000 -0.12131010  
+#            24) include_non_normal=FALSE 441  185.6341000 -0.33971870  
+#              48) include_heter=FALSE 60    8.3705760 -1.15981800 * (null plots, small e_sigma -> large over-prediction)
+#              49) include_heter=TRUE 381  130.5548000 -0.21056910  
+#                98) b< 0.4619795 85   22.8319800 -0.60512210 *
+#                99) b>=0.4619795 296   90.6909300 -0.09726844  
+#                 198) b>=1.080411 252   72.3919200 -0.15604160  
+#                   396) b< 2.878015 65   20.2765800 -0.37135770 *
+#                   397) b>=2.878015 187   48.0544100 -0.08119905 *
+#                 199) b< 1.080411 44   12.4430700  0.23934130 *
+#            25) include_non_normal=TRUE 718   96.4846700  0.01283772 *
+#          13) b>=28.58967 871  235.4979000  0.06585671  
+#            26) include_heter=FALSE 361  111.6977000 -0.10528680  
+#              52) include_non_normal=FALSE 46    6.2078660 -1.13547200 * (null plots, small e_sigma -> large over-prediction)
+#              53) include_non_normal=TRUE 315   49.5417500  0.04515286  
+#               106) e_sigma>=0.2273127 247   17.7293100 -0.01172324 *
+#               107) e_sigma< 0.2273127 68   28.1111100  0.25174690  
+#                 214) measure_splines>=0.03710735 23   10.7097800 -0.09471551 *
+#                 215) measure_splines< 0.03710735 45   13.2294000  0.42882770 *
+#            27) include_heter=TRUE 510  105.7420000  0.18699950  
+#              54) include_non_normal=TRUE 235   41.6282300  0.05417806 *
+#              55) include_non_normal=FALSE 275   56.4252500  0.30050140  
+#               110) include_x2=FALSE 137   20.9155600  0.17983570 *
+#               111) include_x2=TRUE 138   31.5346400  0.42029280  
+#                 222) b< 70.08912 96   17.2702500  0.30834030 *
+#                 223) b>=70.08912 42   10.3110100  0.67618420 *
+#         7) include_z=TRUE 2682 1517.8790000  0.28661890  
+#          14) include_x2=FALSE 1432  377.3551000  0.07228115  
+#            28) measure_splines>=0.07058104 731  178.2000000 -0.01483915  
+#              56) e_sigma>=2.44726 138   40.9281500 -0.18500620  
+#               112) include_heter=FALSE 59   25.9010700 -0.41304000  
+#                 224) include_non_normal=FALSE 38   17.2724700 -0.68141390  
+#                   448) measure_splines>=0.1003984 20    5.6757540 -1.07654300 * (non-linearity only, moderate e_sigma -> large over-prediction)
+#                   449) measure_splines< 0.1003984 18    5.0046810 -0.24238140 *
+#                 225) include_non_normal=TRUE 21    0.9391262  0.07258882 *
+#               113) include_heter=TRUE 79    9.6678490 -0.01470237 *
+#              57) e_sigma< 2.44726 593  132.3458000  0.02476126  
+#               114) e_sigma>=0.158351 527  105.6393000 -0.00147316 *
+#               115) e_sigma< 0.158351 66   23.4476700  0.23423920  
+#                 230) b< 71.04535 55   15.9627800  0.12578350 *
+#                 231) b>=71.04535 11    3.6032230  0.77651720 *
+#            29) measure_splines< 0.07058104 701  187.8212000  0.16312990  
+#              58) n< 62.5 33   14.3260500 -0.31035730 *
+#              59) n>=62.5 668  165.7314000  0.18652070 *
+#          15) include_x2=TRUE 1250  999.3709000  0.53216420  
+#            30) e_sigma>=0.1149429 1202  829.8030000  0.49933640  
+#              60) e_sigma>=2.906419 328  129.1838000  0.27288400  
+#               120) measure_splines>=0.04413018 134   50.5691600  0.09851024  
+#                 240) b< 2.910823 23   10.1825400 -0.20213530  
+#                   480) x2_k>=6.5 15    4.3945260 -0.49662390 *
+#                   481) x2_k< 6.5 8    2.0480610  0.35003080 *
+#                 241) b>=2.910823 111   37.8769400  0.16080620  
+#                   482) b>=7.320158 97   22.8753700  0.08661750 *
+#                   483) b< 7.320158 14   10.7686400  0.67482760 *
+#               121) measure_splines< 0.04413018 194   71.7259100  0.39332780  
+#                 242) j< 7.5 60   32.0807400  0.19368700  
+#                   484) x2_sigma< 0.343438 10   12.3504400 -0.51629950 *
+#                   485) x2_sigma>=0.343438 50   13.6813400  0.33568430 *
+#                 243) j>=7.5 134   36.1830100  0.48271920 *
+#              61) e_sigma< 2.906419 874  677.4868000  0.58432080  
+#               122) measure_splines>=0.04519141 613  424.1886000  0.45960850  
+#                 244) n>=202.5 393  209.4163000  0.32432050  
+#                   488) include_non_normal=TRUE 188   90.6250100  0.18780810  
+#                     976) x2_k< 7.5 101   41.5156400  0.04431936 *
+#                     977) x2_k>=7.5 87   44.6157600  0.35438700  
+#                      1954) measure_sparse>=0.391553 80   35.8617500  0.29057300 *
+#                      1955) measure_sparse< 0.391553 7    4.7050420  1.08369000 * (not sparse, non-normality, large n, small e_sigma, second predictor, non-linearity -> large under-prediction)
+#                   489) include_non_normal=FALSE 205  112.0748000  0.44951240  
+#                     978) e_sigma>=0.2441489 175   86.5106600  0.36295000  
+#                      1956) measure_splines>=0.07729252 127   55.6660800  0.25807550 *
+#                      1957) measure_splines< 0.07729252 48   25.7519700  0.64043040  
+#                        3914) b>=21.95702 24    7.5694370  0.34280130 *
+#                        3915) b< 21.95702 24   13.9305500  0.93805950  
+#                          7830) x2_dist=even_discrete,uniform 13    2.6412500  0.51209270 *
+#                          7831) x2_dist=lognormal,normal 11    6.1427810  1.44147500 *
+#                     979) e_sigma< 0.2441489 30   16.6036900  0.95445990  
+#                      1958) a< -0.6962934 8    1.2723410  0.16916930 *
+#                      1959) a>=-0.6962934 22    8.6039200  1.24002000 *
+#                 245) n< 202.5 220  194.7300000  0.70128210  
+#                   490) measure_sparse>=0.6256282 81   38.6326300  0.36079080  
+#                     980) measure_monotonic>=0.237124 8    0.2987567 -0.26782100 *
+#                     981) measure_monotonic< 0.237124 73   34.8262100  0.42967980 *
+#                   491) measure_sparse< 0.6256282 139  141.2344000  0.89969790  
+#                     982) n>=97 88   72.2102700  0.72131320  
+#                      1964) x1_sigma>=0.4241299 54   26.0940300  0.51943470  
+#                        3928) e_sigma>=0.2485052 45   19.5310800  0.39828410  
+#                          7856) x1_sigma< 0.5588507 38   11.8880800  0.26476900 *
+#                          7857) x1_sigma>=0.5588507 7    3.2882920  1.12308000 *
+#                        3929) e_sigma< 0.2485052 9    2.6000360  1.12518800 *
+#                      1965) x1_sigma< 0.4241299 34   40.4201400  1.04194400  
+#                        3930) j< 7.5 11    8.6545230  0.32377180 *
+#                        3931) j>=7.5 23   23.3787300  1.38541700  
+#                          7862) b< 20.93247 14   10.8261000  0.95400130 *
+#                          7863) b>=20.93247 9    5.8936770  2.05650900 * (moderate b, large j, small x1_sigma, small n, small sparse, second predictor, non-linearity -> large under-prediction)
+#                     983) n< 97 51   61.3921000  1.20749900  
+#                      1966) e_dist=uniform 16    6.5588030  0.49547860 *
+#                      1967) e_dist=even_discrete,lognormal 35   43.0135900  1.53299400  
+#                        3934) measure_monotonic>=0.04508442 17   19.7133200  0.99850550 *
+#                        3935) measure_monotonic< 0.04508442 18   13.8570300  2.03778800 *
+#               123) measure_splines< 0.04519141 261  221.3718000  0.87722750  
+#                 246) measure_sparse>=0.654522 53   26.6382500  0.36906170  
+#                   492) measure_striped>=0.57 11    2.8015390 -0.20941000 *
+#                   493) measure_striped< 0.57 42   19.1917400  0.52056620 *
+#                 247) measure_sparse< 0.654522 208  177.5598000  1.00671200  
+#                   494) e_sigma>=2.064604 95   37.3057400  0.73978960  
+#                     988) n< 94.5 10    2.7534790  0.01814227 *
+#                     989) n>=94.5 85   28.7318400  0.82468930  
+#                      1978) measure_sparse< 0.3468051 11    2.5461540  0.27928850 *
+#                      1979) measure_sparse>=0.3468051 74   22.4272100  0.90576240 *
+#                   495) e_sigma< 2.064604 113  127.7952000  1.23111600  
+#                     990) measure_splines>=0.03040418 40   30.6303500  0.78917000  
+#                      1980) include_non_normal=TRUE 16    4.6253120  0.29159560 *
+#                      1981) include_non_normal=FALSE 24   19.4029000  1.12088600 *
+#                     991) measure_splines< 0.03040418 73   85.0712900  1.47327800  
+#                      1982) b< 90.99624 66   59.3920200  1.35710600  
+#                        3964) a>=-0.6695046 52   41.5200500  1.19796400  
+#                          7928) x2_sigma< 0.4287885 20   11.8256700  0.80120150  
+#                           15856) measure_splines>=0.009178431 7    2.5265790  0.21136300 *
+#                           15857) measure_splines< 0.009178431 13    5.5523770  1.11880700 *
+#                          7929) x2_sigma>=0.4287885 32   24.5782200  1.44594000  
+#                           15858) x2_sigma>=0.4851646 24   12.4345200  1.18424600  
+#                             31716) e_dist=uniform 7    0.5259639  0.41482310 *
+#                             31717) e_dist=even_discrete,lognormal 17    6.0580960  1.50106700 *
+#                           15859) x2_sigma< 0.4851646 8    5.5692240  2.23102400 * (small x2_sigma, small b, small splines dependence, small e_sigma, small sparse, second predictor, non-linearity -> large under-prediction)
+#                        3965) a< -0.6695046 14   11.6634200  1.94820600 *
+#                      1983) b>=90.99624 7   16.3901800  2.56861600 * (large b, small splines dependence, small e_sigma, small sparse, second predictor, non-linearity -> large under-prediction)
+#            31) e_sigma< 0.1149429 48  135.8347000  1.35422700  
+#              62) measure_striped< 0.44 40   54.1808300  0.92957180  
+#               124) include_non_normal=TRUE 21   14.0872600  0.39910760  
+#                 248) include_heter=TRUE 13    5.7917480  0.06683663 *
+#                 249) include_heter=FALSE 8    4.5279800  0.93904800 *
+#               125) include_non_normal=FALSE 19   27.6530600  1.51587400 *
+#              63) measure_striped>=0.44 8   38.3742300  3.47750200 * (high discreteness, small e_sigma, second predictor, non-linearity -> large under-prediction)
+
+model_pred %>%
+  filter(res == 32L) %>%
+  filter(type == "test") %>%
+  mutate(.resid = effect_size - vss) %>%
+  mutate(include_z = factor(include_z),
+         include_heter = factor(include_heter),
+         include_x2 = factor(include_x2),
+         include_non_normal = factor(include_non_normal),
+         e_even = factor(e_even),
+         x1_even = factor(x1_even),
+         x2_even = factor(x2_even)) %>%
+  select(-filename, -vss, -res, -plot_uid, -data_type, -effect_size, -type) %>%
+  rpart::rpart(.resid ~ ., data = ., control = rpart::rpart.control(cp = 0.001), model = TRUE) -> test_tree
+```
+
+```{r performance-sub}
+
+model_pred %>%
+  filter(res == 32L) %>%
+  group_by(type, include_non_normal, include_heter, include_z) %>%
+  summarise(n = n(), 
+            RMSE = yardstick::rmse_vec(effect_size, vss),
+            R2 = yardstick::rsq_vec(effect_size, vss),
+            MAE = yardstick::mae_vec(effect_size, vss),
+            HUBER = yardstick::huber_loss_vec(effect_size, vss)) %>%
+  mutate(violations = ifelse(include_z, "non-linearity", "null")) %>%
+  mutate(violations = ifelse(include_heter, glue("{violations} + heteroskedasticity"), violations)) %>%
+  mutate(violations = ifelse(include_non_normal, glue("{violations} + non-normality"), violations)) %>%
+  mutate(violations = gsub("null \\+ ", "", violations)) %>%
+  ungroup() %>%
+  select(type, violations, n, RMSE, R2, MAE, HUBER) %>%
+  arrange(desc(type)) %>%
+  select(-type) %>%
+  mutate(across(RMSE:MAE, ~format(.x, digits = 3))) %>%
+  mutate(across(HUBER, ~format(.x, digits = 2))) %>%
+  mutate(violations = ifelse(violations == "null", "no violations", violations)) %>%
+  mutate(R2 = gsub(".*NA.*", "", R2)) %>%
+  kableExtra::kable("latex", 
+                    booktabs = TRUE,
+                    escape = FALSE,
+                    col.names = c("Violations", "\\#samples", "RMSE", "$R^2$", "MAE", "Huber loss"),
+                    caption = "The training and test performance of the $32 \\times 32$ model presented with different model violations.  The best metrics are colored in red.") %>%
+  kableExtra::kable_styling(latex_options = c("scale_down")) %>%
+  kableExtra::pack_rows("Training set", 1, 8) %>%
+  kableExtra::pack_rows("Test set", 9, 16) %>%
+  kableExtra::column_spec(3, color = c(rep("black", 4), "red", rep("black", 3),
+                                       rep("black", 4), "red", rep("black", 3))) %>%
+  kableExtra::column_spec(4, color = c(rep("black", 4), "red", rep("black", 3),
+                                       rep("black", 4), "red", rep("black", 3))) %>%
+  kableExtra::column_spec(5, color = c(rep("black", 4), "red", rep("black", 3),
+                                       rep("black", 4), "red", rep("black", 3))) %>%
+  kableExtra::column_spec(6, color = c(rep("black", 4), "red", rep("black", 3),
+                                       rep("black", 4), "red", rep("black", 3)))
 ```
@@
 
-Figure \ref{fig:rd-human} displays the $D - \hat{D}$ vs $D$ plot for actual residual plots used in the experiment.
-The behaviour of the prediction is roughly the same as what we can be observed from the evaluation of the test data.
-The model over-predicts $\hat{D}$ when $D$ is close to zero, and under-predicts $\hat{D}$ when $D$ is greater than $6$.
-However, for residual plots with non-linearity issues, we do not observe the over-prediction for $D > 6$ because there are not many cases in that area.
-And the over-prediction happens at around $2 < D < 4$.
-
 For null plots, a histogram of the prediction is provided in Figure \ref{fig:hist-null-human}.
@@
 And as $\hat{D}$ increases, the number of predictions decreases.
-There are some outliers with $\hat{D} > 5$.
-
-```{r rd-human, fig.cap = "Residuals vs distance plot on lineups used in the human subject experiment. The blue line is a smoothing curve produced by fitting a gnealized additive model."}
+The 90%, 95% and 99% sample quantiles are at around
+
+```{r rd-human}
 vss_32 <- readRDS(here::here("paper/data/vss_32.rds"))
@@
   left_join(readRDS(here::here("paper/data/actual_ss.rds")))
-
-experiment %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  ggplot() +
-  geom_hline(yintercept = 0) +
-  geom_point(aes(actual_ss, actual_ss - vss), alpha = 0.3) +
-  geom_smooth(aes(actual_ss, actual_ss - vss), se = FALSE) +
-  coord_fixed() +
-  ylab(expression(D - hat(D))) +
-  xlab("D") +
-  theme_light() +
-  facet_wrap(~type)
 ```
@@
 
-```{r rd-uniform-human, fig.cap = "Residuals vs distance plot on lineups with uniform predictor distribution used in the human subject experiment.
-The blue line is a smoothing curve produced by fitting a gnealized additive model.", eval = FALSE}
-experiment %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  filter(x_dist == "uniform") %>%
-  ggplot() +
-  geom_hline(yintercept = 0) +
-  geom_point(aes(actual_ss, actual_ss - vss), alpha = 0.6) +
-  geom_smooth(aes(actual_ss, actual_ss - vss), se = FALSE) +
-  coord_fixed() +
-  ylab(expression(D - hat(D))) +
-  xlab("D") +
-  theme_light() +
-  facet_wrap(~type)
-```
-
-```{r hist-null-human, fig.cap = "Histogram of predicted distance on null plots used in the human subject experiment."}
+```{r hist-null-human, fig.cap = "Histogram of predicted distance on null plots used in the human subject experiment. Sample quantiles are drawn in different colors."}
 lineup_vss <- readRDS(here::here("paper/data/lineup_vss.rds"))
@@
   geom_histogram(aes(vss)) +
+  geom_vline(aes(xintercept = quantile(vss, 0.9), col = "Q(0.90)")) +
+  geom_vline(aes(xintercept = quantile(vss, 0.95), col = "Q(0.95)")) +
+  geom_vline(aes(xintercept = quantile(vss, 0.99), col = "Q(0.99)")) +
   xlab(expression(hat(D))) +
   theme_light() +
-  facet_wrap(~type)
+  facet_wrap(~type) +
+  labs(col = "")
 ```
@@
 vi_lineup <- get_vi_lineup()
-```
-
-```{r fig.height = 7/4*2, fig.width = 7/5*3}
-lineup_vss %>%
-  filter(null) %>%
-  left_join(vi_survey %>%
-              group_by(unique_lineup_id) %>% 
-              summarise(type = first(type), 
-                        attention_check = first(attention_check),
-                        null_lineup = first(null_lineup))) %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  filter(vss > 5) -> human_null_outliers
-
-vi_lineup[human_null_outliers$unique_lineup_id] %>%
-  map2_df(human_null_outliers$pos, 
-          ~.x$data %>% 
-            filter(k == .y) %>% 
-            mutate(unique_lineup_id = .x$metadata$name)) %>%
-  ggplot() +
-  geom_hline(yintercept = 0, col = "red") +
-  geom_point(aes(.fitted, .resid), size = 0.5, stroke = 0.5) +
-  theme_light(base_size = 11/5) +
-  theme(axis.ticks = element_blank(), 
-        axis.text.x = element_blank(),
-        axis.text.y = element_blank(), 
-        axis.title.x = element_blank(), 
-        axis.title.y = element_blank(),
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank()) +
-  facet_wrap(~unique_lineup_id, scales = "free")
-
 ```
@@
 There are other kinds of residual departures like autocorrelation that are not considered in this study.
-The primary goal of this paper is to establish a new way of evaluating residual plot and conducting visual test with computer vision models.
+The primary goal of this study is to establish a new way of evaluating residual plot and conducting visual test with computer vision models.
 Building computer vision models for other model violations and other types of diagnostic plots could be future directions of this field.

### 2024-02-12 paper/paper.tex
update paper

@@
 
-\articletype{ARTICLE TEMPLATE}
+\articletype{}
 
@@
 
-\author{\name{Weihao Li$^{a}$}
+\author{\name{Weihao Li$^{a}$, Dianne Cook$^{a}$, Emi Tanaka$^{b,
+c}$, Susan VanderPlas$^{d}$, Klaus Ackermann$^{a}$}
 \affil{$^{a}$Department of Econometrics and Business Statistics, Monash
-University, Clayton, VIC, Australia}
+University, Clayton, VIC, Australia; $^{b}$Biological Data Science
+Institute, Australian National University, Acton, ACT,
+Australia; $^{c}$Research School of Finance, Actuarial Studies and
+Statistics, Australian National University, Acton, ACT,
+Australia; $^{d}$Department of Statistics, University of Nebraska,
+Lincoln, Nebraska, USA}
 }
@@
 \thanks{CONTACT Weihao
-Li. Email: \href{mailto:weihao.li@monash.edu}{\nolinkurl{weihao.li@monash.edu}}}
+Li. Email: \href{mailto:weihao.li@monash.edu}{\nolinkurl{weihao.li@monash.edu}}, Dianne
+Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, Emi
+Tanaka. Email: \href{mailto:emi.tanaka@anu.edu.au}{\nolinkurl{emi.tanaka@anu.edu.au}}, Susan
+VanderPlas. Email: \href{mailto:susan.vanderplas@unl.edu}{\nolinkurl{susan.vanderplas@unl.edu}}, Klaus
+Ackermann. Email: \href{mailto:Klaus.Ackermann@monash.edu}{\nolinkurl{Klaus.Ackermann@monash.edu}}}
 
@@
 suggest heteroskedasticity, it is important to avoid over-interpreting
-this visual pattern. In this case, the fitted model is correctly
-specified, and the triangular shape is actually a result of the skewed
-distribution of the predictors, rather than indicating a flaw in the
-model.
+this visual pattern. In this case, the fitted regression model is
+correctly specified, and the triangular shape is actually a result of
+the skewed distribution of the predictors, rather than indicating a flaw
+in the model.
 
@@
 idea involves testing whether the actual residual plot visually differs
-significantly from null plots, which are created using residuals
-generated from the null distribution. Typically, this is accomplished
-through the lineup protocol. In this approach, the real residual plot is
-embedded within a lineup alongside several null plots. If the real
-residual plot can be distinguished from the lineup, it provides evidence
-for rejecting the null hypothesis.
+significantly from null plots, where null plots are plotted with
+residuals generated from the residual rotation distribution
+\citep{langsrud2005rotation}, which is a distribution consistent with
+the null hypothesis \(H_0\) that the linear regression model is
+correctly specified. Typically, the visual test is accomplished through
+the lineup protocol, where the real residual plot is embedded within a
+lineup alongside several null plots. If the real residual plot can be
+distinguished from the lineup, it provides evidence for rejecting
+\(H_0\).
 
@@
 diverse subjects. For instance,
-\cite{loy2013diagnostic, loy2014hlmdiag, loy2015you} illustrate its
+\cite{loy2013diagnostic, loy2014hlmdiag, loy2015you} illustrated its
 applicability in diagnosing hierarchical linear models. Additionally,
-\citet{widen2016graphical} demonstrates its utility in geographical
-research, while \citet{krishnan2021hierarchical} explores its
+\citet{widen2016graphical} demonstrated its utility in geographical
+research, while \citet{krishnan2021hierarchical} explored its
 effectiveness in forensic examinations.
@@
 on this work, introducing graph theoretic scagnostics, which encompassed
-nine computable measures applied to planar proximity graphs. These
-measures, including, but not limited to, ``Outlying'', ``Skinny'',
-``Stringy'', ``Straight'', ``Monotonic'', ``Skewed'', ``Clumpy'', and
-``Striated'' aimed to characterize outliers, shape, density, trend,
-coherence and other characteristics of the data. While this approach has
-been inspiring, there is a recognition \citep{buja2009statistical} that
-it may not capture all the necessary visual features that differentiate
+computable measures applied to planar proximity graphs. These measures,
+including, but not limited to, ``Outlying'', ``Skinny'', ``Stringy'',
+``Straight'', ``Monotonic'', ``Skewed'', ``Clumpy'', and ``Striated''
+aimed to characterize outliers, shape, density, trend, coherence and
+other characteristics of the data. While this approach has been
+inspiring, there is a recognition \citep{buja2009statistical} that it
+may not capture all the necessary visual features that differentiate
 actual residual plots from null plots. A more promising alternative
-entails enabling computers to learn the function for extracting visual
+entails enabling machines to learn the function for extracting visual
 features from residual plots. Essentially, this means empowering
 computers to discern the crucial visual features for residual
-diagnostics and determining the method to extract them. Modern computer
-vision models are well-suited for addressing this challenge.
-
-Modern computer vision models often rely on deep neural networks with
-convolutional layers \citep{fukushima1982neocognitron}. These layers
-leverage hierarchical patterns in data, downsizing and transforming
-images by summarizing information in a small space. Numerous studies
-have demonstrated the efficacy of convolutional layers in addressing
-various vision tasks, including image recognition \citep{rawat2017deep}.
-Despite the widespread use of computer vision models in fields like
-computer-aided diagnosis \citep{lee2015image}, pedestrian detection
+diagnostics and determining the method to extract them.
+
+Modern computer vision models are well-suited for addressing this
+challenge. They rely on deep neural networks with convolutional layers
+\citep{fukushima1982neocognitron}. These layers leverage hierarchical
+patterns in data, downsizing and transforming images by summarizing
+information in a small space. Numerous studies have demonstrated the
+efficacy of convolutional layers in addressing various vision tasks,
+including image recognition \citep{rawat2017deep}. Despite the
+widespread use of computer vision models in fields like computer-aided
+diagnosis \citep{lee2015image}, pedestrian detection
 \citep{brunetti2018computer}, and facial recognition
@@
 
-\begin{figure}
+\begin{figure}[!h]
 
@@
 
-\hypertarget{different-input-formats}{%
-\subsubsection{Different input formats}\label{different-input-formats}}
+\hypertarget{input-formats}{%
+\subsubsection{Input formats}\label{input-formats}}
 
@@
 The quality and relevance of the input data greatly influence the
-model's capacity to generate insightful and meaningful results.
-
-A straightforward approach involves feeding a vector of residuals along
+model's capacity to generate insightful and meaningful results. There
+are several designs of the input format can be considered.
+
+A straightforward design involves feeding a vector of residuals along
 with a vector of fitted values, essentially providing all the necessary
@@
 histograms, it becomes possible to provide fixed-length input vectors.
-Still, since the histogram only captures the marginal distribution of
-residuals and fitted values respectively, it can not be used to
+Still, since histograms only capture the marginal distribution of
+residuals and fitted values respectively, they can not be used to
 differentiate visual patterns with same marginal distributions but
@@
 
-Another approach is to use an image as input. The primary advantage of
-using an image, as opposed to the vector format, is the utilization of
-the sophisticated image processing architectures developed over the
-years, such as the VGG16 architecture proposed in
+Another design involves using an image as input. The primary advantage
+of this design, as opposed to the vector format, is the availability of
+the existing and sophisticated image processing architectures developed
+over the years, such as the VGG16 architecture proposed in
 \citet{simonyan2014very}. These architectures can effectively capture
 and summarize spatial information from nearby pixels, which is less
-straightforward with vector input. The only considerations are the image
+straightforward with vector input. The main considerations are the image
 resolution and the aesthetics of the residual plot. In general, higher
@@
 implies that the model can not accept arbitrary images as input but
-requires the use of the same preprocessing software to convert provided
-residuals and fitted values into a standardized style for the residual
-plot.
+requires the use of the same preprocessing pipeline to convert residuals
+and fitted values into a standardized-style residual plot.
 
@@
 a triplet or a lineup is also a possible option.
-\citet{chopra2005learning} has shown that computer vision models
+\citet{chopra2005learning} have shown that computer vision models
 designed for image comparison can assess whether a pair of images are
 similar or dissimilar. Applied to our specific problem, we can define
-null plots of a fitted model to be similar to each other, while
-considering actual residual plots to be distinct from null plots of any
-fitted model. A triplet constitutes a set of three images, denoted as
-\(image_1\), \(image_2\) and \(image_3\). It is often used to predict
-whether \(image_2\) or \(image_3\) is more similar to \(image_1\),
-proving particularly useful for establishing rankings between samples.
-For this setup, we can apply the same criteria to define similarity
-between images. However, it's important to note that these two
-approaches usually require additional considerations regarding the loss
-function and, at times, non-standard training processes due to shared
-weights between different convolutional blocks.
-
-Presenting a lineup to a model and asking it to predict which residual
-plot is the most different one aligns closely with the lineup protocol.
-However, if a lineup consists of a large amount of residual plots, the
-resolution of the input image could become quite large, posing
-challenges in training the model. This approach was experimented in a
-pilot study conducted by us but the performance of the trained model was
-sub-optimal.
+null plots of a fitted regression model to be similar to each other,
+while considering actual residual plots to be distinct from null plots
+of any fitted regression model. A triplet constitutes a set of three
+images, denoted as \(image_1\), \(image_2\) and \(image_3\). It is often
+used to predict whether \(image_2\) or \(image_3\) is more similar to
+\(image_1\), proving particularly useful for establishing rankings
+between samples. For this setup, we can apply the same criteria to
+define similarity between images. However, it is important to note that
+these two approaches usually require additional considerations regarding
+the loss function and, at times, non-standard training processes due to
+shared weights between different convolutional blocks.
+
+Presenting a lineup to a model aligns closely with the lineup protocol.
+However, as the number of residual plots in a lineup increases, the
+resolution of the input image grows rapidly, posing challenges in
+training the model. We experimented with this approach in a pilot study,
+but the performance of the trained model was sub-optimal.
 
@@
 
-\hypertarget{different-output-formats}{%
-\subsubsection{Different output
-formats}\label{different-output-formats}}
+\hypertarget{output-formats}{%
+\subsubsection{Output formats}\label{output-formats}}
 
@@
 fixed-resolution image, the output from the computer vision model can
-take one of two formats: binary or numeric. This choice determines
-whether the model belongs to a classification model or a regression
-model. The binary outcome could be designed to indicate whether the
-input image is a null plot, or whether the input image would be rejected
-in a visual test conducted by humans. The latter option requires data
-from prior human subject experiments, presenting difficulties in
-controlling the quality of data due to variations in experimental
-settings across different studies. Additionally, some visual inference
-experiments are unrelated to linear regression models and residual plot
-diagnostics, resulting in a limited amount of available training data.
-
-Alternatively, the output could be designed as a meaningful numeric
-measure, such as the average distance to good residual plots. This
-approach necessitates defining a distance measure between residual
-plots, which may be a non-trivial task. However, once a meaningful
-distance measure is established, the prediction from the computer vision
-model becomes an interpretable value with tangible significance. Studies
-have demonstrated that defining a proper distance between images can
-enhance the matching accuracy in image search compared to a binary
-outcome model (ref here).
+take one of two forms: binary or numeric. This choice determines whether
+the model belongs to a classification model or a regression model. The
+binary outcome encoded as \(0\) and \(1\) could be used to represent
+whether the input image is a null plot, or whether the input image would
+be rejected in a visual test conducted by humans. Training a model
+following the latter option requires data from prior human subject
+experiments, presenting difficulties in controlling the quality of data
+due to variations in experimental settings across different studies.
+Additionally, some visual inference experiments are unrelated to linear
+regression models or residual plot diagnostics, resulting in a limited
+amount of available training data.
+
+Alternatively, the output could be a meaningful and interpretable
+numerical measure useful for assessing residual plots, such as the
+strength of suspicious visual patterns reflecting the extent of model
+violations and the difficulty index for identifying whether a residual
+plot has no issues. However, these numeric measures are often informally
+used in daily communication but are not typically formalized or
+rigorously defined. For the purpose of training a model, this numeric
+measure has to be quantifiable.
+
+In this study, we chose to define and use a distance measure to quantify
+the difference between the residual plot and the null plots.
+\citet{vo2016localizing} have also demonstrated that defining a proper
+distance between images can enhance the matching accuracy in image
+search compared to a binary outcome model.
 
@@
 Section \ref{introduction} could be used to construct distance metrics
-for data plots comparison, but the functional form still needs to be
+for residual plots comparison, but the functional form still needs to be
 carefully refined to accurately reflect the extent of the violations.
@@
 
-The distance metrics proposed in this paper takes into account the fact
-that we try to measure how different a residual plot is from a good
-residual plot, or in other words, how different a given fitted model is
-from a correctly specified model. For the classical normal linear
-regression model, residuals \(\boldsymbol{e}\) are derived from the
-fitted values \(\hat{\boldsymbol{y}}\) and observed values
+The distance measure proposed in this study takes into account the fact
+that we tried to measure how different a residual plot is from a good
+residual plot, or in other words, how different a given fitted
+regression model is from a correctly specified model. For the classical
+normal linear regression model, residuals \(\boldsymbol{e}\) are derived
+from the fitted values \(\hat{\boldsymbol{y}}\) and observed values
 \(\boldsymbol{y}\). Suppose the data generating process is known and the
@@
 distribution \(Q\) summarizes what good residuals should follow given
-the predictors are known and fixed.
+the design matrix \(\boldsymbol{X}\) is known and fixed.
 
@@
 numbers of \(\boldsymbol{\varepsilon}\) and transform it to
-\(\boldsymbol{e}\) to get the empirical estimates. For simplicity, we
-replaced the variance-covariance matrix of residuals
+\(\boldsymbol{e}\) to get the empirical estimates. For simplicity, in
+this study, we replaced the variance-covariance matrix of residuals
 \(\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}' = \boldsymbol{R}\sigma^2\)
@@
 causing an omitted variable problem, \(P\) will be different from \(Q\)
-because the residual operator obtained from the fitted model will not be
-the same as \(\boldsymbol{R}\). Besides, if the
+because the residual operator obtained from the fitted regression model
+will not be the same as \(\boldsymbol{R}\). Besides, if the
 \(\boldsymbol{\varepsilon}\) follows a non-normal distribution such as a
@@
 distribution \(Q\) and the reference residual distribution \(P\), we
-used a distance metric based on Kullback-Leibler divergence
+used a distance measure based on Kullback-Leibler divergence
 \citep{kullback1951information} to quantify the difference between two
@@
 
-This distance metric was first proposed in \citet{li2023plot}. It was
+This distance measure was first proposed in \citet{li2023plot}. It was
 mainly designed for measuring the effect size of non-linearity and
-heteroskedasticity in a residual plot. \citet{li2023plot} showed that,
-for a classical normal linear regression model that omits a necessary
-higher-order predictors \(\boldsymbol{Z}\), and incorrectly assumes
+heteroskedasticity in a residual plot. \citet{li2023plot} have showed
+that, for a classical normal linear regression model that omits a
+necessary higher-order predictors \(\boldsymbol{Z}\), and incorrectly
+assumes
 \(\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)\)
@@
 Thus, equation \ref{eq:kl-2} given in \citet{li2023plot} will not be
-applicable to models with non-normality violations.
+applicable to models violating the normality assumption.
 
@@
 
-The algorithm for approximating equation \ref{eq:kl-1} starts from
-simulating \(m\) sets of \(\boldsymbol{\varepsilon}\) according to the
-error distribution. The simulated errors are stored in a matrix
-\(\boldsymbol{A}\) with \(n\) rows and \(m\) columns, where each column
-of \(\boldsymbol{A}\) is a set of realized values of
-\(\boldsymbol{\varepsilon}\). Then, we can get \(m\) sets of realized
-values of \(\boldsymbol{e}\) stored in the matrix \(\boldsymbol{B}\) by
-applying the residual operator
+Given \(\boldsymbol{X}\) and \(\boldsymbol{\beta}\), the algorithm for
+approximating equation \ref{eq:kl-1} starts from simulating \(m\) sets
+of observed values \(\boldsymbol{y}\) according to the data generating
+process. The observed values are stored in a matrix \(\boldsymbol{A}\)
+with \(n\) rows and \(m\) columns, where each column of
+\(\boldsymbol{A}\) is a set of observed values. Then, we can get \(m\)
+sets of realized values of \(\boldsymbol{e}\) stored in the matrix
+\(\boldsymbol{B}\) by applying the residual operator
 \(\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}\). Furthermore, kernel
@@
 is applied on each row of \(B\) to estimate \(p_i(e_i)\) for
-\(i = 1, ..., n\). The KDE computation is done by the \texttt{density}
-function in R.
+\(i = 1, ..., n\). The KDE computation can be done by the
+\texttt{density} function in R.
 
@@
 
-\hypertarget{approximation-of-the-distance-metric}{%
+\hypertarget{approximation-of-the-distance-measure}{%
 \subsubsection{Approximation of the distance
-metric}\label{approximation-of-the-distance-metric}}
-
-In the previous sections, we have defined a distance metric given in
+measure}\label{approximation-of-the-distance-measure}}
+
+In the previous sections, we have defined a distance measure given in
 equation \ref{eq:kl-0} for quantifying the difference between the actual
 residual distribution \(P\) and an ideal reference distribution \(Q\).
-You may have noticed that this distance metric can only be computed when
-the data generating process is known. In reality, we often have no
+You may have noticed that this distance measure can only be computed
+when the data generating process is known. In reality, we often have no
 knowledge about the data generating process, otherwise we do not need to
@@
 
-What we proposed in this paper is a method to approximate this distance
-with a residual plot. Let \(D\) be the result of equation \ref{eq:kl-0}
-indicating the extent of the model violations, and our estimator
-\(\hat{D}\) is formulated as
+We tried to train a computer vision model to approximate this distance
+measure with a residual plot. Let \(D\) be the result of equation
+\ref{eq:kl-0} indicating the extent of the model violations, and our
+estimator \(\hat{D}\) is formulated as
 
@@
 different the underlying distribution of the residuals is from a good
-residual distribution. This is meaningful way to check if a residual
-plot is a good residual plot, and to know the strength of the visual
-signal embedded in the residual plot.
+residual distribution. It also provides information for the strength of
+the visual signal embedded in the residual plot.
 
@@
 contained in a single residual plot is limited and it may not be able to
-summarise all the characteristics of the residual distribution. For a
-given residual distribution \(P\), we can generate many different
-residual plots. Some of them share similar visual patterns, but some of
-them could be visually very different from the rest, especially for
-models with small \(n\). This suggests the error of the approximation
-will vary depends on whether the observed residual plot is
-representative or not.
+summarise all the important characteristics of the residual
+distribution. For a given residual distribution \(P\), we can generate
+many different residual plots. Some of them share similar visual
+patterns, but some of them could be visually very different from the
+rest, especially for regression models with small \(n\). This suggests
+the error of the approximation will vary depends on whether the observed
+residual plot is representative or not.
 
@@
 the visual pattern is weak or moderate, having \(\hat{D}\) will not be
-sufficient to determine if \(H_0\) that the model is correctly specified
-should be rejected.
+sufficient to determine if \(H_0\) should be rejected.
 
 To solve this issue while aligning with the principle of visual
-inference, the approximated distance \(\hat{D}\) can be viewed as a test
-statistic. And the null distribution of this statistic can be
-approximated by the empirical distribution of \(\hat{D}_{null}\) for a
-large amount of null plots of a fitted model. The procedure involves
+inference, \(\hat{D}\) can be viewed as a test statistic. And the null
+distribution of this statistic can be approximated by the empirical
+distribution of \(\hat{D}_{null}\), where \(\hat{D}_{null}\) is the
+approximated distance for a null plot simulated from the fitted
+regression model. The approximation of the distribution involves
 applying the residual rotation technique \citep{buja2009statistical} on
-the fitted model to obtain null residuals. The null residuals are then
-used to make null plots and fed into the model to get predictions. The
-empirical distribution is constructed with the predicted distance.
+the fitted regression model to obtain null residuals. The null residuals
+are then used to make null plots and fed into the computer vision model
+to get predictions. The empirical distribution is constructed with the
+approximated distance.
 
@@
 
-Let \(\hat{D}_{null}^{(i)}\) be the predicted distance of the \(i\)-th
-null plots, where \(i = 1,...,n_{null}\) and
+Let \(\hat{D}_{null}^{(i)}\) be the approximated distance of the
+\(i\)-th null plots, where \(i = 1,...,n_{null}\) and
 \(n_{null} \in \mathbb{N}^+\) is a sufficiently large number. Quantiles
@@
 quantile, the 95\% quantile and the 99\% quantile. These quantiles are
-used as thresholds to decide if \(H_0\) needs to be rejected. For
+used as critical values to decide if \(H_0\) needs to be rejected. For
 example, if \(\hat{D}\) is greater than and equal to the 95\% sample
-quantile of the null distribution \(Q_{null}(0.95)\), we could say the
-predicted distance for the actual residual plot is significantly
-different from the predicted distance for null plots with 95\%
-significance level. Based on our experience, in order to get a stable
-estimate of the 95\% quantile, \(n_{null}\) usually needs to be at least
-\(100\). And if the null distribution has a long tail, more null plots
-will be needed. Alternatively, a p-value can be used to represents the
-probability of observing an event equally or more extreme than the given
-event under \(H_0\), and it can be estimated by
+quantile \(Q_{null}(0.95)\), we could say the approximated distance for
+the actual residual plot is significantly different from the
+approximated distance for null plots with 95\% significance level. Based
+on our experience, in order to get a stable estimate of the 95\%
+quantile, \(n_{null}\) usually needs to be at least \(100\). And if the
+null distribution has a long tail, more null plots will be needed.
+Alternatively, a p-value can be used to represents the probability of
+observing an event equally or more extreme than the given event under
+\(H_0\), and it can be estimated by
 \(1/m\sum_{i=1}^{n_{null}}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)\).
+
+If precision in sample quantiles is not the main priority, using a
+pre-calculated table of quantiles is an available option. Such a table
+offers pre-determined quantiles for a specified number of observations.
+It is generated by assessing numerous null plots derived from various
+simulated regression models and averaging them. Essentially, this shifts
+the computational burden from the user to the developer.
 
@@
 requires \(n_{boot} \times n_{null}\) times of residual plot assessment.
-In practice, \(Q_{null}(0.95)\) will be used to replace
+In practice, \(Q_{null}(0.95)\) can be used to replace
 \(Q_{boot}^{(i)}(0.95)\) in the computation.
@@
 \(D\) unattainable. Consequently, the computer vision models developed
-in this study were trained using synthetic data. This approach provides
-us with precise label annotations. Additionally, it ensures a large and
-diverse training dataset, as we have control over the data generating
-process, and the simulation of the training data is relatively
+in this study were trained using synthetic data. This approach provided
+us with precise label annotations. Additionally, it ensured a large and
+diverse training dataset, as we had control over the data generating
+process, and the simulation of the training data was relatively
 cost-effective.
@@
 
-The residuals and fitted values of the fitted model is obtained by
-regressing \(\boldsymbol{y}\) on \(\boldsymbol{x}_1\). This data
-generating process is adopted from \citet{li2023plot} where it was used
-to simulate residual plots with non-linearity and heteroskedasticity
-visual patterns for human subject experiments.
-
-In equation \ref{eq:data-sim}, predictor \(\boldsymbol{x}_1\) and
-\(\boldsymbol{x}_2\) are independent of each other, so excluding
-\(\boldsymbol{x}_2\) from the regression formula will not cause any
-problem. However, \(\boldsymbol{z}\) and \(\boldsymbol{w}\) are
-higher-order terms of \(\boldsymbol{x}_1\) and \(\boldsymbol{x}_2\). If
-\(\beta_2 \neq 0\), the regression model will suffer from non-linearity
-issues. Parameter \(j\) is a shape parameter controlling the number of
-tuning points of the non-linearity pattern. Generally, greater values of
-\(j\) will result in more tuning points.
+\begin{table}
+
+\caption{\label{tab:factor}Factors used in the data generating process for synthetic data simulation.
+Factor $j$ and $a$ controls the non-linearity shape and the heteroskedasticity shape respectively.
+Factor $b$, $\sigma_\varepsilon$ and $n$ control the signal strength.
+Factor $\text{dist}_\varepsilon$, $\text{dist}_{x1}$ and $\text{dist}_{x2}$ specifies the distribution of $\varepsilon$, $X_1$ and $X_2$ respectively.}
+\centering
+\begin{tabular}[t]{ll}
+\toprule
+Factor & Domain\\
+\midrule
+j & \{2, 3, ..., 18\}\\
+a & {}[-1, 1]\\
+b & {}[0, 100]\\
+$\beta_1$ & {0, 1}\\
+$\beta_2$ & {0, 1}\\
+\addlinespace
+$\text{dist}_\varepsilon$ & \{discrete, uniform, normal, lognormal\}\\
+$\text{dist}_{x1}$ & \{discrete, uniform, normal, lognormal\}\\
+$\text{dist}_{x2}$ & \{discrete, uniform, normal, lognormal\}\\
+$\sigma_{\varepsilon}$ & {}[0.0625, 9]\\
+$\sigma_{X1}$ & {}[0.3, 0.6]\\
+\addlinespace
+$\sigma_{X2}$ & {}[0.3, 0.6]\\
+n & {}[50, 500]\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+The residuals and fitted values of the fitted model were obtained by
+regressing \(\boldsymbol{y}\) on \(\boldsymbol{x}_1\). If
+\(\beta_1 \neq 0\), \(\boldsymbol{x}_2\) was also included in the design
+matrix. This data generation process was adapted from
+\citet{li2023plot}, where it was utilized to simulate residual plots
+exhibiting non-linearity and heteroskedasticity visual patterns for
+human subject experiments. A summary of the factors utilized in Equation
+\ref{eq:data-sim} is provided in Table \ref{tab:factor}.
+
+In Equation \ref{eq:data-sim}, \(\boldsymbol{z}\) and \(\boldsymbol{w}\)
+represent higher-order terms of \(\boldsymbol{x}_1\) and
+\(\boldsymbol{x}_2\), respectively. If \(\beta_2 \neq 0\), the
+regression model will encounter non-linearity issues. Parameter \(j\)
+serves as a shape parameter that controls the number of tuning points in
+the non-linear pattern. Typically, higher values of \(j\) lead to an
+increase in the number of tuning points, as illustrated in Figure
+\ref{fig:different-j}.
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-j-1}
+
+}
+
+\caption{Non-linearity forms generated for the synthetic data simulation. The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.}\label{fig:different-j}
+\end{figure}
 
@@
 shape parameter controlling the location of the smallest variance in a
-residual plot.
-
-Non-normality violations are introduced by defining a non-normal
-distribution for \(\boldsymbol{\varepsilon}\).
-
-\begin{itemize}
-\tightlist
-\item
-  distribution of x1 and x2
-\item
-  summary table of all the factor values
-\item
-  example plots for the violations
-\end{itemize}
+residual plot as shown in Figure \ref{fig:different-a}.
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-a-1}
+
+}
+
+\caption{Heteroskedasticity forms generated for the synthetic data simulation.
+Different shapes are controlled by the continuous factor $a$ between -1 and 1.
+For $a = -1$, the residual plot exhibits a "left-triangle" shape.
+And for $a = 1$, the residual plot exhibits a "right-triangle" shape.
+}\label{fig:different-a}
+\end{figure}
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-e-1}
+
+}
+
+\caption{Non-normality forms generated for the synthetic data simulation. Four different error distributions including discrete, lognormal, normal and uniform are considered.}\label{fig:different-e}
+\end{figure}
+
+Non-normality violations arise from specifying a non-normal distribution
+for \(\boldsymbol{\varepsilon}\). In the synthetic data simulation, four
+distinct error distributions are considered, including discrete,
+uniform, normal, and lognormal distributions, as presented in Figure
+\ref{fig:different-e}. Each distribution imparts unique characteristics
+to the residuals. The discrete error distribution introduces
+discreteness in residuals, while the lognormal distribution typically
+yields outliers. Uniform error distribution may result in residuals
+filling the entire space of the residual plot. All of these
+distributions exhibit visual distinctions from the normal error
+distribution.
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-j-x2-1}
+
+}
+
+\caption{Residual plots of multiple linear regression models with non-linearity issues.
+The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.
+A second predictor $\boldsymbol{x}_2$ is introduced to the regression model to create complex shapes.}\label{fig:different-j-x2}
+\end{figure}
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-j-heter-1}
+
+}
+
+\caption{Residual plots of models violating both the non-linearity and the heteroskedasticity assumptions.
+The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$, and the "left-triangle" shape is introduced by setting $a = -1$.}\label{fig:different-j-heter}
+\end{figure}
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/different-e-heter-1}
+
+}
+
+\caption{Residual plots of models violating both the non-normality and the heteroskedasticity assumptions.
+The four shapes are generated by using four different error distributions including discrete, lognormal, normal and uniform, and the "left-triangle" shape is introduced by setting $a = -1$.
+}\label{fig:different-e-heter}
+\end{figure}
+
+Equation \ref{eq:data-sim} accommodates the incorporation of the second
+predictor \(\boldsymbol{x}_2\). Introducing it into the data generation
+process by setting \(\beta_1 = 1\) significantly enhances the complexity
+of the shapes, as illustrated in Figure \ref{fig:different-j-x2}. In
+comparison to Figure \ref{fig:different-j}, Figure
+\ref{fig:different-j-x2} demonstrates that the non-linear shape
+resembles a surface rather than a single curve. This augmentation can
+facilitate the computer vision model in learning visual patterns from
+residual plots of the multiple linear regression model.
+
+In real-world analysis, it's not uncommon to encounter instances where
+multiple model violations coexist. In such cases, the residual plots
+often exhibit a mixed pattern of visual anomalies corresponding to
+different types of model violations. Figure \ref{fig:different-j-heter}
+and \ref{fig:different-e-heter} show the visual patterns of models with
+multiple model violations.
 
@@
 
-We have mentioned in Section \ref{introduction} that scagnostics are a
-set of manually designed visual feature extraction functions. Although
+In Section \ref{introduction}, we mentioned that scagnostics consist of
+a set of manually designed visual feature extraction functions. While
 our computer vision model will learn its own feature extraction function
-during the training process, it will be beneficial to utilize additional
-information provided by scagnostics to help computer vision model making
-more accurate predictions.
-
-For each generated residual plot, four scagnostics, namely,
-``Monotonic'', ``Sparse'', ``Splines'' and ``Striped'' are computed
-using the \texttt{cassowaryr} R package (ref here). The computed
-measures along with the number of observations of the fitted model are
-provided as the second input of the computer vision model. Other
-scagnositcs are also informative, but unfortunately, they are
-unavailable due to a fatal bug in the compiled C program of the
-\texttt{interp} R package (ref here) that will crash the process in a
-unpredictable manner. For reproducibility, we do not include these
-scagnostics in the training data.
+during training, leveraging additional information from scagnostics can
+enhance the model's predictive accuracy.
+
+For each generated residual plot, we computed four scagnostics --
+``Monotonic,'' ``Sparse,'' ``Splines,'' and ``Striped'' -- using the
+cassowaryr R package \citep{mason2022cassowaryr}. These computed
+measures, along with the number of observations from the fitted model,
+were provided as the second input for the computer vision model.
+Although other scagnostics are informative, they are currently
+unavailable due to a fatal bug in the compiled C program of the interp R
+package \citep{Albrecht2023interp}, which may unpredictably crash the
+process. For reproducibility, we excluded these scagnostics from the
+training data.
 
@@
 
-In order to train a robust computer vision model, we intentionally
-control the distribution of the target variable \(D\) of the training
-data. Making it uniform between \(0\) and \(7\). This is accomplished by
-preparing \(50\) buckets each only accepts training samples with \(D\)
-between \([7(i - 1)/49, 7i/49)\) for \(i < 50\), where \(i\) is the
+To train a robust computer vision model, we deliberately controlled the
+distribution of the target variable \(D\) in the training data. We
+ensured that it followed a uniform distribution between \(0\) and \(7\).
+This was achieved by organizing \(50\) buckets, each exclusively
+accepting training samples with \(D\) falling within the range
+\([7(i - 1)/49, 7i/49)\) for \(i < 50\), where \(i\) represents the
 index of the \(i\)-th bucket. For the \(50\)-th bucket, any training
-samples with \(D >= 7\) will be accepted. We have prepared 80000
-training images, so each bucket can only contain
-\(80000 \div 50 = 1600\) training samples. The simulator will repeatedly
-sample parameter values from the parameter sample, generate residuals
-and fitted values using the data generating process, compute the
-distance, and check if the sample can be accepted by the corresponding
-bucket, until all the buckets are full.
+samples with \(D \geq 7\) were accepted.
+
+With 80000 training images prepared, each bucket accommodated a maximum
+of \(80000 \div 50 = 1600\) training samples. The simulator iteratively
+sampled parameter values from the parameter space, generated residuals
+and fitted values using the data generation process, computed the
+distance, and checked if the sample fitted within the corresponding
+bucket. This process continued until all buckets were filled.
+
+Similarly, we adopted the same methodology to prepare 8000 test images
+for performance evaluation and model diagnostics.
 
@@
 
-The architecture of the computer vision model is adopted from a existing
-architecture called VGG16 that proven to be performant in image
-classification \citep{simonyan2014very}. The first layer is the input
-layer with shape \(n \times h \times w \times 3\) which could accept
-\(n\) RGB images, followed by a grayscale conversion layer using the
-luma formula under the CCIR 601 standard (ref here) to convert the
-colour image to an grey scale image. Grayscale is enough for our task
-since data points are plotted in black. Three different combinations of
-\(h\) and \(w\) are chosen to seek a sufficient large image resolution
-for this problem. This includes \(32 \times 32\), \(64 \times 64\) and
-\(128 \times 128\).
-
-The processed image is then used as the input of the first
-convoluational block. The model has at most five consecutive
-convolutional blocks which is similar to the original VGG16
-architecture. In each block, there are two 2D convoluational layers
-followed by two activation layers respectively, and there is a 2D max
-pooling layer follows the second activation layer. The 2D convolutional
-layer convolves the input with a fixed number of \(3 \times 3\)
-convolution filters, and the 2D max pooling layer downsamples the input
-along its spatial dimensions by taking the maximum value over a
-\(2 \times 2\) window for each channel of the input. The activation
-layer is set to be the rectified linear unit activation function (ReLu),
-which is a standard practice in deep learning. This introduces a
-non-linear transformation of the output of the 2D convolutional layer.
-Furthermore, to regularize the training, a batch normalization layer is
-added after each of the 2D convolutional layer and before the activation
-layer. A dropout layer is also appended at the end of each convolutional
-block to randomly set some inputs to zeros during training.
-
-The output of the last convolutional block will be summarised by a
-global max pooling layer or a global average pooling layer to gain a
-two-dimensional tensor. Since we want to utilize the information
-contained in scagnostics, the two-dimensional tensor is concatenated
-with an additional \(n \times 5\) tensor, which contains the
-``Monotonic'', ``Sparse'', ``Splines'' and ``Striped'' measures and the
-number of observations for \(n\) residual plots.
-
-The concatenated tensor will be fed into the final prediction block.
-This block contains two fully-connected layers, the first one has at
-least \(128\) units followed by a dropout layer. Sometimes, a batch
-normalization layer will be inserted between the fully-connected layer
-and the dropout layer. The second fully-connected layer has only one
-unit, serves as the output of the model.
-
-The model weights are randomly initialized and they are optimized by the
-Adam optimizer with the mean square error loss function. Therefore, we
-try to find a set of weights \(\hat{\boldsymbol{\theta}}\) by
+The architecture of the computer vision model is adapted from a
+well-established architecture known as VGG16, which has demonstrated
+high performance in image classification \citep{simonyan2014very}.
+
+The model begins with an input layer of shape
+\(n \times h \times w \times 3\), capable of handling \(n\) RGB images.
+This is followed by a grayscale conversion layer utilizing the luma
+formula under the Rec. 601 standard, which converts the color image to
+grayscale. Grayscale suffices for our task since data points are plotted
+in black. We experiment with three combinations of \(h\) and \(w\):
+\(32 \times 32\), \(64 \times 64\), and \(128 \times 128\), aiming to
+achieve sufficiently high image resolution for the problem at hand.
+
+The processed image is used as the input for the first convolutional
+block. The model comprises at most five consecutive convolutional
+blocks, mirroring the original VGG16 architecture. Within each block,
+there are two 2D convolutional layers followed by two activation layers,
+respectively. Subsequently, a 2D max-pooling layer follows the second
+activation layer. The 2D convolutional layer convolves the input with a
+fixed number of \(3 \times 3\) convolution filters, while the 2D
+max-pooling layer downsamples the input along its spatial dimensions by
+taking the maximum value over a \(2 \times 2\) window for each channel
+of the input. The activation layer employs the rectified linear unit
+(ReLU) activation function, a standard practice in deep learning, which
+introduces a non-linear transformation of the output of the 2D
+convolutional layer. Additionally, to regularize training, a batch
+normalization layer is added after each 2D convolutional layer and
+before the activation layer. Finally, a dropout layer is appended at the
+end of each convolutional block to randomly set some inputs to zero
+during training, further aiding in regularization.
+
+The output of the last convolutional block is summarized by either a
+global max pooling layer or a global average pooling layer, resulting in
+a two-dimensional tensor. To leverage the information contained in
+scagnostics, this tensor is concatenated with an additional
+\(n \times 5\) tensor, which contains the ``Monotonic,'' ``Sparse,''
+``Splines,'' and ``Striped'' measures, along with the number of
+observations for \(n\) residual plots.
+
+The concatenated tensor is then fed into the final prediction block.
+This block consists of two fully-connected layers. The first layer
+contains at least \(128\) units, followed by a dropout layer.
+Occasionally, a batch normalization layer is inserted between the
+fully-connected layer and the dropout layer for regularization purposes.
+The second fully-connected layer consists of only one unit, serving as
+the output of the model.
+
+The model weights \(\boldsymbol{\theta}\) were randomly initialized and
+they were optimized by the Adam optimizer with the mean square error
+loss function
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  a diagram for the model
-\end{itemize}
-
 \hypertarget{training-and-hyperparameter-tuning}{%
@@
 
-To obtain an optimal deep learning model, hyperparameters such as
-learning rate often needs to be tuned via a tuner. In this study, we
-used the Bayesian optimization tuner from the \texttt{keras-tuner}
-Python library (ref here) to tune the hyperparameters. The complete list
-of hyperparameters is provided in Table \ref{tab:hyperparameter}.
-
-The number of base filters determines the number of filters for the
-first 2D convolutional layer. In VGG16, the number of filters for the 2D
-convolutional layer in a block is 2 times the number in the previous
-block, except for the last block which has the same number of
-convolution filters as the previous one. This hyperparameter helps to
-control the complexity of the computer vision model. More base filters
-will result in more trainable parameters.
-
-Similarly, the number of units for the first fully-connected layer
-determines the complexity of the final prediction block. And more units
-will lead to more trainable parameters.
-
-Dropout rate and batch normalization are flexible hyperparameters which
-works jointly with other hyperparameters to smooth the training process.
-A high dropout rate is needed when the model learns too much noise from
-the training dataset. And a low dropout rate is needed when the model is
-complex and difficult to converge. Batch normalization on the other hand
-may be needed for solving the internal covariate shift problem due to
-randomness in the weight initialization and randomness in the input
-data.
-
-Having additional inputs including the scagnostics and the number of
-observations could be beneficial for making more accurate prediction. So
-we let the tuner to decide if these inputs are actually needed.
-
-Learning rate is one of the most important hyperparameters, and it
-determines the step size of the optimization algorithm. A high learning
-rate could help the model to avoid getting stuck in a local minimum, but
-it could also ruin the optimization process by jumping over the global
-minimum. A low learning rate could smooth the training process but it
-increases the time to converge and the chance of getting stuck in local
-minimum.
-
-The model is trained on the M3 high-performance computing platform (ref
-here) with TensorFlow (ref here) and Keras (ref here). During the
-training process, 80\% of the training data are used as the actual
-training data, and 20\% of the training data are reserved as validation
-data. We let the Bayesian optimization tuner runs 100 trials to find the
-best hyperparameter values according to the validation root mean square
-error. The tuner will restore the best epoch of the best model from the
-trials. Early stopping is also applied such that the training process
-will halt if the validation root mean square error does not improve for
-50 epochs. The maximum allowed epochs is 2000 where no models reach.
+To achieve an optimal deep learning model, hyperparameters like the
+learning rate often need to be fine-tuned using a tuner. In our study,
+we utilized the Bayesian optimization tuner from the KerasTuner Python
+library \citep{omalley2019kerastuner} for this purpose. A comprehensive
+list of hyperparameters is provided in Table \ref{tab:hyperparameter}.
+
+The ``number of base filters'' determines the number of filters for the
+first 2D convolutional layer. In the VGG16 architecture, the number of
+filters for the 2D convolutional layer in a block is typically twice the
+number in the previous block, except for the last block, which maintains
+the same number of convolution filters as the previous one. This
+hyperparameter aids in controlling the complexity of the computer vision
+model. Higher numbers of base filters result in more trainable
+parameters. Likewise, the ``number of units for the fully-connected
+layer'' determines the complexity of the final prediction block.
+Increasing the number of units enhances model complexity, resulting in
+more trainable parameters.
+
+The dropout rate and batch normalization are flexible hyperparameters
+that work in conjunction with other parameters to facilitate smooth
+training. A higher dropout rate is necessary when the model tends to
+overfit the training dataset by learning too much noise. Conversely, a
+lower dropout rate is preferred when the model is complex and
+challenging to converge. Batch normalization, on the other hand,
+addresses the internal covariate shift problem arising from the
+randomness in weight initialization and input data. It helps stabilize
+and accelerate the training process by normalizing the activations of
+each layer.
+
+Additionally, incorporating additional inputs such as scagnostics and
+the number of observations can potentially enhance prediction accuracy.
+Therefore, we allowed the tuner to determine whether these inputs were
+necessary for optimal model performance.
+
+The learning rate is a crucial hyperparameter, as it dictates the step
+size of the optimization algorithm. A high learning rate can help the
+model avoid local minima but risks overshooting and missing the global
+minimum. Conversely, a low learning rate smoothens the training process
+but makes the convergence time longer and increases the likelihood of
+getting trapped in local minima.
+
+Our model was trained on the M3 high-performance computing platform,
+using TensorFlow and Keras. During training, 80\% of the training data
+was utilized for actual training, while the remaining 20\% was used as
+validation data. The Bayesian optimization tuner conducted 100 trials to
+identify the best hyperparameter values based on validation root mean
+square error. The tuner then restored the best epoch of the best model
+from the trials. Additionally, we applied early stopping, terminating
+the training process if the validation root mean square error fails to
+improve for 50 epochs. The maximum allowed epochs was set at 2000,
+although no models reached this threshold.
 
@@
 \ref{training-and-hyperparameter-tuning}, the optimized hyperparameter
-values are provided in Table \ref{tab:best-hyperparameter}. It can be
-observed that at least \(32\) base filters are needed, and the
-preferable choice for the \(64 \times 64\) model and the
-\(128 \times 128\) model is \(64\) base filters, which is the same as
-the original VGG16 architecture. The optimized values for dropout rate
-for convolutional blocks is around \(0.4\). And applying batch
-normalization for convolutional blocks has a positive impact on the
-performance. All optimized models choose to preserve the additional
-inputs to reduce the validation error. The number of units needed for
-the fully-connected layer is \(256\), which is not a very large number
-comparing to the VGG16 classifier, suggesting that the problem we are
-trying to solve is relatively easy. The optimized learning rates are
-usually smaller than the default value recommended by Keras which is
+values are presented in Table \ref{tab:best-hyperparameter}. It was
+observable that a minimum of \(32\) base filters was necessary, with the
+preferable choice being \(64\) base filters for both the
+\(64 \times 64\) and \(128 \times 128\) models, mirroring the original
+VGG16 architecture. The optimized dropout rate for convolutional blocks
+hovered around \(0.4\), and incorporating batch normalization for
+convolutional blocks proved beneficial for performance.
+
+All optimized models chose to retain the additional inputs, contributing
+to the reduction of validation error. The number of units required for
+the fully-connected layer was \(256\), a relatively modest number
+compared to the VGG16 classifier, suggesting that the problem at hand
+was relatively straightforward. The optimized learning rates were
+typically smaller than the default value recommended by Keras, which is
 0.001.
@@
 
-The training and test performance of optimized models with three
-different input sizes are provided in Table \ref{tab:performance}. Among
-these models, the \(64 \times 64\) model and the \(32 \times 32\) model
-are consistently having the best training and test performance
-respectively. The mean absolute error shows that the difference between
-\(\hat{D}\) and \(D\) is around \(0.43\) on the test set, which is very
-small considering the range of \(D\) is normally between \(0\) and
-\(7\). The high \(R^2\) also suggests the prediction is mostly linearly
-correlated with the target.
-
-Figure \ref{fig:model-performance} displays a hexagonal heatmap for
-\(D - \hat{D}\) vs \(D\). For residual plots with \(D\) close to zero,
-all the optimized models tends to over-predict \(\hat{D}\), leading to
-negative residuals. When \(D\) is between \(1.5\) and \(6\), the smooth
-curves indicate that \(\hat{D}\) is not heavily biased, and there are
-some instances where \(\hat{D}\) is severely under-predicted. And when
-\(D\) is greater than \(6\), there are more instances being
-under-predicted than over-predicted.
-
-Given the model performance metrics, only the best model evaluated on
-the test set, i.e.~the \(32 \times 32\) model, will be used in the
-following analysis.
+The training and test performance for the optimized models with three
+different input sizes are summarized in Table \ref{tab:performance}.
+Among these models, the \(64 \times 64\) model and the \(32 \times 32\)
+model consistently exhibited the best training and test performance,
+respectively. The mean absolute error indicated that the difference
+between \(\hat{D}\) and \(D\) was approximately \(0.43\) on the test
+set, a negligible deviation considering the normal range of \(D\)
+typically falls between \(0\) and \(7\). The high \(R^2\) values also
+suggested that the predictions were largely linearly correlated with the
+target.
+
+Figure \ref{fig:model-performance} presents a hexagonal heatmap for
+\(D - \hat{D}\) versus \(\hat{D}\). The red smoothing curves, fitted by
+generalized additive models \citep{hastie2017generalized}, demonstrate
+that all the optimized models perform admirably on both the training and
+test sets. No structural issues are noticeable in Figure
+\ref{fig:model-performance}, but some minor issues regarding
+over-prediction and under-prediction are observed.
+
+The figure highlights that most under-predictions occurred when
+\(\hat{D} < 3\), while over-predictions occurred predominantly when
+\(3 < \hat{D} < 6\). To comprehend the reasons behind these deviations,
+we conducted a meta-analysis on \(D - \hat{D}\) for all factors used in
+the data generating process. A regression tree was constructed using the
+rpart R package \citep{Terry2022rpart}. The results of the regression
+tree are provided in the Appendix.
+
+The meta-analysis revealed that most issues arose from non-linearity
+problems and the presence of a second predictor in the regression model.
+When the error distribution had a very small variance, all computer
+vision models tended to under-predict the distance. Conversely, when the
+error distribution had a large variance, all models tended to
+over-predict the distance. Therefore, Additionally, for input images
+representing null plots, it was expected that the models will
+over-predict the distance, as explained in Section
+\ref{null-distribution-of-the-approximated-distance}.
+
+Since most of the deviation stemmed from the presence of non-linearity
+violations, to further investigate this, we split the test set and
+re-evaluated the performance, as detailed in Table
+\ref{tab:performance-sub}. It was evident that metrics for null plots
+were notably worse compared to other categories. Furthermore, residual
+plots solely exhibiting non-normality issues were the easiest to
+predict, with very low test root mean square error (RMSE) around
+\(0.3\), very low mean absolute error (MAE) around \(0.2\), and very
+high \(R^2\) around \(0.94\). Residual plots with non-linearity issues
+were more challenging to assess than those with heteroskedasticity or
+non-normality issues. Assessing residual plots with heteroskedasticity
+was not as difficult as assessing those with non-linearity issues. When
+multiple violations were introduced to a residual plot, the performance
+metrics typically lay between the metrics for each individual violation.
+
+Based on the model performance metrics, we will only use the
+best-performing model evaluated on the test set, namely the
+\(32 \times 32\) model, for the subsequent analysis.
 
@@
 
-\caption{\label{tab:performance}The training and test performance of three optimized models with different input sizes.}
+\caption{\label{tab:performance}The training and test performance of three optimized models with different input sizes. The best metrics are colored in red.}
 \centering
-\begin{tabular}[t]{lllll}
+\begin{tabular}[t]{l>{}l>{}l>{}l>{}l}
 \toprule
@@
 \multicolumn{5}{l}{\textbf{Training set}}\\
-\hspace{1em}$32 \times 32$ & 0.531 & 0.937 & 0.364 & 0.126\\
-\hspace{1em}$64 \times 64$ & 0.405 & 0.963 & 0.260 & 0.072\\
-\hspace{1em}$128 \times 128$ & 0.432 & 0.959 & 0.290 & 0.084\\
+\hspace{1em}$32 \times 32$ & \textcolor{black}{0.531} & \textcolor{black}{0.937} & \textcolor{black}{0.364} & \textcolor{black}{0.126}\\
+\hspace{1em}$64 \times 64$ & \textcolor{red}{0.405} & \textcolor{red}{0.963} & \textcolor{red}{0.260} & \textcolor{red}{0.072}\\
+\hspace{1em}$128 \times 128$ & \textcolor{black}{0.432} & \textcolor{black}{0.959} & \textcolor{black}{0.290} & \textcolor{black}{0.084}\\
 \addlinespace[0.3em]
 \multicolumn{5}{l}{\textbf{Test set}}\\
-\hspace{1em}$32 \times 32$ & 0.660 & 0.901 & 0.434 & 0.181\\
-\hspace{1em}$64 \times 64$ & 0.674 & 0.897 & 0.438 & 0.186\\
-\hspace{1em}$128 \times 128$ & 0.692 & 0.892 & 0.460 & 0.199\\
+\hspace{1em}$32 \times 32$ & \textcolor{red}{0.660} & \textcolor{red}{0.901} & \textcolor{red}{0.434} & \textcolor{red}{0.181}\\
+\hspace{1em}$64 \times 64$ & \textcolor{black}{0.674} & \textcolor{black}{0.897} & \textcolor{black}{0.438} & \textcolor{black}{0.186}\\
+\hspace{1em}$128 \times 128$ & \textcolor{black}{0.692} & \textcolor{black}{0.892} & \textcolor{black}{0.460} & \textcolor{black}{0.199}\\
 \bottomrule
@@
 
-\caption{Hexagonal heatmap for residuals vs distance on training and test data for three optimized models with different input sizes.
-The blue lines are smoothing curves produced by fitting gnealized additive models.}\label{fig:model-performance}
+\caption{Hexagonal heatmap for residuals vs predicted distance on training and test data for three optimized models with different input sizes.
+The red lines are smoothing curves produced by fitting gnealized additive models.
+The area over the zero line in light pink indicates under-prediction, and the area under the zero line in light green indicates over-prediction.}\label{fig:model-performance}
 \end{figure}
+
+\begin{table}
+
+\caption{\label{tab:performance-sub}The training and test performance of the $32 \times 32$ model presented with different model violations.  The best metrics are colored in red.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lr>{}l>{}l>{}l>{}l}
+\toprule
+Violations & \#samples & RMSE & $R^2$ & MAE & Huber loss\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Training set}}\\
+\hspace{1em}no violations & 1546 & \textcolor{black}{1.149} & \textcolor{black}{} & \textcolor{black}{1.080} & \textcolor{black}{0.592}\\
+\hspace{1em}non-linearity & 22184 & \textcolor{black}{0.625} & \textcolor{black}{0.851} & \textcolor{black}{0.474} & \textcolor{black}{0.177}\\
+\hspace{1em}heteroskedasticity & 10826 & \textcolor{black}{0.528} & \textcolor{black}{0.942} & \textcolor{black}{0.407} & \textcolor{black}{0.133}\\
+\hspace{1em}non-linearity + heteroskedasticity & 10221 & \textcolor{black}{0.593} & \textcolor{black}{0.920} & \textcolor{black}{0.439} & \textcolor{black}{0.158}\\
+\hspace{1em}non-normality & 10797 & \textcolor{red}{0.279} & \textcolor{red}{0.955} & \textcolor{red}{0.177} & \textcolor{red}{0.038}\\
+\hspace{1em}non-linearity + non-normality & 8835 & \textcolor{black}{0.472} & \textcolor{black}{0.923} & \textcolor{black}{0.274} & \textcolor{black}{0.090}\\
+\hspace{1em}heteroskedasticity + non-normality & 7870 & \textcolor{black}{0.354} & \textcolor{black}{0.905} & \textcolor{black}{0.211} & \textcolor{black}{0.052}\\
+\hspace{1em}non-linearity + heteroskedasticity + non-normality & 7721 & \textcolor{black}{0.432} & \textcolor{black}{0.874} & \textcolor{black}{0.261} & \textcolor{black}{0.078}\\
+\addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Test set}}\\
+\hspace{1em}no violations & 155 & \textcolor{black}{1.267} & \textcolor{black}{} & \textcolor{black}{1.174} & \textcolor{black}{0.685}\\
+\hspace{1em}non-linearity & 2218 & \textcolor{black}{0.787} & \textcolor{black}{0.771} & \textcolor{black}{0.578} & \textcolor{black}{0.257}\\
+\hspace{1em}heteroskedasticity & 1067 & \textcolor{black}{0.602} & \textcolor{black}{0.922} & \textcolor{black}{0.461} & \textcolor{black}{0.170}\\
+\hspace{1em}non-linearity + heteroskedasticity & 985 & \textcolor{black}{0.751} & \textcolor{black}{0.868} & \textcolor{black}{0.541} & \textcolor{black}{0.236}\\
+\hspace{1em}non-normality & 1111 & \textcolor{red}{0.320} & \textcolor{red}{0.942} & \textcolor{red}{0.194} & \textcolor{red}{0.049}\\
+\hspace{1em}non-linearity + non-normality & 928 & \textcolor{black}{0.600} & \textcolor{black}{0.879} & \textcolor{black}{0.335} & \textcolor{black}{0.133}\\
+\hspace{1em}heteroskedasticity + non-normality & 819 & \textcolor{black}{0.489} & \textcolor{black}{0.832} & \textcolor{black}{0.267} & \textcolor{black}{0.093}\\
+\hspace{1em}non-linearity + heteroskedasticity + non-normality & 717 & \textcolor{black}{0.620} & \textcolor{black}{0.762} & \textcolor{black}{0.339} & \textcolor{black}{0.140}\\
+\bottomrule
+\end{tabular}}
+\end{table}
 
@@
 
-Figure \ref{fig:rd-human} displays the \(D - \hat{D}\) vs \(D\) plot for
-actual residual plots used in the experiment. The behaviour of the
-prediction is roughly the same as what we can be observed from the
-evaluation of the test data. The model over-predicts \(\hat{D}\) when
-\(D\) is close to zero, and under-predicts \(\hat{D}\) when \(D\) is
-greater than \(6\). However, for residual plots with non-linearity
-issues, we do not observe the over-prediction for \(D > 6\) because
-there are not many cases in that area. And the over-prediction happens
-at around \(2 < D < 4\).
-
 For null plots, a histogram of the prediction is provided in Figure
@@
 decreases. There are some outliers with \(\hat{D} > 5\).
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/rd-human-1}
-
-}
-
-\caption{Residuals vs distance plot on lineups used in the human subject experiment. The blue line is a smoothing curve produced by fitting a gnealized additive model.}\label{fig:rd-human}
-\end{figure}
 
@@
 
-\caption{Histogram of predicted distance on null plots used in the human subject experiment.}\label{fig:hist-null-human}
+\caption{Histogram of predicted distance on null plots used in the human subject experiment. Sample quantiles are drawn in different colors.}\label{fig:hist-null-human}
 \end{figure}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-3-1} \end{center}
 
@@
 There are other kinds of residual departures like autocorrelation that
-are not considered in this study. The primary goal of this paper is to
+are not considered in this study. The primary goal of this study is to
 establish a new way of evaluating residual plot and conducting visual

### 2024-02-12 paper/paper.Rmd
update paper

@@
   select(-type) %>%
+  mutate(n = format(n)) %>%
   mutate(across(RMSE:MAE, ~format(.x, digits = 3))) %>%
@@
 The main characteristic is the model violations are introduced separately, meaning non-linearity and heteroskedasticity will not co-exist in one lineup but assigned uniformly to all lineups.
-Additionally, non-normality and multiple predictors are not considered in their experimental design.
+Additionally, non-normality and multiple predictors are not considered in the experimental design.
 
 ### Model performance on the human data
-
-For each lineup used in @li2023plot, there are one actual residual plot and $19$ null plots.
-While the distance $D$ for the actual residual plot depends on the underlying data generating process, the distance $D$ for the null plots are zeros.
-We have used our optimized computer vision model to predict $\hat{D}$ for both the actual residual plots and the null plots.
-
-The performance metrics of $\hat{D}$ for actual residual plots are provided in Table \ref{tab:experiment-performance}.
-It can be observed that all the performance metrics are slightly worse than the metrics evaluated on the test data.
-Still, the mean absolute error remains at a low level, and the linearly correlation between the prediction and the true value is still very high.
-Lineups with non-linearity issues are more difficult to be predicted than those with heteroskedasticity issues.
-
-For null plots, a histogram of the prediction is provided in Figure \ref{fig:hist-null-human}.
-In most cases, $\hat{D}$ will be predicted at around $1$.
-And as $\hat{D}$ increases, the number of predictions decreases.
-The 90%, 95% and 99% sample quantiles are at around
 
@@
 
-```{r hist-null-human, fig.cap = "Histogram of predicted distance on null plots used in the human subject experiment. Sample quantiles are drawn in different colors."}
+For each lineup used in @li2023plot, there is one actual residual plot and 19 null plots.
+While the distance $D$ for the actual residual plot depends on the underlying data generating process, the distance $D$ for the null plots is zero.
+We have used our optimized computer vision model to predict $\hat{D}$ for both the actual residual plots and the null plots.
+
+The performance metrics of $\hat{D}$ for actual residual plots are outlined in Table \ref{tab:experiment-performance}.
+It's notable that all performance metrics are slightly worse than those evaluated on the test data.
+Nevertheless, the mean absolute error remains at a low level, and the linear correlation between the prediction and the true value remains very high.
+Lineups with non-linearity issues are more challenging to predict than those with heteroskedasticity issues.
+
+For null plots, a histogram of the predictions is provided in Figure \ref{fig:hist-null-human}.
+In most cases, $\hat{D}$ is predicted to be around $1$.
+As $\hat{D}$ increases, the number of predictions decreases.
+The 90% and 95% sample quantiles are approximately at $2$, and the 99% sample quantile is around $2.5$.
+These sample quantiles can serve as global critical values for determining whether a residual plot needs to be rejected.
+We have also computed the 95% sample quantiles by assessing 200 null plots for each fitted model used in the experiment.
+The difference between decisions based on these two types of critical values is minor, with decisions aligning `r experiment %>% filter(!attention_check) %>% filter(!null_lineup) %>% summarise(q = mean(model_reject == (2.03 <= vss)) * 100) %>% pull(q) %>% format(digits = 3)`% of the time.
+Therefore, we will be using the 95% sample quantiles obtained from each fitted model in the following analysis.
+
+Figure \ref{fig:human-mosaic} and Table \ref{tab:human-table} provide insights into the agreement between decisions made by the computer vision model and visual tests conducted by humans.
+It can be observed that visual tests are less sensitive than the computer vision model, as they reject almost only when the computer vision model does.
+The agreement rates between visual tests and the computer vision model are 67.96% and 66.84% for residual plots containing heteroskedasticity and non-linearity patterns, respectively.
+
+When plotting the decision against the distance, as depicted in Figure \ref{fig:power}, we observe that for the computer vision model, when the distance $D > 3$, almost all residual plots will be rejected.
+However, for visual tests conducted by humans, the threshold is higher at $D > 5$.
+Visual tests also demonstrate low sensitivity to residual plots with small distances compared to the computer vision model.
+
+Figure \ref{fig:delta} displays the scatter plot of the weighted detection rate vs the $\delta$-difference.
+In the experiment conducted in @li2023plot, participants were allowed to make multiple selections for a lineup.
+The weighted detection rate is computed by assigning weights to each detection.
+If the participant selects zero plots, the corresponding weight is 0.05; otherwise, the weight is 1 divided by the number of selections.
+The $\delta$-difference is defined by @chowdhury2018measuring as
+
+\begin{equation}
+\delta = \hat{D} - \underset{j}{max}\left(\hat{D}_{null}^{(j)}\right) \quad \text{for}~j = 1,...,m-1,
+\end{equation}
+
+\noindent where $\hat{D}_{null}^{(j)}$ is the approximated distance for the $j$-th null plot, and $m$ is the number of plots in a lineup.
+
+Figure \ref{fig:delta} indicates that the weighted detection rate increases as the $\delta$-difference increases, particularly when the $\delta$-difference is greater than zero.
+A negative $\delta$-difference suggests that there is at least one null plot in the lineup with a stronger visual signal than the actual residual plot.
+In some instances, the weighted detection rate is close to one, yet the $\delta$-difference is negative.
+This discrepancy implies that the distance measure, or the approximated distance, may not perfectly reflect actual human behavior.
+
+```{r hist-null-human, fig.pos =  "!h", fig.cap = "Histogram of predicted distance on null plots used in the human subject experiment. Sample quantiles are drawn in different colors."}
 lineup_vss <- readRDS(here::here("paper/data/lineup_vss.rds"))
@@
 
-```{r fig.cap = "A weighted detection rate vs $\\delta$-differnence plot."}
-
+```{r human-table}
+experiment %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  group_by(type) %>%
+  summarise(n = format(n()), agree = format(sum(model_reject == reject)), rate = format(mean(model_reject == reject), digits = 4)) %>%
+  kableExtra::kable("latex", 
+                    booktabs = TRUE,
+                    escape = FALSE,
+                    col.names = c("Violations", "\\#Samples", "\\#Agreements", "Agreement rate"),
+                    caption = "Summary of decisions made by the visual test conducted by human and decisions made by the computer vision model.")
+```
+
+```{r human-mosaic, fig.pos = "!h", fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of visual test conditional on the computer vision model decision on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot.
+The visual test rejects less frequently than the computer vision model, and (almost) only rejects when the computer vision model does."}
+library(ggmosaic)
+
+experiment %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  mutate(type = ifelse(type == "non-linearity", "Non-linearity", "Heteroskedasticity")) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
+  mutate(model_reject = ifelse(model_reject, "Reject", "No")) %>%
+  mutate(reject = ifelse(reject, "Reject", "No")) %>%
+  mutate(across(c(model_reject, reject), ~factor(.x, levels = c("Reject", "No")))) %>%
+  ggplot() +
+  geom_mosaic(aes(x = ggmosaic::product(reject, model_reject), 
+                  fill = reject)) +
+  facet_grid(~type) +
+  ylab("Visual tests reject") +
+  xlab("Computer vision model rejects") +
+  # labs(fill = "Visual tests reject") +
+  scale_fill_brewer("", palette = "Dark2") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  coord_fixed()
+```
+
+```{r power, fig.pos = "!h", fig.cap = "Comparison of decisions made by visual tests and the computer vision model.
+The data points are fitted with logistic regression models with no intercept but an offset equals to $log(0.05/0.95)$."}
+
+max_actual_ss <- max(filter(experiment, !attention_check, !null_lineup)$actual_ss)
+
+model_glm_pred_poly <- glm(model_reject ~ actual_ss - 1, 
+    data = filter(experiment, !attention_check, !null_lineup, type == "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
+    offset = offset0,
+    family = binomial()) %>%
+  predict(data.frame(actual_ss = seq(0, max_actual_ss, 0.01), offset0 = log(0.05/0.95)),
+          type = "response") %>%
+  data.frame(d = seq(0, max_actual_ss, 0.01), type = "Non-linearity", name = "Computer vision model", value = .) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
+
+visual_glm_pred_poly <- glm(reject ~ actual_ss - 1, 
+    data = filter(experiment, !attention_check, !null_lineup, type == "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
+    offset = offset0,
+    family = binomial()) %>%
+  predict(data.frame(actual_ss = seq(0, max_actual_ss, 0.01), offset0 = log(0.05/0.95)),
+          type = "response") %>%
+  data.frame(d = seq(0, max_actual_ss, 0.01), type = "Non-linearity", name = "Visual test", value = .) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
+
+conv_glm_pred_poly <- glm(conventional_reject ~ actual_ss - 1, 
+    data = filter(experiment, !attention_check, !null_lineup, type == "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
+    offset = offset0,
+    family = binomial()) %>%
+  predict(data.frame(actual_ss = seq(0, max_actual_ss, 0.01), offset0 = log(0.05/0.95)),
+          type = "response") %>%
+  data.frame(d = seq(0, max_actual_ss, 0.01), type = "Non-linearity", name = "conventional", value = .) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
+
+model_glm_pred_heter <- glm(model_reject ~ actual_ss - 1, 
+    data = filter(experiment, !attention_check, !null_lineup, type != "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
+    offset = offset0,
+    family = binomial()) %>%
+  predict(data.frame(actual_ss = seq(0, max_actual_ss, 0.01), offset0 = log(0.05/0.95)),
+          type = "response") %>%
+  data.frame(d = seq(0, max_actual_ss, 0.01), type = "Heteroskedasticity", name = "Computer vision model", value = .) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
+
+visual_glm_pred_heter <- glm(reject ~ actual_ss - 1, 
+    data = filter(experiment, !attention_check, !null_lineup, type != "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
+    offset = offset0,
+    family = binomial()) %>%
+  predict(data.frame(actual_ss = seq(0, max_actual_ss, 0.01), offset0 = log(0.05/0.95)),
+          type = "response") %>%
+  data.frame(d = seq(0, max_actual_ss, 0.01), type = "Heteroskedasticity", name = "Visual test", value = .) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
+
+conv_glm_pred_heter <- glm(conventional_reject ~ actual_ss - 1, 
+    data = filter(experiment, !attention_check, !null_lineup, type != "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
+    offset = offset0,
+    family = binomial()) %>%
+  predict(data.frame(actual_ss = seq(0, max_actual_ss, 0.01), offset0 = log(0.05/0.95)),
+          type = "response") %>%
+  data.frame(d = seq(0, max_actual_ss, 0.01), type = "Heteroskedasticity", name = "conventional", value = .) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
+
+mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
+  filter(!null_lineup, !attention_check) %>%
+  pivot_longer(c(model_reject, reject)) %>%
+  mutate(name = ifelse(name == "model_reject", "Computer vision model", "Visual test")) %>%
+  ggplot() +
+  ggbeeswarm::geom_quasirandom(aes(actual_ss, value), size = 0.5) +
+  geom_line(data = model_glm_pred_poly, aes(d, value + 1, col = "Computer vision model"), linewidth = 0.8) +
+  geom_line(data = model_glm_pred_heter, aes(d, value + 1, col = "Computer vision model"), linewidth = 0.8) +
+  geom_line(data = visual_glm_pred_poly, aes(d, value + 1, col = "Visual test"), linewidth = 0.8) +
+  geom_line(data = visual_glm_pred_heter, aes(d, value + 1, col = "Visual test"), linewidth = 0.8) +
+  geom_line(data = mutate(model_glm_pred_poly, name = "Visual test"), aes(d, value + 1, col = "Computer vision model"), linetype = 2, linewidth = 0.8) +
+  geom_line(data = mutate(model_glm_pred_heter, name = "Visual test"), aes(d, value + 1, col = "Computer vision model"), linetype = 2, linewidth = 0.8) +
+  geom_line(data = mutate(visual_glm_pred_poly, name = "Computer vision model"), aes(d, value + 1, col = "Visual test"), linetype = 2, linewidth = 0.8) +
+  geom_line(data = mutate(visual_glm_pred_heter, name = "Computer vision model"), aes(d, value + 1, col = "Visual test"), linetype = 2, linewidth = 0.8) +
+  facet_grid(name ~ type) +
+  theme_light() +
+  scale_color_brewer("", palette = "Dark2") +
+  theme(legend.position = "bottom") +
+  xlab("D") +
+  ylab("Reject")
+
+```
+
+```{r delta, fig.pos = "!h", fig.cap = "A weighted detection rate vs $\\delta$-differnence plot."}
 experiment %>%
@@
   ggplot() +
-  geom_point(aes(delta_diff, prop_detect)) +
+  geom_point(aes(delta_diff, prop_detect), alpha = 0.3) +
   geom_smooth(aes(delta_diff, prop_detect), se = FALSE) +
   ylab("Weighted detection rate") +
-  xlab(expression(gamma-difference)) +
+  xlab(expression(delta-difference)) +
   theme_light()
@@
 
-```{r fig.cap = "A weighted detection rate vs $\\gamma$-number of extreme nulls plot."}
+```{r fig.pos = "!h", fig.cap = "A weighted detection rate vs $\\gamma$-number of extreme nulls plot.", eval = FALSE}
 experiment %>%
@@
 
-```{r fig.cap = "A weighted detection rate vs $\\delta$-differnence plot for lineups evaluated by 11 people."}
-experiment %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  filter(x_dist == "uniform") %>%
-  left_join(lineup_vss %>%
-              group_by(unique_lineup_id) %>%
-              summarise(delta_diff = first(delta_diff), gamma_diff = first(gamma_diff))) %>%
-  ggplot() +
-  geom_point(aes(delta_diff, prop_detect)) +
-  geom_smooth(aes(delta_diff, prop_detect), se = FALSE) +
-  ylab("Weighted detection rate") +
-  xlab(expression(delta-difference)) +
-  theme_light()
-```
-
-```{r fig.cap = "A weighted detection rate vs $\\gamma$-number of extreme nulls plot for lineups evaluated by 11 people."}
-experiment %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  filter(x_dist == "uniform") %>%
-  left_join(lineup_vss %>%
-              group_by(unique_lineup_id) %>%
-              summarise(delta_diff = first(delta_diff), gamma_diff = first(gamma_diff))) %>%
-  ggplot() +
-  geom_point(aes(gamma_diff, prop_detect)) +
-  ylab("Weighted detection rate") +
-  xlab(expression(gamma-number~of~extreme~nulls)) +
-  theme_light()
-```
-
-## When the model works
-
-- simple examples (non-linearity, heteroskedasticity, ...)
-- datasaurus
-
-## When the model does not work
-
-- human detect but model does not
-- cartoon residuals?
-
-## Workflow: how one use this model? (small showcase)
-
-# Dicussion
-
-There are other kinds of residual departures like autocorrelation that are not considered in this study.
+## Data examples
+
+### Boston Housing
+
+```{r fig.pos = "!h", fig.cap = "Residual plot of the regression model fitted on the Boston housing data."}
+
+housing <- read_csv(here::here("paper/data/housing.csv"))
+mod <- lm(MEDV ~ ., data = housing)
+
+my_vi <- autovi::auto_vi(fitted_mod = mod)
+```
+
+```{r}
+if (!file.exists(here::here("paper/data/boston_check.rds"))) {
+  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  my_vi$check(null_draws = 200L, boot_draws = 200L)
+  saveRDS(my_vi$check_result, here::here("paper/data/boston_check.rds"))
+}
+
+```
+
+```{r}
+my_vi$check_result <- readRDS(here::here("paper/data/boston_check.rds"))
+```
+
+```{r fig.pos = "!h"}
+p1 <- my_vi$plot_resid()
+p2 <- my_vi$summary_plot() +
+  ggtitle("", subtitle = "") +
+  theme(legend.position = "bottom", legend.box = "vertical") +
+  labs(linetype = "")
+
+patchwork::wrap_plots(p1, p2)
+```
+
+### Insurance
+
+```{r fig.pos = "!h", fig.cap = "Residual plot of the regression model fitted on the Boston housing data."}
+insurance <- read_csv(here::here("paper/data/insurance.csv"))
+mod <- lm(charges ~ ., data = insurance)
+
+my_vi <- autovi::auto_vi(fitted_mod = mod)
+```
+
+```{r}
+if (!file.exists(here::here("paper/data/insurance_check.rds"))) {
+  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  my_vi$check(null_draws = 200L, boot_draws = 200L)
+  saveRDS(my_vi$check_result, here::here("paper/data/insurance_check.rds"))
+}
+
+```
+
+```{r}
+my_vi$check_result <- readRDS(here::here("paper/data/insurance_check.rds"))
+```
+
+```{r fig.pos = "!h"}
+p1 <- my_vi$plot_resid()
+p2 <- my_vi$summary_plot() +
+  ggtitle("", subtitle = "") +
+  theme(legend.position = "bottom", legend.box = "vertical") +
+  labs(linetype = "")
+
+patchwork::wrap_plots(p1, p2)
+```
+
+### Wine Quality
+
+```{r}
+wine <- read_csv(here::here("paper/data/WineQT.csv"))
+mod <- lm(quality ~ ., data = select(wine, -Id))
+
+my_vi <- autovi::auto_vi(fitted_mod = mod)
+```
+
+```{r}
+if (!file.exists(here::here("paper/data/wine_check.rds"))) {
+  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  my_vi$check(null_draws = 200L, boot_draws = 200L)
+  saveRDS(my_vi$check_result, here::here("paper/data/wine_check.rds"))
+}
+```
+
+```{r}
+my_vi$check_result <- readRDS(here::here("paper/data/wine_check.rds"))
+```
+
+```{r fig.pos = "!h"}
+p1 <- my_vi$plot_resid()
+p2 <- my_vi$summary_plot() +
+  ggtitle("", subtitle = "") +
+  theme(legend.position = "bottom", legend.box = "vertical") +
+  labs(linetype = "")
+
+patchwork::wrap_plots(p1, p2)
+```
+
+### Datasaurus
+
+```{r}
+dino <- datasauRus::datasaurus_dozen %>% filter(dataset == "dino")
+mod <- lm(y ~ ., data = select(dino, -dataset))
+
+my_vi <- autovi::auto_vi(fitted_mod = mod)
+```
+
+```{r}
+if (!file.exists(here::here("paper/data/dino_check.rds"))) {
+  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  my_vi$check(null_draws = 200L, boot_draws = 200L)
+  saveRDS(my_vi$check_result, here::here("paper/data/dino_check.rds"))
+}
+```
+
+```{r}
+my_vi$check_result <- readRDS(here::here("paper/data/dino_check.rds"))
+```
+
+```{r fig.pos = "!h"}
+p1 <- my_vi$plot_resid()
+p2 <- my_vi$summary_plot() +
+  ggtitle("", subtitle = "") +
+  theme(legend.position = "bottom", legend.box = "vertical") +
+  labs(linetype = "")
+
+patchwork::wrap_plots(p1, p2)
+```
+
+<!-- ## When the model works -->
+
+<!-- - simple examples (non-linearity, heteroskedasticity, ...) -->
+<!-- - datasaurus -->
+
+<!-- ## When the model does not work -->
+
+<!-- - human detect but model does not -->
+<!-- - cartoon residuals? -->
+
+<!-- ## Workflow: how one use this model? (small showcase)  -->
+
+<!-- # Dicussion -->
+
+<!-- There are other kinds of residual departures like autocorrelation that are not considered in this study.
 The primary goal of this study is to establish a new way of evaluating residual plot and conducting visual test with computer vision models.
 Building computer vision models for other model violations and other types of diagnostic plots could be future directions of this field.
+-->
 

### 2024-02-12 paper/paper.tex
update paper

@@
 \resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lr>{}l>{}l>{}l>{}l}
+\begin{tabular}[t]{ll>{}l>{}l>{}l>{}l}
 \toprule
@@
 Additionally, non-normality and multiple predictors are not considered
-in their experimental design.
+in the experimental design.
 
@@
 data}\label{model-performance-on-the-human-data}}
-
-For each lineup used in \citet{li2023plot}, there are one actual
-residual plot and \(19\) null plots. While the distance \(D\) for the
-actual residual plot depends on the underlying data generating process,
-the distance \(D\) for the null plots are zeros. We have used our
-optimized computer vision model to predict \(\hat{D}\) for both the
-actual residual plots and the null plots.
-
-The performance metrics of \(\hat{D}\) for actual residual plots are
-provided in Table \ref{tab:experiment-performance}. It can be observed
-that all the performance metrics are slightly worse than the metrics
-evaluated on the test data. Still, the mean absolute error remains at a
-low level, and the linearly correlation between the prediction and the
-true value is still very high. Lineups with non-linearity issues are
-more difficult to be predicted than those with heteroskedasticity
-issues.
-
-For null plots, a histogram of the prediction is provided in Figure
-\ref{fig:hist-null-human}. In most cases, \(\hat{D}\) will be predicted
-at around \(1\). And as \(\hat{D}\) increases, the number of predictions
-decreases. There are some outliers with \(\hat{D} > 5\).
 
@@
 
-\begin{figure}
+For each lineup used in \citet{li2023plot}, there is one actual residual
+plot and 19 null plots. While the distance \(D\) for the actual residual
+plot depends on the underlying data generating process, the distance
+\(D\) for the null plots is zero. We have used our optimized computer
+vision model to predict \(\hat{D}\) for both the actual residual plots
+and the null plots.
+
+The performance metrics of \(\hat{D}\) for actual residual plots are
+outlined in Table \ref{tab:experiment-performance}. It's notable that
+all performance metrics are slightly worse than those evaluated on the
+test data. Nevertheless, the mean absolute error remains at a low level,
+and the linear correlation between the prediction and the true value
+remains very high. Lineups with non-linearity issues are more
+challenging to predict than those with heteroskedasticity issues.
+
+For null plots, a histogram of the predictions is provided in Figure
+\ref{fig:hist-null-human}. In most cases, \(\hat{D}\) is predicted to be
+around \(1\). As \(\hat{D}\) increases, the number of predictions
+decreases. The 90\% and 95\% sample quantiles are approximately at
+\(2\), and the 99\% sample quantile is around \(2.5\). These sample
+quantiles can serve as global critical values for determining whether a
+residual plot needs to be rejected. We have also computed the 95\%
+sample quantiles by assessing 200 null plots for each fitted model used
+in the experiment. The difference between decisions based on these two
+types of critical values is minor, with decisions aligning 93.8\% of the
+time. Therefore, we will be using the 95\% sample quantiles obtained
+from each fitted model in the following analysis.
+
+Figure \ref{fig:human-mosaic} and Table \ref{tab:human-table} provide
+insights into the agreement between decisions made by the computer
+vision model and visual tests conducted by humans. It can be observed
+that visual tests are less sensitive than the computer vision model, as
+they reject almost only when the computer vision model does. The
+agreement rates between visual tests and the computer vision model are
+67.96\% and 66.84\% for residual plots containing heteroskedasticity and
+non-linearity patterns, respectively.
+
+When plotting the decision against the distance, as depicted in Figure
+\ref{fig:power}, we observe that for the computer vision model, when the
+distance \(D > 3\), almost all residual plots will be rejected. However,
+for visual tests conducted by humans, the threshold is higher at
+\(D > 5\). Visual tests also demonstrate low sensitivity to residual
+plots with small distances compared to the computer vision model.
+
+Figure \ref{fig:delta} displays the scatter plot of the weighted
+detection rate vs the \(\delta\)-difference. In the experiment conducted
+in \citet{li2023plot}, participants were allowed to make multiple
+selections for a lineup. The weighted detection rate is computed by
+assigning weights to each detection. If the participant selects zero
+plots, the corresponding weight is 0.05; otherwise, the weight is 1
+divided by the number of selections. The \(\delta\)-difference is
+defined by \citet{chowdhury2018measuring} as
+
+\begin{equation}
+\delta = \hat{D} - \underset{j}{max}\left(\hat{D}_{null}^{(j)}\right) \quad \text{for}~j = 1,...,m-1,
+\end{equation}
+
+\noindent where \(\hat{D}_{null}^{(j)}\) is the approximated distance
+for the \(j\)-th null plot, and \(m\) is the number of plots in a
+lineup.
+
+Figure \ref{fig:delta} indicates that the weighted detection rate
+increases as the \(\delta\)-difference increases, particularly when the
+\(\delta\)-difference is greater than zero. A negative
+\(\delta\)-difference suggests that there is at least one null plot in
+the lineup with a stronger visual signal than the actual residual plot.
+In some instances, the weighted detection rate is close to one, yet the
+\(\delta\)-difference is negative. This discrepancy implies that the
+distance measure, or the approximated distance, may not perfectly
+reflect actual human behavior.
+
+\begin{figure}[!h]
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-4-1}
+\begin{table}
+
+\caption{\label{tab:human-table}Summary of decisions made by the visual test conducted by human and decisions made by the computer vision model.}
+\centering
+\begin{tabular}[t]{llll}
+\toprule
+Violations & \#Samples & \#Agreements & Agreement rate\\
+\midrule
+heteroskedasticity & 540 & 367 & 0.6796\\
+non-linearity & 576 & 385 & 0.6684\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/human-mosaic-1}
 
@@
 
-\caption{A weighted detection rate vs $\delta$-differnence plot.}\label{fig:unnamed-chunk-4}
+\caption{Rejection rate ($p$-value $\leq0.05$) of visual test conditional on the computer vision model decision on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot.
+The visual test rejects less frequently than the computer vision model, and (almost) only rejects when the computer vision model does.}\label{fig:human-mosaic}
 \end{figure}
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-5-1}
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/power-1}
 
@@
 
-\caption{A weighted detection rate vs $\gamma$-number of extreme nulls plot.}\label{fig:unnamed-chunk-5}
+\caption{Comparison of decisions made by visual tests and the computer vision model.
+The data points are fitted with logistic regression models with no intercept but an offset equals to $log(0.05/0.95)$.}\label{fig:power}
 \end{figure}
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-6-1}
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/delta-1}
 
@@
 
-\caption{A weighted detection rate vs $\delta$-differnence plot for lineups evaluated by 11 people.}\label{fig:unnamed-chunk-6}
+\caption{A weighted detection rate vs $\delta$-differnence plot.}\label{fig:delta}
 \end{figure}
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-7-1}
-
-}
-
-\caption{A weighted detection rate vs $\gamma$-number of extreme nulls plot for lineups evaluated by 11 people.}\label{fig:unnamed-chunk-7}
-\end{figure}
-
-\hypertarget{when-the-model-works}{%
-\subsection{When the model works}\label{when-the-model-works}}
-
-\begin{itemize}
-\tightlist
-\item
-  simple examples (non-linearity, heteroskedasticity, \ldots)
-\item
-  datasaurus
-\end{itemize}
-
-\hypertarget{when-the-model-does-not-work}{%
-\subsection{When the model does not
-work}\label{when-the-model-does-not-work}}
-
-\begin{itemize}
-\tightlist
-\item
-  human detect but model does not
-\item
-  cartoon residuals?
-\end{itemize}
-
-\hypertarget{workflow-how-one-use-this-model-small-showcase}{%
-\subsection{Workflow: how one use this model? (small
-showcase)}\label{workflow-how-one-use-this-model-small-showcase}}
-
-\hypertarget{dicussion}{%
-\section{Dicussion}\label{dicussion}}
-
-There are other kinds of residual departures like autocorrelation that
-are not considered in this study. The primary goal of this study is to
-establish a new way of evaluating residual plot and conducting visual
-test with computer vision models. Building computer vision models for
-other model violations and other types of diagnostic plots could be
-future directions of this field.
+\hypertarget{data-examples}{%
+\subsection{Data examples}\label{data-examples}}
+
+\hypertarget{boston-housing}{%
+\subsubsection{Boston Housing}\label{boston-housing}}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-8-1} \end{center}
+
+\hypertarget{advertising}{%
+\subsubsection{Advertising}\label{advertising}}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-12-1} \end{center}
+
+\hypertarget{wine-quality}{%
+\subsubsection{Wine Quality}\label{wine-quality}}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-16-1} \end{center}
+
+\hypertarget{datasaurus}{%
+\subsubsection{Datasaurus}\label{datasaurus}}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-20-1} \end{center}
 

### 2024-02-15 paper/paper.Rmd
update paper

@@
   Automated assessment of residual plots with computer vision models
+type: Draft paper (Incomplete)
 author:
@@
 
-\newpage
-\tableofcontents
-\newpage
+<!-- \newpage -->
+<!-- \tableofcontents -->
+<!-- \newpage -->
 
@@
 Unlike conventional statistical tests that can be performed computationally in statistical software, the lineup protocol requires human evaluation of images.
-This characteristic makes it less suitable for large-scale applications, given the associated high labor costs and time requirements.
+This characteristic makes it less suitable for large-scale applications, given the associated high labour costs and time requirements.
 
@@
 
-The utilization of computers to interpret data plots has a rich history, with early efforts such as "Scagnostics" by @tukey1985computer, focusing on scatterplot diagnostics.
+The utilization of computers to interpret data plots has a rich history, with early efforts such as "Scagnostics" by @tukey1985computer, focusing on scatter plot diagnostics.
 @wilkinson2005graph expanded on this work, introducing graph theoretic scagnostics, which encompassed computable measures applied to planar proximity graphs.
@@
 The vertical spread of the data points varies with the fitted values.
-This often indicates the existence of heteroskedasticity."}
-set.seed(10131)
+This often indicates the existence of heteroskedasticity.
+The Breusch-Pagan test rejects this residual plot at 95\\% significance level ($p\\text{-value} = 0.046$)."}
+set.seed(452)
 ori_x <- rand_lognormal()
@@
 As for the aesthetics of the residual plot, a practical solution is to consistently present residual plots in the same style to the model.
-This implies that the model can not accept arbitrary images as input but requires the use of the same preprocessing pipeline to convert residuals and fitted values into a standardized-style residual plot.
+This implies that the model can not accept arbitrary images as input but requires the use of the same pre-processing pipeline to convert residuals and fitted values into a standardized-style residual plot.
 
@@
 For example, if the data generating process contains variables correlated with any column of $\boldsymbol{X}$ but not included in $\boldsymbol{X}$, causing an omitted variable problem, $P$ will be different from $Q$ because the residual operator obtained from the fitted regression model will not be the same as $\boldsymbol{R}$.
-Besides, if the $\boldsymbol{\varepsilon}$ follows a non-normal distribution such as a multivariate lognormal distribution, $P$ will usually be skewed and has a long tail.
+Besides, if the $\boldsymbol{\varepsilon}$ follows a non-normal distribution such as a multivariate log-normal distribution, $P$ will usually be skewed and has a long tail.
 
@@
 
-We tried to train a computer vision model to approximate this distance measure with a residual plot.
-Let $D$ be the result of equation \ref{eq:kl-0} indicating the extent of the model violations, and our estimator $\hat{D}$ is formulated as
+We tried to train a computer vision model to approximate this distance measure with a residual plot. Let $D$ be the result of equation \ref{eq:kl-0}, and our estimator $\hat{D}$ is formulated as
 
@@
 With the approximated distance $\hat{D}$, we will be able to know how different the underlying distribution of the residuals is from a good residual distribution.
+The approximated distance $\hat{D}$ can also be used as an index of the model violations.
 It also provides information for the strength of the visual signal embedded in the residual plot.
@@
 This suggests the error of the approximation will vary depends on whether the observed residual plot is representative or not.
+
+### Approximated distance as an Model Violations Index (MVI)
+
+In Section \ref{output-formats}, we discuss the output formats of the computer vision model presented in this paper.
+One notable advantage highlighted is its interpretability.
+The approximated distance $\hat{D}$ is an estimator of the difference between the actual residual distribution and the reference residual distribution.
+This difference primarily arises from deviations in model assumptions.
+The magnitude of $D$ directly reflects the degree of these deviations, thus making $\hat{D}$ instrumental in forming a model violation index.
+
+Note that since the Kullback-Leibler divergence could be affected by the number of random variables involved in the evaluation.
+In general, more observations will result in greater distance $D$.
+This does not mean $\hat{D}$ can not correctly represent the extent of the model violations since more observations in a residual plot usually lead to stronger visual signal strength as the underlying visual patterns will have higher chance to be revealed except for cases like serious overlapping.
+For the same data generating process, $D$ usually grows logarithmically as the number of observations increases.
+This is because $D = \text{log}(1 + D_{KL})$, and $D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}$ under the independence assumption.
+
+Note that the Kullback-Leibler divergence might be influenced by the number of random variables involved in its evaluation.
+Generally, a larger number of observations will lead to a greater distance $D$.
+However, this does not imply that $\hat{D}$ fails to accurately represent the extent of model violations.
+In fact, when examining residual plots with more observations, we often observe a stronger visual signal strength, as the underlying patterns are more likely to be revealed, except in cases of significant overlapping.
+
+For a consistent data generating process, $D$ typically increases logarithmically with the number of observations.
+This behaviour comes from the relationship $D = \text{log}(1 + D_{KL})$, where $D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}$ under the assumption of independence.
+
+Therefore, the Model Violations Index (MVI) can be proposed as
+
+\begin{equation}
+\label{eq:mvi}
+\text{MVI} = C + \hat{D} - log(n),
+\end{equation}
+
+\noindent where $C$ is a large enough constant keeping the result positive.
 
@@
 Non-normality violations arise from specifying a non-normal distribution for $\boldsymbol{\varepsilon}$.
-In the synthetic data simulation, four distinct error distributions are considered, including discrete, uniform, normal, and lognormal distributions, as presented in Figure \ref{fig:different-e}.
+In the synthetic data simulation, four distinct error distributions are considered, including discrete, uniform, normal, and log-normal distributions, as presented in Figure \ref{fig:different-e}.
 Each distribution imparts unique characteristics to the residuals.
-The discrete error distribution introduces discreteness in residuals, while the lognormal distribution typically yields outliers.
+The discrete error distribution introduces discreteness in residuals, while the log-normal distribution typically yields outliers.
 Uniform error distribution may result in residuals filling the entire space of the residual plot.
@@
 
-For each generated residual plot, we computed four scagnostics – "Monotonic," "Sparse," "Splines," and "Striped" – using the cassowaryr R package [@mason2022cassowaryr].
+For each generated residual plot, we computed four scagnostics – "Monotonic," "Sparse," "Splines," and "Striped" – using the `cassowaryr` R package [@mason2022cassowaryr].
 These computed measures, along with the number of observations from the fitted model, were provided as the second input for the computer vision model.
-Although other scagnostics are informative, they are currently unavailable due to a fatal bug in the compiled C program of the interp R package [@Albrecht2023interp], which may unpredictably crash the process.
+Although other scagnostics are informative, they are currently unavailable due to a fatal bug in the compiled C program of the `interp` R package [@Albrecht2023interp], which may unpredictably crash the process.
 For reproducibility, we excluded these scagnostics from the training data.
@@
 This is followed by a grayscale conversion layer utilizing the luma formula under the Rec.
-601 standard, which converts the color image to grayscale.
+601 standard, which converts the colour image to grayscale.
 Grayscale suffices for our task since data points are plotted in black.
@@
 To achieve an optimal deep learning model, hyperparameters like the learning rate often need to be fine-tuned using a tuner.
-In our study, we utilized the Bayesian optimization tuner from the KerasTuner Python library [@omalley2019kerastuner] for this purpose.
+In our study, we utilized the Bayesian optimization tuner from the `KerasTuner` Python library [@omalley2019kerastuner] for this purpose.
 A comprehensive list of hyperparameters is provided in Table \ref{tab:hyperparameter}.
@@
 
-- RMSE for the test set
-- R^2
-- Mean bias deviation to understand the overall bias
-- quantile loss to understand how well the model captures the entire distribution of D
-- Percentage of predictions within a tolerance interval (like 0.1)
-
-- comparison with human visual inference
-- overview of the human subject experiment
-- metrics
-- agreement rate
-- check cases that are aggreed
-- check cases that are disaggreed
+This section remains incomplete as we may modify the "Results" section later on.
 
@@
 The figure highlights that most under-predictions occurred when $\hat{D} < 3$, while over-predictions occurred predominantly when $3 < \hat{D} < 6$.
-To comprehend the reasons behind these deviations, we conducted a meta-analysis on $D - \hat{D}$ for all factors used in the data generating process.
-A regression tree was constructed using the rpart R package [@Terry2022rpart].
+To comprehend the reasons behind these deviations, we fitted a regression tree using the `rpart` R package [@Terry2022rpart] on $D - \hat{D}$ for all factors used in the data generating process.
 The results of the regression tree are provided in the Appendix.
 
-The meta-analysis revealed that most issues arose from non-linearity problems and the presence of a second predictor in the regression model.
+The regression tree revealed that most issues arose from non-linearity problems and the presence of a second predictor in the regression model.
 When the error distribution had a very small variance, all computer vision models tended to under-predict the distance.
@@
   geom_point(data = NULL, aes(5, 0, col = "under-prediction"), shape = 15) +
-  geom_hex(aes(effect_size, effect_size - vss), bins = 25) +
+  geom_hex(aes(vss, effect_size - vss), bins = 25) +
   geom_smooth(aes(vss, effect_size - vss), se = FALSE, col = "#e34a33") +
@@
 
-## Comparison with human visual inference
+## Model Violations Index (MVI)
+
+Figures \ref{fig:poly-index} and \ref{fig:heter-index} display the residual plots for fitted models exhibiting varying degrees of non-linearity and heteroskedasticity.
+Each residual plot's Model Violation Index (MVI) is computed using Equation \ref{eq:mvi} with $C = 10$.
+When $\text{MVI} > 8$, the visual patterns are notably strong and easily discernible by humans.
+In the range $6 < \text{MVI} < 8$, the visibility of the visual pattern diminishes as MVI decreases.
+Conversely, when $\text{MVI} < 6$, the visual pattern tends to become relatively faint and challenging to observe.
+Table \ref{tab:mvi} provides a summary of the MVI usage and it is applicable to other linear regression models.
+
+```{r mvi}
+data.frame(degree = c("Strong", "Moderate", "Weak"),
+           range = c("$\\text{MVI} > 8$", "$6 < \\text{MVI} < 8$", "$\\text{MVI} < 6$")) %>%
+  kableExtra::kable("latex", 
+                    booktabs = TRUE,
+                    escape = FALSE,
+                    col.names = c("Degree of model violations", "Range ($C$ = 10)"),
+                    caption = "Degree of model violations or the strength of the visual signals according to the Model Violation Index (MVI). The constant $C$ is set to be 10.")
+```
+
+```{r poly-index, fig.pos = "!h", fig.cap = "Residual plots generated from fitted models exhibiting varying degrees of non-linearity violations.
+The Model violations index (MVI) is displayed atop each residual plot.
+The non-linearity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$."}
+set.seed(1495)
+if (!file.exists(here::here("paper/data/poly_index.rds"))) {
+  e_sigma_factor <- 3
+  keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  poly_dgp <- phn_model(j = 3, sigma = 0.1^e_sigma_factor)
+  ori_dat <- poly_dgp$gen(300)
+
+  vss <- c()
+  actual_ss <- c()
+  dat_combined <- tibble()
+
+  for (e_sigma in seq(0.7, 1.8, length.out = 20)^e_sigma_factor) {
+    poly_dgp <- phn_model(j = 3, sigma = e_sigma)
+    dat <- poly_dgp$gen(300, computed = list(x1 = ori_dat$x1, e = ori_dat$e, k = ori_dat$k * e_sigma / 0.1^e_sigma_factor))
+    mod <- lm(y ~ x1, data = dat)
+    my_vi <- autovi::auto_vi(fitted_mod = mod, keras_mod = keras_mod)
+    vss <- c(vss, my_vi$vss())
+    actual_ss <- c(actual_ss, log(poly_dgp$sample_effect_size(dat) + 1))
+
+    dat_combined <- bind_rows(dat_combined, mutate(dat, e_sigma = e_sigma))
+  }
+
+  saveRDS(dat_combined, here::here("paper/data/poly_index.rds"))
+  saveRDS(vss, here::here("paper/data/poly_index_vss.rds"))
+}
+
+dat_combined <- readRDS(here::here("paper/data/poly_index.rds"))
+vss <- readRDS(here::here("paper/data/poly_index_vss.rds"))
+
+dat_combined %>%
+  autovi::AUTO_VI$plot_resid() +
+  facet_wrap(~e_sigma, scales = "free", 
+             labeller = function(...){ 
+               tibble(nb = map(glue("'MVI' == {format(vss - log(300) + 10, digits = 3)}"), str2expression))
+               },
+             ncol = 5) +
+  theme(strip.text = element_text(size = 10))
+```
+
+```{r heter-index, fig.pos = "!h", fig.cap = "Residual plots generated from fitted models exhibiting varying degrees of heteroskedasticity violations.
+The Model violations index (MVI) is displayed atop each residual plot.
+The heteroskedasticity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$."}
+set.seed(1495)
+if (!file.exists(here::here("paper/data/heter_index.rds"))) {
+  b_factor <- 2
+  keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  heter_dgp <- phn_model(include_z = FALSE, b = 0)
+  ori_dat <- poly_dgp$gen(300)
+
+  vss <- c()
+  actual_ss <- c()
+  dat_combined <- tibble()
+
+  for (b in rev(seq(0.2, 4, length.out = 20))^b_factor) {
+    heter_dgp <- phn_model(include_z = FALSE, b = b)
+    dat <- heter_dgp$gen(300, computed = list(x1 = ori_dat$x1, e = ori_dat$e))
+    mod <- lm(y ~ x1, data = dat)
+    my_vi <- autovi::auto_vi(fitted_mod = mod, keras_mod = keras_mod)
+    vss <- c(vss, my_vi$vss())
+    actual_ss <- c(actual_ss, log(heter_dgp$sample_effect_size(dat) + 1))
+
+    dat_combined <- bind_rows(dat_combined, mutate(dat, nb = -b))
+  }
+
+  saveRDS(dat_combined, here::here("paper/data/heter_index.rds"))
+  saveRDS(vss, here::here("paper/data/heter_index_vss.rds"))
+}
+
+dat_combined <- readRDS(here::here("paper/data/heter_index.rds"))
+vss <- readRDS(here::here("paper/data/heter_index_vss.rds"))
+
+dat_combined %>%
+  autovi::AUTO_VI$plot_resid() +
+  facet_wrap(~nb, scales = "free", 
+             labeller = function(...){ 
+               tibble(nb = map(glue("'MVI' == {format(vss - log(300) + 10, digits = 3)}"), str2expression))
+               },
+             ncol = 5) +
+  theme(strip.text = element_text(size = 10))
+```
+
+## Comparison with human visual inference and conventional tests
 
@@
 We have used our optimized computer vision model to predict $\hat{D}$ for both the actual residual plots and the null plots.
+Additionally, the appropriate conventional tests including the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey1969tests] for non-linearity and the Breusch-Pagan test  [@breusch1979simple] for heteroskedasticity were applied on the same data for comparison.
 
@@
 
-Figure \ref{fig:human-mosaic} and Table \ref{tab:human-table} provide insights into the agreement between decisions made by the computer vision model and visual tests conducted by humans.
-It can be observed that visual tests are less sensitive than the computer vision model, as they reject almost only when the computer vision model does.
-The agreement rates between visual tests and the computer vision model are 67.96% and 66.84% for residual plots containing heteroskedasticity and non-linearity patterns, respectively.
-
-When plotting the decision against the distance, as depicted in Figure \ref{fig:power}, we observe that for the computer vision model, when the distance $D > 3$, almost all residual plots will be rejected.
-However, for visual tests conducted by humans, the threshold is higher at $D > 5$.
-Visual tests also demonstrate low sensitivity to residual plots with small distances compared to the computer vision model.
+Likewise, Figure \ref{fig:conv-mosaic} and Table \ref{tab:conv-table} provide a summary of the agreement between decisions made by the computer vision model and conventional tests.
+The agreement rates between conventional tests and the computer vision model are 65.37% and 57.29% for residual plots containing heteroskedasticity and non-linearity patterns, respectively.
+These figures are lower than those presented in Table \ref{tab:human-table}, indicating that the computer vision model exhibits behaviour more akin to visual tests than conventional tests.
+Additionally, conventional tests tend to reject residual plots almost exclusively when the computer vision model does, but they also reject plots when the computer vision model does not.
+This suggests that conventional tests are more sensitive than the computer vision model.
+
+When plotting the decision against the distance, as illustrated in Figure \ref{fig:power}, several observations emerge.
+Firstly, for the computer vision model, when the distance $D > 3$, almost all residual plots are rejected.
+However, for visual tests conducted by humans, the threshold is higher, set at $D > 5$.
+And conventional tests exhibit some non-rejection cases for $D > 3$.
+Moreover, the computer vision model tends to have fewer rejected cases than conventional tests when $D < 2$.
+However, visual tests demonstrate the lowest sensitivity to residual plots with small distances.
+Therefore, tests based on the computer vision model are less sensitive to small deviations from model assumptions than conventional tests but more sensitive to moderate deviations.
+Rejection decisions are fitted by logistic regression models with no intercept terms but an offset equals to $\text{log}(0.05/0/95)$.
+The fitting curves for the computer vision model fall between those of conventional tests and visual tests for non-linearity and are slightly below conventional tests for heteroskedasticity.
+There is still potential to refine the computer vision model to better align its behaviour with visual tests.
 
@@
 In some instances, the weighted detection rate is close to one, yet the $\delta$-difference is negative.
-This discrepancy implies that the distance measure, or the approximated distance, may not perfectly reflect actual human behavior.
+This discrepancy implies that the distance measure, or the approximated distance, may not perfectly reflect actual human behaviour.
 
@@
                     col.names = c("Violations", "\\#Samples", "\\#Agreements", "Agreement rate"),
-                    caption = "Summary of decisions made by the visual test conducted by human and decisions made by the computer vision model.")
+                    caption = "Summary of decisions made by visual tests conducted by human and decisions made by the computer vision model.")
+```
+
+```{r conv-table}
+experiment %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  group_by(type) %>%
+  summarise(n = format(n()), agree = format(sum(model_reject == conventional_reject)), rate = format(mean(conventional_reject == reject), digits = 4)) %>%
+  kableExtra::kable("latex", 
+                    booktabs = TRUE,
+                    escape = FALSE,
+                    col.names = c("Violations", "\\#Samples", "\\#Agreements", "Agreement rate"),
+                    caption = "Summary of decisions made by conventional tests and decisions made by the computer vision model.")
 ```
@@
 
-```{r power, fig.pos = "!h", fig.cap = "Comparison of decisions made by visual tests and the computer vision model.
-The data points are fitted with logistic regression models with no intercept but an offset equals to $log(0.05/0.95)$."}
+```{r conv-mosaic, fig.pos = "!h", fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of conventional test conditional on the computer vision model decision on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot.
+The conventional test rejects more frequently than the computer vision model, and (almost) only rejects when the computer vision model does."}
+library(ggmosaic)
+
+experiment %>%
+  filter(!attention_check) %>%
+  filter(!null_lineup) %>%
+  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  mutate(type = ifelse(type == "non-linearity", "Non-linearity", "Heteroskedasticity")) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
+  mutate(model_reject = ifelse(model_reject, "Reject", "No")) %>%
+  mutate(conventional_reject = ifelse(conventional_reject, "Reject", "No")) %>%
+  mutate(across(c(model_reject, conventional_reject), ~factor(.x, levels = c("Reject", "No")))) %>%
+  ggplot() +
+  geom_mosaic(aes(x = ggmosaic::product(conventional_reject, model_reject), 
+                  fill = conventional_reject)) +
+  facet_grid(~type) +
+  ylab("Conventional tests reject") +
+  xlab("Computer vision model rejects") +
+  # labs(fill = "Visual tests reject") +
+  scale_fill_brewer("", palette = "Dark2") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  coord_fixed()
+```
+
+```{r power, fig.pos = "!h", fig.cap = "Comparison of decisions made by visual tests, conventional tests and the computer vision model.
+The data points are fitted with logistic regression models with no intercept but an offset equals to $\\text{log}(0.05/0.95)$."}
 
 max_actual_ss <- max(filter(experiment, !attention_check, !null_lineup)$actual_ss)
+max_poly_ss <- max(filter(experiment, !attention_check, !null_lineup, type == "polynomial")$actual_ss)
+max_heter_ss <- max(filter(experiment, !attention_check, !null_lineup, type != "polynomial")$actual_ss)
 
@@
     family = binomial()) %>%
-  predict(data.frame(actual_ss = seq(0, max_actual_ss, 0.01), offset0 = log(0.05/0.95)),
+  predict(data.frame(actual_ss = seq(0, max_poly_ss, 0.01), offset0 = log(0.05/0.95)),
           type = "response") %>%
-  data.frame(d = seq(0, max_actual_ss, 0.01), type = "Non-linearity", name = "Computer vision model", value = .) %>%
+  data.frame(d = seq(0, max_poly_ss, 0.01), type = "Non-linearity", name = "Computer vision model", value = .) %>%
   mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
@@
     family = binomial()) %>%
-  predict(data.frame(actual_ss = seq(0, max_actual_ss, 0.01), offset0 = log(0.05/0.95)),
+  predict(data.frame(actual_ss = seq(0, max_poly_ss, 0.01), offset0 = log(0.05/0.95)),
           type = "response") %>%
-  data.frame(d = seq(0, max_actual_ss, 0.01), type = "Non-linearity", name = "Visual test", value = .) %>%
+  data.frame(d = seq(0, max_poly_ss, 0.01), type = "Non-linearity", name = "Visual test", value = .) %>%
   mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
@@
     family = binomial()) %>%
-  predict(data.frame(actual_ss = seq(0, max_actual_ss, 0.01), offset0 = log(0.05/0.95)),
+  predict(data.frame(actual_ss = seq(0, max_poly_ss, 0.01), offset0 = log(0.05/0.95)),
           type = "response") %>%
-  data.frame(d = seq(0, max_actual_ss, 0.01), type = "Non-linearity", name = "conventional", value = .) %>%
+  data.frame(d = seq(0, max_poly_ss, 0.01), type = "Non-linearity", name = "Conventional test", value = .) %>%
   mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
@@
     family = binomial()) %>%
-  predict(data.frame(actual_ss = seq(0, max_actual_ss, 0.01), offset0 = log(0.05/0.95)),
+  predict(data.frame(actual_ss = seq(0, max_heter_ss, 0.01), offset0 = log(0.05/0.95)),
           type = "response") %>%
-  data.frame(d = seq(0, max_actual_ss, 0.01), type = "Heteroskedasticity", name = "Computer vision model", value = .) %>%
+  data.frame(d = seq(0, max_heter_ss, 0.01), type = "Heteroskedasticity", name = "Computer vision model", value = .) %>%
   mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
@@
     family = binomial()) %>%
-  predict(data.frame(actual_ss = seq(0, max_actual_ss, 0.01), offset0 = log(0.05/0.95)),
+  predict(data.frame(actual_ss = seq(0, max_heter_ss, 0.01), offset0 = log(0.05/0.95)),
           type = "response") %>%
-  data.frame(d = seq(0, max_actual_ss, 0.01), type = "Heteroskedasticity", name = "Visual test", value = .) %>%
+  data.frame(d = seq(0, max_heter_ss, 0.01), type = "Heteroskedasticity", name = "Visual test", value = .) %>%
   mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
@@
     family = binomial()) %>%
-  predict(data.frame(actual_ss = seq(0, max_actual_ss, 0.01), offset0 = log(0.05/0.95)),
+  predict(data.frame(actual_ss = seq(0, max_heter_ss, 0.01), offset0 = log(0.05/0.95)),
           type = "response") %>%
-  data.frame(d = seq(0, max_actual_ss, 0.01), type = "Heteroskedasticity", name = "conventional", value = .) %>%
+  data.frame(d = seq(0, max_heter_ss, 0.01), type = "Heteroskedasticity", name = "Conventional test", value = .) %>%
   mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity")))
@@
   filter(!null_lineup, !attention_check) %>%
-  pivot_longer(c(model_reject, reject)) %>%
-  mutate(name = ifelse(name == "model_reject", "Computer vision model", "Visual test")) %>%
+  pivot_longer(c(model_reject, reject, conventional_reject)) %>%
+  mutate(name = ifelse(name == "model_reject", "Computer vision model", name)) %>%
+  mutate(name = ifelse(name == "reject", "Visual test", name)) %>%
+  mutate(name = ifelse(name == "conventional_reject", "Conventional test", name)) %>%
   ggplot() +
-  ggbeeswarm::geom_quasirandom(aes(actual_ss, value), size = 0.5) +
+  ggbeeswarm::geom_quasirandom(aes(actual_ss, value), size = 0.5, alpha = 0.4) +
   geom_line(data = model_glm_pred_poly, aes(d, value + 1, col = "Computer vision model"), linewidth = 0.8) +
@@
   geom_line(data = visual_glm_pred_heter, aes(d, value + 1, col = "Visual test"), linewidth = 0.8) +
+  geom_line(data = conv_glm_pred_poly, aes(d, value + 1, col = "Conventional test"), linewidth = 0.8) +
+  geom_line(data = conv_glm_pred_heter, aes(d, value + 1, col = "Conventional test"), linewidth = 0.8) +
   geom_line(data = mutate(model_glm_pred_poly, name = "Visual test"), aes(d, value + 1, col = "Computer vision model"), linetype = 2, linewidth = 0.8) +
   geom_line(data = mutate(model_glm_pred_heter, name = "Visual test"), aes(d, value + 1, col = "Computer vision model"), linetype = 2, linewidth = 0.8) +
+  geom_line(data = mutate(model_glm_pred_poly, name = "Conventional test"), aes(d, value + 1, col = "Computer vision model"), linetype = 2, linewidth = 0.8) +
+  geom_line(data = mutate(model_glm_pred_heter, name = "Conventional test"), aes(d, value + 1, col = "Computer vision model"), linetype = 2, linewidth = 0.8) +
   geom_line(data = mutate(visual_glm_pred_poly, name = "Computer vision model"), aes(d, value + 1, col = "Visual test"), linetype = 2, linewidth = 0.8) +
   geom_line(data = mutate(visual_glm_pred_heter, name = "Computer vision model"), aes(d, value + 1, col = "Visual test"), linetype = 2, linewidth = 0.8) +
-  facet_grid(name ~ type) +
+  geom_line(data = mutate(visual_glm_pred_poly, name = "Conventional test"), aes(d, value + 1, col = "Visual test"), linetype = 2, linewidth = 0.8) +
+  geom_line(data = mutate(visual_glm_pred_heter, name = "Conventional test"), aes(d, value + 1, col = "Visual test"), linetype = 2, linewidth = 0.8) +
+  geom_line(data = mutate(conv_glm_pred_poly, name = "Computer vision model"), aes(d, value + 1, col = "Conventional test"), linetype = 2, linewidth = 0.8) +
+  geom_line(data = mutate(conv_glm_pred_heter, name = "Computer vision model"), aes(d, value + 1, col = "Conventional test"), linetype = 2, linewidth = 0.8) +
+  geom_line(data = mutate(conv_glm_pred_poly, name = "Visual test"), aes(d, value + 1, col = "Conventional test"), linetype = 2, linewidth = 0.8) +
+  geom_line(data = mutate(conv_glm_pred_heter, name = "Visual test"), aes(d, value + 1, col = "Conventional test"), linetype = 2, linewidth = 0.8) +
+  facet_grid(name ~ type, scales = "free_x") +
   theme_light() +
@@
 
-### Boston Housing
-
-```{r fig.pos = "!h", fig.cap = "Residual plot of the regression model fitted on the Boston housing data."}
+In this section, we will utilize the trained computer vision model on three example datasets.
+These include the dataset associated with the residual plot displaying a "left-triangle" shape, as displayed in Figure \ref{fig:false-finding}, along with the Boston housing dataset, and the "dino" datasets from the `datasaurus` R package.
+Additionally, we will outline the workflow for employing this model with the `autovi` R package.
+
+### Left-triangle
+
+In Section \ref{introduction}, we presented an example residual plot showcased in Figure \ref{fig:false-finding}, illustrating how humans might misinterpret the "left-triangle" shape as indicative of heteroskedasticity.
+Additionally, the Breusch-Pagan test yielded a rejection with a $p$-value of 0.046, despite the residuals originating from a correctly specified model.
+Figure \ref{fig:false-lineup} offers a lineup for this fitted model, showcasing various degrees of "left-triangle" shape across all residual plots.
+This phenomenon is evidently caused by the skewed distribution of the fitted values.
+Notably, if the residual plot were evaluated through a visual test, it would not be rejected since the actual residual plot positioned at 10 can not be distinguished from the others.
+
+Moving forward, we will employ the trained computer vision model to assess the residual plot with the assistance of the `autovi` R package.
+This package is accessible on the GitHub repository `TengMCing/autovi`, and users can install it by executing the command `remotes::install_github("TengMCing/autovi")` in R.
+
+The results of the `list_keras_model()` function provide a list of available trained computer vision models.
+To perform the residual plot assessment, users need to supply a fitted linear regression model and select a trained Keras model to initialize the checker.
+
+```r
+checker <- auto_vi(fitted_mod = mod, 
+                   keras_mod = get_keras_model("vss_phn_32"))
+```
+
+When invoking the `check()` method to execute the assessment, users can specify the number of null plots and bootstrapped samples using the `null_draws` and `boot_draws` arguments, respectively.
+The diagnostic results can be viewed by directly printing the object.
+Alternatively, a summary plot of the assessment can be generated using the `summary_plot()` method.
+
+```r
+checker$check(null_draws = 200L, boot_draws = 200L)
+checker$summary_plot()
+```
+
+Figure \ref{fig:false-check} presents the results of the assessment.
+Notably, the observed visual signal strength is considerably lower than the 95% sample quantile of the null distribution.
+Moreover, the bootstrapped distribution suggests that it is highly improbable for the fitted model to be misspecified as the majority of bootstrapped fitted models will not be rejected.
+
+Thus, for this particular fitted model, both the visual test and the computer vision model will not reject $H_0$.
+However, the Breusch-Pagan test will reject $H_0$ because it can not effectively utilize information from null plots.
+
+```{r}
+set.seed(452)
+ori_x <- rand_lognormal()
+dat <- heter_model(b = 0, x = closed_form(~-ori_x))$gen(300)
+
+mod <- lm(y ~ x, data = dat)
+my_vi <- autovi::auto_vi(fitted_mod = mod)
+```
+
+```{r}
+if (!file.exists(here::here("paper/data/fig1_check.rds"))) {
+  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  my_vi$check(null_draws = 200L, boot_draws = 200L)
+  saveRDS(my_vi$check_result, here::here("paper/data/fig1_check.rds"))
+}
+
+```
+
+```{r}
+my_vi$check_result <- readRDS(here::here("paper/data/fig1_check.rds"))
+```
+
+```{r false-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
+On the left, the residual plot exhibiting "left-triangle" shape is displayed.
+The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots.
+The fitted model will not be rejected since $\\hat{D} < Q_{null}(0.95)$.'}
+
+p1 <- my_vi$plot_resid()
+p2 <- my_vi$summary_plot() +
+  ggtitle("", subtitle = glue::glue("Conventional p-value = {format(HETER_MODEL$test(dat)$p_value, digits = 3)}")) +
+  theme(legend.position = "bottom", legend.box = "vertical") +
+  labs(linetype = "") +
+  xlab(expression(hat(D))) +
+  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D))))
+
+patchwork::wrap_plots(p1, p2)
+```
+
+```{r false-lineup, fig.pos = "!h", fig.cap = 'A lineup of residual plots displaying "left-triangle" visual patterns.
+The actual residual plot occupies position 10, yet there are no discernible visual patterns that distinguish it from the other plots.'}
+pos <- 10
+result <- tibble()
+for (i in c(1:20)[-pos]) {
+  result <- bind_rows(result,
+                      my_vi$rotate_resid() %>%
+                        mutate(k = i))
+}
+result <- bind_rows(my_vi$get_fitted_and_resid() %>% 
+                      mutate(k = pos),
+                    result)
+result %>%
+  VI_MODEL$plot_lineup(remove_grid_line = TRUE,
+                       remove_axis = TRUE,
+                       theme = theme_light())
+```
+
+### Boston housing
+
+The Boston housing dataset, originally published by @harrison1978hedonic, offers insights into housing in the Boston, Massachusetts area.
+For illustration purposes, we will utilize a reduced version from Kaggle, comprising 489 rows and 4 columns: average number of rooms per dwelling (RM), percentage of lower status of the population (LSTAT), pupil-teacher ratio by town (PTRATIO), and Median value of owner-occupied homes in $1000's (MEDV).
+In our analysis, MEDV will serve as the response variable, while the other columns will function as predictors in a linear regression model.
+Our primary focus is to detect non-linearity, because the relationships between RM and MEDV or LSTAT and MEDV are non-linear.
+
+Figure \ref{fig:boston-check} displays the residual plot and the assessment conducted by the computer vision model.
+A clear non-linearity pattern resembling a "U" shape is shown in the residual plot.
+Furthermore, the RESET test yields a very small $p$-value.
+The approximated distance $\hat{D}$ significantly exceeds $Q_{null}(0.95)$, leading to rejection of $H_0$.
+The bootstrapped distribution also suggests that almost all the bootstrapped fitted models will be rejected, indicating that the fitted model is unlikely to be correctly specified.
+
+Additionally, the lineup of this fitted model presented in Figure \ref{fig:boston-lineup} underscores that the "U" shape is a distinctive characteristic of the actual residual plot, absent in any of the null plots.
+Consequently, if a visual test is conducted, $H_0$ will also be rejected.
+
+```{r}
 
@@
 
-```{r fig.pos = "!h"}
+```{r boston-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
+On the left, the residual plot of the Boston housing fitted model is displayed.
+The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots.
+The fitted model will be rejected since $\\hat{D} >= Q_{null}(0.95)$.'}
 p1 <- my_vi$plot_resid()
 p2 <- my_vi$summary_plot() +
-  ggtitle("", subtitle = "") +
+  ggtitle("", subtitle = glue::glue("Conventional p-value = {format(lmtest::resettest(mod)$p.value, digits = 3)}")) +
   theme(legend.position = "bottom", legend.box = "vertical") +
-  labs(linetype = "")
+  labs(linetype = "") +
+  xlab(expression(hat(D))) +
+  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D))))
 
@@
 
-### Insurance
-
-```{r fig.pos = "!h", fig.cap = "Residual plot of the regression model fitted on the Boston housing data."}
-insurance <- read_csv(here::here("paper/data/insurance.csv"))
-mod <- lm(charges ~ ., data = insurance)
-
-my_vi <- autovi::auto_vi(fitted_mod = mod)
-```
-
-```{r}
-if (!file.exists(here::here("paper/data/insurance_check.rds"))) {
-  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
-  my_vi$check(null_draws = 200L, boot_draws = 200L)
-  saveRDS(my_vi$check_result, here::here("paper/data/insurance_check.rds"))
+```{r boston-lineup, fig.pos = "!h", fig.cap = 'A lineup of residual plots for the Boston housing fitted model.
+The actual residual plot is at position 7.
+It can be easily identified as the most different plot.'}
+pos <- 7
+result <- tibble()
+for (i in c(1:20)[-pos]) {
+  result <- bind_rows(result,
+                      my_vi$rotate_resid() %>%
+                        mutate(k = i))
 }
-
-```
-
-```{r}
-my_vi$check_result <- readRDS(here::here("paper/data/insurance_check.rds"))
-```
-
-```{r fig.pos = "!h"}
-p1 <- my_vi$plot_resid()
-p2 <- my_vi$summary_plot() +
-  ggtitle("", subtitle = "") +
-  theme(legend.position = "bottom", legend.box = "vertical") +
-  labs(linetype = "")
-
-patchwork::wrap_plots(p1, p2)
-```
-
-### Wine Quality
-
-```{r}
-wine <- read_csv(here::here("paper/data/WineQT.csv"))
-mod <- lm(quality ~ ., data = select(wine, -Id))
-
-my_vi <- autovi::auto_vi(fitted_mod = mod)
-```
-
-```{r}
-if (!file.exists(here::here("paper/data/wine_check.rds"))) {
-  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
-  my_vi$check(null_draws = 200L, boot_draws = 200L)
-  saveRDS(my_vi$check_result, here::here("paper/data/wine_check.rds"))
-}
-```
-
-```{r}
-my_vi$check_result <- readRDS(here::here("paper/data/wine_check.rds"))
-```
-
-```{r fig.pos = "!h"}
-p1 <- my_vi$plot_resid()
-p2 <- my_vi$summary_plot() +
-  ggtitle("", subtitle = "") +
-  theme(legend.position = "bottom", legend.box = "vertical") +
-  labs(linetype = "")
-
-patchwork::wrap_plots(p1, p2)
+result <- bind_rows(my_vi$get_fitted_and_resid() %>% 
+                      mutate(k = pos),
+                    result)
+result %>%
+  VI_MODEL$plot_lineup(remove_grid_line = TRUE,
+                       remove_axis = TRUE,
+                       theme = theme_light())
 ```
@@
 ### Datasaurus
+
+The computer vision model possesses the capability to detect not only typical issues like non-linearity, heteroskedasticity, and non-normality but also artifact visual patterns resembling real-world objects, as long as they do not appear in null plots.
+These visual patterns can be challenging to categorize in terms of model violations.
+Therefore, we will employ the RESET test, the Breusch-Pagan test, and the Shapiro-Wilk test [@shapiro1965analysis] for comparison.
+
+The "dino" dataset within the `datasaurus` R package exemplifies this scenario.
+With only two columns, x and y, fitting a regression model to this data yields a residual plot resembling a "dinosaur," as displayed in Figure \ref{fig:dino-check}.
+Unsurprisingly, this distinct residual plot stands out in a lineup, as shown in Figure \ref{fig:dino-lineup}.
+A visual test conducted by humans would undoubtedly reject $H_0$.
+
+According to the residual plot assessment by the computer vision model, $\hat{D}$ exceeds $Q_{null}(0.95)$, warranting a rejection of $H_0$.
+Additionally, most of the bootstrapped fitted models will be rejected, indicating an misspecified model.
+However, both the RESET test and the Breusch-Pagan test yield $p$-values greater than 0.3, leading to a non-rejection of $H_0$.
+Only the Shapiro-Wilk test rejects the normality assumption with a small $p$-value.
+
+In practice, without accessing the residual plot, it would be challenging to identify the artificial pattern of the residuals.
+Moreover, conducting a normality test for a fitted regression model is not always standard practice among analysts.
+Even when performed, violating the normality assumption is sometimes deemed acceptable, especially considering the application of quasi-maximum likelihood estimation in linear regression.
+This example underscores the importance of evaluating residual plots and highlights how the proposed computer vision model can facilitate this process.
 
@@
 
-```{r fig.pos = "!h"}
+```{r dino-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
+On the left, the residual plot exhitbiting a "dinosaur" shape is displayed.
+The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots.
+The fitted model will be rejected since $\\hat{D} >= Q_{null}(0.95)$.'}
 p1 <- my_vi$plot_resid()
 p2 <- my_vi$summary_plot() +
-  ggtitle("", subtitle = "") +
+  ggtitle("", subtitle = glue::glue("RESET test p-value = {format(lmtest::resettest(mod)$p.value, digits = 3)}\n Breusch-Pagan test p-value = {format(lmtest::bptest(mod)$p.value, digits = 3)} \n Shapiro-Wilk test p-value = {format(shapiro.test(mod$residuals)$p.value, digits = 3)}")) +
   theme(legend.position = "bottom", legend.box = "vertical") +
-  labs(linetype = "")
+  labs(linetype = "") +
+  xlab(expression(hat(D))) +
+  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D))))
 
 patchwork::wrap_plots(p1, p2)
+```
+
+```{r dino-lineup, fig.pos = "!h", fig.cap = 'A lineup of residual plots for the fitted model on the "dinosaur" dataset.
+The actual residual plot is at position 17.
+It can be easily identified as the most different plot as the visual pattern is extremly artificial.'}
+pos <- 17
+result <- tibble()
+for (i in c(1:20)[-pos]) {
+  result <- bind_rows(result,
+                      my_vi$rotate_resid() %>%
+                        mutate(k = i))
+}
+result <- bind_rows(my_vi$get_fitted_and_resid() %>% 
+                      mutate(k = pos),
+                    result)
+result %>%
+  VI_MODEL$plot_lineup(remove_grid_line = TRUE,
+                       remove_axis = TRUE,
+                       theme = theme_light())
 ```
@@
 
-<!-- # Dicussion -->
+<!-- # Discussion -->
 
@@
 
+- Brief discussion
 - Summary of findings
 - Contributions to the field
-- Future directions for research
-
-# Acknowledgements
+
+In this study, we have introduced a distance metric based on Kullback-Leibler divergence to quantify the disparity between the residual distribution of a fitted linear regression model and the reference residual distribution assumed under correct model specification.
+This distance metric effectively captures the magnitude of model violations in misspecified models.
+We propose a computer vision model to approximate this distance, utilizing the residual plot of the fitted model as input.
+The resulting approximated distance serves as the foundation for constructing a single Model Violation Index (MVI), facilitating the quantification of various model violations.
+
+Moreover, the approximated distance enables the development of a formal statistical testing procedure by evaluating a large number of null plots generated from the fitted model.
+Additionally, employing bootstrapping techniques and refitting the regression model allows us to ascertain how frequently the fitted model is considered misspecified if data were repeatedly obtained from the same data generating process.
+
+The trained computer vision model demonstrates strong performance on both the training and test sets, although it exhibits slightly lower performance on residual plots with non-linearity visual patterns compared to other types of violations.
+The statistical tests relying on the approximated distance predicted by the computer vision model exhibit lower sensitivity compared to conventional tests but higher sensitivity compared to visual tests conducted by humans.
+While the approximated distance generally mirrors the strength of the visual signal perceived by humans, there remains scope for further improvement in its performance.
+
+Several examples are provided to showcase the effectiveness of the proposed method across different scenarios, emphasizing the similarity between visual tests and distance-based tests.
+Overall, both visual tests and distance-based tests can be viewed as combined tests, aiming to assess the correctness of all model assumptions collectively.
+In contrast, individual residual diagnostic tests such as the RESET test and the Breusch-Pagan test only evaluate specific model assumptions.
+
+In practice, selecting an appropriate set of statistical tests for regression diagnostics can be challenging, particularly given the necessity of adjusting the significance level for each test.
+
+Our method holds significant value as it helps alleviate a portion of analysts' workload associated with assessing residual plots.
+While we recommend analysts to continue reading residual plots whenever feasible, as they offer invaluable insights, our approach serves as a valuable tool for automating the diagnostic process or for supplementary purposes when needed.
+
+- Future directions
+
+# Acknowledgements {-}
+
+- Packages used
+- Website link

### 2024-02-15 paper/paper.tex
update paper

@@
 \newtheorem{notation}{Notation}
+
+% Pandoc syntax highlighting
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 
@@
 
-\articletype{}
+\articletype{Draft paper (Incomplete)}
 
@@
 \end{keywords}
-
-\newpage
-\tableofcontents
-\newpage
 
@@
 images. This characteristic makes it less suitable for large-scale
-applications, given the associated high labor costs and time
+applications, given the associated high labour costs and time
 requirements.
@@
 with early efforts such as ``Scagnostics'' by \citet{tukey1985computer},
-focusing on scatterplot diagnostics. \citet{wilkinson2005graph} expanded
-on this work, introducing graph theoretic scagnostics, which encompassed
-computable measures applied to planar proximity graphs. These measures,
-including, but not limited to, ``Outlying'', ``Skinny'', ``Stringy'',
-``Straight'', ``Monotonic'', ``Skewed'', ``Clumpy'', and ``Striated''
-aimed to characterize outliers, shape, density, trend, coherence and
-other characteristics of the data. While this approach has been
-inspiring, there is a recognition \citep{buja2009statistical} that it
-may not capture all the necessary visual features that differentiate
+focusing on scatter plot diagnostics. \citet{wilkinson2005graph}
+expanded on this work, introducing graph theoretic scagnostics, which
+encompassed computable measures applied to planar proximity graphs.
+These measures, including, but not limited to, ``Outlying'', ``Skinny'',
+``Stringy'', ``Straight'', ``Monotonic'', ``Skewed'', ``Clumpy'', and
+``Striated'' aimed to characterize outliers, shape, density, trend,
+coherence and other characteristics of the data. While this approach has
+been inspiring, there is a recognition \citep{buja2009statistical} that
+it may not capture all the necessary visual features that differentiate
 actual residual plots from null plots. A more promising alternative
@@
 The vertical spread of the data points varies with the fitted values.
-This often indicates the existence of heteroskedasticity.}\label{fig:false-finding}
+This often indicates the existence of heteroskedasticity.
+The Breusch-Pagan test rejects this residual plot at 95\% significance level ($p\text{-value} = 0.046$).}\label{fig:false-finding}
 \end{figure}
@@
 implies that the model can not accept arbitrary images as input but
-requires the use of the same preprocessing pipeline to convert residuals
-and fitted values into a standardized-style residual plot.
+requires the use of the same pre-processing pipeline to convert
+residuals and fitted values into a standardized-style residual plot.
 
@@
 \(\boldsymbol{\varepsilon}\) follows a non-normal distribution such as a
-multivariate lognormal distribution, \(P\) will usually be skewed and
+multivariate log-normal distribution, \(P\) will usually be skewed and
 has a long tail.
@@
 measure with a residual plot. Let \(D\) be the result of equation
-\ref{eq:kl-0} indicating the extent of the model violations, and our
-estimator \(\hat{D}\) is formulated as
+\ref{eq:kl-0}, and our estimator \(\hat{D}\) is formulated as
 
@@
 different the underlying distribution of the residuals is from a good
-residual distribution. It also provides information for the strength of
-the visual signal embedded in the residual plot.
+residual distribution. The approximated distance \(\hat{D}\) can also be
+used as an index of the model violations. It also provides information
+for the strength of the visual signal embedded in the residual plot.
 
@@
 residual plot is representative or not.
+
+\hypertarget{approximated-distance-as-an-model-violations-index-mvi}{%
+\subsubsection{Approximated distance as an Model Violations Index
+(MVI)}\label{approximated-distance-as-an-model-violations-index-mvi}}
+
+In Section \ref{output-formats}, we discuss the output formats of the
+computer vision model presented in this paper. One notable advantage
+highlighted is its interpretability. The approximated distance
+\(\hat{D}\) is an estimator of the difference between the actual
+residual distribution and the reference residual distribution. This
+difference primarily arises from deviations in model assumptions. The
+magnitude of \(D\) directly reflects the degree of these deviations,
+thus making \(\hat{D}\) instrumental in forming a model violation index.
+
+Note that since the Kullback-Leibler divergence could be affected by the
+number of random variables involved in the evaluation. In general, more
+observations will result in greater distance \(D\). This does not mean
+\(\hat{D}\) can not correctly represent the extent of the model
+violations since more observations in a residual plot usually lead to
+stronger visual signal strength as the underlying visual patterns will
+have higher chance to be revealed except for cases like serious
+overlapping. For the same data generating process, \(D\) usually grows
+logarithmically as the number of observations increases. This is because
+\(D = \text{log}(1 + D_{KL})\), and
+\(D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}\) under the independence
+assumption.
+
+Note that the Kullback-Leibler divergence might be influenced by the
+number of random variables involved in its evaluation. Generally, a
+larger number of observations will lead to a greater distance \(D\).
+However, this does not imply that \(\hat{D}\) fails to accurately
+represent the extent of model violations. In fact, when examining
+residual plots with more observations, we often observe a stronger
+visual signal strength, as the underlying patterns are more likely to be
+revealed, except in cases of significant overlapping.
+
+For a consistent data generating process, \(D\) typically increases
+logarithmically with the number of observations. This behaviour comes
+from the relationship \(D = \text{log}(1 + D_{KL})\), where
+\(D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}\) under the assumption of
+independence.
+
+Therefore, the Model Violations Index (MVI) can be proposed as
+
+\begin{equation}
+\label{eq:mvi}
+\text{MVI} = C + \hat{D} - log(n),
+\end{equation}
+
+\noindent where \(C\) is a large enough constant keeping the result
+positive.
 
@@
 distinct error distributions are considered, including discrete,
-uniform, normal, and lognormal distributions, as presented in Figure
+uniform, normal, and log-normal distributions, as presented in Figure
 \ref{fig:different-e}. Each distribution imparts unique characteristics
 to the residuals. The discrete error distribution introduces
-discreteness in residuals, while the lognormal distribution typically
+discreteness in residuals, while the log-normal distribution typically
 yields outliers. Uniform error distribution may result in residuals
@@
 ``Monotonic,'' ``Sparse,'' ``Splines,'' and ``Striped'' -- using the
-cassowaryr R package \citep{mason2022cassowaryr}. These computed
-measures, along with the number of observations from the fitted model,
-were provided as the second input for the computer vision model.
+\texttt{cassowaryr} R package \citep{mason2022cassowaryr}. These
+computed measures, along with the number of observations from the fitted
+model, were provided as the second input for the computer vision model.
 Although other scagnostics are informative, they are currently
-unavailable due to a fatal bug in the compiled C program of the interp R
-package \citep{Albrecht2023interp}, which may unpredictably crash the
-process. For reproducibility, we excluded these scagnostics from the
-training data.
+unavailable due to a fatal bug in the compiled C program of the
+\texttt{interp} R package \citep{Albrecht2023interp}, which may
+unpredictably crash the process. For reproducibility, we excluded these
+scagnostics from the training data.
 
@@
 This is followed by a grayscale conversion layer utilizing the luma
-formula under the Rec. 601 standard, which converts the color image to
+formula under the Rec. 601 standard, which converts the colour image to
 grayscale. Grayscale suffices for our task since data points are plotted
@@
 learning rate often need to be fine-tuned using a tuner. In our study,
-we utilized the Bayesian optimization tuner from the KerasTuner Python
-library \citep{omalley2019kerastuner} for this purpose. A comprehensive
-list of hyperparameters is provided in Table \ref{tab:hyperparameter}.
+we utilized the Bayesian optimization tuner from the \texttt{KerasTuner}
+Python library \citep{omalley2019kerastuner} for this purpose. A
+comprehensive list of hyperparameters is provided in Table
+\ref{tab:hyperparameter}.
 
@@
 
-\begin{itemize}
-\item
-  RMSE for the test set
-\item
-  R\^{}2
-\item
-  Mean bias deviation to understand the overall bias
-\item
-  quantile loss to understand how well the model captures the entire
-  distribution of D
-\item
-  Percentage of predictions within a tolerance interval (like 0.1)
-\item
-  comparison with human visual inference
-\item
-  overview of the human subject experiment
-\item
-  metrics
-\item
-  agreement rate
-\item
-  check cases that are aggreed
-\item
-  check cases that are disaggreed
-\end{itemize}
+This section remains incomplete as we may modify the ``Results'' section
+later on.
 
@@
 \(3 < \hat{D} < 6\). To comprehend the reasons behind these deviations,
-we conducted a meta-analysis on \(D - \hat{D}\) for all factors used in
-the data generating process. A regression tree was constructed using the
-rpart R package \citep{Terry2022rpart}. The results of the regression
-tree are provided in the Appendix.
-
-The meta-analysis revealed that most issues arose from non-linearity
+we fitted a regression tree using the \texttt{rpart} R package
+\citep{Terry2022rpart} on \(D - \hat{D}\) for all factors used in the
+data generating process. The results of the regression tree are provided
+in the Appendix.
+
+The regression tree revealed that most issues arose from non-linearity
 problems and the presence of a second predictor in the regression model.
@@
 
-\hypertarget{comparison-with-human-visual-inference}{%
-\subsection{Comparison with human visual
-inference}\label{comparison-with-human-visual-inference}}
+\hypertarget{model-violations-index-mvi}{%
+\subsection{Model Violations Index
+(MVI)}\label{model-violations-index-mvi}}
+
+Figures \ref{fig:poly-index} and \ref{fig:heter-index} display the
+residual plots for fitted models exhibiting varying degrees of
+non-linearity and heteroskedasticity. Each residual plot's Model
+Violation Index (MVI) is computed using Equation \ref{eq:mvi} with
+\(C = 10\). When \(\text{MVI} > 8\), the visual patterns are notably
+strong and easily discernible by humans. In the range
+\(6 < \text{MVI} < 8\), the visibility of the visual pattern diminishes
+as MVI decreases. Conversely, when \(\text{MVI} < 6\), the visual
+pattern tends to become relatively faint and challenging to observe.
+Table \ref{tab:mvi} provides a summary of the MVI usage and it is
+applicable to other linear regression models.
+
+\begin{table}
+
+\caption{\label{tab:mvi}Degree of model violations or the strength of the visual signals according to the Model Violation Index (MVI). The constant $C$ is set to be 10.}
+\centering
+\begin{tabular}[t]{ll}
+\toprule
+Degree of model violations & Range ($C$ = 10)\\
+\midrule
+Strong & $\text{MVI} > 8$\\
+Moderate & $6 < \text{MVI} < 8$\\
+Weak & $\text{MVI} < 6$\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/poly-index-1}
+
+}
+
+\caption{Residual plots generated from fitted models exhibiting varying degrees of non-linearity violations.
+The Model violations index (MVI) is displayed atop each residual plot.
+The non-linearity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$.}\label{fig:poly-index}
+\end{figure}
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heter-index-1}
+
+}
+
+\caption{Residual plots generated from fitted models exhibiting varying degrees of heteroskedasticity violations.
+The Model violations index (MVI) is displayed atop each residual plot.
+The heteroskedasticity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$.}\label{fig:heter-index}
+\end{figure}
+
+\hypertarget{comparison-with-human-visual-inference-and-conventional-tests}{%
+\subsection{Comparison with human visual inference and conventional
+tests}\label{comparison-with-human-visual-inference-and-conventional-tests}}
 
@@
 vision model to predict \(\hat{D}\) for both the actual residual plots
-and the null plots.
+and the null plots. Additionally, the appropriate conventional tests
+including the Ramsey Regression Equation Specification Error Test
+(RESET) \citep{ramsey1969tests} for non-linearity and the Breusch-Pagan
+test \citep{breusch1979simple} for heteroskedasticity were applied on
+the same data for comparison.
 
@@
 
-Figure \ref{fig:human-mosaic} and Table \ref{tab:human-table} provide
-insights into the agreement between decisions made by the computer
-vision model and visual tests conducted by humans. It can be observed
-that visual tests are less sensitive than the computer vision model, as
-they reject almost only when the computer vision model does. The
-agreement rates between visual tests and the computer vision model are
-67.96\% and 66.84\% for residual plots containing heteroskedasticity and
-non-linearity patterns, respectively.
-
-When plotting the decision against the distance, as depicted in Figure
-\ref{fig:power}, we observe that for the computer vision model, when the
-distance \(D > 3\), almost all residual plots will be rejected. However,
-for visual tests conducted by humans, the threshold is higher at
-\(D > 5\). Visual tests also demonstrate low sensitivity to residual
-plots with small distances compared to the computer vision model.
+Likewise, Figure \ref{fig:conv-mosaic} and Table \ref{tab:conv-table}
+provide a summary of the agreement between decisions made by the
+computer vision model and conventional tests. The agreement rates
+between conventional tests and the computer vision model are 65.37\% and
+57.29\% for residual plots containing heteroskedasticity and
+non-linearity patterns, respectively. These figures are lower than those
+presented in Table \ref{tab:human-table}, indicating that the computer
+vision model exhibits behaviour more akin to visual tests than
+conventional tests. Additionally, conventional tests tend to reject
+residual plots almost exclusively when the computer vision model does,
+but they also reject plots when the computer vision model does not. This
+suggests that conventional tests are more sensitive than the computer
+vision model.
+
+When plotting the decision against the distance, as illustrated in
+Figure \ref{fig:power}, several observations emerge. Firstly, for the
+computer vision model, when the distance \(D > 3\), almost all residual
+plots are rejected. However, for visual tests conducted by humans, the
+threshold is higher, set at \(D > 5\). And conventional tests exhibit
+some non-rejection cases for \(D > 3\). Moreover, the computer vision
+model tends to have fewer rejected cases than conventional tests when
+\(D < 2\). However, visual tests demonstrate the lowest sensitivity to
+residual plots with small distances. Therefore, tests based on the
+computer vision model are less sensitive to small deviations from model
+assumptions than conventional tests but more sensitive to moderate
+deviations. Rejection decisions are fitted by logistic regression models
+with no intercept terms but an offset equals to
+\(\text{log}(0.05/0/95)\). The fitting curves for the computer vision
+model fall between those of conventional tests and visual tests for
+non-linearity and are slightly below conventional tests for
+heteroskedasticity. There is still potential to refine the computer
+vision model to better align its behaviour with visual tests.
 
@@
 distance measure, or the approximated distance, may not perfectly
-reflect actual human behavior.
+reflect actual human behaviour.
 
@@
 
-\caption{\label{tab:human-table}Summary of decisions made by the visual test conducted by human and decisions made by the computer vision model.}
+\caption{\label{tab:human-table}Summary of decisions made by visual tests conducted by human and decisions made by the computer vision model.}
 \centering
@@
 
+\begin{table}
+
+\caption{\label{tab:conv-table}Summary of decisions made by conventional tests and decisions made by the computer vision model.}
+\centering
+\begin{tabular}[t]{llll}
+\toprule
+Violations & \#Samples & \#Agreements & Agreement rate\\
+\midrule
+heteroskedasticity & 540 & 464 & 0.6537\\
+non-linearity & 576 & 459 & 0.5729\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \begin{figure}[!h]
@@
 
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/conv-mosaic-1}
+
+}
+
+\caption{Rejection rate ($p$-value $\leq0.05$) of conventional test conditional on the computer vision model decision on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot.
+The conventional test rejects more frequently than the computer vision model, and (almost) only rejects when the computer vision model does.}\label{fig:conv-mosaic}
+\end{figure}
+
+\begin{figure}[!h]
+
 {\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/power-1}
@@
 
-\caption{Comparison of decisions made by visual tests and the computer vision model.
-The data points are fitted with logistic regression models with no intercept but an offset equals to $log(0.05/0.95)$.}\label{fig:power}
+\caption{Comparison of decisions made by visual tests, conventional tests and the computer vision model.
+The data points are fitted with logistic regression models with no intercept but an offset equals to $\text{log}(0.05/0.95)$.}\label{fig:power}
 \end{figure}
@@
 
+In this section, we will utilize the trained computer vision model on
+three example datasets. These include the dataset associated with the
+residual plot displaying a ``left-triangle'' shape, as displayed in
+Figure \ref{fig:false-finding}, along with the Boston housing dataset,
+and the ``dino'' datasets from the \texttt{datasaurus} R package.
+Additionally, we will outline the workflow for employing this model with
+the \texttt{autovi} R package.
+
+\hypertarget{left-triangle}{%
+\subsubsection{Left-triangle}\label{left-triangle}}
+
+In Section \ref{introduction}, we presented an example residual plot
+showcased in Figure \ref{fig:false-finding}, illustrating how humans
+might misinterpret the ``left-triangle'' shape as indicative of
+heteroskedasticity. Additionally, the Breusch-Pagan test yielded a
+rejection with a \(p\)-value of 0.046, despite the residuals originating
+from a correctly specified model. Figure \ref{fig:false-lineup} offers a
+lineup for this fitted model, showcasing various degrees of
+``left-triangle'' shape across all residual plots. This phenomenon is
+evidently caused by the skewed distribution of the fitted values.
+Notably, if the residual plot were evaluated through a visual test, it
+would not be rejected since the actual residual plot positioned at 10
+can not be distinguished from the others.
+
+Moving forward, we will employ the trained computer vision model to
+assess the residual plot with the assistance of the \texttt{autovi} R
+package. This package is accessible on the GitHub repository
+\texttt{TengMCing/autovi}, and users can install it by executing the
+command \texttt{remotes::install\_github("TengMCing/autovi")} in R.
+
+The results of the \texttt{list\_keras\_model()} function provide a list
+of available trained computer vision models. To perform the residual
+plot assessment, users need to supply a fitted linear regression model
+and select a trained Keras model to initialize the checker.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{checker }\OtherTok{\textless{}{-}} \FunctionTok{auto\_vi}\NormalTok{(}\AttributeTok{fitted\_mod =}\NormalTok{ mod, }
+                   \AttributeTok{keras\_mod =} \FunctionTok{get\_keras\_model}\NormalTok{(}\StringTok{"vss\_phn\_32"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+When invoking the \texttt{check()} method to execute the assessment,
+users can specify the number of null plots and bootstrapped samples
+using the \texttt{null\_draws} and \texttt{boot\_draws} arguments,
+respectively. The diagnostic results can be viewed by directly printing
+the object. Alternatively, a summary plot of the assessment can be
+generated using the \texttt{summary\_plot()} method.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{check}\NormalTok{(}\AttributeTok{null\_draws =}\NormalTok{ 200L, }\AttributeTok{boot\_draws =}\NormalTok{ 200L)}
+\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{summary\_plot}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+Figure \ref{fig:false-check} presents the results of the assessment.
+Notably, the observed visual signal strength is considerably lower than
+the 95\% sample quantile of the null distribution. Moreover, the
+bootstrapped distribution suggests that it is highly improbable for the
+fitted model to be misspecified as the majority of bootstrapped fitted
+models will not be rejected.
+
+Thus, for this particular fitted model, both the visual test and the
+computer vision model will not reject \(H_0\). However, the
+Breusch-Pagan test will reject \(H_0\) because it can not effectively
+utilize information from null plots.
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/false-check-1}
+
+}
+
+\caption{A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
+On the left, the residual plot exhibiting "left-triangle" shape is displayed.
+The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots.
+The fitted model will not be rejected since $\hat{D} < Q_{null}(0.95)$.}\label{fig:false-check}
+\end{figure}
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/false-lineup-1}
+
+}
+
+\caption{A lineup of residual plots displaying "left-triangle" visual patterns.
+The actual residual plot occupies position 10, yet there are no discernible visual patterns that distinguish it from the other plots.}\label{fig:false-lineup}
+\end{figure}
+
 \hypertarget{boston-housing}{%
-\subsubsection{Boston Housing}\label{boston-housing}}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-8-1} \end{center}
-
-\hypertarget{advertising}{%
-\subsubsection{Advertising}\label{advertising}}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-12-1} \end{center}
-
-\hypertarget{wine-quality}{%
-\subsubsection{Wine Quality}\label{wine-quality}}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-16-1} \end{center}
+\subsubsection{Boston housing}\label{boston-housing}}
+
+The Boston housing dataset, originally published by
+\citet{harrison1978hedonic}, offers insights into housing in the Boston,
+Massachusetts area. For illustration purposes, we will utilize a reduced
+version from Kaggle, comprising 489 rows and 4 columns: average number
+of rooms per dwelling (RM), percentage of lower status of the population
+(LSTAT), pupil-teacher ratio by town (PTRATIO), and Median value of
+owner-occupied homes in \$1000's (MEDV). In our analysis, MEDV will
+serve as the response variable, while the other columns will function as
+predictors in a linear regression model. Our primary focus is to detect
+non-linearity, because the relationships between RM and MEDV or LSTAT
+and MEDV are non-linear.
+
+Figure \ref{fig:boston-check} displays the residual plot and the
+assessment conducted by the computer vision model. A clear non-linearity
+pattern resembling a ``U'' shape is shown in the residual plot.
+Furthermore, the RESET test yields a very small \(p\)-value. The
+approximated distance \(\hat{D}\) significantly exceeds
+\(Q_{null}(0.95)\), leading to rejection of \(H_0\). The bootstrapped
+distribution also suggests that almost all the bootstrapped fitted
+models will be rejected, indicating that the fitted model is unlikely to
+be correctly specified.
+
+Additionally, the lineup of this fitted model presented in Figure
+\ref{fig:boston-lineup} underscores that the ``U'' shape is a
+distinctive characteristic of the actual residual plot, absent in any of
+the null plots. Consequently, if a visual test is conducted, \(H_0\)
+will also be rejected.
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/boston-check-1}
+
+}
+
+\caption{A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
+On the left, the residual plot of the Boston housing fitted model is displayed.
+The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots.
+The fitted model will be rejected since $\hat{D} >= Q_{null}(0.95)$.}\label{fig:boston-check}
+\end{figure}
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/boston-lineup-1}
+
+}
+
+\caption{A lineup of residual plots for the Boston housing fitted model. The actual residual plot is at position 7. It can be easily identified as the most different plot.}\label{fig:boston-lineup}
+\end{figure}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-20-1} \end{center}
+The computer vision model possesses the capability to detect not only
+typical issues like non-linearity, heteroskedasticity, and non-normality
+but also artifact visual patterns resembling real-world objects, as long
+as they do not appear in null plots. These visual patterns can be
+challenging to categorize in terms of model violations. Therefore, we
+will employ the RESET test, the Breusch-Pagan test, and the Shapiro-Wilk
+test \citep{shapiro1965analysis} for comparison.
+
+The ``dino'' dataset within the \texttt{datasaurus} R package
+exemplifies this scenario. With only two columns, x and y, fitting a
+regression model to this data yields a residual plot resembling a
+``dinosaur,'' as displayed in Figure \ref{fig:dino-check}.
+Unsurprisingly, this distinct residual plot stands out in a lineup, as
+shown in Figure \ref{fig:dino-lineup}. A visual test conducted by humans
+would undoubtedly reject \(H_0\).
+
+According to the residual plot assessment by the computer vision model,
+\(\hat{D}\) exceeds \(Q_{null}(0.95)\), warranting a rejection of
+\(H_0\). Additionally, most of the bootstrapped fitted models will be
+rejected, indicating an misspecified model. However, both the RESET test
+and the Breusch-Pagan test yield \(p\)-values greater than 0.3, leading
+to a non-rejection of \(H_0\). Only the Shapiro-Wilk test rejects the
+normality assumption with a small \(p\)-value.
+
+In practice, without accessing the residual plot, it would be
+challenging to identify the artificial pattern of the residuals.
+Moreover, conducting a normality test for a fitted regression model is
+not always standard practice among analysts. Even when performed,
+violating the normality assumption is sometimes deemed acceptable,
+especially considering the application of quasi-maximum likelihood
+estimation in linear regression. This example underscores the importance
+of evaluating residual plots and highlights how the proposed computer
+vision model can facilitate this process.
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/dino-check-1}
+
+}
+
+\caption{A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
+On the left, the residual plot exhitbiting a "dinosaur" shape is displayed.
+The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots.
+The fitted model will be rejected since $\hat{D} >= Q_{null}(0.95)$.}\label{fig:dino-check}
+\end{figure}
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/dino-lineup-1}
+
+}
+
+\caption{A lineup of residual plots for the fitted model on the "dinosaur" dataset.
+The actual residual plot is at position 17.
+It can be easily identified as the most different plot as the visual pattern is extremly artificial.}\label{fig:dino-lineup}
+\end{figure}
 
@@
 \tightlist
+\item
+  Brief discussion
 \item
@@
   Contributions to the field
+\end{itemize}
+
+In this study, we have introduced a distance metric based on
+Kullback-Leibler divergence to quantify the disparity between the
+residual distribution of a fitted linear regression model and the
+reference residual distribution assumed under correct model
+specification. This distance metric effectively captures the magnitude
+of model violations in misspecified models. We propose a computer vision
+model to approximate this distance, utilizing the residual plot of the
+fitted model as input. The resulting approximated distance serves as the
+foundation for constructing a single Model Violation Index (MVI),
+facilitating the quantification of various model violations.
+
+Moreover, the approximated distance enables the development of a formal
+statistical testing procedure by evaluating a large number of null plots
+generated from the fitted model. Additionally, employing bootstrapping
+techniques and refitting the regression model allows us to ascertain how
+frequently the fitted model is considered misspecified if data were
+repeatedly obtained from the same data generating process.
+
+The trained computer vision model demonstrates strong performance on
+both the training and test sets, although it exhibits slightly lower
+performance on residual plots with non-linearity visual patterns
+compared to other types of violations. The statistical tests relying on
+the approximated distance predicted by the computer vision model exhibit
+lower sensitivity compared to conventional tests but higher sensitivity
+compared to visual tests conducted by humans. While the approximated
+distance generally mirrors the strength of the visual signal perceived
+by humans, there remains scope for further improvement in its
+performance.
+
+Several examples are provided to showcase the effectiveness of the
+proposed method across different scenarios, emphasizing the similarity
+between visual tests and distance-based tests. Overall, both visual
+tests and distance-based tests can be viewed as combined tests, aiming
+to assess the correctness of all model assumptions collectively. In
+contrast, individual residual diagnostic tests such as the RESET test
+and the Breusch-Pagan test only evaluate specific model assumptions.
+
+In practice, selecting an appropriate set of statistical tests for
+regression diagnostics can be challenging, particularly given the
+necessity of adjusting the significance level for each test.
+
+Our method holds significant value as it helps alleviate a portion of
+analysts' workload associated with assessing residual plots. While we
+recommend analysts to continue reading residual plots whenever feasible,
+as they offer invaluable insights, our approach serves as a valuable
+tool for automating the diagnostic process or for supplementary purposes
+when needed.
+
+\begin{itemize}
+\tightlist
 \item
-  Future directions for research
+  Future directions
+\end{itemize}
+
+\hypertarget{acknowledgements}{%
+\section*{Acknowledgements}\label{acknowledgements}}
+\addcontentsline{toc}{section}{Acknowledgements}
+
+\begin{itemize}
+\tightlist
+\item
+  Packages used
+\item
+  Website link
 \end{itemize}

### 2024-04-15 paper/appendix.Rmd
update paper

--- 2024-02-12 paper/paper.Rmd
@@
 title: |
-  Automated assessment of residual plots with computer vision models
+  Appendix: Automated assessment of residual plots with computer vision models
+type: Draft paper (Incomplete)
 author:
@@
 
-```{r}
-# Visual inference models and p-value calculation
-# remotes::install_github("TengMCing/visage)
+# Additional details about data generation
 
-library(tidyverse)
-library(visage)
-library(glue)
-
-# To control the simulation in this file
-set.seed(10086)
-```
-
-\newpage
-\tableofcontents
-\newpage
-
-# Introduction
-
-Plotting residuals is commonly regarded as a standard practice in linear regression diagnostics \citep[see][]{cook1982residuals, belsley1980regression}.
-This visual assessment plays a crucial role in identifying deviations from model assumptions, such as linearity, homoscedasticity, and normality.
-It also helps in understanding the goodness of fit and various characteristics of the model.
-
-Generating a residual plot in most statistical software is often as straightforward as executing a line of code or clicking a button.
-However, accurately interpreting a residual plot can be challenging.
-Consider Figure \ref{fig:false-finding} as an example, the residuals display a triangular shape pointing to the left.
-While this might suggest heteroskedasticity, it is important to avoid over-interpreting this visual pattern.
-In this case, the fitted regression model is correctly specified, and the triangular shape is actually a result of the skewed distribution of the predictors, rather than indicating a flaw in the model.
-
-A residual plot can exhibit various visual features, but it is crucial to recognize that some may arise from the characteristics of predictors and the inherent randomness of the error, rather than indicating a violation of model assumptions [@li2023plot].
-The concept of visual inference, as proposed by @buja2009statistical, provides an inferential framework to assess whether residual plots indeed contain visual patterns inconsistent with the model assumptions.
-The fundamental idea involves testing whether the actual residual plot visually differs significantly from null plots, where null plots are plotted with residuals generated from the residual rotation distribution [@langsrud2005rotation], which is a distribution consistent with the null hypothesis $H_0$ that the linear regression model is correctly specified.
-Typically, the visual test is accomplished through the lineup protocol, where the real residual plot is embedded within a lineup alongside several null plots.
-If the real residual plot can be distinguished from the lineup, it provides evidence for rejecting $H_0$.
-
-The practice of delivering a residual plot as a lineup is generally regarded as a valuable approach.
-Beyond its application in residual diagnostics, the lineup protocol has integrated into the analysis of diverse subjects.
-For instance, \cite{loy2013diagnostic, loy2014hlmdiag, loy2015you} illustrated its applicability in diagnosing hierarchical linear models.
-Additionally, @widen2016graphical demonstrated its utility in geographical research, while @krishnan2021hierarchical explored its effectiveness in forensic examinations.
-
-However, as pointed out by @li2023plot, a primary limitation of the lineup protocol lies in its reliance on human judgments.
-Unlike conventional statistical tests that can be performed computationally in statistical software, the lineup protocol requires human evaluation of images.
-This characteristic makes it less suitable for large-scale applications, given the associated high labor costs and time requirements.
-
-There is a substantial need to develop an approach that alleviates analysts' workload by automating repetitive tasks and providing standardized results in a controlled environment.
-The large-scale evaluation of lineups is impractical without the use of technology and machines.
-
-The utilization of computers to interpret data plots has a rich history, with early efforts such as "Scagnostics" by @tukey1985computer, focusing on scatterplot diagnostics.
-@wilkinson2005graph expanded on this work, introducing graph theoretic scagnostics, which encompassed computable measures applied to planar proximity graphs.
-These measures, including, but not limited to, "Outlying", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy", and "Striated" aimed to characterize outliers, shape, density, trend, coherence and other characteristics of the data.
-While this approach has been inspiring, there is a recognition [@buja2009statistical] that it may not capture all the necessary visual features that differentiate actual residual plots from null plots.
-A more promising alternative entails enabling machines to learn the function for extracting visual features from residual plots.
-Essentially, this means empowering computers to discern the crucial visual features for residual diagnostics and determining the method to extract them.
-
-Modern computer vision models are well-suited for addressing this challenge.
-They rely on deep neural networks with convolutional layers [@fukushima1982neocognitron].
-These layers leverage hierarchical patterns in data, downsizing and transforming images by summarizing information in a small space.
-Numerous studies have demonstrated the efficacy of convolutional layers in addressing various vision tasks, including image recognition [@rawat2017deep].
-Despite the widespread use of computer vision models in fields like computer-aided diagnosis [@lee2015image], pedestrian detection [@brunetti2018computer], and facial recognition [@emami2012facial], their application in reading data plots remains limited.
-While some studies have explored the use of computer vision models for tasks such as reading recurrence plots for time series regression [@ojeda2020multivariate], time series classification [@chu2019automatic; @hailesilassie2019financial; @hatami2018classification; @zhang2020encoding], anomaly detection [@chen2020convolutional], and pairwise causality analysis [@singh2017deep], the application of reading residual plots with computer vision models represents a relatively new field of study.
-
-In this paper, we develop computer vision models and integrate them into the residual plots diagnostics workflow, filling the gap of.... The paper is structured as follows: ...
-
-```{r false-finding, fig.width = 8/5*4, fig.height = 6/4*4, fig.pos="!h", fig.cap="An example residual vs fitted values plot (red line indicates 0).
-The vertical spread of the data points varies with the fitted values.
-This often indicates the existence of heteroskedasticity."}
-set.seed(10131)
-ori_x <- rand_lognormal()
-mod <- heter_model(b = 0, x = closed_form(~-ori_x))
-ori_dat <- mod$gen(300)
-
-ori_dat %>%
-  VI_MODEL$plot(theme = theme_light(), remove_grid_line = TRUE) +
-  xlab("Fitted values") +
-  ylab("Residuals")
-```
-
-# Methodology
-
-## Different possible configurations of the model formula
-
-There are various configurations of the computer vision model that can be used to assess residual plots.
-We discuss these configurations below based on two key components of the model formula: the input and the output format.
-
-### Input formats
-
-Deep learning models are in general very sensitive to the input data.
-The quality and relevance of the input data greatly influence the model's capacity to generate insightful and meaningful results.
-There are several designs of the input format can be considered.
-
-A straightforward design involves feeding a vector of residuals along with a vector of fitted values, essentially providing all the necessary information for creating a residuals vs fitted values plot.
-However, a drawback of this method is the dynamic input size, which changes based on the number of observations.
-For modern computer vision models implemented in mainstream software like TensorFlow [@abadi2016tensorflow], the input shape is typically fixed.
-One solution is to pad the input vectors with leading or trailing zeros when the input tensor expects longer vectors, but it may fail if the input vector surpasses the designed length.
-Another strategy is to summarize the residuals and fitted values separately using histograms and utilize the counts as the input.
-By controlling the number of bins in the histograms, it becomes possible to provide fixed-length input vectors.
-Still, since histograms only capture the marginal distribution of residuals and fitted values respectively, they can not be used to differentiate visual patterns with same marginal distributions but different joint distributions.
-
-Another design involves using an image as input.
-The primary advantage of this design, as opposed to the vector format, is the availability of the existing and sophisticated image processing architectures developed over the years, such as the VGG16 architecture proposed in @simonyan2014very.
-These architectures can effectively capture and summarize spatial information from nearby pixels, which is less straightforward with vector input.
-The main considerations are the image resolution and the aesthetics of the residual plot.
-In general, higher resolution provides more information to the model but comes with the trade-off of increased complexity and greater difficulty in training.
-As for the aesthetics of the residual plot, a practical solution is to consistently present residual plots in the same style to the model.
-This implies that the model can not accept arbitrary images as input but requires the use of the same preprocessing pipeline to convert residuals and fitted values into a standardized-style residual plot.
-
-Providing multiple residual plots to the model, such as a pair of plots, a triplet or a lineup is also a possible option.
-@chopra2005learning have shown that computer vision models designed for image comparison can assess whether a pair of images are similar or dissimilar.
-Applied to our specific problem, we can define null plots of a fitted regression model to be similar to each other, while considering actual residual plots to be distinct from null plots of any fitted regression model.
-A triplet constitutes a set of three images, denoted as $image_1$, $image_2$ and $image_3$.
-It is often used to predict whether $image_2$ or $image_3$ is more similar to $image_1$, proving particularly useful for establishing rankings between samples.
-For this setup, we can apply the same criteria to define similarity between images.
-However, it is important to note that these two approaches usually require additional considerations regarding the loss function and, at times, non-standard training processes due to shared weights between different convolutional blocks.
-
-Presenting a lineup to a model aligns closely with the lineup protocol.
-However, as the number of residual plots in a lineup increases, the resolution of the input image grows rapidly, posing challenges in training the model.
-We experimented with this approach in a pilot study, but the performance of the trained model was sub-optimal.
-
-We did not explore all the mentioned input formats due the considerable costs associated with data preparation and model training.
-Considering the implementation cost and the interpretability of the model, we settled on the single residual plot input format.
-
-### Output formats
-
-Given that the input is a single residual plot represented as a fixed-resolution image, the output from the computer vision model can take one of two forms: binary or numeric.
-This choice determines whether the model belongs to a classification model or a regression model.
-The binary outcome encoded as $0$ and $1$ could be used to represent whether the input image is a null plot, or whether the input image would be rejected in a visual test conducted by humans.
-Training a model following the latter option requires data from prior human subject experiments, presenting difficulties in controlling the quality of data due to variations in experimental settings across different studies.
-Additionally, some visual inference experiments are unrelated to linear regression models or residual plot diagnostics, resulting in a limited amount of available training data.
-
-Alternatively, the output could be a meaningful and interpretable numerical measure useful for assessing residual plots, such as the strength of suspicious visual patterns reflecting the extent of model violations and the difficulty index for identifying whether a residual plot has no issues.
-However, these numeric measures are often informally used in daily communication but are not typically formalized or rigorously defined.
-For the purpose of training a model, this numeric measure has to be quantifiable.
-
-In this study, we chose to define and use a distance measure to quantify the difference between the residual plot and the null plots.
-@vo2016localizing have also demonstrated that defining a proper distance between images can enhance the matching accuracy in image search compared to a binary outcome model.
-
-## Distance from the good residual plots
-
-In a visual test, the observer will be asked to choose one or more plots that stand out as most distinct from others in a given lineup.
-To develop a computer vision model for assessing residual plots within the visual inference framework, it is important to precisely define a numerical measure of "difference" or "distance" between plots.
-This distance can take the form of a basic statistical operation on pixels, such as the sum of square differences.
-Alternatively, it could involve established image similarity metrics like the Structural Similarity Index Measure [@wang2004image].
-The challenge lies in the fact that metrics tailored for image comparison may not be suitable for evaluating data plots, where only essential plot elements require assessment [@chowdhury2018measuring].
-Furthermore, scagnostics mentioned in Section \ref{introduction} could be used to construct distance metrics for residual plots comparison, but the functional form still needs to be carefully refined to accurately reflect the extent of the violations.
-
-### Residual distribution
-
-The distance measure proposed in this study takes into account the fact that we tried to measure how different a residual plot is from a good residual plot, or in other words, how different a given fitted regression model is from a correctly specified model.
-For the classical normal linear regression model, residuals $\boldsymbol{e}$ are derived from the fitted values $\hat{\boldsymbol{y}}$ and observed values $\boldsymbol{y}$.
-Suppose the data generating process is known and the regression model is correctly specified, by the Frisch-Waugh-Lowell theorem [@frisch1933partial], residuals $\boldsymbol{e}$ can also be treated as random variables and written as a linear transformation of the error $\boldsymbol{\varepsilon}$ formulated as $\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon}$, where $\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the residual operator and an idempotent matrix, $\boldsymbol{X}$ is the design matrix, $\boldsymbol{I}_n$ is a $n$ by $n$ identity matrix, and $n$ is the number of observations.
-
-One of the assumptions of the classical normal linear regression model is the error $\boldsymbol{\varepsilon}$ follows a multivariate normal distribution with zero mean and constant variance, i.e., $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$.
-It can be known that residuals $\boldsymbol{e}$ also follow a certain probability distribution transformed from the multivariate normal distribution, which will be denoted as $Q$.
-This reference distribution $Q$ summarizes what good residuals should follow given the design matrix $\boldsymbol{X}$ is known and fixed.
-
-In ordinary least square, the minimization of the sum of square residuals implies  $\sum_{i=1}^{n} e_i = 0$, making any residual value to be a linear combination of the remaining $n - 1$ residuals.
-This effectively means $\text{rank}(\boldsymbol{R}) = n - 1 < n$ and $Q$ is a degenerate multivariate distribution.
-To capture the characteristics of $Q$, such as moments, we can simulate a large numbers of $\boldsymbol{\varepsilon}$ and transform it to $\boldsymbol{e}$ to get the empirical estimates.
-For simplicity, in this study, we replaced the variance-covariance matrix of residuals $\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}' = \boldsymbol{R}\sigma^2$ with a full-rank diagonal matrix $\text{diag}(\boldsymbol{R}\sigma^2)$, where $\text{diag}(.)$ sets the non-diagonal entries of a matrix to zeros.
-The resulting distribution for $\boldsymbol{Q}$ is $N(\boldsymbol{0}_n, \text{diag}(\boldsymbol{R}\sigma^2))$.
-
-Distribution $Q$ is derived from the correctly specified model.
-However, if the model is misspecified, then the actual distribution of residuals denoted as $P$, will be different from $Q$.
-For example, if the data generating process contains variables correlated with any column of $\boldsymbol{X}$ but not included in $\boldsymbol{X}$, causing an omitted variable problem, $P$ will be different from $Q$ because the residual operator obtained from the fitted regression model will not be the same as $\boldsymbol{R}$.
-Besides, if the $\boldsymbol{\varepsilon}$ follows a non-normal distribution such as a multivariate lognormal distribution, $P$ will usually be skewed and has a long tail.
-
-### Kullback-Leibler divergence of $P$ from $Q$
-
-Define a proper distance between distributions is usually easier than define a proper distance between data plots.
-Given the actual residual distribution $Q$ and the reference residual distribution $P$, we used a distance measure based on Kullback-Leibler divergence [@kullback1951information] to quantify the difference between two distributions
-
-\begin{align}
-\label{eq:kl-0}
-D &= \log\left(1 + D_{KL}\right), \\
-\label{eq:kl-1}
-D_{KL} &= \int_{\mathbb{R}^{n}}\log\frac{p(\boldsymbol{e})}{q(\boldsymbol{e})}p(\boldsymbol{e})d\boldsymbol{e},
-\end{align}
-
-\noindent where $p(.)$ is the probability density function for distribution $P$, and $q(.)$ is the probability density function for distribution $Q$.
-
-This distance measure was first proposed in @li2023plot.
-It was mainly designed for measuring the effect size of non-linearity and heteroskedasticity in a residual plot.
-@li2023plot have showed that, for a classical normal linear regression model that omits a necessary higher-order predictors $\boldsymbol{Z}$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$ with $\boldsymbol{V}$ being an arbitrary symmetric positive semi-definite matrix, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}))$.
-Note that the variance-covariance matrix is replaced with the diagonal matrix to ensure it is a full-rank matrix.
-
-Since both $P$ and $Q$ are adjusted to be multivariate normal distributions, equation \ref{eq:kl-1} can be further expanded to
-
-\begin{align}
-\label{eq:kl-2}
-D_{KL} &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
-\end{align}
-
-\noindent where $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$, and $\boldsymbol{W} = \boldsymbol{R}\boldsymbol{V}\boldsymbol{R}$.
-The assumed error variance $\sigma^2$ is set to be $\text{tr}(\boldsymbol{V})/n$, which is the expectation of the estimated variance.
-
-### Evaluation of Kullback-Leibler divergence for non-normal $P$
-
-For non-normal error $\boldsymbol{\varepsilon}$, the actual residual distribution $P$ is unlikely to be a multivariate normal distribution.
-Thus, equation \ref{eq:kl-2} given in @li2023plot will not be applicable to models violating the normality assumption.
-
-To evaluate the Kullback-Leibler divergence of non-normal $P$ from $Q$, the fallback is to solve equation \ref{eq:kl-1} numerically.
-However, since $\boldsymbol{e}$ is a linear transformation of non-normal random variables, it is very common that the general form of $P$ is unknown, meaning that we can not easily compute $p(\boldsymbol{e})$ using a well-known probability density function.
-Additionally, even if $p(\boldsymbol{e})$ can be calculated for any $\boldsymbol{e} \in \mathbb{R}^n$, it will be very difficult to do numerical integration over the $n$ dimensional space, because $n$ could be potentially very large.
-
-In order to approximate $D_{KL}$ in a practically computable manner, the elements of $\boldsymbol{e}$ are assumed to be independent of each other.
-This assumption solves both of the issues mentioned above.
-First, we no longer need to integrate over $n$ random variables.
-The result of equation \ref{eq:kl-1} is now the sum of the Kullback-Leibler divergence evaluated for each individual residual thanks for the independence assumption.
-Second, it is not required to know the joint probability density $p(\boldsymbol{e})$ any more.
-Instead, the evaluation of Kullback-Leibler divergence for an individual residual relies on the knowledge of the marginal density $p_i(e_i)$, where $e_i$ is the $i$-th residual for $i = 1, ..., n$.
-This is much easier to approximate through simulation.
-It is also worth mentioning that this independence assumption generally will not hold, since $\text{cov}(e_i, e_j) \neq 0$ if $\boldsymbol{R}_{ij} \neq 0$ for any $1 \leq i < j \leq n$, but its existence is essential for reducing the computational cost.
-
-Given $\boldsymbol{X}$ and $\boldsymbol{\beta}$, the algorithm for approximating equation \ref{eq:kl-1} starts from simulating $m$ sets of observed values $\boldsymbol{y}$ according to the data generating process.
-The observed values are stored in a matrix $\boldsymbol{A}$ with $n$ rows and $m$ columns, where each column of $\boldsymbol{A}$ is a set of observed values.
-Then, we can get $m$ sets of realized values of $\boldsymbol{e}$ stored in the matrix $\boldsymbol{B}$ by applying the residual operator $\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}$.
-Furthermore, kernel density estimation (KDE) with Gaussian kernel and optimal bandwidth selected by the Silverman's rule of thumb [@silverman2018density] is applied on each row of $B$ to estimate $p_i(e_i)$ for $i = 1, ..., n$.
-The KDE computation can be done by the `density` function in R.
-
-Since the Kullback-Leibler divergence can be viewed as the expectation of the log-likelihood ratio between distribution $P$ and distribution $Q$ evaluated on distribution $P$, we can reuse the simulated residuals in matrix $\boldsymbol{B}$ to estimate the expectation by the sample mean.
-With the independence assumption, for non-normal $P$, $D_{KL}$ can be approximated by
-
-\begin{align}
-\label{eq:kl-3}
-D_{KL} &\approx \sum_{i = 1}^{n} \hat{D}_{KL}^{(i)}, \\
-\hat{D}_{KL}^{(i)} &= \frac{1}{m}\sum_{j = 1}^{m} log\frac{\hat{p_i}(B_{ij})}{q(B_{ij})},
-\end{align}
-
-\noindent where $\hat{D}_{KL}^{(i)}$ is the estimator of the Kullback-Leibler divergence for an individual residual $e_i$, $B_{ij}$ is the $i$-th row and $j$-th column entry of the matrix $B$, $\hat{p_i}(.)$ is the kernel density estimator of $p_i(.)$, $q(.)$ is the normal density function with mean zero and an assumed variance estimated as $\widehat{\sigma^2} = \sum_{b \in vec(B)}(b - \sum_{b \in vec(B)} b/nm)^2/(nm - 1)$, and $vec(.)$ is the vectorization operator which turns a $n \times m$ matrix into a $nm \times 1$ column vector by stacking the columns of the matrix on top of each other.
-
-### Approximation of the distance measure
-
-In the previous sections, we have defined a distance measure given in equation \ref{eq:kl-0} for quantifying the difference between the actual residual distribution $P$ and an ideal reference distribution $Q$.
-You may have noticed that this distance measure can only be computed when the data generating process is known.
-In reality, we often have no knowledge about the data generating process, otherwise we do not need to fit a linear regression model in the first place.
-
-We tried to train a computer vision model to approximate this distance measure with a residual plot.
-Let $D$ be the result of equation \ref{eq:kl-0} indicating the extent of the model violations, and our estimator $\hat{D}$ is formulated as
-
-\begin{equation}
-\label{eq:d-approx}
-\hat{D} = f_{CV}(V_{h \times w}(\boldsymbol{e}, \hat{\boldsymbol{y}})),
-\end{equation}
-
-\noindent where $V_{h \times w}(.)$ is a plotting function that saves a residuals vs fitted values plot with fixed aesthetic as an image with $h \times w$ pixels and three colour channels, $f_{CV}(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance in the domain $[0, +\infty)$.
-
-With the approximated distance $\hat{D}$, we will be able to know how different the underlying distribution of the residuals is from a good residual distribution.
-It also provides information for the strength of the visual signal embedded in the residual plot.
-
-The approximated distance $\hat{D}$ is not expected to be the same as the original distance $D$.
-This is largely because information contained in a single residual plot is limited and it may not be able to summarise all the important characteristics of the residual distribution.
-For a given residual distribution $P$, we can generate many different residual plots.
-Some of them share similar visual patterns, but some of them could be visually very different from the rest, especially for regression models with small $n$.
-This suggests the error of the approximation will vary depends on whether the observed residual plot is representative or not.
-
-## Statistical testing based on the approximated distance
-
-### Null distribution of the approximated distance
-
-Theoretically, the distance $D$ for a correctly specified model is $0$, because $P$ will be the same as $Q$.
-However, the computer vision model may not necessary predict $0$ for a null plot.
-Using Figure \ref{fig:false-finding} as an example, it contains a visual pattern which is an indication of heteroskedasticity.
-We would not expect the model to be able to magically tell if the suspicious pattern is caused by the skewed distribution of the fitted values or the existence of heteroskedasticity.
-Additionally, some null plots could have outliers or strong visual patterns due to randomness, and a reasonable model will try to summarise those information into the prediction, resulting in $\hat{D} > 0$.
-
-This property is not an issue if $\hat{D} \gg 0$ for which the visual signal of the residual plot is very strong, and we usually do not need any further examination of the significance of the result.
-However, if the visual pattern is weak or moderate, having $\hat{D}$ will not be sufficient to determine if $H_0$ should be rejected.
-
-To solve this issue while aligning with the principle of visual inference, $\hat{D}$ can be viewed as a test statistic.
-And the null distribution of this statistic can be approximated by the empirical distribution of $\hat{D}_{null}$, where $\hat{D}_{null}$ is the approximated distance for a null plot simulated from the fitted regression model.
-The approximation of the distribution involves applying the residual rotation technique [@buja2009statistical] on the fitted regression model to obtain null residuals.
-The null residuals are then used to make null plots and fed into the computer vision model to get predictions.
-The empirical distribution is constructed with the approximated distance.
-
-There are two types of error associated with the approximation of the null distribution.
-These include the sampling error and the estimation error for model parameters.
-Increasing the number of null plots within the distribution approximation can effectively reduce sampling error.
-However, the estimation error for model parameters remains irreducible.
-The residual rotation technique operates under the assumption of a correct fitted model, using the estimated variance instead of the unknown true variance to rescale rotated residuals.
-Consequently, under $H_0$, the true distribution of $\hat{D}_{null}$ will slightly deviate from the approximated null distribution.
-
-### Estimation of quantiles of the null distribution
-
-Let $\hat{D}_{null}^{(i)}$ be the approximated distance of the $i$-th null plots, where $i = 1,...,n_{null}$ and $n_{null} \in \mathbb{N}^+$ is a sufficiently large number.
-Quantiles of the null distribution can then be estimated using the sample quantiles available in statistical software such as R.
-The details of the sample quantile computation can be found in @hyndman1996sample.
-In statistical testing, analysts often care about certain quantiles of the null distribution, such as the 90% quantile, the 95% quantile and the 99% quantile.
-These quantiles are used as critical values to decide if $H_0$ needs to be rejected.
-For example, if $\hat{D}$ is greater than and equal to the 95% sample quantile $Q_{null}(0.95)$, we could say the approximated distance for the actual residual plot is significantly different from the approximated distance for null plots with 95% significance level.
-Based on our experience, in order to get a stable estimate of the 95% quantile, $n_{null}$ usually needs to be at least $100$.
-And if the null distribution has a long tail, more null plots will be needed.
-Alternatively, a p-value can be used to represents the probability of observing an event equally or more extreme than the given event under $H_0$, and it can be estimated by $1/m\sum_{i=1}^{n_{null}}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)$.
-
-If precision in sample quantiles is not the main priority, using a pre-calculated table of quantiles is an available option.
-Such a table offers pre-determined quantiles for a specified number of observations.
-It is generated by assessing numerous null plots derived from various simulated regression models and averaging them.
-Essentially, this shifts the computational burden from the user to the developer.
-
-### Bootstrapping the approximated distance
-
-Bootstrap is often employed in linear regression when conducting inference for estimated parameters.
-It is typically done by sampling individual cases with replacement and refitting the regression model.
-If the observed data accurately reflects the true distribution of the population, the bootstrapped estimates can be used to measure the variability of the parameter estimate without making strong distributional assumptions about the data generating process.
-
-Similarly, bootstrap can be applied on the approximated distance $\hat{D}$.
-For each refitted model $M_{boot}^{(i)}$, there will be an associated residual plot $V_{boot}^{(i)}$ which can be fed into the computer vision model to obtain $\hat{D}_{boot}^{(i)}$, where $i = 1,...,n_{boot}$, and $n_{boot}$ is the number of bootstrapped samples.
-If we are interested in the variation of $\hat{D}$, we can use $\hat{D}_{boot}^{(i)}$ to estimate a confidence interval.
-
-Alternatively, since each $M_{boot}^{(i)}$ has a set of estimated coefficients $\hat{\boldsymbol{\beta}}_{boot}^{(i)}$ and an estimated variance $\hat{\sigma^2}_{boot}^{(i)}$, a new approximated null distribution can be construed and the corresponding 95% sample quantile $Q_{boot}^{(i)}(0.95)$ can be computed.
-Then, if $\hat{D}_{boot}^{(i)} \geq Q_{boot}^{(i)}(0.95)$, $H_0$ will be rejected for $M_{boot}^{(i)}$.
-The ratio of rejected $M_{boot}^{(i)}$ among all the refitted models provides an indication of how often the assumed regression model are considered to be incorrect if the data can be obtained repetitively from the same data generating process.
-But this approach is computationally very expensive because it requires $n_{boot} \times n_{null}$ times of residual plot assessment.
-In practice, $Q_{null}(0.95)$ can be used to replace $Q_{boot}^{(i)}(0.95)$ in the computation.
-
-## Generation of training data
-
-### Data generating process
-
-While observational data is frequently employed in training models for real-world applications, the data generating process of observational data often remains unknown, making computation for our target variable $D$ unattainable.
-Consequently, the computer vision models developed in this study were trained using synthetic data.
-This approach provided us with precise label annotations.
-Additionally, it ensured a large and diverse training dataset, as we had control over the data generating process, and the simulation of the training data was relatively cost-effective.
-
-We have incorporated three types of residual departures of linear regression model in the training data, including non-linearity, heteroskedasticity and non-normality.
-All three departures can be summarised by the data generating process formulated as
-
-\begin{align}
-\label{eq:data-sim}
-\boldsymbol{y} &= \boldsymbol{1}_n + \boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 + \beta_2(\boldsymbol{z} + \beta_1\boldsymbol{w}) + \boldsymbol{k} \odot \boldsymbol{\varepsilon}, \\
-\boldsymbol{z} &= \text{He}_j(g(\boldsymbol{x}_1, 2)), \\
-\boldsymbol{w} &= \text{He}_j(g(\boldsymbol{x}_2, 2)), \\
-\boldsymbol{k} &= \sqrt{\boldsymbol{1}_n + b(2 - |a|)(\boldsymbol{x}_1 + \beta_1\boldsymbol{x}_2 - a\boldsymbol{1}_n)^2},
-\end{align}
-
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{z}$, $\boldsymbol{w}$, $\boldsymbol{k}$ and $\boldsymbol{\varepsilon}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$ are two independent predictors, $\text{He}_j(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau], the $\sqrt{(.)}$ and $(.)^2$ operators are element-wise operators, $\odot$ is the Hadamard product, and $g(., k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
-
-$$g(\boldsymbol{x}, k) = 2k \cdot \frac{\boldsymbol{x} - x_{min}\boldsymbol{1}_n}{x_{max} - x_{min}} - k\boldsymbol{1}_n,~for~k > 0,$$
-\noindent where $x_{min} = \underset{i \in \{ 1,...,n\}}{min} x_i$, $x_{max} = \underset{i \in \{ 1,...,n\}}{max} x_i$ and $x_i$ is the $i$-th entry of $\boldsymbol{x}$.
-
-```{r factor}
-tibble(Factor = c("j", "a", "b", "$\\beta_1$", "$\\beta_2$", " $\\text{dist}_\\varepsilon$", "$\\text{dist}_{x1}$", "$\\text{dist}_{x2}$", "$\\sigma_{\\varepsilon}$", "$\\sigma_{X1}$", "$\\sigma_{X2}$", "n"),
-       Domain = c("\\{2, 3, ..., 18\\}", "[-1, 1]", "[0, 100]", "{0, 1}", "{0, 1}", "\\{discrete, uniform, normal, lognormal\\}", "\\{discrete, uniform, normal, lognormal\\}", "\\{discrete, uniform, normal, lognormal\\}", "[0.0625, 9]", "[0.3, 0.6]", "[0.3, 0.6]", "[50, 500]")) %>%
-  kableExtra::kable("latex", 
-                    escape = FALSE,
-                    booktabs = TRUE,
-                    caption = "Factors used in the data generating process for synthetic data simulation.
-Factor $j$ and $a$ controls the non-linearity shape and the heteroskedasticity shape respectively.
-Factor $b$, $\\sigma_\\varepsilon$ and $n$ control the signal strength.
-Factor $\\text{dist}_\\varepsilon$, $\\text{dist}_{x1}$ and $\\text{dist}_{x2}$ specifies the distribution of $\\varepsilon$, $X_1$ and $X_2$ respectively.")
-```
-
-The residuals and fitted values of the fitted model were obtained by regressing $\boldsymbol{y}$ on $\boldsymbol{x}_1$.
-If $\beta_1 \neq 0$, $\boldsymbol{x}_2$ was also included in the design matrix.
-This data generation process was adapted from @li2023plot, where it was utilized to simulate residual plots exhibiting non-linearity and heteroskedasticity visual patterns for human subject experiments.
-A summary of the factors utilized in Equation \ref{eq:data-sim} is provided in Table \ref{tab:factor}.
-
-In Equation \ref{eq:data-sim}, $\boldsymbol{z}$ and $\boldsymbol{w}$ represent higher-order terms of $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$, respectively.
-If $\beta_2 \neq 0$, the regression model will encounter non-linearity issues.
-Parameter $j$ serves as a shape parameter that controls the number of tuning points in the non-linear pattern.
-Typically, higher values of $j$ lead to an increase in the number of tuning points, as illustrated in Figure \ref{fig:different-j}.
-
-```{r different-j, fig.pos="!h", fig.cap = "Non-linearity forms generated for the synthetic data simulation. The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$."}
-
-set.seed(10086)
-
-# Data for shape 1
-dat_shape_1 <- phn_model(j = 2, include_x2 = FALSE, sigma = 0.05)$gen(500) %>%
-  mutate(j = 2)
-
-# Generate data for shape 2, 3 and 4. Reuse x and e.
-map_df(3:18, function(j) {
-  phn_model(j = j, include_x2 = FALSE, sigma = 0.05)$
-    gen(500, computed = select(dat_shape_1, x1, e)) %>%
-  mutate(j = j)
-}) %>%
-
-  # Combined with data for shape 1
-  bind_rows(dat_shape_1) %>%
-  mutate(j = factor(j)) %>%
-  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
-  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
-```
-
-Additionally, Scaling factor $\boldsymbol{k}$ directly affects the error distribution and it is correlated with $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$.
-If $b \neq 0$ and $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \sigma^2\boldsymbol{I}_n)$, the constant variance assumption will be violated.
-Parameter $a$ is a shape parameter controlling the location of the smallest variance in a residual plot as shown in Figure \ref{fig:different-a}.
-
-```{r different-a, fig.width=8/5*3*2, fig.height=6/4*3*2, fig.pos="!h", fig.cap = 'Heteroskedasticity forms generated for the synthetic data simulation.
-Different shapes are controlled by the continuous factor $a$ between -1 and 1.
-For $a = -1$, the residual plot exhibits a "left-triangle" shape.
-And for $a = 1$, the residual plot exhibits a "right-triangle" shape.
-'}
-
-set.seed(10085)
-
-# Generate data for a = -1
-dat_a_n1 <- phn_model(include_z = FALSE,
-                      include_x2 = FALSE,
-                      a = -1,
-                      b = 100)$gen(500) %>%
-  mutate(a = -1)
-
-# Generate data for other a
-map(c(-0.75, -0.5, -0.25, 0, 0.25, 0.5, 0.75, 1), function(a) {
-  phn_model(include_z = FALSE,
-            include_x2 = FALSE,
-            a = a,
-            b = 100)$gen(500) %>%
-  mutate(a = a)
-}) %>%
-
-  # Combined with data for a = -1
-  bind_rows(dat_a_n1) %>%
-  mutate(a = factor(a)) %>%
-  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
-  facet_wrap(~a, scales = "free", ncol = 3) +
-  xlab("Fitted values") +
-  ylab("Residuals")
-```
-
-```{r different-e, fig.pos="!h", fig.cap = 'Non-normality forms generated for the synthetic data simulation.
-Four different error distributions including discrete, lognormal, normal and uniform are considered.'}
-
-set.seed(10086)
-
-# Data for shape 1
-dat_shape_1 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_uniform(-1.4, 1.4))$gen(500) %>%
-  mutate(e_dist = "uniform")
-
-dat_shape_2 <- phn_model(include_z = FALSE, include_x2 = FALSE, sigma = 0.8)$gen(500) %>%
-  mutate(e_dist = "normal")
-
-dat_shape_3 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_lognormal(sigma = 0.6))$gen(500) %>%
-  mutate(e_dist = "lognormal")
-
-dat_shape_4 <- phn_model(include_z = FALSE, include_x2 = FALSE, e = rand_uniform_d(-1.4, 1.4, even = TRUE))$gen(500) %>%
-  mutate(e_dist = "discrete")
-
-# Generate data for shape 2, 3 and 4. Reuse x and e.
-bind_rows(dat_shape_1, dat_shape_2, dat_shape_3, dat_shape_4) %>%
-  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
-  facet_wrap(~e_dist, scales = "free", labeller = label_parsed, ncol = 2)
-```
-
-Non-normality violations arise from specifying a non-normal distribution for $\boldsymbol{\varepsilon}$.
-In the synthetic data simulation, four distinct error distributions are considered, including discrete, uniform, normal, and lognormal distributions, as presented in Figure \ref{fig:different-e}.
-Each distribution imparts unique characteristics to the residuals.
-The discrete error distribution introduces discreteness in residuals, while the lognormal distribution typically yields outliers.
-Uniform error distribution may result in residuals filling the entire space of the residual plot.
-All of these distributions exhibit visual distinctions from the normal error distribution.
-
-```{r different-j-x2, fig.pos="!h", fig.cap = "Residual plots of multiple linear regression models with non-linearity issues.
-The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$.
-A second predictor $\\boldsymbol{x}_2$ is introduced to the regression model to create complex shapes."}
-
-set.seed(10086)
-
-# Data for shape 1
-dat_shape_1 <- phn_model(j = 2, include_x2 = TRUE, sigma = 0.05)$gen(500) %>%
-  mutate(j = 2)
-
-# Generate data for shape 2, 3 and 4. Reuse x and e.
-map_df(3:18, function(j) {
-  phn_model(j = j, include_x2 = TRUE, sigma = 0.05)$
-    gen(500, computed = select(dat_shape_1, x1, e)) %>%
-  mutate(j = j)
-}) %>%
-
-  # Combined with data for shape 1
-  bind_rows(dat_shape_1) %>%
-  mutate(j = factor(j)) %>%
-  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
-  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
-```
-
-```{r different-j-heter, fig.pos="!h", fig.cap = 'Residual plots of models violating both the non-linearity and the heteroskedasticity assumptions.
-The 17 shapes are generated by varying the order of polynomial given by $j$ in $He_j(.)$, and the "left-triangle" shape is introduced by setting $a = -1$.'}
-
-set.seed(10086)
-
-# Data for shape 1
-dat_shape_1 <- phn_model(j = 2, a = -1, b = 100, include_x2 = FALSE, sigma = 0.05)$gen(500) %>%
-  mutate(j = 2)
-
-# Generate data for shape 2, 3 and 4. Reuse x and e.
-map_df(3:18, function(j) {
-  phn_model(j = j, a = -1, b = 100, include_x2 = FALSE, sigma = 0.05)$
-    gen(500, computed = select(dat_shape_1, x1, e)) %>%
-  mutate(j = j)
-}) %>%
-
-  # Combined with data for shape 1
-  bind_rows(dat_shape_1) %>%
-  mutate(j = factor(j)) %>%
-  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
-  facet_wrap(~j, scales = "free", labeller = label_parsed, ncol = 5)
-```
-
-```{r different-e-heter, fig.pos="!h", fig.cap = 'Residual plots of models violating both the non-normality and the heteroskedasticity assumptions.
-The four shapes are generated by using four different error distributions including discrete, lognormal, normal and uniform, and the "left-triangle" shape is introduced by setting $a = -1$.
-'}
-
-set.seed(10085)
-
-# Data for shape 1
-dat_shape_1 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_uniform(-1.4, 1.4))$gen(500) %>%
-  mutate(e_dist = "uniform")
-
-dat_shape_2 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, sigma = 0.8)$gen(500) %>%
-  mutate(e_dist = "normal")
-
-dat_shape_3 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_lognormal(sigma = 0.6))$gen(500) %>%
-  mutate(e_dist = "lognormal")
-
-dat_shape_4 <- phn_model(a = -1, b = 100, include_z = FALSE, include_x2 = FALSE, e = rand_uniform_d(-1.4, 1.4, even = TRUE))$gen(500) %>%
-  mutate(e_dist = "discrete")
-
-# Generate data for shape 2, 3 and 4. Reuse x and e.
-bind_rows(dat_shape_1, dat_shape_2, dat_shape_3, dat_shape_4) %>%
-  VI_MODEL$plot(remove_axis = TRUE, remove_grid_line = TRUE, theme = theme_light()) +
-  facet_wrap(~e_dist, scales = "free", labeller = label_parsed, ncol = 2)
-```
-
-Equation \ref{eq:data-sim} accommodates the incorporation of the second predictor $\boldsymbol{x}_2$.
-Introducing it into the data generation process by setting $\beta_1 = 1$ significantly enhances the complexity of the shapes, as illustrated in Figure \ref{fig:different-j-x2}.
-In comparison to Figure \ref{fig:different-j}, Figure \ref{fig:different-j-x2} demonstrates that the non-linear shape resembles a surface rather than a single curve.
-This augmentation can facilitate the computer vision model in learning visual patterns from residual plots of the multiple linear regression model.
-
-In real-world analysis, it's not uncommon to encounter instances where multiple model violations coexist.
-In such cases, the residual plots often exhibit a mixed pattern of visual anomalies corresponding to different types of model violations.
-Figure \ref{fig:different-j-heter} and \ref{fig:different-e-heter} show the visual patterns of models with multiple model violations.
-
-### Computation of scagnostics
+## Computation of scagnostics
 
@@
 
-For each generated residual plot, we computed four scagnostics – "Monotonic," "Sparse," "Splines," and "Striped" – using the cassowaryr R package [@mason2022cassowaryr].
+For each generated residual plot, we computed four scagnostics – "Monotonic," "Sparse," "Splines," and "Striped" – using the `cassowaryr` R package [@mason2022cassowaryr].
 These computed measures, along with the number of observations from the fitted model, were provided as the second input for the computer vision model.
-Although other scagnostics are informative, they are currently unavailable due to a fatal bug in the compiled C program of the interp R package [@Albrecht2023interp], which may unpredictably crash the process.
+Although other scagnostics are informative, they are currently unavailable due to a fatal bug in the compiled C program of the `interp` R package [@Albrecht2023interp], which may unpredictably crash the process.
 For reproducibility, we excluded these scagnostics from the training data.
-
-### Crafting a balanced training set
-
-To train a robust computer vision model, we deliberately controlled the distribution of the target variable $D$ in the training data.
-We ensured that it followed a uniform distribution between $0$ and $7$.
-This was achieved by organizing $50$ buckets, each exclusively accepting training samples with $D$ falling within the range $[7(i - 1)/49, 7i/49)$ for $i < 50$, where $i$ represents the index of the $i$-th bucket.
-For the $50$-th bucket, any training samples with $D \geq 7$ were accepted.
-
-With 80000 training images prepared, each bucket accommodated a maximum of $80000 \div 50 = 1600$ training samples.
-The simulator iteratively sampled parameter values from the parameter space, generated residuals and fitted values using the data generation process, computed the distance, and checked if the sample fitted within the corresponding bucket.
-This process continued until all buckets were filled.
-
-Similarly, we adopted the same methodology to prepare 8000 test images for performance evaluation and model diagnostics.
-
-## Architecture of the computer vision model
-
-The architecture of the computer vision model is adapted from a well-established architecture known as VGG16, which has demonstrated high performance in image classification [@simonyan2014very].
-
-The model begins with an input layer of shape $n \times h \times w \times 3$, capable of handling $n$ RGB images.
-This is followed by a grayscale conversion layer utilizing the luma formula under the Rec.
-601 standard, which converts the color image to grayscale.
-Grayscale suffices for our task since data points are plotted in black.
-We experiment with three combinations of $h$ and $w$: $32 \times 32$, $64 \times 64$, and $128 \times 128$, aiming to achieve sufficiently high image resolution for the problem at hand.
-
-The processed image is used as the input for the first convolutional block.
-The model comprises at most five consecutive convolutional blocks, mirroring the original VGG16 architecture.
-Within each block, there are two 2D convolutional layers followed by two activation layers, respectively.
-Subsequently, a 2D max-pooling layer follows the second activation layer.
-The 2D convolutional layer convolves the input with a fixed number of $3 \times 3$ convolution filters, while the 2D max-pooling layer downsamples the input along its spatial dimensions by taking the maximum value over a $2 \times 2$ window for each channel of the input.
-The activation layer employs the rectified linear unit (ReLU) activation function, a standard practice in deep learning, which introduces a non-linear transformation of the output of the 2D convolutional layer.
-Additionally, to regularize training, a batch normalization layer is added after each 2D convolutional layer and before the activation layer.
-Finally, a dropout layer is appended at the end of each convolutional block to randomly set some inputs to zero during training, further aiding in regularization.
-
-The output of the last convolutional block is summarized by either a global max pooling layer or a global average pooling layer, resulting in a two-dimensional tensor.
-To leverage the information contained in scagnostics, this tensor is concatenated with an additional $n \times 5$ tensor, which contains the "Monotonic," "Sparse," "Splines," and "Striped" measures, along with the number of observations for $n$ residual plots.
-
-The concatenated tensor is then fed into the final prediction block.
-This block consists of two fully-connected layers.
-The first layer contains at least $128$ units, followed by a dropout layer.
-Occasionally, a batch normalization layer is inserted between the fully-connected layer and the dropout layer for regularization purposes.
-The second fully-connected layer consists of only one unit, serving as the output of the model.
-
-The model weights $\boldsymbol{\theta}$ were randomly initialized and they were optimized by the Adam optimizer with the mean square error loss function
-
-$$\hat{\boldsymbol{\theta}} = \underset{\boldsymbol{\theta}}{argmin}\frac{1}{n_{train}}\sum_{i=1}^{n_{train}}(D_i - f_{\boldsymbol{\theta}}(V_i, S_i))^2,$$
-
-\noindent where $n_{train}$ is the number of training samples, $V_i$ is the $i$-th residual plot and $S_i$ is the additional information about the $i$-th residual plot including four scagnostics and the number of observations.
-
-## Training and hyperparameter tuning
-
-To achieve an optimal deep learning model, hyperparameters like the learning rate often need to be fine-tuned using a tuner.
-In our study, we utilized the Bayesian optimization tuner from the KerasTuner Python library [@omalley2019kerastuner] for this purpose.
-A comprehensive list of hyperparameters is provided in Table \ref{tab:hyperparameter}.
-
-The "number of base filters" determines the number of filters for the first 2D convolutional layer.
-In the VGG16 architecture, the number of filters for the 2D convolutional layer in a block is typically twice the number in the previous block, except for the last block, which maintains the same number of convolution filters as the previous one.
-This hyperparameter aids in controlling the complexity of the computer vision model.
-Higher numbers of base filters result in more trainable parameters.
-Likewise, the "number of units for the fully-connected layer" determines the complexity of the final prediction block.
-Increasing the number of units enhances model complexity, resulting in more trainable parameters.
-
-The dropout rate and batch normalization are flexible hyperparameters that work in conjunction with other parameters to facilitate smooth training.
-A higher dropout rate is necessary when the model tends to overfit the training dataset by learning too much noise.
-Conversely, a lower dropout rate is preferred when the model is complex and challenging to converge.
-Batch normalization, on the other hand, addresses the internal covariate shift problem arising from the randomness in weight initialization and input data.
-It helps stabilize and accelerate the training process by normalizing the activations of each layer.
-
-Additionally, incorporating additional inputs such as scagnostics and the number of observations can potentially enhance prediction accuracy.
-Therefore, we allowed the tuner to determine whether these inputs were necessary for optimal model performance.
-
-The learning rate is a crucial hyperparameter, as it dictates the step size of the optimization algorithm.
-A high learning rate can help the model avoid local minima but risks overshooting and missing the global minimum.
-Conversely, a low learning rate smoothens the training process but makes the convergence time longer and increases the likelihood of getting trapped in local minima.
-
-Our model was trained on the M3 high-performance computing platform, using TensorFlow and Keras.
-During training, 80% of the training data was utilized for actual training, while the remaining 20% was used as validation data.
-The Bayesian optimization tuner conducted 100 trials to identify the best hyperparameter values based on validation root mean square error.
-The tuner then restored the best epoch of the best model from the trials.
-Additionally, we applied early stopping, terminating the training process if the validation root mean square error fails to improve for 50 epochs.
-The maximum allowed epochs was set at 2000, although no models reached this threshold.
-
-```{r hyperparameter}
-data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
-           Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[1e-8, 1e-1]")) %>%
-  kableExtra::kable("latex", booktabs = TRUE, caption = "Name of hyperparameters and their correspoding domain for the computer vision model.")
-```
-
-## Model evaluation methods
-
-- RMSE for the test set
-- R^2
-- Mean bias deviation to understand the overall bias
-- quantile loss to understand how well the model captures the entire distribution of D
-- Percentage of predictions within a tolerance interval (like 0.1)
-
-- comparison with human visual inference
-- overview of the human subject experiment
-- metrics
-- agreement rate
-- check cases that are aggreed
-- check cases that are disaggreed
-
-# Results
-
-## Optimized hyperparameters
-
-Based on the tuning process described in Section \ref{training-and-hyperparameter-tuning}, the optimized hyperparameter values are presented in Table \ref{tab:best-hyperparameter}.
-It was observable that a minimum of $32$ base filters was necessary, with the preferable choice being $64$ base filters for both the $64 \times 64$ and $128 \times 128$ models, mirroring the original VGG16 architecture.
-The optimized dropout rate for convolutional blocks hovered around $0.4$, and incorporating batch normalization for convolutional blocks proved beneficial for performance.
-
-All optimized models chose to retain the additional inputs, contributing to the reduction of validation error.
-The number of units required for the fully-connected layer was $256$, a relatively modest number compared to the VGG16 classifier, suggesting that the problem at hand was relatively straightforward.
-The optimized learning rates were typically smaller than the default value recommended by Keras, which is 0.001.
-
-```{r best-hyperparameter}
-data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
-           `32` = c("32", "0.4", "true", "max", "false", "256", "false", "0.2", "0.0003"),
-           `64` = c("64", "0.3", "true", "average", "false", "256", "true", "0.4", "0.0006"),
-           `128` = c("64", "0.4", "true", "average", "false", "256", "true", "0.1", "0.0052")) %>%
-  kableExtra::kable("latex", 
-                    booktabs = TRUE, 
-                    caption = "Hyperparameters values for the optimized computer vision models with different input sizes.", 
-                    col.names = c("Hyperparameter", "$32 \\times 32$", "$64 \\times 64$", "$128 \\times 128$"),
-                    escape = FALSE) %>%
-  kableExtra::kable_styling(latex_options = c("scale_down"))
-```
-
-## Model performance
-
-The training and test performance for the optimized models with three different input sizes are summarized in Table \ref{tab:performance}.
-Among these models, the $64 \times 64$ model and the $32 \times 32$ model consistently exhibited the best training and test performance, respectively.
-The mean absolute error indicated that the difference between $\hat{D}$ and $D$ was approximately $0.43$ on the test set, a negligible deviation considering the normal range of $D$ typically falls between $0$ and $7$.
-The high $R^2$ values also suggested that the predictions were largely linearly correlated with the target.
-
-Figure \ref{fig:model-performance} presents a hexagonal heatmap for $D - \hat{D}$ versus $\hat{D}$.
-The red smoothing curves, fitted by generalized additive models [@hastie2017generalized], demonstrate that all the optimized models perform admirably on both the training and test sets.
-No structural issues are noticeable in Figure \ref{fig:model-performance}, but some minor issues regarding over-prediction and under-prediction are observed.
-
-The figure highlights that most under-predictions occurred when $\hat{D} < 3$, while over-predictions occurred predominantly when $3 < \hat{D} < 6$.
-To comprehend the reasons behind these deviations, we conducted a meta-analysis on $D - \hat{D}$ for all factors used in the data generating process.
-A regression tree was constructed using the rpart R package [@Terry2022rpart].
-The results of the regression tree are provided in the Appendix.
-
-The meta-analysis revealed that most issues arose from non-linearity problems and the presence of a second predictor in the regression model.
-When the error distribution had a very small variance, all computer vision models tended to under-predict the distance.
-Conversely, when the error distribution had a large variance, all models tended to over-predict the distance.
-Therefore, Additionally, for input images representing null plots, it was expected that the models will over-predict the distance, as explained in Section \ref{null-distribution-of-the-approximated-distance}.
-
-Since most of the deviation stemmed from the presence of non-linearity violations, to further investigate this, we split the test set and re-evaluated the performance, as detailed in Table \ref{tab:performance-sub}.
-It was evident that metrics for null plots were notably worse compared to other categories.
-Furthermore, residual plots solely exhibiting non-normality issues were the easiest to predict, with very low test root mean square error (RMSE) around $0.3$, very low mean absolute error (MAE) around $0.2$, and very high $R^2$ around $0.94$.
-Residual plots with non-linearity issues were more challenging to assess than those with heteroskedasticity or non-normality issues.
-Assessing residual plots with heteroskedasticity was not as difficult as assessing those with non-linearity issues.
-When multiple violations were introduced to a residual plot, the performance metrics typically lay between the metrics for each individual violation.
-
-Based on the model performance metrics, we will only use the best-performing model evaluated on the test set, namely the $32 \times 32$ model, for the subsequent analysis.
-
-```{r performance}
-test_pred <- readr::read_csv(here::here("paper/data/test_pred.csv"))
-train_pred <- readr::read_csv(here::here("paper/data/train_pred.csv"))
-meta <- readr::read_csv(here::here("paper/data/meta.csv"))
-
-test_summary <- test_pred %>% 
-  left_join(meta) %>%
-  group_by(res) %>%
-  summarise(RMSE = yardstick::rmse_vec(effect_size, vss),
-            R2 = yardstick::rsq_vec(effect_size, vss),
-            MAE = yardstick::mae_vec(effect_size, vss),
-            HUBER = yardstick::huber_loss_vec(effect_size, vss)) %>%
-  mutate(type = "test")
-
-train_summary <- train_pred %>% 
-  left_join(meta) %>%
-  group_by(res) %>%
-  summarise(RMSE = yardstick::rmse_vec(effect_size, vss),
-            R2 = yardstick::rsq_vec(effect_size, vss),
-            MAE = yardstick::mae_vec(effect_size, vss),
-            HUBER = yardstick::huber_loss_vec(effect_size, vss)) %>%
-  mutate(type = "train")
-
-train_summary %>%
-  bind_rows(test_summary) %>%
-  select(res, RMSE, R2, MAE, HUBER) %>%
-  mutate(res = glue::glue("${res} \\times {res}$")) %>%
-  mutate(across(RMSE:MAE, ~format(.x, digits = 3))) %>%
-  mutate(across(HUBER, ~format(.x, digits = 2))) %>%
-  kableExtra::kable(format = "latex", 
-                    booktabs = TRUE, 
-                    escape = FALSE,
-                    col.names = c("", "RMSE", "$R^2$", "MAE", "Huber loss"),
-                    caption = "The training and test performance of three optimized models with different input sizes. The best metrics are colored in red.") %>%
-  kableExtra::pack_rows("Training set", 1, 3) %>%
-  kableExtra::pack_rows("Test set", 4, 6) %>%
-  kableExtra::column_spec(2:5, color = c("black", "red", "black",
-                                       "red", "black", "black"))
-
-```
-
-```{r model-performance, fig.pos="!h", fig.cap = "Hexagonal heatmap for residuals vs predicted distance on training and test data for three optimized models with different input sizes.
-The red lines are smoothing curves produced by fitting gnealized additive models.
-The area over the zero line in light pink indicates under-prediction, and the area under the zero line in light green indicates over-prediction.", dev = 'png', dpi = 300}
-model_pred <- train_pred %>% 
-  left_join(meta) %>%
-  mutate(type = "train") %>%
-  bind_rows(test_pred %>% 
-              left_join(meta) %>%
-              mutate(type = "test"))
-
-model_pred %>%
-  ggplot() +
-  geom_hline(yintercept = 0, alpha = 0.5) +
-  # geom_point(aes(effect_size, effect_size - vss), alpha = 0.03) +
-  annotate("rect", 
-           ymin = min(model_pred$effect_size - model_pred$vss),
-                ymax = 0,
-                xmin = min(model_pred$vss),
-                xmax = max(model_pred$vss),
-            fill = "#66c2a5",
-            alpha = 0.3) +
-  geom_point(data = NULL, aes(5, 0, col = "over-prediction"), shape = 15) +
-  annotate("rect", 
-           ymin = 0,
-                ymax = max(model_pred$effect_size - model_pred$vss),
-                xmin = min(model_pred$vss),
-                xmax = max(model_pred$vss),
-            fill = "#fc8d62",
-            alpha = 0.3) +
-  geom_point(data = NULL, aes(5, 0, col = "under-prediction"), shape = 15) +
-  geom_hex(aes(effect_size, effect_size - vss), bins = 25) +
-  geom_smooth(aes(vss, effect_size - vss), se = FALSE, col = "#e34a33") +
-  facet_grid(type ~ res) +
-  coord_fixed() +
-  ylab(expression(D - hat(D))) +
-  xlab(expression(hat(D))) +
-  theme_light() +
-  scale_fill_continuous(limits = c(1, 200), trans = "log10", low = "#56B1F7", high = "#132B43") +
-  scale_color_manual(values = c("under-prediction" = scales::alpha("#fc8d62", 0.3),
-                                "over-prediction" = scales::alpha("#66c2a5", 0.3)),
-                     breaks = c("under-prediction", "over-prediction")) +
-  guides(col = guide_legend(title = NULL, override.aes = list(size = 8), order = 1))
-```
-
-```{r eval = FALSE}
-
- # 1) root 8000 3466.12800  0.04808905  
- #   2) e_sigma>=4.108756 3288 1146.50300 -0.09147268  
- #     4) include_non_normal=FALSE 1983  817.30850 -0.22151770  
- #       8) include_heter=FALSE 1210  416.16970 -0.39866830  
- #        16) include_z=FALSE 49   20.28332 -1.22777500 * (only non-normal and large error sigma -> large over-prediction)
- #        17) include_z=TRUE 1161  360.78130 -0.36367590 *
- #       9) include_heter=TRUE 773  303.72640  0.05578156 *
- #     5) include_non_normal=TRUE 1305  244.69960  0.10613590 *
- #   3) e_sigma< 4.108756 4712 2210.89500  0.14547420  
- #     6) include_z=FALSE 2030  568.99490 -0.04100357 *
- #     7) include_z=TRUE 2682 1517.87900  0.28661890  
- #      14) include_x2=FALSE 1432  377.35510  0.07228115 *
- #      15) include_x2=TRUE 1250  999.37090  0.53216420  
- #        30) e_sigma>=0.1149429 1202  829.80300  0.49933640 *
- #        31) e_sigma< 0.1149429 48  135.83470  1.35422700  (non-linearity, second predictor, small error sigma)
- #          62) measure_striped< 0.44 40   54.18083  0.92957180 *
- #          63) measure_striped>=0.44 8   38.37423  3.47750200 * (high discreteness -> large under-prediction)
-
-# n= 8000 
-# 
-# node), split, n, deviance, yval
-#       * denotes terminal node
-# 
-#     1) root 8000 3466.1280000  0.04808905  
-#       2) e_sigma>=4.108756 3288 1146.5030000 -0.09147268  
-#         4) include_non_normal=FALSE 1983  817.3085000 -0.22151770  
-#           8) include_heter=FALSE 1210  416.1697000 -0.39866830  
-#            16) include_z=FALSE 49   20.2833200 -1.22777500  
-#              32) measure_sparse< 0.3630358 7    8.9396360 -2.12605300 * (large e_sigma, null plots, not very sparse -> large over-prediction)
-#              33) measure_sparse>=0.3630358 42    4.7539680 -1.07806200 * (large e_sigma, null plots, sparse -> large over-prediction)
-#            17) include_z=TRUE 1161  360.7813000 -0.36367590  
-#              34) measure_splines>=0.02606246 376  134.3999000 -0.59957120  
-#                68) n< 143.5 116   39.6936700 -0.84719690  
-#                 136) measure_monotonic>=0.07456592 7    1.3762680 -1.55893700 * (non-linearity only, small n, large e_sigma -> large over-prediction)
-#                 137) measure_monotonic< 0.07456592 109   34.5436600 -0.80148880 *
-#                69) n>=143.5 260   84.4197800 -0.48909200  
-#                 138) e_sigma>=7.367384 63   15.0193100 -0.79693070 *
-#                 139) e_sigma< 7.367384 197   61.5210600 -0.39064610  
-#                   278) include_x2=FALSE 90   19.5837000 -0.53918530 *
-#                   279) include_x2=TRUE 107   38.2813500 -0.26570670  
-#                     558) measure_sparse< 0.5586777 84   27.3861700 -0.37645850 *
-#                     559) measure_sparse>=0.5586777 23    6.1018500  0.13877850 *
-#              35) measure_splines< 0.02606246 785  195.4365000 -0.25068660  
-#                70) e_sigma>=5.64113 529  100.6972000 -0.37916600  
-#                 140) n< 109 41   11.8877000 -0.73213180 *
-#                 141) n>=109 488   83.2723700 -0.34951110  
-#                   282) e_sigma>=7.90547 154   26.5350800 -0.48055150 *
-#                   283) e_sigma< 7.90547 334   52.8735800 -0.28909130 *
-#                71) e_sigma< 5.64113 256   67.9628700  0.01480427  
-#                 142) include_x2=FALSE 113   23.8143300 -0.15341150 *
-#                 143) include_x2=TRUE 143   38.4243300  0.14773000  
-#                   286) n< 72.5 7    2.6335630 -0.69382120 *
-#                   287) n>=72.5 136   30.5781400  0.19104510 *
-#           9) include_heter=TRUE 773  303.7264000  0.05578156  
-#            18) b< 32.23487 487  221.2986000 -0.09202344  
-#              36) b< 0.2107948 67   26.8944500 -0.51926710  
-#                72) n< 126 7    3.3765570 -1.24628700 * (large e_sigma, heteroskedasticity, small b, small n -> over-prediction)
-#                73) n>=126 60   19.3863300 -0.43444810 *
-#              37) b>=0.2107948 420  180.2232000 -0.02386791  
-#                74) measure_splines>=0.05482434 223   77.7820400 -0.18200180  
-#                 148) b< 7.647324 119   51.8788300 -0.32137070  
-#                   296) measure_sparse>=0.4117995 111   43.0150500 -0.38433220 *
-#                   297) measure_sparse< 0.4117995 8    2.3184780  0.55221930 *
-#                 149) b>=7.647324 104   20.9469900 -0.02253160 *
-#                75) measure_splines< 0.05482434 197   90.5523700  0.15513650  
-#                 150) b>=5.165768 63   26.5340300 -0.02749363  
-#                   300) include_x2=FALSE 35    7.9867080 -0.19247040 *
-#                   301) include_x2=TRUE 28   16.4039600  0.17872730  
-#                     602) measure_monotonic< 0.03127537 21    4.3016860 -0.07172379 *
-#                     603) measure_monotonic>=0.03127537 7    6.8333090  0.93008060 *
-#                 151) b< 5.165768 134   60.9291300  0.24099990  
-#                   302) x2_sigma< 0.5865284 126   52.7919200  0.19154170  
-#                     604) x2_k>=5.5 106   43.2640200  0.11519510  
-#                      1208) j< 17.5 93   38.5219100  0.05750563  
-#                        2416) e_even=TRUE 41   15.2536200 -0.17697420  
-#                          4832) measure_sparse>=0.5392181 29    8.8138120 -0.40043160 *
-#                          4833) measure_sparse< 0.5392181 12    1.4922590  0.36304760 *
-#                        2417) e_even=FALSE 52   19.2367200  0.24238400  
-#                          4834) measure_splines>=0.01044483 31    7.8888290 -0.02089042 *
-#                          4835) measure_splines< 0.01044483 21    6.0272570  0.63102720 *
-#                      1209) j>=17.5 13    2.2184070  0.52789690 *
-#                     605) x2_k< 5.5 20    5.6354220  0.59617870 *
-#                   303) x2_sigma>=0.5865284 8    2.9747040  1.01996600 * (large e_sigma, small spline dependence, heteroskedasticity, large x2_sigma -> large under-prediction)
-#            19) b>=32.23487 286   53.6723400  0.30746350  
-#              38) include_x2=FALSE 139   17.0423600  0.13425010 *
-#              39) include_x2=TRUE 147   28.5161400  0.47125030 *
-#         5) include_non_normal=TRUE 1305  244.6996000  0.10613590  
-#          10) measure_splines>=0.08262098 455   47.2275500  0.00622856 *
-#          11) measure_splines< 0.08262098 850  190.4994000  0.15961570  
-#            22) include_heter=FALSE 594   22.8029300  0.08838938 *
-#            23) include_heter=TRUE 256  157.6908000  0.32488300  
-#              46) measure_striped>=0.47 185   31.8984100  0.17734340 *
-#              47) measure_striped< 0.47 71  111.2723000  0.70931710  
-#                94) e_even=FALSE 32   18.0084200  0.22023240  
-#                 188) a>=-0.205314 21    5.0433870 -0.06070074 *
-#                 189) a< -0.205314 11    8.1435320  0.75655920 *
-#                95) e_even=TRUE 39   79.3286800  1.11061700  
-#                 190) b< 2.373839 11    0.2423938  0.06215229 *
-#                 191) b>=2.373839 28   62.2437600  1.52251400  (discrete e_dist,  high discreteness, heteroskedasticity, small splines dependence, non-normality, large e_sigma -> large under-prediction)
-#                   382) x2_dist=even_discrete,lognormal,uniform 19   33.0782400  1.01336800 *
-#                   383) x2_dist=normal 9   13.8421700  2.59737800 *
-#       3) e_sigma< 4.108756 4712 2210.8950000  0.14547420  
-#         6) include_z=FALSE 2030  568.9949000 -0.04100357  
-#          12) b< 28.58967 1159  316.0764000 -0.12131010  
-#            24) include_non_normal=FALSE 441  185.6341000 -0.33971870  
-#              48) include_heter=FALSE 60    8.3705760 -1.15981800 * (null plots, small e_sigma -> large over-prediction)
-#              49) include_heter=TRUE 381  130.5548000 -0.21056910  
-#                98) b< 0.4619795 85   22.8319800 -0.60512210 *
-#                99) b>=0.4619795 296   90.6909300 -0.09726844  
-#                 198) b>=1.080411 252   72.3919200 -0.15604160  
-#                   396) b< 2.878015 65   20.2765800 -0.37135770 *
-#                   397) b>=2.878015 187   48.0544100 -0.08119905 *
-#                 199) b< 1.080411 44   12.4430700  0.23934130 *
-#            25) include_non_normal=TRUE 718   96.4846700  0.01283772 *
-#          13) b>=28.58967 871  235.4979000  0.06585671  
-#            26) include_heter=FALSE 361  111.6977000 -0.10528680  
-#              52) include_non_normal=FALSE 46    6.2078660 -1.13547200 * (null plots, small e_sigma -> large over-prediction)
-#              53) include_non_normal=TRUE 315   49.5417500  0.04515286  
-#               106) e_sigma>=0.2273127 247   17.7293100 -0.01172324 *
-#               107) e_sigma< 0.2273127 68   28.1111100  0.25174690  
-#                 214) measure_splines>=0.03710735 23   10.7097800 -0.09471551 *
-#                 215) measure_splines< 0.03710735 45   13.2294000  0.42882770 *
-#            27) include_heter=TRUE 510  105.7420000  0.18699950  
-#              54) include_non_normal=TRUE 235   41.6282300  0.05417806 *
-#              55) include_non_normal=FALSE 275   56.4252500  0.30050140  
-#               110) include_x2=FALSE 137   20.9155600  0.17983570 *
-#               111) include_x2=TRUE 138   31.5346400  0.42029280  
-#                 222) b< 70.08912 96   17.2702500  0.30834030 *
-#                 223) b>=70.08912 42   10.3110100  0.67618420 *
-#         7) include_z=TRUE 2682 1517.8790000  0.28661890  
-#          14) include_x2=FALSE 1432  377.3551000  0.07228115  
-#            28) measure_splines>=0.07058104 731  178.2000000 -0.01483915  
-#              56) e_sigma>=2.44726 138   40.9281500 -0.18500620  
-#               112) include_heter=FALSE 59   25.9010700 -0.41304000  
-#                 224) include_non_normal=FALSE 38   17.2724700 -0.68141390  
-#                   448) measure_splines>=0.1003984 20    5.6757540 -1.07654300 * (non-linearity only, moderate e_sigma -> large over-prediction)
-#                   449) measure_splines< 0.1003984 18    5.0046810 -0.24238140 *
-#                 225) include_non_normal=TRUE 21    0.9391262  0.07258882 *
-#               113) include_heter=TRUE 79    9.6678490 -0.01470237 *
-#              57) e_sigma< 2.44726 593  132.3458000  0.02476126  
-#               114) e_sigma>=0.158351 527  105.6393000 -0.00147316 *
-#               115) e_sigma< 0.158351 66   23.4476700  0.23423920  
-#                 230) b< 71.04535 55   15.9627800  0.12578350 *
-#                 231) b>=71.04535 11    3.6032230  0.77651720 *
-#            29) measure_splines< 0.07058104 701  187.8212000  0.16312990  
-#              58) n< 62.5 33   14.3260500 -0.31035730 *
-#              59) n>=62.5 668  165.7314000  0.18652070 *
-#          15) include_x2=TRUE 1250  999.3709000  0.53216420  
-#            30) e_sigma>=0.1149429 1202  829.8030000  0.49933640  
-#              60) e_sigma>=2.906419 328  129.1838000  0.27288400  
-#               120) measure_splines>=0.04413018 134   50.5691600  0.09851024  
-#                 240) b< 2.910823 23   10.1825400 -0.20213530  
-#                   480) x2_k>=6.5 15    4.3945260 -0.49662390 *
-#                   481) x2_k< 6.5 8    2.0480610  0.35003080 *
-#                 241) b>=2.910823 111   37.8769400  0.16080620  
-#                   482) b>=7.320158 97   22.8753700  0.08661750 *
-#                   483) b< 7.320158 14   10.7686400  0.67482760 *
-#               121) measure_splines< 0.04413018 194   71.7259100  0.39332780  
-#                 242) j< 7.5 60   32.0807400  0.19368700  
-#                   484) x2_sigma< 0.343438 10   12.3504400 -0.51629950 *
-#                   485) x2_sigma>=0.343438 50   13.6813400  0.33568430 *
-#                 243) j>=7.5 134   36.1830100  0.48271920 *
-#              61) e_sigma< 2.906419 874  677.4868000  0.58432080  
-#               122) measure_splines>=0.04519141 613  424.1886000  0.45960850  
-#                 244) n>=202.5 393  209.4163000  0.32432050  
-#                   488) include_non_normal=TRUE 188   90.6250100  0.18780810  
-#                     976) x2_k< 7.5 101   41.5156400  0.04431936 *
-#                     977) x2_k>=7.5 87   44.6157600  0.35438700  
-#                      1954) measure_sparse>=0.391553 80   35.8617500  0.29057300 *
-#                      1955) measure_sparse< 0.391553 7    4.7050420  1.08369000 * (not sparse, non-normality, large n, small e_sigma, second predictor, non-linearity -> large under-prediction)
-#                   489) include_non_normal=FALSE 205  112.0748000  0.44951240  
-#                     978) e_sigma>=0.2441489 175   86.5106600  0.36295000  
-#                      1956) measure_splines>=0.07729252 127   55.6660800  0.25807550 *
-#                      1957) measure_splines< 0.07729252 48   25.7519700  0.64043040  
-#                        3914) b>=21.95702 24    7.5694370  0.34280130 *
-#                        3915) b< 21.95702 24   13.9305500  0.93805950  
-#                          7830) x2_dist=even_discrete,uniform 13    2.6412500  0.51209270 *
-#                          7831) x2_dist=lognormal,normal 11    6.1427810  1.44147500 *
-#                     979) e_sigma< 0.2441489 30   16.6036900  0.95445990  
-#                      1958) a< -0.6962934 8    1.2723410  0.16916930 *
-#                      1959) a>=-0.6962934 22    8.6039200  1.24002000 *
-#                 245) n< 202.5 220  194.7300000  0.70128210  
-#                   490) measure_sparse>=0.6256282 81   38.6326300  0.36079080  
-#                     980) measure_monotonic>=0.237124 8    0.2987567 -0.26782100 *
-#                     981) measure_monotonic< 0.237124 73   34.8262100  0.42967980 *
-#                   491) measure_sparse< 0.6256282 139  141.2344000  0.89969790  
-#                     982) n>=97 88   72.2102700  0.72131320  
-#                      1964) x1_sigma>=0.4241299 54   26.0940300  0.51943470  
-#                        3928) e_sigma>=0.2485052 45   19.5310800  0.39828410  
-#                          7856) x1_sigma< 0.5588507 38   11.8880800  0.26476900 *
-#                          7857) x1_sigma>=0.5588507 7    3.2882920  1.12308000 *
-#                        3929) e_sigma< 0.2485052 9    2.6000360  1.12518800 *
-#                      1965) x1_sigma< 0.4241299 34   40.4201400  1.04194400  
-#                        3930) j< 7.5 11    8.6545230  0.32377180 *
-#                        3931) j>=7.5 23   23.3787300  1.38541700  
-#                          7862) b< 20.93247 14   10.8261000  0.95400130 *
-#                          7863) b>=20.93247 9    5.8936770  2.05650900 * (moderate b, large j, small x1_sigma, small n, small sparse, second predictor, non-linearity -> large under-prediction)
-#                     983) n< 97 51   61.3921000  1.20749900  
-#                      1966) e_dist=uniform 16    6.5588030  0.49547860 *
-#                      1967) e_dist=even_discrete,lognormal 35   43.0135900  1.53299400  
-#                        3934) measure_monotonic>=0.04508442 17   19.7133200  0.99850550 *
-#                        3935) measure_monotonic< 0.04508442 18   13.8570300  2.03778800 *
-#               123) measure_splines< 0.04519141 261  221.3718000  0.87722750  
-#                 246) measure_sparse>=0.654522 53   26.6382500  0.36906170  
-#                   492) measure_striped>=0.57 11    2.8015390 -0.20941000 *
-#                   493) measure_striped< 0.57 42   19.1917400  0.52056620 *
-#                 247) measure_sparse< 0.654522 208  177.5598000  1.00671200  
-#                   494) e_sigma>=2.064604 95   37.3057400  0.73978960  
-#                     988) n< 94.5 10    2.7534790  0.01814227 *
-#                     989) n>=94.5 85   28.7318400  0.82468930  
-#                      1978) measure_sparse< 0.3468051 11    2.5461540  0.27928850 *
-#                      1979) measure_sparse>=0.3468051 74   22.4272100  0.90576240 *
-#                   495) e_sigma< 2.064604 113  127.7952000  1.23111600  
-#                     990) measure_splines>=0.03040418 40   30.6303500  0.78917000  
-#                      1980) include_non_normal=TRUE 16    4.6253120  0.29159560 *
-#                      1981) include_non_normal=FALSE 24   19.4029000  1.12088600 *
-#                     991) measure_splines< 0.03040418 73   85.0712900  1.47327800  
-#                      1982) b< 90.99624 66   59.3920200  1.35710600  
-#                        3964) a>=-0.6695046 52   41.5200500  1.19796400  
-#                          7928) x2_sigma< 0.4287885 20   11.8256700  0.80120150  
-#                           15856) measure_splines>=0.009178431 7    2.5265790  0.21136300 *
-#                           15857) measure_splines< 0.009178431 13    5.5523770  1.11880700 *
-#                          7929) x2_sigma>=0.4287885 32   24.5782200  1.44594000  
-#                           15858) x2_sigma>=0.4851646 24   12.4345200  1.18424600  
-#                             31716) e_dist=uniform 7    0.5259639  0.41482310 *
-#                             31717) e_dist=even_discrete,lognormal 17    6.0580960  1.50106700 *
-#                           15859) x2_sigma< 0.4851646 8    5.5692240  2.23102400 * (small x2_sigma, small b, small splines dependence, small e_sigma, small sparse, second predictor, non-linearity -> large under-prediction)
-#                        3965) a< -0.6695046 14   11.6634200  1.94820600 *
-#                      1983) b>=90.99624 7   16.3901800  2.56861600 * (large b, small splines dependence, small e_sigma, small sparse, second predictor, non-linearity -> large under-prediction)
-#            31) e_sigma< 0.1149429 48  135.8347000  1.35422700  
-#              62) measure_striped< 0.44 40   54.1808300  0.92957180  
-#               124) include_non_normal=TRUE 21   14.0872600  0.39910760  
-#                 248) include_heter=TRUE 13    5.7917480  0.06683663 *
-#                 249) include_heter=FALSE 8    4.5279800  0.93904800 *
-#               125) include_non_normal=FALSE 19   27.6530600  1.51587400 *
-#              63) measure_striped>=0.44 8   38.3742300  3.47750200 * (high discreteness, small e_sigma, second predictor, non-linearity -> large under-prediction)
-
-model_pred %>%
-  filter(res == 32L) %>%
-  filter(type == "test") %>%
-  mutate(.resid = effect_size - vss) %>%
-  mutate(include_z = factor(include_z),
-         include_heter = factor(include_heter),
-         include_x2 = factor(include_x2),
-         include_non_normal = factor(include_non_normal),
-         e_even = factor(e_even),
-         x1_even = factor(x1_even),
-         x2_even = factor(x2_even)) %>%
-  select(-filename, -vss, -res, -plot_uid, -data_type, -effect_size, -type) %>%
-  rpart::rpart(.resid ~ ., data = ., control = rpart::rpart.control(cp = 0.001), model = TRUE) -> test_tree
-```
-
-```{r performance-sub}
-
-model_pred %>%
-  filter(res == 32L) %>%
-  group_by(type, include_non_normal, include_heter, include_z) %>%
-  summarise(n = n(), 
-            RMSE = yardstick::rmse_vec(effect_size, vss),
-            R2 = yardstick::rsq_vec(effect_size, vss),
-            MAE = yardstick::mae_vec(effect_size, vss),
-            HUBER = yardstick::huber_loss_vec(effect_size, vss)) %>%
-  mutate(violations = ifelse(include_z, "non-linearity", "null")) %>%
-  mutate(violations = ifelse(include_heter, glue("{violations} + heteroskedasticity"), violations)) %>%
-  mutate(violations = ifelse(include_non_normal, glue("{violations} + non-normality"), violations)) %>%
-  mutate(violations = gsub("null \\+ ", "", violations)) %>%
-  ungroup() %>%
-  select(type, violations, n, RMSE, R2, MAE, HUBER) %>%
-  arrange(desc(type)) %>%
-  select(-type) %>%
-  mutate(across(RMSE:MAE, ~format(.x, digits = 3))) %>%
-  mutate(across(HUBER, ~format(.x, digits = 2))) %>%
-  mutate(violations = ifelse(violations == "null", "no violations", violations)) %>%
-  mutate(R2 = gsub(".*NA.*", "", R2)) %>%
-  kableExtra::kable("latex", 
-                    booktabs = TRUE,
-                    escape = FALSE,
-                    col.names = c("Violations", "\\#samples", "RMSE", "$R^2$", "MAE", "Huber loss"),
-                    caption = "The training and test performance of the $32 \\times 32$ model presented with different model violations.  The best metrics are colored in red.") %>%
-  kableExtra::kable_styling(latex_options = c("scale_down")) %>%
-  kableExtra::pack_rows("Training set", 1, 8) %>%
-  kableExtra::pack_rows("Test set", 9, 16) %>%
-  kableExtra::column_spec(3, color = c(rep("black", 4), "red", rep("black", 3),
-                                       rep("black", 4), "red", rep("black", 3))) %>%
-  kableExtra::column_spec(4, color = c(rep("black", 4), "red", rep("black", 3),
-                                       rep("black", 4), "red", rep("black", 3))) %>%
-  kableExtra::column_spec(5, color = c(rep("black", 4), "red", rep("black", 3),
-                                       rep("black", 4), "red", rep("black", 3))) %>%
-  kableExtra::column_spec(6, color = c(rep("black", 4), "red", rep("black", 3),
-                                       rep("black", 4), "red", rep("black", 3)))
-```
-
-## Comparison with human visual inference
-
-### Overview of the human subject experiment
-
-In order to check the validity of the proposed computer vision model, residual plots presented in the human subject experiment conducted by @li2023plot will be assessed.
-This study has collected $7974$ human responses to $1152$ lineups.
-Each lineup contains one randomly placed actual residual plot and 19 null plots.
-Among the $1152$ lineups, $24$ are attention check lineups in which the visual patterns are designed to be extremely obvious and very different from the corresponding to null plots, $36$ are null lineups where all the lineups consist of only null plots, $279$ are lineups with uniform predictor distribution evaluated by $11$ participants, and the remaining $813$ are lineups with discrete, skewed or normal predictor distribution evaluated by $5$ participants.
-Attention check lineups and null lineups will not be assessed in the following analysis.
-
-In @li2023plot, the residual plots are simulated from a data generating process which is a special case of Equation \ref{eq:data-sim}.
-The main characteristic is the model violations are introduced separately, meaning non-linearity and heteroskedasticity will not co-exist in one lineup but assigned uniformly to all lineups.
-Additionally, non-normality and multiple predictors are not considered in their experimental design.
-
-### Model performance on the human data
-
-For each lineup used in @li2023plot, there are one actual residual plot and $19$ null plots.
-While the distance $D$ for the actual residual plot depends on the underlying data generating process, the distance $D$ for the null plots are zeros.
-We have used our optimized computer vision model to predict $\hat{D}$ for both the actual residual plots and the null plots.
-
-The performance metrics of $\hat{D}$ for actual residual plots are provided in Table \ref{tab:experiment-performance}.
-It can be observed that all the performance metrics are slightly worse than the metrics evaluated on the test data.
-Still, the mean absolute error remains at a low level, and the linearly correlation between the prediction and the true value is still very high.
-Lineups with non-linearity issues are more difficult to be predicted than those with heteroskedasticity issues.
-
-For null plots, a histogram of the prediction is provided in Figure \ref{fig:hist-null-human}.
-In most cases, $\hat{D}$ will be predicted at around $1$.
-And as $\hat{D}$ increases, the number of predictions decreases.
-The 90%, 95% and 99% sample quantiles are at around
-
-```{r rd-human}
-vss_32 <- readRDS(here::here("paper/data/vss_32.rds"))
-
-experiment <- vi_survey %>%
-  group_by(unique_lineup_id) %>%
-  summarise(across(everything(), first)) %>%
-  select(unique_lineup_id, attention_check, null_lineup, prop_detect,
-         answer, effect_size, conventional_p_value,
-         p_value, type, shape, a, b, x_dist, e_dist,
-         e_sigma, include_z, k, n)
-
-experiment <- vss_32 %>%
-  left_join(experiment) %>%
-  mutate(conventional_reject = conventional_p_value <= 0.05) %>%
-  mutate(reject = p_value <= 0.05) %>%
-  mutate(model_reject = vss_p_value <= 0.05) %>%
-  mutate(model_boot_reject = vss_boot_p_value <= 0.05)
-
-experiment <- experiment %>%
-  left_join(readRDS(here::here("paper/data/actual_ss.rds")))
-```
-
-```{r experiment-performance}
-experiment %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  group_by(type) %>%
-  summarise(RMSE = yardstick::rmse_vec(actual_ss, vss),
-            R2 = yardstick::rsq_vec(actual_ss, vss),
-            MAE = yardstick::mae_vec(actual_ss, vss),
-            HUBER = yardstick::huber_loss_vec(actual_ss, vss)) %>%
-  mutate(across(RMSE:HUBER, ~format(.x, digits = 3))) %>%
-  kableExtra::kable(format = "latex", 
-                    booktabs = TRUE, 
-                    escape = FALSE,
-                    col.names = c("Voilation", "RMSE", "$R^2$", "MAE", "Huber loss"),
-                    caption = "The performance of the $32 \\times 32$ model on the data used in the human subject experiment.")
-```
-
-```{r hist-null-human, fig.cap = "Histogram of predicted distance on null plots used in the human subject experiment. Sample quantiles are drawn in different colors."}
-lineup_vss <- readRDS(here::here("paper/data/lineup_vss.rds"))
-lineup_vss <- lineup_vss %>%
-  group_by(unique_lineup_id) %>%
-  mutate(delta_diff = vss[!null] - max(vss[null])) %>%
-  mutate(gamma_diff = sum(vss[null] > vss[!null]))
-
-lineup_vss %>%
-  filter(null) %>%
-  left_join(vi_survey %>%
-              group_by(unique_lineup_id) %>% 
-              summarise(type = first(type), 
-                        attention_check = first(attention_check),
-                        null_lineup = first(null_lineup))) %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  ggplot() +
-  geom_histogram(aes(vss)) +
-  geom_vline(aes(xintercept = quantile(vss, 0.9), col = "Q(0.90)")) +
-  geom_vline(aes(xintercept = quantile(vss, 0.95), col = "Q(0.95)")) +
-  geom_vline(aes(xintercept = quantile(vss, 0.99), col = "Q(0.99)")) +
-  xlab(expression(hat(D))) +
-  theme_light() +
-  facet_wrap(~type) +
-  labs(col = "")
-```
-
-```{r cache = TRUE}
-vi_lineup <- get_vi_lineup()
-```
-
-```{r fig.cap = "A weighted detection rate vs $\\delta$-differnence plot."}
-
-experiment %>%
-  left_join(lineup_vss %>%
-              group_by(unique_lineup_id) %>%
-              summarise(delta_diff = first(delta_diff), gamma_diff = first(gamma_diff))) %>%
-  ggplot() +
-  geom_point(aes(delta_diff, prop_detect)) +
-  geom_smooth(aes(delta_diff, prop_detect), se = FALSE) +
-  ylab("Weighted detection rate") +
-  xlab(expression(gamma-difference)) +
-  theme_light()
-```
-
-```{r fig.cap = "A weighted detection rate vs $\\gamma$-number of extreme nulls plot."}
-experiment %>%
-  left_join(lineup_vss %>%
-              group_by(unique_lineup_id) %>%
-              summarise(delta_diff = first(delta_diff), gamma_diff = first(gamma_diff))) %>%
-  ggplot() +
-  geom_point(aes(gamma_diff, prop_detect)) +
-  ylab("Weighted detection rate") +
-  xlab(expression(gamma-number~of~extreme~nulls)) +
-  theme_light()
-```
-
-```{r fig.cap = "A weighted detection rate vs $\\delta$-differnence plot for lineups evaluated by 11 people."}
-experiment %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  filter(x_dist == "uniform") %>%
-  left_join(lineup_vss %>%
-              group_by(unique_lineup_id) %>%
-              summarise(delta_diff = first(delta_diff), gamma_diff = first(gamma_diff))) %>%
-  ggplot() +
-  geom_point(aes(delta_diff, prop_detect)) +
-  geom_smooth(aes(delta_diff, prop_detect), se = FALSE) +
-  ylab("Weighted detection rate") +
-  xlab(expression(delta-difference)) +
-  theme_light()
-```
-
-```{r fig.cap = "A weighted detection rate vs $\\gamma$-number of extreme nulls plot for lineups evaluated by 11 people."}
-experiment %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  filter(x_dist == "uniform") %>%
-  left_join(lineup_vss %>%
-              group_by(unique_lineup_id) %>%
-              summarise(delta_diff = first(delta_diff), gamma_diff = first(gamma_diff))) %>%
-  ggplot() +
-  geom_point(aes(gamma_diff, prop_detect)) +
-  ylab("Weighted detection rate") +
-  xlab(expression(gamma-number~of~extreme~nulls)) +
-  theme_light()
-```
-
-## When the model works
-
-- simple examples (non-linearity, heteroskedasticity, ...)
-- datasaurus
-
-## When the model does not work
-
-- human detect but model does not
-- cartoon residuals?
-
-## Workflow: how one use this model? (small showcase)
-
-# Dicussion
-
-There are other kinds of residual departures like autocorrelation that are not considered in this study.
-The primary goal of this study is to establish a new way of evaluating residual plot and conducting visual test with computer vision models.
-Building computer vision models for other model violations and other types of diagnostic plots could be future directions of this field.
-
-# Conclusion
-
-- Summary of findings
-- Contributions to the field
-- Future directions for research
-
-# Acknowledgements

### 2024-04-15 paper/paper.Rmd
update paper

@@
 
-### Approximated distance as an Model Violations Index (MVI)
+### Model Violations Index (MVI) {#sec:model-violations-index}
 
@@
 
-Note that since the Kullback-Leibler divergence could be affected by the number of random variables involved in the evaluation.
-In general, more observations will result in greater distance $D$.
-This does not mean $\hat{D}$ can not correctly represent the extent of the model violations since more observations in a residual plot usually lead to stronger visual signal strength as the underlying visual patterns will have higher chance to be revealed except for cases like serious overlapping.
-For the same data generating process, $D$ usually grows logarithmically as the number of observations increases.
-This is because $D = \text{log}(1 + D_{KL})$, and $D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}$ under the independence assumption.
-
 Note that the Kullback-Leibler divergence might be influenced by the number of random variables involved in its evaluation.
@@
 
-### Null distribution of the approximated distance
+### Lineup evaluation
 
@@
 
-To solve this issue while aligning with the principle of visual inference, $\hat{D}$ can be viewed as a test statistic.
-And the null distribution of this statistic can be approximated by the empirical distribution of $\hat{D}_{null}$, where $\hat{D}_{null}$ is the approximated distance for a null plot simulated from the fitted regression model.
-The approximation of the distribution involves applying the residual rotation technique [@buja2009statistical] on the fitted regression model to obtain null residuals.
-The null residuals are then used to make null plots and fed into the computer vision model to get predictions.
-The empirical distribution is constructed with the approximated distance.
-
-There are two types of error associated with the approximation of the null distribution.
-These include the sampling error and the estimation error for model parameters.
-Increasing the number of null plots within the distribution approximation can effectively reduce sampling error.
-However, the estimation error for model parameters remains irreducible.
-The residual rotation technique operates under the assumption of a correct fitted model, using the estimated variance instead of the unknown true variance to rescale rotated residuals.
-Consequently, under $H_0$, the true distribution of $\hat{D}_{null}$ will slightly deviate from the approximated null distribution.
-
-### Estimation of quantiles of the null distribution
-
-Let $\hat{D}_{null}^{(i)}$ be the approximated distance of the $i$-th null plots, where $i = 1,...,n_{null}$ and $n_{null} \in \mathbb{N}^+$ is a sufficiently large number.
-Quantiles of the null distribution can then be estimated using the sample quantiles available in statistical software such as R.
+To address this issue while adhering to the principle of visual inference, we can compare the estimated distance $\hat{D}$ to the approximated distances for the null plots in a lineup.
+Specifically, if a lineup comprises 20 plots, the null hypothesis $H_0$ will be rejected if $\hat{D}$ exceeds the maximum approximated distance among the $m - 1$ null plots, denoted as $\max\limits_{1 \leq i \leq m-1} {\hat{D}_{null}^{(i)}}$, where $\hat{D}_{null}^{(i)}$ represents the approximated distance for the $i$-th null plot.
+This approach is equivalent to the typical lineup protocol requiring a 95% significance level, where $H_0$ is rejected if the data plot is identified as the most distinct plot by the sole participant.
+The approximated distance serves as a metric to quantify the difference between the data plot and the null plots, as intended.
+
+For lineups consisting of more than 20 plots, the 95% significance level can be maintained if the number of plots is a multiple of 20.
+Specifically, for lineups comprising $20p$ plots, where $p$ is a positive integer, we reject $H_0$ if $\hat{D}$ exceeds 95% ${\hat{D}_{null}^{(i)}}$ for $i = 1, \ldots, 20p-1$.
+The p-value in this case is given by $\frac{1}{m} + \frac{1}{m}\sum{i=1}^{m-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)$, where $I(\cdot)$ is the indicator function.
+
+Moreover, if the number of plots in a lineup, denoted by $m$, is sufficiently large, the empirical distribution of ${\hat{D}_{null}^{(i)}}$ can be viewed as an approximation of the null distribution of the approximated distance.
+Consequently, quantiles of the null distribution can be estimated using the sample quantiles available in statistical software such as R, and these quantiles can be utilized for decision-making purposes.
 The details of the sample quantile computation can be found in @hyndman1996sample.
-In statistical testing, analysts often care about certain quantiles of the null distribution, such as the 90% quantile, the 95% quantile and the 99% quantile.
-These quantiles are used as critical values to decide if $H_0$ needs to be rejected.
-For example, if $\hat{D}$ is greater than and equal to the 95% sample quantile $Q_{null}(0.95)$, we could say the approximated distance for the actual residual plot is significantly different from the approximated distance for null plots with 95% significance level.
-Based on our experience, in order to get a stable estimate of the 95% quantile, $n_{null}$ usually needs to be at least $100$.
-And if the null distribution has a long tail, more null plots will be needed.
-Alternatively, a p-value can be used to represents the probability of observing an event equally or more extreme than the given event under $H_0$, and it can be estimated by $1/m\sum_{i=1}^{n_{null}}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)$.
+For instance, if $\hat{D}$ is greater than or equal to the 95% sample quantile, denoted as $Q_{null}(0.95)$, we can conclude that the approximated distance for the actual residual plot is significantly different from the approximated distance for null plots with a 95% significance level.
+Based on our experience, to obtain a stable estimate of the 95% quantile, the number of null plots, $n_{null}$, typically needs to be at least 100.
+However, if the null distribution exhibits a long tail, a larger number of null plots may be required.
+Alternatively, a p-value can be used to represent the probability of observing an event equally or more extreme than the given event under the null hypothesis $H_0$, and it can be estimated by $\frac{1}{m}\sum_{i=1}^{m-1}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)$.
 
@@
 All optimized models chose to retain the additional inputs, contributing to the reduction of validation error.
-The number of units required for the fully-connected layer was $256$, a relatively modest number compared to the VGG16 classifier, suggesting that the problem at hand was relatively straightforward.
-The optimized learning rates were typically smaller than the default value recommended by Keras, which is 0.001.
+The number of units required for the fully-connected layer was $256$, a relatively modest number compared to the VGG16 classifier, suggesting that the problem at hand was less complex.
+The optimized learning rates were smaller than the default value recommended by Keras, which is 0.001.
 
@@
 
-## Model performance
+## Summary of model performance
 
@@
 Conversely, when the error distribution had a large variance, all models tended to over-predict the distance.
-Therefore, Additionally, for input images representing null plots, it was expected that the models will over-predict the distance, as explained in Section \ref{null-distribution-of-the-approximated-distance}.
+Furthermore, for input images representing null plots, it was expected that the models will over-predict the distance, as explained in Section \ref{lineup-evaluation}.
 
@@
 It was evident that metrics for null plots were notably worse compared to other categories.
-Furthermore, residual plots solely exhibiting non-normality issues were the easiest to predict, with very low test root mean square error (RMSE) around $0.3$, very low mean absolute error (MAE) around $0.2$, and very high $R^2$ around $0.94$.
+Furthermore, residual plots solely exhibiting non-normality issues were the easiest to predict, with very low test root mean square error (RMSE) at around $0.3$.
 Residual plots with non-linearity issues were more challenging to assess than those with heteroskedasticity or non-normality issues.
@@
 
-Based on the model performance metrics, we will only use the best-performing model evaluated on the test set, namely the $32 \times 32$ model, for the subsequent analysis.
+Based on the model performance metrics, we chose to use the best-performing model evaluated on the test set, namely the $32 \times 32$ model, for the subsequent analysis.
 
@@
                     col.names = c("", "RMSE", "$R^2$", "MAE", "Huber loss"),
-                    caption = "The training and test performance of three optimized models with different input sizes. The best metrics are colored in red.") %>%
+                    caption = "The training and test performance of three optimized models with different input sizes.") %>%
   kableExtra::pack_rows("Training set", 1, 3) %>%
-  kableExtra::pack_rows("Test set", 4, 6) %>%
-  kableExtra::column_spec(2:5, color = c("black", "red", "black",
-                                       "red", "black", "black"))
-
+  kableExtra::pack_rows("Test set", 4, 6)
 ```
@@
 
-```{r eval = FALSE}
-
- # 1) root 8000 3466.12800  0.04808905  
- #   2) e_sigma>=4.108756 3288 1146.50300 -0.09147268  
- #     4) include_non_normal=FALSE 1983  817.30850 -0.22151770  
- #       8) include_heter=FALSE 1210  416.16970 -0.39866830  
- #        16) include_z=FALSE 49   20.28332 -1.22777500 * (only non-normal and large error sigma -> large over-prediction)
- #        17) include_z=TRUE 1161  360.78130 -0.36367590 *
- #       9) include_heter=TRUE 773  303.72640  0.05578156 *
- #     5) include_non_normal=TRUE 1305  244.69960  0.10613590 *
- #   3) e_sigma< 4.108756 4712 2210.89500  0.14547420  
- #     6) include_z=FALSE 2030  568.99490 -0.04100357 *
- #     7) include_z=TRUE 2682 1517.87900  0.28661890  
- #      14) include_x2=FALSE 1432  377.35510  0.07228115 *
- #      15) include_x2=TRUE 1250  999.37090  0.53216420  
- #        30) e_sigma>=0.1149429 1202  829.80300  0.49933640 *
- #        31) e_sigma< 0.1149429 48  135.83470  1.35422700  (non-linearity, second predictor, small error sigma)
- #          62) measure_striped< 0.44 40   54.18083  0.92957180 *
- #          63) measure_striped>=0.44 8   38.37423  3.47750200 * (high discreteness -> large under-prediction)
-
-# n= 8000 
-# 
-# node), split, n, deviance, yval
-#       * denotes terminal node
-# 
-#     1) root 8000 3466.1280000  0.04808905  
-#       2) e_sigma>=4.108756 3288 1146.5030000 -0.09147268  
-#         4) include_non_normal=FALSE 1983  817.3085000 -0.22151770  
-#           8) include_heter=FALSE 1210  416.1697000 -0.39866830  
-#            16) include_z=FALSE 49   20.2833200 -1.22777500  
-#              32) measure_sparse< 0.3630358 7    8.9396360 -2.12605300 * (large e_sigma, null plots, not very sparse -> large over-prediction)
-#              33) measure_sparse>=0.3630358 42    4.7539680 -1.07806200 * (large e_sigma, null plots, sparse -> large over-prediction)
-#            17) include_z=TRUE 1161  360.7813000 -0.36367590  
-#              34) measure_splines>=0.02606246 376  134.3999000 -0.59957120  
-#                68) n< 143.5 116   39.6936700 -0.84719690  
-#                 136) measure_monotonic>=0.07456592 7    1.3762680 -1.55893700 * (non-linearity only, small n, large e_sigma -> large over-prediction)
-#                 137) measure_monotonic< 0.07456592 109   34.5436600 -0.80148880 *
-#                69) n>=143.5 260   84.4197800 -0.48909200  
-#                 138) e_sigma>=7.367384 63   15.0193100 -0.79693070 *
-#                 139) e_sigma< 7.367384 197   61.5210600 -0.39064610  
-#                   278) include_x2=FALSE 90   19.5837000 -0.53918530 *
-#                   279) include_x2=TRUE 107   38.2813500 -0.26570670  
-#                     558) measure_sparse< 0.5586777 84   27.3861700 -0.37645850 *
-#                     559) measure_sparse>=0.5586777 23    6.1018500  0.13877850 *
-#              35) measure_splines< 0.02606246 785  195.4365000 -0.25068660  
-#                70) e_sigma>=5.64113 529  100.6972000 -0.37916600  
-#                 140) n< 109 41   11.8877000 -0.73213180 *
-#                 141) n>=109 488   83.2723700 -0.34951110  
-#                   282) e_sigma>=7.90547 154   26.5350800 -0.48055150 *
-#                   283) e_sigma< 7.90547 334   52.8735800 -0.28909130 *
-#                71) e_sigma< 5.64113 256   67.9628700  0.01480427  
-#                 142) include_x2=FALSE 113   23.8143300 -0.15341150 *
-#                 143) include_x2=TRUE 143   38.4243300  0.14773000  
-#                   286) n< 72.5 7    2.6335630 -0.69382120 *
-#                   287) n>=72.5 136   30.5781400  0.19104510 *
-#           9) include_heter=TRUE 773  303.7264000  0.05578156  
-#            18) b< 32.23487 487  221.2986000 -0.09202344  
-#              36) b< 0.2107948 67   26.8944500 -0.51926710  
-#                72) n< 126 7    3.3765570 -1.24628700 * (large e_sigma, heteroskedasticity, small b, small n -> over-prediction)
-#                73) n>=126 60   19.3863300 -0.43444810 *
-#              37) b>=0.2107948 420  180.2232000 -0.02386791  
-#                74) measure_splines>=0.05482434 223   77.7820400 -0.18200180  
-#                 148) b< 7.647324 119   51.8788300 -0.32137070  
-#                   296) measure_sparse>=0.4117995 111   43.0150500 -0.38433220 *
-#                   297) measure_sparse< 0.4117995 8    2.3184780  0.55221930 *
-#                 149) b>=7.647324 104   20.9469900 -0.02253160 *
-#                75) measure_splines< 0.05482434 197   90.5523700  0.15513650  
-#                 150) b>=5.165768 63   26.5340300 -0.02749363  
-#                   300) include_x2=FALSE 35    7.9867080 -0.19247040 *
-#                   301) include_x2=TRUE 28   16.4039600  0.17872730  
-#                     602) measure_monotonic< 0.03127537 21    4.3016860 -0.07172379 *
-#                     603) measure_monotonic>=0.03127537 7    6.8333090  0.93008060 *
-#                 151) b< 5.165768 134   60.9291300  0.24099990  
-#                   302) x2_sigma< 0.5865284 126   52.7919200  0.19154170  
-#                     604) x2_k>=5.5 106   43.2640200  0.11519510  
-#                      1208) j< 17.5 93   38.5219100  0.05750563  
-#                        2416) e_even=TRUE 41   15.2536200 -0.17697420  
-#                          4832) measure_sparse>=0.5392181 29    8.8138120 -0.40043160 *
-#                          4833) measure_sparse< 0.5392181 12    1.4922590  0.36304760 *
-#                        2417) e_even=FALSE 52   19.2367200  0.24238400  
-#                          4834) measure_splines>=0.01044483 31    7.8888290 -0.02089042 *
-#                          4835) measure_splines< 0.01044483 21    6.0272570  0.63102720 *
-#                      1209) j>=17.5 13    2.2184070  0.52789690 *
-#                     605) x2_k< 5.5 20    5.6354220  0.59617870 *
-#                   303) x2_sigma>=0.5865284 8    2.9747040  1.01996600 * (large e_sigma, small spline dependence, heteroskedasticity, large x2_sigma -> large under-prediction)
-#            19) b>=32.23487 286   53.6723400  0.30746350  
-#              38) include_x2=FALSE 139   17.0423600  0.13425010 *
-#              39) include_x2=TRUE 147   28.5161400  0.47125030 *
-#         5) include_non_normal=TRUE 1305  244.6996000  0.10613590  
-#          10) measure_splines>=0.08262098 455   47.2275500  0.00622856 *
-#          11) measure_splines< 0.08262098 850  190.4994000  0.15961570  
-#            22) include_heter=FALSE 594   22.8029300  0.08838938 *
-#            23) include_heter=TRUE 256  157.6908000  0.32488300  
-#              46) measure_striped>=0.47 185   31.8984100  0.17734340 *
-#              47) measure_striped< 0.47 71  111.2723000  0.70931710  
-#                94) e_even=FALSE 32   18.0084200  0.22023240  
-#                 188) a>=-0.205314 21    5.0433870 -0.06070074 *
-#                 189) a< -0.205314 11    8.1435320  0.75655920 *
-#                95) e_even=TRUE 39   79.3286800  1.11061700  
-#                 190) b< 2.373839 11    0.2423938  0.06215229 *
-#                 191) b>=2.373839 28   62.2437600  1.52251400  (discrete e_dist,  high discreteness, heteroskedasticity, small splines dependence, non-normality, large e_sigma -> large under-prediction)
-#                   382) x2_dist=even_discrete,lognormal,uniform 19   33.0782400  1.01336800 *
-#                   383) x2_dist=normal 9   13.8421700  2.59737800 *
-#       3) e_sigma< 4.108756 4712 2210.8950000  0.14547420  
-#         6) include_z=FALSE 2030  568.9949000 -0.04100357  
-#          12) b< 28.58967 1159  316.0764000 -0.12131010  
-#            24) include_non_normal=FALSE 441  185.6341000 -0.33971870  
-#              48) include_heter=FALSE 60    8.3705760 -1.15981800 * (null plots, small e_sigma -> large over-prediction)
-#              49) include_heter=TRUE 381  130.5548000 -0.21056910  
-#                98) b< 0.4619795 85   22.8319800 -0.60512210 *
-#                99) b>=0.4619795 296   90.6909300 -0.09726844  
-#                 198) b>=1.080411 252   72.3919200 -0.15604160  
-#                   396) b< 2.878015 65   20.2765800 -0.37135770 *
-#                   397) b>=2.878015 187   48.0544100 -0.08119905 *
-#                 199) b< 1.080411 44   12.4430700  0.23934130 *
-#            25) include_non_normal=TRUE 718   96.4846700  0.01283772 *
-#          13) b>=28.58967 871  235.4979000  0.06585671  
-#            26) include_heter=FALSE 361  111.6977000 -0.10528680  
-#              52) include_non_normal=FALSE 46    6.2078660 -1.13547200 * (null plots, small e_sigma -> large over-prediction)
-#              53) include_non_normal=TRUE 315   49.5417500  0.04515286  
-#               106) e_sigma>=0.2273127 247   17.7293100 -0.01172324 *
-#               107) e_sigma< 0.2273127 68   28.1111100  0.25174690  
-#                 214) measure_splines>=0.03710735 23   10.7097800 -0.09471551 *
-#                 215) measure_splines< 0.03710735 45   13.2294000  0.42882770 *
-#            27) include_heter=TRUE 510  105.7420000  0.18699950  
-#              54) include_non_normal=TRUE 235   41.6282300  0.05417806 *
-#              55) include_non_normal=FALSE 275   56.4252500  0.30050140  
-#               110) include_x2=FALSE 137   20.9155600  0.17983570 *
-#               111) include_x2=TRUE 138   31.5346400  0.42029280  
-#                 222) b< 70.08912 96   17.2702500  0.30834030 *
-#                 223) b>=70.08912 42   10.3110100  0.67618420 *
-#         7) include_z=TRUE 2682 1517.8790000  0.28661890  
-#          14) include_x2=FALSE 1432  377.3551000  0.07228115  
-#            28) measure_splines>=0.07058104 731  178.2000000 -0.01483915  
-#              56) e_sigma>=2.44726 138   40.9281500 -0.18500620  
-#               112) include_heter=FALSE 59   25.9010700 -0.41304000  
-#                 224) include_non_normal=FALSE 38   17.2724700 -0.68141390  
-#                   448) measure_splines>=0.1003984 20    5.6757540 -1.07654300 * (non-linearity only, moderate e_sigma -> large over-prediction)
-#                   449) measure_splines< 0.1003984 18    5.0046810 -0.24238140 *
-#                 225) include_non_normal=TRUE 21    0.9391262  0.07258882 *
-#               113) include_heter=TRUE 79    9.6678490 -0.01470237 *
-#              57) e_sigma< 2.44726 593  132.3458000  0.02476126  
-#               114) e_sigma>=0.158351 527  105.6393000 -0.00147316 *
-#               115) e_sigma< 0.158351 66   23.4476700  0.23423920  
-#                 230) b< 71.04535 55   15.9627800  0.12578350 *
-#                 231) b>=71.04535 11    3.6032230  0.77651720 *
-#            29) measure_splines< 0.07058104 701  187.8212000  0.16312990  
-#              58) n< 62.5 33   14.3260500 -0.31035730 *
-#              59) n>=62.5 668  165.7314000  0.18652070 *
-#          15) include_x2=TRUE 1250  999.3709000  0.53216420  
-#            30) e_sigma>=0.1149429 1202  829.8030000  0.49933640  
-#              60) e_sigma>=2.906419 328  129.1838000  0.27288400  
-#               120) measure_splines>=0.04413018 134   50.5691600  0.09851024  
-#                 240) b< 2.910823 23   10.1825400 -0.20213530  
-#                   480) x2_k>=6.5 15    4.3945260 -0.49662390 *
-#                   481) x2_k< 6.5 8    2.0480610  0.35003080 *
-#                 241) b>=2.910823 111   37.8769400  0.16080620  
-#                   482) b>=7.320158 97   22.8753700  0.08661750 *
-#                   483) b< 7.320158 14   10.7686400  0.67482760 *
-#               121) measure_splines< 0.04413018 194   71.7259100  0.39332780  
-#                 242) j< 7.5 60   32.0807400  0.19368700  
-#                   484) x2_sigma< 0.343438 10   12.3504400 -0.51629950 *
-#                   485) x2_sigma>=0.343438 50   13.6813400  0.33568430 *
-#                 243) j>=7.5 134   36.1830100  0.48271920 *
-#              61) e_sigma< 2.906419 874  677.4868000  0.58432080  
-#               122) measure_splines>=0.04519141 613  424.1886000  0.45960850  
-#                 244) n>=202.5 393  209.4163000  0.32432050  
-#                   488) include_non_normal=TRUE 188   90.6250100  0.18780810  
-#                     976) x2_k< 7.5 101   41.5156400  0.04431936 *
-#                     977) x2_k>=7.5 87   44.6157600  0.35438700  
-#                      1954) measure_sparse>=0.391553 80   35.8617500  0.29057300 *
-#                      1955) measure_sparse< 0.391553 7    4.7050420  1.08369000 * (not sparse, non-normality, large n, small e_sigma, second predictor, non-linearity -> large under-prediction)
-#                   489) include_non_normal=FALSE 205  112.0748000  0.44951240  
-#                     978) e_sigma>=0.2441489 175   86.5106600  0.36295000  
-#                      1956) measure_splines>=0.07729252 127   55.6660800  0.25807550 *
-#                      1957) measure_splines< 0.07729252 48   25.7519700  0.64043040  
-#                        3914) b>=21.95702 24    7.5694370  0.34280130 *
-#                        3915) b< 21.95702 24   13.9305500  0.93805950  
-#                          7830) x2_dist=even_discrete,uniform 13    2.6412500  0.51209270 *
-#                          7831) x2_dist=lognormal,normal 11    6.1427810  1.44147500 *
-#                     979) e_sigma< 0.2441489 30   16.6036900  0.95445990  
-#                      1958) a< -0.6962934 8    1.2723410  0.16916930 *
-#                      1959) a>=-0.6962934 22    8.6039200  1.24002000 *
-#                 245) n< 202.5 220  194.7300000  0.70128210  
-#                   490) measure_sparse>=0.6256282 81   38.6326300  0.36079080  
-#                     980) measure_monotonic>=0.237124 8    0.2987567 -0.26782100 *
-#                     981) measure_monotonic< 0.237124 73   34.8262100  0.42967980 *
-#                   491) measure_sparse< 0.6256282 139  141.2344000  0.89969790  
-#                     982) n>=97 88   72.2102700  0.72131320  
-#                      1964) x1_sigma>=0.4241299 54   26.0940300  0.51943470  
-#                        3928) e_sigma>=0.2485052 45   19.5310800  0.39828410  
-#                          7856) x1_sigma< 0.5588507 38   11.8880800  0.26476900 *
-#                          7857) x1_sigma>=0.5588507 7    3.2882920  1.12308000 *
-#                        3929) e_sigma< 0.2485052 9    2.6000360  1.12518800 *
-#                      1965) x1_sigma< 0.4241299 34   40.4201400  1.04194400  
-#                        3930) j< 7.5 11    8.6545230  0.32377180 *
-#                        3931) j>=7.5 23   23.3787300  1.38541700  
-#                          7862) b< 20.93247 14   10.8261000  0.95400130 *
-#                          7863) b>=20.93247 9    5.8936770  2.05650900 * (moderate b, large j, small x1_sigma, small n, small sparse, second predictor, non-linearity -> large under-prediction)
-#                     983) n< 97 51   61.3921000  1.20749900  
-#                      1966) e_dist=uniform 16    6.5588030  0.49547860 *
-#                      1967) e_dist=even_discrete,lognormal 35   43.0135900  1.53299400  
-#                        3934) measure_monotonic>=0.04508442 17   19.7133200  0.99850550 *
-#                        3935) measure_monotonic< 0.04508442 18   13.8570300  2.03778800 *
-#               123) measure_splines< 0.04519141 261  221.3718000  0.87722750  
-#                 246) measure_sparse>=0.654522 53   26.6382500  0.36906170  
-#                   492) measure_striped>=0.57 11    2.8015390 -0.20941000 *
-#                   493) measure_striped< 0.57 42   19.1917400  0.52056620 *
-#                 247) measure_sparse< 0.654522 208  177.5598000  1.00671200  
-#                   494) e_sigma>=2.064604 95   37.3057400  0.73978960  
-#                     988) n< 94.5 10    2.7534790  0.01814227 *
-#                     989) n>=94.5 85   28.7318400  0.82468930  
-#                      1978) measure_sparse< 0.3468051 11    2.5461540  0.27928850 *
-#                      1979) measure_sparse>=0.3468051 74   22.4272100  0.90576240 *
-#                   495) e_sigma< 2.064604 113  127.7952000  1.23111600  
-#                     990) measure_splines>=0.03040418 40   30.6303500  0.78917000  
-#                      1980) include_non_normal=TRUE 16    4.6253120  0.29159560 *
-#                      1981) include_non_normal=FALSE 24   19.4029000  1.12088600 *
-#                     991) measure_splines< 0.03040418 73   85.0712900  1.47327800  
-#                      1982) b< 90.99624 66   59.3920200  1.35710600  
-#                        3964) a>=-0.6695046 52   41.5200500  1.19796400  
-#                          7928) x2_sigma< 0.4287885 20   11.8256700  0.80120150  
-#                           15856) measure_splines>=0.009178431 7    2.5265790  0.21136300 *
-#                           15857) measure_splines< 0.009178431 13    5.5523770  1.11880700 *
-#                          7929) x2_sigma>=0.4287885 32   24.5782200  1.44594000  
-#                           15858) x2_sigma>=0.4851646 24   12.4345200  1.18424600  
-#                             31716) e_dist=uniform 7    0.5259639  0.41482310 *
-#                             31717) e_dist=even_discrete,lognormal 17    6.0580960  1.50106700 *
-#                           15859) x2_sigma< 0.4851646 8    5.5692240  2.23102400 * (small x2_sigma, small b, small splines dependence, small e_sigma, small sparse, second predictor, non-linearity -> large under-prediction)
-#                        3965) a< -0.6695046 14   11.6634200  1.94820600 *
-#                      1983) b>=90.99624 7   16.3901800  2.56861600 * (large b, small splines dependence, small e_sigma, small sparse, second predictor, non-linearity -> large under-prediction)
-#            31) e_sigma< 0.1149429 48  135.8347000  1.35422700  
-#              62) measure_striped< 0.44 40   54.1808300  0.92957180  
-#               124) include_non_normal=TRUE 21   14.0872600  0.39910760  
-#                 248) include_heter=TRUE 13    5.7917480  0.06683663 *
-#                 249) include_heter=FALSE 8    4.5279800  0.93904800 *
-#               125) include_non_normal=FALSE 19   27.6530600  1.51587400 *
-#              63) measure_striped>=0.44 8   38.3742300  3.47750200 * (high discreteness, small e_sigma, second predictor, non-linearity -> large under-prediction)
+```{r performance-sub}
 
@@
   filter(res == 32L) %>%
-  filter(type == "test") %>%
-  mutate(.resid = effect_size - vss) %>%
-  mutate(include_z = factor(include_z),
-         include_heter = factor(include_heter),
-         include_x2 = factor(include_x2),
-         include_non_normal = factor(include_non_normal),
-         e_even = factor(e_even),
-         x1_even = factor(x1_even),
-         x2_even = factor(x2_even)) %>%
-  select(-filename, -vss, -res, -plot_uid, -data_type, -effect_size, -type) %>%
-  rpart::rpart(.resid ~ ., data = ., control = rpart::rpart.control(cp = 0.001), model = TRUE) -> test_tree
-```
-
-```{r performance-sub}
-
-model_pred %>%
-  filter(res == 32L) %>%
-  group_by(type, include_non_normal, include_heter, include_z) %>%
-  summarise(n = n(), 
-            RMSE = yardstick::rmse_vec(effect_size, vss),
-            R2 = yardstick::rsq_vec(effect_size, vss),
-            MAE = yardstick::mae_vec(effect_size, vss),
-            HUBER = yardstick::huber_loss_vec(effect_size, vss)) %>%
+  group_by(include_non_normal, include_heter, include_z) %>%
+  summarise(train_n = sum(type == "train"),
+            train_RMSE = yardstick::rmse_vec(effect_size[type == "train"], vss[type == "train"]),
+            test_n = sum(type == "test"),
+            test_RMSE = yardstick::rmse_vec(effect_size[type == "test"], vss[type == "test"])) %>%
   mutate(violations = ifelse(include_z, "non-linearity", "null")) %>%
@@
   ungroup() %>%
-  select(type, violations, n, RMSE, R2, MAE, HUBER) %>%
-  arrange(desc(type)) %>%
-  select(-type) %>%
-  mutate(n = format(n)) %>%
-  mutate(across(RMSE:MAE, ~format(.x, digits = 3))) %>%
-  mutate(across(HUBER, ~format(.x, digits = 2))) %>%
+  select(violations, train_n, train_RMSE, test_n, test_RMSE) %>%
+  mutate(train_n = format(train_n)) %>%
+  mutate(test_n = format(test_n)) %>%
+  mutate(across(c(train_RMSE, test_RMSE), ~format(.x, digits = 3))) %>%
   mutate(violations = ifelse(violations == "null", "no violations", violations)) %>%
-  mutate(R2 = gsub(".*NA.*", "", R2)) %>%
   kableExtra::kable("latex", 
@@
                     escape = FALSE,
-                    col.names = c("Violations", "\\#samples", "RMSE", "$R^2$", "MAE", "Huber loss"),
-                    caption = "The training and test performance of the $32 \\times 32$ model presented with different model violations.  The best metrics are colored in red.") %>%
-  kableExtra::kable_styling(latex_options = c("scale_down")) %>%
-  kableExtra::pack_rows("Training set", 1, 8) %>%
-  kableExtra::pack_rows("Test set", 9, 16) %>%
-  kableExtra::column_spec(3, color = c(rep("black", 4), "red", rep("black", 3),
-                                       rep("black", 4), "red", rep("black", 3))) %>%
-  kableExtra::column_spec(4, color = c(rep("black", 4), "red", rep("black", 3),
-                                       rep("black", 4), "red", rep("black", 3))) %>%
-  kableExtra::column_spec(5, color = c(rep("black", 4), "red", rep("black", 3),
-                                       rep("black", 4), "red", rep("black", 3))) %>%
-  kableExtra::column_spec(6, color = c(rep("black", 4), "red", rep("black", 3),
-                                       rep("black", 4), "red", rep("black", 3)))
-```
-
-## Model Violations Index (MVI)
+                    col.names = c("Violations", "\\#training samples", "Training RMSE", "\\#test samples", "Test RMSE"),
+                    caption = "The training and test performance of the $32 \\times 32$ model presented with different model violations.") %>%
+  kableExtra::kable_styling(latex_options = c("scale_down"))
+```
+
+## Demostration of MVI
+
+As described in Section \ref{sec:model-violations-index} a Model Violations Index (MVI) can be constructed using $\hat{D}$. This section provides guidelines and examples of MVI.
 
 Figures \ref{fig:poly-index} and \ref{fig:heter-index} display the residual plots for fitted models exhibiting varying degrees of non-linearity and heteroskedasticity.
-Each residual plot's Model Violation Index (MVI) is computed using Equation \ref{eq:mvi} with $C = 10$.
+Each residual plot's MVI is computed using Equation \ref{eq:mvi} with $C = 10$.
 When $\text{MVI} > 8$, the visual patterns are notably strong and easily discernible by humans.
@@
 While the distance $D$ for the actual residual plot depends on the underlying data generating process, the distance $D$ for the null plots is zero.
-We have used our optimized computer vision model to predict $\hat{D}$ for both the actual residual plots and the null plots.
+We have used our optimized computer vision model to predict the approximated distance for both the actual residual plots and the null plots.
+To have a fair comparison, $H_0$ will be rejected if the actual residual plot has the greatest approximated distance among all plots in a lineup.
 Additionally, the appropriate conventional tests including the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey1969tests] for non-linearity and the Breusch-Pagan test  [@breusch1979simple] for heteroskedasticity were applied on the same data for comparison.
@@
 
-For null plots, a histogram of the predictions is provided in Figure \ref{fig:hist-null-human}.
+<!-- For null plots, a histogram of the predictions is provided in Figure \ref{fig:hist-null-human}.
 In most cases, $\hat{D}$ is predicted to be around $1$.
@@
 Therefore, we will be using the 95% sample quantiles obtained from each fitted model in the following analysis.
-
-Likewise, Figure \ref{fig:conv-mosaic} and Table \ref{tab:conv-table} provide a summary of the agreement between decisions made by the computer vision model and conventional tests.
-The agreement rates between conventional tests and the computer vision model are 65.37% and 57.29% for residual plots containing heteroskedasticity and non-linearity patterns, respectively.
-These figures are lower than those presented in Table \ref{tab:human-table}, indicating that the computer vision model exhibits behaviour more akin to visual tests than conventional tests.
-Additionally, conventional tests tend to reject residual plots almost exclusively when the computer vision model does, but they also reject plots when the computer vision model does not.
+-->
+
+Table \ref{tab:human-conv-table} provides a summary of the agreement between decisions made by the computer vision model and conventional tests.
+The agreement rates between conventional tests and the computer vision model are 85.95% and 79.69% for residual plots containing heteroskedasticity and non-linearity patterns, respectively.
+These figures are higher than those calculated for visual tests conducted by human, indicating that the computer vision model exhibits behaviour more akin to the best available conventional tests.
+However, Figure \ref{fig:conv-mosaic} shows that conventional tests tend to reject residual plots almost exclusively when the computer vision model does, but they also reject plots when the computer vision model does not.
 This suggests that conventional tests are more sensitive than the computer vision model.
@@
 When plotting the decision against the distance, as illustrated in Figure \ref{fig:power}, several observations emerge.
-Firstly, for the computer vision model, when the distance $D > 3$, almost all residual plots are rejected.
+Firstly, for the computer vision model, when the distance $D > 4$, almost all residual plots are rejected.
 However, for visual tests conducted by humans, the threshold is higher, set at $D > 5$.
@@
 Rejection decisions are fitted by logistic regression models with no intercept terms but an offset equals to $\text{log}(0.05/0/95)$.
-The fitting curves for the computer vision model fall between those of conventional tests and visual tests for non-linearity and are slightly below conventional tests for heteroskedasticity.
+The fitting curves for the computer vision model fall between those of conventional tests and visual tests for both non-linearity and heteroskedasticity.
 There is still potential to refine the computer vision model to better align its behaviour with visual tests.
@@
 
-lineup_vss %>%
-  filter(null) %>%
-  left_join(vi_survey %>%
-              group_by(unique_lineup_id) %>% 
-              summarise(type = first(type), 
-                        attention_check = first(attention_check),
-                        null_lineup = first(null_lineup))) %>%
+experiment <- experiment %>%
+  left_join(lineup_vss %>%
+  filter(null == FALSE) %>%
+  mutate(model_reject_20 = rank == 1) %>%
+  select(unique_lineup_id, rank, model_reject_20))
+
+# lineup_vss %>%
+#   filter(null) %>%
+#   left_join(vi_survey %>%
+#               group_by(unique_lineup_id) %>% 
+#               summarise(type = first(type), 
+#                         attention_check = first(attention_check),
+#                         null_lineup = first(null_lineup))) %>%
+#   filter(!attention_check) %>%
+#   filter(!null_lineup) %>%
+#   mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+#   ggplot() +
+#   geom_histogram(aes(vss)) +
+#   geom_vline(aes(xintercept = quantile(vss, 0.9), col = "Q(0.90)")) +
+#   geom_vline(aes(xintercept = quantile(vss, 0.95), col = "Q(0.95)")) +
+#   geom_vline(aes(xintercept = quantile(vss, 0.99), col = "Q(0.99)")) +
+#   xlab(expression(hat(D))) +
+#   theme_light() +
+#   facet_wrap(~type) +
+#   labs(col = "")
+```
+
+```{r cache = TRUE}
+vi_lineup <- get_vi_lineup()
+```
+
+```{r eval = FALSE}
+library(visage)
+conv_experiment <- vi_lineup %>%
+  map_df(function(.x) {
+    reset_p_value <- c()
+    bp_p_value <- c()
+    sw_p_value <- c()
+
+    for (i in 1:20) {
+      this_dat <- .x$data %>% filter(k == i)
+      reset_p_value <- c(reset_p_value, POLY_MODEL$test(this_dat, test = "RESET")$p_value)
+      bp_p_value <- c(bp_p_value, HETER_MODEL$test(this_dat)$p_value)
+      sw_p_value <- c(sw_p_value, shapiro.test(this_dat$.resid)$p.value)
+    }
+    return(tibble(unique_lineup_id = rep(.x$metadata$name, 20),
+                  ans = .x$metadata$answer,
+                  pos = 1:20,
+                  reset_p_value = reset_p_value,
+                  reset_rank = rank(reset_p_value, ties.method = "first"),
+                  bp_p_value = bp_p_value,
+                  bp_rank = rank(bp_p_value, ties.method = "first"),
+                  sw_p_value = sw_p_value,
+                  sw_rank = rank(sw_p_value, ties.method = "first")))
+  })
+
+conv_experiment %>%
+  filter(ans == pos) %>%
+  mutate(reset_reject_20 = reset_rank == 1,
+         bp_reject_20 = bp_rank == 1,
+         sw_reject_20 = sw_rank == 1) %>%
+  select(-ans, -pos, -c(reset_p_value:sw_rank)) %>%
+  right_join(experiment) %>%
   filter(!attention_check) %>%
   filter(!null_lineup) %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
+  pivot_longer(c(reject, model_reject_20, reset_reject_20, bp_reject_20, conventional_reject, model_reject)) %>%
+  select(type, actual_ss, name, value) %>%
+  mutate(offset0 = log(0.05/0.95)) %>%
+  nest(dat = c(offset0, actual_ss, value)) %>%
+  mutate(mod = map(dat, 
+                   ~glm(value ~ actual_ss - 1, 
+                        family = binomial(), 
+                        data = .x,
+                        offset = offset0))) %>%
+  mutate(power = map(mod, function(mod) {
+    result <- data.frame(actual_ss = seq(0, 10, 0.1),
+                         offset0 = log(0.05/0.95))
+    result$power <- predict(mod, type = "response", newdata = result)
+    result
+  })) %>%
+  select(-dat, -mod) %>%
+  unnest(power) %>%
   ggplot() +
-  geom_histogram(aes(vss)) +
-  geom_vline(aes(xintercept = quantile(vss, 0.9), col = "Q(0.90)")) +
-  geom_vline(aes(xintercept = quantile(vss, 0.95), col = "Q(0.95)")) +
-  geom_vline(aes(xintercept = quantile(vss, 0.99), col = "Q(0.99)")) +
-  xlab(expression(hat(D))) +
-  theme_light() +
-  facet_wrap(~type) +
-  labs(col = "")
-```
-
-```{r cache = TRUE}
-vi_lineup <- get_vi_lineup()
-```
-
-```{r human-table}
+  geom_line(aes(actual_ss, power, col = name)) +
+  facet_wrap(~type, scales = "free_x")
+```
+
+```{r human-conv-table, eval = TRUE}
+bind_rows(
   experiment %>%
@@
     group_by(type) %>%
-  summarise(n = format(n()), agree = format(sum(model_reject == reject)), rate = format(mean(model_reject == reject), digits = 4)) %>%
-  kableExtra::kable("latex", 
-                    booktabs = TRUE,
-                    escape = FALSE,
-                    col.names = c("Violations", "\\#Samples", "\\#Agreements", "Agreement rate"),
-                    caption = "Summary of decisions made by visual tests conducted by human and decisions made by the computer vision model.")
-```
-
-```{r conv-table}
+    summarise(n = format(n()), agree = format(sum(model_reject == conventional_reject)), rate = format(mean(conventional_reject == model_reject), digits = 4)),
   experiment %>%
@@
     group_by(type) %>%
-  summarise(n = format(n()), agree = format(sum(model_reject == conventional_reject)), rate = format(mean(conventional_reject == reject), digits = 4)) %>%
+    summarise(n = format(n()), agree = format(sum(model_reject == reject)), rate = format(mean(model_reject == reject), digits = 4))
+  ) %>%
   kableExtra::kable("latex", 
@@
                     col.names = c("Violations", "\\#Samples", "\\#Agreements", "Agreement rate"),
-                    caption = "Summary of decisions made by conventional tests and decisions made by the computer vision model.")
+                    caption = "Summary of the comparison of decisions made by computer vision model with decisions made by conventional tests and visual tests conducted by human.") %>%
+  kableExtra::pack_rows("Compared with conventional tests", 1, 2) %>%
+  kableExtra::pack_rows("Compared with visual tests conducted by human", 3, 4)
 ```
@@
 ```{r human-mosaic, fig.pos = "!h", fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of visual test conditional on the computer vision model decision on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot.
-The visual test rejects less frequently than the computer vision model, and (almost) only rejects when the computer vision model does."}
+The visual test rejects less frequently than the computer vision model, and (almost) only rejects when the computer vision model does.", eval = FALSE}
 library(ggmosaic)
@@
   mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
-  mutate(model_reject = ifelse(model_reject, "Reject", "No")) %>%
+  mutate(model_reject_20 = ifelse(model_reject_20, "Reject", "No")) %>%
   mutate(reject = ifelse(reject, "Reject", "No")) %>%
-  mutate(across(c(model_reject, reject), ~factor(.x, levels = c("Reject", "No")))) %>%
+  mutate(across(c(model_reject_20, reject), ~factor(.x, levels = c("Reject", "No")))) %>%
   ggplot() +
-  geom_mosaic(aes(x = ggmosaic::product(reject, model_reject), 
-                  fill = reject)) +
+  geom_mosaic(aes(x = ggmosaic::product(model_reject_20, reject), 
+                  fill = model_reject_20)) +
   facet_grid(~type) +
-  ylab("Visual tests reject") +
-  xlab("Computer vision model rejects") +
+  xlab("Visual tests reject") +
+  ylab("Computer vision model rejects") +
   # labs(fill = "Visual tests reject") +
@@
   mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
-  mutate(model_reject = ifelse(model_reject, "Reject", "No")) %>%
+  mutate(model_reject_20 = ifelse(model_reject_20, "Reject", "No")) %>%
   mutate(conventional_reject = ifelse(conventional_reject, "Reject", "No")) %>%
-  mutate(across(c(model_reject, conventional_reject), ~factor(.x, levels = c("Reject", "No")))) %>%
+  mutate(across(c(model_reject_20, conventional_reject), ~factor(.x, levels = c("Reject", "No")))) %>%
   ggplot() +
-  geom_mosaic(aes(x = ggmosaic::product(conventional_reject, model_reject), 
+  geom_mosaic(aes(x = ggmosaic::product(conventional_reject, model_reject_20), 
                   fill = conventional_reject)) +
   facet_grid(~type) +
-  ylab("Conventional tests reject") +
   xlab("Computer vision model rejects") +
+  ylab("Conventional tests reject ") +
   # labs(fill = "Visual tests reject") +
@@
 
-```{r power, fig.pos = "!h", fig.cap = "Comparison of decisions made by visual tests, conventional tests and the computer vision model.
-The data points are fitted with logistic regression models with no intercept but an offset equals to $\\text{log}(0.05/0.95)$."}
+```{r power, fig.pos = "!h", fig.cap = "Comparison of power of visual tests, conventional tests and the computer vision model.
+Marks along the x-axis at the bottom of the plot represent rejections made by each type of test.
+Marks at the top of the plot represent acceptances.
+Power curves are fitted by logistic regression models with no intercept but an offset equals to $\\text{log}(0.05/0.95)$."}
 
@@
 
-model_glm_pred_poly <- glm(model_reject ~ actual_ss - 1, 
+model_glm_pred_poly <- glm(model_reject_20 ~ actual_ss - 1, 
     data = filter(experiment, !attention_check, !null_lineup, type == "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
@@
 
-model_glm_pred_heter <- glm(model_reject ~ actual_ss - 1, 
+model_glm_pred_heter <- glm(model_reject_20 ~ actual_ss - 1, 
     data = filter(experiment, !attention_check, !null_lineup, type != "polynomial") %>% mutate(offset0 = log(0.05/0.95)), 
@@
   filter(!null_lineup, !attention_check) %>%
-  pivot_longer(c(model_reject, reject, conventional_reject)) %>%
-  mutate(name = ifelse(name == "model_reject", "Computer vision model", name)) %>%
+ggplot() +
+  geom_line(data = model_glm_pred_poly, aes(d, value, col = "Computer vision model"), linewidth = 0.8) +
+  geom_line(data = model_glm_pred_heter, aes(d, value, col = "Computer vision model"), linewidth = 0.8) +
+  geom_line(data = visual_glm_pred_poly, aes(d, value, col = "Visual test"), linewidth = 0.8) +
+  geom_line(data = visual_glm_pred_heter, aes(d, value, col = "Visual test"), linewidth = 0.8) +
+  geom_line(data = conv_glm_pred_poly, aes(d, value, col = "Conventional test"), linewidth = 0.8) +
+  geom_line(data = conv_glm_pred_heter, aes(d, value, col = "Conventional test"), linewidth = 0.8) +
+  geom_segment(data = ~filter(.x, !reject), aes(x = actual_ss, xend = actual_ss, y = -0.1, yend = -0.1 + 0.03, col = "Visual test"), alpha = 0.4) +
+  geom_segment(data = ~filter(.x, reject), aes(x = actual_ss, xend = actual_ss, y = 1.1 - 0.03 * 2, yend = 1.1 - 0.03 * 3, col = "Visual test"), alpha = 0.4) +
+  geom_segment(data = ~filter(.x, !model_reject_20), aes(x = actual_ss, xend = actual_ss, y = -0.1 + 0.03, yend = -0.1 + 0.03 * 2, col = "Computer vision model"), alpha = 0.4) +
+  geom_segment(data = ~filter(.x, model_reject_20), aes(x = actual_ss, xend = actual_ss, y = 1.1 - 0.03, yend = 1.1 - 0.03 * 2, col = "Computer vision model"), alpha = 0.4) +
+  geom_segment(data = ~filter(.x, !conventional_reject), aes(x = actual_ss, xend = actual_ss, y = -0.1 + 0.03 * 2, yend = -0.1 + 0.03 * 3, col = "Conventional test"), alpha = 0.4) +
+  geom_segment(data = ~filter(.x, conventional_reject), aes(x = actual_ss, xend = actual_ss, y = 1.1, yend = 1.1 - 0.03, col = "Conventional test"), alpha = 0.4) +
+  facet_grid(~type, scales = "free_x") +
+  theme_light() +
+  scale_color_brewer("", palette = "Dark2") +
+  theme(legend.position = "bottom") +
+  xlab("D") +
+  ylab("Reject")
+```
+
+```{r pcp, fig.cap = "Parallel coordinate plots of decisions made by computer vision model, conventional tests and visual tests made by human.", eval = TRUE}
+library(ggpcp)
+mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
+  filter(!null_lineup, !attention_check) %>%
+  select(type, actual_ss, model_reject_20, reject, conventional_reject) %>%
+  mutate(test = rnorm(n())) %>%
+  mutate(across(c(model_reject_20, conventional_reject, reject), ~ifelse(.x, "Reject", "Not reject"))) %>%
+  mutate(agree = model_reject_20 == reject) %>%
+  pcp_select(c(reject, model_reject_20, conventional_reject)) %>%
+  group_by(type) %>%
+  pcp_scale() %>%
+  ungroup() %>%
+  pcp_arrange(method = "from-right") %>%
+  ggplot() +
+  geom_pcp(aes_pcp(), linewidth = 0.1) +
+  geom_pcp_labels(aes_pcp()) +
+  geom_pcp_boxes(aes_pcp(), boxwidth = 0.1) +
+  facet_wrap(~type) +
+  theme_light() +
+  scale_x_discrete(labels = c("Visual test", "Computer vision model", "Conventional test")) +
+  labs(col = "D") +
+  ylab("") +
+  xlab("") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
+        axis.text.y = element_blank(), 
+        axis.ticks.y = element_blank(), 
+        axis.title = element_blank())
+```
+
+```{r results = "hide"}
+library(VennDiagram)
+mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
+  filter(type == "Non-linearity") %>%
+  filter(!null_lineup, !attention_check) %>%
+  select(unique_lineup_id, type, actual_ss, model_reject_20, reject, conventional_reject) %>%
+  mutate(across(c(model_reject_20, conventional_reject, reject), ~ifelse(.x, "Reject", "Not reject"))) %>%
+  pivot_longer(c(model_reject_20, reject, conventional_reject)) %>%
+  group_by(name) %>%
+  summarise(ids = list(unique_lineup_id[value != "Reject"])) %>%
+  pull(ids) %>%
+  venn.diagram(
+    category.names = c("Conventional test" , "Computer vision model" , "Visual tests"),
+    filename = here::here('paper/diagram/poly_venn_diagramm.png'),
+    disable.logging = TRUE,
+    lwd = 1,
+    col = RColorBrewer::brewer.pal(3, "Dark2"),
+    fill = alpha(RColorBrewer::brewer.pal(3, "Dark2"), 0.3),
+    cex = 0.5,
+    fontfamily = "sans",
+    cat.cex = 0.5,
+    cat.default.pos = "outer",
+    cat.fontfamily = "sans",
+    cat.pos = c(330, 30, 180),
+    cat.dist = c(0, 0, 0.03),
+    cat.col = RColorBrewer::brewer.pal(3, "Dark2"),
+    height = 800,
+    width = 800,
+    resolution = 300,
+    main.fontfamily = "sans",
+    main.cex = 0.6,
+    main = "Non-linearity",
+    main.pos = c(0.25, 1.1)
+  )
+
+mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
+  filter(type != "Non-linearity") %>%
+  filter(!null_lineup, !attention_check) %>%
+  select(unique_lineup_id, type, actual_ss, model_reject_20, reject, conventional_reject) %>%
+  mutate(across(c(model_reject_20, conventional_reject, reject), ~ifelse(.x, "Reject", "Not reject"))) %>%
+  pivot_longer(c(model_reject_20, reject, conventional_reject)) %>%
+  group_by(name) %>%
+  summarise(ids = list(unique_lineup_id[value != "Reject"])) %>%
+  pull(ids) %>%
+  venn.diagram(
+    category.names = c("Conventional test" , "Computer vision model" , "Visual tests"),
+    filename = here::here('paper/diagram/heter_venn_diagramm.png'),
+    disable.logging = TRUE,
+    lwd = 1,
+    col = RColorBrewer::brewer.pal(3, "Dark2"),
+    fill = alpha(RColorBrewer::brewer.pal(3, "Dark2"), 0.3),
+    cex = 0.5,
+    fontfamily = "sans",
+    cat.cex = 0.5,
+    cat.default.pos = "outer",
+    cat.fontfamily = "sans",
+    cat.pos = c(330, 30, 180),
+    cat.dist = c(0, 0, 0.03),
+    cat.col = RColorBrewer::brewer.pal(3, "Dark2"),
+    height = 800,
+    width = 800,
+    resolution = 300,
+    main.fontfamily = "sans",
+    main.cex = 0.6,
+    main = "Heteroskedasticity",
+    main.pos = c(0.25, 1.1)
+  )
+```
+
+```{r venn, fig.pos = "!h", fig.cap = "Venn diagram for non-rejection decisions made by conventional tests, computer vision model, and visual tests conducted by human."}
+poly_venn <- magick::image_read(here::here("paper/diagram/poly_venn_diagramm.png")) %>%
+  magick::image_ggplot()
+heter_venn <- magick::image_read(here::here("paper/diagram/heter_venn_diagramm.png")) %>%
+  magick::image_ggplot()
+patchwork::wrap_plots(poly_venn, heter_venn, ncol = 2)
+```
+
+```{r eval = FALSE}
+mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
+  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
+  filter(!null_lineup, !attention_check) %>%
+  pivot_longer(c(model_reject_20, reject, conventional_reject)) %>%
+  mutate(name = ifelse(name == "model_reject_20", "Computer vision model", name)) %>%
   mutate(name = ifelse(name == "reject", "Visual test", name)) %>%
@@
   ggplot() +
-  ggbeeswarm::geom_quasirandom(aes(actual_ss, value), size = 0.5, alpha = 0.4) +
+  ggbeeswarm::geom_quasirandom(aes(actual_ss, value, col = name), size = 0.5, alpha = 0.4) +
   geom_line(data = model_glm_pred_poly, aes(d, value + 1, col = "Computer vision model"), linewidth = 0.8) +
@@
 
+```{r eval = FALSE}
+plot_path <- my_vi$plot_resid() %>%
+  my_vi$save_plot()
+input_auxiliary <- my_vi$auxiliary()
+
+reticulate::py$get_raw_gradient(keras_mod = keras_mod, 
+                                plot_path = plot_path, 
+                                target_size = c(32L, 32L), 
+                                input_auxiliary = input_auxiliary) -> raw_g
+
+reticulate::py$get_smooth_gradient(keras_mod = keras_mod, 
+                                   plot_path = plot_path, 
+                                   target_size = c(32L, 32L), 
+                                   input_auxiliary = input_auxiliary,
+                                   noise_level = 0.45,
+                                   n = 200L) -> smooth_g
+
+x_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$x.range
+y_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$y.range
+
+raw_att_map <- raw_g[[1]] |> 
+  as.data.frame() |>
+  dplyr::mutate(row = rev(1:32)) |>
+  tidyr::pivot_longer(V1:V32, names_to = "column", values_to = "gradient") |>
+  dplyr::mutate(column = as.integer(gsub("V", "", column))) |>
+  dplyr::mutate(column = x_range[1] + column * (x_range[2] - x_range[1])/(32 - 1)) |>
+  dplyr::mutate(row = y_range[1] + row * (y_range[2] - y_range[1])/(32 - 1))
+
+my_vi$plot_resid() +
+  ggplot() +
+  ggplot2::geom_raster(data = raw_att_map, ggplot2::aes(column, row, fill = gradient), alpha = 0.7) +
+  scale_fill_gradient(low = "black", high = "white") +
+  # viridis::scale_fill_viridis(option = "D") +
+  ggplot2::theme_void() +
+  ggplot2::ggtitle(paste0(gsub("measure_", "", names(input_auxiliary)), ": ", format(raw_g[[2]], digits = 2), 
+                          collapse = ", ")) -> p1
+
+smooth_att_map <- smooth_g[[1]] |> 
+  as.data.frame() |>
+  dplyr::mutate(row = rev(1:32)) |>
+  tidyr::pivot_longer(V1:V32, names_to = "column", values_to = "gradient") |>
+  dplyr::mutate(column = as.integer(gsub("V", "", column))) |>
+  dplyr::mutate(column = x_range[1] + column * (x_range[2] - x_range[1])/(32 - 1)) |>
+  dplyr::mutate(row = y_range[1] + row * (y_range[2] - y_range[1])/(32 - 1))
+
+scale_zero_one <- function(x) (x - min(x))/(max(x) - min(x))
+
+# my_vi$plot_resid() +
+  ggplot() +
+  ggplot2::geom_raster(data = mutate(smooth_att_map, gradient = scale_zero_one(gradient)), ggplot2::aes(column, row, fill = gradient), alpha = 0.7) +
+  scale_fill_gradient(low = "black", high = "white") +
+  # viridis::scale_fill_viridis(option = "D") +
+  ggplot2::theme_void() +
+  ggplot2::ggtitle(paste0(gsub("measure_", "", names(input_auxiliary)), ": ", format(raw_g[[2]], digits = 2), 
+                          collapse = ", ")) -> p2
+
+p1 + p2
+```
+
 ### Datasaurus
@@
 - Website link
+
+```{r eval = FALSE}
+
+# This shows that even if the model can 100% accurately predict the distance for not null plots, 
+# the statistical tests based on approximated distance will still reject more cases than visual tests
+# So it has very little to do with the accuracy of the prediction
+experiment %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(model_reject_true = actual_ss > vss_null_95) %>%
+  filter(model_reject_true != reject) %>%
+  ggplot() +
+  ggbeeswarm::geom_quasirandom(aes(actual_ss, reject)) +
+  ggtitle("Unmatched cases") +
+  ylab("Visual tests reject") +
+  xlab("D")
+```
+
+```{r eval = FALSE}
+
+map_dbl(10^seq(-10, 0, 0.01), function(crt) {
+  experiment %>%
+    filter(!null_lineup) %>%
+    filter(!attention_check) %>%
+    mutate(mvi = vss - log(n) + 10) %>%
+    mutate(actual_ss_cut = conventional_p_value <= crt) %>%
+    mutate(agree = actual_ss_cut == reject) %>%
+    pull(agree) %>%
+    mean()
+}) %>%
+  max()
+
+experiment %>%
+  filter(!null_lineup) %>%
+  filter(!attention_check) %>%
+  mutate(mvi = vss - log(n) + 10) %>%
+  ggplot() +
+  ggbeeswarm::geom_quasirandom(aes(vss, reject)) +
+  ggtitle("Unmatched cases") +
+  ylab("Visual tests reject") +
+  xlab("D")
+  mutate(model_reject_true = actual_ss > vss_null_95) %>%
+  filter(model_reject_true != reject)
+```

### 2024-04-15 paper/paper.tex
update paper

@@
 \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
 \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
@@
 \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
 \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
@@
 \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
 \newcommand{\ImportTok}[1]{#1}
@@
 \newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
 \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 
-\hypertarget{methodology}{%
-\section{Methodology}\label{methodology}}
-
-\hypertarget{different-possible-configurations-of-the-model-formula}{%
+\section{Methodology}\label{methodology}
+
 \subsection{Different possible configurations of the model
-formula}\label{different-possible-configurations-of-the-model-formula}}
+formula}\label{different-possible-configurations-of-the-model-formula}
 
@@
 
-\hypertarget{input-formats}{%
-\subsubsection{Input formats}\label{input-formats}}
+\subsubsection{Input formats}\label{input-formats}
 
@@
 
-\hypertarget{output-formats}{%
-\subsubsection{Output formats}\label{output-formats}}
+\subsubsection{Output formats}\label{output-formats}
 
@@
 
-\hypertarget{distance-from-the-good-residual-plots}{%
 \subsection{Distance from the good residual
-plots}\label{distance-from-the-good-residual-plots}}
+plots}\label{distance-from-the-good-residual-plots}
 
@@
 
-\hypertarget{residual-distribution}{%
-\subsubsection{Residual distribution}\label{residual-distribution}}
+\subsubsection{Residual distribution}\label{residual-distribution}
 
@@
 
-\hypertarget{kullback-leibler-divergence-of-p-from-q}{%
 \subsubsection{\texorpdfstring{Kullback-Leibler divergence of \(P\) from
-\(Q\)}{Kullback-Leibler divergence of P from Q}}\label{kullback-leibler-divergence-of-p-from-q}}
+\(Q\)}{Kullback-Leibler divergence of P from Q}}\label{kullback-leibler-divergence-of-p-from-q}
 
@@
 
-\hypertarget{evaluation-of-kullback-leibler-divergence-for-non-normal-p}{%
 \subsubsection{\texorpdfstring{Evaluation of Kullback-Leibler divergence
 for non-normal
-\(P\)}{Evaluation of Kullback-Leibler divergence for non-normal P}}\label{evaluation-of-kullback-leibler-divergence-for-non-normal-p}}
+\(P\)}{Evaluation of Kullback-Leibler divergence for non-normal P}}\label{evaluation-of-kullback-leibler-divergence-for-non-normal-p}
 
@@
 
-\hypertarget{approximation-of-the-distance-measure}{%
 \subsubsection{Approximation of the distance
-measure}\label{approximation-of-the-distance-measure}}
+measure}\label{approximation-of-the-distance-measure}
 
@@
 
-\hypertarget{approximated-distance-as-an-model-violations-index-mvi}{%
-\subsubsection{Approximated distance as an Model Violations Index
-(MVI)}\label{approximated-distance-as-an-model-violations-index-mvi}}
+\subsubsection{Model Violations Index
+(MVI)}\label{sec:model-violations-index}
 
@@
 
-Note that since the Kullback-Leibler divergence could be affected by the
-number of random variables involved in the evaluation. In general, more
-observations will result in greater distance \(D\). This does not mean
-\(\hat{D}\) can not correctly represent the extent of the model
-violations since more observations in a residual plot usually lead to
-stronger visual signal strength as the underlying visual patterns will
-have higher chance to be revealed except for cases like serious
-overlapping. For the same data generating process, \(D\) usually grows
-logarithmically as the number of observations increases. This is because
-\(D = \text{log}(1 + D_{KL})\), and
-\(D_{KL} = \sum_{i=1}^{n}D_{KL}^{(i)}\) under the independence
-assumption.
-
 Note that the Kullback-Leibler divergence might be influenced by the
@@
 
-\hypertarget{statistical-testing-based-on-the-approximated-distance}{%
 \subsection{Statistical testing based on the approximated
-distance}\label{statistical-testing-based-on-the-approximated-distance}}
-
-\hypertarget{null-distribution-of-the-approximated-distance}{%
-\subsubsection{Null distribution of the approximated
-distance}\label{null-distribution-of-the-approximated-distance}}
+distance}\label{statistical-testing-based-on-the-approximated-distance}
+
+\subsubsection{Lineup evaluation}\label{lineup-evaluation}
 
@@
 
-To solve this issue while aligning with the principle of visual
-inference, \(\hat{D}\) can be viewed as a test statistic. And the null
-distribution of this statistic can be approximated by the empirical
-distribution of \(\hat{D}_{null}\), where \(\hat{D}_{null}\) is the
-approximated distance for a null plot simulated from the fitted
-regression model. The approximation of the distribution involves
-applying the residual rotation technique \citep{buja2009statistical} on
-the fitted regression model to obtain null residuals. The null residuals
-are then used to make null plots and fed into the computer vision model
-to get predictions. The empirical distribution is constructed with the
-approximated distance.
-
-There are two types of error associated with the approximation of the
-null distribution. These include the sampling error and the estimation
-error for model parameters. Increasing the number of null plots within
-the distribution approximation can effectively reduce sampling error.
-However, the estimation error for model parameters remains irreducible.
-The residual rotation technique operates under the assumption of a
-correct fitted model, using the estimated variance instead of the
-unknown true variance to rescale rotated residuals. Consequently, under
-\(H_0\), the true distribution of \(\hat{D}_{null}\) will slightly
-deviate from the approximated null distribution.
-
-\hypertarget{estimation-of-quantiles-of-the-null-distribution}{%
-\subsubsection{Estimation of quantiles of the null
-distribution}\label{estimation-of-quantiles-of-the-null-distribution}}
-
-Let \(\hat{D}_{null}^{(i)}\) be the approximated distance of the
-\(i\)-th null plots, where \(i = 1,...,n_{null}\) and
-\(n_{null} \in \mathbb{N}^+\) is a sufficiently large number. Quantiles
-of the null distribution can then be estimated using the sample
-quantiles available in statistical software such as R. The details of
-the sample quantile computation can be found in
-\citet{hyndman1996sample}. In statistical testing, analysts often care
-about certain quantiles of the null distribution, such as the 90\%
-quantile, the 95\% quantile and the 99\% quantile. These quantiles are
-used as critical values to decide if \(H_0\) needs to be rejected. For
-example, if \(\hat{D}\) is greater than and equal to the 95\% sample
-quantile \(Q_{null}(0.95)\), we could say the approximated distance for
-the actual residual plot is significantly different from the
-approximated distance for null plots with 95\% significance level. Based
-on our experience, in order to get a stable estimate of the 95\%
-quantile, \(n_{null}\) usually needs to be at least \(100\). And if the
-null distribution has a long tail, more null plots will be needed.
-Alternatively, a p-value can be used to represents the probability of
-observing an event equally or more extreme than the given event under
+To address this issue while adhering to the principle of visual
+inference, we can compare the estimated distance \(\hat{D}\) to the
+approximated distances for the null plots in a lineup. Specifically, if
+a lineup comprises 20 plots, the null hypothesis \(H_0\) will be
+rejected if \(\hat{D}\) exceeds the maximum approximated distance among
+the \(m - 1\) null plots, denoted as
+\(\max\limits_{1 \leq i \leq m-1} {\hat{D}_{null}^{(i)}}\), where
+\(\hat{D}_{null}^{(i)}\) represents the approximated distance for the
+\(i\)-th null plot. This approach is equivalent to the typical lineup
+protocol requiring a 95\% significance level, where \(H_0\) is rejected
+if the data plot is identified as the most distinct plot by the sole
+participant. The approximated distance serves as a metric to quantify
+the difference between the data plot and the null plots, as intended.
+
+For lineups consisting of more than 20 plots, the 95\% significance
+level can be maintained if the number of plots is a multiple of 20.
+Specifically, for lineups comprising \(20p\) plots, where \(p\) is a
+positive integer, we reject \(H_0\) if \(\hat{D}\) exceeds 95\%
+\({\hat{D}_{null}^{(i)}}\) for \(i = 1, \ldots, 20p-1\). The p-value in
+this case is given by
+\(\frac{1}{m} + \frac{1}{m}\sum{i=1}^{m-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)\),
+where \(I(\cdot)\) is the indicator function.
+
+Moreover, if the number of plots in a lineup, denoted by \(m\), is
+sufficiently large, the empirical distribution of
+\({\hat{D}_{null}^{(i)}}\) can be viewed as an approximation of the null
+distribution of the approximated distance. Consequently, quantiles of
+the null distribution can be estimated using the sample quantiles
+available in statistical software such as R, and these quantiles can be
+utilized for decision-making purposes. The details of the sample
+quantile computation can be found in \citet{hyndman1996sample}. For
+instance, if \(\hat{D}\) is greater than or equal to the 95\% sample
+quantile, denoted as \(Q_{null}(0.95)\), we can conclude that the
+approximated distance for the actual residual plot is significantly
+different from the approximated distance for null plots with a 95\%
+significance level. Based on our experience, to obtain a stable estimate
+of the 95\% quantile, the number of null plots, \(n_{null}\), typically
+needs to be at least 100. However, if the null distribution exhibits a
+long tail, a larger number of null plots may be required. Alternatively,
+a p-value can be used to represent the probability of observing an event
+equally or more extreme than the given event under the null hypothesis
 \(H_0\), and it can be estimated by
-\(1/m\sum_{i=1}^{n_{null}}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)\).
+\(\frac{1}{m}\sum_{i=1}^{m-1}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)\).
 
@@
 
-\hypertarget{bootstrapping-the-approximated-distance}{%
 \subsubsection{Bootstrapping the approximated
-distance}\label{bootstrapping-the-approximated-distance}}
+distance}\label{bootstrapping-the-approximated-distance}
 
@@
 
-\hypertarget{generation-of-training-data}{%
 \subsection{Generation of training
-data}\label{generation-of-training-data}}
-
-\hypertarget{data-generating-process}{%
-\subsubsection{Data generating process}\label{data-generating-process}}
+data}\label{generation-of-training-data}
+
+\subsubsection{Data generating process}\label{data-generating-process}
 
@@
 
-\hypertarget{computation-of-scagnostics}{%
 \subsubsection{Computation of
-scagnostics}\label{computation-of-scagnostics}}
+scagnostics}\label{computation-of-scagnostics}
 
@@
 
-\hypertarget{crafting-a-balanced-training-set}{%
 \subsubsection{Crafting a balanced training
-set}\label{crafting-a-balanced-training-set}}
+set}\label{crafting-a-balanced-training-set}
 
@@
 
-\hypertarget{architecture-of-the-computer-vision-model}{%
 \subsection{Architecture of the computer vision
-model}\label{architecture-of-the-computer-vision-model}}
+model}\label{architecture-of-the-computer-vision-model}
 
@@
 
-\hypertarget{training-and-hyperparameter-tuning}{%
 \subsection{Training and hyperparameter
-tuning}\label{training-and-hyperparameter-tuning}}
+tuning}\label{training-and-hyperparameter-tuning}
 
@@
 
-\hypertarget{model-evaluation-methods}{%
-\subsection{Model evaluation methods}\label{model-evaluation-methods}}
+\subsection{Model evaluation methods}\label{model-evaluation-methods}
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{optimized-hyperparameters}{%
-\subsection{Optimized hyperparameters}\label{optimized-hyperparameters}}
+\section{Results}\label{results}
+
+\subsection{Optimized hyperparameters}\label{optimized-hyperparameters}
 
@@
 compared to the VGG16 classifier, suggesting that the problem at hand
-was relatively straightforward. The optimized learning rates were
-typically smaller than the default value recommended by Keras, which is
-0.001.
+was less complex. The optimized learning rates were smaller than the
+default value recommended by Keras, which is 0.001.
 
@@
 
-\hypertarget{model-performance}{%
-\subsection{Model performance}\label{model-performance}}
+\subsection{Summary of model
+performance}\label{summary-of-model-performance}
 
@@
 error distribution had a large variance, all models tended to
-over-predict the distance. Therefore, Additionally, for input images
-representing null plots, it was expected that the models will
-over-predict the distance, as explained in Section
-\ref{null-distribution-of-the-approximated-distance}.
+over-predict the distance. Furthermore, for input images representing
+null plots, it was expected that the models will over-predict the
+distance, as explained in Section \ref{lineup-evaluation}.
 
@@
 plots solely exhibiting non-normality issues were the easiest to
-predict, with very low test root mean square error (RMSE) around
-\(0.3\), very low mean absolute error (MAE) around \(0.2\), and very
-high \(R^2\) around \(0.94\). Residual plots with non-linearity issues
-were more challenging to assess than those with heteroskedasticity or
-non-normality issues. Assessing residual plots with heteroskedasticity
-was not as difficult as assessing those with non-linearity issues. When
-multiple violations were introduced to a residual plot, the performance
-metrics typically lay between the metrics for each individual violation.
-
-Based on the model performance metrics, we will only use the
+predict, with very low test root mean square error (RMSE) at around
+\(0.3\). Residual plots with non-linearity issues were more challenging
+to assess than those with heteroskedasticity or non-normality issues.
+Assessing residual plots with heteroskedasticity was not as difficult as
+assessing those with non-linearity issues. When multiple violations were
+introduced to a residual plot, the performance metrics typically lay
+between the metrics for each individual violation.
+
+Based on the model performance metrics, we chose to use the
 best-performing model evaluated on the test set, namely the
@@
 
-\caption{\label{tab:performance}The training and test performance of three optimized models with different input sizes. The best metrics are colored in red.}
+\caption{\label{tab:performance}The training and test performance of three optimized models with different input sizes.}
 \centering
-\begin{tabular}[t]{l>{}l>{}l>{}l>{}l}
+\begin{tabular}[t]{lllll}
 \toprule
@@
 \multicolumn{5}{l}{\textbf{Training set}}\\
-\hspace{1em}$32 \times 32$ & \textcolor{black}{0.531} & \textcolor{black}{0.937} & \textcolor{black}{0.364} & \textcolor{black}{0.126}\\
-\hspace{1em}$64 \times 64$ & \textcolor{red}{0.405} & \textcolor{red}{0.963} & \textcolor{red}{0.260} & \textcolor{red}{0.072}\\
-\hspace{1em}$128 \times 128$ & \textcolor{black}{0.432} & \textcolor{black}{0.959} & \textcolor{black}{0.290} & \textcolor{black}{0.084}\\
+\hspace{1em}$32 \times 32$ & 0.531 & 0.937 & 0.364 & 0.126\\
+\hspace{1em}$64 \times 64$ & 0.405 & 0.963 & 0.260 & 0.072\\
+\hspace{1em}$128 \times 128$ & 0.432 & 0.959 & 0.290 & 0.084\\
 \addlinespace[0.3em]
 \multicolumn{5}{l}{\textbf{Test set}}\\
-\hspace{1em}$32 \times 32$ & \textcolor{red}{0.660} & \textcolor{red}{0.901} & \textcolor{red}{0.434} & \textcolor{red}{0.181}\\
-\hspace{1em}$64 \times 64$ & \textcolor{black}{0.674} & \textcolor{black}{0.897} & \textcolor{black}{0.438} & \textcolor{black}{0.186}\\
-\hspace{1em}$128 \times 128$ & \textcolor{black}{0.692} & \textcolor{black}{0.892} & \textcolor{black}{0.460} & \textcolor{black}{0.199}\\
+\hspace{1em}$32 \times 32$ & 0.660 & 0.901 & 0.434 & 0.181\\
+\hspace{1em}$64 \times 64$ & 0.674 & 0.897 & 0.438 & 0.186\\
+\hspace{1em}$128 \times 128$ & 0.692 & 0.892 & 0.460 & 0.199\\
 \bottomrule
@@
 
-\caption{\label{tab:performance-sub}The training and test performance of the $32 \times 32$ model presented with different model violations.  The best metrics are colored in red.}
+\caption{\label{tab:performance-sub}The training and test performance of the $32 \times 32$ model presented with different model violations.}
 \centering
 \resizebox{\linewidth}{!}{
-\begin{tabular}[t]{ll>{}l>{}l>{}l>{}l}
+\begin{tabular}[t]{lllll}
 \toprule
-Violations & \#samples & RMSE & $R^2$ & MAE & Huber loss\\
+Violations & \#training samples & Training RMSE & \#test samples & Test RMSE\\
 \midrule
-\addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Training set}}\\
-\hspace{1em}no violations & 1546 & \textcolor{black}{1.149} & \textcolor{black}{} & \textcolor{black}{1.080} & \textcolor{black}{0.592}\\
-\hspace{1em}non-linearity & 22184 & \textcolor{black}{0.625} & \textcolor{black}{0.851} & \textcolor{black}{0.474} & \textcolor{black}{0.177}\\
-\hspace{1em}heteroskedasticity & 10826 & \textcolor{black}{0.528} & \textcolor{black}{0.942} & \textcolor{black}{0.407} & \textcolor{black}{0.133}\\
-\hspace{1em}non-linearity + heteroskedasticity & 10221 & \textcolor{black}{0.593} & \textcolor{black}{0.920} & \textcolor{black}{0.439} & \textcolor{black}{0.158}\\
-\hspace{1em}non-normality & 10797 & \textcolor{red}{0.279} & \textcolor{red}{0.955} & \textcolor{red}{0.177} & \textcolor{red}{0.038}\\
-\hspace{1em}non-linearity + non-normality & 8835 & \textcolor{black}{0.472} & \textcolor{black}{0.923} & \textcolor{black}{0.274} & \textcolor{black}{0.090}\\
-\hspace{1em}heteroskedasticity + non-normality & 7870 & \textcolor{black}{0.354} & \textcolor{black}{0.905} & \textcolor{black}{0.211} & \textcolor{black}{0.052}\\
-\hspace{1em}non-linearity + heteroskedasticity + non-normality & 7721 & \textcolor{black}{0.432} & \textcolor{black}{0.874} & \textcolor{black}{0.261} & \textcolor{black}{0.078}\\
-\addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Test set}}\\
-\hspace{1em}no violations & 155 & \textcolor{black}{1.267} & \textcolor{black}{} & \textcolor{black}{1.174} & \textcolor{black}{0.685}\\
-\hspace{1em}non-linearity & 2218 & \textcolor{black}{0.787} & \textcolor{black}{0.771} & \textcolor{black}{0.578} & \textcolor{black}{0.257}\\
-\hspace{1em}heteroskedasticity & 1067 & \textcolor{black}{0.602} & \textcolor{black}{0.922} & \textcolor{black}{0.461} & \textcolor{black}{0.170}\\
-\hspace{1em}non-linearity + heteroskedasticity & 985 & \textcolor{black}{0.751} & \textcolor{black}{0.868} & \textcolor{black}{0.541} & \textcolor{black}{0.236}\\
-\hspace{1em}non-normality & 1111 & \textcolor{red}{0.320} & \textcolor{red}{0.942} & \textcolor{red}{0.194} & \textcolor{red}{0.049}\\
-\hspace{1em}non-linearity + non-normality & 928 & \textcolor{black}{0.600} & \textcolor{black}{0.879} & \textcolor{black}{0.335} & \textcolor{black}{0.133}\\
-\hspace{1em}heteroskedasticity + non-normality & 819 & \textcolor{black}{0.489} & \textcolor{black}{0.832} & \textcolor{black}{0.267} & \textcolor{black}{0.093}\\
-\hspace{1em}non-linearity + heteroskedasticity + non-normality & 717 & \textcolor{black}{0.620} & \textcolor{black}{0.762} & \textcolor{black}{0.339} & \textcolor{black}{0.140}\\
+no violations & 1546 & 1.149 & 155 & 1.267\\
+non-linearity & 22184 & 0.625 & 2218 & 0.787\\
+heteroskedasticity & 10826 & 0.528 & 1067 & 0.602\\
+non-linearity + heteroskedasticity & 10221 & 0.593 & 985 & 0.751\\
+non-normality & 10797 & 0.279 & 1111 & 0.320\\
+\addlinespace
+non-linearity + non-normality & 8835 & 0.472 & 928 & 0.600\\
+heteroskedasticity + non-normality & 7870 & 0.354 & 819 & 0.489\\
+non-linearity + heteroskedasticity + non-normality & 7721 & 0.432 & 717 & 0.620\\
 \bottomrule
@@
 
-\hypertarget{model-violations-index-mvi}{%
-\subsection{Model Violations Index
-(MVI)}\label{model-violations-index-mvi}}
+\subsection{Demostration of MVI}\label{demostration-of-mvi}
+
+As described in Section \ref{sec:model-violations-index} a Model
+Violations Index (MVI) can be constructed using \(\hat{D}\). This
+section provides guidelines and examples of MVI.
 
@@
 residual plots for fitted models exhibiting varying degrees of
-non-linearity and heteroskedasticity. Each residual plot's Model
-Violation Index (MVI) is computed using Equation \ref{eq:mvi} with
-\(C = 10\). When \(\text{MVI} > 8\), the visual patterns are notably
-strong and easily discernible by humans. In the range
-\(6 < \text{MVI} < 8\), the visibility of the visual pattern diminishes
-as MVI decreases. Conversely, when \(\text{MVI} < 6\), the visual
-pattern tends to become relatively faint and challenging to observe.
-Table \ref{tab:mvi} provides a summary of the MVI usage and it is
-applicable to other linear regression models.
+non-linearity and heteroskedasticity. Each residual plot's MVI is
+computed using Equation \ref{eq:mvi} with \(C = 10\). When
+\(\text{MVI} > 8\), the visual patterns are notably strong and easily
+discernible by humans. In the range \(6 < \text{MVI} < 8\), the
+visibility of the visual pattern diminishes as MVI decreases.
+Conversely, when \(\text{MVI} < 6\), the visual pattern tends to become
+relatively faint and challenging to observe. Table \ref{tab:mvi}
+provides a summary of the MVI usage and it is applicable to other linear
+regression models.
 
@@
 
-\hypertarget{comparison-with-human-visual-inference-and-conventional-tests}{%
 \subsection{Comparison with human visual inference and conventional
-tests}\label{comparison-with-human-visual-inference-and-conventional-tests}}
-
-\hypertarget{overview-of-the-human-subject-experiment}{%
+tests}\label{comparison-with-human-visual-inference-and-conventional-tests}
+
 \subsubsection{Overview of the human subject
-experiment}\label{overview-of-the-human-subject-experiment}}
+experiment}\label{overview-of-the-human-subject-experiment}
 
@@
 
-\hypertarget{model-performance-on-the-human-data}{%
 \subsubsection{Model performance on the human
-data}\label{model-performance-on-the-human-data}}
+data}\label{model-performance-on-the-human-data}
 
@@
 \(D\) for the null plots is zero. We have used our optimized computer
-vision model to predict \(\hat{D}\) for both the actual residual plots
-and the null plots. Additionally, the appropriate conventional tests
-including the Ramsey Regression Equation Specification Error Test
-(RESET) \citep{ramsey1969tests} for non-linearity and the Breusch-Pagan
-test \citep{breusch1979simple} for heteroskedasticity were applied on
-the same data for comparison.
+vision model to predict the approximated distance for both the actual
+residual plots and the null plots. To have a fair comparison, \(H_0\)
+will be rejected if the actual residual plot has the greatest
+approximated distance among all plots in a lineup. Additionally, the
+appropriate conventional tests including the Ramsey Regression Equation
+Specification Error Test (RESET) \citep{ramsey1969tests} for
+non-linearity and the Breusch-Pagan test \citep{breusch1979simple} for
+heteroskedasticity were applied on the same data for comparison.
 
@@
 
-For null plots, a histogram of the predictions is provided in Figure
-\ref{fig:hist-null-human}. In most cases, \(\hat{D}\) is predicted to be
-around \(1\). As \(\hat{D}\) increases, the number of predictions
-decreases. The 90\% and 95\% sample quantiles are approximately at
-\(2\), and the 99\% sample quantile is around \(2.5\). These sample
-quantiles can serve as global critical values for determining whether a
-residual plot needs to be rejected. We have also computed the 95\%
-sample quantiles by assessing 200 null plots for each fitted model used
-in the experiment. The difference between decisions based on these two
-types of critical values is minor, with decisions aligning 93.8\% of the
-time. Therefore, we will be using the 95\% sample quantiles obtained
-from each fitted model in the following analysis.
-
-Likewise, Figure \ref{fig:conv-mosaic} and Table \ref{tab:conv-table}
-provide a summary of the agreement between decisions made by the
-computer vision model and conventional tests. The agreement rates
-between conventional tests and the computer vision model are 65.37\% and
-57.29\% for residual plots containing heteroskedasticity and
-non-linearity patterns, respectively. These figures are lower than those
-presented in Table \ref{tab:human-table}, indicating that the computer
-vision model exhibits behaviour more akin to visual tests than
-conventional tests. Additionally, conventional tests tend to reject
+Table \ref{tab:human-conv-table} provides a summary of the agreement
+between decisions made by the computer vision model and conventional
+tests. The agreement rates between conventional tests and the computer
+vision model are 85.95\% and 79.69\% for residual plots containing
+heteroskedasticity and non-linearity patterns, respectively. These
+figures are higher than those calculated for visual tests conducted by
+human, indicating that the computer vision model exhibits behaviour more
+akin to the best available conventional tests. However, Figure
+\ref{fig:conv-mosaic} shows that conventional tests tend to reject
 residual plots almost exclusively when the computer vision model does,
@@
 Figure \ref{fig:power}, several observations emerge. Firstly, for the
-computer vision model, when the distance \(D > 3\), almost all residual
+computer vision model, when the distance \(D > 4\), almost all residual
 plots are rejected. However, for visual tests conducted by humans, the
@@
 \(\text{log}(0.05/0/95)\). The fitting curves for the computer vision
-model fall between those of conventional tests and visual tests for
-non-linearity and are slightly below conventional tests for
-heteroskedasticity. There is still potential to refine the computer
-vision model to better align its behaviour with visual tests.
+model fall between those of conventional tests and visual tests for both
+non-linearity and heteroskedasticity. There is still potential to refine
+the computer vision model to better align its behaviour with visual
+tests.
 
@@
 
-\begin{figure}[!h]
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/hist-null-human-1}
-
-}
-
-\caption{Histogram of predicted distance on null plots used in the human subject experiment. Sample quantiles are drawn in different colors.}\label{fig:hist-null-human}
-\end{figure}
-
 \begin{table}
 
-\caption{\label{tab:human-table}Summary of decisions made by visual tests conducted by human and decisions made by the computer vision model.}
+\caption{\label{tab:human-conv-table}Summary of the comparison of decisions made by computer vision model with decisions made by conventional tests and visual tests conducted by human.}
 \centering
@@
 \midrule
-heteroskedasticity & 540 & 367 & 0.6796\\
-non-linearity & 576 & 385 & 0.6684\\
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Compared with conventional tests}}\\
+\hspace{1em}heteroskedasticity & 540 & 464 & 0.8593\\
+\hspace{1em}non-linearity & 576 & 459 & 0.7969\\
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Compared with visual tests conducted by human}}\\
+\hspace{1em}heteroskedasticity & 540 & 367 & 0.6796\\
+\hspace{1em}non-linearity & 576 & 385 & 0.6684\\
 \bottomrule
@@
 
-\begin{table}
-
-\caption{\label{tab:conv-table}Summary of decisions made by conventional tests and decisions made by the computer vision model.}
-\centering
-\begin{tabular}[t]{llll}
-\toprule
-Violations & \#Samples & \#Agreements & Agreement rate\\
-\midrule
-heteroskedasticity & 540 & 464 & 0.6537\\
-non-linearity & 576 & 459 & 0.5729\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{figure}[!h]
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/human-mosaic-1}
-
-}
-
-\caption{Rejection rate ($p$-value $\leq0.05$) of visual test conditional on the computer vision model decision on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot.
-The visual test rejects less frequently than the computer vision model, and (almost) only rejects when the computer vision model does.}\label{fig:human-mosaic}
-\end{figure}
-
 \begin{figure}[!h]
@@
 
-\caption{Comparison of decisions made by visual tests, conventional tests and the computer vision model.
-The data points are fitted with logistic regression models with no intercept but an offset equals to $\text{log}(0.05/0.95)$.}\label{fig:power}
+\caption{Comparison of power of visual tests, conventional tests and the computer vision model.
+Marks along the x-axis at the bottom of the plot represent rejections made by each type of test.
+Marks at the top of the plot represent acceptances.
+Power curves are fitted by logistic regression models with no intercept but an offset equals to $\text{log}(0.05/0.95)$.}\label{fig:power}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/pcp-1}
+
+}
+
+\caption{Parallel coordinate plots of decisions made by computer vision model, conventional tests and visual tests made by human.}\label{fig:pcp}
+\end{figure}
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/venn-1}
+
+}
+
+\caption{Venn diagram for non-rejection decisions made by conventional tests, computer vision model, and visual tests conducted by human.}\label{fig:venn}
 \end{figure}
@@
 
-\hypertarget{data-examples}{%
-\subsection{Data examples}\label{data-examples}}
+\subsection{Data examples}\label{data-examples}
 
@@
 
-\hypertarget{left-triangle}{%
-\subsubsection{Left-triangle}\label{left-triangle}}
+\subsubsection{Left-triangle}\label{left-triangle}
 
@@
 \begin{Highlighting}[]
-\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{check}\NormalTok{(}\AttributeTok{null\_draws =}\NormalTok{ 200L, }\AttributeTok{boot\_draws =}\NormalTok{ 200L)}
+\NormalTok{checker}\SpecialCharTok{$}\FunctionTok{check}\NormalTok{(}\AttributeTok{null\_draws =} \DecValTok{200}\DataTypeTok{L}\NormalTok{, }\AttributeTok{boot\_draws =} \DecValTok{200}\DataTypeTok{L}\NormalTok{)}
 \NormalTok{checker}\SpecialCharTok{$}\FunctionTok{summary\_plot}\NormalTok{()}
@@
 
-\hypertarget{boston-housing}{%
-\subsubsection{Boston housing}\label{boston-housing}}
+\subsubsection{Boston housing}\label{boston-housing}
 
@@
 
-\hypertarget{datasaurus}{%
-\subsubsection{Datasaurus}\label{datasaurus}}
+\subsubsection{Datasaurus}\label{datasaurus}
 
@@
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
+\section{Conclusion}\label{conclusion}
 
@@
 
-\hypertarget{acknowledgements}{%
-\section*{Acknowledgements}\label{acknowledgements}}
+\section*{Acknowledgements}\label{acknowledgements}
 \addcontentsline{toc}{section}{Acknowledgements}

### 2024-04-15 paper/data/smoothn.py

@@ -1,0 +1,354 @@
+# Robust smoothing algorithm with missing values
+# From: https://github.com/SanParraguez/smoothn/blob/main/smoothn.py
+# Author: Santiago Parraguez
+# 
+# This code is a Python implementation of the smoothn(y) algorithm 
+# described by García (2010). The original source code (Matlab) can 
+# be found at https://www.biomecardio.com/en/download.html
+# licensed under the GNU General Public License v3.0
+#
+# Reference:
+# Garcia, D. (2010). Robust smoothing of gridded data in one and higher 
+# dimensions with missing values. Computational Statistics & Data Analysis; 
+# 54:1167-1178. doi:10.1016/j.csda.2009.09.020.
+
+
+# ====== IMPORT LIBRARIES ======
+
+import numpy as np
+from scipy.fft import dctn, idctn
+from scipy.optimize import minimize
+from scipy.ndimage.morphology import distance_transform_edt
+
+
+# ================================================================================================
+def smoothn(y, s=None, tolz=1e-3, z0=None, w=None, di=None,
+            robust=False, max_iter=1000, robust_iter=3, workers=1,
+            return_s=False):
+    """
+    Fast smooths an array of data based on the cosine discrete transform. Allows to choose the smoothing parameter,
+    or calculate it using a GCVscore.
+    Also, it implements a robust iteration (False by default) to ignore outlier points.
+
+    Parameters
+    ----------
+    y : np.ma.MaskedArray or np.ndarray
+        Data to be smoothed and filled.
+    s : None or float
+        Smoothing parameter, calculated automatically if not provided to minimize gcv score.
+    tolz : float
+        Tolerance of iteration over z (prediction of data).
+    z0 : np.ndarray
+        Initial guess for z, calculated with nearest neighbor if not provided
+    w : np.ndarray
+        Weights array, if not provided assumes all data has the same confidence
+    di : tuple
+        Grid dimensions, assumed regular if not given
+    robust : bool
+        Indicates if the robust iteration is executed
+    workers : int
+        Number of workers assigned to Scipy functions `fft.dctn` and `fft.idctn`
+    max_iter : int
+        Maximum number of iterations. Default: 1,000
+    robust_iter : int
+        Number of robust iterations to perform. Notice that the max_iter value works inside the number of robust
[…]

### 2024-04-22 paper/cnn.tex
update diagram

\documentclass[border=15pt, multi, tikz]{standalone}
%\usepackage{blocks}
\usepackage{import}
\subimport{layers/}{init}
\usetikzlibrary{positioning}
\usetikzlibrary{3d}

\def\ConvColor{rgb:yellow,5;red,2.5;white,5}
\def\ConvReluColor{rgb:yellow,5;red,5;white,5}
\def\PoolColor{rgb:red,1;black,0.3}
\def\FcColor{rgb:blue,5;red,2.5;white,5}
\def\FcReluColor{rgb:blue,5;red,5;white,4}
\def\SoftmaxColor{rgb:magenta,5;black,7}
\def\SumColor{rgb:blue,5;green,15}

\begin{document}
\begin{tikzpicture}
\tikzstyle{connection}=[ultra thick,every node/.style={sloped,allow upside down},draw=\edgecolor,opacity=0.7]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Draw Layer Blocks
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\node[canvas is zy plane at x=0] (temp) at (-3,0,0) {\includegraphics[width=8cm,height=8cm]{layers/1793.png}};

% conv1_1,conv1_2
\pic[shift={(0,0,0)}] at (0,0,0) {RightBandedBox={name=cr1,caption=Convolutional block 1,%
        xlabel={{"32","32"}},ylabel=32,zlabel=32,fill=\ConvColor,bandfill=\ConvReluColor,%
        height=40,width={2,2},depth=40}};
%pool1
\pic[shift={(0,0,0)}] at (cr1-east) {Box={name=p1,%
        fill=\PoolColor,opacity=0.5,height=35,width=1,depth=35}};
%%%%%%%%%%
% conv2_1,conv2_2
\pic[shift={(2,0,0)}] at (p1-east) {RightBandedBox={name=cr2,caption=Convolutional block 2,%
        xlabel={{"64","64"}},zlabel=16,fill=\ConvColor,bandfill=\ConvReluColor,%
        height=35,width={3,3},depth=35}};
%pool2
\pic[shift={(0,0,0)}] at (cr2-east) {Box={name=p2,%
        fill=\PoolColor,opacity=0.5,height=30,width=1,depth=30}};
%%%%%%%%%%
% conv3_1,conv3_2
\pic[shift={(2,0,0)}] at (p2-east) {RightBandedBox={name=cr3,caption=Convolutional block 3,%
        xlabel={{"128","128","128"}},zlabel=8,fill=\ConvColor,bandfill=\ConvReluColor,%
        height=30,width={4,4,4},depth=30}};
%pool3
\pic[shift={(0,0,0)}] at (cr3-east) {Box={name=p3,%
        fill=\PoolColor,opacity=0.5,height=23,width=1,depth=23}};
%%%%%%%%%%
% conv4_1,conv4_2,conv4_3
\pic[shift={(1.8,0,0)}] at (p3-east) {RightBandedBox={name=cr4,caption=Convolutional block 4,%
        xlabel={{"256","256","256"}},zlabel=4,fill=\ConvColor,bandfill=\ConvReluColor,%
        height=23,width={7,7,7},depth=23}};
%pool4
\pic[shift={(0,0,0)}] at (cr4-east) {Box={name=p4,%
        fill=\PoolColor,opacity=0.5,height=15,width=1,depth=15}};
%%%%%%%%%%
% conv5_1,conv5_2,conv5_3
\pic[shift={(1.5,0,0)}] at (p4-east) {RightBandedBox={name=cr5,caption=Convolutional block 5,%
        xlabel={{"256","256","256"}},zlabel=2,fill=\ConvColor,bandfill=\ConvReluColor,%
        height=15,width={7,7,7},depth=15}};
%pool5
\pic[shift={(0,0,0)}] at (cr5-east) {Box={name=p5,%
        fill=\PoolColor,opacity=0.5,height=10,width=1,depth=10}};
%%%%%%%%%%
% fc6
\pic[shift={(3,0,0)}] at (p5-east) {RightBandedBox={name=fc6,caption=FC1,%
        xlabel={{"1",""}},zlabel=256,fill=\FcColor,bandfill=\FcReluColor,%
        height=3,width=3,depth=100}};

\pic[shift={(3.5,0,15)}] at (p5-east) {Ball={name=elt1,%
fill=\SumColor,opacity=0.6,%
radius=2.5,logo=$+$}};

\pic[shift={(3,0,20)}] at (p5-east) {Box={name=fc66,caption=Auxiliary input,%
        xlabel={{"1",""}},zlabel=5,fill=\FcColor,%
        height=3,width=3,depth=10}};

%%%%%%%%%%
% fc7
\pic[shift={(2,0,0)}] at (fc6-east) {RightBandedBox={name=fc7,caption=FC2,%
        xlabel={{"1","dummy"}},zlabel=256,fill=\FcColor,bandfill=\FcReluColor,%
        height=3,width=3,depth=100}};
%%%%%%%%%%
% fc8
\pic[shift={(1.5,0,0)}] at (fc7-east) {RightBandedBox={name=fc8,caption=FC3 + relu,%
        xlabel={{"1","dummy"}},fill=\FcColor,bandfill=\FcReluColor,%
        height=3,width=3,depth=5}};

%%%%%%%%%%
% softmax
\pic[shift={(0,0,0)}] at (fc8-east) {Box={name=softmax,%
        xlabel={{"","dummy"}},zlabel=1,opacity=0.8,fill=\SoftmaxColor,%
        height=3,width=1.5,depth=5}};

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Draw Arrow Connections
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\draw [connection]  (p1-east)        -- node {\midarrow} (cr2-west);
\draw [connection]  (p2-east)        -- node {\midarrow} (cr3-west);
\draw [connection]  (p3-east)        -- node {\midarrow} (cr4-west);
\draw [connection]  (p4-east)        -- node {\midarrow} (cr5-west);
\draw [connection]  (p5-east)        -- node {\midarrow} (fc6-west);
\draw [connection]  (p5-east)        -- node {\midarrow} (fc6-west);
\draw [connection]  (fc6-east)       -- node {\midarrow} (fc7-west);
\draw [connection]  (fc7-east)       -- node {\midarrow} (fc8-west);
\draw [connection]  (fc66-east)      -- node {} (fc6-near);
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Draw Dotted Edges 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\draw[densely dashed]
    (fc6-west)++(0, 1.5*.2, 1.5*.2) coordinate(a) -- (p5-nearnortheast)
    (fc6-west)++(0,-1.5*.2, 1.5*.2) coordinate(b) -- (p5-nearsoutheast)
    (fc6-west)++(0,-1.5*.2,-1.5*.2) coordinate(c) -- (p5-farsoutheast)
    (fc6-west)++(0, 1.5*.2,-1.5*.2) coordinate(d) -- (p5-farnortheast)

    (a)--(b)--(c)--(d)
    ;
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{tikzpicture}
\end{document}

### 2024-04-22 paper/paper.Rmd
update diagram

@@
 
+```{r out.width = "100%", fig.cap = "Diagram of the architecture of the optimized computer vision model."}
+knitr::include_graphics("cnn.pdf")
+```
+
 The architecture of the computer vision model is adapted from a well-established architecture known as VGG16, which has demonstrated high performance in image classification [@simonyan2014very].
@@
   geom_pcp_boxes(aes_pcp(), boxwidth = 0.1) +
-  facet_wrap(~type) +
+  facet_wrap(~type, scales = "free_y") +
   theme_light() +
@@
 
+# Extension to other types of regression model
+
+- Use raw residuals.
+  - D for null plots will be greater than zero.
+  - Can not assume D for actual plots are greater than D for null plots any more.
+  - Need to use a two-sided test instead of a one-sided test.
+  - It may fail completely due to the identification issue.
+- Use transformed residuals that roughly follow a multivariate normal distribution.
+  - The one-sided test will still work. 
+  - Need to ensure that the transformed actual plot will have residual distribution different from the transformed null distribution. 
+- Train a new model.
+  - KL divergence may not be the appropriate measure for the difference between the residual distributions. Could consider using a new distance measure.
+  -
+
 # Conclusion

### 2024-04-29 paper/cnn.tex
revise up to model arch

@@
 \def\FcColor{rgb:blue,5;red,2.5;white,5}
-\def\FcReluColor{rgb:blue,5;red,5;white,4}
+\def\FcReluColor{rgb:blue,5;red,5;white,5}
 \def\SoftmaxColor{rgb:magenta,5;black,7}
-\def\SumColor{rgb:blue,5;green,15}
+\def\AuxColor{rgb:blue,5;green,15;white,5}
+\def\AuxReluColor{rgb:blue,5;green,7.5;white,5}
 
@@
 % fc6
-\pic[shift={(3,0,0)}] at (p5-east) {RightBandedBox={name=fc6,caption=FC1,%
-        xlabel={{"1",""}},zlabel=256,fill=\FcColor,bandfill=\FcReluColor,%
+\pic[shift={(3,0,0)}] at (p5-east) {Box={name=fc6,caption=Concatenation layer,%
+        xlabel={{"1",""}},zlabel=256,fill=\FcColor,%
         height=3,width=3,depth=100}};
 
-\pic[shift={(3.5,0,15)}] at (p5-east) {Ball={name=elt1,%
-fill=\SumColor,opacity=0.6,%
-radius=2.5,logo=$+$}};
-
-\pic[shift={(3,0,20)}] at (p5-east) {Box={name=fc66,caption=Auxiliary input,%
-        xlabel={{"1",""}},zlabel=5,fill=\FcColor,%
+\pic[shift={(3,0,-11)}] at (p5-east) {Box={name=fc66,caption=Auxiliary input,%
+        xlabel={{"",""}},zlabel=5,fill=\AuxColor,%
         height=3,width=3,depth=10}};
@@
 % fc7
-\pic[shift={(2,0,0)}] at (fc6-east) {RightBandedBox={name=fc7,caption=FC2,%
+\pic[shift={(2,0,0)}] at (fc6-east) {RightBandedBox={name=fc7,caption=Dense layer,%
         xlabel={{"1","dummy"}},zlabel=256,fill=\FcColor,bandfill=\FcReluColor,%
@@
 % fc8
-\pic[shift={(1.5,0,0)}] at (fc7-east) {RightBandedBox={name=fc8,caption=FC3 + relu,%
+\pic[shift={(1.5,0,0)}] at (fc7-east) {RightBandedBox={name=fc8,caption=Output layer,%
         xlabel={{"1","dummy"}},fill=\FcColor,bandfill=\FcReluColor,%
         height=3,width=3,depth=5}};
-
-%%%%%%%%%%
-% softmax
-\pic[shift={(0,0,0)}] at (fc8-east) {Box={name=softmax,%
-        xlabel={{"","dummy"}},zlabel=1,opacity=0.8,fill=\SoftmaxColor,%
-        height=3,width=1.5,depth=5}};
 
@@
 \draw [connection]  (fc7-east)       -- node {\midarrow} (fc8-west);
-\draw [connection]  (fc66-east)      -- node {} (fc6-near);
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-04-29 paper/paper.Rmd
revise up to model arch

@@
   \usepackage{setspace}
-  \doublespacing
 output: rticles::tf_article
@@
 
-<!-- \newpage -->
-<!-- \tableofcontents -->
-<!-- \newpage -->
-
 # Introduction
@@
 However, accurately interpreting a residual plot can be challenging.
+A residual plot can exhibit various visual features, but it is crucial to recognize that some may arise from the characteristics of predictors and the inherent randomness of the error, rather than indicating a violation of model assumptions [@li2023plot].
 Consider Figure \ref{fig:false-finding} as an example, the residuals display a triangular shape pointing to the left.
@@
 
-A residual plot can exhibit various visual features, but it is crucial to recognize that some may arise from the characteristics of predictors and the inherent randomness of the error, rather than indicating a violation of model assumptions [@li2023plot].
 The concept of visual inference, as proposed by @buja2009statistical, provides an inferential framework to assess whether residual plots indeed contain visual patterns inconsistent with the model assumptions.
-The fundamental idea involves testing whether the actual residual plot visually differs significantly from null plots, where null plots are plotted with residuals generated from the residual rotation distribution [@langsrud2005rotation], which is a distribution consistent with the null hypothesis $H_0$ that the linear regression model is correctly specified.
+The fundamental idea involves testing whether the true residual plot visually differs significantly from null plots, where null plots are plotted with residuals generated from the residual rotation distribution [@langsrud2005rotation], which is a distribution consistent with the null hypothesis $H_0$ that the linear regression model is correctly specified.
 Typically, the visual test is accomplished through the lineup protocol, where the real residual plot is embedded within a lineup alongside several null plots.
@@
 These measures, including, but not limited to, "Outlying", "Skinny", "Stringy", "Straight", "Monotonic", "Skewed", "Clumpy", and "Striated" aimed to characterize outliers, shape, density, trend, coherence and other characteristics of the data.
-While this approach has been inspiring, there is a recognition [@buja2009statistical] that it may not capture all the necessary visual features that differentiate actual residual plots from null plots.
+While this approach has been inspiring, there is a recognition [@buja2009statistical] that it may not capture all the necessary visual features that differentiate true residual plots from null plots.
 A more promising alternative entails enabling machines to learn the function for extracting visual features from residual plots.
@@
 
-## Different possible configurations of the model formula
+## Model Formulas
 
 There are various configurations of the computer vision model that can be used to assess residual plots.
-We discuss these configurations below based on two key components of the model formula: the input and the output format.
-
-### Input formats
+We discuss these configurations below focusing on two key components of the model formula: the input and the output format.
+
+### Input Formats
 
@@
 @chopra2005learning have shown that computer vision models designed for image comparison can assess whether a pair of images are similar or dissimilar.
-Applied to our specific problem, we can define null plots of a fitted regression model to be similar to each other, while considering actual residual plots to be distinct from null plots of any fitted regression model.
+Applied to our specific problem, we can define null plots of a fitted regression model to be similar to each other, while considering true residual plots to be distinct from null plots of any fitted regression model.
 A triplet constitutes a set of three images, denoted as $image_1$, $image_2$ and $image_3$.
@@
 
-We did not explore all the mentioned input formats due the considerable costs associated with data preparation and model training.
-Considering the implementation cost and the interpretability of the model, we settled on the single residual plot input format.
-
-### Output formats
+We did not explore all the mentioned input formats due to the significant costs tied to data preparation and model training.
+Taking into account the implementation cost and the need for model interpretability, we settled on the single residual plot input format.
+
+### Output Formats
 
@@
 
-Alternatively, the output could be a meaningful and interpretable numerical measure useful for assessing residual plots, such as the strength of suspicious visual patterns reflecting the extent of model violations and the difficulty index for identifying whether a residual plot has no issues.
+Alternatively, the output could be a meaningful and interpretable numerical measure useful for assessing residual plots, such as the strength of suspicious visual patterns reflecting the extent of model violations, or the difficulty index for identifying whether a residual plot has no issues.
 However, these numeric measures are often informally used in daily communication but are not typically formalized or rigorously defined.
@@
 
-## Distance from the good residual plots
+## Distance Between Residual Plots
 
@@
 
-### Residual distribution
+### Residual Distribution
 
 The distance measure proposed in this study takes into account the fact that we tried to measure how different a residual plot is from a good residual plot, or in other words, how different a given fitted regression model is from a correctly specified model.
-For the classical normal linear regression model, residuals $\boldsymbol{e}$ are derived from the fitted values $\hat{\boldsymbol{y}}$ and observed values $\boldsymbol{y}$.
+
+For a classical normal linear regression model, residuals $\boldsymbol{e}$ are derived from the fitted values $\hat{\boldsymbol{y}}$ and observed values $\boldsymbol{y}$.
 Suppose the data generating process is known and the regression model is correctly specified, by the Frisch-Waugh-Lowell theorem [@frisch1933partial], residuals $\boldsymbol{e}$ can also be treated as random variables and written as a linear transformation of the error $\boldsymbol{\varepsilon}$ formulated as $\boldsymbol{e} = \boldsymbol{R}\boldsymbol{\varepsilon}$, where $\boldsymbol{R}=\boldsymbol{I}_n -\boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'$ is the residual operator and an idempotent matrix, $\boldsymbol{X}$ is the design matrix, $\boldsymbol{I}_n$ is a $n$ by $n$ identity matrix, and $n$ is the number of observations.
@@
 
-In ordinary least square, the minimization of the sum of square residuals implies  $\sum_{i=1}^{n} e_i = 0$, making any residual value to be a linear combination of the remaining $n - 1$ residuals.
-This effectively means $\text{rank}(\boldsymbol{R}) = n - 1 < n$ and $Q$ is a degenerate multivariate distribution.
+Suppose the design matrix $\boldsymbol{X}$ has linearly independent columns, the trace of the hat matrix will equal to the number of columns in $\boldsymbol{X}$ denoted as $k$.
+As a result, the rank of $\boldsymbol{R}$ is $n - k$, and $Q$ is a degenerate multivariate distribution.
 To capture the characteristics of $Q$, such as moments, we can simulate a large numbers of $\boldsymbol{\varepsilon}$ and transform it to $\boldsymbol{e}$ to get the empirical estimates.
@@
 However, if the model is misspecified, then the actual distribution of residuals denoted as $P$, will be different from $Q$.
-For example, if the data generating process contains variables correlated with any column of $\boldsymbol{X}$ but not included in $\boldsymbol{X}$, causing an omitted variable problem, $P$ will be different from $Q$ because the residual operator obtained from the fitted regression model will not be the same as $\boldsymbol{R}$.
+For example, if the data generating process contains variables correlated with any column of $\boldsymbol{X}$ but missing from $\boldsymbol{X}$, causing an omitted variable problem, $P$ will be different from $Q$ because the residual operator obtained from the fitted regression model will not be the same as $\boldsymbol{R}$.
 Besides, if the $\boldsymbol{\varepsilon}$ follows a non-normal distribution such as a multivariate log-normal distribution, $P$ will usually be skewed and has a long tail.
 
-### Kullback-Leibler divergence of $P$ from $Q$
+### Distance of $P$ from $Q$
 
 Define a proper distance between distributions is usually easier than define a proper distance between data plots.
-Given the actual residual distribution $Q$ and the reference residual distribution $P$, we used a distance measure based on Kullback-Leibler divergence [@kullback1951information] to quantify the difference between two distributions
+Given the true residual distribution $Q$ and the reference residual distribution $P$, we used a distance measure based on Kullback-Leibler divergence [@kullback1951information] to quantify the difference between two distributions
 
@@
 
-\noindent where $p(.)$ is the probability density function for distribution $P$, and $q(.)$ is the probability density function for distribution $Q$.
+\noindent where $p(.)$ and $q(.)$ are the probability density functions for distribution $P$ and distribution $Q$ respectively.
 
@@
 It was mainly designed for measuring the effect size of non-linearity and heteroskedasticity in a residual plot.
-@li2023plot have showed that, for a classical normal linear regression model that omits a necessary higher-order predictors $\boldsymbol{Z}$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$ with $\boldsymbol{V}$ being an arbitrary symmetric positive semi-definite matrix, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}))$.
+@li2023plot have derived that, for a classical normal linear regression model that omits a necessary higher-order predictors $\boldsymbol{Z}$, and incorrectly assumes $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n,\sigma^2\boldsymbol{I}_n)$ while in fact $\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}_n, \boldsymbol{V})$ with $\boldsymbol{V}$ being an arbitrary symmetric positive semi-definite matrix, $Q$ can be represented as $N(\boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z, \text{diag}(\boldsymbol{R}\boldsymbol{V}\boldsymbol{R}))$.
 Note that the variance-covariance matrix is replaced with the diagonal matrix to ensure it is a full-rank matrix.
@@
 \label{eq:kl-2}
-D_{KL} &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + \text{tr}(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
+D_{KL} &= \frac{1}{2}\left(\log\frac{|\text{diag}(\boldsymbol{W})|}{|\text{diag}(\boldsymbol{R}\sigma^2)|} - n + tr(\text{diag}(\boldsymbol{W})^{-1}\text{diag}(\boldsymbol{R}\sigma^2)) + \boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{W}))^{-1}\boldsymbol{\mu}_z\right),
 \end{align}
@@
 \noindent where $\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z$, and $\boldsymbol{W} = \boldsymbol{R}\boldsymbol{V}\boldsymbol{R}$.
-The assumed error variance $\sigma^2$ is set to be $\text{tr}(\boldsymbol{V})/n$, which is the expectation of the estimated variance.
-
-### Evaluation of Kullback-Leibler divergence for non-normal $P$
-
-For non-normal error $\boldsymbol{\varepsilon}$, the actual residual distribution $P$ is unlikely to be a multivariate normal distribution.
+The assumed error variance $\sigma^2$ is set to be $tr(\boldsymbol{V})/n$, which is the expectation of the estimated variance.
+
+### Non-normal $P$
+
+For non-normal error $\boldsymbol{\varepsilon}$, the true residual distribution $P$ is unlikely to be a multivariate normal distribution.
 Thus, equation \ref{eq:kl-2} given in @li2023plot will not be applicable to models violating the normality assumption.
@@
 
-### Approximation of the distance measure
-
-In the previous sections, we have defined a distance measure given in equation \ref{eq:kl-0} for quantifying the difference between the actual residual distribution $P$ and an ideal reference distribution $Q$.
+## Distance Estimation
+
+In the previous sections, we have defined a distance measure given in equation \ref{eq:kl-0} for quantifying the difference between the true residual distribution $P$ and an ideal reference distribution $Q$.
 You may have noticed that this distance measure can only be computed when the data generating process is known.
@@
 
-We tried to train a computer vision model to approximate this distance measure with a residual plot. Let $D$ be the result of equation \ref{eq:kl-0}, and our estimator $\hat{D}$ is formulated as
+We tried to train a computer vision model to estimate this distance measure with a residual plot. Let $D$ be the result of equation \ref{eq:kl-0}, and our estimator $\hat{D}$ is formulated as
 
@@
 
-\noindent where $V_{h \times w}(.)$ is a plotting function that saves a residuals vs fitted values plot with fixed aesthetic as an image with $h \times w$ pixels and three colour channels, $f_{CV}(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance in the domain $[0, +\infty)$.
-
-With the approximated distance $\hat{D}$, we will be able to know how different the underlying distribution of the residuals is from a good residual distribution.
-The approximated distance $\hat{D}$ can also be used as an index of the model violations.
-It also provides information for the strength of the visual signal embedded in the residual plot.
-
-The approximated distance $\hat{D}$ is not expected to be the same as the original distance $D$.
+\noindent where $V_{h \times w}(.)$ is a plotting function that saves a residuals vs fitted values plot with fixed aesthetic as an RGB image with $h \times w$ pixels, $f_{CV}(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance in the domain $[0, +\infty)$.
+
+With the estimated distance $\hat{D}$, we will be able to know how different the underlying distribution of the residuals is from a good residual distribution.
+$\hat{D}$ can also be used as an index of the model violations indicating the strength of the visual signal embedded in the residual plot.
+
+It is not expected that $\hat{D}$ will be equal to original distance $D$.
 This is largely because information contained in a single residual plot is limited and it may not be able to summarise all the important characteristics of the residual distribution.
-For a given residual distribution $P$, we can generate many different residual plots.
-Some of them share similar visual patterns, but some of them could be visually very different from the rest, especially for regression models with small $n$.
-This suggests the error of the approximation will vary depends on whether the observed residual plot is representative or not.
-
-### Model Violations Index (MVI) {#sec:model-violations-index}
-
-In Section \ref{output-formats}, we discuss the output formats of the computer vision model presented in this paper.
-One notable advantage highlighted is its interpretability.
-The approximated distance $\hat{D}$ is an estimator of the difference between the actual residual distribution and the reference residual distribution.
+For a given residual distribution $P$, many different residual plots can be simulated, where many will share similar visual patterns, but some of them could be visually very different from the rest, especially for regression models with small $n$.
+This suggests the error of the estimation will vary depends on whether the input residual plot is representative or not.
+
+## Model Violations Index (MVI) {#sec:model-violations-index}
+
+$\hat{D}$ is an estimator of the difference between the true residual distribution and the reference residual distribution.
 This difference primarily arises from deviations in model assumptions.
-The magnitude of $D$ directly reflects the degree of these deviations, thus making $\hat{D}$ instrumental in forming a model violation index.
-
-Note that the Kullback-Leibler divergence might be influenced by the number of random variables involved in its evaluation.
+The magnitude of $D$ directly reflects the degree of these deviations, thus making $\hat{D}$ instrumental in forming a model violations index.
+
+Note that Equation \ref{eq:kl-0} might be influenced by the number of random variables involved in its evaluation.
 Generally, a larger number of observations will lead to a greater distance $D$.
@@
 
-## Statistical testing based on the approximated distance
-
-### Lineup evaluation
+Figures \ref{fig:poly-index} and \ref{fig:heter-index} display the residual plots for fitted models exhibiting varying degrees of non-linearity and heteroskedasticity.
+Each residual plot's MVI is computed using Equation \ref{eq:mvi} with $C = 10$.
+When $\text{MVI} > 8$, the visual patterns are notably strong and easily discernible by humans.
+In the range $6 < \text{MVI} < 8$, the visibility of the visual pattern diminishes as MVI decreases.
+Conversely, when $\text{MVI} < 6$, the visual pattern tends to become relatively faint and challenging to observe.
+Table \ref{tab:mvi} provides a summary of the MVI usage and it is applicable to other linear regression models.
+
+```{r mvi}
+data.frame(degree = c("Strong", "Moderate", "Weak"),
+           range = c("$\\text{MVI} > 8$", "$6 < \\text{MVI} < 8$", "$\\text{MVI} < 6$")) %>%
+  kableExtra::kable("latex", 
+                    booktabs = TRUE,
+                    escape = FALSE,
+                    col.names = c("Degree of model violations", "Range ($C$ = 10)"),
+                    caption = "Degree of model violations or the strength of the visual signals according to the Model Violation Index (MVI). The constant $C$ is set to be 10.")
+```
+
+```{r poly-index, fig.pos = "!h", fig.cap = "Residual plots generated from fitted models exhibiting varying degrees of non-linearity violations.
+The Model violations index (MVI) is displayed atop each residual plot.
+The non-linearity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$."}
+set.seed(1495)
+if (!file.exists(here::here("paper/data/poly_index.rds"))) {
+  e_sigma_factor <- 3
+  keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  poly_dgp <- phn_model(j = 3, sigma = 0.1^e_sigma_factor)
+  ori_dat <- poly_dgp$gen(300)
+
+  vss <- c()
+  actual_ss <- c()
+  dat_combined <- tibble()
+
+  for (e_sigma in seq(0.7, 1.8, length.out = 20)^e_sigma_factor) {
+    poly_dgp <- phn_model(j = 3, sigma = e_sigma)
+    dat <- poly_dgp$gen(300, computed = list(x1 = ori_dat$x1, e = ori_dat$e, k = ori_dat$k * e_sigma / 0.1^e_sigma_factor))
+    mod <- lm(y ~ x1, data = dat)
+    my_vi <- autovi::auto_vi(fitted_mod = mod, keras_mod = keras_mod)
+    vss <- c(vss, my_vi$vss())
+    actual_ss <- c(actual_ss, log(poly_dgp$sample_effect_size(dat) + 1))
+
+    dat_combined <- bind_rows(dat_combined, mutate(dat, e_sigma = e_sigma))
+  }
+
+  saveRDS(dat_combined, here::here("paper/data/poly_index.rds"))
+  saveRDS(vss, here::here("paper/data/poly_index_vss.rds"))
+}
+
+dat_combined <- readRDS(here::here("paper/data/poly_index.rds"))
+vss <- readRDS(here::here("paper/data/poly_index_vss.rds"))
+
+dat_combined %>%
+  autovi::AUTO_VI$plot_resid() +
+  facet_wrap(~e_sigma, scales = "free", 
+             labeller = function(...){ 
+               tibble(nb = map(glue("'MVI' == {format(vss - log(300) + 10, digits = 3)}"), str2expression))
+               },
+             ncol = 5) +
+  theme(strip.text = element_text(size = 10))
+```
+
+```{r heter-index, fig.pos = "!h", fig.cap = "Residual plots generated from fitted models exhibiting varying degrees of heteroskedasticity violations.
+The Model violations index (MVI) is displayed atop each residual plot.
+The heteroskedasticity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$."}
+set.seed(1495)
+if (!file.exists(here::here("paper/data/heter_index.rds"))) {
+  b_factor <- 2
+  keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  heter_dgp <- phn_model(include_z = FALSE, b = 0)
+  ori_dat <- poly_dgp$gen(300)
+
+  vss <- c()
+  actual_ss <- c()
+  dat_combined <- tibble()
+
+  for (b in rev(seq(0.2, 4, length.out = 20))^b_factor) {
+    heter_dgp <- phn_model(include_z = FALSE, b = b)
+    dat <- heter_dgp$gen(300, computed = list(x1 = ori_dat$x1, e = ori_dat$e))
+    mod <- lm(y ~ x1, data = dat)
+    my_vi <- autovi::auto_vi(fitted_mod = mod, keras_mod = keras_mod)
+    vss <- c(vss, my_vi$vss())
+    actual_ss <- c(actual_ss, log(heter_dgp$sample_effect_size(dat) + 1))
+
+    dat_combined <- bind_rows(dat_combined, mutate(dat, nb = -b))
+  }
+
+  saveRDS(dat_combined, here::here("paper/data/heter_index.rds"))
+  saveRDS(vss, here::here("paper/data/heter_index_vss.rds"))
+}
+
+dat_combined <- readRDS(here::here("paper/data/heter_index.rds"))
+vss <- readRDS(here::here("paper/data/heter_index_vss.rds"))
+
+dat_combined %>%
+  autovi::AUTO_VI$plot_resid() +
+  facet_wrap(~nb, scales = "free", 
+             labeller = function(...){ 
+               tibble(nb = map(glue("'MVI' == {format(vss - log(300) + 10, digits = 3)}"), str2expression))
+               },
+             ncol = 5) +
+  theme(strip.text = element_text(size = 10))
+```
+
+## Statistical testing
+
+### Lineup Evaluation
 
@@
 
-To address this issue while adhering to the principle of visual inference, we can compare the estimated distance $\hat{D}$ to the approximated distances for the null plots in a lineup.
-Specifically, if a lineup comprises 20 plots, the null hypothesis $H_0$ will be rejected if $\hat{D}$ exceeds the maximum approximated distance among the $m - 1$ null plots, denoted as $\max\limits_{1 \leq i \leq m-1} {\hat{D}_{null}^{(i)}}$, where $\hat{D}_{null}^{(i)}$ represents the approximated distance for the $i$-th null plot.
+To address this issue while adhering to the principle of visual inference, we can compare the estimated distance $\hat{D}$ to the estimated distances for the null plots in a lineup.
+Specifically, if a lineup comprises 20 plots, the null hypothesis $H_0$ will be rejected if $\hat{D}$ exceeds the maximum estimated distance among the $m - 1$ null plots, denoted as $\max\limits_{1 \leq i \leq m-1} {\hat{D}_{null}^{(i)}}$, where $\hat{D}_{null}^{(i)}$ represents the estimated distance for the $i$-th null plot.
 This approach is equivalent to the typical lineup protocol requiring a 95% significance level, where $H_0$ is rejected if the data plot is identified as the most distinct plot by the sole participant.
-The approximated distance serves as a metric to quantify the difference between the data plot and the null plots, as intended.
+The estimated distance serves as a metric to quantify the difference between the data plot and the null plots, as intended.
 
 For lineups consisting of more than 20 plots, the 95% significance level can be maintained if the number of plots is a multiple of 20.
-Specifically, for lineups comprising $20p$ plots, where $p$ is a positive integer, we reject $H_0$ if $\hat{D}$ exceeds 95% ${\hat{D}_{null}^{(i)}}$ for $i = 1, \ldots, 20p-1$.
-The p-value in this case is given by $\frac{1}{m} + \frac{1}{m}\sum{i=1}^{m-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)$, where $I(\cdot)$ is the indicator function.
+Specifically, for lineups comprising $20t$ plots, where $l$ is a positive integer, we reject $H_0$ if $\hat{D}$ exceeds 95% ${\hat{D}_{null}^{(i)}}$ for $i = 1, \ldots, 20t-1$.
+The $p$-value in this case is given by $\frac{1}{20t} + \frac{1}{20t}\sum{i=1}^{20t-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)$, where $I(\cdot)$ is the indicator function.
 
@@
 The details of the sample quantile computation can be found in @hyndman1996sample.
-For instance, if $\hat{D}$ is greater than or equal to the 95% sample quantile, denoted as $Q_{null}(0.95)$, we can conclude that the approximated distance for the actual residual plot is significantly different from the approximated distance for null plots with a 95% significance level.
+For instance, if $\hat{D}$ is greater than or equal to the 95% sample quantile, denoted as $Q_{null}(0.95)$, we can conclude that the estimated distance for the true residual plot is significantly different from the estimated distance for null plots with a 95% significance level.
 Based on our experience, to obtain a stable estimate of the 95% quantile, the number of null plots, $n_{null}$, typically needs to be at least 100.
 However, if the null distribution exhibits a long tail, a larger number of null plots may be required.
-Alternatively, a p-value can be used to represent the probability of observing an event equally or more extreme than the given event under the null hypothesis $H_0$, and it can be estimated by $\frac{1}{m}\sum_{i=1}^{m-1}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)$.
+Alternatively, a $p$-value can be used to represent the probability of observing an event equally or more extreme than the given event under the null hypothesis $H_0$, and it can be estimated by $\frac{1}{m}\sum_{i=1}^{m-1}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)$.
 
@@
 
-### Bootstrapping the approximated distance
+### Bootstrapping
 
@@
 
-Similarly, bootstrap can be applied on the approximated distance $\hat{D}$.
+Similarly, bootstrap can be applied on the estimated distance $\hat{D}$.
 For each refitted model $M_{boot}^{(i)}$, there will be an associated residual plot $V_{boot}^{(i)}$ which can be fed into the computer vision model to obtain $\hat{D}_{boot}^{(i)}$, where $i = 1,...,n_{boot}$, and $n_{boot}$ is the number of bootstrapped samples.
@@
 
-## Generation of training data
-
-### Data generating process
+## Data Genearation
+
+### Simulation Scheme
 
@@
 
-### Computation of scagnostics
+### Scagnostics
 
@@
 These computed measures, along with the number of observations from the fitted model, were provided as the second input for the computer vision model.
-Although other scagnostics are informative, they are currently unavailable due to a fatal bug in the compiled C program of the `interp` R package [@Albrecht2023interp], which may unpredictably crash the process.
-For reproducibility, we excluded these scagnostics from the training data.
-
-### Crafting a balanced training set
+While other scagnostics provide valuable insights, they come with high computational costs and are not suitable for quick inference.
+
+### Balanced Dataset
 
@@
 
-## Architecture of the computer vision model
+## Model Architecture
 
@@
 \noindent where $n_{train}$ is the number of training samples, $V_i$ is the $i$-th residual plot and $S_i$ is the additional information about the $i$-th residual plot including four scagnostics and the number of observations.
-
-## Training and hyperparameter tuning
-
-To achieve an optimal deep learning model, hyperparameters like the learning rate often need to be fine-tuned using a tuner.
-In our study, we utilized the Bayesian optimization tuner from the `KerasTuner` Python library [@omalley2019kerastuner] for this purpose.
-A comprehensive list of hyperparameters is provided in Table \ref{tab:hyperparameter}.
-
-The "number of base filters" determines the number of filters for the first 2D convolutional layer.
-In the VGG16 architecture, the number of filters for the 2D convolutional layer in a block is typically twice the number in the previous block, except for the last block, which maintains the same number of convolution filters as the previous one.
-This hyperparameter aids in controlling the complexity of the computer vision model.
-Higher numbers of base filters result in more trainable parameters.
-Likewise, the "number of units for the fully-connected layer" determines the complexity of the final prediction block.
-Increasing the number of units enhances model complexity, resulting in more trainable parameters.
-
-The dropout rate and batch normalization are flexible hyperparameters that work in conjunction with other parameters to facilitate smooth training.
-A higher dropout rate is necessary when the model tends to overfit the training dataset by learning too much noise.
-Conversely, a lower dropout rate is preferred when the model is complex and challenging to converge.
-Batch normalization, on the other hand, addresses the internal covariate shift problem arising from the randomness in weight initialization and input data.
-It helps stabilize and accelerate the training process by normalizing the activations of each layer.
-
-Additionally, incorporating additional inputs such as scagnostics and the number of observations can potentially enhance prediction accuracy.
-Therefore, we allowed the tuner to determine whether these inputs were necessary for optimal model performance.
-
-The learning rate is a crucial hyperparameter, as it dictates the step size of the optimization algorithm.
-A high learning rate can help the model avoid local minima but risks overshooting and missing the global minimum.
-Conversely, a low learning rate smoothens the training process but makes the convergence time longer and increases the likelihood of getting trapped in local minima.
-
-Our model was trained on the M3 high-performance computing platform, using TensorFlow and Keras.
-During training, 80% of the training data was utilized for actual training, while the remaining 20% was used as validation data.
-The Bayesian optimization tuner conducted 100 trials to identify the best hyperparameter values based on validation root mean square error.
-The tuner then restored the best epoch of the best model from the trials.
-Additionally, we applied early stopping, terminating the training process if the validation root mean square error fails to improve for 50 epochs.
-The maximum allowed epochs was set at 2000, although no models reached this threshold.
-
-```{r hyperparameter}
-data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
-           Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[1e-8, 1e-1]")) %>%
-  kableExtra::kable("latex", booktabs = TRUE, caption = "Name of hyperparameters and their correspoding domain for the computer vision model.")
-```
-
-## Model evaluation methods
-
-This section remains incomplete as we may modify the "Results" section later on.
-
-# Results
-
-## Optimized hyperparameters
 
@@
 ```
+
+## Model Training
+
+To achieve an optimal deep learning model, hyperparameters like the learning rate often need to be fine-tuned using a tuner.
+In our study, we utilized the Bayesian optimization tuner from the `KerasTuner` Python library [@omalley2019kerastuner] for this purpose.
+A comprehensive list of hyperparameters is provided in Table \ref{tab:hyperparameter}.
+
+The "number of base filters" determines the number of filters for the first 2D convolutional layer.
+In the VGG16 architecture, the number of filters for the 2D convolutional layer in a block is typically twice the number in the previous block, except for the last block, which maintains the same number of convolution filters as the previous one.
+This hyperparameter aids in controlling the complexity of the computer vision model.
+Higher numbers of base filters result in more trainable parameters.
+Likewise, the "number of units for the fully-connected layer" determines the complexity of the final prediction block.
+Increasing the number of units enhances model complexity, resulting in more trainable parameters.
+
+The dropout rate and batch normalization are flexible hyperparameters that work in conjunction with other parameters to facilitate smooth training.
+A higher dropout rate is necessary when the model tends to overfit the training dataset by learning too much noise.
+Conversely, a lower dropout rate is preferred when the model is complex and challenging to converge.
+Batch normalization, on the other hand, addresses the internal covariate shift problem arising from the randomness in weight initialization and input data.
+It helps stabilize and accelerate the training process by normalizing the activations of each layer.
+
+Additionally, incorporating additional inputs such as scagnostics and the number of observations can potentially enhance prediction accuracy.
+Therefore, we allowed the tuner to determine whether these inputs were necessary for optimal model performance.
+
+The learning rate is a crucial hyperparameter, as it dictates the step size of the optimization algorithm.
+A high learning rate can help the model avoid local minima but risks overshooting and missing the global minimum.
+Conversely, a low learning rate smoothens the training process but makes the convergence time longer and increases the likelihood of getting trapped in local minima.
+
+Our model was trained on the M3 high-performance computing platform, using TensorFlow and Keras.
+During training, 80% of the training data was utilized for actual training, while the remaining 20% was used as validation data.
+The Bayesian optimization tuner conducted 100 trials to identify the best hyperparameter values based on validation root mean square error.
+The tuner then restored the best epoch of the best model from the trials.
+Additionally, we applied early stopping, terminating the training process if the validation root mean square error fails to improve for 50 epochs.
+The maximum allowed epochs was set at 2000, although no models reached this threshold.
+
+```{r hyperparameter}
+data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
+           Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[1e-8, 1e-1]")) %>%
+  kableExtra::kable("latex", booktabs = TRUE, caption = "Name of hyperparameters and their correspoding domain for the computer vision model.")
+```
+
+## Evaluation Methods
+
+This section remains incomplete as we may modify the "Results" section later on.
+
+# Results
+
+understand what the convolutional layer is actually doing (read vgg16 doc)
 
@@
 
-## Demostration of MVI
-
-As described in Section \ref{sec:model-violations-index} a Model Violations Index (MVI) can be constructed using $\hat{D}$. This section provides guidelines and examples of MVI.
-
-Figures \ref{fig:poly-index} and \ref{fig:heter-index} display the residual plots for fitted models exhibiting varying degrees of non-linearity and heteroskedasticity.
-Each residual plot's MVI is computed using Equation \ref{eq:mvi} with $C = 10$.
-When $\text{MVI} > 8$, the visual patterns are notably strong and easily discernible by humans.
-In the range $6 < \text{MVI} < 8$, the visibility of the visual pattern diminishes as MVI decreases.
-Conversely, when $\text{MVI} < 6$, the visual pattern tends to become relatively faint and challenging to observe.
-Table \ref{tab:mvi} provides a summary of the MVI usage and it is applicable to other linear regression models.
-
-```{r mvi}
-data.frame(degree = c("Strong", "Moderate", "Weak"),
-           range = c("$\\text{MVI} > 8$", "$6 < \\text{MVI} < 8$", "$\\text{MVI} < 6$")) %>%
-  kableExtra::kable("latex", 
-                    booktabs = TRUE,
-                    escape = FALSE,
-                    col.names = c("Degree of model violations", "Range ($C$ = 10)"),
-                    caption = "Degree of model violations or the strength of the visual signals according to the Model Violation Index (MVI). The constant $C$ is set to be 10.")
-```
-
-```{r poly-index, fig.pos = "!h", fig.cap = "Residual plots generated from fitted models exhibiting varying degrees of non-linearity violations.
-The Model violations index (MVI) is displayed atop each residual plot.
-The non-linearity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$."}
-set.seed(1495)
-if (!file.exists(here::here("paper/data/poly_index.rds"))) {
-  e_sigma_factor <- 3
-  keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
-  poly_dgp <- phn_model(j = 3, sigma = 0.1^e_sigma_factor)
-  ori_dat <- poly_dgp$gen(300)
-
-  vss <- c()
-  actual_ss <- c()
-  dat_combined <- tibble()
-
-  for (e_sigma in seq(0.7, 1.8, length.out = 20)^e_sigma_factor) {
-    poly_dgp <- phn_model(j = 3, sigma = e_sigma)
-    dat <- poly_dgp$gen(300, computed = list(x1 = ori_dat$x1, e = ori_dat$e, k = ori_dat$k * e_sigma / 0.1^e_sigma_factor))
-    mod <- lm(y ~ x1, data = dat)
-    my_vi <- autovi::auto_vi(fitted_mod = mod, keras_mod = keras_mod)
-    vss <- c(vss, my_vi$vss())
-    actual_ss <- c(actual_ss, log(poly_dgp$sample_effect_size(dat) + 1))
-
-    dat_combined <- bind_rows(dat_combined, mutate(dat, e_sigma = e_sigma))
-  }
-
-  saveRDS(dat_combined, here::here("paper/data/poly_index.rds"))
-  saveRDS(vss, here::here("paper/data/poly_index_vss.rds"))
-}
-
-dat_combined <- readRDS(here::here("paper/data/poly_index.rds"))
-vss <- readRDS(here::here("paper/data/poly_index_vss.rds"))
-
-dat_combined %>%
-  autovi::AUTO_VI$plot_resid() +
-  facet_wrap(~e_sigma, scales = "free", 
-             labeller = function(...){ 
-               tibble(nb = map(glue("'MVI' == {format(vss - log(300) + 10, digits = 3)}"), str2expression))
-               },
-             ncol = 5) +
-  theme(strip.text = element_text(size = 10))
-```
-
-```{r heter-index, fig.pos = "!h", fig.cap = "Residual plots generated from fitted models exhibiting varying degrees of heteroskedasticity violations.
-The Model violations index (MVI) is displayed atop each residual plot.
-The heteroskedasticity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$."}
-set.seed(1495)
-if (!file.exists(here::here("paper/data/heter_index.rds"))) {
-  b_factor <- 2
-  keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
-  heter_dgp <- phn_model(include_z = FALSE, b = 0)
-  ori_dat <- poly_dgp$gen(300)
-
-  vss <- c()
-  actual_ss <- c()
-  dat_combined <- tibble()
-
-  for (b in rev(seq(0.2, 4, length.out = 20))^b_factor) {
-    heter_dgp <- phn_model(include_z = FALSE, b = b)
-    dat <- heter_dgp$gen(300, computed = list(x1 = ori_dat$x1, e = ori_dat$e))
-    mod <- lm(y ~ x1, data = dat)
-    my_vi <- autovi::auto_vi(fitted_mod = mod, keras_mod = keras_mod)
-    vss <- c(vss, my_vi$vss())
-    actual_ss <- c(actual_ss, log(heter_dgp$sample_effect_size(dat) + 1))
-
-    dat_combined <- bind_rows(dat_combined, mutate(dat, nb = -b))
-  }
-
-  saveRDS(dat_combined, here::here("paper/data/heter_index.rds"))
-  saveRDS(vss, here::here("paper/data/heter_index_vss.rds"))
-}
-
-dat_combined <- readRDS(here::here("paper/data/heter_index.rds"))
-vss <- readRDS(here::here("paper/data/heter_index_vss.rds"))
-
-dat_combined %>%
-  autovi::AUTO_VI$plot_resid() +
-  facet_wrap(~nb, scales = "free", 
-             labeller = function(...){ 
-               tibble(nb = map(glue("'MVI' == {format(vss - log(300) + 10, digits = 3)}"), str2expression))
-               },
-             ncol = 5) +
-  theme(strip.text = element_text(size = 10))
-```
-
 ## Comparison with human visual inference and conventional tests
@@
 This study has collected $7974$ human responses to $1152$ lineups.
-Each lineup contains one randomly placed actual residual plot and 19 null plots.
+Each lineup contains one randomly placed true residual plot and 19 null plots.
 Among the $1152$ lineups, $24$ are attention check lineups in which the visual patterns are designed to be extremely obvious and very different from the corresponding to null plots, $36$ are null lineups where all the lineups consist of only null plots, $279$ are lineups with uniform predictor distribution evaluated by $11$ participants, and the remaining $813$ are lineups with discrete, skewed or normal predictor distribution evaluated by $5$ participants.
@@
 
-For each lineup used in @li2023plot, there is one actual residual plot and 19 null plots.
-While the distance $D$ for the actual residual plot depends on the underlying data generating process, the distance $D$ for the null plots is zero.
-We have used our optimized computer vision model to predict the approximated distance for both the actual residual plots and the null plots.
-To have a fair comparison, $H_0$ will be rejected if the actual residual plot has the greatest approximated distance among all plots in a lineup.
+For each lineup used in @li2023plot, there is one true residual plot and 19 null plots.
+While the distance $D$ for the true residual plot depends on the underlying data generating process, the distance $D$ for the null plots is zero.
+We have used our optimized computer vision model to predict the approximated distance for both the true residual plots and the null plots.
+To have a fair comparison, $H_0$ will be rejected if the true residual plot has the greatest approximated distance among all plots in a lineup.
 Additionally, the appropriate conventional tests including the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey1969tests] for non-linearity and the Breusch-Pagan test  [@breusch1979simple] for heteroskedasticity were applied on the same data for comparison.
 
-The performance metrics of $\hat{D}$ for actual residual plots are outlined in Table \ref{tab:experiment-performance}.
+The performance metrics of $\hat{D}$ for true residual plots are outlined in Table \ref{tab:experiment-performance}.
 It's notable that all performance metrics are slightly worse than those evaluated on the test data.
@@
 Figure \ref{fig:delta} indicates that the weighted detection rate increases as the $\delta$-difference increases, particularly when the $\delta$-difference is greater than zero.
-A negative $\delta$-difference suggests that there is at least one null plot in the lineup with a stronger visual signal than the actual residual plot.
+A negative $\delta$-difference suggests that there is at least one null plot in the lineup with a stronger visual signal than the true residual plot.
 In some instances, the weighted detection rate is close to one, yet the $\delta$-difference is negative.
@@
 This phenomenon is evidently caused by the skewed distribution of the fitted values.
-Notably, if the residual plot were evaluated through a visual test, it would not be rejected since the actual residual plot positioned at 10 can not be distinguished from the others.
+Notably, if the residual plot were evaluated through a visual test, it would not be rejected since the true residual plot positioned at 10 can not be distinguished from the others.
 
@@
 ```{r false-lineup, fig.pos = "!h", fig.cap = 'A lineup of residual plots displaying "left-triangle" visual patterns.
-The actual residual plot occupies position 10, yet there are no discernible visual patterns that distinguish it from the other plots.'}
+The true residual plot occupies position 10, yet there are no discernible visual patterns that distinguish it from the other plots.'}
 pos <- 10
@@
 
-Additionally, the lineup of this fitted model presented in Figure \ref{fig:boston-lineup} underscores that the "U" shape is a distinctive characteristic of the actual residual plot, absent in any of the null plots.
+Additionally, the lineup of this fitted model presented in Figure \ref{fig:boston-lineup} underscores that the "U" shape is a distinctive characteristic of the true residual plot, absent in any of the null plots.
 Consequently, if a visual test is conducted, $H_0$ will also be rejected.
@@
 ```{r boston-lineup, fig.pos = "!h", fig.cap = 'A lineup of residual plots for the Boston housing fitted model.
-The actual residual plot is at position 7.
+The true residual plot is at position 7.
 It can be easily identified as the most different plot.'}
@@
 ```{r dino-lineup, fig.pos = "!h", fig.cap = 'A lineup of residual plots for the fitted model on the "dinosaur" dataset.
-The actual residual plot is at position 17.
+The true residual plot is at position 17.
 It can be easily identified as the most different plot as the visual pattern is extremly artificial.'}
@@
 
-- Use raw residuals.
-  - D for null plots will be greater than zero.
-  - Can not assume D for actual plots are greater than D for null plots any more.
+1. Use raw residuals.
+  - D (effect size) for null plots will be greater than zero.
+  - Can not assume D for actual plots are greater than D for null plots any more. -> KL null - normal, KL actual - normal
   - Need to use a two-sided test instead of a one-sided test.
   - It may fail completely due to the identification issue.
-- Use transformed residuals that roughly follow a multivariate normal distribution.
+2. Use transformed residuals that roughly follow a multivariate normal distribution.
   - The one-sided test will still work. 
   - Need to ensure that the transformed actual plot will have residual distribution different from the transformed null distribution. 
-- Train a new model.
+3. Train a new model.
   - KL divergence may not be the appropriate measure for the difference between the residual distributions. Could consider using a new distance measure.
-  -
 

### 2024-04-29 paper/paper.tex
update diagram
revise up to model arch

@@
 \usepackage{setspace}
-\doublespacing
 
@@
 straightforward as executing a line of code or clicking a button.
-However, accurately interpreting a residual plot can be challenging.
-Consider Figure \ref{fig:false-finding} as an example, the residuals
-display a triangular shape pointing to the left. While this might
-suggest heteroskedasticity, it is important to avoid over-interpreting
-this visual pattern. In this case, the fitted regression model is
-correctly specified, and the triangular shape is actually a result of
-the skewed distribution of the predictors, rather than indicating a flaw
-in the model.
-
-A residual plot can exhibit various visual features, but it is crucial
-to recognize that some may arise from the characteristics of predictors
-and the inherent randomness of the error, rather than indicating a
-violation of model assumptions \citep{li2023plot}. The concept of visual
-inference, as proposed by \citet{buja2009statistical}, provides an
-inferential framework to assess whether residual plots indeed contain
-visual patterns inconsistent with the model assumptions. The fundamental
-idea involves testing whether the actual residual plot visually differs
-significantly from null plots, where null plots are plotted with
-residuals generated from the residual rotation distribution
-\citep{langsrud2005rotation}, which is a distribution consistent with
-the null hypothesis \(H_0\) that the linear regression model is
-correctly specified. Typically, the visual test is accomplished through
-the lineup protocol, where the real residual plot is embedded within a
-lineup alongside several null plots. If the real residual plot can be
-distinguished from the lineup, it provides evidence for rejecting
-\(H_0\).
+However, accurately interpreting a residual plot can be challenging. A
+residual plot can exhibit various visual features, but it is crucial to
+recognize that some may arise from the characteristics of predictors and
+the inherent randomness of the error, rather than indicating a violation
+of model assumptions \citep{li2023plot}. Consider Figure
+\ref{fig:false-finding} as an example, the residuals display a
+triangular shape pointing to the left. While this might suggest
+heteroskedasticity, it is important to avoid over-interpreting this
+visual pattern. In this case, the fitted regression model is correctly
+specified, and the triangular shape is actually a result of the skewed
+distribution of the predictors, rather than indicating a flaw in the
+model.
+
+The concept of visual inference, as proposed by
+\citet{buja2009statistical}, provides an inferential framework to assess
+whether residual plots indeed contain visual patterns inconsistent with
+the model assumptions. The fundamental idea involves testing whether the
+true residual plot visually differs significantly from null plots, where
+null plots are plotted with residuals generated from the residual
+rotation distribution \citep{langsrud2005rotation}, which is a
+distribution consistent with the null hypothesis \(H_0\) that the linear
+regression model is correctly specified. Typically, the visual test is
+accomplished through the lineup protocol, where the real residual plot
+is embedded within a lineup alongside several null plots. If the real
+residual plot can be distinguished from the lineup, it provides evidence
+for rejecting \(H_0\).
 
@@
 it may not capture all the necessary visual features that differentiate
-actual residual plots from null plots. A more promising alternative
+true residual plots from null plots. A more promising alternative
 entails enabling machines to learn the function for extracting visual
@@
 
-\subsection{Different possible configurations of the model
-formula}\label{different-possible-configurations-of-the-model-formula}
+\subsection{Model Formulas}\label{model-formulas}
 
@@
 be used to assess residual plots. We discuss these configurations below
-based on two key components of the model formula: the input and the
+focusing on two key components of the model formula: the input and the
 output format.
 
-\subsubsection{Input formats}\label{input-formats}
+\subsubsection{Input Formats}\label{input-formats}
 
@@
 null plots of a fitted regression model to be similar to each other,
-while considering actual residual plots to be distinct from null plots
-of any fitted regression model. A triplet constitutes a set of three
+while considering true residual plots to be distinct from null plots of
+any fitted regression model. A triplet constitutes a set of three
 images, denoted as \(image_1\), \(image_2\) and \(image_3\). It is often
@@
 
-We did not explore all the mentioned input formats due the considerable
-costs associated with data preparation and model training. Considering
-the implementation cost and the interpretability of the model, we
-settled on the single residual plot input format.
-
-\subsubsection{Output formats}\label{output-formats}
+We did not explore all the mentioned input formats due to the
+significant costs tied to data preparation and model training. Taking
+into account the implementation cost and the need for model
+interpretability, we settled on the single residual plot input format.
+
+\subsubsection{Output Formats}\label{output-formats}
 
@@
 strength of suspicious visual patterns reflecting the extent of model
-violations and the difficulty index for identifying whether a residual
+violations, or the difficulty index for identifying whether a residual
 plot has no issues. However, these numeric measures are often informally
@@
 
-\subsection{Distance from the good residual
-plots}\label{distance-from-the-good-residual-plots}
+\subsection{Distance Between Residual
+Plots}\label{distance-between-residual-plots}
 
@@
 
-\subsubsection{Residual distribution}\label{residual-distribution}
+\subsubsection{Residual Distribution}\label{residual-distribution}
 
@@
 residual plot, or in other words, how different a given fitted
-regression model is from a correctly specified model. For the classical
-normal linear regression model, residuals \(\boldsymbol{e}\) are derived
-from the fitted values \(\hat{\boldsymbol{y}}\) and observed values
-\(\boldsymbol{y}\). Suppose the data generating process is known and the
-regression model is correctly specified, by the Frisch-Waugh-Lowell
-theorem \citep{frisch1933partial}, residuals \(\boldsymbol{e}\) can also
-be treated as random variables and written as a linear transformation of
+regression model is from a correctly specified model.
+
+For a classical normal linear regression model, residuals
+\(\boldsymbol{e}\) are derived from the fitted values
+\(\hat{\boldsymbol{y}}\) and observed values \(\boldsymbol{y}\). Suppose
+the data generating process is known and the regression model is
+correctly specified, by the Frisch-Waugh-Lowell theorem
+\citep{frisch1933partial}, residuals \(\boldsymbol{e}\) can also be
+treated as random variables and written as a linear transformation of
 the error \(\boldsymbol{\varepsilon}\) formulated as
@@
 
-In ordinary least square, the minimization of the sum of square
-residuals implies \(\sum_{i=1}^{n} e_i = 0\), making any residual value
-to be a linear combination of the remaining \(n - 1\) residuals. This
-effectively means \(\text{rank}(\boldsymbol{R}) = n - 1 < n\) and \(Q\)
-is a degenerate multivariate distribution. To capture the
-characteristics of \(Q\), such as moments, we can simulate a large
-numbers of \(\boldsymbol{\varepsilon}\) and transform it to
-\(\boldsymbol{e}\) to get the empirical estimates. For simplicity, in
-this study, we replaced the variance-covariance matrix of residuals
+Suppose the design matrix \(\boldsymbol{X}\) has linearly independent
+columns, the trace of the hat matrix will equal to the number of columns
+in \(\boldsymbol{X}\) denoted as \(k\). As a result, the rank of
+\(\boldsymbol{R}\) is \(n - k\), and \(Q\) is a degenerate multivariate
+distribution. To capture the characteristics of \(Q\), such as moments,
+we can simulate a large numbers of \(\boldsymbol{\varepsilon}\) and
+transform it to \(\boldsymbol{e}\) to get the empirical estimates. For
+simplicity, in this study, we replaced the variance-covariance matrix of
+residuals
 \(\text{cov}(\boldsymbol{e}, \boldsymbol{e}) = \boldsymbol{R}\sigma^2\boldsymbol{R}' = \boldsymbol{R}\sigma^2\)
@@
 if the data generating process contains variables correlated with any
-column of \(\boldsymbol{X}\) but not included in \(\boldsymbol{X}\),
+column of \(\boldsymbol{X}\) but missing from \(\boldsymbol{X}\),
 causing an omitted variable problem, \(P\) will be different from \(Q\)
@@
 
-\subsubsection{\texorpdfstring{Kullback-Leibler divergence of \(P\) from
-\(Q\)}{Kullback-Leibler divergence of P from Q}}\label{kullback-leibler-divergence-of-p-from-q}
+\subsubsection{\texorpdfstring{Distance of \(P\) from
+\(Q\)}{Distance of P from Q}}\label{distance-of-p-from-q}
 
 Define a proper distance between distributions is usually easier than
-define a proper distance between data plots. Given the actual residual
+define a proper distance between data plots. Given the true residual
 distribution \(Q\) and the reference residual distribution \(P\), we
@@
 
-\noindent where \(p(.)\) is the probability density function for
-distribution \(P\), and \(q(.)\) is the probability density function for
-distribution \(Q\).
+\noindent where \(p(.)\) and \(q(.)\) are the probability density
+functions for distribution \(P\) and distribution \(Q\) respectively.
 
@@
 mainly designed for measuring the effect size of non-linearity and
-heteroskedasticity in a residual plot. \citet{li2023plot} have showed
+heteroskedasticity in a residual plot. \citet{li2023plot} have derived
 that, for a classical normal linear regression model that omits a
@@
 assumed error variance \(\sigma^2\) is set to be
-\(\text{tr}(\boldsymbol{V})/n\), which is the expectation of the
-estimated variance.
-
-\subsubsection{\texorpdfstring{Evaluation of Kullback-Leibler divergence
-for non-normal
-\(P\)}{Evaluation of Kullback-Leibler divergence for non-normal P}}\label{evaluation-of-kullback-leibler-divergence-for-non-normal-p}
-
-For non-normal error \(\boldsymbol{\varepsilon}\), the actual residual
+\(tr(\boldsymbol{V})/n\), which is the expectation of the estimated
+variance.
+
+\subsubsection{\texorpdfstring{Non-normal
+\(P\)}{Non-normal P}}\label{non-normal-p}
+
+For non-normal error \(\boldsymbol{\varepsilon}\), the true residual
 distribution \(P\) is unlikely to be a multivariate normal distribution.
@@
 
-\subsubsection{Approximation of the distance
-measure}\label{approximation-of-the-distance-measure}
+\subsection{Distance Estimation}\label{distance-estimation}
 
 In the previous sections, we have defined a distance measure given in
-equation \ref{eq:kl-0} for quantifying the difference between the actual
+equation \ref{eq:kl-0} for quantifying the difference between the true
 residual distribution \(P\) and an ideal reference distribution \(Q\).
@@
 
-We tried to train a computer vision model to approximate this distance
+We tried to train a computer vision model to estimate this distance
 measure with a residual plot. Let \(D\) be the result of equation
@@
 \noindent where \(V_{h \times w}(.)\) is a plotting function that saves
-a residuals vs fitted values plot with fixed aesthetic as an image with
-\(h \times w\) pixels and three colour channels, \(f_{CV}(.)\) is a
-computer vision model which takes an \(h \times w\) image as input and
-predicts the distance in the domain \([0, +\infty)\).
-
-With the approximated distance \(\hat{D}\), we will be able to know how
+a residuals vs fitted values plot with fixed aesthetic as an RGB image
+with \(h \times w\) pixels, \(f_{CV}(.)\) is a computer vision model
+which takes an \(h \times w\) image as input and predicts the distance
+in the domain \([0, +\infty)\).
+
+With the estimated distance \(\hat{D}\), we will be able to know how
 different the underlying distribution of the residuals is from a good
-residual distribution. The approximated distance \(\hat{D}\) can also be
-used as an index of the model violations. It also provides information
-for the strength of the visual signal embedded in the residual plot.
-
-The approximated distance \(\hat{D}\) is not expected to be the same as
-the original distance \(D\). This is largely because information
-contained in a single residual plot is limited and it may not be able to
-summarise all the important characteristics of the residual
-distribution. For a given residual distribution \(P\), we can generate
-many different residual plots. Some of them share similar visual
-patterns, but some of them could be visually very different from the
-rest, especially for regression models with small \(n\). This suggests
-the error of the approximation will vary depends on whether the observed
-residual plot is representative or not.
-
-\subsubsection{Model Violations Index
+residual distribution. \(\hat{D}\) can also be used as an index of the
+model violations indicating the strength of the visual signal embedded
+in the residual plot.
+
+It is not expected that \(\hat{D}\) will be equal to original distance
+\(D\). This is largely because information contained in a single
+residual plot is limited and it may not be able to summarise all the
+important characteristics of the residual distribution. For a given
+residual distribution \(P\), many different residual plots can be
+simulated, where many will share similar visual patterns, but some of
+them could be visually very different from the rest, especially for
+regression models with small \(n\). This suggests the error of the
+estimation will vary depends on whether the input residual plot is
+representative or not.
+
+\subsection{Model Violations Index
 (MVI)}\label{sec:model-violations-index}
 
-In Section \ref{output-formats}, we discuss the output formats of the
-computer vision model presented in this paper. One notable advantage
-highlighted is its interpretability. The approximated distance
-\(\hat{D}\) is an estimator of the difference between the actual
-residual distribution and the reference residual distribution. This
-difference primarily arises from deviations in model assumptions. The
-magnitude of \(D\) directly reflects the degree of these deviations,
-thus making \(\hat{D}\) instrumental in forming a model violation index.
-
-Note that the Kullback-Leibler divergence might be influenced by the
-number of random variables involved in its evaluation. Generally, a
-larger number of observations will lead to a greater distance \(D\).
-However, this does not imply that \(\hat{D}\) fails to accurately
-represent the extent of model violations. In fact, when examining
-residual plots with more observations, we often observe a stronger
-visual signal strength, as the underlying patterns are more likely to be
-revealed, except in cases of significant overlapping.
+\(\hat{D}\) is an estimator of the difference between the true residual
+distribution and the reference residual distribution. This difference
+primarily arises from deviations in model assumptions. The magnitude of
+\(D\) directly reflects the degree of these deviations, thus making
+\(\hat{D}\) instrumental in forming a model violations index.
+
+Note that Equation \ref{eq:kl-0} might be influenced by the number of
+random variables involved in its evaluation. Generally, a larger number
+of observations will lead to a greater distance \(D\). However, this
+does not imply that \(\hat{D}\) fails to accurately represent the extent
+of model violations. In fact, when examining residual plots with more
+observations, we often observe a stronger visual signal strength, as the
+underlying patterns are more likely to be revealed, except in cases of
+significant overlapping.
 
@@
 
-\subsection{Statistical testing based on the approximated
-distance}\label{statistical-testing-based-on-the-approximated-distance}
-
-\subsubsection{Lineup evaluation}\label{lineup-evaluation}
+Figures \ref{fig:poly-index} and \ref{fig:heter-index} display the
+residual plots for fitted models exhibiting varying degrees of
+non-linearity and heteroskedasticity. Each residual plot's MVI is
+computed using Equation \ref{eq:mvi} with \(C = 10\). When
+\(\text{MVI} > 8\), the visual patterns are notably strong and easily
+discernible by humans. In the range \(6 < \text{MVI} < 8\), the
+visibility of the visual pattern diminishes as MVI decreases.
+Conversely, when \(\text{MVI} < 6\), the visual pattern tends to become
+relatively faint and challenging to observe. Table \ref{tab:mvi}
+provides a summary of the MVI usage and it is applicable to other linear
+regression models.
+
+\begin{table}
+
+\caption{\label{tab:mvi}Degree of model violations or the strength of the visual signals according to the Model Violation Index (MVI). The constant $C$ is set to be 10.}
+\centering
+\begin{tabular}[t]{ll}
+\toprule
+Degree of model violations & Range ($C$ = 10)\\
+\midrule
+Strong & $\text{MVI} > 8$\\
+Moderate & $6 < \text{MVI} < 8$\\
+Weak & $\text{MVI} < 6$\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/poly-index-1}
+
+}
+
+\caption{Residual plots generated from fitted models exhibiting varying degrees of non-linearity violations.
+The Model violations index (MVI) is displayed atop each residual plot.
+The non-linearity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$.}\label{fig:poly-index}
+\end{figure}
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heter-index-1}
+
+}
+
+\caption{Residual plots generated from fitted models exhibiting varying degrees of heteroskedasticity violations.
+The Model violations index (MVI) is displayed atop each residual plot.
+The heteroskedasticity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$.}\label{fig:heter-index}
+\end{figure}
+
+\subsection{Statistical testing}\label{statistical-testing}
+
+\subsubsection{Lineup Evaluation}\label{lineup-evaluation}
 
@@
 inference, we can compare the estimated distance \(\hat{D}\) to the
-approximated distances for the null plots in a lineup. Specifically, if
-a lineup comprises 20 plots, the null hypothesis \(H_0\) will be
-rejected if \(\hat{D}\) exceeds the maximum approximated distance among
-the \(m - 1\) null plots, denoted as
+estimated distances for the null plots in a lineup. Specifically, if a
+lineup comprises 20 plots, the null hypothesis \(H_0\) will be rejected
+if \(\hat{D}\) exceeds the maximum estimated distance among the
+\(m - 1\) null plots, denoted as
 \(\max\limits_{1 \leq i \leq m-1} {\hat{D}_{null}^{(i)}}\), where
-\(\hat{D}_{null}^{(i)}\) represents the approximated distance for the
+\(\hat{D}_{null}^{(i)}\) represents the estimated distance for the
 \(i\)-th null plot. This approach is equivalent to the typical lineup
@@
 if the data plot is identified as the most distinct plot by the sole
-participant. The approximated distance serves as a metric to quantify
-the difference between the data plot and the null plots, as intended.
+participant. The estimated distance serves as a metric to quantify the
+difference between the data plot and the null plots, as intended.
 
@@
 level can be maintained if the number of plots is a multiple of 20.
-Specifically, for lineups comprising \(20p\) plots, where \(p\) is a
+Specifically, for lineups comprising \(20t\) plots, where \(l\) is a
 positive integer, we reject \(H_0\) if \(\hat{D}\) exceeds 95\%
-\({\hat{D}_{null}^{(i)}}\) for \(i = 1, \ldots, 20p-1\). The p-value in
-this case is given by
-\(\frac{1}{m} + \frac{1}{m}\sum{i=1}^{m-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)\),
+\({\hat{D}_{null}^{(i)}}\) for \(i = 1, \ldots, 20t-1\). The \(p\)-value
+in this case is given by
+\(\frac{1}{20t} + \frac{1}{20t}\sum{i=1}^{20t-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)\),
 where \(I(\cdot)\) is the indicator function.
@@
 quantile, denoted as \(Q_{null}(0.95)\), we can conclude that the
-approximated distance for the actual residual plot is significantly
-different from the approximated distance for null plots with a 95\%
-significance level. Based on our experience, to obtain a stable estimate
-of the 95\% quantile, the number of null plots, \(n_{null}\), typically
-needs to be at least 100. However, if the null distribution exhibits a
-long tail, a larger number of null plots may be required. Alternatively,
-a p-value can be used to represent the probability of observing an event
-equally or more extreme than the given event under the null hypothesis
-\(H_0\), and it can be estimated by
+estimated distance for the true residual plot is significantly different
+from the estimated distance for null plots with a 95\% significance
+level. Based on our experience, to obtain a stable estimate of the 95\%
+quantile, the number of null plots, \(n_{null}\), typically needs to be
+at least 100. However, if the null distribution exhibits a long tail, a
+larger number of null plots may be required. Alternatively, a
+\(p\)-value can be used to represent the probability of observing an
+event equally or more extreme than the given event under the null
+hypothesis \(H_0\), and it can be estimated by
 \(\frac{1}{m}\sum_{i=1}^{m-1}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)\).
@@
 
-\subsubsection{Bootstrapping the approximated
-distance}\label{bootstrapping-the-approximated-distance}
+\subsubsection{Bootstrapping}\label{bootstrapping}
 
@@
 
-Similarly, bootstrap can be applied on the approximated distance
+Similarly, bootstrap can be applied on the estimated distance
 \(\hat{D}\). For each refitted model \(M_{boot}^{(i)}\), there will be
@@
 
-\subsection{Generation of training
-data}\label{generation-of-training-data}
-
-\subsubsection{Data generating process}\label{data-generating-process}
+\subsection{Data Genearation}\label{data-genearation}
+
+\subsubsection{Simulation Scheme}\label{simulation-scheme}
 
@@
 
-\subsubsection{Computation of
-scagnostics}\label{computation-of-scagnostics}
+\subsubsection{Scagnostics}\label{scagnostics}
 
@@
 model, were provided as the second input for the computer vision model.
-Although other scagnostics are informative, they are currently
-unavailable due to a fatal bug in the compiled C program of the
-\texttt{interp} R package \citep{Albrecht2023interp}, which may
-unpredictably crash the process. For reproducibility, we excluded these
-scagnostics from the training data.
-
-\subsubsection{Crafting a balanced training
-set}\label{crafting-a-balanced-training-set}
+While other scagnostics provide valuable insights, they come with high
+computational costs and are not suitable for quick inference.
+
+\subsubsection{Balanced Dataset}\label{balanced-dataset}
 
@@
 
-\subsection{Architecture of the computer vision
-model}\label{architecture-of-the-computer-vision-model}
+\subsection{Model Architecture}\label{model-architecture}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{cnn}
+
+}
+
+\caption{Diagram of the architecture of the optimized computer vision model.}\label{fig:unnamed-chunk-2}
+\end{figure}
 
@@
 number of observations.
-
-\subsection{Training and hyperparameter
-tuning}\label{training-and-hyperparameter-tuning}
-
-To achieve an optimal deep learning model, hyperparameters like the
-learning rate often need to be fine-tuned using a tuner. In our study,
-we utilized the Bayesian optimization tuner from the \texttt{KerasTuner}
-Python library \citep{omalley2019kerastuner} for this purpose. A
-comprehensive list of hyperparameters is provided in Table
-\ref{tab:hyperparameter}.
-
-The ``number of base filters'' determines the number of filters for the
-first 2D convolutional layer. In the VGG16 architecture, the number of
-filters for the 2D convolutional layer in a block is typically twice the
-number in the previous block, except for the last block, which maintains
-the same number of convolution filters as the previous one. This
-hyperparameter aids in controlling the complexity of the computer vision
-model. Higher numbers of base filters result in more trainable
-parameters. Likewise, the ``number of units for the fully-connected
-layer'' determines the complexity of the final prediction block.
-Increasing the number of units enhances model complexity, resulting in
-more trainable parameters.
-
-The dropout rate and batch normalization are flexible hyperparameters
-that work in conjunction with other parameters to facilitate smooth
-training. A higher dropout rate is necessary when the model tends to
-overfit the training dataset by learning too much noise. Conversely, a
-lower dropout rate is preferred when the model is complex and
-challenging to converge. Batch normalization, on the other hand,
-addresses the internal covariate shift problem arising from the
-randomness in weight initialization and input data. It helps stabilize
-and accelerate the training process by normalizing the activations of
-each layer.
-
-Additionally, incorporating additional inputs such as scagnostics and
-the number of observations can potentially enhance prediction accuracy.
-Therefore, we allowed the tuner to determine whether these inputs were
-necessary for optimal model performance.
-
-The learning rate is a crucial hyperparameter, as it dictates the step
-size of the optimization algorithm. A high learning rate can help the
-model avoid local minima but risks overshooting and missing the global
-minimum. Conversely, a low learning rate smoothens the training process
-but makes the convergence time longer and increases the likelihood of
-getting trapped in local minima.
-
-Our model was trained on the M3 high-performance computing platform,
-using TensorFlow and Keras. During training, 80\% of the training data
-was utilized for actual training, while the remaining 20\% was used as
-validation data. The Bayesian optimization tuner conducted 100 trials to
-identify the best hyperparameter values based on validation root mean
-square error. The tuner then restored the best epoch of the best model
-from the trials. Additionally, we applied early stopping, terminating
-the training process if the validation root mean square error fails to
-improve for 50 epochs. The maximum allowed epochs was set at 2000,
-although no models reached this threshold.
-
-\begin{table}
-
-\caption{\label{tab:hyperparameter}Name of hyperparameters and their correspoding domain for the computer vision model.}
-\centering
-\begin{tabular}[t]{ll}
-\toprule
-Hyperparameter & Domain\\
-\midrule
-Number of base filters & \{4, 8, 16, 32, 64\}\\
-Dropout rate for convolutional blocks & {}[0.1, 0.6]\\
-Batch normalization for convolutional blocks & \{false, true\}\\
-Type of global pooling & \{max, average\}\\
-Ignore additional inputs & \{false, true\}\\
-\addlinespace
-Number of units for the fully-connected layer & \{128, 256, 512, 1024, 2048\}\\
-Batch normalization for the fully-connected layer & \{false, true\}\\
-Dropout rate for the fully-connected layer & {}[0.1, 0.6]\\
-Learning rate & {}[1e-8, 1e-1]\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\subsection{Model evaluation methods}\label{model-evaluation-methods}
-
-This section remains incomplete as we may modify the ``Results'' section
-later on.
-
-\section{Results}\label{results}
-
-\subsection{Optimized hyperparameters}\label{optimized-hyperparameters}
 
@@
 
+\subsection{Model Training}\label{model-training}
+
+To achieve an optimal deep learning model, hyperparameters like the
+learning rate often need to be fine-tuned using a tuner. In our study,
+we utilized the Bayesian optimization tuner from the \texttt{KerasTuner}
+Python library \citep{omalley2019kerastuner} for this purpose. A
+comprehensive list of hyperparameters is provided in Table
+\ref{tab:hyperparameter}.
+
+The ``number of base filters'' determines the number of filters for the
+first 2D convolutional layer. In the VGG16 architecture, the number of
+filters for the 2D convolutional layer in a block is typically twice the
+number in the previous block, except for the last block, which maintains
+the same number of convolution filters as the previous one. This
+hyperparameter aids in controlling the complexity of the computer vision
+model. Higher numbers of base filters result in more trainable
+parameters. Likewise, the ``number of units for the fully-connected
+layer'' determines the complexity of the final prediction block.
+Increasing the number of units enhances model complexity, resulting in
+more trainable parameters.
+
+The dropout rate and batch normalization are flexible hyperparameters
+that work in conjunction with other parameters to facilitate smooth
+training. A higher dropout rate is necessary when the model tends to
+overfit the training dataset by learning too much noise. Conversely, a
+lower dropout rate is preferred when the model is complex and
+challenging to converge. Batch normalization, on the other hand,
+addresses the internal covariate shift problem arising from the
+randomness in weight initialization and input data. It helps stabilize
+and accelerate the training process by normalizing the activations of
+each layer.
+
+Additionally, incorporating additional inputs such as scagnostics and
+the number of observations can potentially enhance prediction accuracy.
+Therefore, we allowed the tuner to determine whether these inputs were
+necessary for optimal model performance.
+
+The learning rate is a crucial hyperparameter, as it dictates the step
+size of the optimization algorithm. A high learning rate can help the
+model avoid local minima but risks overshooting and missing the global
+minimum. Conversely, a low learning rate smoothens the training process
+but makes the convergence time longer and increases the likelihood of
+getting trapped in local minima.
+
+Our model was trained on the M3 high-performance computing platform,
+using TensorFlow and Keras. During training, 80\% of the training data
+was utilized for actual training, while the remaining 20\% was used as
+validation data. The Bayesian optimization tuner conducted 100 trials to
+identify the best hyperparameter values based on validation root mean
+square error. The tuner then restored the best epoch of the best model
+from the trials. Additionally, we applied early stopping, terminating
+the training process if the validation root mean square error fails to
+improve for 50 epochs. The maximum allowed epochs was set at 2000,
+although no models reached this threshold.
+
+\begin{table}
+
+\caption{\label{tab:hyperparameter}Name of hyperparameters and their correspoding domain for the computer vision model.}
+\centering
+\begin{tabular}[t]{ll}
+\toprule
+Hyperparameter & Domain\\
+\midrule
+Number of base filters & \{4, 8, 16, 32, 64\}\\
+Dropout rate for convolutional blocks & {}[0.1, 0.6]\\
+Batch normalization for convolutional blocks & \{false, true\}\\
+Type of global pooling & \{max, average\}\\
+Ignore additional inputs & \{false, true\}\\
+\addlinespace
+Number of units for the fully-connected layer & \{128, 256, 512, 1024, 2048\}\\
+Batch normalization for the fully-connected layer & \{false, true\}\\
+Dropout rate for the fully-connected layer & {}[0.1, 0.6]\\
+Learning rate & {}[1e-8, 1e-1]\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\subsection{Evaluation Methods}\label{evaluation-methods}
+
+This section remains incomplete as we may modify the ``Results'' section
+later on.
+
+\section{Results}\label{results}
+
+understand what the convolutional layer is actually doing (read vgg16
+doc)
+
 \subsection{Summary of model
@@
 
-\subsection{Demostration of MVI}\label{demostration-of-mvi}
-
-As described in Section \ref{sec:model-violations-index} a Model
-Violations Index (MVI) can be constructed using \(\hat{D}\). This
-section provides guidelines and examples of MVI.
-
-Figures \ref{fig:poly-index} and \ref{fig:heter-index} display the
-residual plots for fitted models exhibiting varying degrees of
-non-linearity and heteroskedasticity. Each residual plot's MVI is
-computed using Equation \ref{eq:mvi} with \(C = 10\). When
-\(\text{MVI} > 8\), the visual patterns are notably strong and easily
-discernible by humans. In the range \(6 < \text{MVI} < 8\), the
-visibility of the visual pattern diminishes as MVI decreases.
-Conversely, when \(\text{MVI} < 6\), the visual pattern tends to become
-relatively faint and challenging to observe. Table \ref{tab:mvi}
-provides a summary of the MVI usage and it is applicable to other linear
-regression models.
-
-\begin{table}
-
-\caption{\label{tab:mvi}Degree of model violations or the strength of the visual signals according to the Model Violation Index (MVI). The constant $C$ is set to be 10.}
-\centering
-\begin{tabular}[t]{ll}
-\toprule
-Degree of model violations & Range ($C$ = 10)\\
-\midrule
-Strong & $\text{MVI} > 8$\\
-Moderate & $6 < \text{MVI} < 8$\\
-Weak & $\text{MVI} < 6$\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{figure}[!h]
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/poly-index-1}
-
-}
-
-\caption{Residual plots generated from fitted models exhibiting varying degrees of non-linearity violations.
-The Model violations index (MVI) is displayed atop each residual plot.
-The non-linearity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$.}\label{fig:poly-index}
-\end{figure}
-
-\begin{figure}[!h]
-
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/heter-index-1}
-
-}
-
-\caption{Residual plots generated from fitted models exhibiting varying degrees of heteroskedasticity violations.
-The Model violations index (MVI) is displayed atop each residual plot.
-The heteroskedasticity patterns are relatively strong for $MVI > 8$, and relatively weak for $MVI < 6$.}\label{fig:heter-index}
-\end{figure}
-
 \subsection{Comparison with human visual inference and conventional
@@
 human responses to \(1152\) lineups. Each lineup contains one randomly
-placed actual residual plot and 19 null plots. Among the \(1152\)
-lineups, \(24\) are attention check lineups in which the visual patterns
-are designed to be extremely obvious and very different from the
+placed true residual plot and 19 null plots. Among the \(1152\) lineups,
+\(24\) are attention check lineups in which the visual patterns are
+designed to be extremely obvious and very different from the
 corresponding to null plots, \(36\) are null lineups where all the
@@
 
-For each lineup used in \citet{li2023plot}, there is one actual residual
-plot and 19 null plots. While the distance \(D\) for the actual residual
+For each lineup used in \citet{li2023plot}, there is one true residual
+plot and 19 null plots. While the distance \(D\) for the true residual
 plot depends on the underlying data generating process, the distance
 \(D\) for the null plots is zero. We have used our optimized computer
-vision model to predict the approximated distance for both the actual
+vision model to predict the approximated distance for both the true
 residual plots and the null plots. To have a fair comparison, \(H_0\)
-will be rejected if the actual residual plot has the greatest
-approximated distance among all plots in a lineup. Additionally, the
-appropriate conventional tests including the Ramsey Regression Equation
+will be rejected if the true residual plot has the greatest approximated
+distance among all plots in a lineup. Additionally, the appropriate
+conventional tests including the Ramsey Regression Equation
 Specification Error Test (RESET) \citep{ramsey1969tests} for
@@
 
-The performance metrics of \(\hat{D}\) for actual residual plots are
+The performance metrics of \(\hat{D}\) for true residual plots are
 outlined in Table \ref{tab:experiment-performance}. It's notable that
@@
 \(\delta\)-difference suggests that there is at least one null plot in
-the lineup with a stronger visual signal than the actual residual plot.
-In some instances, the weighted detection rate is close to one, yet the
+the lineup with a stronger visual signal than the true residual plot. In
+some instances, the weighted detection rate is close to one, yet the
 \(\delta\)-difference is negative. This discrepancy implies that the
@@
 Notably, if the residual plot were evaluated through a visual test, it
-would not be rejected since the actual residual plot positioned at 10
-can not be distinguished from the others.
+would not be rejected since the true residual plot positioned at 10 can
+not be distinguished from the others.
 
@@
 \caption{A lineup of residual plots displaying "left-triangle" visual patterns.
-The actual residual plot occupies position 10, yet there are no discernible visual patterns that distinguish it from the other plots.}\label{fig:false-lineup}
+The true residual plot occupies position 10, yet there are no discernible visual patterns that distinguish it from the other plots.}\label{fig:false-lineup}
 \end{figure}
@@
 \ref{fig:boston-lineup} underscores that the ``U'' shape is a
-distinctive characteristic of the actual residual plot, absent in any of
+distinctive characteristic of the true residual plot, absent in any of
 the null plots. Consequently, if a visual test is conducted, \(H_0\)
@@
 
-\caption{A lineup of residual plots for the Boston housing fitted model. The actual residual plot is at position 7. It can be easily identified as the most different plot.}\label{fig:boston-lineup}
+\caption{A lineup of residual plots for the Boston housing fitted model. The true residual plot is at position 7. It can be easily identified as the most different plot.}\label{fig:boston-lineup}
 \end{figure}
@@
 \caption{A lineup of residual plots for the fitted model on the "dinosaur" dataset.
-The actual residual plot is at position 17.
+The true residual plot is at position 17.
 It can be easily identified as the most different plot as the visual pattern is extremly artificial.}\label{fig:dino-lineup}
 \end{figure}
+
+\section{Extension to other types of regression
+model}\label{extension-to-other-types-of-regression-model}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Use raw residuals.
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  D (effect size) for null plots will be greater than zero.
+\item
+  Can not assume D for actual plots are greater than D for null plots
+  any more. -\textgreater{} KL null - normal, KL actual - normal
+\item
+  Need to use a two-sided test instead of a one-sided test.
+\item
+  It may fail completely due to the identification issue.
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  Use transformed residuals that roughly follow a multivariate normal
+  distribution.
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  The one-sided test will still work.
+\item
+  Need to ensure that the transformed actual plot will have residual
+  distribution different from the transformed null distribution.
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  Train a new model.
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  KL divergence may not be the appropriate measure for the difference
+  between the residual distributions. Could consider using a new
+  distance measure.
+\end{itemize}
 

### 2024-05-13 paper/appendix.Rmd
revise up to model arch
up to examples part

@@
 For reproducibility, we excluded these scagnostics from the training data.
+
+# Neural Network Layers Used in the Study
+
+This study employs seven types of neural network layers, all of which are standard components frequently found in modern deep learning models.
+These layers are well-documented in textbooks like @goodfellow2016deep, which offer thorough explanations and mathematical insights.
+In this section, we will offer a concise overview of these layers, drawing primarily from the insights provided in @goodfellow2016deep.
+
+## Dense Layer
+
+The Dense layer, also known as the fully-connected layer, is the fundamental unit in neural networks.
+It conducts a matrix multiplication operation between the input matrix $\boldsymbol{X}$ and a weight matrix $\boldsymbol{W}$ to generate the output matrix $\boldsymbol{O}$, which can be written as
+
+$$\boldsymbol{O} = \boldsymbol{X}\boldsymbol{W} + b,$$
+
+where $b$ is the intercept.
+
+## ReLu Layer
+
+The ReLU layer, short for rectified linear unit, is an element-wise non-linear function introduced by @nair2010rectified.
+It sets the output elements to zero if the corresponding input element is negative; otherwise, it retains the original input.
+Mathematically, it can be expressed as:
+
+$$\boldsymbol{O}(i,j) = max\{0, \boldsymbol{X}(i,j)\},$$
+
+where $\boldsymbol{O}(i,j)$ is the $i$th row and $j$th column entry of matrix $\boldsymbol{O}$, and $\boldsymbol{X}(i,j)$ is the $i$th row and $j$th column entry of matrix $\boldsymbol{X}$.
+
+## Convolutaional Layer
+
+In Dense layers, matrix multiplication leads to each output unit interacting with every input unit, whereas convolutional layers operate differently with sparse interactions.
+Here, an output unit in a convolutional layer is connected solely to a subset of input units, and the weight is shared across all input units.
+Achieving this involves using a kernel, typically a small square matrix, to conduct matrix multiplication across all input units.
+Precisely, this concept can be formulated as:
+
+$$O(i, j) = \sum_m\sum_nI(i - m, j - n)K(m, n),$$
+
+where $m$ and $n$ are the row and columns indices of the kernel $K$.
+
+If there are multiple kernels used in one covolutional layer, then each kernel will have its own weights and the output will be a three-dimensional tensor, where the length of the third channel is the number of kernels.
+
+## Pooling Layer
+
+A pooling layer substitutes the input at a specific location with a summary statistic derived from nearby input units.
+Typically, there are two types of pooling layers: max pooling and average pooling.
+Max pooling computes the maximum value within a rectangular neighborhood, while average pooling calculates their average.
+Pooling layers helps making the representation approximately invariant to minor translations of the input.
+The output matrix of a pooling layer is approximately $s$ times smaller than the input matrix, where $s$ represents the length of the rectangular area.
+This can be formulated as:
+
+$$O(i, j) = \underset{m,n}{\max} I(si + m,sj+n).$$
+
+## Global Pooling Layer
+
+A global pooling layer condenses an input matrix into a scalar value by either extracting the maximum or computing the average across all elements.
+This layer acts as a crucial link between the convolutional structure and the subsequent dense layers in a neural network architecture.
+When convolutional layers utilize multiple kernels, the output becomes a three-dimensional tensor with numerous channels.
+In this scenario, the global pooling layer treats each channel individually, much like distinct features in a conventional classifier.
+This approach facilitates the extraction of essential features from complex data representations, enhancing the network's ability to learn meaningful patterns.
+
+## Batch Normalization Layer
+
+## Dropout Layer
+
+<!-- ## visualize the convolutional blocks -->

### 2024-05-13 paper/appendix.tex
up to examples part

% interactcadsample.tex
% v1.03 - April 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{lscape}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\usepackage{setspace}
\doublespacing

\begin{document}

\articletype{Draft paper (Incomplete)}

\title{Appendix: Automated assessment of residual plots with computer
vision models}

\author{\name{Weihao Li$^{a}$, Dianne Cook$^{a}$, Emi Tanaka$^{b,
c}$, Susan VanderPlas$^{d}$, Klaus Ackermann$^{a}$}
\affil{$^{a}$Department of Econometrics and Business Statistics, Monash
University, Clayton, VIC, Australia; $^{b}$Biological Data Science
Institute, Australian National University, Acton, ACT,
Australia; $^{c}$Research School of Finance, Actuarial Studies and
Statistics, Australian National University, Acton, ACT,
Australia; $^{d}$Department of Statistics, University of Nebraska,
Lincoln, Nebraska, USA}
}

\thanks{CONTACT Weihao
Li. Email: \href{mailto:weihao.li@monash.edu}{\nolinkurl{weihao.li@monash.edu}}, Dianne
Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, Emi
Tanaka. Email: \href{mailto:emi.tanaka@anu.edu.au}{\nolinkurl{emi.tanaka@anu.edu.au}}, Susan
VanderPlas. Email: \href{mailto:susan.vanderplas@unl.edu}{\nolinkurl{susan.vanderplas@unl.edu}}, Klaus
Ackermann. Email: \href{mailto:Klaus.Ackermann@monash.edu}{\nolinkurl{Klaus.Ackermann@monash.edu}}}

\maketitle

\begin{abstract}
TBD.
\end{abstract}

\begin{keywords}
TBD
\end{keywords}

\section{Additional details about data
generation}\label{additional-details-about-data-generation}

\subsection{Computation of
scagnostics}\label{computation-of-scagnostics}

In Section \ref{introduction}, we mentioned that scagnostics consist of
a set of manually designed visual feature extraction functions. While
our computer vision model will learn its own feature extraction function
during training, leveraging additional information from scagnostics can
enhance the model's predictive accuracy.

For each generated residual plot, we computed four scagnostics --
``Monotonic,'' ``Sparse,'' ``Splines,'' and ``Striped'' -- using the
\texttt{cassowaryr} R package \citep{mason2022cassowaryr}. These
computed measures, along with the number of observations from the fitted
model, were provided as the second input for the computer vision model.
Although other scagnostics are informative, they are currently
unavailable due to a fatal bug in the compiled C program of the
\texttt{interp} R package \citep{Albrecht2023interp}, which may
unpredictably crash the process. For reproducibility, we excluded these
scagnostics from the training data.

\section{Neural Network Layers Used in the
Study}\label{neural-network-layers-used-in-the-study}

This study employs seven types of neural network layers, all of which
are standard components frequently found in modern deep learning models.
These layers are well-documented in textbooks like
\citet{goodfellow2016deep}, which offer thorough explanations and
mathematical insights. In this section, we will offer a concise overview
of these layers, drawing primarily from the insights provided in
\citet{goodfellow2016deep}.

\subsection{Dense Layer}\label{dense-layer}

The Dense layer, also known as the fully-connected layer, is the
fundamental unit in neural networks. It conducts a matrix multiplication
operation between the input matrix \(\boldsymbol{X}\) and a weight
matrix \(\boldsymbol{W}\) to generate the output matrix
\(\boldsymbol{O}\), which can be written as

\[\boldsymbol{O} = \boldsymbol{X}\boldsymbol{W} + b,\]

where \(b\) is the intercept.

\subsection{ReLu Layer}\label{relu-layer}

The ReLU layer, short for rectified linear unit, is an element-wise
non-linear function introduced by \citet{nair2010rectified}. It sets the
output elements to zero if the corresponding input element is negative;
otherwise, it retains the original input. Mathematically, it can be
expressed as:

\[\boldsymbol{O}(i,j) = max\{0, \boldsymbol{X}(i,j)\},\]

where \(\boldsymbol{O}(i,j)\) is the \(i\)th row and \(j\)th column
entry of matrix \(\boldsymbol{O}\), and \(\boldsymbol{X}(i,j)\) is the
\(i\)th row and \(j\)th column entry of matrix \(\boldsymbol{X}\).

\subsection{Convolutaional Layer}\label{convolutaional-layer}

In Dense layers, matrix multiplication leads to each output unit
interacting with every input unit, whereas convolutional layers operate
differently with sparse interactions. Here, an output unit in a
convolutional layer is connected solely to a subset of input units, and
the weight is shared across all input units. Achieving this involves
using a kernel, typically a small square matrix, to conduct matrix
multiplication across all input units. Precisely, this concept can be
formulated as:

\[O(i, j) = \sum_m\sum_nI(i - m, j - n)K(m, n),\]

where \(m\) and \(n\) are the row and columns indices of the kernel
\(K\).

If there are multiple kernels used in one covolutional layer, then each
kernel will have its own weights and the output will be a
three-dimensional tensor, where the length of the third channel is the
number of kernels.

\subsection{Pooling Layer}\label{pooling-layer}

A pooling layer substitutes the input at a specific location with a
summary statistic derived from nearby input units. Typically, there are
two types of pooling layers: max pooling and average pooling. Max
pooling computes the maximum value within a rectangular neighborhood,
while average pooling calculates their average. Pooling layers helps
making the representation approximately invariant to minor translations
of the input. The output matrix of a pooling layer is approximately
\(s\) times smaller than the input matrix, where \(s\) represents the
length of the rectangular area. This can be formulated as:

\[O(i, j) = \underset{m,n}{\max} I(si + m,sj+n).\]

\subsection{Global Pooling Layer}\label{global-pooling-layer}

A global pooling layer condenses an input matrix into a scalar value by
either extracting the maximum or computing the average across all
elements. This layer acts as a crucial link between the convolutional
structure and the subsequent dense layers in a neural network
architecture. When convolutional layers utilize multiple kernels, the
output becomes a three-dimensional tensor with numerous channels. In
this scenario, the global pooling layer treats each channel
individually, much like distinct features in a conventional classifier.
This approach facilitates the extraction of essential features from
complex data representations, enhancing the network's ability to learn
meaningful patterns.

\subsection{Batch Normalization Layer}\label{batch-normalization-layer}

\subsection{Dropout Layer}\label{dropout-layer}

\bibliographystyle{tfcad}
\bibliography{ref.bib}

\end{document}

### 2024-05-13 paper/paper.Rmd
up to examples part

@@
 
-## Model Formulas
-
-There are various configurations of the computer vision model that can be used to assess residual plots.
-We discuss these configurations below focusing on two key components of the model formula: the input and the output format.
+## Model Specifications
+
+There are various specifications of the computer vision model that can be used to assess residual plots.
+We discuss these specifications below focusing on two key components of the model formula: the input and the output format.
 
@@
 
-```{r out.width = "100%", fig.cap = "Diagram of the architecture of the optimized computer vision model."}
+```{r cnn-diag, out.width = "100%", fig.cap = "Diagram of the architecture of the optimized computer vision model."}
 knitr::include_graphics("cnn.pdf")
@@
 The architecture of the computer vision model is adapted from a well-established architecture known as VGG16, which has demonstrated high performance in image classification [@simonyan2014very].
+Figure \ref{fig:cnn-diag} provides a diagram of the architecture.
+More details about the neural network layers used in this study are provided in the Appendix.
 
@@
 
-Based on the tuning process described in Section \ref{training-and-hyperparameter-tuning}, the optimized hyperparameter values are presented in Table \ref{tab:best-hyperparameter}.
+## Model Training
+
+To achieve an optimal deep learning model, hyperparameters like the learning rate often need to be fine-tuned using a tuner.
+In our study, we utilized the Bayesian optimization tuner from the `KerasTuner` Python library [@omalley2019kerastuner] for this purpose.
+A comprehensive list of hyperparameters is provided in Table \ref{tab:hyperparameter}.
+
+The "number of base filters" determines the number of filters for the first 2D convolutional layer.
+In the VGG16 architecture, the number of filters for the 2D convolutional layer in a block is typically twice the number in the previous block, except for the last block, which maintains the same number of convolution filters as the previous one.
+This hyperparameter aids in controlling the complexity of the computer vision model.
+Higher numbers of base filters result in more trainable parameters.
+Likewise, the "number of units for the fully-connected layer" determines the complexity of the final prediction block.
+Increasing the number of units enhances model complexity, resulting in more trainable parameters.
+
+The dropout rate and batch normalization are flexible hyperparameters that work in conjunction with other parameters to facilitate smooth training.
+A higher dropout rate is necessary when the model tends to overfit the training dataset by learning too much noise.
+Conversely, a lower dropout rate is preferred when the model is complex and challenging to converge.
+Batch normalization, on the other hand, addresses the internal covariate shift problem arising from the randomness in weight initialization and input data.
+It helps stabilize and accelerate the training process by normalizing the activations of each layer.
+
+Additionally, incorporating additional inputs such as scagnostics and the number of observations can potentially enhance prediction accuracy.
+Therefore, we allowed the tuner to determine whether these inputs were necessary for optimal model performance.
+
+The learning rate is a crucial hyperparameter, as it dictates the step size of the optimization algorithm.
+A high learning rate can help the model avoid local minima but risks overshooting and missing the global minimum.
+Conversely, a low learning rate smoothens the training process but makes the convergence time longer and increases the likelihood of getting trapped in local minima.
+
+Our model was trained on the M3 high-performance computing platform, using TensorFlow and Keras.
+During training, 80% of the training data was utilized for actual training, while the remaining 20% was used as validation data.
+The Bayesian optimization tuner conducted 100 trials to identify the best hyperparameter values based on validation root mean square error.
+The tuner then restored the best epoch of the best model from the trials.
+Additionally, we applied early stopping, terminating the training process if the validation root mean square error fails to improve for 50 epochs.
+The maximum allowed epochs was set at 2000, although no models reached this threshold.
+
+```{r hyperparameter}
+data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
+           Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[1e-8, 1e-1]")) %>%
+  kableExtra::kable("latex", booktabs = TRUE, caption = "Name of hyperparameters and their correspoding domain for the computer vision model.")
+```
+
+Based on the tuning process described above, the optimized hyperparameter values are presented in Table \ref{tab:best-hyperparameter}.
 It was observable that a minimum of $32$ base filters was necessary, with the preferable choice being $64$ base filters for both the $64 \times 64$ and $128 \times 128$ models, mirroring the original VGG16 architecture.
@@
 
-## Model Training
-
-To achieve an optimal deep learning model, hyperparameters like the learning rate often need to be fine-tuned using a tuner.
-In our study, we utilized the Bayesian optimization tuner from the `KerasTuner` Python library [@omalley2019kerastuner] for this purpose.
-A comprehensive list of hyperparameters is provided in Table \ref{tab:hyperparameter}.
-
-The "number of base filters" determines the number of filters for the first 2D convolutional layer.
-In the VGG16 architecture, the number of filters for the 2D convolutional layer in a block is typically twice the number in the previous block, except for the last block, which maintains the same number of convolution filters as the previous one.
-This hyperparameter aids in controlling the complexity of the computer vision model.
-Higher numbers of base filters result in more trainable parameters.
-Likewise, the "number of units for the fully-connected layer" determines the complexity of the final prediction block.
-Increasing the number of units enhances model complexity, resulting in more trainable parameters.
-
-The dropout rate and batch normalization are flexible hyperparameters that work in conjunction with other parameters to facilitate smooth training.
-A higher dropout rate is necessary when the model tends to overfit the training dataset by learning too much noise.
-Conversely, a lower dropout rate is preferred when the model is complex and challenging to converge.
-Batch normalization, on the other hand, addresses the internal covariate shift problem arising from the randomness in weight initialization and input data.
-It helps stabilize and accelerate the training process by normalizing the activations of each layer.
-
-Additionally, incorporating additional inputs such as scagnostics and the number of observations can potentially enhance prediction accuracy.
-Therefore, we allowed the tuner to determine whether these inputs were necessary for optimal model performance.
-
-The learning rate is a crucial hyperparameter, as it dictates the step size of the optimization algorithm.
-A high learning rate can help the model avoid local minima but risks overshooting and missing the global minimum.
-Conversely, a low learning rate smoothens the training process but makes the convergence time longer and increases the likelihood of getting trapped in local minima.
-
-Our model was trained on the M3 high-performance computing platform, using TensorFlow and Keras.
-During training, 80% of the training data was utilized for actual training, while the remaining 20% was used as validation data.
-The Bayesian optimization tuner conducted 100 trials to identify the best hyperparameter values based on validation root mean square error.
-The tuner then restored the best epoch of the best model from the trials.
-Additionally, we applied early stopping, terminating the training process if the validation root mean square error fails to improve for 50 epochs.
-The maximum allowed epochs was set at 2000, although no models reached this threshold.
-
-```{r hyperparameter}
-data.frame(`Hyperparameter` = c("Number of base filters", "Dropout rate for convolutional blocks", "Batch normalization for convolutional blocks", "Type of global pooling", "Ignore additional inputs", "Number of units for the fully-connected layer", "Batch normalization for the fully-connected layer", "Dropout rate for the fully-connected layer", "Learning rate"),
-           Domain = c("{4, 8, 16, 32, 64}", "[0.1, 0.6]", "{false, true}", "{max, average}", "{false, true}", "{128, 256, 512, 1024, 2048}", "{false, true}", "[0.1, 0.6]", "[1e-8, 1e-1]")) %>%
-  kableExtra::kable("latex", booktabs = TRUE, caption = "Name of hyperparameters and their correspoding domain for the computer vision model.")
-```
-
-## Evaluation Methods
-
-This section remains incomplete as we may modify the "Results" section later on.
+<!-- ## Evaluation Methods -->
+
+<!-- This section remains incomplete as we may modify the "Results" section later on. -->
 
@@
 
-understand what the convolutional layer is actually doing (read vgg16 doc)
-
-## Summary of model performance
+## Model Performance
 
@@
 The figure highlights that most under-predictions occurred when $\hat{D} < 3$, while over-predictions occurred predominantly when $3 < \hat{D} < 6$.
-To comprehend the reasons behind these deviations, we fitted a regression tree using the `rpart` R package [@Terry2022rpart] on $D - \hat{D}$ for all factors used in the data generating process.
-The results of the regression tree are provided in the Appendix.
-
-The regression tree revealed that most issues arose from non-linearity problems and the presence of a second predictor in the regression model.
-When the error distribution had a very small variance, all computer vision models tended to under-predict the distance.
-Conversely, when the error distribution had a large variance, all models tended to over-predict the distance.
-Furthermore, for input images representing null plots, it was expected that the models will over-predict the distance, as explained in Section \ref{lineup-evaluation}.
-
-Since most of the deviation stemmed from the presence of non-linearity violations, to further investigate this, we split the test set and re-evaluated the performance, as detailed in Table \ref{tab:performance-sub}.
+For input images representing null plots, it was expected that the models will over-predict the distance, as explained in Section \ref{lineup-evaluation}.
+However, it can not explain the under-prediction issue.
+Therefore, we analysed the relationship between residuals and all the factors involved in the data generating process.
+We found that most issues actually arose from non-linearity problems and the presence of a second predictor in the regression model as illustrated in Figure \ref{fig:over-under}.
+When the variance for the error distribution was small, the optimized model tended to under-predict the distance.
+Conversely, when the error distribution had a large variance, the model tended to over-predict the distance.
+
+Since most of the deviation stemmed from the presence of non-linearity violations, to further investigate this, we split the test set based on violation types and re-evaluated the performance, as detailed in Table \ref{tab:performance-sub}.
 It was evident that metrics for null plots were notably worse compared to other categories.
@@
 
+```{r over-under, fig.pos="!h", fig.cap = "Scatter plots for residuals vs $\\sigma$ on test data for the $32 \\times 32$ optimized model.
+The data is grouped by whether the regression has only non-linearity violation, and whether it includes a second predictor in the regression formula.
+The red lines are smoothing curves produced by fitting gnealized additive models.
+The area over the zero line in light pink indicates under-prediction, and the area under the zero line in light green indicates over-prediction.", dev = 'png', dpi = 300}
+model_pred %>%
+  filter(res == 32L) %>%
+  filter(type == "test") %>%
+  mutate(only_z = include_z & !include_heter & !include_non_normal) %>%
+  rename(`Only has non-linearity violation` = only_z, `Has second predictor` = include_x2) %>%
+  ggplot() +
+  geom_hline(yintercept = 0, alpha = 0.5) +
+  annotate("rect", 
+           ymin = min(model_pred$effect_size[model_pred$type == "test"] - model_pred$vss[model_pred$type == "test"]),
+                ymax = 0,
+                xmin = min(model_pred$e_sigma),
+                xmax = max(model_pred$e_sigma),
+            fill = "#66c2a5",
+            alpha = 0.3) +
+  annotate("rect", 
+           ymin = 0,
+                ymax = max(model_pred$effect_size[model_pred$type == "test"] - model_pred$vss[model_pred$type == "test"]),
+                xmin = min(model_pred$e_sigma),
+                xmax = max(model_pred$e_sigma),
+            fill = "#fc8d62",
+            alpha = 0.3) +
+  geom_point(aes(e_sigma, effect_size - vss), alpha = 0.2) +
+  geom_smooth(aes(e_sigma, effect_size - vss), se = FALSE, col = "#e34a33") +
+  facet_grid(`Has second predictor` ~ `Only has non-linearity violation`,
+             labeller = label_both) +
+  theme_light() +
+  ylab(expression(D - hat(D))) +
+  xlab(expression(sigma))
+```
+
 ```{r performance-sub}
@@
 
-## Comparison with human visual inference and conventional tests
-
-### Overview of the human subject experiment
+<!-- - check if j has an impact on residuals more carefully -->
+<!-- - the RMSE has very little to do with j if x2 is not included. And its impact is not as large as sigma. -->
+
+## Comparison with Human Visual Inference and Conventional Tests
+
+### Overview of the Human Subject experiment
 
 In order to check the validity of the proposed computer vision model, residual plots presented in the human subject experiment conducted by @li2023plot will be assessed.
+
+<!-- The experiment revealed that conventional tests are more sensitive to weak departures from model assumptions than visual test, and they often reject the null hypothesis when departures are not visibly different from null residual plots.
+-->
+
 This study has collected $7974$ human responses to $1152$ lineups.
@@
 While the distance $D$ for the true residual plot depends on the underlying data generating process, the distance $D$ for the null plots is zero.
-We have used our optimized computer vision model to predict the approximated distance for both the true residual plots and the null plots.
+We have used our optimized computer vision model to estimate distance for both the true residual plots and the null plots.
 To have a fair comparison, $H_0$ will be rejected if the true residual plot has the greatest approximated distance among all plots in a lineup.
@@
 These figures are higher than those calculated for visual tests conducted by human, indicating that the computer vision model exhibits behaviour more akin to the best available conventional tests.
-However, Figure \ref{fig:conv-mosaic} shows that conventional tests tend to reject residual plots almost exclusively when the computer vision model does, but they also reject plots when the computer vision model does not.
+However, Figure \ref{fig:conv-mosaic} shows that the computer vision model does not always reject when the conventional tests reject.
+And a small number of plots will be rejected by computer vision model but not by conventional tests.
 This suggests that conventional tests are more sensitive than the computer vision model.
+
+Figure \ref{fig:pcp} further illustrates the decisions made by visual tests conducted by human, computer vision models, and conventional tests, using a parallel coordinate plots.
+It can be observed that all three tests will agree with each other for around 50% of the cases.
+When visual tests conducted by human do not reject, there are substantial amount of cases where computer vision model also not reject but conventional tests reject.
+There are much fewer cases that do not reject by visual tests and conventional tests, but reject by computer vision models.
+This indicates computer vision model can behave like visual tests conducted by human better than conventional tests.
+Moreover, there are great proportion of cases where visual tests conducted by human is the only test who does not reject, .
 
@@
   ggplot() +
-  geom_mosaic(aes(x = ggmosaic::product(model_reject_20, reject), 
+  geom_mosaic(aes(x = ggmosaic::product(reject, model_reject_20), 
                   fill = model_reject_20)) +
   facet_grid(~type) +
-  xlab("Visual tests reject") +
-  ylab("Computer vision model rejects") +
+  ylab("Visual tests reject") +
+  xlab("Computer vision model rejects") +
   # labs(fill = "Visual tests reject") +
@@
 
-```{r conv-mosaic, fig.pos = "!h", fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of conventional test conditional on the computer vision model decision on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot.
-The conventional test rejects more frequently than the computer vision model, and (almost) only rejects when the computer vision model does."}
+```{r conv-mosaic, fig.pos = "!h", fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of computer vision models conditional on conventional tests on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot."}
 library(ggmosaic)
@@
   ggplot() +
-  geom_mosaic(aes(x = ggmosaic::product(conventional_reject, model_reject_20), 
-                  fill = conventional_reject)) +
+  geom_mosaic(aes(x = ggmosaic::product(model_reject_20, conventional_reject), 
+                  fill = model_reject_20)) +
   facet_grid(~type) +
-  xlab("Computer vision model rejects") +
-  ylab("Conventional tests reject ") +
+  ylab("Computer vision model rejects") +
+  xlab("Conventional tests reject ") +
   # labs(fill = "Visual tests reject") +
@@
 
-```{r venn, fig.pos = "!h", fig.cap = "Venn diagram for non-rejection decisions made by conventional tests, computer vision model, and visual tests conducted by human."}
+```{r venn, fig.pos = "!h", fig.cap = "Venn diagram for non-rejection decisions made by conventional tests, computer vision model, and visual tests conducted by human.", eval = FALSE}
 poly_venn <- magick::image_read(here::here("paper/diagram/poly_venn_diagramm.png")) %>%
@@
 
-## Data examples
+# Examples
+
+<!-- - explain the purpose of each case here as a summary -->
+<!-- - left-triangle: conventional reject, other two do not reject -->
+<!-- - Boston: all three reject -->
+<!-- - datasaurus: conv not reject -->
+
+<!-- - we don't have one conv reject, model reject, human not -->
+
+<!-- conv: reject 1, not 0 -->
+<!-- model: reject 1, not 0 -->
+<!-- human: reject 1, not 0 -->
+
+<!-- left:       100 -->
+<!-- boston:     111 -->
+<!-- datasaurus: 011 -->
+
+<!-- example?:   001 (probably because of outliers) -->
+<!-- try "baker" dataset -->
 
@@
 These include the dataset associated with the residual plot displaying a "left-triangle" shape, as displayed in Figure \ref{fig:false-finding}, along with the Boston housing dataset, and the "dino" datasets from the `datasaurus` R package.
+
+The first example illustrates a scenario where both the computer vision model and human visual inspection successfully avoid rejecting $H_0$ when $H_0$ is true, contrary to conventional tests.
+This underscores the necessity of visually examining the residual plot.
+
+In the second example, we encounter a more pronounced violation of the model, resulting in rejection of $H_0$ by all three tests.
+This highlights the practicality of the computer vision model, particularly for less intricate tasks.
+
+The third example presents a situation where the model deviation is non-typical.
+Here, the computer vision model and human visual inspection reject $H_0$, whereas conventional tests do not.
+This emphasizes the benefits of visual inspection and the unique advantage of the computer vision model, which, like humans, makes decisions based on visual discoveries.
+
 Additionally, we will outline the workflow for employing this model with the `autovi` R package.
 
-### Left-triangle
+## Left-triangle
 
@@
 Moreover, the bootstrapped distribution suggests that it is highly improbable for the fitted model to be misspecified as the majority of bootstrapped fitted models will not be rejected.
-
 Thus, for this particular fitted model, both the visual test and the computer vision model will not reject $H_0$.
 However, the Breusch-Pagan test will reject $H_0$ because it can not effectively utilize information from null plots.
+
+The attention map at plot B of Figure \ref{fig:false-check} suggests that the estimation is highly influenced by the top-right and bottom-right part of the residual plot, as it forms two vertices of the triangular shape.
+A principle component analysis is also performed on the output of the global pooling layer of the computer vision model.
+As mentioned in @simonyan2014very, a computer vision model built upon the convolutional blocks can be viewed as a feature extractor.
+For the $32 \times 32$ model, there are 256 features outputted from the global pooling layer, which can be further used for different visual tasks not limited to distance prediction.
+To see if these features can be effectively used for distinguishing null plots and true residual plot, we linearly project them into the PC1 vs PC2 space as shown in Figure \ref{fig:false-check}.
+It can be observed that because the bootstrapped plots are mostly similar to the null plots, the points drawn in different colours are mixed together.
+The true residual plot is also covered by both the cluster of null plots and cluster of bootstrapped plots.
+This accurately reflects our understanding of Figure \ref{fig:false-lineup}.
 
@@
   my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
-  my_vi$check(null_draws = 200L, boot_draws = 200L)
+  my_vi$check(null_draws = 200L, boot_draws = 200L, extract_feature_from_layer = "global_max_pooling2d")
   saveRDS(my_vi$check_result, here::here("paper/data/fig1_check.rds"))
@@
 
+```{r}
+define_gradient <- function() {
+  reticulate::py_run_string(glue::glue(
+"
+import tensorflow as tf
+import PIL
+import numpy as np
+
+def get_smooth_gradient(keras_mod, plot_path, target_size, input_auxiliary = None, noise_level = 0.1, n = 50):
+    im = PIL.Image.open(plot_path)
+    im = im.resize(target_size)
+    input_im = tf.keras.utils.img_to_array(im)
+    input_im = np.reshape(input_im, tuple([1]) + tuple(target_size) + tuple([3]))
+    if input_auxiliary is not None:
+        input_auxiliary = np.reshape(np.array(input_auxiliary), (1, 5))
+
+    if input_auxiliary is not None:
+        input_auxiliary = tf.Variable(input_auxiliary)
+
+    im_grad = None
+    sigma = noise_level * (np.max(input_im) - np.min(input_im))
+
+    for i in range(n):    
+        new_input_im = input_im + np.random.normal(0, sigma, input_im.shape)
+        new_input_im = tf.Variable(new_input_im)
+
+        with tf.GradientTape() as tape:
+            if input_auxiliary is not None:
+                pred = keras_mod([new_input_im, input_auxiliary])
+            else:
+                pred = keras_mod(new_input_im)
+        new_im_grad = tape.gradient(pred, new_input_im)
+        new_im_grad = tf.image.rgb_to_grayscale(new_im_grad)[0].numpy()
+        if im_grad is None:
+            im_grad = new_im_grad
+        else:
+            im_grad = im_grad + new_im_grad
+
+    im_grad = im_grad / n
+
+    input_im = tf.Variable(input_im)
+    if input_auxiliary is not None:
+        with tf.GradientTape() as tape:
+            pred = keras_mod([input_im, input_auxiliary])
+        auxiliary_grad = tape.gradient(pred, input_auxiliary)
+    if input_auxiliary is not None:
+        return (im_grad, auxiliary_grad[0].numpy())
+    else:
+        return im_grad
+
+def get_raw_gradient(keras_mod, plot_path, target_size, input_auxiliary = None):
+    im = PIL.Image.open(plot_path)
+    im = im.resize(target_size)
+    input_im = tf.keras.utils.img_to_array(im)
+    input_im = np.reshape(input_im, tuple([1]) + tuple(target_size) + tuple([3]))
+    if input_auxiliary is not None:
+        input_auxiliary = np.reshape(np.array(input_auxiliary), (1, 5))
+
+    input_im = tf.Variable(input_im)
+    if input_auxiliary is not None:
+        input_auxiliary = tf.Variable(input_auxiliary)
+
+    with tf.GradientTape() as tape:
+        if input_auxiliary is not None:
+            pred = keras_mod([input_im, input_auxiliary])
+        else:
+            pred = keras_mod(input_im)
+    im_grad = tape.gradient(pred, input_im)
+
+    if input_auxiliary is not None:
+        with tf.GradientTape() as tape:
+            pred = keras_mod([input_im, input_auxiliary])
+        auxiliary_grad = tape.gradient(pred, input_auxiliary)
+
+    im_grad = tf.image.rgb_to_grayscale(im_grad)
+    if input_auxiliary is not None:
+        return (im_grad[0].numpy(), auxiliary_grad[0].numpy())
+    else:
+        return im_grad[0].numpy()
+"
+))
+}
+
+```
+
+```{r}
+if (!file.exists(here::here("paper/data/fig1_attention_map.rds"))) {
+  define_gradient()
+  keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  plot_path <- my_vi$plot_resid() %>%
+  my_vi$save_plot()
+  input_auxiliary <- my_vi$auxiliary()
+
+  reticulate::py$get_smooth_gradient(keras_mod = keras_mod, 
+                                     plot_path = plot_path, 
+                                     target_size = c(32L, 32L), 
+                                     input_auxiliary = input_auxiliary,
+                                     noise_level = 0.5,
+                                     n = 1000L) -> smooth_g
+
+  x_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$x.range
+  y_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$y.range
+
+  smooth_att_map <- smooth_g[[1]] |> 
+    as.data.frame() |>
+    dplyr::mutate(row = rev(1:32)) |>
+    tidyr::pivot_longer(V1:V32, names_to = "column", values_to = "gradient") |>
+    dplyr::mutate(column = as.integer(gsub("V", "", column))) |>
+    dplyr::mutate(column = x_range[1] + column * (x_range[2] - x_range[1])/(32 - 1)) |>
+    dplyr::mutate(row = y_range[1] + row * (y_range[2] - y_range[1])/(32 - 1))
+
+  scale_zero_one <- function(x) (x - min(x))/(max(x) - min(x))
+
+  p <- ggplot() +
+    ggplot2::geom_raster(data = mutate(smooth_att_map, gradient = scale_zero_one(gradient)), ggplot2::aes(column, row, fill = gradient), alpha = 0.7) +
+    scale_fill_gradient(low = "black", high = "white") +
+    ggplot2::theme_void() +
+    theme(legend.position = "none")
+
+  saveRDS(p, here::here("paper/data/fig1_attention_map.rds"))
+
+}
+```
+
 ```{r false-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
-On the left, the residual plot exhibiting "left-triangle" shape is displayed.
+(A) The true residual plot exhibiting "left-triangle" shape.
+(B) The attention map produced by computing the gradient of the output with respect to the greyscale input.
+(C) The density plot of estimated distnace for null plots and bootstrapped plots.
 The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots.
-The fitted model will not be rejected since $\\hat{D} < Q_{null}(0.95)$.'}
-
-p1 <- my_vi$plot_resid()
-p2 <- my_vi$summary_plot() +
-  ggtitle("", subtitle = glue::glue("Conventional p-value = {format(HETER_MODEL$test(dat)$p_value, digits = 3)}")) +
+The fitted model will not be rejected since $\\hat{D} < Q_{null}(0.95)$.
+(D) PC2 vs PC1 plot for features extracted from the global pooling layer of the computer vision model.
+'}
+p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot") 
+p2 <- readRDS(here::here("paper/data/fig1_attention_map.rds")) + ggtitle("(B) Attention map")
+p3 <- my_vi$summary_plot() +
+  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("Conventional p-value = {format(HETER_MODEL$test(dat)$p_value, digits = 3)}")) +
   theme(legend.position = "bottom", legend.box = "vertical") +
@@
   scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D))))
-
-patchwork::wrap_plots(p1, p2)
+p4 <- my_vi$feature_pc() %>%
+  mutate(type = ifelse(type == "boot", "Boot", type)) %>%
+  mutate(type = ifelse(type == "null", "Null", type)) %>%
+  ggplot() +
+  geom_point(data = ~filter(.x, type != "observed"), aes(PC1, PC2, col = type, shape = "Simulated"), size = 2, alpha = 0.3) +
+  geom_point(data = ~filter(.x, type == "observed"), aes(PC1, PC2, shape = "Observed"), size = 4, col = "black", fill = "black") +
+  theme_light() +
+  scale_color_brewer(palette = "Dark2") +
+  scale_shape_manual(values = c("Simulated" = 19, "Observed" = 17)) +
+  xlab("PC1") +
+  ylab("PC2") +
+  labs(col = "", shape = "") +
+  theme(legend.position = "bottom", legend.box="vertical") +
+  ggtitle("(D) PCA for extracted features")
+
+patchwork::wrap_plots(p1, p2, p3, p4, nrow = 2, widths = c(1, 1), heights = c(1, 1))
 ```
@@
 
-### Boston housing
+## Boston housing
 
@@
 ```{r}
-
 housing <- read_csv(here::here("paper/data/housing.csv"))
@@
   my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
-  my_vi$check(null_draws = 200L, boot_draws = 200L)
+  my_vi$check(null_draws = 200L, boot_draws = 200L, extract_feature_from_layer = "global_max_pooling2d")
   saveRDS(my_vi$check_result, here::here("paper/data/boston_check.rds"))
@@
 my_vi$check_result <- readRDS(here::here("paper/data/boston_check.rds"))
+```
+
+```{r}
+if (!file.exists(here::here("paper/data/boston_attention_map.rds"))) {
+  define_gradient()
+  keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  plot_path <- my_vi$plot_resid() %>%
+  my_vi$save_plot()
+  input_auxiliary <- my_vi$auxiliary()
+
+  reticulate::py$get_smooth_gradient(keras_mod = keras_mod, 
+                                     plot_path = plot_path, 
+                                     target_size = c(32L, 32L), 
+                                     input_auxiliary = input_auxiliary,
+                                     noise_level = 0.5,
+                                     n = 1000L) -> smooth_g
+
+  x_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$x.range
+  y_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$y.range
+
+  smooth_att_map <- smooth_g[[1]] |> 
+    as.data.frame() |>
+    dplyr::mutate(row = rev(1:32)) |>
+    tidyr::pivot_longer(V1:V32, names_to = "column", values_to = "gradient") |>
+    dplyr::mutate(column = as.integer(gsub("V", "", column))) |>
+    dplyr::mutate(column = x_range[1] + column * (x_range[2] - x_range[1])/(32 - 1)) |>
+    dplyr::mutate(row = y_range[1] + row * (y_range[2] - y_range[1])/(32 - 1))
+
+  scale_zero_one <- function(x) (x - min(x))/(max(x) - min(x))
+
+  p <- ggplot() +
+    ggplot2::geom_raster(data = mutate(smooth_att_map, gradient = scale_zero_one(gradient)), ggplot2::aes(column, row, fill = gradient), alpha = 0.7) +
+    scale_fill_gradient(low = "black", high = "white") +
+    ggplot2::theme_void() +
+    theme(legend.position = "none")
+
+  saveRDS(p, here::here("paper/data/boston_attention_map.rds"))
+
+}
 ```
@@
 The fitted model will be rejected since $\\hat{D} >= Q_{null}(0.95)$.'}
-p1 <- my_vi$plot_resid()
-p2 <- my_vi$summary_plot() +
-  ggtitle("", subtitle = glue::glue("Conventional p-value = {format(lmtest::resettest(mod)$p.value, digits = 3)}")) +
+p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot") 
+p2 <- readRDS(here::here("paper/data/boston_attention_map.rds")) + ggtitle("(B) Attention map")
+p3 <- my_vi$summary_plot() +
+  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("Conventional p-value = {format(lmtest::resettest(mod)$p.value, digits = 3)}")) +
   theme(legend.position = "bottom", legend.box = "vertical") +
@@
 
-```{r eval = FALSE}
+## Datasaurus
+
+The computer vision model possesses the capability to detect not only typical issues like non-linearity, heteroskedasticity, and non-normality but also artifact visual patterns resembling real-world objects, as long as they do not appear in null plots.
+These visual patterns can be challenging to categorize in terms of model violations.
+Therefore, we will employ the RESET test, the Breusch-Pagan test, and the Shapiro-Wilk test [@shapiro1965analysis] for comparison.
+
+The "dino" dataset within the `datasaurus` R package exemplifies this scenario.
+With only two columns, x and y, fitting a regression model to this data yields a residual plot resembling a "dinosaur," as displayed in Figure \ref{fig:dino-check}.
+Unsurprisingly, this distinct residual plot stands out in a lineup, as shown in Figure \ref{fig:dino-lineup}.
+A visual test conducted by humans would undoubtedly reject $H_0$.
+
+According to the residual plot assessment by the computer vision model, $\hat{D}$ exceeds $Q_{null}(0.95)$, warranting a rejection of $H_0$.
+Additionally, most of the bootstrapped fitted models will be rejected, indicating an misspecified model.
+However, both the RESET test and the Breusch-Pagan test yield $p$-values greater than 0.3, leading to a non-rejection of $H_0$.
+Only the Shapiro-Wilk test rejects the normality assumption with a small $p$-value.
+
+In practice, without accessing the residual plot, it would be challenging to identify the artificial pattern of the residuals.
+Moreover, conducting a normality test for a fitted regression model is not always standard practice among analysts.
+Even when performed, violating the normality assumption is sometimes deemed acceptable, especially considering the application of quasi-maximum likelihood estimation in linear regression.
+This example underscores the importance of evaluating residual plots and highlights how the proposed computer vision model can facilitate this process.
+
+```{r}
+dino <- datasauRus::datasaurus_dozen %>% filter(dataset == "dino")
+mod <- lm(y ~ ., data = select(dino, -dataset))
+
+my_vi <- autovi::auto_vi(fitted_mod = mod)
+```
+
+```{r}
+if (!file.exists(here::here("paper/data/dino_check.rds"))) {
+  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  my_vi$check(null_draws = 200L, boot_draws = 200L, extract_feature_from_layer = "global_max_pooling2d")
+  saveRDS(my_vi$check_result, here::here("paper/data/dino_check.rds"))
+}
+```
+
+```{r}
+my_vi$check_result <- readRDS(here::here("paper/data/dino_check.rds"))
+```
+
+```{r}
+if (!file.exists(here::here("paper/data/dino_attention_map.rds"))) {
+  define_gradient()
+  keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
   plot_path <- my_vi$plot_resid() %>%
@@
   input_auxiliary <- my_vi$auxiliary()
-
-reticulate::py$get_raw_gradient(keras_mod = keras_mod, 
-                                plot_path = plot_path, 
-                                target_size = c(32L, 32L), 
-                                input_auxiliary = input_auxiliary) -> raw_g
 
@@
                                      input_auxiliary = input_auxiliary,
-                                   noise_level = 0.45,
-                                   n = 200L) -> smooth_g
+                                     noise_level = 0.5,
+                                     n = 1000L) -> smooth_g
 
@@
   y_range <- ggplot2::ggplot_build(my_vi$plot_resid())$layout$panel_params[[1]]$y.range
-
-raw_att_map <- raw_g[[1]] |> 
-  as.data.frame() |>
-  dplyr::mutate(row = rev(1:32)) |>
-  tidyr::pivot_longer(V1:V32, names_to = "column", values_to = "gradient") |>
-  dplyr::mutate(column = as.integer(gsub("V", "", column))) |>
-  dplyr::mutate(column = x_range[1] + column * (x_range[2] - x_range[1])/(32 - 1)) |>
-  dplyr::mutate(row = y_range[1] + row * (y_range[2] - y_range[1])/(32 - 1))
-
-my_vi$plot_resid() +
-  ggplot() +
-  ggplot2::geom_raster(data = raw_att_map, ggplot2::aes(column, row, fill = gradient), alpha = 0.7) +
-  scale_fill_gradient(low = "black", high = "white") +
-  # viridis::scale_fill_viridis(option = "D") +
-  ggplot2::theme_void() +
-  ggplot2::ggtitle(paste0(gsub("measure_", "", names(input_auxiliary)), ": ", format(raw_g[[2]], digits = 2), 
-                          collapse = ", ")) -> p1
 
@@
 
-# my_vi$plot_resid() +
-  ggplot() +
+  p <- ggplot() +
     ggplot2::geom_raster(data = mutate(smooth_att_map, gradient = scale_zero_one(gradient)), ggplot2::aes(column, row, fill = gradient), alpha = 0.7) +
     scale_fill_gradient(low = "black", high = "white") +
-  # viridis::scale_fill_viridis(option = "D") +
     ggplot2::theme_void() +
-  ggplot2::ggtitle(paste0(gsub("measure_", "", names(input_auxiliary)), ": ", format(raw_g[[2]], digits = 2), 
-                          collapse = ", ")) -> p2
-
-p1 + p2
-```
-
-### Datasaurus
-
-The computer vision model possesses the capability to detect not only typical issues like non-linearity, heteroskedasticity, and non-normality but also artifact visual patterns resembling real-world objects, as long as they do not appear in null plots.
-These visual patterns can be challenging to categorize in terms of model violations.
-Therefore, we will employ the RESET test, the Breusch-Pagan test, and the Shapiro-Wilk test [@shapiro1965analysis] for comparison.
-
-The "dino" dataset within the `datasaurus` R package exemplifies this scenario.
-With only two columns, x and y, fitting a regression model to this data yields a residual plot resembling a "dinosaur," as displayed in Figure \ref{fig:dino-check}.
-Unsurprisingly, this distinct residual plot stands out in a lineup, as shown in Figure \ref{fig:dino-lineup}.
-A visual test conducted by humans would undoubtedly reject $H_0$.
-
-According to the residual plot assessment by the computer vision model, $\hat{D}$ exceeds $Q_{null}(0.95)$, warranting a rejection of $H_0$.
-Additionally, most of the bootstrapped fitted models will be rejected, indicating an misspecified model.
-However, both the RESET test and the Breusch-Pagan test yield $p$-values greater than 0.3, leading to a non-rejection of $H_0$.
-Only the Shapiro-Wilk test rejects the normality assumption with a small $p$-value.
-
-In practice, without accessing the residual plot, it would be challenging to identify the artificial pattern of the residuals.
-Moreover, conducting a normality test for a fitted regression model is not always standard practice among analysts.
-Even when performed, violating the normality assumption is sometimes deemed acceptable, especially considering the application of quasi-maximum likelihood estimation in linear regression.
-This example underscores the importance of evaluating residual plots and highlights how the proposed computer vision model can facilitate this process.
-
-```{r}
-dino <- datasauRus::datasaurus_dozen %>% filter(dataset == "dino")
-mod <- lm(y ~ ., data = select(dino, -dataset))
-
-my_vi <- autovi::auto_vi(fitted_mod = mod)
-```
-
-```{r}
-if (!file.exists(here::here("paper/data/dino_check.rds"))) {
-  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
-  my_vi$check(null_draws = 200L, boot_draws = 200L)
-  saveRDS(my_vi$check_result, here::here("paper/data/dino_check.rds"))
+    theme(legend.position = "none")
+
+  saveRDS(p, here::here("paper/data/dino_attention_map.rds"))
+
 }
-```
-
-```{r}
-my_vi$check_result <- readRDS(here::here("paper/data/dino_check.rds"))
 ```
@@
 The fitted model will be rejected since $\\hat{D} >= Q_{null}(0.95)$.'}
-p1 <- my_vi$plot_resid()
-p2 <- my_vi$summary_plot() +
-  ggtitle("", subtitle = glue::glue("RESET test p-value = {format(lmtest::resettest(mod)$p.value, digits = 3)}\n Breusch-Pagan test p-value = {format(lmtest::bptest(mod)$p.value, digits = 3)} \n Shapiro-Wilk test p-value = {format(shapiro.test(mod$residuals)$p.value, digits = 3)}")) +
+
+p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot")
+p2 <- readRDS(here::here("paper/data/dino_attention_map.rds")) + ggtitle("(B) Attention map")
+p3 <- my_vi$summary_plot() +
+  ggtitle("(C) Density plot of estimated D", subtitle = glue::glue("RESET test p-value = {format(lmtest::resettest(mod)$p.value, digits = 3)}\n Breusch-Pagan test p-value = {format(lmtest::bptest(mod)$p.value, digits = 3)} \n Shapiro-Wilk test p-value = {format(shapiro.test(mod$residuals)$p.value, digits = 3)}")) +
   theme(legend.position = "bottom", legend.box = "vertical") +

### 2024-05-13 paper/paper.tex
up to examples part

@@
 
-\subsection{Model Formulas}\label{model-formulas}
-
-There are various configurations of the computer vision model that can
-be used to assess residual plots. We discuss these configurations below
+\subsection{Model Specifications}\label{model-specifications}
+
+There are various specifications of the computer vision model that can
+be used to assess residual plots. We discuss these specifications below
 focusing on two key components of the model formula: the input and the
@@
 
-\caption{Diagram of the architecture of the optimized computer vision model.}\label{fig:unnamed-chunk-2}
+\caption{Diagram of the architecture of the optimized computer vision model.}\label{fig:cnn-diag}
 \end{figure}
@@
 high performance in image classification \citep{simonyan2014very}.
+Figure \ref{fig:cnn-diag} provides a diagram of the architecture. More
+details about the neural network layers used in this study are provided
+in the Appendix.
 
@@
 
-Based on the tuning process described in Section
-\ref{training-and-hyperparameter-tuning}, the optimized hyperparameter
-values are presented in Table \ref{tab:best-hyperparameter}. It was
-observable that a minimum of \(32\) base filters was necessary, with the
-preferable choice being \(64\) base filters for both the
-\(64 \times 64\) and \(128 \times 128\) models, mirroring the original
-VGG16 architecture. The optimized dropout rate for convolutional blocks
-hovered around \(0.4\), and incorporating batch normalization for
-convolutional blocks proved beneficial for performance.
+\subsection{Model Training}\label{model-training}
+
+To achieve an optimal deep learning model, hyperparameters like the
+learning rate often need to be fine-tuned using a tuner. In our study,
+we utilized the Bayesian optimization tuner from the \texttt{KerasTuner}
+Python library \citep{omalley2019kerastuner} for this purpose. A
+comprehensive list of hyperparameters is provided in Table
+\ref{tab:hyperparameter}.
+
+The ``number of base filters'' determines the number of filters for the
+first 2D convolutional layer. In the VGG16 architecture, the number of
+filters for the 2D convolutional layer in a block is typically twice the
+number in the previous block, except for the last block, which maintains
+the same number of convolution filters as the previous one. This
+hyperparameter aids in controlling the complexity of the computer vision
+model. Higher numbers of base filters result in more trainable
+parameters. Likewise, the ``number of units for the fully-connected
+layer'' determines the complexity of the final prediction block.
+Increasing the number of units enhances model complexity, resulting in
+more trainable parameters.
+
+The dropout rate and batch normalization are flexible hyperparameters
+that work in conjunction with other parameters to facilitate smooth
+training. A higher dropout rate is necessary when the model tends to
+overfit the training dataset by learning too much noise. Conversely, a
+lower dropout rate is preferred when the model is complex and
+challenging to converge. Batch normalization, on the other hand,
+addresses the internal covariate shift problem arising from the
+randomness in weight initialization and input data. It helps stabilize
+and accelerate the training process by normalizing the activations of
+each layer.
+
+Additionally, incorporating additional inputs such as scagnostics and
+the number of observations can potentially enhance prediction accuracy.
+Therefore, we allowed the tuner to determine whether these inputs were
+necessary for optimal model performance.
+
+The learning rate is a crucial hyperparameter, as it dictates the step
+size of the optimization algorithm. A high learning rate can help the
+model avoid local minima but risks overshooting and missing the global
+minimum. Conversely, a low learning rate smoothens the training process
+but makes the convergence time longer and increases the likelihood of
+getting trapped in local minima.
+
+Our model was trained on the M3 high-performance computing platform,
+using TensorFlow and Keras. During training, 80\% of the training data
+was utilized for actual training, while the remaining 20\% was used as
+validation data. The Bayesian optimization tuner conducted 100 trials to
+identify the best hyperparameter values based on validation root mean
+square error. The tuner then restored the best epoch of the best model
+from the trials. Additionally, we applied early stopping, terminating
+the training process if the validation root mean square error fails to
+improve for 50 epochs. The maximum allowed epochs was set at 2000,
+although no models reached this threshold.
+
+\begin{table}
+
+\caption{\label{tab:hyperparameter}Name of hyperparameters and their correspoding domain for the computer vision model.}
+\centering
+\begin{tabular}[t]{ll}
+\toprule
+Hyperparameter & Domain\\
+\midrule
+Number of base filters & \{4, 8, 16, 32, 64\}\\
+Dropout rate for convolutional blocks & {}[0.1, 0.6]\\
+Batch normalization for convolutional blocks & \{false, true\}\\
+Type of global pooling & \{max, average\}\\
+Ignore additional inputs & \{false, true\}\\
+\addlinespace
+Number of units for the fully-connected layer & \{128, 256, 512, 1024, 2048\}\\
+Batch normalization for the fully-connected layer & \{false, true\}\\
+Dropout rate for the fully-connected layer & {}[0.1, 0.6]\\
+Learning rate & {}[1e-8, 1e-1]\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Based on the tuning process described above, the optimized
+hyperparameter values are presented in Table
+\ref{tab:best-hyperparameter}. It was observable that a minimum of
+\(32\) base filters was necessary, with the preferable choice being
+\(64\) base filters for both the \(64 \times 64\) and \(128 \times 128\)
+models, mirroring the original VGG16 architecture. The optimized dropout
+rate for convolutional blocks hovered around \(0.4\), and incorporating
+batch normalization for convolutional blocks proved beneficial for
+performance.
 
@@
 
-\subsection{Model Training}\label{model-training}
-
-To achieve an optimal deep learning model, hyperparameters like the
-learning rate often need to be fine-tuned using a tuner. In our study,
-we utilized the Bayesian optimization tuner from the \texttt{KerasTuner}
-Python library \citep{omalley2019kerastuner} for this purpose. A
-comprehensive list of hyperparameters is provided in Table
-\ref{tab:hyperparameter}.
-
-The ``number of base filters'' determines the number of filters for the
-first 2D convolutional layer. In the VGG16 architecture, the number of
-filters for the 2D convolutional layer in a block is typically twice the
-number in the previous block, except for the last block, which maintains
-the same number of convolution filters as the previous one. This
-hyperparameter aids in controlling the complexity of the computer vision
-model. Higher numbers of base filters result in more trainable
-parameters. Likewise, the ``number of units for the fully-connected
-layer'' determines the complexity of the final prediction block.
-Increasing the number of units enhances model complexity, resulting in
-more trainable parameters.
-
-The dropout rate and batch normalization are flexible hyperparameters
-that work in conjunction with other parameters to facilitate smooth
-training. A higher dropout rate is necessary when the model tends to
-overfit the training dataset by learning too much noise. Conversely, a
-lower dropout rate is preferred when the model is complex and
-challenging to converge. Batch normalization, on the other hand,
-addresses the internal covariate shift problem arising from the
-randomness in weight initialization and input data. It helps stabilize
-and accelerate the training process by normalizing the activations of
-each layer.
-
-Additionally, incorporating additional inputs such as scagnostics and
-the number of observations can potentially enhance prediction accuracy.
-Therefore, we allowed the tuner to determine whether these inputs were
-necessary for optimal model performance.
-
-The learning rate is a crucial hyperparameter, as it dictates the step
-size of the optimization algorithm. A high learning rate can help the
-model avoid local minima but risks overshooting and missing the global
-minimum. Conversely, a low learning rate smoothens the training process
-but makes the convergence time longer and increases the likelihood of
-getting trapped in local minima.
-
-Our model was trained on the M3 high-performance computing platform,
-using TensorFlow and Keras. During training, 80\% of the training data
-was utilized for actual training, while the remaining 20\% was used as
-validation data. The Bayesian optimization tuner conducted 100 trials to
-identify the best hyperparameter values based on validation root mean
-square error. The tuner then restored the best epoch of the best model
-from the trials. Additionally, we applied early stopping, terminating
-the training process if the validation root mean square error fails to
-improve for 50 epochs. The maximum allowed epochs was set at 2000,
-although no models reached this threshold.
-
-\begin{table}
-
-\caption{\label{tab:hyperparameter}Name of hyperparameters and their correspoding domain for the computer vision model.}
-\centering
-\begin{tabular}[t]{ll}
-\toprule
-Hyperparameter & Domain\\
-\midrule
-Number of base filters & \{4, 8, 16, 32, 64\}\\
-Dropout rate for convolutional blocks & {}[0.1, 0.6]\\
-Batch normalization for convolutional blocks & \{false, true\}\\
-Type of global pooling & \{max, average\}\\
-Ignore additional inputs & \{false, true\}\\
-\addlinespace
-Number of units for the fully-connected layer & \{128, 256, 512, 1024, 2048\}\\
-Batch normalization for the fully-connected layer & \{false, true\}\\
-Dropout rate for the fully-connected layer & {}[0.1, 0.6]\\
-Learning rate & {}[1e-8, 1e-1]\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\subsection{Evaluation Methods}\label{evaluation-methods}
-
-This section remains incomplete as we may modify the ``Results'' section
-later on.
-
 \section{Results}\label{results}
 
-understand what the convolutional layer is actually doing (read vgg16
-doc)
-
-\subsection{Summary of model
-performance}\label{summary-of-model-performance}
+\subsection{Model Performance}\label{model-performance}
 
@@
 \(\hat{D} < 3\), while over-predictions occurred predominantly when
-\(3 < \hat{D} < 6\). To comprehend the reasons behind these deviations,
-we fitted a regression tree using the \texttt{rpart} R package
-\citep{Terry2022rpart} on \(D - \hat{D}\) for all factors used in the
-data generating process. The results of the regression tree are provided
-in the Appendix.
-
-The regression tree revealed that most issues arose from non-linearity
-problems and the presence of a second predictor in the regression model.
-When the error distribution had a very small variance, all computer
-vision models tended to under-predict the distance. Conversely, when the
-error distribution had a large variance, all models tended to
-over-predict the distance. Furthermore, for input images representing
-null plots, it was expected that the models will over-predict the
-distance, as explained in Section \ref{lineup-evaluation}.
+\(3 < \hat{D} < 6\). For input images representing null plots, it was
+expected that the models will over-predict the distance, as explained in
+Section \ref{lineup-evaluation}. However, it can not explain the
+under-prediction issue. Therefore, we analysed the relationship between
+residuals and all the factors involved in the data generating process.
+We found that most issues actually arose from non-linearity problems and
+the presence of a second predictor in the regression model as
+illustrated in Figure \ref{fig:over-under}. When the variance for the
+error distribution was small, the optimized model tended to
+under-predict the distance. Conversely, when the error distribution had
+a large variance, the model tended to over-predict the distance.
 
 Since most of the deviation stemmed from the presence of non-linearity
-violations, to further investigate this, we split the test set and
-re-evaluated the performance, as detailed in Table
+violations, to further investigate this, we split the test set based on
+violation types and re-evaluated the performance, as detailed in Table
 \ref{tab:performance-sub}. It was evident that metrics for null plots
@@
 
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/over-under-1}
+
+}
+
+\caption{Scatter plots for residuals vs $\sigma$ on test data for the $32 \times 32$ optimized model.
+The data is grouped by whether the regression has only non-linearity violation, and whether it includes a second predictor in the regression formula.
+The red lines are smoothing curves produced by fitting gnealized additive models.
+The area over the zero line in light pink indicates under-prediction, and the area under the zero line in light green indicates over-prediction.}\label{fig:over-under}
+\end{figure}
+
 \begin{table}
@@
 
-\subsection{Comparison with human visual inference and conventional
-tests}\label{comparison-with-human-visual-inference-and-conventional-tests}
-
-\subsubsection{Overview of the human subject
+\subsection{Comparison with Human Visual Inference and Conventional
+Tests}\label{comparison-with-human-visual-inference-and-conventional-tests}
+
+\subsubsection{Overview of the Human Subject
 experiment}\label{overview-of-the-human-subject-experiment}
@@
 residual plots presented in the human subject experiment conducted by
-\citet{li2023plot} will be assessed. This study has collected \(7974\)
-human responses to \(1152\) lineups. Each lineup contains one randomly
-placed true residual plot and 19 null plots. Among the \(1152\) lineups,
-\(24\) are attention check lineups in which the visual patterns are
-designed to be extremely obvious and very different from the
-corresponding to null plots, \(36\) are null lineups where all the
-lineups consist of only null plots, \(279\) are lineups with uniform
-predictor distribution evaluated by \(11\) participants, and the
-remaining \(813\) are lineups with discrete, skewed or normal predictor
-distribution evaluated by \(5\) participants. Attention check lineups
-and null lineups will not be assessed in the following analysis.
+\citet{li2023plot} will be assessed.
+
+This study has collected \(7974\) human responses to \(1152\) lineups.
+Each lineup contains one randomly placed true residual plot and 19 null
+plots. Among the \(1152\) lineups, \(24\) are attention check lineups in
+which the visual patterns are designed to be extremely obvious and very
+different from the corresponding to null plots, \(36\) are null lineups
+where all the lineups consist of only null plots, \(279\) are lineups
+with uniform predictor distribution evaluated by \(11\) participants,
+and the remaining \(813\) are lineups with discrete, skewed or normal
+predictor distribution evaluated by \(5\) participants. Attention check
+lineups and null lineups will not be assessed in the following analysis.
 
@@
 \(D\) for the null plots is zero. We have used our optimized computer
-vision model to predict the approximated distance for both the true
-residual plots and the null plots. To have a fair comparison, \(H_0\)
-will be rejected if the true residual plot has the greatest approximated
-distance among all plots in a lineup. Additionally, the appropriate
-conventional tests including the Ramsey Regression Equation
-Specification Error Test (RESET) \citep{ramsey1969tests} for
-non-linearity and the Breusch-Pagan test \citep{breusch1979simple} for
-heteroskedasticity were applied on the same data for comparison.
+vision model to estimate distance for both the true residual plots and
+the null plots. To have a fair comparison, \(H_0\) will be rejected if
+the true residual plot has the greatest approximated distance among all
+plots in a lineup. Additionally, the appropriate conventional tests
+including the Ramsey Regression Equation Specification Error Test
+(RESET) \citep{ramsey1969tests} for non-linearity and the Breusch-Pagan
+test \citep{breusch1979simple} for heteroskedasticity were applied on
+the same data for comparison.
 
@@
 akin to the best available conventional tests. However, Figure
-\ref{fig:conv-mosaic} shows that conventional tests tend to reject
-residual plots almost exclusively when the computer vision model does,
-but they also reject plots when the computer vision model does not. This
-suggests that conventional tests are more sensitive than the computer
-vision model.
+\ref{fig:conv-mosaic} shows that the computer vision model does not
+always reject when the conventional tests reject. And a small number of
+plots will be rejected by computer vision model but not by conventional
+tests. This suggests that conventional tests are more sensitive than the
+computer vision model.
+
+Figure \ref{fig:pcp} further illustrates the decisions made by visual
+tests conducted by human, computer vision models, and conventional
+tests, using a parallel coordinate plots. It can be observed that all
+three tests will agree with each other for around 50\% of the cases.
+When visual tests conducted by human do not reject, there are
+substantial amount of cases where computer vision model also not reject
+but conventional tests reject. There are much fewer cases that do not
+reject by visual tests and conventional tests, but reject by computer
+vision models. This indicates computer vision model can behave like
+visual tests conducted by human better than conventional tests.
+Moreover, there are great proportion of cases where visual tests
+conducted by human is the only test who does not reject, .
 
@@
 
-\caption{Rejection rate ($p$-value $\leq0.05$) of conventional test conditional on the computer vision model decision on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot.
-The conventional test rejects more frequently than the computer vision model, and (almost) only rejects when the computer vision model does.}\label{fig:conv-mosaic}
+\caption{Rejection rate ($p$-value $\leq0.05$) of computer vision models conditional on conventional tests on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot.}\label{fig:conv-mosaic}
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/venn-1}
-
-}
-
-\caption{Venn diagram for non-rejection decisions made by conventional tests, computer vision model, and visual tests conducted by human.}\label{fig:venn}
-\end{figure}
-
-\begin{figure}[!h]
-
 {\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/delta-1}
@@
 
-\subsection{Data examples}\label{data-examples}
+\section{Examples}\label{examples}
 
@@
 and the ``dino'' datasets from the \texttt{datasaurus} R package.
+
+The first example illustrates a scenario where both the computer vision
+model and human visual inspection successfully avoid rejecting \(H_0\)
+when \(H_0\) is true, contrary to conventional tests. This underscores
+the necessity of visually examining the residual plot.
+
+In the second example, we encounter a more pronounced violation of the
+model, resulting in rejection of \(H_0\) by all three tests. This
+highlights the practicality of the computer vision model, particularly
+for less intricate tasks.
+
+The third example presents a situation where the model deviation is
+non-typical. Here, the computer vision model and human visual inspection
+reject \(H_0\), whereas conventional tests do not. This emphasizes the
+benefits of visual inspection and the unique advantage of the computer
+vision model, which, like humans, makes decisions based on visual
+discoveries.
+
 Additionally, we will outline the workflow for employing this model with
@@
 
-\subsubsection{Left-triangle}\label{left-triangle}
+\subsection{Left-triangle}\label{left-triangle}
 
@@
 fitted model to be misspecified as the majority of bootstrapped fitted
-models will not be rejected.
-
-Thus, for this particular fitted model, both the visual test and the
-computer vision model will not reject \(H_0\). However, the
-Breusch-Pagan test will reject \(H_0\) because it can not effectively
-utilize information from null plots.
+models will not be rejected. Thus, for this particular fitted model,
+both the visual test and the computer vision model will not reject
+\(H_0\). However, the Breusch-Pagan test will reject \(H_0\) because it
+can not effectively utilize information from null plots.
+
+The attention map at plot B of Figure \ref{fig:false-check} suggests
+that the estimation is highly influenced by the top-right and
+bottom-right part of the residual plot, as it forms two vertices of the
+triangular shape. A principle component analysis is also performed on
+the output of the global pooling layer of the computer vision model. As
+mentioned in \citet{simonyan2014very}, a computer vision model built
+upon the convolutional blocks can be viewed as a feature extractor. For
+the \(32 \times 32\) model, there are 256 features outputted from the
+global pooling layer, which can be further used for different visual
+tasks not limited to distance prediction. To see if these features can
+be effectively used for distinguishing null plots and true residual
+plot, we linearly project them into the PC1 vs PC2 space as shown in
+Figure \ref{fig:false-check}. It can be observed that because the
+bootstrapped plots are mostly similar to the null plots, the points
+drawn in different colours are mixed together. The true residual plot is
+also covered by both the cluster of null plots and cluster of
+bootstrapped plots. This accurately reflects our understanding of Figure
+\ref{fig:false-lineup}.
 
@@
 \caption{A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
-On the left, the residual plot exhibiting "left-triangle" shape is displayed.
+(A) The true residual plot exhibiting "left-triangle" shape.
+(B) The attention map produced by computing the gradient of the output with respect to the greyscale input.
+(C) The density plot of estimated distnace for null plots and bootstrapped plots.
 The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots.
-The fitted model will not be rejected since $\hat{D} < Q_{null}(0.95)$.}\label{fig:false-check}
+The fitted model will not be rejected since $\hat{D} < Q_{null}(0.95)$.
+(D) PC2 vs PC1 plot for features extracted from the global pooling layer of the computer vision model.
+}\label{fig:false-check}
 \end{figure}
@@
 
-\subsubsection{Boston housing}\label{boston-housing}
+\subsection{Boston housing}\label{boston-housing}
 
@@
 
-\subsubsection{Datasaurus}\label{datasaurus}
+\subsection{Datasaurus}\label{datasaurus}
 

### 2024-05-27 paper/appendix.Rmd
wait for review

@@
   Appendix: Automated assessment of residual plots with computer vision models
-type: Draft paper (Incomplete)
 author:
@@
 bibliography: ref.bib
-abstract: |
-  TBD.
-keywords: |
-  TBD
 header-includes: |
@@
 
-# Additional details about data generation
-
-## Computation of scagnostics
-
-In Section \ref{introduction}, we mentioned that scagnostics consist of a set of manually designed visual feature extraction functions.
-While our computer vision model will learn its own feature extraction function during training, leveraging additional information from scagnostics can enhance the model's predictive accuracy.
-
-For each generated residual plot, we computed four scagnostics – "Monotonic," "Sparse," "Splines," and "Striped" – using the `cassowaryr` R package [@mason2022cassowaryr].
-These computed measures, along with the number of observations from the fitted model, were provided as the second input for the computer vision model.
-Although other scagnostics are informative, they are currently unavailable due to a fatal bug in the compiled C program of the `interp` R package [@Albrecht2023interp], which may unpredictably crash the process.
-For reproducibility, we excluded these scagnostics from the training data.
-
 # Neural Network Layers Used in the Study
 
-This study employs seven types of neural network layers, all of which are standard components frequently found in modern deep learning models.
-These layers are well-documented in textbooks like @goodfellow2016deep, which offer thorough explanations and mathematical insights.
+This study used seven types of neural network layers, all of which are standard components frequently found in modern deep learning models.
+These layers are well-documented in textbooks like @goodfellow2016deep and @chollet2021deep, which offer thorough explanations and mathematical insights.
 In this section, we will offer a concise overview of these layers, drawing primarily from the insights provided in @goodfellow2016deep.
@@
 The Dense layer, also known as the fully-connected layer, is the fundamental unit in neural networks.
-It conducts a matrix multiplication operation between the input matrix $\boldsymbol{X}$ and a weight matrix $\boldsymbol{W}$ to generate the output matrix $\boldsymbol{O}$, which can be written as
+It conducts a matrix multiplication operation between the input matrix $\boldsymbol{I}$ and a weight matrix $\boldsymbol{W}$ to generate the output matrix $\boldsymbol{O}$, which can be written as
 
-$$\boldsymbol{O} = \boldsymbol{X}\boldsymbol{W} + b,$$
+$$\boldsymbol{O} = \boldsymbol{I}\boldsymbol{W} + b,$$
 
@@
 
-$$\boldsymbol{O}(i,j) = max\{0, \boldsymbol{X}(i,j)\},$$
+$$\boldsymbol{O}(i,j) = max\{0, \boldsymbol{I}(i,j)\},$$
 
-where $\boldsymbol{O}(i,j)$ is the $i$th row and $j$th column entry of matrix $\boldsymbol{O}$, and $\boldsymbol{X}(i,j)$ is the $i$th row and $j$th column entry of matrix $\boldsymbol{X}$.
+where $\boldsymbol{O}(i,j)$ is the $i$th row and $j$th column entry of matrix $\boldsymbol{O}$, and $\boldsymbol{I}(i,j)$ is the $i$th row and $j$th column entry of matrix $\boldsymbol{I}$.
 
@@
 In Dense layers, matrix multiplication leads to each output unit interacting with every input unit, whereas convolutional layers operate differently with sparse interactions.
-Here, an output unit in a convolutional layer is connected solely to a subset of input units, and the weight is shared across all input units.
+An output unit in a convolutional layer is connected solely to a subset of input units, and the weight is shared across all input units.
 Achieving this involves using a kernel, typically a small square matrix, to conduct matrix multiplication across all input units.
@@
 
-$$O(i, j) = \sum_m\sum_nI(i - m, j - n)K(m, n),$$
+$$\boldsymbol{O}(i, j) = \sum_m\sum_n\boldsymbol{I}(i - m, j - n)K(m, n),$$
 
@@
 The output matrix of a pooling layer is approximately $s$ times smaller than the input matrix, where $s$ represents the length of the rectangular area.
-This can be formulated as:
+A max pooling layer can be formulated as:
 
-$$O(i, j) = \underset{m,n}{\max} I(si + m,sj+n).$$
+$$\boldsymbol{O}(i, j) = \underset{m,n}{\max} \boldsymbol{I}(si + m,sj+n).$$
 
@@
 This layer acts as a crucial link between the convolutional structure and the subsequent dense layers in a neural network architecture.
-When convolutional layers utilize multiple kernels, the output becomes a three-dimensional tensor with numerous channels.
+When convolutional layers uses multiple kernels, the output becomes a three-dimensional tensor with numerous channels.
 In this scenario, the global pooling layer treats each channel individually, much like distinct features in a conventional classifier.
 This approach facilitates the extraction of essential features from complex data representations, enhancing the network's ability to learn meaningful patterns.
+A global max pooling layer can be formulated as
+
+$$O(i, j) = \underset{m,n,k}{\max} I(si + m,sj+n,k),$$
+where $k$ is the kernel index.
 
@@
 
+Batch normalization is a method of adaptive reparametrization.
+One of the issues it adjusts is the simultaneous update of parameters in different layers, especially for network with a large number layers.
+At training time, the batch normalization layer normalizes the input matrix $I$ using the formula
+
+$$\boldsymbol{O} = \frac{\boldsymbol{I} - \boldsymbol{\mu}_I}{\boldsymbol{\sigma}_I},$$
+
+where $\boldsymbol{\mu}_I$ and $\boldsymbol{\sigma}_I$ are the mean and the standard deviation of each unit respectively.
+
+It reparametrizes the model to make some units always be standardized by definition, such that the model training is stabilized.
+At inference time, $\boldsymbol{\mu}_I$ and $\boldsymbol{\sigma}_I$ are usually replaced with the running mean and running average obtained during training.
+
 ## Dropout Layer
 
-<!-- ## visualize the convolutional blocks -->
+Dropout is a computationally inexpensive way to apply regularization on neural network.
+For each input unit, it randomly sets to be zero during training, effectively training a large number of subnetworks simultaneously, but these subnetworks share weights and each will only be trained for a single steps in a large network.
+It is essentially a different implementation of the bagging algorithm.
+Mathematically, it is formulated as
+
+$$\boldsymbol{O}(i,j) = \boldsymbol{D}(i,j)\boldsymbol{I}(i,j),$$
+
+where $\boldsymbol{D}(i,j) \sim B(1, p)$ and $p$ is a hyperparameter that can be tuned.

### 2024-05-27 paper/appendix.tex
wait for review

@@
 
-\articletype{Draft paper (Incomplete)}
+\articletype{}
 
@@
 
-\begin{abstract}
-TBD.
-\end{abstract}
-
-\begin{keywords}
-TBD
-\end{keywords}
-
-\section{Additional details about data
-generation}\label{additional-details-about-data-generation}
-
-\subsection{Computation of
-scagnostics}\label{computation-of-scagnostics}
-
-In Section \ref{introduction}, we mentioned that scagnostics consist of
-a set of manually designed visual feature extraction functions. While
-our computer vision model will learn its own feature extraction function
-during training, leveraging additional information from scagnostics can
-enhance the model's predictive accuracy.
-
-For each generated residual plot, we computed four scagnostics --
-``Monotonic,'' ``Sparse,'' ``Splines,'' and ``Striped'' -- using the
-\texttt{cassowaryr} R package \citep{mason2022cassowaryr}. These
-computed measures, along with the number of observations from the fitted
-model, were provided as the second input for the computer vision model.
-Although other scagnostics are informative, they are currently
-unavailable due to a fatal bug in the compiled C program of the
-\texttt{interp} R package \citep{Albrecht2023interp}, which may
-unpredictably crash the process. For reproducibility, we excluded these
-scagnostics from the training data.
-
 \section{Neural Network Layers Used in the
@@
 
-This study employs seven types of neural network layers, all of which
-are standard components frequently found in modern deep learning models.
+This study used seven types of neural network layers, all of which are
+standard components frequently found in modern deep learning models.
 These layers are well-documented in textbooks like
-\citet{goodfellow2016deep}, which offer thorough explanations and
-mathematical insights. In this section, we will offer a concise overview
-of these layers, drawing primarily from the insights provided in
-\citet{goodfellow2016deep}.
+\citet{goodfellow2016deep} and \citet{chollet2021deep}, which offer
+thorough explanations and mathematical insights. In this section, we
+will offer a concise overview of these layers, drawing primarily from
+the insights provided in \citet{goodfellow2016deep}.
 
@@
 fundamental unit in neural networks. It conducts a matrix multiplication
-operation between the input matrix \(\boldsymbol{X}\) and a weight
+operation between the input matrix \(\boldsymbol{I}\) and a weight
 matrix \(\boldsymbol{W}\) to generate the output matrix
@@
 
-\[\boldsymbol{O} = \boldsymbol{X}\boldsymbol{W} + b,\]
+\[\boldsymbol{O} = \boldsymbol{I}\boldsymbol{W} + b,\]
 
@@
 
-\[\boldsymbol{O}(i,j) = max\{0, \boldsymbol{X}(i,j)\},\]
+\[\boldsymbol{O}(i,j) = max\{0, \boldsymbol{I}(i,j)\},\]
 
 where \(\boldsymbol{O}(i,j)\) is the \(i\)th row and \(j\)th column
-entry of matrix \(\boldsymbol{O}\), and \(\boldsymbol{X}(i,j)\) is the
-\(i\)th row and \(j\)th column entry of matrix \(\boldsymbol{X}\).
+entry of matrix \(\boldsymbol{O}\), and \(\boldsymbol{I}(i,j)\) is the
+\(i\)th row and \(j\)th column entry of matrix \(\boldsymbol{I}\).
 
@@
 interacting with every input unit, whereas convolutional layers operate
-differently with sparse interactions. Here, an output unit in a
-convolutional layer is connected solely to a subset of input units, and
-the weight is shared across all input units. Achieving this involves
-using a kernel, typically a small square matrix, to conduct matrix
-multiplication across all input units. Precisely, this concept can be
-formulated as:
+differently with sparse interactions. An output unit in a convolutional
+layer is connected solely to a subset of input units, and the weight is
+shared across all input units. Achieving this involves using a kernel,
+typically a small square matrix, to conduct matrix multiplication across
+all input units. Precisely, this concept can be formulated as:
 
-\[O(i, j) = \sum_m\sum_nI(i - m, j - n)K(m, n),\]
+\[\boldsymbol{O}(i, j) = \sum_m\sum_n\boldsymbol{I}(i - m, j - n)K(m, n),\]
 
@@
 \(s\) times smaller than the input matrix, where \(s\) represents the
-length of the rectangular area. This can be formulated as:
+length of the rectangular area. A max pooling layer can be formulated
+as:
 
-\[O(i, j) = \underset{m,n}{\max} I(si + m,sj+n).\]
+\[\boldsymbol{O}(i, j) = \underset{m,n}{\max} \boldsymbol{I}(si + m,sj+n).\]
 
@@
 structure and the subsequent dense layers in a neural network
-architecture. When convolutional layers utilize multiple kernels, the
+architecture. When convolutional layers uses multiple kernels, the
 output becomes a three-dimensional tensor with numerous channels. In
@@
 complex data representations, enhancing the network's ability to learn
-meaningful patterns.
+meaningful patterns. A global max pooling layer can be formulated as
+
+\[O(i, j) = \underset{m,n,k}{\max} I(si + m,sj+n,k),\] where \(k\) is
+the kernel index.
 
@@
 
+Batch normalization is a method of adaptive reparametrization. One of
+the issues it adjusts is the simultaneous update of parameters in
+different layers, especially for network with a large number layers. At
+training time, the batch normalization layer normalizes the input matrix
+\(I\) using the formula
+
+\[\boldsymbol{O} = \frac{\boldsymbol{I} - \boldsymbol{\mu}_I}{\boldsymbol{\sigma}_I},\]
+
+where \(\boldsymbol{\mu}_I\) and \(\boldsymbol{\sigma}_I\) are the mean
+and the standard deviation of each unit respectively.
+
+It reparametrizes the model to make some units always be standardized by
+definition, such that the model training is stabilized. At inference
+time, \(\boldsymbol{\mu}_I\) and \(\boldsymbol{\sigma}_I\) are usually
+replaced with the running mean and running average obtained during
+training.
+
 \subsection{Dropout Layer}\label{dropout-layer}
+
+Dropout is a computationally inexpensive way to apply regularization on
+neural network. For each input unit, it randomly sets to be zero during
+training, effectively training a large number of subnetworks
+simultaneously, but these subnetworks share weights and each will only
+be trained for a single steps in a large network. It is essentially a
+different implementation of the bagging algorithm. Mathematically, it is
+formulated as
+
+\[\boldsymbol{O}(i,j) = \boldsymbol{D}(i,j)\boldsymbol{I}(i,j),\]
+
+where \(\boldsymbol{D}(i,j) \sim B(1, p)\) and \(p\) is a hyperparameter
+that can be tuned.
 

### 2024-05-27 paper/cnn.tex
wait for review

@@
 \def\PoolColor{rgb:red,1;black,0.3}
+\def\ConColor{rgb:black,1;white,1}
 \def\FcColor{rgb:blue,5;red,2.5;white,5}
@@
 \pic[shift={(3,0,0)}] at (p5-east) {Box={name=fc6,caption=Concatenation layer,%
-        xlabel={{"1",""}},zlabel=256,fill=\FcColor,%
+        xlabel={{"1",""}},zlabel=256,fill=\ConColor,%
         height=3,width=3,depth=100}};

### 2024-05-27 paper/paper.Rmd
wait for review

@@
   Automated assessment of residual plots with computer vision models
-type: Draft paper (Incomplete)
 author:
@@
 abstract: |
-  TBD.
+  Plotting residuals is a standard practice in linear regression diagnostics, essential for identifying deviations from model assumptions such as linearity, homoscedasticity, and normality.
+Visual inference provides an inferential framework to assess whether residual plots contain patterns inconsistent with model assumptions, typically using a lineup protocol.
+However, the lineup protocol's reliance on human judgment limits its scalability.
+This study addresses this limitation by automating the interpretation of residual plots using computer vision models.
+We develop a distance measure based on Kullback-Leibler divergence to quantify the disparity between the residual distribution of a fitted classical normal linear regression model and the reference distribution.
+We propose a computer vision model to estimate this distance from residual plots, facilitating formal statistical testing and bootstrapping techniques to assess model specification.
+Our computer vision model shows strong performance, though it performs slightly less effectively on non-linearity visual patterns.
+The statistical tests based on the estimated distance exhibit lower sensitivity than conventional tests but higher sensitivity than human visual tests.
+Examples demonstrate the method's effectiveness across different scenarios, highlighting its value in automating the diagnostic process and supplementing traditional methods.
 keywords: |
@@
 
-In this paper, we develop computer vision models and integrate them into the residual plots diagnostics workflow, filling the gap of.... The paper is structured as follows: ...
-
-```{r false-finding, fig.width = 8/5*4, fig.height = 6/4*4, fig.pos="!h", fig.cap="An example residual vs fitted values plot (red line indicates 0).
+In this paper, we develop computer vision models and integrate them into the residual plots diagnostics workflow, addressing the need for automated visual inference.
+The paper is structured as follows.
+Section \ref{model-specifications} discusses various specifications of the computer vision models.
+Section \ref{distance-between-residual-plots} defines the distance measure used to measure model violations.
+Section \ref{distance-estimation} explains how the computer vision models estimate this distance measure.
+Section \ref{statistical-testing} covers the statistical testing based on the estimated distance.
+Sections \ref{data-generation}, \ref{model-architecture}, and \ref{model-training} detail the data preparation, model architecture, and training process, respectively.
+The results are presented in Section \ref{results}.
+Finally, we conclude with a discussion of our findings and propose ideas for future research directions.
+
+```{r false-finding, fig.width = 8, fig.height = 8*4/5, fig.pos="!h", fig.cap="An example residual vs fitted values plot (red line indicates 0).
 The vertical spread of the data points varies with the fitted values.
@@
 
-# Methodology
-
-## Model Specifications
+# Model Specifications
 
@@
 
-### Input Formats
+## Input Formats
 
@@
 
-### Output Formats
+## Output Formats
 
@@
 
-## Distance Between Residual Plots
+# Distance Between Residual Plots
 
@@
 The challenge lies in the fact that metrics tailored for image comparison may not be suitable for evaluating data plots, where only essential plot elements require assessment [@chowdhury2018measuring].
-Furthermore, scagnostics mentioned in Section \ref{introduction} could be used to construct distance metrics for residual plots comparison, but the functional form still needs to be carefully refined to accurately reflect the extent of the violations.
-
-### Residual Distribution
+Furthermore, scagnostics mentioned in Section \ref{introduction} could be used to construct distance measures for residual plots comparison, but the functional form still needs to be carefully refined to accurately reflect the extent of the violations.
+
+## Residual Distribution
 
@@
 For example, if the data generating process contains variables correlated with any column of $\boldsymbol{X}$ but missing from $\boldsymbol{X}$, causing an omitted variable problem, $P$ will be different from $Q$ because the residual operator obtained from the fitted regression model will not be the same as $\boldsymbol{R}$.
-Besides, if the $\boldsymbol{\varepsilon}$ follows a non-normal distribution such as a multivariate log-normal distribution, $P$ will usually be skewed and has a long tail.
-
-### Distance of $P$ from $Q$
+Besides, if the $\boldsymbol{\varepsilon}$ follows a non-normal distribution such as a multivariate lognormal distribution, $P$ will usually be skewed and has a long tail.
+
+## Distance of $P$ from $Q$
 
@@
 
-### Non-normal $P$
+## Non-normal $P$
 
@@
 
-## Distance Estimation
+# Distance Estimation
 
 In the previous sections, we have defined a distance measure given in equation \ref{eq:kl-0} for quantifying the difference between the true residual distribution $P$ and an ideal reference distribution $Q$.
-You may have noticed that this distance measure can only be computed when the data generating process is known.
+It can be noticed that this distance measure can only be computed when the data generating process is known.
 In reality, we often have no knowledge about the data generating process, otherwise we do not need to fit a linear regression model in the first place.
@@
 
-## Model Violations Index (MVI) {#sec:model-violations-index}
+# Model Violations Index (MVI) {#sec:model-violations-index}
 
@@
 
-## Statistical testing
-
-### Lineup Evaluation
+# Statistical testing
+
+## Lineup Evaluation
 
@@
 Specifically, if a lineup comprises 20 plots, the null hypothesis $H_0$ will be rejected if $\hat{D}$ exceeds the maximum estimated distance among the $m - 1$ null plots, denoted as $\max\limits_{1 \leq i \leq m-1} {\hat{D}_{null}^{(i)}}$, where $\hat{D}_{null}^{(i)}$ represents the estimated distance for the $i$-th null plot.
-This approach is equivalent to the typical lineup protocol requiring a 95% significance level, where $H_0$ is rejected if the data plot is identified as the most distinct plot by the sole participant.
+This approach is equivalent to the typical lineup protocol requiring a 95% significance level, where $H_0$ is rejected if the data plot is identified as the most distinct plot by the sole observer.
 The estimated distance serves as a metric to quantify the difference between the data plot and the null plots, as intended.
@@
 For lineups consisting of more than 20 plots, the 95% significance level can be maintained if the number of plots is a multiple of 20.
-Specifically, for lineups comprising $20t$ plots, where $l$ is a positive integer, we reject $H_0$ if $\hat{D}$ exceeds 95% ${\hat{D}_{null}^{(i)}}$ for $i = 1, \ldots, 20t-1$.
-The $p$-value in this case is given by $\frac{1}{20t} + \frac{1}{20t}\sum{i=1}^{20t-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)$, where $I(\cdot)$ is the indicator function.
-
-Moreover, if the number of plots in a lineup, denoted by $m$, is sufficiently large, the empirical distribution of ${\hat{D}_{null}^{(i)}}$ can be viewed as an approximation of the null distribution of the approximated distance.
+Specifically, for lineups comprising $20t$ plots, where $t$ is a positive integer, we reject $H_0$ if $\hat{D}$ exceeds 95% ${\hat{D}_{null}^{(i)}}$ for $i = 1, \ldots, 20t-1$.
+The $p$-value in this case is given by $\frac{1}{20t} + \frac{1}{20t}\sum_{i=1}^{20t-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)$, where $I(\cdot)$ is the indicator function.
+
+Moreover, if the number of plots in a lineup, denoted by $m$, is sufficiently large, the empirical distribution of ${\hat{D}_{null}^{(i)}}$ can be viewed as an approximation of the null distribution of the estimated distance.
 Consequently, quantiles of the null distribution can be estimated using the sample quantiles available in statistical software such as R, and these quantiles can be utilized for decision-making purposes.
@@
 
-### Bootstrapping
+## Bootstrapping
 
@@
 
-## Data Genearation
-
-### Simulation Scheme
+# Data Generation
+
+## Simulation Scheme
 
 While observational data is frequently employed in training models for real-world applications, the data generating process of observational data often remains unknown, making computation for our target variable $D$ unattainable.
-Consequently, the computer vision models developed in this study were trained using synthetic data.
+Consequently, the computer vision models developed in this study were trained using synthetic data, including 80000 training images and 8000 test images.
 This approach provided us with precise label annotations.
@@
 
-\noindent where $\boldsymbol{y}$, $\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{z}$, $\boldsymbol{w}$, $\boldsymbol{k}$ and $\boldsymbol{\varepsilon}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$ are two independent predictors, $\text{He}_j(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau], the $\sqrt{(.)}$ and $(.)^2$ operators are element-wise operators, $\odot$ is the Hadamard product, and $g(., k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
+\noindent where $\boldsymbol{y}$, $\boldsymbol{x}_1$, $\boldsymbol{x}_2$, $\boldsymbol{z}$, $\boldsymbol{w}$, $\boldsymbol{k}$ and $\boldsymbol{\varepsilon}$ are vectors of size $n$, $\boldsymbol{1}_n$ is a vector of ones of size $n$, $\boldsymbol{x}_1$ and $\boldsymbol{x}_2$ are two independent predictors, $\text{He}_j(.)$ is the $j$th-order probabilist's Hermite polynomials [@hermite1864nouveau], the $\sqrt{(.)}$ and $(.)^2$ operators are element-wise operators, $\odot$ is the Hadamard product, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as
 
@@
 tibble(Factor = c("j", "a", "b", "$\\beta_1$", "$\\beta_2$", " $\\text{dist}_\\varepsilon$", "$\\text{dist}_{x1}$", "$\\text{dist}_{x2}$", "$\\sigma_{\\varepsilon}$", "$\\sigma_{X1}$", "$\\sigma_{X2}$", "n"),
-       Domain = c("\\{2, 3, ..., 18\\}", "[-1, 1]", "[0, 100]", "{0, 1}", "{0, 1}", "\\{discrete, uniform, normal, lognormal\\}", "\\{discrete, uniform, normal, lognormal\\}", "\\{discrete, uniform, normal, lognormal\\}", "[0.0625, 9]", "[0.3, 0.6]", "[0.3, 0.6]", "[50, 500]")) %>%
+       Domain = c("\\{2, 3, ..., 18\\}", "[-1, 1]", "[0, 100]", "\\{0, 1\\}", "\\{0, 1\\}", "\\{discrete, uniform, normal, lognormal\\}", "\\{discrete, uniform, normal, lognormal\\}", "\\{discrete, uniform, normal, lognormal\\}", "[0.0625, 9]", "[0.3, 0.6]", "[0.3, 0.6]", "[50, 500]")) %>%
   kableExtra::kable("latex", 
@@
 Non-normality violations arise from specifying a non-normal distribution for $\boldsymbol{\varepsilon}$.
-In the synthetic data simulation, four distinct error distributions are considered, including discrete, uniform, normal, and log-normal distributions, as presented in Figure \ref{fig:different-e}.
+In the synthetic data simulation, four distinct error distributions are considered, including discrete, uniform, normal, and lognormal distributions, as presented in Figure \ref{fig:different-e}.
 Each distribution imparts unique characteristics to the residuals.
-The discrete error distribution introduces discreteness in residuals, while the log-normal distribution typically yields outliers.
+The discrete error distribution introduces discreteness in residuals, while the lognormal distribution typically yields outliers.
 Uniform error distribution may result in residuals filling the entire space of the residual plot.
@@
 
-### Scagnostics
+## Scagnostics
 
@@
 
-For each generated residual plot, we computed four scagnostics – "Monotonic," "Sparse," "Splines," and "Striped" – using the `cassowaryr` R package [@mason2022cassowaryr].
+For each generated residual plot, we computed four scagnostics – "Monotonic", "Sparse", "Splines", and "Striped" – using the `cassowaryr` R package [@mason2022cassowaryr].
 These computed measures, along with the number of observations from the fitted model, were provided as the second input for the computer vision model.
@@
 
-### Balanced Dataset
+## Balanced Dataset
 
@@
 
-## Model Architecture
-
-```{r cnn-diag, out.width = "100%", fig.cap = "Diagram of the architecture of the optimized computer vision model."}
+# Model Architecture
+
+```{r cnn-diag, out.width = "100%", fig.cap = "Diagram of the architecture of the optimized computer vision model.
+Numbers at the bottom of each box show the shape of the output of each layer.
+The band of each box drawn in a darker colour indicates the use of the rectified linear unit activation function.
+Yellow boxes are 2D convolutional layers, orange boxes are pooling layers, the grey box is the concatenation layer, and the purple boxes are dense layers."}
 knitr::include_graphics("cnn.pdf")
@@
 
-## Model Training
-
-To achieve an optimal deep learning model, hyperparameters like the learning rate often need to be fine-tuned using a tuner.
+# Model Training
+
+To achieve a near-optimal deep learning model, hyperparameters like the learning rate often need to be fine-tuned using a tuner.
 In our study, we utilized the Bayesian optimization tuner from the `KerasTuner` Python library [@omalley2019kerastuner] for this purpose.
@@
 
-Our model was trained on the M3 high-performance computing platform, using TensorFlow and Keras.
+Our model was trained on the MASSIVE M3 high-performance computing platform [@goscinski2014multi], using TensorFlow [@abadi2016tensorflow] and Keras [@chollet2015keras].
 During training, 80% of the training data was utilized for actual training, while the remaining 20% was used as validation data.
@@
 ```
-
-<!-- ## Evaluation Methods -->
-
-<!-- This section remains incomplete as we may modify the "Results" section later on. -->
 
@@
 ```{r model-performance, fig.pos="!h", fig.cap = "Hexagonal heatmap for residuals vs predicted distance on training and test data for three optimized models with different input sizes.
-The red lines are smoothing curves produced by fitting gnealized additive models.
-The area over the zero line in light pink indicates under-prediction, and the area under the zero line in light green indicates over-prediction.", dev = 'png', dpi = 300}
+The brown lines are smoothing curves produced by fitting gnealized additive models.
+The area over the zero line in light yellow indicates under-prediction, and the area under the zero line in light green indicates over-prediction.", dev = 'png', dpi = 300}
 model_pred <- train_pred %>% 
@@
                 xmax = max(model_pred$vss),
-            fill = "#66c2a5",
+            fill = "#40B0A6",
             alpha = 0.3) +
@@
                 xmax = max(model_pred$vss),
-            fill = "#fc8d62",
+            fill = "#E1BE6A",
             alpha = 0.3) +
   geom_point(data = NULL, aes(5, 0, col = "under-prediction"), shape = 15) +
-  geom_hex(aes(vss, effect_size - vss), bins = 25) +
-  geom_smooth(aes(vss, effect_size - vss), se = FALSE, col = "#e34a33") +
+  geom_hex(aes(vss, effect_size - vss), bins = 20) +
+  geom_smooth(aes(vss, effect_size - vss), se = FALSE, col = "#994F00") +
   facet_grid(type ~ res) +
@@
   theme_light() +
-  scale_fill_continuous(limits = c(1, 200), trans = "log10", low = "#56B1F7", high = "#132B43") +
-  scale_color_manual(values = c("under-prediction" = scales::alpha("#fc8d62", 0.3),
-                                "over-prediction" = scales::alpha("#66c2a5", 0.3)),
+  scale_fill_continuous(limits = c(1, 6000), trans = "log10", low = "#56B1F7", high = "#132B43") +
+  scale_color_manual(values = c("under-prediction" = scales::alpha("#E1BE6A", 0.3),
+                                "over-prediction" = scales::alpha("#40B0A6", 0.3)),
                      breaks = c("under-prediction", "over-prediction")) +
@@
 The data is grouped by whether the regression has only non-linearity violation, and whether it includes a second predictor in the regression formula.
-The red lines are smoothing curves produced by fitting gnealized additive models.
-The area over the zero line in light pink indicates under-prediction, and the area under the zero line in light green indicates over-prediction.", dev = 'png', dpi = 300}
+The brown lines are smoothing curves produced by fitting gnealized additive models.
+The area over the zero line in light yellow indicates under-prediction, and the area under the zero line in light green indicates over-prediction.", dev = 'png', dpi = 300}
 model_pred %>%
@@
                 xmax = max(model_pred$e_sigma),
-            fill = "#66c2a5",
+            fill = "#40B0A6",
             alpha = 0.3) +
@@
                 xmax = max(model_pred$e_sigma),
-            fill = "#fc8d62",
+            fill = "#E1BE6A",
             alpha = 0.3) +
   geom_point(aes(e_sigma, effect_size - vss), alpha = 0.2) +
-  geom_smooth(aes(e_sigma, effect_size - vss), se = FALSE, col = "#e34a33") +
+  geom_smooth(aes(e_sigma, effect_size - vss), se = FALSE, col = "#994F00") +
   facet_grid(`Has second predictor` ~ `Only has non-linearity violation`,
@@
 
-<!-- For null plots, a histogram of the predictions is provided in Figure \ref{fig:hist-null-human}.
-In most cases, $\hat{D}$ is predicted to be around $1$.
-As $\hat{D}$ increases, the number of predictions decreases.
-The 90% and 95% sample quantiles are approximately at $2$, and the 99% sample quantile is around $2.5$.
-These sample quantiles can serve as global critical values for determining whether a residual plot needs to be rejected.
-We have also computed the 95% sample quantiles by assessing 200 null plots for each fitted model used in the experiment.
-The difference between decisions based on these two types of critical values is minor, with decisions aligning `r experiment %>% filter(!attention_check) %>% filter(!null_lineup) %>% summarise(q = mean(model_reject == (2.03 <= vss)) * 100) %>% pull(q) %>% format(digits = 3)`% of the time.
-Therefore, we will be using the 95% sample quantiles obtained from each fitted model in the following analysis.
--->
-
 Table \ref{tab:human-conv-table} provides a summary of the agreement between decisions made by the computer vision model and conventional tests.
@@
 
-Figure \ref{fig:delta} displays the scatter plot of the weighted detection rate vs the $\delta$-difference.
 In the experiment conducted in @li2023plot, participants were allowed to make multiple selections for a lineup.
@@
 
-Figure \ref{fig:delta} indicates that the weighted detection rate increases as the $\delta$-difference increases, particularly when the $\delta$-difference is greater than zero.
+Figure \ref{fig:delta} displays the scatter plot of the weighted detection rate vs the $\delta$-difference.
+It indicates that the weighted detection rate increases as the $\delta$-difference increases, particularly when the $\delta$-difference is greater than zero.
 A negative $\delta$-difference suggests that there is at least one null plot in the lineup with a stronger visual signal than the true residual plot.
 In some instances, the weighted detection rate is close to one, yet the $\delta$-difference is negative.
-This discrepancy implies that the distance measure, or the approximated distance, may not perfectly reflect actual human behaviour.
+This discrepancy implies that the distance measure, or the estimated distance, may not perfectly reflect actual human behaviour.
 
@@
   select(unique_lineup_id, rank, model_reject_20))
-
-# lineup_vss %>%
-#   filter(null) %>%
-#   left_join(vi_survey %>%
-#               group_by(unique_lineup_id) %>% 
-#               summarise(type = first(type), 
-#                         attention_check = first(attention_check),
-#                         null_lineup = first(null_lineup))) %>%
-#   filter(!attention_check) %>%
-#   filter(!null_lineup) %>%
-#   mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-#   ggplot() +
-#   geom_histogram(aes(vss)) +
-#   geom_vline(aes(xintercept = quantile(vss, 0.9), col = "Q(0.90)")) +
-#   geom_vline(aes(xintercept = quantile(vss, 0.95), col = "Q(0.95)")) +
-#   geom_vline(aes(xintercept = quantile(vss, 0.99), col = "Q(0.99)")) +
-#   xlab(expression(hat(D))) +
-#   theme_light() +
-#   facet_wrap(~type) +
-#   labs(col = "")
 ```
@@
 vi_lineup <- get_vi_lineup()
-```
-
-```{r eval = FALSE}
-library(visage)
-conv_experiment <- vi_lineup %>%
-  map_df(function(.x) {
-    reset_p_value <- c()
-    bp_p_value <- c()
-    sw_p_value <- c()
-
-    for (i in 1:20) {
-      this_dat <- .x$data %>% filter(k == i)
-      reset_p_value <- c(reset_p_value, POLY_MODEL$test(this_dat, test = "RESET")$p_value)
-      bp_p_value <- c(bp_p_value, HETER_MODEL$test(this_dat)$p_value)
-      sw_p_value <- c(sw_p_value, shapiro.test(this_dat$.resid)$p.value)
-    }
-    return(tibble(unique_lineup_id = rep(.x$metadata$name, 20),
-                  ans = .x$metadata$answer,
-                  pos = 1:20,
-                  reset_p_value = reset_p_value,
-                  reset_rank = rank(reset_p_value, ties.method = "first"),
-                  bp_p_value = bp_p_value,
-                  bp_rank = rank(bp_p_value, ties.method = "first"),
-                  sw_p_value = sw_p_value,
-                  sw_rank = rank(sw_p_value, ties.method = "first")))
-  })
-
-conv_experiment %>%
-  filter(ans == pos) %>%
-  mutate(reset_reject_20 = reset_rank == 1,
-         bp_reject_20 = bp_rank == 1,
-         sw_reject_20 = sw_rank == 1) %>%
-  select(-ans, -pos, -c(reset_p_value:sw_rank)) %>%
-  right_join(experiment) %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  pivot_longer(c(reject, model_reject_20, reset_reject_20, bp_reject_20, conventional_reject, model_reject)) %>%
-  select(type, actual_ss, name, value) %>%
-  mutate(offset0 = log(0.05/0.95)) %>%
-  nest(dat = c(offset0, actual_ss, value)) %>%
-  mutate(mod = map(dat, 
-                   ~glm(value ~ actual_ss - 1, 
-                        family = binomial(), 
-                        data = .x,
-                        offset = offset0))) %>%
-  mutate(power = map(mod, function(mod) {
-    result <- data.frame(actual_ss = seq(0, 10, 0.1),
-                         offset0 = log(0.05/0.95))
-    result$power <- predict(mod, type = "response", newdata = result)
-    result
-  })) %>%
-  select(-dat, -mod) %>%
-  unnest(power) %>%
-  ggplot() +
-  geom_line(aes(actual_ss, power, col = name)) +
-  facet_wrap(~type, scales = "free_x")
 ```
@@
 
-```{r pcp, fig.cap = "Parallel coordinate plots of decisions made by computer vision model, conventional tests and visual tests made by human.", eval = TRUE}
+```{r pcp, fig.cap = "Parallel coordinate plots of decisions made by computer vision model, conventional tests and visual tests made by human.", eval = TRUE, fig.height = 4}
 library(ggpcp)
@@
 
-```{r results = "hide"}
-library(VennDiagram)
-mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
-  filter(type == "Non-linearity") %>%
-  filter(!null_lineup, !attention_check) %>%
-  select(unique_lineup_id, type, actual_ss, model_reject_20, reject, conventional_reject) %>%
-  mutate(across(c(model_reject_20, conventional_reject, reject), ~ifelse(.x, "Reject", "Not reject"))) %>%
-  pivot_longer(c(model_reject_20, reject, conventional_reject)) %>%
-  group_by(name) %>%
-  summarise(ids = list(unique_lineup_id[value != "Reject"])) %>%
-  pull(ids) %>%
-  venn.diagram(
-    category.names = c("Conventional test" , "Computer vision model" , "Visual tests"),
-    filename = here::here('paper/diagram/poly_venn_diagramm.png'),
-    disable.logging = TRUE,
-    lwd = 1,
-    col = RColorBrewer::brewer.pal(3, "Dark2"),
-    fill = alpha(RColorBrewer::brewer.pal(3, "Dark2"), 0.3),
-    cex = 0.5,
-    fontfamily = "sans",
-    cat.cex = 0.5,
-    cat.default.pos = "outer",
-    cat.fontfamily = "sans",
-    cat.pos = c(330, 30, 180),
-    cat.dist = c(0, 0, 0.03),
-    cat.col = RColorBrewer::brewer.pal(3, "Dark2"),
-    height = 800,
-    width = 800,
-    resolution = 300,
-    main.fontfamily = "sans",
-    main.cex = 0.6,
-    main = "Non-linearity",
-    main.pos = c(0.25, 1.1)
-  )
-
-mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
-  filter(type != "Non-linearity") %>%
-  filter(!null_lineup, !attention_check) %>%
-  select(unique_lineup_id, type, actual_ss, model_reject_20, reject, conventional_reject) %>%
-  mutate(across(c(model_reject_20, conventional_reject, reject), ~ifelse(.x, "Reject", "Not reject"))) %>%
-  pivot_longer(c(model_reject_20, reject, conventional_reject)) %>%
-  group_by(name) %>%
-  summarise(ids = list(unique_lineup_id[value != "Reject"])) %>%
-  pull(ids) %>%
-  venn.diagram(
-    category.names = c("Conventional test" , "Computer vision model" , "Visual tests"),
-    filename = here::here('paper/diagram/heter_venn_diagramm.png'),
-    disable.logging = TRUE,
-    lwd = 1,
-    col = RColorBrewer::brewer.pal(3, "Dark2"),
-    fill = alpha(RColorBrewer::brewer.pal(3, "Dark2"), 0.3),
-    cex = 0.5,
-    fontfamily = "sans",
-    cat.cex = 0.5,
-    cat.default.pos = "outer",
-    cat.fontfamily = "sans",
-    cat.pos = c(330, 30, 180),
-    cat.dist = c(0, 0, 0.03),
-    cat.col = RColorBrewer::brewer.pal(3, "Dark2"),
-    height = 800,
-    width = 800,
-    resolution = 300,
-    main.fontfamily = "sans",
-    main.cex = 0.6,
-    main = "Heteroskedasticity",
-    main.pos = c(0.25, 1.1)
-  )
-```
-
-```{r venn, fig.pos = "!h", fig.cap = "Venn diagram for non-rejection decisions made by conventional tests, computer vision model, and visual tests conducted by human.", eval = FALSE}
-poly_venn <- magick::image_read(here::here("paper/diagram/poly_venn_diagramm.png")) %>%
-  magick::image_ggplot()
-heter_venn <- magick::image_read(here::here("paper/diagram/heter_venn_diagramm.png")) %>%
-  magick::image_ggplot()
-patchwork::wrap_plots(poly_venn, heter_venn, ncol = 2)
-```
-
-```{r eval = FALSE}
-mutate(experiment, type = ifelse(type == "polynomial", "Non-linearity", "Heteroskedasticity")) %>%
-  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
-  filter(!null_lineup, !attention_check) %>%
-  pivot_longer(c(model_reject_20, reject, conventional_reject)) %>%
-  mutate(name = ifelse(name == "model_reject_20", "Computer vision model", name)) %>%
-  mutate(name = ifelse(name == "reject", "Visual test", name)) %>%
-  mutate(name = ifelse(name == "conventional_reject", "Conventional test", name)) %>%
-  ggplot() +
-  ggbeeswarm::geom_quasirandom(aes(actual_ss, value, col = name), size = 0.5, alpha = 0.4) +
-  geom_line(data = model_glm_pred_poly, aes(d, value + 1, col = "Computer vision model"), linewidth = 0.8) +
-  geom_line(data = model_glm_pred_heter, aes(d, value + 1, col = "Computer vision model"), linewidth = 0.8) +
-  geom_line(data = visual_glm_pred_poly, aes(d, value + 1, col = "Visual test"), linewidth = 0.8) +
-  geom_line(data = visual_glm_pred_heter, aes(d, value + 1, col = "Visual test"), linewidth = 0.8) +
-  geom_line(data = conv_glm_pred_poly, aes(d, value + 1, col = "Conventional test"), linewidth = 0.8) +
-  geom_line(data = conv_glm_pred_heter, aes(d, value + 1, col = "Conventional test"), linewidth = 0.8) +
-  geom_line(data = mutate(model_glm_pred_poly, name = "Visual test"), aes(d, value + 1, col = "Computer vision model"), linetype = 2, linewidth = 0.8) +
-  geom_line(data = mutate(model_glm_pred_heter, name = "Visual test"), aes(d, value + 1, col = "Computer vision model"), linetype = 2, linewidth = 0.8) +
-  geom_line(data = mutate(model_glm_pred_poly, name = "Conventional test"), aes(d, value + 1, col = "Computer vision model"), linetype = 2, linewidth = 0.8) +
-  geom_line(data = mutate(model_glm_pred_heter, name = "Conventional test"), aes(d, value + 1, col = "Computer vision model"), linetype = 2, linewidth = 0.8) +
-  geom_line(data = mutate(visual_glm_pred_poly, name = "Computer vision model"), aes(d, value + 1, col = "Visual test"), linetype = 2, linewidth = 0.8) +
-  geom_line(data = mutate(visual_glm_pred_heter, name = "Computer vision model"), aes(d, value + 1, col = "Visual test"), linetype = 2, linewidth = 0.8) +
-  geom_line(data = mutate(visual_glm_pred_poly, name = "Conventional test"), aes(d, value + 1, col = "Visual test"), linetype = 2, linewidth = 0.8) +
-  geom_line(data = mutate(visual_glm_pred_heter, name = "Conventional test"), aes(d, value + 1, col = "Visual test"), linetype = 2, linewidth = 0.8) +
-  geom_line(data = mutate(conv_glm_pred_poly, name = "Computer vision model"), aes(d, value + 1, col = "Conventional test"), linetype = 2, linewidth = 0.8) +
-  geom_line(data = mutate(conv_glm_pred_heter, name = "Computer vision model"), aes(d, value + 1, col = "Conventional test"), linetype = 2, linewidth = 0.8) +
-  geom_line(data = mutate(conv_glm_pred_poly, name = "Visual test"), aes(d, value + 1, col = "Conventional test"), linetype = 2, linewidth = 0.8) +
-  geom_line(data = mutate(conv_glm_pred_heter, name = "Visual test"), aes(d, value + 1, col = "Conventional test"), linetype = 2, linewidth = 0.8) +
-  facet_grid(name ~ type, scales = "free_x") +
-  theme_light() +
-  scale_color_brewer("", palette = "Dark2") +
-  theme(legend.position = "bottom") +
-  xlab("D") +
-  ylab("Reject")
-
-```
-
-```{r delta, fig.pos = "!h", fig.cap = "A weighted detection rate vs $\\delta$-differnence plot."}
+```{r delta, fig.pos = "!h", fig.cap = "A weighted detection rate vs $\\delta$-differnence plot. The brown line is smoothing curve produced by fitting gnealized additive models", fig.height = 4}
 experiment %>%
@@
   ggplot() +
+  geom_vline(xintercept = 0, alpha = 0.4, linetype = 2) +
   geom_point(aes(delta_diff, prop_detect), alpha = 0.3) +
-  geom_smooth(aes(delta_diff, prop_detect), se = FALSE) +
+  geom_smooth(aes(delta_diff, prop_detect), se = FALSE, col = "#994F00") +
   ylab("Weighted detection rate") +
@@
 
-```{r fig.pos = "!h", fig.cap = "A weighted detection rate vs $\\gamma$-number of extreme nulls plot.", eval = FALSE}
-experiment %>%
-  left_join(lineup_vss %>%
-              group_by(unique_lineup_id) %>%
-              summarise(delta_diff = first(delta_diff), gamma_diff = first(gamma_diff))) %>%
-  ggplot() +
-  geom_point(aes(gamma_diff, prop_detect)) +
-  ylab("Weighted detection rate") +
-  xlab(expression(gamma-number~of~extreme~nulls)) +
-  theme_light()
-```
-
 # Examples
 
-<!-- - explain the purpose of each case here as a summary -->
-<!-- - left-triangle: conventional reject, other two do not reject -->
-<!-- - Boston: all three reject -->
-<!-- - datasaurus: conv not reject -->
-
-<!-- - we don't have one conv reject, model reject, human not -->
-
-<!-- conv: reject 1, not 0 -->
-<!-- model: reject 1, not 0 -->
-<!-- human: reject 1, not 0 -->
-
-<!-- left:       100 -->
-<!-- boston:     111 -->
-<!-- datasaurus: 011 -->
-
-<!-- example?:   001 (probably because of outliers) -->
-<!-- try "baker" dataset -->
-
-In this section, we will utilize the trained computer vision model on three example datasets.
+In this section, we will present the performance of trained computer vision model on three example datasets.
 These include the dataset associated with the residual plot displaying a "left-triangle" shape, as displayed in Figure \ref{fig:false-finding}, along with the Boston housing dataset, and the "dino" datasets from the `datasaurus` R package.
@@
 The third example presents a situation where the model deviation is non-typical.
-Here, the computer vision model and human visual inspection reject $H_0$, whereas conventional tests do not.
+Here, the computer vision model and human visual inspection reject $H_0$, whereas some commonly used conventional tests do not.
 This emphasizes the benefits of visual inspection and the unique advantage of the computer vision model, which, like humans, makes decisions based on visual discoveries.
@@
 
-Figure \ref{fig:false-check} presents the results of the assessment.
+Figure \ref{fig:false-check} presents the results of the assessment by the computer vision model.
 Notably, the observed visual signal strength is considerably lower than the 95% sample quantile of the null distribution.
@@
 mod <- lm(y ~ x, data = dat)
-my_vi <- autovi::auto_vi(fitted_mod = mod)
+my_vi <- autovi::auto_vi(fitted_model = mod)
 ```
@@
 if (!file.exists(here::here("paper/data/fig1_check.rds"))) {
-  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
-  my_vi$check(null_draws = 200L, boot_draws = 200L, extract_feature_from_layer = "global_max_pooling2d")
+  keras_model <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  my_vi$check(null_draws = 200L, boot_draws = 200L, keras_model = keras_model, extract_feature_from_layer = "global_max_pooling2d")
   saveRDS(my_vi$check_result, here::here("paper/data/fig1_check.rds"))
@@
 if (!file.exists(here::here("paper/data/fig1_attention_map.rds"))) {
+  reticulate::py_set_seed(10086, disable_hash_randomization = TRUE)
   define_gradient()
@@
   plot_path <- my_vi$plot_resid() %>%
-  my_vi$save_plot()
+    autovi::save_plot()
   input_auxiliary <- my_vi$auxiliary()
@@
 ```{r false-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
-(A) The true residual plot exhibiting "left-triangle" shape.
+(A) The true residual plot exhibiting a "left-triangle" shape.
 (B) The attention map produced by computing the gradient of the output with respect to the greyscale input.
 (C) The density plot of estimated distnace for null plots and bootstrapped plots.
-The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots.
+The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots.
 The fitted model will not be rejected since $\\hat{D} < Q_{null}(0.95)$.
@@
   xlab(expression(hat(D))) +
-  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D))))
-p4 <- my_vi$feature_pc() %>%
-  mutate(type = ifelse(type == "boot", "Boot", type)) %>%
-  mutate(type = ifelse(type == "null", "Null", type)) %>%
+  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D)))) +
+  scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) +
+  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
+p4 <- my_vi$feature_pca() %>%
+  mutate(set = ifelse(set == "boot", "Boot", set)) %>%
+  mutate(set = ifelse(set == "null", "Null", set)) %>%
   ggplot() +
-  geom_point(data = ~filter(.x, type != "observed"), aes(PC1, PC2, col = type, shape = "Simulated"), size = 2, alpha = 0.3) +
-  geom_point(data = ~filter(.x, type == "observed"), aes(PC1, PC2, shape = "Observed"), size = 4, col = "black", fill = "black") +
+  geom_point(data = ~filter(.x, set != "observed"), aes(PC1, PC2, col = set), size = 2, alpha = 0.3) +
+  geom_point(data = ~filter(.x, set == "observed"), aes(PC1, PC2), size = 4, col = "black", fill = "black") +
+  geom_text(data = ~filter(.x, set == "observed"), aes(PC1, PC2, label = "Observed"), nudge_x = 10) +
   theme_light() +
   scale_color_brewer(palette = "Dark2") +
-  scale_shape_manual(values = c("Simulated" = 19, "Observed" = 17)) +
   xlab("PC1") +
   ylab("PC2") +
-  labs(col = "", shape = "") +
-  theme(legend.position = "bottom", legend.box="vertical") +
-  ggtitle("(D) PCA for extracted features")
+  labs(col = "") +
+  theme(legend.position = "bottom") +
+  ggtitle("(D) PCA for extracted features") +
+  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
 
@@
 Figure \ref{fig:boston-check} displays the residual plot and the assessment conducted by the computer vision model.
-A clear non-linearity pattern resembling a "U" shape is shown in the residual plot.
+A clear non-linearity pattern resembling a "U" shape is shown in the plot A.
 Furthermore, the RESET test yields a very small $p$-value.
-The approximated distance $\hat{D}$ significantly exceeds $Q_{null}(0.95)$, leading to rejection of $H_0$.
+The estimated distance $\hat{D}$ significantly exceeds $Q_{null}(0.95)$, leading to rejection of $H_0$.
 The bootstrapped distribution also suggests that almost all the bootstrapped fitted models will be rejected, indicating that the fitted model is unlikely to be correctly specified.
-
-Additionally, the lineup of this fitted model presented in Figure \ref{fig:boston-lineup} underscores that the "U" shape is a distinctive characteristic of the true residual plot, absent in any of the null plots.
-Consequently, if a visual test is conducted, $H_0$ will also be rejected.
+The attention map in plot B suggests the center of the image has higher leverage than other areas, and it is the turning point of the "U" shape.
+The principal component analysis provided in plot D shows two distinct clusters of data points, further underlines the visual differences between bootstrapped plots and null plots.
+This coincides the findings from Figure  \ref{fig:boston-lineup}, where the true plot exhibiting a "U" shape is visually distinctive from null plots.
+If a visual test is conducted by human, $H_0$ will also be rejected.
 
@@
 if (!file.exists(here::here("paper/data/boston_check.rds"))) {
-  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
-  my_vi$check(null_draws = 200L, boot_draws = 200L, extract_feature_from_layer = "global_max_pooling2d")
+  keras_model <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  my_vi$check(null_draws = 200L, boot_draws = 200L, keras_model = keras_model, extract_feature_from_layer = "global_max_pooling2d")
   saveRDS(my_vi$check_result, here::here("paper/data/boston_check.rds"))
@@
 if (!file.exists(here::here("paper/data/boston_attention_map.rds"))) {
+  reticulate::py_set_seed(10086, disable_hash_randomization = TRUE)
   define_gradient()
@@
 
-```{r boston-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
-On the left, the residual plot of the Boston housing fitted model is displayed.
-The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots.
-The fitted model will be rejected since $\\hat{D} >= Q_{null}(0.95)$.'}
+```{r boston-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment for the Boston housing fitted model evaluted on 200 null plots and 200 bootstrapped plots.
+(A) The true residual plot exhibiting a "U" shape.
+(B) The attention map produced by computing the gradient of the output with respect to the greyscale input.
+(C) The density plot of estimated distnace for null plots and bootstrapped plots.
+The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots.
+The fitted model will be rejected since $\\hat{D} \\geq Q_{null}(0.95)$.
+(D) PC2 vs PC1 plot for features extracted from the global pooling layer of the computer vision model.
+'}
 p1 <- my_vi$plot_resid(theme = theme_light()) + ggtitle("(A) Residual plot") 
@@
   xlab(expression(hat(D))) +
-  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D))))
-p4 <- my_vi$feature_pc() %>%
-  mutate(type = ifelse(type == "boot", "Boot", type)) %>%
-  mutate(type = ifelse(type == "null", "Null", type)) %>%
+  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D)))) +
+  scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) +
+  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
+p4 <- my_vi$feature_pca() %>%
+  mutate(set = ifelse(set == "boot", "Boot", set)) %>%
+  mutate(set = ifelse(set == "null", "Null", set)) %>%
   ggplot() +
-  geom_point(data = ~filter(.x, type != "observed"), aes(PC1, PC2, col = type, shape = "Simulated"), size = 2, alpha = 0.3) +
-  geom_point(data = ~filter(.x, type == "observed"), aes(PC1, PC2, shape = "Observed"), size = 4, col = "black", fill = "black") +
+  geom_point(data = ~filter(.x, set != "observed"), aes(PC1, PC2, col = set), size = 2, alpha = 0.3) +
+  geom_point(data = ~filter(.x, set == "observed"), aes(PC1, PC2), size = 4, col = "black", fill = "black") +
+  geom_text(data = ~filter(.x, set == "observed"), aes(PC1, PC2, label = "Observed"), nudge_x = 7) +
   theme_light() +
   scale_color_brewer(palette = "Dark2") +
-  scale_shape_manual(values = c("Simulated" = 19, "Observed" = 17)) +
   xlab("PC1") +
   ylab("PC2") +
-  labs(col = "", shape = "") +
-  theme(legend.position = "bottom", legend.box="vertical") +
-  ggtitle("(D) PCA for extracted features")
+  labs(col = "") +
+  theme(legend.position = "bottom") +
+  ggtitle("(D) PCA for extracted features") +
+  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
 
@@
 The "dino" dataset within the `datasaurus` R package exemplifies this scenario.
-With only two columns, x and y, fitting a regression model to this data yields a residual plot resembling a "dinosaur," as displayed in Figure \ref{fig:dino-check}.
+With only two columns, x and y, fitting a regression model to this data yields a residual plot resembling a "dinosaur," as displayed in plot A of Figure \ref{fig:dino-check}.
 Unsurprisingly, this distinct residual plot stands out in a lineup, as shown in Figure \ref{fig:dino-lineup}.
@@
 
+More importantly, the attention map in plot B of Figure \ref{fig:dino-check} clearly exhibits a "dinosaur" shape, strongly suggesting the prediction of the distance is based on human perceptible visual patterns.
+The computer vision model is also capable of capturing the contour or the outline of the embedded shape, just like human being reading residual plots.
+The principle component analysis in plot D of Figure \ref{fig:dino-check} also shows that the cluster of bootstrapped plots is at the corner of the cluster of null plots.
+
+More importantly, the attention map in plot B of Figure \ref{fig:dino-check} clearly exhibits a "dinosaur" shape, strongly suggesting that the distance prediction is based on human-perceptible visual patterns.
+The computer vision model effectively captures the contour or outline of the embedded shape, similar to how humans interpret residual plots.
+Additionally, the principal component analysis in plot D of Figure \ref{fig:dino-check} demonstrates that the cluster of bootstrapped plots is positioned at the corner of the cluster of null plots.
+
 In practice, without accessing the residual plot, it would be challenging to identify the artificial pattern of the residuals.
@@
 if (!file.exists(here::here("paper/data/dino_check.rds"))) {
-  my_vi$keras_mod <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
-  my_vi$check(null_draws = 200L, boot_draws = 200L, extract_feature_from_layer = "global_max_pooling2d")
+  keras_model <- keras::load_model_tf(here::here("paper/data/phn_v2_32.keras"))
+  my_vi$check(null_draws = 200L, boot_draws = 200L, keras_model = keras_model, extract_feature_from_layer = "global_max_pooling2d")
   saveRDS(my_vi$check_result, here::here("paper/data/dino_check.rds"))
@@
 if (!file.exists(here::here("paper/data/dino_attention_map.rds"))) {
+  reticulate::py_set_seed(10086, disable_hash_randomization = TRUE)
   define_gradient()
@@
 
-```{r dino-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
-On the left, the residual plot exhitbiting a "dinosaur" shape is displayed.
-The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots.
-The fitted model will be rejected since $\\hat{D} >= Q_{null}(0.95)$.'}
+```{r dino-check, fig.pos = "!h", fig.cap = 'A summary of the residual plot assessment for the Datasaurus fitted model evaluted on 200 null plots and 200 bootstrapped plots.
+(A) The true residual plot exhibiting a "dinosaur" shape.
+(B) The attention map produced by computing the gradient of the output with respect to the greyscale input.
+(C) The density plot of estimated distnace for null plots and bootstrapped plots.
+The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots.
+The fitted model will be rejected since $\\hat{D} \\geq Q_{null}(0.95)$.
+(D) PC2 vs PC1 plot for features extracted from the global pooling layer of the computer vision model.'}
 
@@
   xlab(expression(hat(D))) +
-  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D))))
-p4 <- my_vi$feature_pc() %>%
-  mutate(type = ifelse(type == "boot", "Boot", type)) %>%
-  mutate(type = ifelse(type == "null", "Null", type)) %>%
+  scale_linetype(labels = c("95% quantile of the null distribution", expression(hat(D)))) +
+  scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) +
+  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
+p4 <- my_vi$feature_pca() %>%
+  mutate(set = ifelse(set == "boot", "Boot", set)) %>%
+  mutate(set = ifelse(set == "null", "Null", set)) %>%
   ggplot() +
-  geom_point(data = ~filter(.x, type != "observed"), aes(PC1, PC2, col = type, shape = "Simulated"), size = 2, alpha = 0.3) +
-  geom_point(data = ~filter(.x, type == "observed"), aes(PC1, PC2, shape = "Observed"), size = 4, col = "black", fill = "black") +
+  geom_point(data = ~filter(.x, set != "observed"), aes(PC1, PC2, col = set), size = 2, alpha = 0.3) +
+  geom_point(data = ~filter(.x, set == "observed"), aes(PC1, PC2), size = 4, col = "black", fill = "black") +
+  geom_text(data = ~filter(.x, set == "observed"), aes(PC1, PC2, label = "Observed"), nudge_x = -10) +
   theme_light() +
   scale_color_brewer(palette = "Dark2") +
-  scale_shape_manual(values = c("Simulated" = 19, "Observed" = 17)) +
   xlab("PC1") +
@@
   theme(legend.position = "bottom", legend.box="vertical") +
-  ggtitle("(D) PCA for extracted features")
+  ggtitle("(D) PCA for extracted features") +
+  scale_color_manual(values = c("#40B0A6", "#E1BE6A"))
 
@@
 
-<!-- ## When the model works -->
-
-<!-- - simple examples (non-linearity, heteroskedasticity, ...) -->
-<!-- - datasaurus -->
-
-<!-- ## When the model does not work -->
-
-<!-- - human detect but model does not -->
-<!-- - cartoon residuals? -->
-
-<!-- ## Workflow: how one use this model? (small showcase)  -->
-
-<!-- # Discussion -->
-
-<!-- There are other kinds of residual departures like autocorrelation that are not considered in this study.
-The primary goal of this study is to establish a new way of evaluating residual plot and conducting visual test with computer vision models.
-Building computer vision models for other model violations and other types of diagnostic plots could be future directions of this field.
--->
-
-# Extension to other types of regression model
-
-1. Use raw residuals.
-  - D (effect size) for null plots will be greater than zero.
-  - Can not assume D for actual plots are greater than D for null plots any more. -> KL null - normal, KL actual - normal
-  - Need to use a two-sided test instead of a one-sided test.
-  - It may fail completely due to the identification issue.
-2. Use transformed residuals that roughly follow a multivariate normal distribution.
-  - The one-sided test will still work. 
-  - Need to ensure that the transformed actual plot will have residual distribution different from the transformed null distribution. 
-3. Train a new model.
-  - KL divergence may not be the appropriate measure for the difference between the residual distributions. Could consider using a new distance measure.
+# Limitations and Future Work
+
+Despite the computer vision model performing well with general cases under the synthetic data generation scheme and the three examples used in this paper, this study has several limitations that could guide future work.
+
+The proposed distance measure assumes that the true model is a classical normal linear regression model, which can be too restrictive.
+Although this paper does not address the relaxation of this assumption, there are potential methods to evaluate other types of regression models.
+The most comprehensive approach would be to define a distance measure for each different class of regression model and then train the computer vision model following the methodology described in this paper.
+To accelerate training, one could use the convolutional blocks of our trained model as a feature extractor and perform transfer learning on top of it, as these blocks effectively capture shapes in residual plots.
+Another approach would be to transform the residuals so they are roughly normally distributed and have constant variance.
+If only raw residuals are used, the distance-based statistical testing compares the difference in distance to a classical normal linear regression model for the true plot and null plots.
+This comparison is meaningful only if the difference can be identified by the distance measure proposed in this paper.
+
+There are other types of residual plots commonly used in diagnostics, such as residuals vs.
+predictor and quantile-quantile plots.
+In this study, we focused on the most commonly used residual plot as a starting point for exploring the new field of automated visual inference.
+Similarly, we did not explore other, more sophisticated computer vision model architectures and specifications for the same reason.
+While the performance of the computer vision model is acceptable, there is still room for improvement to achieve behavior more closely resembling that of humans interpreting residual plots.
+This may require external survey data or human subject experiment data to understand the fundamental differences between our implementation and human evaluation.
 
@@
 
-- Brief discussion
-- Summary of findings
-- Contributions to the field
-
-In this study, we have introduced a distance metric based on Kullback-Leibler divergence to quantify the disparity between the residual distribution of a fitted linear regression model and the reference residual distribution assumed under correct model specification.
-This distance metric effectively captures the magnitude of model violations in misspecified models.
-We propose a computer vision model to approximate this distance, utilizing the residual plot of the fitted model as input.
-The resulting approximated distance serves as the foundation for constructing a single Model Violation Index (MVI), facilitating the quantification of various model violations.
-
-Moreover, the approximated distance enables the development of a formal statistical testing procedure by evaluating a large number of null plots generated from the fitted model.
+In this study, we have introduced a distance measure based on Kullback-Leibler divergence to quantify the disparity between the residual distribution of a fitted classical normal linear regression model and the reference residual distribution assumed under correct model specification.
+This distance measure effectively captures the magnitude of model violations in misspecified models.
+We propose a computer vision model to estimate this distance, utilizing the residual plot of the fitted model as input.
+The resulting estimated distance serves as the foundation for constructing a single Model Violation Index (MVI), facilitating the quantification of various model violations.
+
+Moreover, the estimated distance enables the development of a formal statistical testing procedure by evaluating a large number of null plots generated from the fitted model.
 Additionally, employing bootstrapping techniques and refitting the regression model allows us to ascertain how frequently the fitted model is considered misspecified if data were repeatedly obtained from the same data generating process.
@@
 The trained computer vision model demonstrates strong performance on both the training and test sets, although it exhibits slightly lower performance on residual plots with non-linearity visual patterns compared to other types of violations.
-The statistical tests relying on the approximated distance predicted by the computer vision model exhibit lower sensitivity compared to conventional tests but higher sensitivity compared to visual tests conducted by humans.
-While the approximated distance generally mirrors the strength of the visual signal perceived by humans, there remains scope for further improvement in its performance.
+The statistical tests relying on the estimated distance predicted by the computer vision model exhibit lower sensitivity compared to conventional tests but higher sensitivity compared to visual tests conducted by humans.
+While the estimated distance generally mirrors the strength of the visual signal perceived by humans, there remains scope for further improvement in its performance.
 
@@
 In contrast, individual residual diagnostic tests such as the RESET test and the Breusch-Pagan test only evaluate specific model assumptions.
-
 In practice, selecting an appropriate set of statistical tests for regression diagnostics can be challenging, particularly given the necessity of adjusting the significance level for each test.
@@
 
-- Future directions
-
 # Acknowledgements {-}
 
-- Packages used
-- Website link
-
-```{r eval = FALSE}
-
-# This shows that even if the model can 100% accurately predict the distance for not null plots, 
-# the statistical tests based on approximated distance will still reject more cases than visual tests
-# So it has very little to do with the accuracy of the prediction
-experiment %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  mutate(model_reject_true = actual_ss > vss_null_95) %>%
-  filter(model_reject_true != reject) %>%
-  ggplot() +
-  ggbeeswarm::geom_quasirandom(aes(actual_ss, reject)) +
-  ggtitle("Unmatched cases") +
-  ylab("Visual tests reject") +
-  xlab("D")
-```
-
-```{r eval = FALSE}
-
-map_dbl(10^seq(-10, 0, 0.01), function(crt) {
-  experiment %>%
-    filter(!null_lineup) %>%
-    filter(!attention_check) %>%
-    mutate(mvi = vss - log(n) + 10) %>%
-    mutate(actual_ss_cut = conventional_p_value <= crt) %>%
-    mutate(agree = actual_ss_cut == reject) %>%
-    pull(agree) %>%
-    mean()
-}) %>%
-  max()
-
-experiment %>%
-  filter(!null_lineup) %>%
-  filter(!attention_check) %>%
-  mutate(mvi = vss - log(n) + 10) %>%
-  ggplot() +
-  ggbeeswarm::geom_quasirandom(aes(vss, reject)) +
-  ggtitle("Unmatched cases") +
-  ylab("Visual tests reject") +
-  xlab("D")
-  mutate(model_reject_true = actual_ss > vss_null_95) %>%
-  filter(model_reject_true != reject)
-```
+These `R` packages were used for the work: `tidyverse` [@tidyverse], `lmtest` [@lmtest], `mpoly` [@mpoly], `ggmosaic` [@ggmosaic], `kableExtra` [@kableextra], `patchwork` [@patchwork], `cassowaryr` [@mason2022cassowaryr], `bandicoot` [@bandicoot], `ggpcp` [@ggpcp], `keras` [@keras] and `here` [@here].
+
+The article was created with R packages `rticles` [@rticles], `knitr` [@knitr] and `rmarkdown` [@rmarkdown].
+The project's GitHub repository (https://github.com/TengMCing/auto_residual_reading/paper) contains all materials required to reproduce this article.
+
+# Disclosure Statement {-}
+
+No potential conflict of interest was reported by the author(s).
+
+# Supplementary Materials {-}
+
+\begin{description}
+\item{Appendix:} The appendix includes more details about the neural network layers used for constructing the computer vision models.
+\end{description}

### 2024-05-27 paper/paper.tex
wait for review

@@
 \begin{abstract}
-TBD.
+Plotting residuals is a standard practice in linear regression
+diagnostics, essential for identifying deviations from model assumptions
+such as linearity, homoscedasticity, and normality. Visual inference
+provides an inferential framework to assess whether residual plots
+contain patterns inconsistent with model assumptions, typically using a
+lineup protocol. However, the lineup protocol's reliance on human
+judgment limits its scalability. This study addresses this limitation by
+automating the interpretation of residual plots using computer vision
+models. We develop a distance measure based on Kullback-Leibler
+divergence to quantify the disparity between the residual distribution
+of a fitted classical normal linear regression model and the reference
+distribution. We propose a computer vision model to estimate this
+distance from residual plots, facilitating formal statistical testing
+and bootstrapping techniques to assess model specification. Our computer
+vision model shows strong performance, though it performs slightly less
+effectively on non-linearity visual patterns. The statistical tests
+based on the estimated distance exhibit lower sensitivity than
+conventional tests but higher sensitivity than human visual tests.
+Examples demonstrate the method's effectiveness across different
+scenarios, highlighting its value in automating the diagnostic process
+and supplementing traditional methods.
 \end{abstract}
@@
 In this paper, we develop computer vision models and integrate them into
-the residual plots diagnostics workflow, filling the gap of\ldots. The
-paper is structured as follows: \ldots{}
+the residual plots diagnostics workflow, addressing the need for
+automated visual inference. The paper is structured as follows. Section
+\ref{model-specifications} discusses various specifications of the
+computer vision models. Section \ref{distance-between-residual-plots}
+defines the distance measure used to measure model violations. Section
+\ref{distance-estimation} explains how the computer vision models
+estimate this distance measure. Section \ref{statistical-testing} covers
+the statistical testing based on the estimated distance. Sections
+\ref{data-generation}, \ref{model-architecture}, and
+\ref{model-training} detail the data preparation, model architecture,
+and training process, respectively. The results are presented in Section
+\ref{results}. Finally, we conclude with a discussion of our findings
+and propose ideas for future research directions.
 
@@
 
-\section{Methodology}\label{methodology}
-
-\subsection{Model Specifications}\label{model-specifications}
+\section{Model Specifications}\label{model-specifications}
 
@@
 
-\subsubsection{Input Formats}\label{input-formats}
+\subsection{Input Formats}\label{input-formats}
 
@@
 
-\subsubsection{Output Formats}\label{output-formats}
+\subsection{Output Formats}\label{output-formats}
 
@@
 
-\subsection{Distance Between Residual
+\section{Distance Between Residual
 Plots}\label{distance-between-residual-plots}
@@
 \citep{chowdhury2018measuring}. Furthermore, scagnostics mentioned in
-Section \ref{introduction} could be used to construct distance metrics
+Section \ref{introduction} could be used to construct distance measures
 for residual plots comparison, but the functional form still needs to be
@@
 
-\subsubsection{Residual Distribution}\label{residual-distribution}
+\subsection{Residual Distribution}\label{residual-distribution}
 
@@
 \(\boldsymbol{\varepsilon}\) follows a non-normal distribution such as a
-multivariate log-normal distribution, \(P\) will usually be skewed and
+multivariate lognormal distribution, \(P\) will usually be skewed and
 has a long tail.
 
-\subsubsection{\texorpdfstring{Distance of \(P\) from
+\subsection{\texorpdfstring{Distance of \(P\) from
 \(Q\)}{Distance of P from Q}}\label{distance-of-p-from-q}
@@
 
-\subsubsection{\texorpdfstring{Non-normal
+\subsection{\texorpdfstring{Non-normal
 \(P\)}{Non-normal P}}\label{non-normal-p}
@@
 
-\subsection{Distance Estimation}\label{distance-estimation}
+\section{Distance Estimation}\label{distance-estimation}
 
@@
 residual distribution \(P\) and an ideal reference distribution \(Q\).
-You may have noticed that this distance measure can only be computed
-when the data generating process is known. In reality, we often have no
+It can be noticed that this distance measure can only be computed when
+the data generating process is known. In reality, we often have no
 knowledge about the data generating process, otherwise we do not need to
@@
 
-\subsection{Model Violations Index
-(MVI)}\label{sec:model-violations-index}
+\section{Model Violations Index (MVI)}\label{sec:model-violations-index}
 
@@
 
-\subsection{Statistical testing}\label{statistical-testing}
-
-\subsubsection{Lineup Evaluation}\label{lineup-evaluation}
+\section{Statistical testing}\label{statistical-testing}
+
+\subsection{Lineup Evaluation}\label{lineup-evaluation}
 
@@
 if the data plot is identified as the most distinct plot by the sole
-participant. The estimated distance serves as a metric to quantify the
+observer. The estimated distance serves as a metric to quantify the
 difference between the data plot and the null plots, as intended.
@@
 level can be maintained if the number of plots is a multiple of 20.
-Specifically, for lineups comprising \(20t\) plots, where \(l\) is a
+Specifically, for lineups comprising \(20t\) plots, where \(t\) is a
 positive integer, we reject \(H_0\) if \(\hat{D}\) exceeds 95\%
@@
 in this case is given by
-\(\frac{1}{20t} + \frac{1}{20t}\sum{i=1}^{20t-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)\),
+\(\frac{1}{20t} + \frac{1}{20t}\sum_{i=1}^{20t-1} I\left(\hat{D}_{null}^{(i)} > \hat{D}\right)\),
 where \(I(\cdot)\) is the indicator function.
@@
 \({\hat{D}_{null}^{(i)}}\) can be viewed as an approximation of the null
-distribution of the approximated distance. Consequently, quantiles of
-the null distribution can be estimated using the sample quantiles
-available in statistical software such as R, and these quantiles can be
-utilized for decision-making purposes. The details of the sample
-quantile computation can be found in \citet{hyndman1996sample}. For
-instance, if \(\hat{D}\) is greater than or equal to the 95\% sample
-quantile, denoted as \(Q_{null}(0.95)\), we can conclude that the
-estimated distance for the true residual plot is significantly different
-from the estimated distance for null plots with a 95\% significance
-level. Based on our experience, to obtain a stable estimate of the 95\%
-quantile, the number of null plots, \(n_{null}\), typically needs to be
-at least 100. However, if the null distribution exhibits a long tail, a
-larger number of null plots may be required. Alternatively, a
-\(p\)-value can be used to represent the probability of observing an
-event equally or more extreme than the given event under the null
-hypothesis \(H_0\), and it can be estimated by
+distribution of the estimated distance. Consequently, quantiles of the
+null distribution can be estimated using the sample quantiles available
+in statistical software such as R, and these quantiles can be utilized
+for decision-making purposes. The details of the sample quantile
+computation can be found in \citet{hyndman1996sample}. For instance, if
+\(\hat{D}\) is greater than or equal to the 95\% sample quantile,
+denoted as \(Q_{null}(0.95)\), we can conclude that the estimated
+distance for the true residual plot is significantly different from the
+estimated distance for null plots with a 95\% significance level. Based
+on our experience, to obtain a stable estimate of the 95\% quantile, the
+number of null plots, \(n_{null}\), typically needs to be at least 100.
+However, if the null distribution exhibits a long tail, a larger number
+of null plots may be required. Alternatively, a \(p\)-value can be used
+to represent the probability of observing an event equally or more
+extreme than the given event under the null hypothesis \(H_0\), and it
+can be estimated by
 \(\frac{1}{m}\sum_{i=1}^{m-1}I\left(\hat{D}_{null}^{(i)} \geq \hat{D}\right)\).
@@
 
-\subsubsection{Bootstrapping}\label{bootstrapping}
+\subsection{Bootstrapping}\label{bootstrapping}
 
@@
 
-\subsection{Data Genearation}\label{data-genearation}
-
-\subsubsection{Simulation Scheme}\label{simulation-scheme}
+\section{Data Generation}\label{data-generation}
+
+\subsection{Simulation Scheme}\label{simulation-scheme}
 
@@
 \(D\) unattainable. Consequently, the computer vision models developed
-in this study were trained using synthetic data. This approach provided
-us with precise label annotations. Additionally, it ensured a large and
-diverse training dataset, as we had control over the data generating
-process, and the simulation of the training data was relatively
-cost-effective.
+in this study were trained using synthetic data, including 80000
+training images and 8000 test images. This approach provided us with
+precise label annotations. Additionally, it ensured a large and diverse
+training dataset, as we had control over the data generating process,
+and the simulation of the training data was relatively cost-effective.
 
@@
 \((.)^2\) operators are element-wise operators, \(\odot\) is the
-Hadamard product, and \(g(., k)\) is a scaling function to enforce the
-support of the random vector to be \([-k, k]^n\) defined as
+Hadamard product, and \(g(\boldsymbol{x}, k)\) is a scaling function to
+enforce the support of the random vector to be \([-k, k]^n\) defined as
 
@@
 b & {}[0, 100]\\
-$\beta_1$ & {0, 1}\\
-$\beta_2$ & {0, 1}\\
+$\beta_1$ & \{0, 1\}\\
+$\beta_2$ & \{0, 1\}\\
 \addlinespace
@@
 distinct error distributions are considered, including discrete,
-uniform, normal, and log-normal distributions, as presented in Figure
+uniform, normal, and lognormal distributions, as presented in Figure
 \ref{fig:different-e}. Each distribution imparts unique characteristics
 to the residuals. The discrete error distribution introduces
-discreteness in residuals, while the log-normal distribution typically
+discreteness in residuals, while the lognormal distribution typically
 yields outliers. Uniform error distribution may result in residuals
@@
 
-\subsubsection{Scagnostics}\label{scagnostics}
+\subsection{Scagnostics}\label{scagnostics}
 
@@
 For each generated residual plot, we computed four scagnostics --
-``Monotonic,'' ``Sparse,'' ``Splines,'' and ``Striped'' -- using the
+``Monotonic'', ``Sparse'', ``Splines'', and ``Striped'' -- using the
 \texttt{cassowaryr} R package \citep{mason2022cassowaryr}. These
@@
 
-\subsubsection{Balanced Dataset}\label{balanced-dataset}
+\subsection{Balanced Dataset}\label{balanced-dataset}
 
@@
 
-\subsection{Model Architecture}\label{model-architecture}
+\section{Model Architecture}\label{model-architecture}
 
@@
 
-\caption{Diagram of the architecture of the optimized computer vision model.}\label{fig:cnn-diag}
+\caption{Diagram of the architecture of the optimized computer vision model.
+Numbers at the bottom of each box show the shape of the output of each layer.
+The band of each box drawn in a darker colour indicates the use of the rectified linear unit activation function.
+Yellow boxes are 2D convolutional layers, orange boxes are pooling layers, the grey box is the concatenation layer, and the purple boxes are dense layers.}\label{fig:cnn-diag}
 \end{figure}
@@
 
-\subsection{Model Training}\label{model-training}
-
-To achieve an optimal deep learning model, hyperparameters like the
+\section{Model Training}\label{model-training}
+
+To achieve a near-optimal deep learning model, hyperparameters like the
 learning rate often need to be fine-tuned using a tuner. In our study,
@@
 
-Our model was trained on the M3 high-performance computing platform,
-using TensorFlow and Keras. During training, 80\% of the training data
-was utilized for actual training, while the remaining 20\% was used as
-validation data. The Bayesian optimization tuner conducted 100 trials to
-identify the best hyperparameter values based on validation root mean
-square error. The tuner then restored the best epoch of the best model
-from the trials. Additionally, we applied early stopping, terminating
-the training process if the validation root mean square error fails to
-improve for 50 epochs. The maximum allowed epochs was set at 2000,
-although no models reached this threshold.
+Our model was trained on the MASSIVE M3 high-performance computing
+platform \citep{goscinski2014multi}, using TensorFlow
+\citep{abadi2016tensorflow} and Keras \citep{chollet2015keras}. During
+training, 80\% of the training data was utilized for actual training,
+while the remaining 20\% was used as validation data. The Bayesian
+optimization tuner conducted 100 trials to identify the best
+hyperparameter values based on validation root mean square error. The
+tuner then restored the best epoch of the best model from the trials.
+Additionally, we applied early stopping, terminating the training
+process if the validation root mean square error fails to improve for 50
+epochs. The maximum allowed epochs was set at 2000, although no models
+reached this threshold.
 
@@
 \caption{Hexagonal heatmap for residuals vs predicted distance on training and test data for three optimized models with different input sizes.
-The red lines are smoothing curves produced by fitting gnealized additive models.
-The area over the zero line in light pink indicates under-prediction, and the area under the zero line in light green indicates over-prediction.}\label{fig:model-performance}
+The brown lines are smoothing curves produced by fitting gnealized additive models.
+The area over the zero line in light yellow indicates under-prediction, and the area under the zero line in light green indicates over-prediction.}\label{fig:model-performance}
 \end{figure}
@@
 The data is grouped by whether the regression has only non-linearity violation, and whether it includes a second predictor in the regression formula.
-The red lines are smoothing curves produced by fitting gnealized additive models.
-The area over the zero line in light pink indicates under-prediction, and the area under the zero line in light green indicates over-prediction.}\label{fig:over-under}
+The brown lines are smoothing curves produced by fitting gnealized additive models.
+The area over the zero line in light yellow indicates under-prediction, and the area under the zero line in light green indicates over-prediction.}\label{fig:over-under}
 \end{figure}
@@
 
-Figure \ref{fig:delta} displays the scatter plot of the weighted
-detection rate vs the \(\delta\)-difference. In the experiment conducted
-in \citet{li2023plot}, participants were allowed to make multiple
-selections for a lineup. The weighted detection rate is computed by
-assigning weights to each detection. If the participant selects zero
-plots, the corresponding weight is 0.05; otherwise, the weight is 1
-divided by the number of selections. The \(\delta\)-difference is
-defined by \citet{chowdhury2018measuring} as
+In the experiment conducted in \citet{li2023plot}, participants were
+allowed to make multiple selections for a lineup. The weighted detection
+rate is computed by assigning weights to each detection. If the
+participant selects zero plots, the corresponding weight is 0.05;
+otherwise, the weight is 1 divided by the number of selections. The
+\(\delta\)-difference is defined by \citet{chowdhury2018measuring} as
 
@@
 
-Figure \ref{fig:delta} indicates that the weighted detection rate
-increases as the \(\delta\)-difference increases, particularly when the
-\(\delta\)-difference is greater than zero. A negative
-\(\delta\)-difference suggests that there is at least one null plot in
-the lineup with a stronger visual signal than the true residual plot. In
-some instances, the weighted detection rate is close to one, yet the
-\(\delta\)-difference is negative. This discrepancy implies that the
-distance measure, or the approximated distance, may not perfectly
-reflect actual human behaviour.
+Figure \ref{fig:delta} displays the scatter plot of the weighted
+detection rate vs the \(\delta\)-difference. It indicates that the
+weighted detection rate increases as the \(\delta\)-difference
+increases, particularly when the \(\delta\)-difference is greater than
+zero. A negative \(\delta\)-difference suggests that there is at least
+one null plot in the lineup with a stronger visual signal than the true
+residual plot. In some instances, the weighted detection rate is close
+to one, yet the \(\delta\)-difference is negative. This discrepancy
+implies that the distance measure, or the estimated distance, may not
+perfectly reflect actual human behaviour.
 
@@
 
-\caption{A weighted detection rate vs $\delta$-differnence plot.}\label{fig:delta}
+\caption{A weighted detection rate vs $\delta$-differnence plot. The brown line is smoothing curve produced by fitting gnealized additive models}\label{fig:delta}
 \end{figure}
@@
 
-In this section, we will utilize the trained computer vision model on
-three example datasets. These include the dataset associated with the
-residual plot displaying a ``left-triangle'' shape, as displayed in
-Figure \ref{fig:false-finding}, along with the Boston housing dataset,
-and the ``dino'' datasets from the \texttt{datasaurus} R package.
+In this section, we will present the performance of trained computer
+vision model on three example datasets. These include the dataset
+associated with the residual plot displaying a ``left-triangle'' shape,
+as displayed in Figure \ref{fig:false-finding}, along with the Boston
+housing dataset, and the ``dino'' datasets from the \texttt{datasaurus}
+R package.
 
@@
 non-typical. Here, the computer vision model and human visual inspection
-reject \(H_0\), whereas conventional tests do not. This emphasizes the
-benefits of visual inspection and the unique advantage of the computer
-vision model, which, like humans, makes decisions based on visual
-discoveries.
+reject \(H_0\), whereas some commonly used conventional tests do not.
+This emphasizes the benefits of visual inspection and the unique
+advantage of the computer vision model, which, like humans, makes
+decisions based on visual discoveries.
 
@@
 
-Figure \ref{fig:false-check} presents the results of the assessment.
-Notably, the observed visual signal strength is considerably lower than
-the 95\% sample quantile of the null distribution. Moreover, the
-bootstrapped distribution suggests that it is highly improbable for the
-fitted model to be misspecified as the majority of bootstrapped fitted
-models will not be rejected. Thus, for this particular fitted model,
-both the visual test and the computer vision model will not reject
-\(H_0\). However, the Breusch-Pagan test will reject \(H_0\) because it
-can not effectively utilize information from null plots.
+Figure \ref{fig:false-check} presents the results of the assessment by
+the computer vision model. Notably, the observed visual signal strength
+is considerably lower than the 95\% sample quantile of the null
+distribution. Moreover, the bootstrapped distribution suggests that it
+is highly improbable for the fitted model to be misspecified as the
+majority of bootstrapped fitted models will not be rejected. Thus, for
+this particular fitted model, both the visual test and the computer
+vision model will not reject \(H_0\). However, the Breusch-Pagan test
+will reject \(H_0\) because it can not effectively utilize information
+from null plots.
 
@@
 \caption{A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
-(A) The true residual plot exhibiting "left-triangle" shape.
+(A) The true residual plot exhibiting a "left-triangle" shape.
 (B) The attention map produced by computing the gradient of the output with respect to the greyscale input.
 (C) The density plot of estimated distnace for null plots and bootstrapped plots.
-The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots.
+The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots.
 The fitted model will not be rejected since $\hat{D} < Q_{null}(0.95)$.
@@
 assessment conducted by the computer vision model. A clear non-linearity
-pattern resembling a ``U'' shape is shown in the residual plot.
-Furthermore, the RESET test yields a very small \(p\)-value. The
-approximated distance \(\hat{D}\) significantly exceeds
-\(Q_{null}(0.95)\), leading to rejection of \(H_0\). The bootstrapped
-distribution also suggests that almost all the bootstrapped fitted
-models will be rejected, indicating that the fitted model is unlikely to
-be correctly specified.
-
-Additionally, the lineup of this fitted model presented in Figure
-\ref{fig:boston-lineup} underscores that the ``U'' shape is a
-distinctive characteristic of the true residual plot, absent in any of
-the null plots. Consequently, if a visual test is conducted, \(H_0\)
-will also be rejected.
+pattern resembling a ``U'' shape is shown in the plot A. Furthermore,
+the RESET test yields a very small \(p\)-value. The estimated distance
+\(\hat{D}\) significantly exceeds \(Q_{null}(0.95)\), leading to
+rejection of \(H_0\). The bootstrapped distribution also suggests that
+almost all the bootstrapped fitted models will be rejected, indicating
+that the fitted model is unlikely to be correctly specified. The
+attention map in plot B suggests the center of the image has higher
+leverage than other areas, and it is the turning point of the ``U''
+shape. The principal component analysis provided in plot D shows two
+distinct clusters of data points, further underlines the visual
+differences between bootstrapped plots and null plots. This coincides
+the findings from Figure \ref{fig:boston-lineup}, where the true plot
+exhibiting a ``U'' shape is visually distinctive from null plots. If a
+visual test is conducted by human, \(H_0\) will also be rejected.
 
@@
 
-\caption{A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
-On the left, the residual plot of the Boston housing fitted model is displayed.
-The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots.
-The fitted model will be rejected since $\hat{D} >= Q_{null}(0.95)$.}\label{fig:boston-check}
+\caption{A summary of the residual plot assessment for the Boston housing fitted model evaluted on 200 null plots and 200 bootstrapped plots.
+(A) The true residual plot exhibiting a "U" shape.
+(B) The attention map produced by computing the gradient of the output with respect to the greyscale input.
+(C) The density plot of estimated distnace for null plots and bootstrapped plots.
+The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots.
+The fitted model will be rejected since $\hat{D} \geq Q_{null}(0.95)$.
+(D) PC2 vs PC1 plot for features extracted from the global pooling layer of the computer vision model.
+}\label{fig:boston-check}
 \end{figure}
@@
 regression model to this data yields a residual plot resembling a
-``dinosaur,'' as displayed in Figure \ref{fig:dino-check}.
+``dinosaur,'' as displayed in plot A of Figure \ref{fig:dino-check}.
 Unsurprisingly, this distinct residual plot stands out in a lineup, as
@@
 normality assumption with a small \(p\)-value.
+
+More importantly, the attention map in plot B of Figure
+\ref{fig:dino-check} clearly exhibits a ``dinosaur'' shape, strongly
+suggesting the prediction of the distance is based on human perceptible
+visual patterns. The computer vision model is also capable of capturing
+the contour or the outline of the embedded shape, just like human being
+reading residual plots. The principle component analysis in plot D of
+Figure \ref{fig:dino-check} also shows that the cluster of bootstrapped
+plots is at the corner of the cluster of null plots.
+
+More importantly, the attention map in plot B of Figure
+\ref{fig:dino-check} clearly exhibits a ``dinosaur'' shape, strongly
+suggesting that the distance prediction is based on human-perceptible
+visual patterns. The computer vision model effectively captures the
+contour or outline of the embedded shape, similar to how humans
+interpret residual plots. Additionally, the principal component analysis
+in plot D of Figure \ref{fig:dino-check} demonstrates that the cluster
+of bootstrapped plots is positioned at the corner of the cluster of null
+plots.
 
@@
 
-\caption{A summary of the residual plot assessment evaluted on 200 null plots and 200 bootstrapped plots.
-On the left, the residual plot exhitbiting a "dinosaur" shape is displayed.
-The blue area indicates the distribution of approximated distances for null plots, while the red area represents the distribution of approximated distances for bootstrapped plots.
-The fitted model will be rejected since $\hat{D} >= Q_{null}(0.95)$.}\label{fig:dino-check}
+\caption{A summary of the residual plot assessment for the Datasaurus fitted model evaluted on 200 null plots and 200 bootstrapped plots.
+(A) The true residual plot exhibiting a "dinosaur" shape.
+(B) The attention map produced by computing the gradient of the output with respect to the greyscale input.
+(C) The density plot of estimated distnace for null plots and bootstrapped plots.
+The blue area indicates the distribution of estimated distances for bootstrapped plots, while the yellow area represents the distribution of estimated distances for null plots.
+The fitted model will be rejected since $\hat{D} \geq Q_{null}(0.95)$.
+(D) PC2 vs PC1 plot for features extracted from the global pooling layer of the computer vision model.}\label{fig:dino-check}
 \end{figure}
@@
 
-\section{Extension to other types of regression
-model}\label{extension-to-other-types-of-regression-model}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Use raw residuals.
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  D (effect size) for null plots will be greater than zero.
-\item
-  Can not assume D for actual plots are greater than D for null plots
-  any more. -\textgreater{} KL null - normal, KL actual - normal
-\item
-  Need to use a two-sided test instead of a one-sided test.
-\item
-  It may fail completely due to the identification issue.
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  Use transformed residuals that roughly follow a multivariate normal
-  distribution.
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  The one-sided test will still work.
-\item
-  Need to ensure that the transformed actual plot will have residual
-  distribution different from the transformed null distribution.
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{2}
-\tightlist
-\item
-  Train a new model.
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  KL divergence may not be the appropriate measure for the difference
-  between the residual distributions. Could consider using a new
-  distance measure.
-\end{itemize}
+\section{Limitations and Future Work}\label{limitations-and-future-work}
+
+Despite the computer vision model performing well with general cases
+under the synthetic data generation scheme and the three examples used
+in this paper, this study has several limitations that could guide
+future work.
+
+The proposed distance measure assumes that the true model is a classical
+normal linear regression model, which can be too restrictive. Although
+this paper does not address the relaxation of this assumption, there are
+potential methods to evaluate other types of regression models. The most
+comprehensive approach would be to define a distance measure for each
+different class of regression model and then train the computer vision
+model following the methodology described in this paper. To accelerate
+training, one could use the convolutional blocks of our trained model as
+a feature extractor and perform transfer learning on top of it, as these
+blocks effectively capture shapes in residual plots. Another approach
+would be to transform the residuals so they are roughly normally
+distributed and have constant variance. If only raw residuals are used,
+the distance-based statistical testing compares the difference in
+distance to a classical normal linear regression model for the true plot
+and null plots. This comparison is meaningful only if the difference can
+be identified by the distance measure proposed in this paper.
+
+There are other types of residual plots commonly used in diagnostics,
+such as residuals vs.~predictor and quantile-quantile plots. In this
+study, we focused on the most commonly used residual plot as a starting
+point for exploring the new field of automated visual inference.
+Similarly, we did not explore other, more sophisticated computer vision
+model architectures and specifications for the same reason. While the
+performance of the computer vision model is acceptable, there is still
+room for improvement to achieve behavior more closely resembling that of
+humans interpreting residual plots. This may require external survey
+data or human subject experiment data to understand the fundamental
+differences between our implementation and human evaluation.
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Brief discussion
-\item
-  Summary of findings
-\item
-  Contributions to the field
-\end{itemize}
-
-In this study, we have introduced a distance metric based on
+In this study, we have introduced a distance measure based on
 Kullback-Leibler divergence to quantify the disparity between the
-residual distribution of a fitted linear regression model and the
-reference residual distribution assumed under correct model
-specification. This distance metric effectively captures the magnitude
-of model violations in misspecified models. We propose a computer vision
-model to approximate this distance, utilizing the residual plot of the
-fitted model as input. The resulting approximated distance serves as the
-foundation for constructing a single Model Violation Index (MVI),
-facilitating the quantification of various model violations.
-
-Moreover, the approximated distance enables the development of a formal
+residual distribution of a fitted classical normal linear regression
+model and the reference residual distribution assumed under correct
+model specification. This distance measure effectively captures the
+magnitude of model violations in misspecified models. We propose a
+computer vision model to estimate this distance, utilizing the residual
+plot of the fitted model as input. The resulting estimated distance
+serves as the foundation for constructing a single Model Violation Index
+(MVI), facilitating the quantification of various model violations.
+
+Moreover, the estimated distance enables the development of a formal
 statistical testing procedure by evaluating a large number of null plots
@@
 compared to other types of violations. The statistical tests relying on
-the approximated distance predicted by the computer vision model exhibit
+the estimated distance predicted by the computer vision model exhibit
 lower sensitivity compared to conventional tests but higher sensitivity
-compared to visual tests conducted by humans. While the approximated
+compared to visual tests conducted by humans. While the estimated
 distance generally mirrors the strength of the visual signal perceived
@@
 contrast, individual residual diagnostic tests such as the RESET test
-and the Breusch-Pagan test only evaluate specific model assumptions.
-
-In practice, selecting an appropriate set of statistical tests for
+and the Breusch-Pagan test only evaluate specific model assumptions. In
+practice, selecting an appropriate set of statistical tests for
 regression diagnostics can be challenging, particularly given the
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Future directions
-\end{itemize}
-
 \section*{Acknowledgements}\label{acknowledgements}
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Packages used
-\item
-  Website link
-\end{itemize}
+These \texttt{R} packages were used for the work: \texttt{tidyverse}
+\citep{tidyverse}, \texttt{lmtest} \citep{lmtest}, \texttt{mpoly}
+\citep{mpoly}, \texttt{ggmosaic} \citep{ggmosaic}, \texttt{kableExtra}
+\citep{kableextra}, \texttt{patchwork} \citep{patchwork},
+\texttt{cassowaryr} \citep{mason2022cassowaryr}, \texttt{bandicoot}
+\citep{bandicoot}, \texttt{ggpcp} \citep{ggpcp}, \texttt{keras}
+\citep{keras} and \texttt{here} \citep{here}.
+
+The article was created with R packages \texttt{rticles}
+\citep{rticles}, \texttt{knitr} \citep{knitr} and \texttt{rmarkdown}
+\citep{rmarkdown}. The project's GitHub repository
+(\url{https://github.com/TengMCing/auto_residual_reading/paper})
+contains all materials required to reproduce this article.
+
+\section*{Disclosure Statement}\label{disclosure-statement}
+\addcontentsline{toc}{section}{Disclosure Statement}
+
+No potential conflict of interest was reported by the author(s).
+
+\section*{Supplementary Materials}\label{supplementary-materials}
+\addcontentsline{toc}{section}{Supplementary Materials}
+
+\begin{description}
+\item{Appendix:} The appendix includes more details about the neural network layers used for constructing the computer vision models.
+\end{description}
 

### 2024-05-29 paper/paper.Rmd
fix dark2 col

@@
 
+```{r}
+# check if package installed, version
+```
+
 # Introduction
@@
 
-```{r human-mosaic, fig.pos = "!h", fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of visual test conditional on the computer vision model decision on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot.
-The visual test rejects less frequently than the computer vision model, and (almost) only rejects when the computer vision model does.", eval = FALSE}
-library(ggmosaic)
-
-experiment %>%
-  filter(!attention_check) %>%
-  filter(!null_lineup) %>%
-  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
-  mutate(type = ifelse(type == "non-linearity", "Non-linearity", "Heteroskedasticity")) %>%
-  mutate(type = factor(type, levels = c("Non-linearity", "Heteroskedasticity"))) %>%
-  mutate(model_reject_20 = ifelse(model_reject_20, "Reject", "No")) %>%
-  mutate(reject = ifelse(reject, "Reject", "No")) %>%
-  mutate(across(c(model_reject_20, reject), ~factor(.x, levels = c("Reject", "No")))) %>%
-  ggplot() +
-  geom_mosaic(aes(x = ggmosaic::product(reject, model_reject_20), 
-                  fill = model_reject_20)) +
-  facet_grid(~type) +
-  ylab("Visual tests reject") +
-  xlab("Computer vision model rejects") +
-  # labs(fill = "Visual tests reject") +
-  scale_fill_brewer("", palette = "Dark2") +
-  theme_bw() +
-  theme(legend.position = "none") +
-  coord_fixed()
-```
-
 ```{r conv-mosaic, fig.pos = "!h", fig.cap = "Rejection rate ($p$-value $\\leq0.05$) of computer vision models conditional on conventional tests on non-linearity (left) and heteroskedasticity (right) lineups displayed using a mosaic plot."}
@@
   # labs(fill = "Visual tests reject") +
-  scale_fill_brewer("", palette = "Dark2") +
+  scale_fill_manual(values = c("#40B0A6", "#E1BE6A")) + 
   theme_bw() +
